### 2016-01-07 Rmd/cost-distance.Rmd
First commit of various Rmd files, up through first part of vegetation-types.Rmd

---
title: "Cost distance inputs"
author: "Lyndon Estes"
date: "December 9, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Creating a cost distance surface for Zambia

A summary of the datasets and steps needs to create a cost distance surface for Zambia.

## Approach

Following Tim Thomas's (IFPRI) basic approaches and datasets for making friction/access surfaces. We are going to use isotropic distances, and not worry about slopes and directional variation in speed.

### Datasets needed

1. Roads - these will vary the most, and will draw on three datasets
    + Urban Links shapefiles from Zambia--these will define urban areas
    + PFRLinks - Zambia Roads Department data, I think for provincial roads 
    + TMDLinks merged - RDA data these are national/trunk roads
    + roads_trunk; roads_primary; roads_secondary; roads_tertiary - these were downloaded from OpenStreetMap (using a QGIS utility). They provide almost the same coverage (and a bit better in some places then the TMD and PFRlinks data) and are accurate, and show consistent transport linkages into other countries. 
    + The trick is to compare the Zambia RDA datasets with the OpenStreetMap ones and ancillary data about the roads projects to figure out what are new roads, and what are just upgrades. The class of road and the upgrades (or whether they are new or not) will determine how quickly travel time will be on them. 
2. Rivers - These will provide a barrier except where roads cross. Using the USGS hydrosheds, selecting cells with more than 1000 cells of contributing areas upstream as major rivers.  
3. Lakes - Two sets of lakes data, one for all of Africa from FAO and one from the Zambia datasets I have received from ZARI. I will merge them both. 
4. Towns/cities - Three datasets - grump (NASA) urban centers for region outside of Zambia (towns that might represent the nearest town to some Zambia border regions), and the district head towns from the ZARI datasets.
5. International boundaries (Zambia shape). Will provide a barrier or friction.
6. Background frame - a rectangle covering Zambia and parts of neighboring countries, to have some connections between them.

```{r, eval = FALSE}
library(lmisc)
library(raster)
library(rgdal)
library(rgeos)
library(gdalUtils)
p_root <- set_base_path("zambiaToff")
p_roads <- full_path(p_root, "external/input_devel/roads/")
p_roads2 <- paste0("~/Dropbox/data/zari/infrastructure/zambia/rda_dump/", 
                   "Final Shape files RDA/")
p_trade <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/"

# Read in data
setwd(p_roads)

# Africa grid, for grid cells size and projection
mgrid <- raster(full_path(p_trade, "africa_grid.tif"))

# zambia and frame, for projecting
setwd("..")
zframe <- readOGR("zambia-frame.sqlite", layer = "zambia-frame")
zfr <- crop(mgrid, spTransform(zframe, projection(mgrid)))
zfr[!is.na(zfr)] <- 1
# plot(spTransform(zframe, projection(mgrid)), add = TRUE)

# zfr <- rasterize(spTransform(zframe, projection(mgrid)), mgrid)
# zfr <- raster(round(extent(zframe), 1))  # use extent to create raster
# res(zfr) <- 0.00833
# zfr[] <- 1
# projection(zfr) <- zframe@proj4string

setwd(p_roads2)
zambia <- readOGR("zam_boundary_region_ll.shp", "zam_boundary_region_ll")
zambia@proj4string <- zframe@proj4string

# towns
setwd(p_roads)
ztowns <- readOGR(dsn = "district_towns.shp", layer = "district_towns")
ntowns <- readOGR("zambia-neighbor-cities.shp", 
                  layer = "zambia-neighbor-cities")
ntowns <- ntowns[ntowns$ES00POP > 10000, ]  # filter down to 10,000 or greater
ztowns$cid <- 1:nrow(ztowns)
ztowns@data <- ztowns@data[, "cid", drop = FALSE]  # reduce for binding
ntowns$cid <- (max(ztowns$cid) + 1):(nrow(ntowns) + (max(ztowns$cid)))
ntowns@data <- ntowns@data[, "cid", drop = FALSE]  # reduce for binding
ntowns@proj4string <- ztowns@proj4string
alltowns <- spTransform(rbind(ztowns, ntowns), CRSobj = projection(zfr))

# write this out--needed for determining nodes in cost distance analysis
writeOGR(alltowns, dsn = "alltowns.shp", layer = "alltowns", 
         driver = "ESRI Shapefile", overwrite = TRUE)

# rivers & lakes
setwd(p_roads)
riv <- readOGR("rivers-gt1000.sqlite", layer = "rivers_gt1000")
riv@proj4string <- zframe@proj4string
riv <- readOGR("rivers-gt1000.sqlite", layer = "rivers_gt1000")

# QGIS merge of africa water-bodies clipped to frame and Zambia lakes layer
# both have inaccuracies against satellite
lakes <- readOGR("lakesall.shp", "lakesall")

# roads
rtypes <- c("trunk", "primary", "second", "tertiary")

r3 <- "roads_tmdlinks.sqlite"
r4 <- "roads_pfrlinks.sqlite"

roadnms <- sapply(rtypes, function(x) {
  dir(p_roads, pattern = paste0("^roads_", x))
})
rdrcl <- c(1, 2, 4, 8)
roads <- lapply(1:length(roads), function(x) {
  nm <- unname(roadnms)[x]
  rdspoly <- readOGR(nm, layer = gsub("\\.sqlite", "", nm))
  rdspoly@proj4string <- zframe@proj4string
  rdspoly$code <- rdrcl[x]  # assign new variable with code for road coding
  rdspoly@data <- rdspoly@data[, c("full_id", "osm_id", "code")]
  rdspoly
})

# 
tmd <- readOGR(r3, layer = gsub("\\.sqlite", "", r3))
tmd@proj4string <- zambia@proj4string
pfr <- readOGR(r4, layer = gsub("\\.sqlite", "", r4))
pfr@proj4string <- zambia@proj4string

# plot(zframe)
# plot(lakes, add = TRUE)
# plot(ztowns, pch = 20, cex = 0.5)
# points(ntowns, pch = 20, cex = 0.5)

```

### Rasterize datasets

We'll start with the frame for the region - a box defining Zambia and its surrounding regions. Next we'll do the roads. This is the one to pay attention to, because we are going to have use the TMDlinks and pfrlinks to update the values of these.  
```{r, eval = FALSE}
# Start with the frame
zamr <- rasterize(as(spTransform(zambia, projection(zfr)),"SpatialLines"), zfr)
zamr[is.na(zamr)] <- 0

# roads
# want to avoid super high numbers where roads overlap--always assign
# lowest value to intersecting roads
# e.g. 1 where primary, secondary overlap
# 1 <= 1 + 1 (2)
# 1 <= 1 + 2 | 1 + 1 + 1 (3)
# 1 <= 1 + 2 + 4 (7)
# 1 <= 1 + 2 + 4 + 8 (15)
# 2 <= 2 + 4 (6)
# 2 <= 2 + 4 + 8 (14)
# 3 <= 4 + 8 (12)
# 4 <= 8
# now read in roads data in a loop. This is the place where you will need to
# update values. Here we are just reading in 
# roads_trunk|primary|secondary|tertiary. 
roadsr <- lapply(1:length(roads), function(x) {
  print(paste("processing", x))
  rp <- spTransform(roads[[x]], projection(zfr))
  writeOGR(rp, dsn = "rdtmp.sqlite", layer = "rdtmp", driver = "SQLite")
  # r <- rasterize(x = rp, y = zfr, field = "code", progress = "text")
  fnm <- paste0("roads_", names(roadnms)[x], ".tif")
  gdal_rasterize(src_datasource = "rdtmp.sqlite", l = "rdtmp", ot = "Int16", 
                 a = "code", tr = res(zfr), at = TRUE, init = 0,
                 dst_filename = fnm, te = bbox(zfr)[1:4], of = "GTiff")
  file.remove("rdtmp.sqlite")
  r <- raster(fnm)
})

# start to add roads grids, and reclassify as we go. The fastest road is dominant
# where roads intersect
# roadsr[[1]] > 0 & roadsr[[2]] > 0
roadsrs <- roadsr[[1]] + roadsr[[2]]  # trunk and primary
roadsrs[roadsrs == 3] <- 1  # if 3 (trunk and primary), set to 1
roadsrs <- roadsrs + roadsr[[3]]  # sum of that plus secondary
roadsrs[roadsrs == 5] <- 1  # if 5 (trunk + secondary) set to 1
roadsrs[roadsrs == 6] <- 2  # if 6 (primary + secondary) set to 2
roadsrs <- roadsrs + roadsr[[4]]  # sum of that plus tertiary
roadsrs[roadsrs == 9] <- 1  # if 9 (trunk + tertiary) set to 1
roadsrs[roadsrs == 10] <- 2  # if 10 (primary + tertiary) set to 2
roadsrs[roadsrs == 12] <- 3  # if 12 (secondary + tertiary) set to 3
roadsrs[roadsrs == 4] <- 3  # reset 4 to 3
roadsrs[roadsrs == 8] <- 4  # reset 8 to 4
# freq(roadsr[[3]] > 0 & roadsr[[4]] > 0)
# freq(roadsrs)

# rivers
# 3 classes of rivers, based on upstream contributing areas. This will affect 
# crossing times. These are worth playing with
# gt > 20000. Let's say 90 minutes to cross, to make a barrier
# gt > 10000 < 20000. Let's say 60 minutes to cross, to make a barrier
# gt < 10000. Let's say 30 minutes to cross

# plots
# hist(riv$up_cells, plot = FALSE)
# plot(zambia)
# plot(riv[riv$up_cells > 20000, ], add = TRUE, col = "red")
# plot(riv[riv$up_cells > 10000 & riv$up_cells <= 20000, ], add = TRUE, 
#      col = "blue")
# plot(riv[riv$up_cells <= 10000, ], add = TRUE, 
#      col = "green")

# project rivers shapefile, give crossing time code, rasterize
rp <- spTransform(riv, projection(zfr))  # project to albers
rp$code <- 2  # assign code for setting times (this will be 10000-20000)
rp$code[rp$up_cells > 20000] <- 3  # 3 for greater than 20000
rp$code[rp$up_cells < 10000] <- 1  # 1 for lt 10000
writeOGR(rp, dsn = "tmp.sqlite", layer = "tmp", driver = "SQLite")
# r <- rasterize(x = rp, y = zfr, field = "code", progress = "text")
fnm <- "rivers.tif"
gdal_rasterize(src_datasource = "tmp.sqlite", l = "tmp", ot = "Int16", 
               a = "code", tr = res(zfr), at = TRUE, init = 0,
               dst_filename = fnm, te = bbox(zfr)[1:4], of = "GTiff")
file.remove("tmp.sqlite")
rivr <- raster(fnm)

# lakes
lakes$code <- 1
rp <- spTransform(lakes, projection(zfr))  # project to albers
fnm <- "lakes.tif"
writeOGR(rp, dsn = "tmp.sqlite", layer = "tmp", driver = "SQLite")
gdal_rasterize(src_datasource = "tmp.sqlite", l = "tmp", ot = "Int16", 
               a = "code", tr = res(zfr), at = TRUE, init = 0,
               dst_filename = fnm, te = bbox(zfr)[1:4], of = "GTiff")
file.remove("tmp.sqlite")
laker <- raster(fnm)

# assemble grid layers into a single friction/cost surface
roadsrs
laker[laker > 0] <- 50  # give lake a value easy to distinguish
zamr[zamr > 0] <- 100  # give border a value
rivr2 <- rivr * 200

# find overlaps to correct
bridges <- roadsrs > 0 & rivr > 0  # find places where roads cross rivers
rivinlake <- laker > 0 & rivr > 0  # places where rivers are in lakes
roadsinlake <- roadsrs > 0 & laker > 0  # roads in lakes - set to roads vals

# create friction surface and fix overlaps
fric <- laker + rivr2
fric2 <- (rivinlake == 0) * fric + (rivinlake * laker)  # lake over rivers
fric3 <- fric2 * (roadsrs == 0) + roadsrs  # roads over lakes and rivers
fric4 <- fric3 * (zamr == 0) + zamr  # border over lakes, rivers, roads

# give time values to the friction raster - time in minutes to cross 1 km grid 
fric4[fric4 == 1] <- 1 # 1 minutes for trunk road (60 km/h)
fric4[fric4 == 2] <- 1.333  # 1.333 minutes for primary road (45 km/h)
fric4[fric4 == 3] <- 2  # 2 minutes for secondary road (30 km/h)
fric4[fric4 == 4] <- 4 # 4 minutes for tertiary road (15 km/h)
fric4[fric4 == 50] <- 6  # 6 minutes on lake (10 km/h)
fric4[fric4 == 100] <- 60  # 60 minutes to cross border
fric4[fric4 == 200] <- 30   # 30 minutes to cross smaller river
fric4[fric4 == 400] <- 60   # 60 minutes to cross next size river
fric4[fric4 == 600] <- 90   # 1.5 hours to cross next size river
fric4[fric4 == 0] <- 20  # 20 minutes off-road

fric4 <- fric4 / 1000 / 60  # divide by 1000 / 60 to get cost in hrs / meter

# write it out
writeRaster(fric4, file = "friction.tif", overwrite = TRUE)

# copy over to ArcGIS folder (if you have it set-up)
cnm <- "/Volumes/Data/loMadzala/GIS/Zambia/costdistance/friction.tif"
file.copy("friction.tif", cnm, overwrite = TRUE)

```

The raster that was written out now becomes the cost surface for cost distance analysis. Do that in ArcGIS, because R doesn't have a great algorithm for that yet.

You will the alltowns.shp dataset as well for your analysis, which are the towns that will serve as the focal points for calculating travel costs

Send over the friction surface to ArcGIS, and then go into ArcToolBox > Spatial Analyst Tools > Distance > Cost Distance and the input alltowns into the input raster or feature source dialog, and friction into input cost raster dialog, and then choose an output name.

To bring up back into R, which you will need to do to work with it, export to the arcgis grid to a geotiff (right click on the name of the cost grid in the left hand menu of arc, save it as a geotiff, and copy it to where you will work with it in R)

That will give you a cost of movement in hours. You might want to play around with the values I have provided for the friction surface to see if they are reasonable.  Some end up being quite high.

# Processing cost grid to make it usable by model
```{r, eval = FALSE}
# Some code I use to quickly copy data out of the ArcGIS processing folder
# back to my machine
cnm <- "/Volumes/Data/loMadzala/GIS/Zambia/costdistance/zamcost2.tif"
file.copy(cnm, "cost.tif", overwrite = TRUE)

# read in the cost raster
costr <- raster("cost.tif")

# first we have to cut this down for R, and then convert it to a data.table
# get the zambia mask for AgroEcoTradeoffs
# find your install of AgroEcoTradeoffs
# read in the "ZA-mask.csv" file, which is a data.table
library(agroEcoTradeoff)
dt_path <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/"
msk <- fread(paste0(dt_path, "ZA-mask.csv"))  # read it in
mskr <- dt_to_raster(msk, CRSobj = projection(zamr))  # convert to raster

# now crop the cost surface, then resample it to  
# line up with our mask raster, since they are slightly offset from one 
# another, and then mask it
costrcr <- crop(costr, mskr)
costcrrs <- resample(costrcr, mskr)
costrmsk <- mask(costcrrs, mskr)
plot(costrmsk)  # voila

# then convert it to a data.table as follows
costdt <- na.omit(as.data.table(costrmsk, xy = FALSE))

# and then you will write it out as a csv file.  
fnm <- "../../data/ZA-cost.csv"
write.table(costdt, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
```

Your final output csv file will want to replace the ZA-cost.csv file in agroEcoTradeoffs folder `external/data/dt`, and then the model will work nicely.

### 2016-01-07 Rmd/maize-soy-grid.Rmd
First commit of various Rmd files, up through first part of vegetation-types.Rmd

---
title: "Zambia Potential Yield Grids"
author: "Lyndon Estes"
date: "December 3, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Prepare Maize and Soy Yield Grids

Based on simulated potential yields run by Di Tian. 
```{r, eval = FALSE}
library(data.table)
library(lmisc)
library(dtraster)
# devtools::install_github("PrincetonUniversity/lmisc")

p_root <- set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_agro <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/"

# Read in spatial reference grid and chop it down to size
ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
ID[, V1 := NULL]
setkeyv(ID, c("WTH", "SOL", "Zone"))
ID[, SharePct := as.numeric(SharePct)]
IDr <- subset(unique(ID[, .(x, y, WTH, Zone)]))
setkeyv(IDr, c("WTH", "Zone"))
# unique(ID[, .(WTH, SOL)])
# gr <- dt_to_raster(IDr[, .(ID, x, y)], CRS(prj))
# plot(gr)

plot(dt_to_raster(IDr[, .(x, y, as.factor(Zone))], CRS(prj)))
prj <- paste0("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0",
              " +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")
gcs <- "+proj=longlat +datum=WGS84"

# Maize yield in each grid cell varies as a function of soil, year, and 
# management inputs
# Need to capture that uncertainty
# Start with climate uncertainty by mean soil. 
# Then look at uncertainty in average year by soil type

# Weighted variance functions, from here: 
# https://rulesofreason.wordpress.com/2012/02/13/weighted-variance-and-weighted-coefficient-of-variation/
weighted.var <- function(x, w, na.rm = FALSE) {
    if (na.rm) {
        w <- w[i <- !is.na(x)]
        x <- x[i]
    }
    sum.w <- sum(w)
    sum.w2 <- sum(w^2)
    mean.w <- sum(x * w) / sum(w)
    (sum.w / (sum.w^2 - sum.w2)) * sum(w * (x - mean.w)^2, na.rm = na.rm)
}

weighted.var2 <- function(x, w, na.rm = FALSE) {
  if (na.rm) {
    w <- w[i <- !is.na(x)]
    x <- x[i]
  }
  sum.w <- sum(w)
  (sum(w*x^2) * sum.w - sum(w*x)^2) / (sum.w^2 - sum(w^2))
}

# We're going to start by mapping the CV of the soil versus the inter-annual
# variability in the data, just to the 0.25 degree resolution
# read in different grids, start with Maize
crops <- list.dirs(p_crop, full.names = FALSE)[-1]
cv_rasts <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X))
  fl <- unlist(lapply(c("S", "R", "I"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  idt <- copy(IDr)
  for(y in 1:length(fl)) { # y <- 2
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    # crss <- cr[YR == 1, ]
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "TRNO") := NULL]
    setkeyv(cr, c("WTH", "SOL", "Zone"))

    # unique(ID$WTH)[!unique(ID$WTH) %in% unique(substr(cr$WTH, 1, 4))]

    # cr[, .N, by = .(TNAM, SOL, Zone)][which(N > 62), ]
    # tst <- cr[TNAM == "AAAX0003" & SOL == "WI_FLSO001", ][order(YR), ]
    # plot(tst[YR > 31, HWAH], tst[YR <= 31, HWAH])
    # cr[TNAM == "AAEA0001" & SOL == "WI_CMZR003", ]

    crm <- cr[ID[, .(WTH, SOL, SharePct, Zone)], allow.cartesian = TRUE]

    # CV due to interannual variability
    solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
                 by = .(WTH, Zone, YR)]

    snm <- gsub("*.*_|\\.csv", "", fl[y])
    anncv <- solmu[, round(cv(V1)), by = .(WTH, Zone)]
    setnames(anncv, "V1", paste0("acv", snm))
    idt <- idt[anncv]
    setkeyv(idt, c("WTH", "Zone"))

    # anncv <- solmu[, round(cv(V1)), by = .(WTH, Zone)][, V1]
    # setnames(anncv, "V1", "acv")

    # annwv2[WTH == "AERM" & Zone == "Zone3",]
    # crm[WTH == "AERM" & Zone == "Zone3" & YR == 1, 
    # weighted.var2(HWAH, SharePct)]
    # annwv2[, sqrt(V1)] / solmu[, V1]  # checks

    # CV due to soil
    annmu <- crm[, round(mean(HWAH)), by = .(WTH, SOL, Zone, SharePct)]
    # cr[, .N, by = .(TNAM, SOL)][which(N > 62), ]
    # crm[WTH == "AAAI" & SOL == "WI_LVUY032" & Zone == "Zone2a", ]
    # crm[WTH == "AAAI" & SOL == "WI_LVUY032" & Zone == "Zone3", mean(HWAH)]
    # annmu[WTH == "AAAI" & Zone == "Zone3", ]
    # crm[WTH == "AAAI" & Zone == "Zone3" & SOL == "WI_LVUY032", mean(HWAH)]
    # range(crm[, round(mean(HWAH)), by = .(WTH, SOL, Zone)][, V1] - 
    # crm[, round(mean(HWAH)), by = .(WTH, SOL, Zone, SharePct)][, V1])
    # crm[WTH == "AAAA" & Zone == "Zone3" & SOL == "WI_CLBW256", mean(HWAH)]
    # annmu[WTH == "AAAA" & Zone == "Zone3" & SOL == "WI_CLBW256", ]
    annsolmu <- annmu[, weighted.mean(V1, SharePct), by = .(WTH, Zone)]
    # sum(annsolmu$V1 - annmu[, weighted.mean(V1, SharePct), 
    # by = .(WTH, Zone, SOL)][, V1])
    solcv <- annmu[, sqrt(weighted.var(V1, w = SharePct, na.rm = TRUE)), 
                   by = .(WTH, Zone)][, V1 / annsolmu[, V1] * 100]
    scvnm <- paste0("scv", snm)
    annsolmu[, c(scvnm) := round(solcv)]
    annsolmu[, V1 := NULL]
    setkeyv(annsolmu, c("WTH", "Zone"))
    idt <- idt[annsolmu]
    setkeyv(idt, c("WTH", "Zone"))
    idt
  }
  idt
})
names(cv_rasts) <- crops

# Mean CVs by cultivar
mucvs <- lapply(crops, function(x) {
  cv_rasts[[x]][, lapply(.SD, function(x) mean(x, na.rm = TRUE)), 
                 .SDcols = grep("cv", names(cv_rasts[[x]]))]
})
# cv_rasts$soybean[, lapply(.SD, function(x) length(which(is.na(x)))), 
#                  .SDcols = grep("cv", names(cv_rasts$soybean))]  # only 2 nas
# plot(sapply(nms1, function(x) mucvs[[1]][, get(x)]), 
#      sapply(nms2, function(x) mucvs[[1]][, get(x)]))

# Calculate mean CV across cultivar types
dtr <- lapply(crops, function(x) {
  nms1 <- names(cv_rasts[[x]])[grep("scv", names(cv_rasts[[x]]))]
  nms2 <- names(cv_rasts[[x]])[grep("acv", names(cv_rasts[[x]]))]
  cv_rasts[[x]][,  scvmu := rowMeans(.SD, na.rm = TRUE), .SDcols = nms1]
  cv_rasts[[x]][,  acvmu := rowMeans(.SD, na.rm = TRUE), .SDcols = nms2]
  dt_to_raster(cv_rasts[[x]][, .(x, y, scvmu, acvmu)], CRSobj = gcs)
})

# cv_rasts[[1]][order(x, -y), .(x, y)]

# CV due to soil and climate
png(full_path(p_root, "external/input_devel/crop-cv.png"), width = 700, 
    height = 700)
plot(stack(dtr), zlim = c(0, 100),  
     main = c("maize soil CV", "maize annual CV", "soy soil CV", 
              "soy annual CV"))
dev.off()
```

![figure](../external/input_devel/crop-cv.png)

These maps display the coefficient of variation in simulated maize (top row) and soybean (bottom row) yields, for variability due to soil differences (left column) and inter-annual variability in weather (right column). Soil differences within a given grid cause more yield variability than seasonal differences in weather.

<a href="#top">Back to top</a>

# Yield grids
## 10 km grids

Creating 10 km grids for 10th, 50th, 90th percentile yields for each cultivar under each management type (subsistence, rainfed commercial, irrigated). 
```{r, eval=FALSE}
nmso <- c("10", "50", "90")  # naming vector for quantile variables
probs <- c(0.1, 0.5, 0.9)  # quantiles to get
crops <- list.dirs(p_crop, full.names = FALSE)[-1]  # dir names
yld_rasts <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X))
  fl <- unlist(lapply(c("S", "R", "I"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  idt <- copy(IDr)
  for(y in 1:length(fl)) { # y <- 1
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "TRNO") := NULL]
    setkeyv(cr, c("WTH", "SOL", "Zone"))

    # merge IDs
    crm <- cr[ID[, .(WTH, SOL, SharePct, Zone)], allow.cartesian = TRUE]

    # Soil averaged yields
    solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
                 by = .(WTH, Zone, YR)]
    snm <- gsub("*.*_|\\.csv", "", fl[y])

    # Yield quantiles and means
    for(i in 1:3) {
      # solmu[WTH == "AAAA" & Zone == "Zone2b", box_stats(V1), by = .(WTH, Zone)]
      v <- solmu[, quantile(V1, probs = probs[i]), by = .(WTH, Zone)]
      setnames(v, "V1", paste0("y", nmso[i], "_", snm))
      idt <- idt[v]
      setkeyv(idt, c("WTH", "Zone"))
    }
    v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
    setnames(v, "V1", paste0("ymu_", snm))
    idt <- idt[v]
    setkeyv(idt, c("WTH", "Zone"))
    idt
  }
  idt
})
names(yld_rasts) <- crops

# Write out 10 km yield grids to csvs, for safe keeping
lapply(crops, function(x) {
  fnm <- full_path(p_root,
                   paste0("external/data/", x, "_yields_10km.csv"))
  write.csv(yld_rasts[[x]], file = fnm, row.names = FALSE)
})

# Take average of mean yields across the different cultivar types. 
# This will give 1 set of yields for each of the three management scenarios
# Subsistence, rainfed commercial, irrigated
# The same approach can be used to estimate the average of the 10th percentile 
# across cultivars, or median, or 90th, etc. 
yld_mus <- lapply(crops, function(x) {  # x <- crops[[1]]
  ymu <- lapply(c("S", "R", "I"), function(y) {  #y <- "S"
    cnms <- names(yld_rasts[[x]])[grep(paste0("mu_", y), names(yld_rasts[[x]]))]
    ylds <- copy(yld_rasts[[x]][, c("x", "y", cnms), with = FALSE])
    ylds[, c(x) := round(rowMeans(.SD)), .SDcols = cnms]
    ylds[, c(cnms) := NULL]
    ylds
  })
  named_out(ymu, c("S", "R", "I"))
})
names(yld_mus) <- crops

plot(dt_to_raster(yld_mus[[2]]$R, CRSobj = gcs))
```
<a href="#top">Back to top</a>

# 1 km grids

Average mean yields across cultivars within management treatments, and then reseample to 1 km grids. These become new inputs for the tradeoff model.  
```{r, eval=FALSE}
mgrid <- raster(full_path(p_agro, "ZA-grid.tif"))
msk <- raster(full_path(p_agro, "ZA-mask.tif"))
lapply(c("S", "R", "I"), function(x) { # x <- "S" 
  print(paste("Processing", x))
  DTL <- lapply(crops, function(y) {  # y <- crops[1]
    # fname <- full_path("external/base_data/", paste0(x, "-crop-yields.tif"))
    print(paste("..Processing", y))
    r <- dt_to_raster(yld_mus[[y]][[x]], CRSobj = gcs)
    r2 <- disaggregate(r, fact = 10)
    r3 <- resample(r, r2)
    r4 <- projectRaster(r3, msk)
    r4 <- mask(r4, msk)
    # plot(r4)
    valinds <- which(!is.na(values(msk)))
    DT <- as.data.table.raster(r4, xy = FALSE)[valinds, ]
    DT[, c(y) := round(get(y))]
    DT
  })
  DTdt <- do.call(cbind, DTL)
  fnm <- full_path(p_root,
                   paste0("external/data/ZA-potential-yields-", x, ".csv"))
  write.table(DTdt, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
})
```
<a href="#top">Back to top</a>

### 2016-01-07 Rmd/to-do.Rmd
First commit of various Rmd files, up through first part of vegetation-types.Rmd

---
title: "To-Do List"
author: "Lyndon Estes"
date: "January 4, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Data and Improvements Needed

For the next iteration of the model and the ProcRoyalB paper.

## Improvements

A table of things to do, stating their progress and likelihood of achieving ("D" = doing, "L" is likely to get done, "U" is unlikely to get done, "F" is finished, "A" is abandoned) during the month of January.  __Table last updated: `r Sys.Date()`__

```{r, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
# http://stackoverflow.com/questions/19997242/simple-manual-rmarkdown-tables-that-look-good-in-html-pdf-and-docx
# tabl <- "  
# | Item        | Description           | Doing or Done?  |
# |---------------|:-------------|------|
# | Plausible scenario | A single production target to follow (Alastair)  | Doing |
# | Gridded yield estimates | Maize and soybean, exact method for handling variability & production zones (e.g. irrigation regions) |  Doing |
# | Gridded yield estimates | Run with improved soil profiles (Di) | Unlikely |
# | Upgrade soils | With ILUA database (with Nate?) | Unlikely |
# | Filter terrain | Get rid of unfarmable bits, probably needs new model logic (cell fraction multiplier) | Doing | 
# 
# "
# cat(tabl) # output the table in a format good for HTML/PDF/docx conversion

# make an html table
# http://r-norberg.blogspot.com/2013/06/generating-tables-using-pander-knitr.html
tabl <- cbind(
  "Item" = c("Plausible Scenario", 
             "Gridded yields",
             "Gridded yields",
             "Upgrade soils", 
             "Filter terrain", 
             "Biodiversity impacts", 
             "Impacts weighting (per Tim)", 
             "Work in economics", 
             "Explore econ versus enviro goals", 
             "Better cropland map, yield estimates"), 
  "Notes" = c("Single production to follow (Alastair)", 
              "Exact method for variability and production zones, including inability to reach potential yields", 
              "Run with improved soils maps", 
              "With ILUA database (and Nate's help)", 
              "Removable unfarmable areas, add new supporting logic", 
              "Habitat rarity/threatenedness", 
              "Weight carbon and biodiversity equally", 
              "Estimate profitability per pixel. Either input modification or code modification", 
              "See where map overlaps are greatest, most divergent", 
              "Upgrade all maps to SERVIR landcover, and rebase yield estimates on that"),
  "Doing" = c("D", "D", "U", "U", "D", "D", "D", "L", "D", "D")
)
print(xtable::xtable(tabl), align = rep("c", dim(table)[2] + 1), type = "html", 
      include.rownames = FALSE)
```

### 2016-01-07 Rmd/vegetation-types.Rmd
First commit of various Rmd files, up through first part of vegetation-types.Rmd

---
title: "Vegetation types"
author: "Lyndon Estes"
date: "January 4, 2016"
output: html_document
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Preparing a habitat rarity layer based on vegetation types

A more straightforward way for calculating biodiversity impacts, which can be based on hectares

## Data
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(RPostgreSQL)
library(rmapaccuracy)
library(gisutils)

# p_dat <- "/Users/lestes/Dropbox/data/zari/landcover/Spatial Data_FD_ILUAII/"
p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_dat <- paste0("/Users/lestes/Dropbox/data/distributions/vegetation/",
                "potential/vecea/Zambia/")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_devel <- full_path(p_proj, "external/input_devel/biodiversity")

# Load in Vegetation Map of Zambia
veg <- readOGR(dsn = full_path(p_dat, "pnv_vecea_v2_0_Zambia.shp"), 
               layer = "pnv_vecea_v2_0_Zambia", stringsAsFactors = FALSE)
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
msk <- raster(full_path(p_basedat, "ZA-mask.tif"))
cntry <- readOGR(full_path(p_basedat, "africa_countries_alb.sqlite"), 
                 layer = "africa_countries_alb")
zam <- cntry[cntry@data$fips_cntry == "ZA", ]
```

## Intersect grid with vegetation types

To 1 km resolution, in order to turn it into percent coverage

```{r, eval=FALSE}
# transform vegetation map to albers
vegalb <- spTransform(veg, mgrid@crs)

# Test intersects
vtypes <- unique(vegalb$CAT)
vegss <- vegalb[vegalb$CAT == 75, ]
tcrp <- crop(mgrid, extent(vegss))
tint <- intersect(vegss, tcrp)
# plot(tint[1:100, ])
# plot(tcrp)
# tcrp <- rasterToPolygons(tcrp)
# plot(tcrp[1:10, ])
# plot(vegss)
# plot(zam, add = TRUE)
# rm(tint)

```

Initial tests shows this will be tedious in R, so will run this through postgis, for speed (although given the install hassles I could have already run this and been done in R)

Notes on fixing a broken postgis install are appended to postgis

```{r, eval = FALSE}
# Set up new postigs database
Sys.setenv(PATH=paste(Sys.getenv("PATH"),"/usr/local/pgsql/bin/",
                      sep=":"))
# Sys.getenv("PATH")
dbname <- "zambiatoff"
# system(paste("createdb", dbname))
# system("psql -U lestes -h localhost zambiatoff")  # just hangs
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, host = "localhost", dbname = dbname, port = 5432)
# dbSendQuery(con, "CREATE EXTENSION postgis;")
# dbSendQuery(con, "CREATE EXTENSION postgis_topology;")
# dbSendQuery(con, "CREATE EXTENSION fuzzystrmatch;")
# dbSendQuery(con, "CREATE EXTENSION postgis_tiger_geocoder;")

# Need to add back in Africa Albers projection
# p_sql <- "~/Dropbox/projects/mappingafrica/external/postgis/af-alb-ref.sql"
# system(paste("psql -f", p_sql, dbname))  # add in prj
srid <- 102022
sql <- paste0("select proj4text from spatial_ref_sys WHERE srid=", srid)
rs <- dbGetQuery(con, sql)

# First convert zambia grid to a set of polygons, for loading into postgis
# Resurrect old gdal_polygonize wrapper
gdal_polygonize <- function(inraster, outshape, attname, 
                            gdalformat = "SQLite") {
  gdal_setInstallation()
  ext_tab <- cbind(c("SQLite", "ESRI Shapefile"), 
                   c(".sqlite", ".shp"))
  ext <- ext_tab[which(ext_tab[, 1] == gdalformat), 2]
  gdal_opts <- getOption("gdalUtils_gdalPath")[[1]]

  pyv <- system("which python", intern = TRUE)
  py_c <- paste(pyv, paste0(gdal_opts$path, "gdal_polygonize.py"))
  rnm <- inraster@file@name
  # rast_nm <- unlist(strsplit(rnm, "/"))[length(unlist(strsplit(rnm, "/")))]
  fullc <- paste(py_c, rnm, "-f", paste0("'", gdalformat, "'"), 
                 paste0(outshape, ext), outshape, attname)
  system(fullc)
  # shp <- readOGR(dsn = paste0(outshape, ext), layer = outshape)
  # return(shp)
}

# Convert to shapefile using gdal_polygonize
setwd(p_devel)
gdal_polygonize(mgrid, "zamgrid", "ID", "ESRI Shapefile")

r <- crop(mgrid, extent(-200000, -100000, -2000000, -1900000))

# Small test grid
# plot(zam)
# plot(as(extent(-200000, -100000, -2000000, -1900000), "SpatialPolygons"), 
#      add = TRUE)
r <- writeRaster(r, file = "testraster.tif", overwrite = TRUE)
gdal_polygonize(r, "testgrid", "ID", "ESRI Shapefile")
polygonizer(r@file@name, outshape = "testgrid2.shp")  # doesn't work with asc

# Send to postgis
postgis_vectab("testgrid.shp", "testgrid", "zambiatoff", "geom", "lestes", con)
# dbRemoveTable(con, "testgrid")

# system("shp2pgsql zamgrid.shp zamgrid zambiatoff > zamgrid.sql")
# system("psql -U lestes -d zambiatoff -f zamgrid.sql")

# Index it
grname_out <- "zamgrid"
dbSendQuery(con, paste0("CREATE INDEX gix_",  grname_out, " ON ", grname_out, 
                        " USING GIST(geom)"))
dbSendQuery(con, paste("VACUUM ANALYZE",  grname_out))
dbSendQuery(con, paste0("CLUSTER ", grname_out, " USING gix_", grname_out))
dbSendQuery(con, paste("ANALYZE",  grname_out))

# A little test, using rmapaccuracy functions
# sql <- paste0("SELECT ID, ST_ASeWKT(geom) FROM ", grname_out, 
#              " WHERE gid in (", paste0(1:10, collapse = ","), ")") 
# qgrid <- dbGetQuery(con, sql) 
# plot(polyFromWkt(qgrid, rs$proj4text))

# Vacuum analyze
create_idx <- function(tabname, idxvarname, idxvar, con) {
  # this first one should be revised to check if there is an index variable or 
  # not first
  dbSendQuery(con, paste0("CREATE INDEX gix_", idxvarname, " ON ",
                          tabname, " USING GIST(", idxvar, ")"))
  dbSendQuery(con, paste("VACUUM ANALYZE", tabname))
  dbSendQuery(con, paste0("CLUSTER ", tabname, " USING gix_", idxvarname))
  dbSendQuery(con, paste("ANALYZE",  tabname))
}

# A function to create tables from shapefiles
postgis_vectab <- function(shp, tabname, dbname, idxvarname, idxvar, user, 
                           con) {
  pgcalls <- list(paste("shp2pgsql", shp, tabname, dbname, ">",
                        paste0(tabname, ".sql")), 
                  paste("psql -U", user, "-d", dbname, "-f", 
                        paste0(tabname, ".sql")))
  for(i in pgcalls) system(i)  # create table in db
  create_idx(tabname, idxvarname, idxvar, con)
}

# send vegetation types to postgis
writeOGR(vegalb, dsn = full_path(p_devel, "vecea.shp"), 
         layer = "vecea", driver = "ESRI Shapefile")
create_pg_vtable("vecea.shp", "vtypes", "zambiatoff", "geom", "lestes", con)

# Update srids for the two fields
sql <- paste0("SELECT UpdateGeometrySRID('vtypes', 'geom', ", srid, ")")  
dbSendQuery(con, sql)
dbClearResult(dbListResults(con)[[1]])
sql <- paste0("SELECT UpdateGeometrySRID('zamgrid', 'geom', ", srid, ")")  
dbSendQuery(con, sql)
dbClearResult(dbListResults(con)[[1]])

# Intersection
isectname <- "vtypegrids"
sql <- paste("CREATE TABLE",  isectname, 
             "(gid serial primary key,", 
             "grid_id integer,", 
             "v_cat integer,",  
             "geom geometry(multipolygon,", srid, ")",
             ");")
d <- dbSendQuery(con, sql)

a <- "zamgrid"
b <- "vtypes"
tick <- Sys.time()
print(paste("Processing intersection, starting at time", tick))
sql <- paste("INSERT INTO", isectname, "(grid_id, v_cat, geom)", 
             "SELECT", 
             "a.id AS grid_id,", 
             "b.gid AS v_cat,",  
             "CASE", 
             "WHEN ST_Within(a.geom,b.geom)", 
             "THEN a.geom", 
             "ELSE ST_Multi(ST_Intersection(a.geom,b.geom))", 
             "END AS geom", 
             "FROM", a, "a",
             "JOIN", b, "b", 
             "ON ST_Intersects(a.geom, b.geom)")
system.time(d <- dbSendQuery(con, sql))  # 6.211
tock <- Sys.time()
print(paste0("Finished intersections at ", tock, ", total time = ", tock - tick))

create_idx()

for(i in vtypes) {
  v <- vegalb[vegalb@data$CAT == i, ]

}
vtypes

dbDisconnect(con)

```

```{r, eval=FALSE, echo=FALSE}
# For cleaning up zari-sourced vegetation map

# trim down some names
# unique(veg@data[, c(1:2, 4)])
ftype <- unique(veg@data$FEATURE_TY)
ftyper <- ftype
ftyper[1] <- "treeless-grassy"
ftyper[2] <- "termitary"
ftyper[3:4] <- c("miombo hilly", "miombo plateau")
ftyper[10] <- "Parinari-copperbelt chipya"
ftyper[15] <- "Munga heavy soils"
ftyper[16] <- "Kalahari woodland"
ftyper[17] <- "Mopane"
ftyper[18] <- "Baikiaea"

# Replace in table
ftype_tab <- cbind(ftype, ftyper)
for(i in 1:nrow(ftype_tab)) {  # i <- 1
  ind <- which(veg@data$FEATURE_TY == ftype_tab[i, 1])
  veg@data$FEATURE_TY[ind] <- ftype_tab[i, 2]
}
colnames(veg@data) <- c("code", "vtype1", "trapnell", "vtype2")
unique(veg@data[, c(1:2, 4)])
# plot(veg[veg$FORESTCLAS == "Inland water", ])
# plot(veg[veg@data$vtype1 == "Itigi forest", ])

```

### 2016-01-15 Rmd/cropland.Rmd
Various scripts for creating terrain, cropland, and actual vegetation layers

---
title: "cropland.Rmd"
author: "Lyndon Estes"
date: "January 14, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# New cropland map of Zambia

Cropland class from the SERVIR landcover map (2010)

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(rworldxtra)
data("countriesHigh")

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
lc <- raster(full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif"))

```

## Extract cropland class and resample
```{r, eval = FALSE}

# pullout cropland class 8
system.time(cropland <- rast_math(lc == 8, tmpdir = p_devel))

onm <- full_path(p_devel, "cropland-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = cropland@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
r <- raster(onm)
par(mar = rep(0, 4)); plot(r)

```

### 2016-01-15 Rmd/terrain.Rmd
Various scripts for creating terrain, cropland, and actual vegetation layers

---
title: "terrain"
author: "Lyndon Estes"
date: "January 8, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Filter terrain by workability

The first step entailed getting a percent slope map from EarthEngine, using script XX, from which I exported a map showing areas less than 20% slope built from the new SRTM 30 m dataset for Africa.

## Background data for Earth Engine
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(rworldxtra)
data("countriesHigh")

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
# p_dat <- paste0("/Users/lestes/Dropbox/data/distributions/vegetation/",
#                 "potential/vecea/Zambia/")

p_devel <- full_path(p_proj, "external/input_devel/terrain")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")

# Load in Vegetation Map of Zambia
cntry <- readOGR(full_path(p_basedat, "africa_countries_alb.sqlite"), 
                 layer = "africa_countries_alb")
cntrygcs <- spTransform(cntry, raster()@crs)
zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
zambia@data <- zambia@data[, c("OID_", "NAME")]
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))

writeOGR(cntrygcs, dsn = full_path(p_devel, "africa.kml"), 
         layer = "africa", driver = "KML")
writeOGR(zambia, dsn = full_path(p_devel, "zambia.kml"), 
         layer = "zambia", driver = "KML")
```

## Process slope filter from Earth Engine
```{r, eval = FALSE}

farmable <- raster(full_path(p_devel, "farmable.tif"))

onm <- full_path(p_devel, "farmable-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = farmable@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
r <- raster(onm)
par(mar = rep(0, 4)); plot(r)

```

### 2016-01-15 Rmd/vegetation-types.Rmd
Various scripts for creating terrain, cropland, and actual vegetation layers

@@
 date: "January 4, 2016"
-output: html_document
+output: 
+  html_document:
     highlight: tango
@@
 library(gisutils)
+library(rworldxtra)
+data("countriesHigh")
 
@@
 mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
+# length(which(!is.na(values(mgrid)))) # diff(cellStats(mgrid, range))
+# plot(mgrid)
 msk <- raster(full_path(p_basedat, "ZA-mask.tif"))
@@
 
-Notes on fixing a broken postgis install are appended to postgis
+Notes on fixing a broken postgis install are in evernote.
 
@@
 setwd(p_devel)
-gdal_polygonize(mgrid, "zamgrid", "ID", "ESRI Shapefile")
-
-r <- crop(mgrid, extent(-200000, -100000, -2000000, -1900000))
 
@@
 #      add = TRUE)
+r <- crop(mgrid, extent(-200000, -100000, -2000000, -1900000))
 r <- writeRaster(r, file = "testraster.tif", overwrite = TRUE)
-gdal_polygonize(r, "testgrid", "ID", "ESRI Shapefile")
-polygonizer(r@file@name, outshape = "testgrid2.shp")  # doesn't work with asc
+# gdal_polygonize(r, "testgrid", "ID", "ESRI Shapefile")
+polygonizer(r@file@name, outshape = "testgrid.shp")  # doesn't work with asc
+# gdal_polygonize(mgrid, "zamgrid", "ID", "ESRI Shapefile")
+polygonizer(mgrid@file@name, outshape = "zamgrid.shp")  # doesn't work with asc
 
@@
 
-# system("shp2pgsql zamgrid.shp zamgrid zambiatoff > zamgrid.sql")
+# system("shp2pgsql zamgrid2.shp zamgrid zambiatoff > zamgrid.sql")
 # system("psql -U lestes -d zambiatoff -f zamgrid.sql")
@@
 # Index it
-grname_out <- "zamgrid"
-dbSendQuery(con, paste0("CREATE INDEX gix_",  grname_out, " ON ", grname_out, 
-                        " USING GIST(geom)"))
-dbSendQuery(con, paste("VACUUM ANALYZE",  grname_out))
-dbSendQuery(con, paste0("CLUSTER ", grname_out, " USING gix_", grname_out))
-dbSendQuery(con, paste("ANALYZE",  grname_out))
+postgis_vectab("zamgrid.shp", "zamgrid", "zambiatoff", "geom", "lestes", con)
+# create_idx("zamgrid", idxvar = "geom", con)
+# dbRemoveTable(con, "zamgrid")
 
@@
 
-# Vacuum analyze
-create_idx <- function(tabname, idxvarname, idxvar, con) {
-  # this first one should be revised to check if there is an index variable or 
-  # not first
-  dbSendQuery(con, paste0("CREATE INDEX gix_", idxvarname, " ON ",
-                          tabname, " USING GIST(", idxvar, ")"))
-  dbSendQuery(con, paste("VACUUM ANALYZE", tabname))
-  dbSendQuery(con, paste0("CLUSTER ", tabname, " USING gix_", idxvarname))
-  dbSendQuery(con, paste("ANALYZE",  tabname))
-}
-
-# A function to create tables from shapefiles
-postgis_vectab <- function(shp, tabname, dbname, idxvarname, idxvar, user, 
-                           con) {
-  pgcalls <- list(paste("shp2pgsql", shp, tabname, dbname, ">",
-                        paste0(tabname, ".sql")), 
-                  paste("psql -U", user, "-d", dbname, "-f", 
-                        paste0(tabname, ".sql")))
-  for(i in pgcalls) system(i)  # create table in db
-  create_idx(tabname, idxvarname, idxvar, con)
-}
-
 # send vegetation types to postgis
@@
          layer = "vecea", driver = "ESRI Shapefile")
-create_pg_vtable("vecea.shp", "vtypes", "zambiatoff", "geom", "lestes", con)
+postgis_vectab("vecea.shp", "vtypes", "zambiatoff", "geom", "lestes", con)
 
@@
 dbClearResult(dbListResults(con)[[1]])
+# sql <- paste0("SELECT UpdateGeometrySRID('testgrid', 'geom', ", srid, ")")  
 sql <- paste0("SELECT UpdateGeometrySRID('zamgrid', 'geom', ", srid, ")")  
@@
 # Intersection
-isectname <- "vtypegrids"
-sql <- paste("CREATE TABLE",  isectname, 
-             "(gid serial primary key,", 
-             "grid_id integer,", 
-             "v_cat integer,",  
+isectname <- "gridveg"
+# dbRemoveTable(con, isectname)
+sql <- paste0("CREATE TABLE ",  isectname, 
+             "(gid serial primary key, ", 
+             "grid_id integer, ", 
+             "vgid integer, ", 
+             "veg_cat integer, ",
              "geom geometry(multipolygon,", srid, ")",
@@
 print(paste("Processing intersection, starting at time", tick))
-sql <- paste("INSERT INTO", isectname, "(grid_id, v_cat, geom)", 
+sql <- paste("INSERT INTO", isectname, "(grid_id, vgid, veg_cat, geom)", 
              "SELECT", 
              "a.id AS grid_id,", 
-             "b.gid AS v_cat,",  
+             "b.gid AS vgid,", "b.cat AS veg_cat,",  
              "CASE", 
-             "WHEN ST_Within(a.geom,b.geom)", 
+             "WHEN ST_CoveredBy(a.geom,b.geom)", 
              "THEN a.geom", 
@@
              "END AS geom", 
-             "FROM", a, "a",
-             "JOIN", b, "b", 
+             "FROM", a, "AS a",
+             "INNER JOIN", b, "AS b", 
              "ON ST_Intersects(a.geom, b.geom)")
@@
 tock <- Sys.time()
-print(paste0("Finished intersections at ", tock, ", total time = ", tock - tick))
-
-create_idx()
-
-for(i in vtypes) {
-  v <- vegalb[vegalb@data$CAT == i, ]
-
-}
-vtypes
+print(paste0("Finished at ", tock, ", total time = ", tock - tick))
 
 dbDisconnect(con)
+```
+Intersection in postgis taking far too long (it ran for more than 15 hours before I killed it).
+Switching to Grass, which is much faster.
+Table updates of area calculations took unacceptably long there also.
+So I ended up doing the full workflow in ArcGIS (!), in script `vegridToRaster.py`.
+
+# Convert vegetation types
+
+## Come up with percent remaining first
+
+First I sent both Zambia landcover and a rasterized version of vecea (in EPSG:4326) created by ArcGIS (`veceaToRaster.py`).  I then 
+```{r, eval = FALSE}
+# Read in extent of Zambia landcover grid
+p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"
+lcnm <- full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif")
+lc <- raster(full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif"))
+vegrem <- raster(full_path(p_devel, fname = "vegremain.tif"))
+
+lt <- 8
+fnm <- full_path(p_devel, "cropland.tif")
+pcalc <- paste0("gdal_calc.py -A ", lcnm, " --outfile=", fnm, " --type='Byte'",
+                " --overwrite --calc=", "'", lt, "'", " --co 'COMPRESS=LZW'")
+system.time(system(pcalc))  # 49 seconds
+
+# Read in unique vegetation types from "actual" vegetation map
+v <- gdalinfo(vegrem@file@name, hist=TRUE)[31]  # gdalinfo histogram
+v <- as.numeric(strsplit(v, " ")[[1]])  # convert to numeric, get numbers
+v <- v[!is.na(v)]  # remove NAs
+lcvals <- c(0:(length(v) - 1))[which(v != 0)]  # what are the actuals, less 0
+
+# ran raster calculator - this proves much faster than gdal_calc.py, which is 
+# a dog by comparison and always converts to float, despite passing byte type
+p_temp <- full_path(p_devel, "temp")
+dir.create(p_temp)
+
+# Processed converted areas also
+vtr <- lapply(lcvals, function(x) {
+  print(paste("Processing veg type", x))
+  dang <- Sys.time()
+  r <- rast_math(vegrem == x, tmpdir = p_temp)
+  print(Sys.time() - dang)
+  r
+})
+# length(vtr) * 4  # rough time to process, about 88 minutes for all types
+names(vtr) <- paste0("veg", lcvals)
+
+# aggregate and reproject these types to 1 km resolution
+ext <- sapply(c("xmin", "ymin", "xmax", "ymax"), function(x) {
+  slot(extent(sa_r), x)
+})
+
+p_aveg <- full_path(p_devel, "act_veg")
+dir.create(p_aveg)
+
+atypes <- names(vtr)[names(vtr) != "veg0"]  # drop type 0, which are the converted areas
+avr <- lapply(1:length(atypes), function(x) {  # x <- 1
+  ri <- vtr[atypes[x]]
+  nm <- paste0(names(ri), ".tif")
+  onm <- full_path(p_aveg, nm)
+  ext <- bbox(mgrid)[1:4]  # extent   
+  print(paste("warping and masking", nm))
+  dang <- Sys.time()
+  gdalwarp(srcfile = ri[[1]]@file@name, t_srs = projection(mgrid), 
+           dstfile = onm, r = "average", ot = "Float32", te = ext, 
+           srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
+           of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
+  # file.remove(lcnm[j])
+  print(ptime <- Sys.time() - dang)
+  l <- list(raster(onm), as.numeric(gsub("[[:alpha:]]", "", ptime)))
+  names(l) <- c(names(ri), "minutes")
+  return(l)
+})
+sum(sapply(avr, function(x) x$minutes)) / 60  # actually in seconds, not mins
+
+# getOption("gdalUtils_gdalPath")
+# getOption("")
+
+# stack the rasters
+s <- stack(lapply(avr, function(x) x[[1]]))
+
+# Sum for total - the inverse of this will be the farmable areas
+remfrac <- calc(s, sum)
+
+veg_dt <- dtraster::as.data.table(s)  # stack and convert to dt
+veg_dt[, lapply(.SD, function(x) max(x, na.rm = TRUE))]
+veg_dt <- na.omit(veg_dt)
+
+```
+
+# Various rubbish I tried
+```{r, eval = FALSE}
+# Convert to kmls for creating fusion tables
+# http://gis.stackexchange.com/questions/43848/how-to-re-project-shapefiles
+system("ogr2ogr -f KML -t_srs EPSG:4326 zamgrid.kml zamgrid.shp")
+system("ogr2ogr -f KML -t_srs EPSG:4326 vecea.kml vecea.shp")
+system("ogr2ogr -t_srs EPSG:4326 veceagcs.shp vecea.shp")
+
+options("gdal")
+
+fnm <- full_path(p_proj, "external/input_devel/roads/tmdpfr_links.sqlite")
+onm <- full_path(p_proj, "external/input_devel/roads/tmdpfr_links.kml")
+system(paste("ogr2ogr -f KML", onm, fnm))
+
+zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
+zambia@data <- zambia@data[, c("OID_", "NAME")]
+r <- raster(extent(zambia))
+# res(r) <- c(0.008333333, 0.008333333)
+res(r) <- c(0.00899, 0.00899)
+projection(r) <- raster()@crs
+r[] <- 1
+dataType(r) <- "INT4S"
+
+zambiar <- rasterize(zambia, r, dataType = "INT4S")
+dataType(zambiar) <- "INT4S"
+vi <- which(!is.na(values(zambiar)))
+zambiar[vi] <- as.integer(1:length(vi))
+# plot(zambiar)
+length(which(!is.na(values(zambiar))))
+dataType(zambiar)
+
+zambiawr <- writeRaster(zambiar, format = "GTiff",
+                        filename = full_path(p_devel, "zamgridgcs.tif"),
+                        datatype = "INT4S", overwrite = TRUE)
+dataType(zambiawr)
+
+round(res(r)[1] / res(lc)[1])
+res(lc)[1] * 31
+
+# First rasterize landcover types at 30 m resolution
+# inm <- full_path(p_dat, "pnv_vecea_v2_0_Zambia.shp")
+# fnm <- full_path(p_devel, "vecea.tif")
+# gdal_rasterize(src_datasource = inm, l = "pnv_vecea_v2_0_Zambia", ot = "Byte", 
+#                a = "CAT", tr = res(lc), at = TRUE, co = "COMPRESS=LZW",
+#                dst_filename = fnm, te = bbox(lc)[1:4], of = "GTiff", 
+#                verbose = TRUE)
+# 
+# veg@data <- veg@data[, "CAT", drop = FALSE]
+# writeOGR(veg, dsn = full_path(p_devel, "vecea.kml"), layer = "vecea", 
+#          driver = "KML")
+
+# a <- rep(0, 40)
+# a[sample(1:length(a), size = 40, replace = FALSE)] <- 1
+# mean(a)
 

### 2016-01-15 js/farmable.js

@@ -1,0 +1,24 @@
+// EarthEngine code to create steep slope mask for Zambia
+
+// Load in SRTM 30 m for Africa and Zambia shapefile
+var dem = ee.Image('USGS/SRTMGL1_003');
+var zambia = ee.FeatureCollection('ft:10YdSHdN3XgDoRFDyPLecjjyV_5iQqGEymorahxMD', 'geometry');
+var zamgrid = ee.Image('users/lyndonestes/africa/zamgrid');
+
+// Calculate slope, clip down to Zambia, and then convert to percent
+var slope = ee.Terrain.slope(dem);
+var slopedeg = slope.clip(zambia);
+var slopepct = slopedeg.expression('tan(slope * 3.141593 / 180) * 100', {
+  'slope': slopedeg.select('slope')
+});
+
+var zoneScale = slopepct.projection().nominalScale().getInfo();
+
+// Filter out steep areas
+var farmable = slopepct.lte(20);
+
+Export.image(farmable, 'farmable', {
+  region: zambia,
+  scale: zoneScale,
+  maxPixels: 1500000000
+});

### 2016-01-15 js/rasterZoneStats.js

@@ -1,0 +1,61 @@
+//
+// Function to apply a reducer/stat to raster zones
+//
+function RasterZoneStats(imCollection,zones,reducer,clipRegions,timeStampName,
+timeStampFormat,scale){
+  // decide reduceRegion scale which impacts the computation time significantly
+  if (scale === undefined){
+    //use the finer scale of imCollection and zones
+    var icScale = ee.Image(imCollection.first()).projection().nominalScale().getInfo();
+    //print('Image Collection Scale',icScale);
+    var zoneScale = zones.projection().nominalScale().getInfo();
+    //print('Zone raster scale',zoneScale);
+    scale =Math.min(icScale,zoneScale);
+  }
+  print("RasterZoneStats() scale:", scale);
+
+  // get the footprints of the image collection and zone raster
+  var fpIc=ee.Image(imCollection.first()).geometry();
+
+  var fpZone=zones.geometry();
+  print("IC footprint",fpIc);
+  print('Zone footprint',fpZone);
+
+  // map function that reduce the regions for each image
+  var fcOfFcs=imCollection.map(function(im){
+    var numOfBands=im.bandNames().length();
+    var imWithZones = im.addBands(zones);
+
+    // Using group reducer to reduce on raster zones.
+    //var clipRegions=zones.geometry();
+    var d = imWithZones.reduceRegion(reducer.group({groupField: numOfBands,groupName: 'Zone',}),
+            clipRegions, scale);
+
+    //make a list of features (with no geometry) from returned dictionary
+    var zoneList=ee.List(d.get('groups'));
+    var fList=zoneList.map(function(e){
+      return ee.Feature(null,e);
+    });
+    // make a feature collection from the dumb features
+    var fcOut = ee.FeatureCollection(fList);
+
+    // add image start and end time stamp to the features
+    var imPropertyNames=im.propertyNames();
+    if(imPropertyNames.contains("system:time_start")){
+      var imStartDate = ee.Date(im.get("system:time_start")).format(timeStampFormat);
+      fcOut = fcOut.map(function(feat) {
+        return feat.set(timeStampName+"_start",imStartDate);
+      });
+    }
+    if(imPropertyNames.contains("system:time_end")){
+      var imEndDate = ee.Date(im.get("system:time_end")).format(timeStampFormat);
+      fcOut = fcOut.map(function(feat) {
+          return feat.set(timeStampName+"_end",imEndDate);
+        });
[…]

### 2016-01-15 js/vegRemain.js

@@ -1,0 +1,23 @@
+// EarthEngine script to filter out converted areas from VECEA potential veg
+var vecea = ee.Image('users/lyndonestes/africa/vecea');
+var lc = ee.Image('users/lyndonestes/africa/zambia_lc_2010_II');
+var zamgrid = ee.Image('users/lyndonestes/africa/zamgrid');
+var zambia = ee.FeatureCollection('ft:10YdSHdN3XgDoRFDyPLecjjyV_5iQqGEymorahxMD', 'geometry');
+
+// select out converted areas or water bodies
+var cropland = lc.eq(8);
+var settled = lc.eq(11);
+var water = lc.eq(10);
+var convtotal = cropland.add(settled); // converted areas
+
+// start changing potential vegetation
+var vegactual = vecea.where(convtotal.eq(1), 0); // remove crop+urban
+var vegactual = vegactual.where(vegactual.eq(255), 0); // set 255 to 0
+
+var zoneScale = vegactual.projection().nominalScale().getInfo();
+
+Export.image(vegactual, 'vegremain', {
+  region: zambia,
+  scale: zoneScale,
+  maxPixels: 1500000000
+});

### 2016-01-15 python/veceaToRaster.py

@@ -1,0 +1,21 @@
+import arcpy
+import time
+
+try:
+    rpath = 'C:\\Users\\lestes\\Dropbox\\'
+    vpath = rpath + 'publications\\zambiaToff\\private\\zambiaToff\\' + \
+        'external\\input_devel\\vegetation\\'
+    itiff = rpath + 'data\\landcover\\' + 'Zambia_LandCover_2010_Scheme_II\\'
+    shp = rpath + 'data\\distributions\\vegetation\\' + \
+        'potential\\vecea\\Zambia\\pnv_vecea_v2_0_Zambia.shp'
+    otiff = rpath + vpath + 'vecea.tif'
+
+    e=arcpy.Describe(gtiff).extent
+    arcpy.env.extent = "%s %s %s %s" %(e.XMin, e.YMin, e.XMax, e.YMax)
+
+    arcpy.PolygonToRaster_conversion(shp, "CAT", otiff, "MAXIMUM_AREA", "CAT", \
+        0.0002778558)
+
+except Exception:
+    e = sys.exc_info()[1]
+    print(e.args[0])

### 2016-01-15 python/vegridToRaster.py

@@ -1,0 +1,41 @@
+# Calculate areas in veggrid intersect and create raster of areas for each
+# vegetation type in each 1km^2 cell
+
+import arcpy
+import time
+
+try:
+    bpath='C:\\Users\\lestes\\Dropbox\\publications\\zambiaToff\\private\\zambiaToff\\external\\input_devel\\biodiversity'
+    inshp=bpath+ "\\veggrid.shp"
+
+    print "Started script at " + time.strftime("%d %H:%M:%S")
+    fldname="CAT"
+    myList = set((row.getValue(fldname) for row in arcpy.SearchCursor(inshp, fields=fldname)))
+    e=arcpy.Describe(inshp).extent  # extent of inshape
+    arcpy.env.extent = "%s %s %s %s" %(e.XMin, e.YMin, e.XMax, e.YMax)
+
+    for i in myList:
+
+        print "Processing veg type " + str(i)
+
+        # Variables
+        fstr = "veg" + str(i)
+        fnm = bpath + "\\" + fstr + ".tif"
+        shpnm = "results" + str(i)# bpath + "\\" + fstr + ".shp"
+        query = "\"CAT\"="+str(i)
+
+        # Subset polygons and add area column to it, calculate area
+        print "..calculating areas"
+        arcpy.FeatureClassToFeatureClass_conversion(inshp, "in_memory", shpnm, query)
+        arcpy.AddField_management(shpnm, "area", "INTEGER", 0, 3)
+        arcpy.CalculateField_management(shpnm, "area", "!shape.Area@hectares!", "PYTHON_9.3")
+
+        # Finally, write out area to raster
+        print "..writing raster"
+        arcpy.PolygonToRaster_conversion(shpnm, "area", fnm, "MAXIMUM_AREA", "area", 1000)
+
+    print "Finished script at " + time.strftime("%d %H:%M:%S")
+
+except Exception:
+    e = sys.exc_info()[1]
+    print(e.args[0])

### 2016-01-18 Rmd/biodiversity-layer.Rmd
Created new biodiversity priority raster

---
title: "Biodiversity layer prep"
author: "Lyndon Estes"
date: "January 15, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Assembling new biodiversity layer

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_bd <- full_path(p_proj, "external/input_devel/biodiversity")
p_dat <- "/Users/lestes/Dropbox/data/zari/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_findat <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/"
p_pas <- full_path(p_dat, "landcover/Spatial Data_FD_ILUAII")

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
vegrem <- raster(full_path(p_bd, fname = "vegremain.tif"))
msk <- raster(full_path(p_basedat, "ZA-mask.tif"))
mskdt <- fread(full_path(p_findat, "ZA-mask.csv"))
nps <- spTransform(readOGR(full_path(p_pas, "Nat_Parks_GMAs.shp"), 
                           layer = "Nat_Parks_GMAs"), mgrid@crs)
freserves <- spTransform(readOGR(full_path(p_pas, "Forest_Reserves.shp"), 
                                 layer = "Forest_Reserves"), mgrid@crs)
vecea <- readOGR(full_path(p_bd, "veceagcs.shp"), layer = "veceagcs",
                 stringsAsFactors = FALSE)

# assign threat matrix rank from van Breugel et al (2014) 
veceatypes <- unique(vecea@data)
veceatypes <- veceatypes[, c(1, 3)]
veceatypes[order(veceatypes$CODE), ]
vecea@data[vecea@data$CODE == "wd", ][1, ]
endang <- c("VU", "EN", "", "VU", "NA", "VU", "", "", "VU", "EN", "NA", "drop",
            "VU", "VU", "", "VU", "VU", "VU", "", "", "VU")
veceatypes$THRT <- endang
thrtmat <- cbind(unique(endang), c(3, 2, 4, 5, 6))
veceatypes$erank <- 0
for(i in 1:nrow(thrtmat)) {
  veceatypes[veceatypes$THRT == thrtmat[i, 1], "erank"] <- thrtmat[i, 2]
}

```

## Prepare impacts layer
```{r, eval = FALSE}
fnms <- full_path(full_path(p_bd, "act_veg"), 
                  dir(full_path(p_bd, "act_veg"), pattern = ".tif$"))
s <- stack(lapply(fnms, function(x) raster(x)))
sm <- mask(s, mgrid)

# Sum for total - the inverse of this will be the farmable areas
remfrac <- calc(s, sum)
remfrac <- mask(remfrac, mgrid)

# Calculate intactness
w <- matrix(nrow = 11, ncol = 11)
w[] <- 1
intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
# intact <- mask(intact, mgrid)
# plot(intact)

# Calculate a veg type weight based on relative remaining area
veg_dt <- dtraster::as.data.table(sm)  # stack and convert to dt
veg_dt <- na.omit(veg_dt)
vnms <- names(veg_dt)[-c(1:2)]
vtots <- veg_dt[, lapply(.SD, function(x) sum(x * 100)), 
                .SDcols = vnms]  
wgti <- round(1 / (vtots / nrow(veg_dt)), 2)  # bd weights (1 = rarest)
# wgt <- round(vtots / nrow(veg_dt), 4)  # biodiversity weights (1 = most common)
# plot(unlist(wgti), unlist(wgt))

# Check the relative areas against histograms from 30 m layer
v <- gdalinfo(vegrem@file@name, hist=TRUE)[31]  # gdalinfo histogram
v <- as.numeric(strsplit(v, " ")[[1]])  # convert to numeric, get numbers
v <- v[!is.na(v)]  # remove NAs
lcvals <- c(0:(length(v) - 1))[which(v != 0)]  # what are the actuals, less 0
hist30 <- cbind(lcvals, v[v != 0])

# check to see if similar to data.table derived versions (checks out)
plot(round(hist30[-1, 2] / sum(hist30[-1, 2]) * 100, 2), vtots / nrow(veg_dt))

```

## New impacts layer

Will be intactness, PAs, and rarity weights combined

### Weighting in relation to rarity and protectedness
```{r, eval = TRUE}
# Combine protected areas (NPs and GMAs, forest reserves)
nps$code <- ifelse(nps$TYPE == "gma", 2, 1) 
npsr <- rasterize(nps, mgrid, field = "code")  # rasterize NPs/GMAs
npsr[is.na(npsr)] <- 0  
freserves$code <- 3  
freservesr <- rasterize(freserves, mgrid, field = "code")  # forest reserves
freservesr[is.na(freservesr)] <- 0
pas <- npsr + freservesr  # sum them
reclmat <- cbind(0:5, c(0:3, 1, 2))
pas <- reclassify(pas, reclmat)
pas <- mask(pas, mgrid)
# plot(pas)
# plot(freservesr)

# Let's look at how much veg is protected, and maybe use that as basis for 
# weights - rarity + proportion protected
smpa <- mask(stack(s, pas), mgrid)
vegpa_dt <- as.data.table(smpa)
vegpa_dt <- na.omit(vegpa_dt)
setnames(vegpa_dt, "layer", "PA")
protected <- vegpa_dt[, lapply(.SD, sum(x * 100)), by = PA, .SDcols = vnms]
protected <- cbind("PA" = protected[, PA], 
                   protected[, lapply(.SD, function(x) {
                       round(x / sum(x), 3)
                   }), .SDcols = vnms])
# proportion protected in both NPs and GMA, and just in NPs
prop_prot2 <- protected[PA %in% c(1:2), lapply(.SD, sum), .SDcols = vnms]
prop_prot1 <- protected[PA == 1, lapply(.SD, sum), .SDcols = vnms]  # nps only

# stack PAS with intact area
pasconvs <- mask(stack(pas, remfrac), mgrid)
pa_dt <- as.data.table(pasconvs)
pa_dt <- na.omit(pa_dt)
setnames(pa_dt, c("layer.1", "layer.2"), c("pas", "nveg"))
pa_dt[, (1 - sum(nveg) / .N) * 100 , by = "pas"]
pa_dt[, .N, by = "pas"]

# how much of PAs converted
# 15.8 % outside of PAs converted (472,429 km2)
# 0.9 % NPs converted (60,201 km2)
# 3.8 % in GMAs converted (165228 km2)
# 8.3 % forest reserves converted (54091 km2)

# Weighting scheme, factoring in protection status
wgts <- unlist(log(wgti))  # log transform weights
wgtsn <- (wgts - min(wgts)) / diff(range(wgts))
normalize <- function(x) (x - min(x)) / diff(range(x))
wgts_prop <- normalize(unlist(wgtsn + (1 - prop_prot2)))
wgts_propl <- normalize(log(unlist(wgtsn + (1 - prop_prot2))))
wgts_prop_np <- normalize(unlist(wgtsn + (1 - prop_prot1)))

# plots to examine them
labs <- gsub("veg", "", names(wgtsn))
# par(mfrow = c(2, 1), mar = c(2, 2, 2, 2))
# weights against protected proportion (NPs + GMAs)
plot(wgtsn, prop_prot2, pch = "", ylab = "")
text(wgtsn, prop_prot2, labels = labs, cex = 0.75)
# weights against protected proportion (NPs only)
plot(wgtsn, prop_prot1, pch = "", ylab = "")
text(wgtsn, prop_prot1, labels = labs, cex = 0.75)
# additive weights factoring in protected proportion versus protected proportion
plot(wgts_prop2, prop_prot, pch = "", ylab = "")
text(wgts_prop2,  prop_prot, labels = labs, cex = 0.75)
# BD + NP protected proportion versus NP protected proportion
plot(wgts_prop_np, prop_prot1, pch = "", ylab = "")
text(wgts_prop_np,  prop_prot1, labels = labs, cex = 0.75)
# protected proportion weighted versus same but log-transformed
plot(wgts_prop, wgts_propl, pch = "", ylab = "")
text(wgts_prop,  wgts_propl, labels = labs, cex = 0.75)
# Bd weights against BD + protected proportion weights
plot(wgtsn, wgts_prop, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop, labels = labs, cex = 0.75)
# Bd weights against BD + NP protected proportion weights
plot(wgtsn, wgts_prop_np, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn, wgts_prop_np, labels = labs, cex = 0.75)
# Bd weights versus log-transformed BD + protected proportion weights
plot(wgtsn, wgts_propl, pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgts_propl, labels = labs, cex = 0.75)
# The same but against multiplicative weights
plot(wgtsn, wgtsn * unlist(1 - prop_prot), pch = "", ylab = "")
lines(c(0, 1), c(0, 1))
text(wgtsn,  wgtsn * unlist(1 - prop_prot), labels = labs, cex = 0.75)
# sort(wgts_prop); sort(wgts_prop_np)
# sort(wgtsn)

# check against veceatypes von Breugel categories
veceatypes$wgtsn <- wgtsn[match(veceatypes$CAT, labs)]
veceatypes$wgtsnp <- wgts_prop[match(veceatypes$CAT, labs)]
plot(veceatypes$erank, veceatypes$wgtsn)  # not much connection
plot(veceatypes$erank, veceatypes$wgtsnp)  # not much connection
```

We will use the weights for rarity adjusted by how much the vegetation type falls in parks -- this corresponds to von Breugel risk class A C2 in Table 2.

### Reweight with intactness
```{r, eval = FALSE}
# assign weights to veg types
bdw <- wgts_prop_np
bd1 <- copy(veg_dt)
for(i in names(bdw)) {
  wi <- bdw[i]
  di <- eval(parse(text = paste0("quote(", i, ")")))
  bd1[eval(di) > 0, (i) := wi]
}

# Reduce to single column that takes the maximum weight across vegetation types
bd1[, TID := 1:.N ]
setkey(bd1, "TID")
bd1[, MWGT := apply(.SD, 1, max), .SDcols = vnms]
bd1r <- dt_to_raster(bd1[, .(x, y, MWGT)], mgrid@crs)
# plot(bd1r)
# plot(remfrac == 0)
# plot(smpa[["veg68"]])

# Stack vprior
bdstack <- stack(list("veg" = bd1r, "intact" = crop(intact, bd1r)))
bdr <- calc(bdstack, sum)
bdr <- (bdr - minValue(bdr)) / diff(cellStats(bdr, range, na.rm = TRUE))
# plot(bdr)

# Filter out PAs
# 0.9 % NPs converted (60,201 km2) - set NPs to 0
# 3.8 % in GMAs converted (165228 km2)  - set them to 4 %
# 8.3 % forest reserves converted (54091 km2) - set them to 8.3
pamat <- cbind(1:3, c(1, 1 - 0.038, 1 - 0.083))
pafix <- lapply(1:3, function(x) {  # x <- 2
  pamsk <- crop((pas == x) * pamat[x, 2], bdr)
  bdmsk <- mask(bdr, pamsk, maskvalue = 0)
  bdcalc <- calc(stack(bdmsk, pamsk), max)
})
pafix <- calc(stack(pafix), sum)  # sum those

# final biodiversity layer
bdf <- calc(stack(bdr, pafix), max)
plot(remfrac)
plot(bdf)

writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"))

```

### 2016-01-19 Rmd/carbon.Rmd
Created new carbon layers

---
title: "Carbon datasets, updated"
author: "Lyndon Estes"
date: "January 18, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# New carbon layer
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(GSIF)

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_dat <- "/Users/lestes/Dropbox/data/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
# p_findat <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/"
p_c <- full_path(p_dat, "carbon/baccini/africa/")
p_devel <- full_path(p_proj, "external/input_devel/carbon")
p_soc <- full_path(p_dat, "soil/carbon/SoilGrids1km")
p_bd <- full_path(p_dat, "soil/bulkdens/SoilGrids1km")

# load datasets
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
vcarb <- raster(full_path(p_c, "biomass-C.tif"))
ocarb <- brick(full_path(p_basedat, "ZA-carbon.tif"))  # original carbon
names(ocarb) <- c("veg", "soil", "total")
# plot(ocarb < 30)

```

## Vegetation carbon
```{r, eval=FALSE}
onm <- full_path(p_devel, "veg-carbon.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = vcarb@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
zvcarb <- raster(onm)  # vegetation carbon
plot(zvcarb)

```

## Soil carbon

ISRIC WISE grids, from [here](http://soilgrids1km.isric.org), with units and ancillary information from [here](http://www.isric.org/content/faq-soilgrids#How_much_does_it_cost). Conversion to t/ha from [here](http://www.soilquality.org.au/factsheets/organic-carbon) and [here](https://www.agric.wa.gov.au/climate-change/what-soil-organic-carbon?page=0%2C1).
```{r, eval = FALSE}
crpshp <- as(extent(mgrid), "SpatialPolygons")
crpshp@proj4string <- mgrid@crs

# organic carbon 
socr <- lapply(dir(p_soc, full.names = TRUE), function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs))
  r / 10  # convert from permille to percent
})
# plot(socr[[1]])

# bulk density converted kg / m3 (or g / cm3)
bd <- lapply(dir(p_bd, full.names = TRUE), function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs))
  r  / 1000  # convert from Mg/m3 to kg/m3
})
# plot(bd[[1]] * 1000)

# layer depth
depths <- c(0, 5, 15, 30, 60, 100, 200) 
thickness <- depths[-1] - depths[-length(depths)]

# plot(round(bd[[1]], 1) == 1.2)
# tst <- (socr[[1]] == 1.3 & round(bd[[1]], 1) == 1.2)
# 1.2 * 1.2 * 10
# 10000 * 0.1
# depths / 100

# calculate carbon in each layer
soct <- lapply(1:length(thickness), function(x) {
  socr[[x]] * bd[[x]] * thickness[x]
})

# then total for top 1 m
soct_1m <- calc(stack(soct[1:5]), sum)
soct_1ma <- projectRaster(soct_1m, mgrid)
writeRaster(soct_1ma, filename = full_path(p_devel, "soil-C-1m.tif"))

# plot(mask(soct_1ma, mgrid))
# plot(ocarb$soil)
# plot(mask(soct_1ma, mgrid) - ocarb$soil)  # same order of magnitude

zamcarb <- brick(stack(zvcarb, soct_1ma), 
                 filename = full_path(p_devel, "ZA-carbon.tif"))

```

### 2016-01-19 Rmd/to-do.Rmd
Created new carbon layers

@@
              "Explore econ versus enviro goals", 
-             "Better cropland map, yield estimates"), 
+             "Better cropland map, yield estimates", 
+             "Updated carbon layers"), 
   "Notes" = c("Single production to follow (Alastair)", 
@@
               "See where map overlaps are greatest, most divergent", 
-              "Upgrade all maps to SERVIR landcover, and rebase yield estimates on that"),
-  "Doing" = c("D", "D", "U", "U", "D", "D", "D", "L", "D", "D")
+              "Upgrade all maps to SERVIR landcover, and rebase yield estimates on that", "Baccini vegetation carbon + ISRIC grids"),
+  "Doing" = c("D", "D", "U", "U", "D", "D", "D", "L", "D", "D", "F")
 )

### 2016-01-19 Rmd/biodiversity-layer.Rmd
Created new carbon layers
Updated cropland datasets, calculating current productivity for maize and soybean from FAO and USDA statistics.

@@
 
+# keep for assessing impacts
+writeRaster(sm, filename = full_path(p_bd, "ZA-vegremain.tif"), overwrite = TRUE)
+
 # Sum for total - the inverse of this will be the farmable areas
@@
 remfrac <- mask(remfrac, mgrid)
+writeRaster(remfrac, filename = full_path(p_bd, "ZA-remfrac.tif"))
 
@@
 intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
+writeRaster(intact, filename = full_path(p_bd, "intactness.tif"))
+
 # intact <- mask(intact, mgrid)
@@
                 .SDcols = vnms]  
+1 / wgti
 wgti <- round(1 / (vtots / nrow(veg_dt)), 2)  # bd weights (1 = rarest)
@@
 # The same but against multiplicative weights
-plot(wgtsn, wgtsn * unlist(1 - prop_prot), pch = "", ylab = "")
+plot(wgtsn, wgtsn * unlist(1 - prop_prot1), pch = "", ylab = "")
 lines(c(0, 1), c(0, 1))
-text(wgtsn,  wgtsn * unlist(1 - prop_prot), labels = labs, cex = 0.75)
+text(wgtsn,  wgtsn * unlist(1 - prop_prot1), labels = labs, cex = 0.75)
 # sort(wgts_prop); sort(wgts_prop_np)
@@
 bdf <- calc(stack(bdr, pafix), max)
-plot(remfrac)
-plot(bdf)
+# plot(remfrac)
+# plot(bdf)
 
-writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"))
+writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"), overwrite = TRUE)
+
+vprops <- cbind("var" = c("area", "rarity", "priority"),
+                rbind(vtots / nrow(veg_dt), wgti, data.table(t(wgts_prop_np))))
+save(vprops, file = full_path(p_bd, "vprops.rda"))
 

### 2016-01-19 Rmd/cropland.Rmd
Updated cropland datasets, calculating current productivity for maize and soybean from FAO and USDA statistics.

@@
 ---
-title: "cropland.Rmd"
+title: "Current Cropland, Production, and Yield"
 author: "Lyndon Estes"
@@
                     "base_data/")
+p_dat <- "~/Dropbox/data/agricultural/fao/countrystat/zambia/"
+p_usda <- "~/Dropbox/data/agricultural/usda/fas/zambia/"
 p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"
@@
          of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
-r <- raster(onm)
-par(mar = rep(0, 4)); plot(r)
+cropland <- raster(onm)
+par(mar = rep(0, 4)); plot(cropland)
 
 ```
+
+## Cropland statistics
+
+From FAO countrystat at this [link](http://www.countrystat.org/home.aspx?c=ZMB&tr=21).
+```{r, eval = FALSE}
+# select crop stats
+fnms <- dir(p_dat, pattern = "10|16", full.names = TRUE)
+vars <- c("prod", "pa")
+fao <- rbindlist(lapply(1:length(fnms), function(x){
+  DT <- fread(fnms[x], stringsAsFactors = FALSE)
+  cbind(vars[x], DT)
+})) 
+fao[, Value := as.numeric(Value)]
+setnames(fao, old = c("V1", "administrative level 1 (code)", 
+                      "administrative level 1", "Crop (code)"), 
+         new = c("var", "provcode", "prov", "cropcode"))
+setkeyv(fao, c("var", "Year"))
+
+# missing 2008 NW province maize and 2011 Southern Province, same for soybean
+fao[Crop == "Maize" & var == "prod" & is.na(Value)]  
+fao[Crop == "Soybeans" & var == "prod" & is.na(Value)]
+
+# corresponding FAOStat full country values are 1211566 and 3020380 for maize
+# 56839 and 116539 for soybean
+nwrepl <- fao[Crop == "Maize" & var == "prod" & Year == "2008" & 
+                provcode != 3431, 1211566 - sum(Value)]
+fao[Crop == "Maize" & var == "prod" & Year == "2008" & 
+                provcode == 3431, Value := nwrepl]
+srepl <- fao[Crop == "Maize" & var == "prod" & Year == "2011" & 
+               provcode != 3433, 3020380 - sum(Value)]
+fao[Crop == "Maize" & var == "prod" & Year == "2011" & 
+                provcode == 3433, Value := srepl]
+nwrepl <- fao[Crop == "Soybeans" & var == "prod" & Year == "2008" & 
+                provcode != 3431, 56839 - sum(Value)]
+fao[Crop == "Soybeans" & var == "prod" & Year == "2008" & 
+                provcode == 3431, Value := nwrepl]
+srepl <- fao[Crop == "Soybeans" & var == "prod" & Year == "2011" & 
+               provcode != 3433, 116539 - sum(Value)]
+fao[Crop == "Soybeans" & var == "prod" & Year == "2011" & 
+                provcode == 3433, Value := srepl]
+
+# Planted area and production 2010-2014 for Zambia
+muprod <- lapply(c("Maize", "Soybeans"), function(x) {
+  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
+      by = Year]
+})
+names(muprod) <- c("maize", "soy")
+mupa <- lapply(c("Maize", "Soybeans"), function(x) {
+  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
+      by = Year]
+})
+names(mupa) <- c("maize", "soy")
+
+# Read in USDA harvested area and production to get difference in harvested area
+# and planted area
+fnms <- dir(p_usda, full.names = TRUE)
+vars <- c("maize", "soy")
+usda <- rbindlist(lapply(1:length(fnms), function(x){
+  DT <- fread(fnms[x], stringsAsFactors = FALSE, header = TRUE)
+  # cbind(vars[x], DT)
+}))
+cols <- names(usda)[4:9]
+usda_ylds <- usda[var == "prod", cols, with = FALSE] / 
+  usda[var == "harvarea", cols, with = FALSE]
+
+# FAO references start year, it seems for maize, but soybean is correct year
+yp <- unlist(usda[var == "prod" & crop == "maize", cols, with = FALSE] * 1000)
+plot(yp, muprod[[1]]$V1[-c(1, 8)])  # but essentially same yields
+yp <- unlist(usda[var == "prod" & crop == "soybean", cols, with =FALSE] * 1000)
+plot(yp, muprod[[2]]$V1[-c(1:2)])
+
+# Production estimates ~2010/1-2015
+currprod <- round(c("maize" = muprod$maize[Year %in% 2010:2014, mean(V1)], 
+                    "soy" = muprod$soy[Year %in% 2011:2015, mean(V1)]))
+save(currprod, file = full_path(p_devel, "currprod.rda"))
+
+```
+
+## Resulting model inputs
+
+`currprod.rda`, `cropland-1k.tif`
+
+Eventually might want to disaggregate yields onto cropland map, but will need provincial harvested area for that.  Model might not really need that, however.

### 2016-01-20 Rmd/all-inputs.Rmd
Many updates as getting ready to assemble and mask inputs

---
title: "Assemble Inputs"
author: "Lyndon Estes"
date: "January 20, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Assemble Inputs

Bring all datasets in, create farmable area mask, mask them, and convert to data.tables for model input

1. For running model, non-grid: 
    + `currprod.rda` - For input, not data.table
    + `vprops.rda` - table of veg type area, rarity value, and prioritization value.

2. For running model:
    + Prioritization (P) and impact (I)
        + `ZA-carbon.tif` - Carbon impacts, to data.table (P|I)
        + `bdprior.tif` - biodiversity prioritization grid (P)
        + `ZA-cost.tif` - cost distance
        + `maize|soybean_S|R|I_10km,tif` - variants thereof (P|I). Separate variants of these must be created prior to model runtime. A function will be creatd here to facilitate that.  
    + Impact only:
        + `ZA-vegremain.tif` - vegetation remaining grid, for assessing conversion impacts. 
    + Masking/allocation
        + `farmable-1k.tif` - slope-based farmable area mask
        + `water-1k.tif` - water-based farmable area mask
        + `ZA-remfrac.tif` - cropland and urban-based farmable area mask
3. For displaying model outputs: 
    + `cropland-1k.tif` - cropland mask

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/cropland")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_crop <- full_path(p_proj, "external/input_devel/cropland")
p_bd <- full_path(p_proj, "external/input_devel/biodiversity")
p_cost <- full_path(p_proj, "external/input_devel/roads/cost")

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
cropland <- raster(full_path(p_crop, fname = "cropland-1k.tif"))
bd <- raster(full_path(p_bd, fname = "bdprior.tif"))
vegr <- raster(full_path(p_bd, fname = "ZA-vegremain.tif"))
cost <- raster(full_path(p_cost, fname = "zamcost.tif"))

```

### 2016-01-20 Rmd/biodiversity-layer.Rmd
Many updates as getting ready to assemble and mask inputs

--- 2016-01-18 Rmd/biodiversity-layer.Rmd
@@
 ```{r, eval = FALSE}
-fnms <- full_path(full_path(p_bd, "act_veg"), 
-                  dir(full_path(p_bd, "act_veg"), pattern = ".tif$"))
+vnms <- dir(full_path(p_bd, "act_veg"), pattern = ".tif$")
+vnms <- vnms[vnms != "veg88.tif"]  # drop water bodies
+fnms <- full_path(full_path(p_bd, "act_veg"), vnms)
 s <- stack(lapply(fnms, function(x) raster(x)))
 sm <- mask(s, mgrid)
+
+# keep for assessing impacts
+writeRaster(sm, filename = full_path(p_bd, "ZA-vegremain.tif"), 
+            overwrite = TRUE)
 
@@
 remfrac <- mask(remfrac, mgrid)
+writeRaster(remfrac, filename = full_path(p_bd, "ZA-remfrac.tif"), 
+            overwrite = TRUE)
 
@@
 intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
+writeRaster(intact, filename = full_path(p_bd, "intactness.tif"), 
+            overwrite = TRUE)
+
 # intact <- mask(intact, mgrid)
@@
 hist30 <- cbind(lcvals, v[v != 0])
+hist30 <- hist30[-21, ]  # type 88
 
@@
 ### Weighting in relation to rarity and protectedness
-```{r, eval = TRUE}
+```{r, eval = FALSE}
 # Combine protected areas (NPs and GMAs, forest reserves)
@@
 # additive weights factoring in protected proportion versus protected proportion
-plot(wgts_prop2, prop_prot, pch = "", ylab = "")
-text(wgts_prop2,  prop_prot, labels = labs, cex = 0.75)
+plot(wgts_prop, prop_prot2, pch = "", ylab = "")
+text(wgts_prop,  prop_prot2, labels = labs, cex = 0.75)
 # BD + NP protected proportion versus NP protected proportion
@@
 bdf <- calc(stack(bdr, pafix), max)
-plot(remfrac)
-plot(bdf)
+# plot(remfrac)
+# plot(bdf)
 
-writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"))
+writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"), overwrite = TRUE)
+vprops <- cbind("var" = c("area", "rarity", "priority"),
+                rbind(vtots / nrow(veg_dt), wgti, data.table(t(wgts_prop_np))))
+save(vprops, file = full_path(p_bd, "vprops.rda"))
 
 ```
+
+## Resulting model inputs
+
+`vprops.rda` - table of veg type area, rarity value, and prioritization value
+
+`bdprior.tif` - biodiversity prioritization grid
+
+`ZA-vegremain.tif` - vegetation remaining grid, for assessing conversion impacts
+
+`ZA-remfrac.tif` - to combine with other farmable area masks

### 2016-01-20 Rmd/cost-distance.Rmd
Various scripts for creating terrain, cropland, and actual vegetation layers
Many updates as getting ready to assemble and mask inputs

@@
 library(gdalUtils)
-p_root <- set_base_path("zambiaToff")
+p_root <- lmisc::set_base_path("zambiaToff")
 p_roads <- full_path(p_root, "external/input_devel/roads/")
@@
 p_trade <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/"
+p_cost <- full_path(p_root, "external/input_devel/roads/cost")
 
@@
 cnm <- "/Volumes/Data/loMadzala/GIS/Zambia/costdistance/zamcost2.tif"
-file.copy(cnm, "cost.tif", overwrite = TRUE)
+file.copy(cnm, full_path(p_cost, "zamcost.tif"), overwrite = TRUE)
 
 # read in the cost raster
-costr <- raster("cost.tif")
+costr <- raster(full_path(p_cost, "zamcost.tif"))
 
@@
 library(agroEcoTradeoff)
-dt_path <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/"
-msk <- fread(paste0(dt_path, "ZA-mask.csv"))  # read it in
-mskr <- dt_to_raster(msk, CRSobj = projection(zamr))  # convert to raster
+# dt_path <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/"
+# msk <- fread(paste0(dt_path, "ZA-mask.csv"))  # read it in
+# mskr <- dt_to_raster(msk, CRSobj = projection(zamr))  # convert to raster
+mgrid <- raster(full_path(p_trade, "ZA-grid.tif"))
 
@@
 # another, and then mask it
-costrcr <- crop(costr, mskr)
-costcrrs <- resample(costrcr, mskr)
-costrmsk <- mask(costcrrs, mskr)
-plot(costrmsk)  # voila
+costrcr <- crop(costr, mgrid)
+costcrrs <- resample(costrcr, mgrid)
+# costrmsk <- mask(costcrrs, )
+# plot(costcrrs)  # voila
+
+writeRaster(costcrrs, filename = full_path(p_cost, "ZA-cost.tif"))
 
 # then convert it to a data.table as follows
-costdt <- na.omit(as.data.table(costrmsk, xy = FALSE))
+# costdt <- na.omit(as.data.table(costrmsk, xy = FALSE))
 
 # and then you will write it out as a csv file.  
-fnm <- "../../data/ZA-cost.csv"
-write.table(costdt, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+# fnm <- "../../data/ZA-cost.csv"
+# write.table(costdt, file = fnm, sep = ",", col.names = TRUE, 
+#             row.names = FALSE)
+
 ```
 
+## Resulting model inputs
+
+Alistair's project:
 Your final output csv file will want to replace the ZA-cost.csv file in agroEcoTradeoffs folder `external/data/dt`, and then the model will work nicely.
+
+Full analysis: 
+`ZA-cost.tif` - for masking, etc

### 2016-01-20 Rmd/maize-soy-grid.Rmd
Many updates as getting ready to assemble and mask inputs

@@
 ```{r, eval = FALSE}
-library(data.table)
+library(agroEcoTradeoff)
+# library(data.table)
 library(lmisc)
-library(dtraster)
+# library(dtraster)
 # devtools::install_github("PrincetonUniversity/lmisc")
 
-p_root <- set_base_path()
+p_root <- lmisc::set_base_path()
 p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
@@
 p_agro <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/"
+p_yield <- full_path(p_root, "external/input_devel/yield")
 
@@
 
-plot(dt_to_raster(IDr[, .(x, y, as.factor(Zone))], CRS(prj)))
 prj <- paste0("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0",
@@
 gcs <- "+proj=longlat +datum=WGS84"
+plot(dt_to_raster(IDr[, .(x, y, as.factor(Zone))], CRS(prj)))
 
@@
 # across cultivars, or median, or 90th, etc.
+
+# fill in yield at edges
+yield_fix <- function(x = r[[1]], w, na.rm = TRUE) {
+  r <- focal(x, w = w, fun = mean, na.rm = na.rm)
+  rmsk <- (!is.na(r) & is.na(x)) * r
+  x[is.na(x)] <- 0
+  rfix <- x + rmsk  # add back in edge bits
+}
+
+# ra
+yld_rl <- lapply(crops, function(x) {  # x <- crops[[1]]
+  yr <- lapply(c("S", "R", "I"), function(y) {  #y <- "S"
+    ylds <- yld_rasts[[x]]
+    nms <- names(ylds)
+    cnms <- nms[grep(paste0("y.*._", y), nms)]
+    r <- dt_to_raster(ylds[, c("x", "y", cnms), with = FALSE], CRSobj = gcs)
+    w <- matrix(ncol = 3, nrow = 3, rep(1, 9))
+    rfoc <- sapply(1:nlayers(r), function(i) rfoc <- yield_fix(r[[i]], w = w))
+    names(rfoc) <- names(r)
+    rs <- stack(rfoc)
+    rs
+  })
+  named_out(yr, c("S", "R", "I"))
+})
+names(yld_rl) <- crops
+
+for(i in 1:length(yld_rl)) {
+  rl <- yld_rl[[i]]
+  for(j in 1:length(rl)) {
+    fnm <- full_path(p_yield, paste0(crops[i], "_", names(rl)[j], "_10K.tif"))
+    writeRaster(rl[[j]], filename = fnm)
+  }
+}
+```
+<a href="#top">Back to top</a>
+
+## How to make 1 km grids/data.tables
+
+This provides the methods for taking 10 km grids, averaging them across cultivars within management treatments, and then reseampling to 1 km grids.
+This is approximately the approach that will be used to make inputs for the tradeoff model.
+Other methods to explore include blending different types of management practice, e.g.
+X% of each cell is irrigated, Y% is rainfed commercial, Z% is subsistence.
+```{r, eval=FALSE}
 yld_mus <- lapply(crops, function(x) {  # x <- crops[[1]]
@@
 names(yld_mus) <- crops
+plot(dt_to_raster(yld_mus[[2]]$R, CRSobj = gcs))
 
-plot(dt_to_raster(yld_mus[[2]]$R, CRSobj = gcs))
-```
-<a href="#top">Back to top</a>
-
-# 1 km grids
-
-Average mean yields across cultivars within management treatments, and then reseample to 1 km grids. These become new inputs for the tradeoff model.  
-```{r, eval=FALSE}
+# 1 km resample
 mgrid <- raster(full_path(p_agro, "ZA-grid.tif"))
@@
 <a href="#top">Back to top</a>
+
+## Resulting model inputs
+
+`maize_S|R|I_10K.tif` - 10 km maize grids for different management
+`soybean_S|R|I_10K.tif` - 10 km soy grids for different management
+
+These will have to go through intermediate, and perhaps on the fly, processing in batch mode to represent different yield scenarios.

### 2016-01-20 Rmd/to-do.Rmd
Many updates as getting ready to assemble and mask inputs

@@
               "Upgrade all maps to SERVIR landcover, and rebase yield estimates on that", "Baccini vegetation carbon + ISRIC grids"),
-  "Doing" = c("D", "D", "U", "U", "D", "D", "D", "L", "D", "D", "F")
+  "Doing" = c("D", "D", "U", "U", "D", "D", "D", "L", "D", "F*", "F")
 )
@@
 ```
+
+*Just using total production estimates, which will not be mapped for now. These will be basis for estimating production targets.

### 2016-01-20 Rmd/water.Rmd
Many updates as getting ready to assemble and mask inputs

---
title: "Water mask"
author: "Lyndon Estes"
date: "January 19, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Water mask of Zambia

Water class from the SERVIR landcover map (2010)

## Inputs
```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(rworldxtra)
data("countriesHigh")

p_proj <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_devel <- full_path(p_proj, "external/input_devel/water")
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_dat <- "~/Dropbox/data/agricultural/fao/countrystat/zambia/"
p_usda <- "~/Dropbox/data/agricultural/usda/fas/zambia/"
p_lc <- "~/Dropbox/data/landcover/Zambia_LandCover_2010_Scheme_II/"

# Load in data
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
lc <- raster(full_path(p_lc, fname = "Zambia_LandCover_2010_Scheme_II.tif"))

```

## Extract water class and resample
```{r, eval = FALSE}

# pullout cropland class 8
system.time(water <- rast_math(lc == 10, tmpdir = p_devel))

onm <- full_path(p_devel, "water-1k.tif")
ext <- bbox(mgrid)[1:4]  # extent   
gdalwarp(srcfile = water@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
water <- raster(onm)
par(mar = rep(0, 4)); plot(water)

```

### 2016-01-21 Rmd/all-inputs.Rmd
Temporary fix for low-biased soybean yield with FAO GAEZ yields (bias-correction). In maize-soy-grid.Rmd. Processed resulting surfaces in all-inputs.Rmd to data.table inputs

@@
 p_cost <- full_path(p_proj, "external/input_devel/roads/cost")
+p_yield <- full_path(p_proj, "external/input_devel/yield")
+p_h20 <- full_path(p_proj, "external/input_devel/water")
+p_terr <- full_path(p_proj, "external/input_devel/terrain")
+p_carb <- full_path(p_proj, "external/input_devel/carbon")
+p_production <- full_path(p_proj, "external/data/production")
 
@@
 bd <- raster(full_path(p_bd, fname = "bdprior.tif"))
-vegr <- raster(full_path(p_bd, fname = "ZA-vegremain.tif"))
-cost <- raster(full_path(p_cost, fname = "zamcost.tif"))
+vegr <- brick(full_path(p_bd, fname = "ZA-vegremain.tif"))
+load(full_path(p_bd, fname = "ZA-vegremain.rda"))
+names(vegr) <- vegnames
+carbon <- brick(full_path(p_carb, fname = "ZA-carbon.tif"))
+names(carbon) <- c("veg", "soil")
+cost <- raster(full_path(p_cost, fname = "ZA-cost.tif"))
+ynms <- dir(p_yield, pattern = "tif")
+ynms <- sapply(c("S", "R", "I"), function(x) ynms[grep(x, ynms)])
+ynms <- ynms[c(1, 3, 5, 2, 4, 6)]
+ylds <- lapply(ynms, function(x) {
+  b <- brick(full_path(p_yield, x))
+  ylnms <- gsub("\\.tif", "\\.rda", x)
+  load(full_path(p_yield, ylnms))
+  names(b) <- layernms
+  b
+})
+names(ylds) <- gsub("_10K", "", ynms)
+
+# mask materials
+remfrac <- raster(full_path(p_bd, fname = "ZA-remfrac.tif"))
+water <- raster(full_path(p_h20, fname = "water-1k.tif"))
+farmable <- raster(full_path(p_terr, fname = "farmable-1k.tif"))
 
 ```
+
+## Create farmable area
+
+```{r, eval=FALSE}
+# merge together the three farmland available masks
+avail <- remfrac - water
+avail[avail < 0] <- 0
+avail <- avail - (1 - farmable)
+avail[avail < 0] <- 0
+avail <- avail * (bd < 0.962)  # remove PAs entirely (1 if nps, 0.962 if GMAs)
+# plot(avail)
+```
+
+That creates the farmland availability mask that is needed.
+
+## Stack and mask all grids
+```{r, eval = FALSE}
+
+# first get rid of all NA area in all grids
+# start with yield grids
+yldmask <- calc(stack(ylds), sum)
+yldmask2 <- disaggregate(yldmask, fact = 10)
+yldmask3 <- resample(yldmask, yldmask2)
+yldmask4 <- projectRaster(yldmask3, mgrid)
+yldmask5 <- mask(yldmask4, mgrid)
+plot(yldmask5)
+
+# now with other
+carbonf <- stack(lapply(1:nlayers(carbon), function(x) {
+  r <- carbon[[x]]
+  r[is.na(r)] <- 0
+  r
+}))
+
+mgridc <- crop(mgrid, avail)
+inlist <- list("convertible" = avail, "vegremain" = vegr, "carbon" = carbonf, 
+               "cost" = cost, "bd" = bd, "cropland" = cropland)
+inlist <- lapply(inlist, function(x) crop(x, mgridc))
+inlist2 <- lapply(inlist, function(x) calc(x, mean))
+inlist3 <- calc(stack(inlist2), mean)
+
+# filter NA values
+nafilt <- calc(stack(!is.na(inlist3), !is.na(crop(yldmask5, mgridc))), sum)
+# plot(nafilt)
+# plot(!is.na(calc(vegr, mean)))
+# plot(!is.na(crop(yldmask5, mgridc)), col = c("transparent", "red"), add = TRUE)
+msk <- nafilt == maxValue(nafilt)
+msk[msk == 0] <- NA
+
+# updated mask
+mgridf <- mask(mgridc, msk)
+valinds <- which(!is.na(values(mgridf)))
+outlist <- inlist[names(inlist) != "cropland"]
+
+# write out new data.tables
+cnames <- "ZA"  
+onms <- names(inlist)
+onms[onms == "bd"] <- "cons-priorities"
+for(i in 1:length(outlist)) { # i <- 1
+  fnm <- full_path(p_production, paste0(cnames, "-", onms[i], ".csv"))
+  DT <- as.data.table(outlist[[i]], xy = FALSE)[valinds, ]
+  write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+}
+
+# write out mask raster
+base <- as.data.table(msk, xy = TRUE)
+setnames(base, old = names(msk), new = "val")
+base[, ind := 1:nrow(base)]
+base <- base[!is.na(val)][, val := NULL]
+fnm <- full_path(p_production, paste0(cnames, "-mask.csv"))
+write.table(base, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+# plot(dt_to_raster(base, mgridf@crs))
+
+```
+
+## Yield grids
+```{r, eval = FALSE}
+# Function to disaggregate and reproject yield rasters
+yld_disagg <- function(x, msk, fact = 10) {
+  r <- disaggregate(x, fact = 10)
+  r2 <- resample(x, r)
+  r3 <- projectRaster(r2, msk)
+  r4 <- mask(r3, msk)
+  r4
+}
+
+# Function to aggregate yields within a particular yield brick
+# stat <- "50"; cults <- paste0("9900", c("08", "09", "10"))
+yield_stat <- function(x, stat = "mu", mgmnt = "S", cults, fun = mean) {
+  repstr <- paste0(paste0(stat, "_", mgmnt), cults, collapse = "|")
+  cnms <- names(x)[grep(repstr, names(x))]
+  print(cnms)
+  calc(x[[cnms]], fun = fun)
+}
+# rm(stat, mgmnt, cults, fun, x, cnms, repstr)
+
+# Let's make an initial yield combination for model development that is
+# is composed of: 
+# maize
+# 1. Subsistence management, all four cultivars, median yields, 50%
+# 2. Rainfed commerical, hybrid cultivars, median yields, 40%
+# 3. Irrigated maize, hybrid cultivars, median yields, 10%
+
+maize <- list("S" = yield_stat(x = ylds$maize_S.tif, "50", "S", 
+                               paste0("000", 1:4)), 
+              "R" = yield_stat(x = ylds$maize_R.tif, "50", "R", 
+                               paste0("000", 3:4)), 
+              "I" = yield_stat(x = ylds$maize_I.tif, "50", "I", 
+                               paste0("000", 3:4)))
+wgts <- c(0.5, 0.4, 0.1)
+maizewgt <- lapply(1:length(wgts), function(x) maize[[x]] * wgts[x])
+potmaize <- calc(stack(maizewgt), sum)  # potential maize yield
+
+# Soybean
+# 1. Subsistence management, all 3 cultivars, median yields, 20%
+# 2. Rainfed commerical, all 3cultivars, median yields, 70%
+# 3. Irrigated maize, all 3 cultivars, median yields, 10%
+soy <- list("S" = yield_stat(x = ylds$soybean_S.tif, "50", "S", 
+                             paste0("9900", c("08", "09", "10"))), 
+            "R" = yield_stat(x = ylds$soybean_R.tif, "50", "R", 
+                             paste0("9900", c("08", "09", "10"))), 
+            "I" = yield_stat(x = ylds$soybean_I.tif, "50", "I", 
+                             paste0("9900", c("08", "09", "10"))))
+wgts <- c(0.2, 0.7, 0.1)
+soywgt <- lapply(1:length(wgts), function(x) soy[[x]] * wgts[x])
+potsoy <- calc(stack(soywgt), sum)  # potential maize yield
+# plot(potmaize); cellStats(potmaize, mean)
+# plot(potsoy); cellStats(potsoy, mean)
+# plot(potmaize)
+# plot(potsoy)
+
+# Disaggregate potential yields, stack, and round
+ylds_1k <- lapply(list(potmaize, potsoy), function(x) {
+  yld_disagg(x, msk = mgridf)
+})
+ylds_s <- stack(ylds_1k) 
+names(ylds_s) <- c("maize", "soy")
+ylds_s <- round(ylds_s)
+
+# data.table them
+cnames <- "ZA"  
+fnm <- full_path(p_production, paste0(cnames, "-potential-yields.csv"))
+DT <- as.data.table(ylds_s, xy = FALSE)[valinds, ]
+write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+
+```

### 2016-01-21 Rmd/biodiversity-layer.Rmd
Temporary fix for low-biased soybean yield with FAO GAEZ yields (bias-correction). In maize-soy-grid.Rmd. Processed resulting surfaces in all-inputs.Rmd to data.table inputs

@@
             overwrite = TRUE)
+vegnames <- gsub("\\.tif", "", gsub("veg", "v", vnms))
+save(vegnames, file = full_path(p_bd, "ZA-vegremain.rda"))
 

### 2016-01-21 Rmd/cropland.Rmd
Temporary fix for low-biased soybean yield with FAO GAEZ yields (bias-correction). In maize-soy-grid.Rmd. Processed resulting surfaces in all-inputs.Rmd to data.table inputs

@@
                 provcode == 3433, Value := srepl]
+save(fao, file = full_path(p_devel, "fao-yields.rda"))
 
@@
 names(mupa) <- c("maize", "soy")
+# mean(muprod$soy[, V1] / mupa$soy[, V1])
 
@@
   usda[var == "harvarea", cols, with = FALSE]
+save(usda, file = full_path(p_devel, "usda-yields.rda"))
 

### 2016-01-21 Rmd/maize-soy-grid.Rmd
Temporary fix for low-biased soybean yield with FAO GAEZ yields (bias-correction). In maize-soy-grid.Rmd. Processed resulting surfaces in all-inputs.Rmd to data.table inputs

@@
 p_yield <- full_path(p_root, "external/input_devel/yield")
+p_cropland <- full_path(p_root, "external/input_devel/cropland")
+p_fao <- "~/Dropbox/data/agricultural/fao/aez/soybean/"
 
@@
 
-    # unique(ID$WTH)[!unique(ID$WTH) %in% unique(substr(cr$WTH, 1, 4))]
-
-    # cr[, .N, by = .(TNAM, SOL, Zone)][which(N > 62), ]
-    # tst <- cr[TNAM == "AAAX0003" & SOL == "WI_FLSO001", ][order(YR), ]
-    # plot(tst[YR > 31, HWAH], tst[YR <= 31, HWAH])
-    # cr[TNAM == "AAEA0001" & SOL == "WI_CMZR003", ]
-
     crm <- cr[ID[, .(WTH, SOL, SharePct, Zone)], allow.cartesian = TRUE]
@@
 
-    # anncv <- solmu[, round(cv(V1)), by = .(WTH, Zone)][, V1]
-    # setnames(anncv, "V1", "acv")
-
-    # annwv2[WTH == "AERM" & Zone == "Zone3",]
-    # crm[WTH == "AERM" & Zone == "Zone3" & YR == 1, 
-    # weighted.var2(HWAH, SharePct)]
-    # annwv2[, sqrt(V1)] / solmu[, V1]  # checks
-
     # CV due to soil
     annmu <- crm[, round(mean(HWAH)), by = .(WTH, SOL, Zone, SharePct)]
-    # cr[, .N, by = .(TNAM, SOL)][which(N > 62), ]
-    # crm[WTH == "AAAI" & SOL == "WI_LVUY032" & Zone == "Zone2a", ]
-    # crm[WTH == "AAAI" & SOL == "WI_LVUY032" & Zone == "Zone3", mean(HWAH)]
-    # annmu[WTH == "AAAI" & Zone == "Zone3", ]
-    # crm[WTH == "AAAI" & Zone == "Zone3" & SOL == "WI_LVUY032", mean(HWAH)]
-    # range(crm[, round(mean(HWAH)), by = .(WTH, SOL, Zone)][, V1] - 
-    # crm[, round(mean(HWAH)), by = .(WTH, SOL, Zone, SharePct)][, V1])
-    # crm[WTH == "AAAA" & Zone == "Zone3" & SOL == "WI_CLBW256", mean(HWAH)]
-    # annmu[WTH == "AAAA" & Zone == "Zone3" & SOL == "WI_CLBW256", ]
     annsolmu <- annmu[, weighted.mean(V1, SharePct), by = .(WTH, Zone)]
-    # sum(annsolmu$V1 - annmu[, weighted.mean(V1, SharePct), 
-    # by = .(WTH, Zone, SOL)][, V1])
     solcv <- annmu[, sqrt(weighted.var(V1, w = SharePct, na.rm = TRUE)), 
@@
 # ra
+w <- matrix(ncol = 3, nrow = 3, rep(1, 9))
 yld_rl <- lapply(crops, function(x) {  # x <- crops[[1]]
-  yr <- lapply(c("S", "R", "I"), function(y) {  #y <- "S"
+  yr <- lapply(c("S", "R", "I"), function(y) {  # y <- "S"
     ylds <- yld_rasts[[x]]
@@
     r <- dt_to_raster(ylds[, c("x", "y", cnms), with = FALSE], CRSobj = gcs)
-    w <- matrix(ncol = 3, nrow = 3, rep(1, 9))
     rfoc <- sapply(1:nlayers(r), function(i) rfoc <- yield_fix(r[[i]], w = w))
@@
 
-for(i in 1:length(yld_rl)) {
-  rl <- yld_rl[[i]]
+```
+
+## Short-term soybean fix
+
+Potential soybean yields are too low for Zambia, so we will adjust them up with FAO-GAEZ yields
+
+```{r, eval = FALSE}
+faonms <- dir(p_fao, full.names = TRUE)
+file.rename(faonms, gsub("res02crav6190|yld_package", "", faonms))
+faonms <- dir(p_fao, full.names = TRUE)
+p_faosoy <- full_path(p_fao, c("soybean_I", "soybean_R", "soybean_S"))
+file.rename(faonms, full_path(p_fao, c("soybean_I", "soybean_R", "soybean_S")))
+faonms <- dir(p_faosoy, full.names = TRUE)
+renm <- paste0(p_faosoy, c("soybean_I", "soybean_R", "soybean_S"), ".tif")
+file.rename(faonms[grep("res02_crav6190", faonms)], renm)
+
+# load in tiffs, crop and mask to Zambia
+faosoy <- lapply(renm, function(x) raster(x))
+names(faosoy) <- c("soybean_I", "soybean_R", "soybean_S")
+faosoy <- lapply(faosoy, function(x) {
+  r <- crop(x, yld_rl$soybean$S)
+  ro <- mask(r, yld_rl$soybean$S[[1]])
+})
+faosoy <- faosoy[names(faosoy)[3:1]]  # rearrange
+plot(faosoy[[2]])
+
+# dssat soy mean yield grids
+dssatsoy_mu <- sapply(yld_rl$soybean, function(x) {
+  r <- calc(x[[grep("ymu", names(x))]], mean)
+})
+sapply(dssatsoy_mu, function(x) cellStats(x, mean))
+
+# Find mean yield differences
+soydiff <- lapply(1:3, function(x) {
+  faosoy[[x]] - dssatsoy_mu[[x]]
+})
+soydiffmsk <- lapply(soydiff, function(x) x < 0) # poss and negative
+
+# Liuwa plains and wetland yields much higher than should be. Don't adjust
+
+bias <- sapply(1:3, function(x) {
+  r1 <- mask(soydiff[[x]], soydiffmsk[[x]] == 0, maskvalue = 0)
+  r2 <- mask(soydiff[[x]], soydiffmsk[[x]] == 1, maskvalue = 0)
+  c("mu1" = cellStats(r1, mean), "mu2" = cellStats(r2, mean), 
+    "med1" =  quantile(r1, probs = 0.5), "med2" =  quantile(r2, probs = 0.5))
+})
+
+# calculate bias offset for different management types, filtering out wetland
+# areas. Interpolate between rainfed subsistence and commercial because of 
+# of overly high (3,500 kg/ha) rainfed high input bias
+biasu <- bias[c(1, 3), ]
+biasu[, 2] <- rowMeans(bias[c(1, 3), c(1, 3)]) # replace with interpolated
+biasd <- bias[c(2, 4), ]  # down adjustment for wetlands
+biasd[, 2] <- rowMeans(bias[c(2, 4), c(1, 3)]) # replace with interpolated
+
+# median and mean bias about the same
+# fix dssat yields, test first on just the mean yields
+par(mfrow = c(2, 2))
+soyadjtest <- lapply(1:3, function(x) {  # x = 1
+  r1 <- (dssatsoy_mu[[x]] + biasu[2, x]) * (soydiffmsk[[x]] == 0)
+  r2 <- (dssatsoy_mu[[x]]) * (soydiffmsk[[x]] == 1)
+  r3 <- r1 + r2
+  plot(r3)
+})
+
+# Fix all of them now
+soyfix <- lapply(1:3, function(x) {
+  ri <- yld_rl$soybean[[x]]
+  r1 <- (ri + biasu[2, x]) * (soydiffmsk[[x]] == 0)
+  r2 <- ri * (soydiffmsk[[x]] == 1)
+  r3 <- r1 + r2
+  names(r3) <- names(ri)
+  r3
+})
+names(soyfix) <- names(yld_rl$soybean)
+```
+
+```{r, eval = FALSE}
+yld_rl2 <- list(yld_rl$maize, soyfix)
+names(yld_rl2) <- crops
+
+for(i in 1:length(yld_rl2)) {
+  rl <- yld_rl2[[i]]
   for(j in 1:length(rl)) {
     fnm <- full_path(p_yield, paste0(crops[i], "_", names(rl)[j], "_10K.tif"))
-    writeRaster(rl[[j]], filename = fnm)
+    writeRaster(rl[[j]], filename = fnm, overwrite = TRUE)
+    layernms <- names(rl[[j]])
+    save(layernms, file = gsub("\\.tif", "\\.rda", fnm))
   }
@@
 
-This provides the methods for taking 10 km grids, averaging them across cultivars within management treatments, and then reseampling to 1 km grids.
+This provides the methods for taking 10 km grids, averaging them across cultivars within management treatments, and then resampling to 1 km grids.
 This is approximately the approach that will be used to make inputs for the tradeoff model.

### 2016-01-21 Rmd/to-do.Rmd
Updated to-do

--- 2016-01-19 Rmd/to-do.Rmd
@@
               "Upgrade all maps to SERVIR landcover, and rebase yield estimates on that", "Baccini vegetation carbon + ISRIC grids"),
-  "Doing" = c("D", "D", "U", "U", "D", "D", "D", "L", "D", "D", "F")
+  "Doing" = c("D", "F*", "U", "U", "D", "D", "D", "L", "D", "F**", "F")
 )
@@
 ```
+*Temporary soybean grid with bias adjustment from FAO GAEZ yields. Developed weighted merged method for different types of management.
+
+**Just using total production estimates, which will not be mapped for now. These will be basis for estimating production targets.

### 2016-01-28 Rmd/all-inputs.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

--- 2016-01-20 Rmd/all-inputs.Rmd
@@
     + Impact only:
-        + `ZA-vegremain.tif` - vegetation remaining grid, for assessing conversion impacts. 
+        + `ZA-bd.tif` - vegetation remaining grid, for assessing conversion impacts. 
+        + `intactness.tif` - measure of intactness
+        + `ZA-pas.tif` - protected areas, to see how many forest reserves get hit
     + Masking/allocation
@@
 cropland <- raster(full_path(p_crop, fname = "cropland-1k.tif"))
-bd <- raster(full_path(p_bd, fname = "bdprior.tif"))
-vegr <- raster(full_path(p_bd, fname = "ZA-vegremain.tif"))
-cost <- raster(full_path(p_cost, fname = "zamcost.tif"))
+cons <- raster(full_path(p_bd, fname = "cons-priorities.tif"))
+bd <- brick(full_path(p_bd, fname = "ZA-vegremain.tif"))
+load(full_path(p_bd, fname = "ZA-vegremain.rda"))
+names(bd) <- vegnames
+carbon <- brick(full_path(p_carb, fname = "ZA-carbon.tif"))
+names(carbon) <- c("veg", "soil")
+cost <- raster(full_path(p_cost, fname = "ZA-cost.tif"))
+pas <- raster(full_path(p_bd, "ZA-pas.tif"))
+intact <- raster(full_path(p_bd, "intactness.tif"))
+ynms <- dir(p_yield, pattern = "FAO.tif")  # FAO adjusted now
+ynms <- sapply(c("S", "R", "I"), function(x) ynms[grep(x, ynms)])
+ynms <- ynms[c(1, 3, 5, 2, 4, 6)]
+ylds <- lapply(ynms, function(x) {
+  b <- brick(full_path(p_yield, x))
+  ylnms <- gsub("\\.tif", "\\.rda", x)
+  load(full_path(p_yield, ylnms))
+  names(b) <- layernms
+  b
+})
+names(ylds) <- gsub("_10K|\\.tif|_FAO", "", ynms)
+
+# mask materials
+remfrac <- raster(full_path(p_bd, fname = "ZA-remfrac.tif"))
+water <- raster(full_path(p_h20, fname = "water-1k.tif"))
+farmable <- raster(full_path(p_terr, fname = "farmable-1k.tif"))
 
 ```
+
+## Create farmable area
+
+```{r, eval=FALSE}
+# merge together the three farmland available masks
+avail <- remfrac - water
+avail[avail < 0] <- 0
+avail <- avail - (1 - farmable)
+avail[avail < 0] <- 0
+avail <- avail * (cons < 0.962)  # remove PAs entirely (1 if nps, 0.962 if GMAs)
+names(avail) <- "convertible"
+# plot(avail)
+```
+
+That creates the farmland availability mask that is needed.
+
+## Stack and mask all grids
+```{r, eval = FALSE}
+# deal with yields
+potmaize <- ylds$maize_R  # just taking rainfed commercial potential to start
+w <- matrix(nrow = 5, ncol = 5, rep(1, 5 * 5))  # have to filter b/c artifacts
+potmaizef <- focal(potmaize[[1]], w = w, fun = mean, na.rm = TRUE)
+potsoy <- ylds$soybean_R  # just taking rainfed commercial potential to start
+potsoy[potsoy > 5] <- 5 #round(cellStats(potsoy, mean), 1)  # really high fix
+potsoyf <- focal(potsoy[[1]], w = w, fun = mean, na.rm = TRUE)
+# potsoy <- mask(potsoyf, potsoy)  # crop yield
+
+yield_fix <- function(x = r[[1]], w, na.rm = TRUE) {
+  r <- focal(x, w = w, fun = mean, na.rm = na.rm)
+  rmsk <- (!is.na(r) & is.na(x)) * r
+  x[is.na(x)] <- 0
+  rfix <- x + rmsk  # add back in edge bits
+}
+
+w <- matrix(nrow = 3, ncol = 3, rep(1, 9))
+potsoyf2 <- yield_fix(potsoyf[[1]], w)
+potmaizef2 <- yield_fix(potmaizef[[1]], w)
+plot(potmaizef2)
+# first get rid of all NA area in all grids
+# start with yield grids
+# yldmask <- calc(stack(ylds), sum)
+
+# Function to disaggregate and reproject yield rasters
+yld_disagg <- function(x, msk, fact = 10) {
+  r <- disaggregate(x, fact = 10)
+  r2 <- resample(x, r)
+  r3 <- projectRaster(r2, msk)
+  r4 <- mask(r3, msk)
+  r4
+}
+
+ylds_1k <- lapply(list(potmaizef2, potsoyf2), function(x) {
+  yld_disagg(x, msk = mgrid)
+})
+ylds_s <- stack(ylds_1k) 
+names(ylds_s) <- c("maize", "soy")
+ylds_s <- round(ylds_s, 1)  
+# plot(ylds_s)
+
+# yldmask <- calc(stack(potmaizef2, potsoyf2), sum)
+# yldmask2 <- disaggregate(yldmask, fact = 10)
+# yldmask3 <- resample(yldmask, yldmask2)
+# yldmask4 <- projectRaster(yldmask3, mgrid)
+# yldmask5 <- mask(yldmask4, mgrid)
+# plot(yldmask5)
+
+# now with other
+carbonf <- stack(lapply(1:nlayers(carbon), function(x) {
+  r <- carbon[[x]]
+  r[is.na(r)] <- 0
+  r
+}))
+
+# pas and intactness
+intpa <- stack(intact, pas)
+
+mgridc <- crop(mgrid, avail)
+inlist <- list("convertible" = avail, "bd" = bd, "carbon" = carbonf, 
+               "cost" = cost, "cons" = cons, "cropland" = cropland, 
+               "intpa" = intpa, "p_yield" = ylds_s)
+inlist <- lapply(inlist, function(x) crop(x, mgridc))
+inlist2 <- lapply(inlist, function(x) calc(x, mean))
+inlist3 <- calc(stack(inlist2), mean)
+# inlist4 <- calc(stack(inlist3, crop(ylds_s, mgridc)), mean)
+
+# filter NA values
+# nafilt <- calc(stack(!is.na(inlist3), !is.na(crop(yldmask5, mgridc))), sum)
+nafilt <- !is.na(inlist3)
+# plot(nafilt)
+# plot(!is.na(calc(vegr, mean)))
+# plot(!is.na(crop(yldmask5, mgridc)), col = c("transparent", "red"), add = TRUE)
+msk <- nafilt == maxValue(nafilt)  # freq(nafilt)
+msk[msk == 0] <- NA
+# plot(msk)
+
+# updated mask
+mgridf <- mask(mgridc, msk)
+valinds <- which(!is.na(values(mgridf)))
+outlist <- inlist[names(inlist) != "cropland"]
+
+# write out new data.tables
+### make sure convertible writes out column header name properly
+cnames <- "ZA"  
+onms <- names(outlist)
+onms[onms == "cons"] <- "cons-priorities"
+onms[onms == "p_yield"] <- "potential-yields"
+
+for(i in 1:length(outlist)) { # i <- 1
+  fnm <- full_path(p_production, paste0(cnames, "-", onms[i], ".csv"))
+  DT <- as.data.table(outlist[[i]], xy = FALSE)[valinds, ]
+  write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+}
+
+# write out mask raster
+base <- as.data.table(msk, xy = TRUE)
+setnames(base, old = names(msk), new = "val")
+base[, ind := 1:nrow(base)]
+base <- base[!is.na(val)][, val := NULL]
+fnm <- full_path(p_production, paste0(cnames, "-mask.csv"))
+write.table(base, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+# plot(dt_to_raster(base, mgridf@crs))
+
+# # data.table them
+# cnames <- "ZA"  
+# fnm <- full_path(p_production, paste0(cnames, "-potential-yields.csv"))
+# DT <- as.data.table(ylds_s, xy = FALSE)[valinds, ]
+# # DT[which(is.na(DT))]
+# write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+```
+
+# Copy over files to the correct directory
+```{r, eval}
+p_mod <- full_path(p_proj, "agroEcoTradeoff/external/data/dt/latest/")
+p_master <- "~/Dropbox/projects/agroEcoTradeoff/external/data/dt/latest/"
+fnms <- dir(p_production)
+file.copy(dir(p_production, full.names = TRUE), full_path(p_mod, fnms),
+          overwrite = TRUE)
+file.copy(dir(p_production, full.names = TRUE), full_path(p_master, fnms),
+          overwrite = TRUE)
+
+```
+
+## Yield grids
+```{r, eval = FALSE}
+# # Function to disaggregate and reproject yield rasters
+# yld_disagg <- function(x, msk, fact = 10) {
+#   r <- disaggregate(x, fact = 10)
+#   r2 <- resample(x, r)
+#   r3 <- projectRaster(r2, msk)
+#   r4 <- mask(r3, msk)
+#   r4
+# }
+
+# # Function to aggregate yields within a particular yield brick
+# # stat <- "50"; cults <- paste0("9900", c("08", "09", "10"))
+# yield_stat <- function(x, stat = "mu", mgmnt = "S", cults, fun = mean) {
+#   repstr <- paste0(paste0(stat, "_", mgmnt), cults, collapse = "|")
+#   cnms <- names(x)[grep(repstr, names(x))]
+#   print(cnms)
+#   calc(x[[cnms]], fun = fun)
+# }
+# rm(stat, mgmnt, cults, fun, x, cnms, repstr)
+
+# Let's make an initial yield combination for model development that is
+# is composed of: 
+# maize
+# 1. Subsistence management, all four cultivars, median yields, 50%
+# 2. Rainfed commerical, hybrid cultivars, median yields, 40%
+# 3. Irrigated maize, hybrid cultivars, median yields, 10%
+
+# maize <- list("S" = yield_stat(x = ylds$maize_S.tif, "50", "S", 
+#                                paste0("000", 1:4)), 
+#               "R" = yield_stat(x = ylds$maize_R.tif, "50", "R", 
+#                                paste0("000", 3:4)), 
+#               "I" = yield_stat(x = ylds$maize_I.tif, "50", "I", 
+#                                paste0("000", 3:4)))
+# wgts <- c(0.5, 0.4, 0.1)
+# maizewgt <- lapply(1:length(wgts), function(x) maize[[x]] * wgts[x])
+# potmaize <- calc(stack(maizewgt), sum)  # potential maize yield
+
+# wgts <- c(0.2, 0.7, 0.1)
+# mznms <- names(ylds)[grep("maize", names(ylds))]
+# # maizewgt <- lapply(1:length(wgts), function(x) ylds[[mznms[x]]] * wgts[x])
+# # potmaize <- calc(stack(maizewgt), sum)  # potential maize yield
+# potmaize <- ylds$maize_R  # just taking rainfed commercial potential to start
+# plot(ylds$maize_I)
+
+# Soybean
+# 1. Subsistence management, all 3 cultivars, median yields, 20%
+# 2. Rainfed commerical, all 3cultivars, median yields, 70%
+# 3. Irrigated maize, all 3 cultivars, median yields, 10%
+# soy <- list("S" = yield_stat(x = ylds$soybean_S.tif, "50", "S", 
+#                              paste0("9900", c("08", "09", "10"))), 
+#             "R" = yield_stat(x = ylds$soybean_R.tif, "50", "R", 
+#                              paste0("9900", c("08", "09", "10"))), 
+#             "I" = yield_stat(x = ylds$soybean_I.tif, "50", "I", 
+#                              paste0("9900", c("08", "09", "10"))))
+# wgts <- c(0.2, 0.7, 0.1)
+# soywgt <- lapply(1:length(wgts), function(x) soy[[x]] * wgts[x])
+# potsoy <- calc(stack(soywgt), sum)  # potential maize yield
+# plot(potmaize); cellStats(potmaize, mean)
+# plot(potsoy); cellStats(potsoy, mean)
+# plot(potmaize)
+# plot(potsoy)
+# potsoy <- ylds$soybean_R  # just taking rainfed commercial potential to start
+# potsoy[potsoy > 5] <- round(cellStats(potsoy, mean), 1)
+# w <- matrix(nrow = 5, ncol = 5, rep(1, 5 * 5))  # have to filter b/c artifacts
+# potsoyf <- focal(potsoy[[1]], w = w, fun = mean, na.rm = TRUE)
+# potsoy <- mask(potsoyf, potsoy)  # crop yield
+# 
+# yield_fix <- function(x = r[[1]], w, na.rm = TRUE) {
+#   r <- focal(x, w = w, fun = mean, na.rm = na.rm)
+#   rmsk <- (!is.na(r) & is.na(x)) * r
+#   x[is.na(x)] <- 0
+#   rfix <- x + rmsk  # add back in edge bits
+# }
+# 
+# w <- matrix(3, 3, rep(1, 9))
+# potsoyf2 <- yield_fix(potsoy[[1]], w)
+# potmaizef2 <- yield_fix(potmaize[[1]], w)
+
+# Disaggregate potential yields, stack, and round
+# ylds_1k <- lapply(list(potmaizef2, potsoyf2), function(x) {
+#   yld_disagg(x, msk = mgridf)
+# })
+# ylds_s <- stack(ylds_1k) 
+# names(ylds_s) <- c("maize", "soy")
+# ylds_s <- round(ylds_s, 1)  
+# # plot(ylds_s)
+# 
+# # data.table them
+# cnames <- "ZA"  
+# fnm <- full_path(p_production, paste0(cnames, "-potential-yields.csv"))
+# DT <- as.data.table(ylds_s, xy = FALSE)[valinds, ]
+# # DT[which(is.na(DT))]
+# write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+#
+
+```

### 2016-01-28 Rmd/biodiversity-layer.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

@@
 pas <- mask(pas, mgrid)
+writeRaster(pas, filename = full_path(p_bd, "ZA-pas.tif"))
 # plot(pas)
@@
 
-writeRaster(bdf, file = full_path(p_bd, "bdprior.tif"), overwrite = TRUE)
-vprops <- cbind("var" = c("area", "rarity", "priority"),
-                rbind(vtots / nrow(veg_dt), wgti, data.table(t(wgts_prop_np))))
-save(vprops, file = full_path(p_bd, "vprops.rda"))
+writeRaster(bdf, file = full_path(p_bd, "cons-priorities.tif"), 
+            overwrite = TRUE)
+bdprops <- cbind("var" = c("area", "rarity", "priority"),
+                 rbind(vtots / nrow(veg_dt), data.table(t(wgtsn)), 
+                       data.table(t(wgts_prop_np))))
+names(bdprops) <- gsub("veg", "v", names(bdprops))
+save(bdprops, file = full_path(p_bd, "ZA-bdprops.rda"))
+# load(full_path(p_bd, "ZA-bdprops.rda"))
 
@@
 
-`vprops.rda` - table of veg type area, rarity value, and prioritization value
+`ZA-bdprops.rda` - table of veg type area, rarity value, & prioritization value
 
-`bdprior.tif` - biodiversity prioritization grid
+`cons-priorities.tif` - biodiversity prioritization grid
 

### 2016-01-28 Rmd/cropland.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

@@
 # mean(muprod$soy[, V1] / mupa$soy[, V1])
+# mean(muprod$maize[, V1] / mupa$maize[, V1])
 

### 2016-01-28 Rmd/demo-figures.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

---
title: "WB-IFAD presentation"
author: "Lyndon Estes"
date: "January 26, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes

---

# Initial slides for running model
```{r}
library(agroEcoTradeoff)
setwd("agroEcoTradeoff")
p_model <- set_base_path()
p_main <- "/Users/lestes/Dropbox/publications/zambiaToff/private/zambiaToff/"
p_figures <- full_path(p_main, "paper/figures/presentation/")
p_bd <- full_path(p_main, "external/input_devel/biodiversity")
library(rworldxtra)
data("countriesHigh")
# load other data here

```

## To make

### Meeting demand on existing land versus getting it all from new land

Existing yields are 2.3 t/ha for maize and 2 for soy.  Potential mean yields are 4.43 for maize and 3.4 for soybean, based on 2015-2050 projected changes from Bruinsma.

Scenario: maize demand 2X, soy demand 5X

```{r, eval=FALSE}
path <- "external/data/dt/latest/"
il <- input_handler(input_key = "ZA", ybetas = list(1, 1), ybeta_update = 0, 
                    path = path)

targ <- c("maize" = 3, "soy" = 6)
currprodmod <- c(1, 1)
target <- targets(prod_targ = targ, currprod = il$currprod, 
                  currprodmod = currprodmod)
mzgap <- 4.43 / 2.3
soygap <- 3.4 / 2
# 3224 / 1674; 2124 / 1257

# gap closing test
# y2 <- 7.5  # future yield
# y1 <- 5  # current yield
# ha <- 10  # hectares
# prod <- c(y2, y1) * ha  # future and current production
# (y1 + (2.5 * 0.25)) * ha  
# # future production with 25% of possible yield increase met
# (((prod[2] * y2 / y1) - prod[2]) * 0.25 + prod[2]) / prod[2] * prod[2]

# closing 75% of yield gap
cprodmod_h <- ((((il$currprod * c(mzgap, soygap)) - il$currprod) * 
                  1) + il$currprod) / il$currprod
# assume 75% gap for each
targ_high <- targets(targ, il$currprod, cprodmod_h)
# (1 + ((mzgap - 1) * 0.75) * mzgap) * il$currprod
# (il$currprod - 1 + ((mzgap - 1) * 0.75) * il$currprod)

# closing 25% of yield gap
cprodmod_l <- ((((il$currprod * c(mzgap, soygap)) - il$currprod) * 
                  0.25) + il$currprod) / il$currprod
# assume 75% gap for each
# targ_low <- targets(targ, il$currprod, cprodmod_l)
targ_low <- targets(targ, il$currprod, currprodmod = 1)

# Run model
cbetas <- c("Y" = 0.5, "C" = 0, "BD" = 0, "COST" = 0.5)
thigh <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                      ybetas = list(1, 1), exist_list = il, 
                      currprodmod = cprodmod_h, input_key = "ZA", 
                      silent = FALSE)
tlow <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                     ybetas = list(1, 1), exist_list = il, 
                     #currprodmod = cprodmod_l, input_key = "ZA", 
                     currprodmod = 1, input_key = "ZA", 
                     silent = FALSE)
thigh$impacts
tlow$impacts

CRSobj <- CRS(il$sp$crs)

# rasters
rhigh <- dt_to_raster(thigh$conv, CRSobj)
rhigh <- calc(rhigh * c(1, 2), sum)
rlow <- dt_to_raster(tlow$conv, CRSobj)
rlow <- calc(rlow * c(1, 2), sum)

zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
zambia <- spTransform(zambia, CRSobj)
p_figs <- full_path(p_main, "paper/figures/presentation/")
# p_roads <- full_path(p_main, "external/input_devel/roads/tmdpfr_links.sqlite")
load(full_path(p_model, "external/data/parks_roads.rda"))
# roads <- readOGR(p_roads, layer = "tmdpfr_links")
# pas <- raster(full_path(p_bd, "ZA-pas.tif"))
# pmask <- !is.na(pas)
# pmask[pmask == 0] <- NA

# impact plots
i1 <- tlow$impacts
imp_plot(dtr = rlow, impacts = i1, fnm = "low_closure.png", 
         ftitle = "Current Yield")
i2 <- thigh$impacts
imp_plot(dtr = rhigh, impacts = i2, fnm = "high_closure.png", 
         ftitle = "FAO-Projected Yield")
dr <- (rlow > 0) - (rhigh > 0)
plot(dr) #; freq(dr)
```

### Optimal for each parameter
```{r, eval=FALSE}
cbetas <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
Y <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                  exist_list = il, currprodmod = 1, input_key = "ZA",
                  silent = FALSE)
cbetas <- c("Y" = 0, "C" = 1, "BD" = 0, "COST" = 0)
C <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                  exist_list = il, currprodmod = 1, input_key = "ZA",
                  silent = FALSE)
cbetas <- c("Y" = 0, "C" = 0, "BD" = 1, "COST" = 0)
BD <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                   exist_list = il, currprodmod = 1, input_key = "ZA",
                   silent = FALSE)
consr <- dt_list_to_raster(il$mask, list(il$cons), CRSobj)
cbetas <- c("Y" = 0, "C" = 0, "BD" = 0, "COST" = 1)
COST <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                     exist_list = il, currprodmod = 1, 
                     input_key = "ZA", silent = FALSE)

Yr <- calc(dt_to_raster(Y$conv, CRSobj) * c(1, 2), sum)
Cr <- calc(dt_to_raster(C$conv, CRSobj) * c(1, 2), sum)
BDr <- calc(dt_to_raster(BD$conv, CRSobj) * c(1, 2), sum)
COSTr <- calc(dt_to_raster(COST$conv, CRSobj) * c(1, 2), sum)

cnms <- names(cbetas)
mnms <- paste(c("Yield", "Carbon Protection", "Biodiversity Protection", 
                "Cost Minimization"), "Prioritized")
for(i in 1:length(cnms)) { # i = 1
  print(cnms[i])
  i1 <- get(cnms[i])$impacts
  imp_plot(dtr = get(paste0(cnms[i], "r")), impacts = i1, 
           fnm = paste0(cnms[i], ".png"), ftitle = mnms[i])  
}

```

### Equal Weights
```{r, eval = FALSE}
cbetas <- c("Y" = 0.25, "C" = 0.25, "BD" = 0.25, "COST" = 0.25)
YBCD <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, ybetas = list(1, 1),
                     exist_list = il, currprodmod = 1, input_key = "ZA",
                     silent = FALSE)
YBCDr <- calc(dt_to_raster(YBCD$conv, CRSobj) * c(1, 2), sum)
imp_plot(dtr = YBCDr, impacts = YBCD$impacts, 
         fnm = "YBCD.png", ftitle = "Ag 25%, C 25%, BIOD 25%, COST 25%")

```

### Optimal expansion based on selected weights
```{r, eval = FALSE}
yblist <- list(yb1 <- c(1, 1))
step <- 0.1
# cnames <- c("Y", "C", "BD", "COST"); step = 0.05
cnames <- c("Y", "C", "BD", "COST"); step = 0.2
# cnames <- c("Y", "C", "BD"); step = 0.25#, "BD", "COST"); step = 0.2
pareto_steps(cnames, step)
Yv = "conv_area"; Cv = "tot_C"; BDv = "priority"; COSTv = "mu_cost"
Yst = sum; Cst = sum; BDst = mean; COSTst = mean
a <- Sys.time()
print(a)
pare <- pareto(cnames, step, targ, yblist = list(1, 1),
               input_key = "ZA", Yv, Cv, BDv, COSTv, Yst, Cst, BDst, COSTst,
               todisk = TRUE, silent = FALSE, ncl = 8, path = path)
Sys.time() - a  # 1.75 minutes
pare[, plot(Y, C)]

parms <- batch_params(list(c(1, 1)), list(targ), pareto_steps(cnames, step))

ott <- data.frame(pare)
ott$colors <- rainbow(length(ott$ind))
ott$ind <- seq(1, length(ott$ind))

odfs <- lapply(cnames, function(x) {
  if(x %in% colnames(ott)) {
    odf <- ott[order(ott[, x]), ]
  } else{ 
    odf <- NULL
  }
  odf
})

pvec <- cnames
tvec <- c("by Area", "by C", "by BIOD", "by Cost")
yvec <- c("Area km^2", "C Mt", "BIOD", "Time")
div <- c(1 / 100, 1 / 100000, 100, 1)
dfs <- which(sapply(odfs, is.data.frame))
png(full_path(p_figs, "pareto_weights.png"), height = 700, width = 1000)
if(length(dfs) == 2) {
  par(mfcol = c(2, 2), oma = c(2, 5, 1, 0), mar = c(1, 1, 1, 1) + 0.1, 
      xaxs = "i", mgp = c(1, 0.1, 0), tcl = -0.2)
} else if(length(dfs) == 3) {
  par(mfcol = c(3, 2), oma = c(2, 5, 1, 0), mar = c(1, 1, 1, 1) + 0.1, 
      xaxs = "i", mgp = c(1, 0.1, 0), tcl = -0.2)
} else if(length(dfs) == 4) {
  par(mfcol = c(4, 2), oma = c(2, 5, 1, 0), mar = c(1, 1, 1, 1) + 0.1, 
      xaxs = "i", mgp = c(1, 0.1, 0), tcl = -0.2)
}
for(i in 1:2) {  #i <- 1
  d <- odfs[[i]]
  if(!is.null(d)) {
    for(j in 1:length(pvec)) {  # j <- 1
      if(pvec[[j]] %in% colnames(d)) {
        barplot(round(d[, pvec[j]] * div[j]), names.arg = d$ind, col = d$colors,
                yaxt = "n", las = 2, cex.axis = 0.3)
        if(j == 1) mtext(tvec[i], side = 3)
        if(i == 1) mtext(yvec[j], side = 2, cex = 0.8, line = 3)
        if(i == 1) {
          axis(2, las = 2, cex = 0.6)
        } else {
          axis(2, labels = FALSE)
        }
      }
    } 
  }
}
dev.off()
quantile(0:100, probs = seq(0, 1, 0.1))

# Figure out which meets multiple objective simultaneously
pdeciles <- pare[, lapply(.SD, function(x) {
  quantile(x, probs = seq(0, 1, 0.1))[-1]
  }), .SD = 1:4]
# pdeciles <- cbind("deciles" = paste0(seq(10, 100, 10), "th"), pdeciles)

prange <- pare[, lapply(.SD, range), .SD = 1:4]
pdiff <- prange[, lapply(.SD, function(x) diff(x) * 0.4 + min(x))]
swvec <- as.matrix(pdiff)[1, ]
tmat <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
tmat[tmat == "TRUE"] <- 1

best_ind <- which(rowSums(tmat[, 1:4]) == 4)
parms[best_ind, 1:4]
pare[best_ind, ]

# alternative scenario
pdiff2 <- prange[, lapply(.SD, function(x) diff(x) * 0.2 + min(x))]
pdiff3 <- prange[, lapply(.SD, function(x) diff(x) * 0.6 + min(x))]
pdiff4 <- cbind(pdiff2[, 1:2, with = FALSE], pdiff3[, 3:4, with = FALSE])

swvec <- as.matrix(pdiff4)[1, ]
tmat <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
tmat[tmat == "TRUE"] <- 1
ind1 <- which(rowSums(tmat[, 1:4]) == 4)  # no solution

swvec <- as.matrix(pdiff2)[1, 1:2]
tmat <- sweep(as.matrix(pare)[, 1:2], MARGIN = 2, swvec, `<=`)
tmat[tmat == "TRUE"] <- 1
second_ind <- which(rowSums(tmat[, 1:2]) == 2)

parms[second_ind, 1:4]
pare[best_ind, ]

# sapply(1:nrow(pdeciles) {
#   #swvec <- as.matrix(pdeciles[deciles == "20th", cnames, with = FALSE])[1, ]
#   swvec <- as.matrix(pdeciles[deciles == "20th", cnames, with = FALSE])[1, ]
#   tmat <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
#   parms
# })
# pare[, lapply(.SD, function(x) )]

```

### Create map of those weight combinations
```{r, eval = FALSE}
cbetas <- parms[best_ind, 1:4]
tobest <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                       ybetas = list(1, 1), exist_list = il, 
                       currprodmod = 1, input_key = "ZA", silent = FALSE)
rbest <- calc(dt_to_raster(tobest$conv, CRSobj) * c(1, 2), sum)
ib <- tobest$impacts
imp_plot(dtr = rbest, impacts = ib, fnm = "preferred_scenario.png", 
         ftitle = "Ag 40%, C 0%, BIOD 40%, COST 20%")

cbetas <- parms[second_ind, 1:4]
tosecond <- tradeoff_mod(prod_targ = targ, cbetas = cbetas, 
                         ybetas = list(1, 1), exist_list = il, 
                         currprodmod = 1, input_key = "ZA", 
                         silent = FALSE)
rsecond <- calc(dt_to_raster(tosecond$conv, CRSobj) * c(1, 2), sum)
i2b <- tosecond$impacts
imp_plot(dtr = rsecond, impacts = i2b, fnm = "second_scenario.png", 
         ftitle = "Ag 60%, C 40%, BIOD 0%, COST 0%")

```

### Heat map from all combinations
```{r, eval = FALSE}
bpath <- full_path(p_model, "external/output/batch/dt/")
load(full_path(bpath, "bcode.rda"))
fnames <- dir(full_path(bpath, bcode), pattern = "csv", full.names = TRUE)
DT <- lapply(fnames, function(x) fread(x)[, rowSums(.SD)])
DTbind <- data.table(do.call(cbind, DT))
DTsum <- data.table(DTbind[, rowSums(.SD)])
DTspat <- cbind(il$mask[, .(x, y)], DTsum)
DTb <- dt_to_raster(DTspat, CRSobj)
DTbpct <- DTb / maxValue(DTb) * 100

DTbpct[DTbpct == 0] <- NA

png(full_path(p_figs, "heat.png"), height = 700, width = 700)
ccols <- c("grey90", "orange", "red")
pacols <- c("transparent", rgb(1, 0, 0, alpha = 0.1),
            rgb(0, 0, 1, alpha = 0.1), "transparent")
par(mar = c(1, 1, 1, 3))
plot(zambia, col = "grey90", border = "transparent")
plot(DTbpct, col = rev(heat.colors(20)), add = TRUE)
# plot(dtr, legend = FALSE, col = ccols, add = TRUE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
     border = FALSE)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
     border = FALSE)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
dev.off()

```

### Slide showing an area equal to 200% of need
```{r, eval = FALSE}
cbetas <- parms[best_ind, 1:4]
targ200 <- c("maize" = 6, "soy" = 12)
to200 <- tradeoff_mod(prod_targ = targ200, cbetas = cbetas, 
                      ybetas = list(1, 1), exist_list = il, 
                      currprodmod = 1, input_key = "ZA", silent = FALSE)
r200 <- calc(dt_to_raster(to200$conv, CRSobj) * c(1, 2), sum)
ib <- to200$impacts
imp_plot(dtr = r200, impacts = ib, fnm = "demand2X.png", 
         ftitle = "Maize 6X, Soy 12X")
```

### Our areas versus proposed farm blocks
```{r, eval = FALSE}
# fb <- "~/Dropbox/projects/agroEcoTradeoff/external/base_data/farm-blocks.tif"
# fblocks <- raster(fb)
fbp <- "/Users/lestes/Dropbox/data/zari/misc/NEW4GEO/Fblocks.sqlite"
fb <- readOGR(fbp, layer = "fblocks")
fbalb <- spTransform(fb, CRSobj)
fbalb$ID <- 1:nrow(fbalb)
fbalb@data <- fbalb@data[, "ID", drop = FALSE]
fnm <- full_path(p_main, "external/input_devel/other/fblocks_alb.sqlite")
writeOGR(fbalb, dsn = fnm, layer = "fblocks_alb", driver = "SQLite")
# First rasterize farm blocks
# require(gdalUtils)
# onm <- full_path(p_main, "external/input_devel/other/fblocks_alb.tif")
# gdal_rasterize(src_datasource = fnm, l = "fblocks_alb", ot = "Byte", 
#                a = "ID", tr = res(r200), at = TRUE, co = "COMPRESS=LZW",
#                dst_filename = fnm, te = bbox(r200)[1:4], of = "GTiff", 
#                verbose = TRUE)
fbr <- rasterize(fbalb, r200)
fbr[is.na(fbr)] <- 0
fbr[fbr > 0] <- 1
fbr <- mask(fbr, r200)

r200sum <- r200 > 0
fbint <- (r200sum == 1 & fbr == 1) * 3 + (r200sum == 1 & fbr == 0) * 2 +
  (r200sum == 0 & fbr == 1) * 1
png(full_path(p_figs, "fblock-overlap.png"), height = 700, width = 700)
ccols <- c("grey90", "tan", "green3", "red")
pacols <- c("transparent", rgb(1, 0, 0, alpha = 0.1),
            rgb(0, 0, 1, alpha = 0.1), "transparent")
par(mar = c(1, 1, 1, 3))
plot(zambia, col = "grey90", border = "transparent")
plot(fbint, col = ccols, add = TRUE, legend = FALSE)
# plot(dtr, legend = FALSE, col = ccols, add = TRUE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
     border = FALSE)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
     border = FALSE)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", 
       legend = c("Farm blocks", "Converted", "Converted in farm blocks", 
                  "NatParks", "GMAs"),
       pch = 15, col = c(ccols[2:4], pacols[2:3]), bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()

```

### Slide showing clustering

### Slide showing clustering along proposed road improvement areas

### Remaining Natural Vegetation Types

```{r, eval = FALSE}
p_bd <- full_path(p_main, "external/input_devel/biodiversity")
remfrac <- raster(full_path(p_bd, "ZA-remfrac.tif"))
png(full_path(p_figs, "remveg.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 1))
plot(remfrac, axes = FALSE, box = FALSE)
dev.off()

```

### Background Slides
```{r, eval = FALSE}
brasts <- dt_list_to_raster(il$mask, 
                            il[c("p_yield", "carbon", "cons", "cost")], 
                            CRSobj)
brasts <- lapply(brasts, function(x) x[[-c(1:2)]])

# Yields
png(full_path(p_figs, "yields.png"), height = 350, width = 700)
par(mfrow = c(1, 2), mar = c(0, 0, 0, 3), oma = c(0, 0, 0, 2))
for(i in 1:2) {
  plot(zambia, col = "grey90", border = "transparent")
  plot(brasts$p_yield[[i]], add = TRUE)
  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
  if(i == 2) {
    legend("bottomright", legend = c("NatParks", "GMAs"),
           pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
           cex = 1.5)
  }
}
dev.off()

# Carbon loss
closs <- brasts$carbon[[1]] + brasts$carbon[[1]] * 0.25
png(full_path(p_figs, "pot-C-loss.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 3))
plot(closs, axes = FALSE, box = FALSE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", legend = c("NatParks", "GMAs"),
       pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()

# Cost
png(full_path(p_figs, "cost.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 3))
plot(brasts$cost, axes = FALSE, box = FALSE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", legend = c("NatParks", "GMAs"),
       pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()

# Biodiversity
png(full_path(p_figs, "biodiversity.png"), height = 700, width = 700)
par(mar = c(1, 1, 1, 3))
plot(brasts$cons, axes = FALSE, box = FALSE)
plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2], border = F)
plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3], border = F)
plot(roads, add = TRUE, lwd = 0.1, col = "grey")
legend("bottomright", legend = c("NatParks", "GMAs"),
       pch = 15, col = pacols[2:3], bty = "n", pt.cex = 3,
       cex = 1.5)
dev.off()

```

### 2016-01-28 Rmd/maize-soy-grid.Rmd
Bunch of stuff, up through creation of figures for presentation, in demo-figures.Rmd

@@
 # read in different grids, start with Maize
-crops <- list.dirs(p_crop, full.names = FALSE)[-1]
+crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
 cv_rasts <- lapply(crops, function(X) {
-  # X <- crops[1]
-  fnms <- dir(full_path(p_crop, X))
+  # X <- crops[2]
+  fnms <- dir(full_path(p_crop, X), pattern = X)
   fl <- unlist(lapply(c("S", "R", "I"), function(x) {
@@
   idt <- copy(IDr)
-  for(y in 1:length(fl)) { # y <- 2
+
+  # idt[, zoneid := rep(0, nrow(idt))]
+  # for(i in 1:length(unique(idt$Zone))) {
+  #  idt[Zone == unique(idt$Zone)[i], zoneid := i]
+  # }
+
+  for(y in 1:length(fl)) { # y <- 3
     print(paste("Processing", fl[y]))
@@
 probs <- c(0.1, 0.5, 0.9)  # quantiles to get
-crops <- list.dirs(p_crop, full.names = FALSE)[-1]  # dir names
+crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
 yld_rasts <- lapply(crops, function(X) {
-  # X <- crops[1]
-  fnms <- dir(full_path(p_crop, X))
+  # X <- crops[2]
+  fnms <- dir(full_path(p_crop, X), pattern = X)
   fl <- unlist(lapply(c("S", "R", "I"), function(x) {
@@
 })
+yld_rasts <- lapply(crops, function(x) {
+  fnm <- full_path(p_root,
+                   paste0("external/data/", x, "_yields_10km.csv"))
+  fread(fnm)
+})
+names(yld_rasts) <- crops
 
@@
     cnms <- nms[grep(paste0("y.*._", y), nms)]
-    r <- dt_to_raster(ylds[, c("x", "y", cnms), with = FALSE], CRSobj = gcs)
+    r <- dt_to_raster(ylds[, c("x", "y", cnms), with = FALSE], 
+                      CRSobj = CRS(gcs))
     rfoc <- sapply(1:nlayers(r), function(i) rfoc <- yield_fix(r[[i]], w = w))
@@
 
+# cellStats(yld_rl$maize$R[["ymu_R0002"]], mean)
+# cellStats(yld_rl$soybean$I[["ymu_I990008"]], mean)
+plot(yld_rl$soybean$S[["ymu_S990008"]])
+plot(yld_rl$soybean$R[["ymu_R990008"]])
+plot(yld_rl$maize$R[["y50_R0003"]])
 ```
 
-## Short-term soybean fix
+## Yield bias correction
+### Soybean fix based on GAEZ data
 
-Potential soybean yields are too low for Zambia, so we will adjust them up with FAO-GAEZ yields
+Potential soybean yields are too low for Zambia, so we will adjust them up with FAO-GAEZ yields.
+This doesn't seem necessary for maize, because our potential yield averages are at or below the GYGA rainfed yield potentials, per the figures and tables below.
 
 ```{r, eval = FALSE} 
-faonms <- dir(p_fao, full.names = TRUE)
-file.rename(faonms, gsub("res02crav6190|yld_package", "", faonms))
-faonms <- dir(p_fao, full.names = TRUE)
+cellStats(yld_rl$maize$S[["ymu_S0001"]], mean)
+plot(yld_rl$maize$R[["ymu_R0002"]])
+plot(yld_rl$maize$S[["ymu_S0002"]])
+
+```
+
+This is one approach for bias-correction.
+
+```{r, eval = FALSE}
+# faonms <- dir(p_fao, full.names = TRUE)
+# file.rename(faonms, gsub("res02crav6190|yld_package", "", faonms))
+# faonms <- dir(p_fao, full.names = TRUE)
 p_faosoy <- full_path(p_fao, c("soybean_I", "soybean_R", "soybean_S"))
@@
 names(soyfix) <- names(yld_rl$soybean)
+
 ```
 
+Write that out. 
 ```{r, eval = FALSE}
@@
 <a href="#top">Back to top</a>
+
+But now do the second fix.  Let's adjust yield relative to the actual and baseline and projected growth.
+
+FAO projected growth to 2000-2050 is 69.4% for maize and 219% for soybean. Current mean yields are from FAO 2010-2014 data and USDA
+
+```{r, eval = FALSE}
+load(full_path(p_cropland, "currprod.rda"))
+load(full_path(p_cropland, "usda-yields.rda"))
+load(full_path(p_cropland, "fao-yields.rda"))
+cols <- names(usda)[4:9]
+usda_ylds <- usda[var == "prod", cols, with = FALSE] / 
+  usda[var == "harvarea", cols, with = FALSE]
+usda_mu <- round(rowMeans(usda_ylds), 1)
+
+muprod <- lapply(c("Maize", "Soybeans"), function(x) {
+  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
+      by = Year]
+})
+names(muprod) <- c("maize", "soy")
+mupa <- lapply(c("Maize", "Soybeans"), function(x) {
+  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
+      by = Year]
+})
+names(mupa) <- c("maize", "soy")
+fao_mu <- c("maize" = mean(muprod$maize[, V1] / mupa$maize[, V1]), 
+            "soy" = mean(muprod$soy[, V1] / mupa$soy[, V1]))
+
+# current yields, mean of USDA and FAO
+curr_mu <- round(colMeans(rbind(usda_mu, fao_mu)), 1)  # 2.3 maize, 2.0 soy
+
+# Calculate the mean yields for the country in each management scenario (across 
+# cultivar types)
+ydiffs <- lapply(crops, function(x) {
+  yr <- lapply(c("S", "R", "I"), function(y) {
+    r <- yld_rl[[x]][[y]]
+    nms <- grep("mu", names(r))
+    rmu <- cellStats(calc(r[[nms]], mean), mean)
+    rvar <- r / rmu
+    names(rvar) <- names(r)
+    list("ratio" = rvar, "mu" = rmu)
+  })
+  names(yr) <- c("S", "R", "I")
+  yr
+})
+names(ydiffs) <- crops
+
+# calculate the mean ratio
+yrats <- lapply(1:2, function(x) { # x <- 1
+  yr <- lapply(c("S", "R", "I"), function(y) {  # y <- 1
+     # inner loop if needed
+     X <- crops[x]
+     nms <- grep("mu", names(ydiffs[[X]][[y]]$ratio))
+     r <- ydiffs[[X]][[y]]$ratio[[nms]]
+     rmu <- calc(r, mean)
+     rmu 
+  })
+  names(yr) <- c("S", "R", "I")
+  yr
+})
+names(yrats) <- crops
+
+# Take the DSSAT means and figure out how much they differ relative to their 
+# overall mean across management types - use as management specific ratio
+dssat_mus <- sapply(ydiffs, function(x) sapply(x, function(y) y$mu))
+dssat_adj <- sweep(dssat_mus, MARGIN = 2,  colMeans(dssat_mus), FUN = `/`)
+
+# Bruinsma 2007-2050 growth
+bruinsma <- c(maize = 3224 / 1674, soy = 2124 / 1257)
+
+# adjusted DSSAT factors
+potmus <- curr_mu * c(bruinsma)  # 4.43 maize, 3.4 soy
+potmu_ds <- sweep(dssat_adj, MARGIN = 2,  potmus, FUN = `*`)
+
+# w <- matrix(ncol = 5, nrow = 5, rep(1, 25))
+# plot(focal(yrats$maize$S, w), fun = mean, na.rm = TRUE)
+# plot(yrats$maize$S * potmu_ds["S", "maize"])
+# plot(yrats$maize$R * potmu_ds["R", "maize"])
+# plot(yrats$soybean$S * potmu_ds["S", "soybean"])
+
+# finally create adjust yield rasters
+ynew <- lapply(crops, function(x) { # x <- 1
+  yr <- lapply(c("S", "R", "I"), function(y) {  # y <- 1
+     r <- yrats[[x]][[y]] * potmu_ds[y, x]
+  })
+  names(yr) <- c("S", "R", "I")
+  yr
+})
+names(ynew) <- crops
+
+# writout
+for(i in 1:length(ynew)) {
+  rl <- ynew[[i]]
+  for(j in 1:length(rl)) {
+    fnm <- full_path(p_yield, 
+                     paste0(crops[i], "_", names(rl)[j], "_10K_FAO.tif"))
+    writeRaster(rl[[j]], filename = fnm, overwrite = TRUE)
+    layernms <- names(rl[[j]])
+    save(layernms, file = gsub("\\.tif", "\\.rda", fnm))
+  }
+}
+
+```
 

### 2016-02-08 Rmd/demo-figures.Rmd
Many changes, including development of yield-predictors, empirical-yields (GAM fit to DSSAT yields), paper to-do list, etc.

@@
 ```{r, eval=FALSE}
-path <- "external/data/dt/latest/"
-il <- input_handler(input_key = "ZA", ybetas = list(1, 1), ybeta_update = 0, 
-                    path = path)
+setwd("agroEcoTradeoff")
+path <- full_path(set_base_path(), "external/data/ZA")
+code <- run_code("ZA")
+il <- input_handler(path, ybetas = list(1, 1), ybeta_update = 0, code = code)
 
@@
 # p_roads <- full_path(p_main, "external/input_devel/roads/tmdpfr_links.sqlite")
-load(full_path(p_model, "external/data/parks_roads.rda"))
+load(full_path(p_model, "external/data/ZA/parks_roads.rda"))
+save(pas, roads, zambia, 
+     file = full_path(p_model, "external/data/ZA/parks_roads.rda"))
 # roads <- readOGR(p_roads, layer = "tmdpfr_links")

### 2016-02-08 Rmd/empirical-yields.Rmd
Many changes, including development of yield-predictors, empirical-yields (GAM fit to DSSAT yields), paper to-do list, etc.

---
title: "Stage 2 Yield Prediction"
author: "Lyndon Estes"
date: "February 5, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Empirical prediction of yield

A yield transfer function for DSSAT yields

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(matrixStats)
library(doMC)
library(mgcv)

p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")

# Zambia provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Regression variables
load(fp(p_yield, "empirical/regdat.rda"))

```

# Fit model
## Exploratory plots
```{r, eval = FALSE}
plot(prov)
regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
regdt[CROP == "maize" & YLDS > 8000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
regdt[CROP == "maize" & YLDS > 8000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
regdt[CROP == "soybean" & YLDS > 1700, 
      points(x, y, pch = 16, cex = 0.5, col = "blue")]

# Prune out yields > 8000 maize > 1700 soybean in subsistence case
mzfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
                .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
soyfilt <- regdt[CROP == "maize" & YLDS > 8000, unique(.SD), 
                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
# regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]

# regnew[CROP == "maize", hist(YLDS)]
# regnew[CROP == "maize", hist(YLDR)]

# summary(lm(YLDS ~ poly(MAP, 2), data = regnew[CROP == "maize", ]))
# summary(lm(YLDR ~ poly(MSP, 2), data = regnew[CROP == "maize", ]))
regnew[CROP == "maize", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(FP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sloc, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sloc, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slcl, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slcl, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slsi, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slsi, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sbdm, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sbdm, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slhw, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(slhw, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(GDD, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(KDD, YLDS, pch = 16, cex = 0.5)]
# regnew[CROP == "maize", plot(log(KDD), YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKJ, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKF, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ATMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRM, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRM, YLDS, pch = 16, cex = 0.5)]

regnew[, round(cor(.SD), 2), .SDcols = names(regnew)[-c(1:13)]]

summary(lm(YLDS ~ poly(sbdm, 2) + poly(slhw, 2), 
           data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(slhw, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(sbdm, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(KDD, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDR ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ KDD, data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(GDD, 2), data = regnew[CROP == "maize", ]))

# p <- "sbdm"; d <- "YLDS"
# regnew[CROP == "maize", {
#   lm1 <- lm(get(d) ~ get(p))
#   lm2 <- lm(get(d) ~ poly(get(p), 2))
#   plot(get(p), get(d), pch = 16, cex = 0.1)
#   abline(lm1)
#   dum <- data.frame(floor(min(get(p))):ceiling(max(get(p))))
#   colnames(dum) <- p
#   lines(dum[, 1], predict(lm2, newdata = dum), col = "red")
#   summary(lm1)
#   # summary(lm2)
# }]

reg_plotter <- function(dep, pred) {
  lm1 <- lm(dep ~ pred)
  lm2 <- lm(dep ~ poly(pred, 2))
  plot(pred, dep, pch = 16, cex = 0.1)
  abline(lm1)
  prng <- c(floor(min(pred)), ceiling(max(pred)))
  dum <- data.frame(seq(prng[1], prng[2], by = diff(prng) / 100))
  # dum <- data.frame(floor(min(pred)):ceiling(max(pred)))
  colnames(dum) <- "pred"
  preds <- predict(lm2, newdata = dum)
  lines(dum[, 1], preds, col = "red")
  mtext(round(summary(lm1)$adj.r.squared, 4), side = 3, adj = 0)
  mtext(round(summary(lm2)$adj.r.squared, 4), side = 3, adj = 0.5)
}
# regnew[CROP == "maize", reg_plotter(YLDR, GDD)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.045
regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.012
# regnew[CROP == "maize", reg_plotter(YLDS, JP - mean(JP))]  # P 0.012
regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.005
regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.003

regnew[CROP == "maize", reg_plotter(YLDS, slhw)]  # P 0.17
regnew[CROP == "maize", reg_plotter(YLDS, sloc)]  # P 0.266
# regnew[CROP == "maize", reg_plotter(YLDR, sloc)]  # P 
regnew[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.185
regnew[CROP == "maize", reg_plotter(YLDS, sbdm)]  # 0.181
# mtax pairings
regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # P 0.07
regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.056 
# NK
regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.054
regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.027 
# GDD/KDD
regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.04
regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.027
# srad
regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.012 
regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # Same, picked up wrong one
regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.0288 
regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.0299

# best predictors look to be
# MAP, SLOC (but polynomial relationship odd), SLHW, slcl, sbdm, all good
# STMX (linear KDD as a possible alternate), GDD, ASRM

#
regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4279

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4199

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(sloc, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4044 # SLOC doesn't give as much value, NS on linear

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4293  # Doesn't add much

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4293  # Oct precip add a tiny bit, significant

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4312  # Jan precip adds more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4374  # Feb precip even more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

# regnew[CROP == "maize", {
#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
#               poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
#              family = gaussian(link = "log"), method = "ML")
#   summary(gam1)
#   # plot(YLDS, predict(lm1))
# }]  # 0.4301  # Oct precip add a tiny bit, significant

```

## Reduce variables and fit models

Reduce down to means across years, and select 10% and 90% percentile years

### Maize
```{r, eval = FALSE}
regred <- copy(regnew[CROP == "maize"])
regred[, mid2 := 1:.N]

# regred[, c("Y10S", "Y90S") := as.integer(0)]

regmu <- regnew[CROP == "maize", lapply(.SD, mean), 
                by = list(CROP, WTH, Zone, SOL), 
                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
regmu[, YST := "mu"]
# regnew[CROP == "maize" & Zone == "Zone2a" & WTH == "AABS" & 
#          SOL == "WI_CMZR003", mean(YLDS)]
qs <- function(x, y, ptile) {
  DF <- cbind(1:length(x), x, y)
  DF <- DF[order(DF[, 2]), ]
  DF[round(nrow(DF) * ptile), 3]
}

y <- 30:60
DF <- cbind(1:31, sample(2000:3000, size = 31), y)
DF <- DF[order(DF[, 2]), ]
DF[round(nrow(DF) * 0.9), 3]
sort(DF[, 2])

a <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
            SOL == "WI_RGTZ054", YLDS]
b <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
            SOL == "WI_RGTZ054", mid]
qs(a, b, 0.9)
sort(a)

# Select out quantile values
m10 <- regred[,
  qs(YLDS, mid2, 0.1), by = list(CROP, WTH, Zone, SOL)
][, V1]
m90 <- regred[,
  qs(YLDS, mid2, 0.9), by = list(CROP, WTH, Zone, SOL) 
][, V1]

# 
y10 <- regred[mid2 %in% m10]
y90 <- regred[mid2 %in% m90]
y10[, YST := "10"]
y90[, YST := "90"]

y1090 <- rbind(y10, y90)
cnames <- colnames(y1090)[colnames(y1090) %in% colnames(regmu)]
setcolorder(regmu, cnames)
regred2 <- rbind(regmu, y1090[, cnames, with = FALSE])

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regred2[CROP == "maize" & YST == i, hist(YLDS)]
}
regred2[CROP == "maize", mean(YLDS), by = YST]

regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.071
regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.020
regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.017
regred2[CROP == "maize",  summary(lm(YLDS ~ poly(FP, 2)))]
regred2[CROP == "maize", reg_plotter(YLDS, 100 - slcl + slsi)]  # P 0.102
regred2[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.1879
regred2[CROP == "maize", reg_plotter(YLDS, slsi)]  # P 0.03

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize" & YST == "mu", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(sloc, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  gam1 <- gam(YLDS ~ s(MAP) + s(JP) + s(slhw) + s(slcl) + (sbdm) + s(GDD) +
                KDD + s(ASRM), 
            family = gaussian(link = "log"), method = "ML")
}]

gam1 <- gam(YLDS ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
              s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
              s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam2 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam3 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2) + s(x, y), 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              s(x, y), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam6 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam7 <- gam(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam8 <- gam(YLDS ~ poly(MAP, 2) + slhw +  poly(slcl, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam9 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
              poly(100 - (slcl + slsi), 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam10 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(sloc, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam11 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               sloc + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam12 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               sloc + poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])

summary(gam3)
plot.gam(gam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam4)
plot.gam(gam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam5)
plot.gam(gam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam6)
plot.gam(gam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam7)
plot.gam(gam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam8)
plot.gam(gam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam9)
plot.gam(gam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam10)
plot.gam(gam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam11)  # This one or
plot.gam(gam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(gam12)  # of this one, with smooth x, y terms
plot.gam(gam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

par(mfrow = c(2, 2))
plot(regred2[CROP == "maize", YLDS], predict(gam11, type = "response"))
plot(regred2[CROP == "maize", YLDS], predict(gam12, type = "response"))
plot(regred2[CROP == "maize", YLDS], predict(gam4, type = "response"))

# Reduce down to model for mean, 10th, 90th)
gam12mu <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[YST == "mu"])
summary(gam12mu)
plot.gam(gam12mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
gam1210 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[YST == "10"])
summary(gam1210)
plot.gam(gam1210, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
gam1290 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[YST == "90"])
summary(gam1290)
plot.gam(gam1290, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

```

### Soybean

```{r, eval = FALSE}
regreds <- copy(regnew[CROP == "soybean"])
regreds[, mid2 := 1:.N]

# regred[, c("Y10S", "Y90S") := as.integer(0)]

regmus <- regreds[, lapply(.SD, mean), by = list(CROP, WTH, Zone, SOL), 
                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
regmus[, YST := "mu"]

# Select out quantile values
m10 <- regreds[,
  qs(YLDR, mid2, 0.1), by = list(CROP, WTH, Zone, SOL)
][, V1]
m90 <- regreds[,
  qs(YLDR, mid2, 0.9), by = list(CROP, WTH, Zone, SOL) 
][, V1]

# Select out 10th and 90th percentile years
y10s <- regreds[mid2 %in% m10]
y90s <- regreds[mid2 %in% m90]
y10s[, YST := "10"]
y90s[, YST := "90"]

y1090s <- rbind(y10s, y90s)
cnames <- colnames(y1090s)[colnames(y1090s) %in% colnames(regmus)]
setcolorder(regmus, cnames)
regreds2 <- rbind(regmus, y1090s[, cnames, with = FALSE])
regreds2[, slcf := 100 - (slcl + slsi)]

par(mfrow = c(2, 2))
for(i in c("10", "mu", "90")) {
  regreds2[CROP == "soybean" & YST == i, hist(YLDR)]
}
regreds2[CROP == "soybean", mean(YLDR), by = YST]

regreds2[, reg_plotter(YLDR, MAP)]  # L 0.065
regreds2[, reg_plotter(YLDR, MSP)]  # L 0.1105
regreds2[, reg_plotter(YLDR, JP)]  # L 0.08
regreds2[, reg_plotter(YLDR, FP)]  # P 0.016
regreds2[, reg_plotter(YLDR, 100 - slcl + slsi)]  # P 0.12
regreds2[, reg_plotter(YLDR, slcl)]  # P 0.31
regreds2[, reg_plotter(YLDR, slsi)]  # P 0.003
regreds2[, reg_plotter(YLDR, sloc)]  # P 0.31
regreds2[, reg_plotter(YLDR, slhw)]  # P 0.27
regreds2[, reg_plotter(YLDR, GDD)]  # P 0.006
regreds2[, reg_plotter(YLDR, KDD)]  # P 0.006
regreds2[, reg_plotter(YLDR, STMX)]  # L 0.029
regreds2[, reg_plotter(YLDR, NKF)]  # L 0.032
regreds2[, reg_plotter(YLDR, NKJ)]  # P 0.029
regreds2[, reg_plotter(YLDR, SSRT)]  # L 0.023
regreds2[, reg_plotter(YLDR, ASRM)]  # L 0.055

sgam1 <- gam(YLDR ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
               s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
               s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regreds2)

sgam2 <- gam(YLDR ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam3 <- gam(YLDR ~ poly(MAP, 2) + poly(FP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam4 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam5 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(sloc, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam6 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + sloc + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam7 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
              poly(slcl, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regreds2)
sgam8 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + poly(ASRM, 2), 
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(ASRM, 2), 
             family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam10 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(ASRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam11 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(sbdm, 2) + 
               poly(GDD, 2) + poly(ASRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam12 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam13 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + sloc +  
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam14 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam15 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + KDD + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam16 <- gam(YLDR ~ poly(MSP, 2) + poly(OP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam17 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam18 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + NKJ + poly(SSRM, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
sgam19 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
               poly(GDD, 2) + NKF + poly(SSRM, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)

summary(sgam1)
plot.gam(sgam1, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam2)
plot.gam(sgam2, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam3)
plot.gam(sgam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam4)
plot.gam(sgam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam5)
plot.gam(sgam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam6)
plot.gam(sgam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam7)
plot.gam(sgam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam8)
plot.gam(sgam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam9)  # MSP stronger than MAP
plot.gam(sgam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam10)  # MSP stronger than MAP
plot.gam(sgam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam11)  # SBDM much weaker SLCL
plot.gam(sgam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam12)  # SSRM a tad better than 
plot.gam(sgam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam13)  # SLOC goes in the wrong direction
plot.gam(sgam13, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam14)  # OP adds a bit
plot.gam(sgam14, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam15)  # KDD NS
plot.gam(sgam15, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam16)  # Polynomial term on October precip doesn't add much
plot.gam(sgam16, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam17)  # xy term added to model 15, all remain signif, seems way to go
plot.gam(sgam17, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam18)  # NKJ NS
plot.gam(sgam18, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
summary(sgam19)  # NKF NS
plot.gam(sgam19, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

# Now see about taking same model and fitting separately to 10th/90th/mean years
sgam17mu <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
                  poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "mu"])
summary(sgam17mu)  # OP drops out, SSRM drops out
plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
sgam17mu <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + 
                  poly(slcl, 2) + poly(GDD, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "mu"])
summary(sgam17mu)  # OP drops out, SSRM drops out
plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
sgam1710 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
                + poly(GDD, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "10"])
summary(sgam1710)  # OP drops out, SSRM drops out
plot.gam(sgam1710, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
sgam1790 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
                + poly(GDD, 2) + s(x, y), 
                family = gaussian(link = "log"), method = "REML", 
                data = regreds2[YST == "90"])
summary(sgam1790)  # OP drops out, SSRM drops out
plot.gam(sgam1790, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)

```

### Save GAMs outputs
```{r, eval = FALSE}
save(gam12, gam12mu, gam1210, gam1290, sgam17, sgam17mu, sgam1710, sgam1790, 
     file = fp(p_yield, "gams.rda"))

```

### 2016-02-08 Rmd/maize-soy-grid.Rmd
Many changes, including development of yield-predictors, empirical-yields (GAM fit to DSSAT yields), paper to-do list, etc.

@@
 library(agroEcoTradeoff)
-# library(data.table)
-library(lmisc)
-# library(dtraster)
-# devtools::install_github("PrincetonUniversity/lmisc")
 
@@
 # cellStats(yld_rl$soybean$I[["ymu_I990008"]], mean)
-plot(yld_rl$soybean$S[["ymu_S990008"]])
-plot(yld_rl$soybean$R[["ymu_R990008"]])
-plot(yld_rl$maize$R[["y50_R0003"]])
 ```
@@
 ## Yield bias correction
-### Soybean fix based on GAEZ data
+### Approach 1: soybean fix based on GAEZ data
 
@@
 ```{r, eval = FALSE} 
-cellStats(yld_rl$maize$S[["ymu_S0001"]], mean)
-plot(yld_rl$maize$R[["ymu_R0002"]])
-plot(yld_rl$maize$S[["ymu_S0002"]])
+# cellStats(yld_rl$maize$S[["ymu_S0001"]], mean)
+p_scratch <- fp(p_root, "paper/figures/scratch")
+for(i in crops) {  # i <- "maize"
+  for(j in c("S", "R", "I")) {  # j <- "S"
+    yr <- yld_rl[[i]][[j]]
+    yrmu <- calc(yr[[grep("mu", names(yr))]], mean)
+    png(fp(p_scratch, paste0(i, "_", j, "_mu.png")), width = 700, height = 700)
+    par(mar = c(0, 0, 2, 0))
+    plot(yrmu, axes = FALSE, box = FALSE)
+    dev.off()
+  }
+}
+# plot(yld_rl$soybean$R[["ymu_R990008"]])
+# plot(yld_rl$maize$R[["y50_R0003"]])
+# plot(yld_rl$maize$R[["ymu_R0002"]])
+# plot(yld_rl$maize$S[["ymu_S0002"]])
 
@@
 <a href="#top">Back to top</a>
+
+### Approach 2: Bias correct against FAO Stats
 
@@
 
+### Approach 3: Bias-correct against MODIS NPP
+```{r, eval = FALSE}
+# MODIS NPP
+p_npp <- "~/Dropbox/data/npp/mod17a3/"
+npp <- lapply(dir(p_npp, full.names = TRUE), function(x) {
+  r <- crop(raster(x), yld_rl$maize$S)
+  named_out(r, paste0("NPP", gsub("MOD.*.NPP_|.tif", "", basename(x))))
+})
+names(npp) <- sapply(npp, function(x) names(x))
+
+# Zambia provinces
+p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
+                 "rda_dump/Final Shape files RDA/")
+prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")
+prov@proj4string <- npp[[1]]@crs
+prov$ID <- 1:nrow(prov)
+
+# load in cropland 1 km layer
+cropland <- raster(fp(p_cropland, "cropland-1k.tif"))
+croplandgcs <- projectRaster(cropland, npp[[1]])
+croplandmsk <- croplandgcs > 0  # select pixels having any cropland
+provr <- rasterize(prov, croplandmsk, field = "ID")
+
+# select out those pixels from NPP
+nppcrops <- lapply(npp, function(x) croplandmsk * x * 0.1)
+
+s <- stack(provr, croplandgcs, stack(npp))
+npp_dt <- as.data.table(s)
+setnames(npp_dt, c("layer", "cropland.1k"), c("prov", "cland"))
+npp_dt <- npp_dt[!is.na(cland) | !is.na(prov), ]
+npp_dt <- npp_dt[!is.na(prov), ]
+npp_dt <- npp_dt[!is.na(cland), ]
+# plot(dt_to_raster(npp_dt[, .(x, y, cland)], prov@proj4string))
+
+# calculate weighted means by province
+npp_mu <- npp_dt[, lapply(.SD, function(x) {
+  sum(x * 0.1 * cland, na.rm = TRUE) / sum(cland)
+}), by = prov, .SDcols = c("cland", names(npp))]
+npp_dt[, sum(cland), by = prov]
+
+npp_mu2 <- npp_dt[, lapply(.SD, function(x) {
+  weighted.mean(x * 0.1, w = cland, na.rm = TRUE)
+}), by = prov, .SDcols = c("cland", names(npp))]  # slightly diff. rounding?
+
+npp_mu <- merge(npp_mu, prov@data[, c("PROVINCE", "ID")], by.x = "prov", 
+                by.y = "ID")
+
+# FAO country-stat data, which have problems, it seems (see checks below)
+prod <- fao[Crop %in% c("Maize", "Soybeans") & var == "prod", ]
+setnames(prod, "Value", "prod")
+setkeyv(prod, c("prov", "Year", "Crop"))
+prod[, c("var", "provcode", "cropcode", "Flag") := NULL]
+pa <- fao[Crop %in% c("Maize", "Soybeans")  & var == "pa", ]
+pa[Crop == "Soybeans", hist(pa)]
+setnames(pa, "Value", "pa")
+setkeyv(pa, c("prov", "Year", "Crop"))
+pa[, c("var", "provcode", "cropcode", "Flag") := NULL]
+
+# prodpa <- prod[pa]
+# prodpa[, yld := prod / pa]
+# prodpa[Crop == "Soybeans" & Year %in% 2007:2009 & prov == "Western"]
+# fao[Year == 2011 & Crop == "Soybeans"]
+
+# plot(npp_mu2$NPP2008, npp_mu$NPP2008)
+# prodpa[Crop == "Maize" & Year == 2011, ]
+# prodpa[Crop == "Maize" & prov == "Southern", 
+#        plot(Year, yld, type = "l", ylim = c(0, 6))]
+# cols <- rainbow(9)
+# for(i in 1:9) {
+#   prodpa[Crop == "Maize" & prov == unique(prodpa$prov)[i], 
+#          lines(Year, yld, col = cols[i])]
+# }
+# prodpa[Crop == "Soybeans" & prov == "Southern", 
+#        plot(Year, yld, type = "l", ylim = c(0, 6))]
+# for(i in 1:9) {
+#   prodpa[Crop == "Soybeans" & prov == unique(prodpa$prov)[i], 
+#          lines(Year, yld, col = cols[i])]
+# }
+
+# Actually forget the above, too unreliable. Use the PHS data I found from ZARI, # cross-checked against FAO country level production and harvested area data 
+# (to make PHS and CFS)
+p_phs <- "~/Dropbox/data/agricultural/zambia/phs/"
+phs <- data.table(readxl::read_excel(fp(p_phs, "phs-1986-2014.xlsx")))
+setnames(phs, "YEAR", "Year")
+
+# reshape it
+phs_rs <- rbindlist(lapply(names(phs)[-c(1:3)], function(x) {
+  DT <- phs[, c("Year", "crop", x), with = FALSE]
+  setnames(DT, x, "yld")
+  DT[, prov := x]
+}))
+setkeyv(phs_rs, c("Year", "prov", "crop"))
+# phs_rs[crop == "maize", mean(yld, na.rm = TRUE), by = Year][, plot(Year,V1)]
+# plot(phs_rs$Year[phs_rs$crop == "maize"], phs_rs$yld[phs_rs$crop == "maize"])
+
+npp_murs <- rbindlist(lapply(names(npp), function(x) { 
+  DT <- npp_mu[, c("PROVINCE", x), with = FALSE]
+  DT[, Year := gsub("NPP", "", x)]
+  setnames(DT, c("PROVINCE", x), c("prov", "NPP"))
+  DT
+}))
+npp_murs[, Year := as.numeric(Year)]
+setkeyv(npp_murs, c("Year", "prov"))
+
+# prodmg <- prodpa[npp_murs]
+tlm <- lm((a = 1:10) ~ (b = 1:10))
+plot(tlm$model[, 2], tlm$model[, 1])
+str(tlm)
+prodmg <- phs_rs[npp_murs]
+mznpp <- prodmg[crop == "maize", lapply(.SD, mean), .SDcols = c("yld", "NPP"),
+                by = prov]
+prodmg[crop == "soybean", lapply(.SD, mean), .SDcols = c("yld", "NPP"),
+       by = prov][, summary(lm(yld ~ NPP))]
+prodmg[crop == "soybean", lapply(.SD, mean), .SDcols = c("yld", "NPP"),
+       by = prov][, plot(NPP, yld)]
+
+prodmg[crop == "maize", lapply(.SD, mean), .SDcols = c("yld", "NPP"),
+                by = prov][,  {
+                  plot(yld ~ NPP)
+                  npplm = lm(yld ~ NPP)
+                  abline(npplm)
+                  summary(npplm)
+                }]
+```
+
+The relationship between mean annual NPP over croplands and maize, both by province, is:
+
+```
+Call:
+lm(formula = yld ~ NPP)
+
+Residuals:
+     Min       1Q   Median       3Q      Max 
+-0.85083 -0.10149  0.05117  0.26377  0.40294
+
+Coefficients:
+             Estimate Std. Error t value Pr(>|t|)  
+(Intercept) -2.369097   1.531678  -1.547   0.1658  
+NPP          0.006685   0.002326   2.874   0.0238 *
+---
+Signif. codes:  
+0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
+
+Residual standard error: 0.4277 on 7 degrees of freedom
+Multiple R-squared:  0.5413,	Adjusted R-squared:  0.4758 
+F-statistic: 8.262 on 1 and 7 DF,  p-value: 0.02384
+```
+Soy on the other hand has no clear relationship, probably because it is a minor crop and covers only a small portion of the area.
+
+Now to correct DSSAT yields using these NPP
+```{r, eval=FALSE}
+p_bd <- fp(p_root, "external/input_devel/biodiversity")
+remfrac <- raster(full_path(p_bd, fname = "ZA-remfrac.tif"))
+
+# calculate mean NPP of relatively undisturbed pixels in neighborhoods 
+remmsk <- remfrac > 0.9  # areas having greater > 90% intact
+
+w <- matrix(nrow = 5, ncol = 5, rep(1, 25))
+remmsk[remmsk == 0] <- NA
+
+nppmu <- calc(stack(npp) * 0.1, mean)
+nppalb <- projectRaster(nppmu, remmsk)
+nppalb <- raster::mask(nppalb, remmsk)
+
+# moving window
+npp_fmu <- focal(nppalb, w, fun = mean, na.rm = TRUE)
+npp_fmx <- focal(nppalb, w, fun = max, na.rm = TRUE)
+plot(npp_fmx)
+plot(npp_fmu)
+
+# NPP to Z-score
+npp_z <- (npp_fmu - cellStats(npp_fmu, mean)) / cellStats(npp_fmu, sd)
+
+dr <- disaggregate(yld_rl[["maize"]][["R"]]$ymu_R0002, fact = 10)
+dralb <- projectRaster(dr, npp_fmu)
+par(mfrow = c(1, 2), mar = rep(0, 4))
+plot(npp_z); plot(dralb)
+plot(dralb, npp_fmu)
+
+plot(dr2alb)
+
+dr_z <- (dralb - cellStats(dralb, mean)) / cellStats(dralb, sd)
+plot(npp_z)
+
+dr2 <- disaggregate(yld_rl[["soybean"]][["R"]]$ymu_R990008, fact = 10)
+dr2alb <- projectRaster(dr2, npp_fmu)
+par(mar = rep(0, 4))
+plot(dr2alb, npp_fmu)
+
+ydiffs <- lapply(crops, function(x) {
+  yr <- lapply(c("S", "R", "I"), function(y) {
+    r <- yld_rl[[x]][[y]]
+    nms <- grep("mu", names(r))
+    rmu <- cellStats(calc(r[[nms]], mean), mean)
+    rvar <- r / rmu
+    names(rvar) <- names(r)
+    list("ratio" = rvar, "mu" = rmu)
+  })
+  names(yr) <- c("S", "R", "I")
+  yr
+})
+names(ydiffs) <- crops
+
+plot(fpp_mu)
+
+```
+
 ## How to make 1 km grids/data.tables

### 2016-02-08 Rmd/paper-todo.Rmd
Many changes, including development of yield-predictors, empirical-yields (GAM fit to DSSAT yields), paper to-do list, etc.

---
title: "Paper To-Do"
author: "Lyndon Estes"
date: "January 29, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# To-Do List for Paper

A summary of data and model upgrades and analyses to-do.

# Data fixes remaining

## Crops

+ Near-term: See if we can bias-correct using NPP.  
+ Longer-term: DSSAT yields -> machine learning (Nate)

## Biodiversity:

+ Near-term: Code in impact based on intactness + protected weight
+ Longer-term: Code in Newbold model (Laura and Tobias)

# Analyses and timeline

To figure out the analysis, we need to know what questions we are asking first:

1. Is there a way for Zambia to meet a potentially much larger agricultural development target but for substantially much lower environmental cost than conventional development approaches?

    + Supporting analysis: 
        + Run model over multiple permutations of weights for just cost and yield benefit.  
            i. Probably separately for each crop. 
            ii. Counting tonnage of overlap for both crops could be way of further adjusting demand projections. i.e. You figure that if an area shows up as allocated to each crop, you need to increase production targets by twice that amount, because 1) the crops will have to be grown in rotation and 2) that means you are only getting that amount of crop out of that land once in every two (or more, maybe 3, depending on typical rotation) year.  
            iii. This approach avoids problem of optimal allocation
        + Do the same but for just carbon and biodiversity objectives.  
        + Assess overlaps between both sets of heat maps. 
            i. This will give insight into the degree to which agricultural and conservation goals intersect.
                a. We can also look at how far apart goals are within each two subsets. How much range is there in results for the different combinations of benefit and cost for agriculture (this is analagous to considering variations in prices and how they affect profitability)? How much do carbon and biodiversity conservation objectives agree? 
            ii. How far away does the overlap between both objectives leave us from where we need to be in hitting the target, if we only developed those areas?             iii. Then we ask what we ask how much do we have to enlarge the area to hit the targets, and what is the universe of weight combinations that allow you to achieve to achieve 50% of the production target, 75%, 95%, when sum weight is being put on at least one of the agricultural (yield + cost) and one of the conservation measures.

2. What will the costs be under the selected universe of weights that allow you to achieve the 75%, 95%, etc. of production targets?

   + Ranges of costs in terms of lost carbon, biodiversity, and production/earnings.

Analysis/fix | Drop-dead date | Date Finished
------------- | -------------  | -------------
NPP-crop correct  | 29 Jan |
Biodiversity fix  | 31 Jan |
Crop overlaps | ??? |
Yield variability |  ??? |
Combined optimization, back out weights? |  ???

### 2016-02-08 Rmd/yield-predictors.rmd
Many changes, including development of yield-predictors, empirical-yields (GAM fit to DSSAT yields), paper to-do list, etc.

---
title: "Yield Mapping"
author: "Lyndon Estes"
date: "January 31, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Predict DSSAT yields onto finer soil map

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(rCropMod)
library(matrixStats)
library(doMC)

p_root <- lmisc::set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_soil <- "~/DSSAT45/SOIL/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"

# Read in Zambia grid
mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))

# Zambia provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Read in original soil-climate data from rCropMod
soilgrid <- raster(fp(p_cropmod, "af_soil_grid.tif"))  # soil grid
soilshare <- fread(fp(p_cropmod, "af_soil_share.csv"))  # share of profiles

# Read in soil-climate grid from Di's work
ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
ID[, V1 := NULL]
setkeyv(ID, c("WTH", "SOL", "Zone"))
ID[, SharePct := as.numeric(SharePct)]
zone <- sort(unique(ID$Zone))
for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
setkeyv(ID, c("WTH", "Zone"))
# ID[, FID := 1:.N]
IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
setkeyv(IDr, c("WTH", "Zone"))
# ID[SOL == "WI_ARBW438"]
gcs <- "+proj=longlat +datum=WGS84"
idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
# plot(idgr)

# Yield data selection points
# GSOD points
gsod <- fread(fp(p_clim, "gsod.txt"), stringsAsFactors = FALSE)
setnames(gsod, names(gsod), c("nms", "x", "y", "n"))
gsod <- gsod[, 1:3, with = FALSE]
gsod[, nms := paste0("g", nms)]
# plot(idgr)
# points(zamgsod)  # not representative enough

# ZMD rainfall stations
nms <- c('CHIPAT01', 'CHIPEP01', 'CHOMA001', 'ISOKA001', 'KABOMP01', 'KABWE001',
         'KAFIRO01', 'KAFUE001', 'KALABO01', 'KAOMA001', 'KASAMA01', 'KASEMP01',          'KAWAMB01', 'LIVING01', 'LUNDAZ01', 'LUSAKA01', 'LUSAKA02', 'LUSITU01',          'MAGOYE01', 'MANSA001', 'MBALA001', 'MFUWE001', 'MISAMF01', 'MKUSHI01',          'MONGU002', 'MPIKA001') 
x <- c(32.58, 28.38, 27.07, 32.63, 24.20, 28.47, 28.12, 27.92, 22.70, 24.80,
       31.13, 25.85, 29.08, 25.82, 33.20, 28.32, 28.45, 29.37, 27.60, 28.85, 
       31.33, 31.93, 31.25, 29.33, 23.15, 31.43)  
y <- c(-13.55, -17.33, -16.83, -10.12, -13.60, -14.45, -12.60, -15.77, -14.95,
       -14.80, -10.22, -13.53, -9.80, -17.82, -12.28, -15.42, -15.32, -16.30,
       -16.00, -11.10, -8.85, -13.27, -10.10, -14.00, -15.25, -11.90)
zmd <- cbind.data.frame(nms = as.character(nms), x, y, 
                        stringsAsFactors = FALSE)
write.csv(zmd, file = fp(p_climo, "zmd-stations.csv"))

# bring together GSOD and ZMD stations
zmd_gsod <- rbind(zmd, gsod)
coordinates(zmd_gsod) <- c("x", "y")

# Select those intersecting Zambia
gint <- rgeos::gIntersects(prov, zmd_gsod, byid = TRUE)
sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
zmet <- zmd_gsod[sel, ]
# str(zmet)

# buffer the ZMD stations by ~20 km to see if any GSOD stations fall within that
zbuffer <- rgeos::gBuffer(zmet[1:26, ], byid = TRUE, width = 0.3)
zbuffer <- spChFIDs(zbuffer, as.character(1:nrow(zbuffer)))
gint <- rgeos::gIntersects(zbuffer, zmet[27:nrow(zmet), ], byid = TRUE)
sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
desel <- c(27:nrow(zmet))[sel]
# plot(zmet[-desel, ], pch = 16, cex = 0.5)
# points(zmet[desel, ], pch = 16, cex = 0.5, col = "red")
# plot(prov, add = TRUE)
zmetf <- zmet[-desel, ]

zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)

zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
setkeyv(zmet_dt, c("x", "y"))
setkeyv(IDr, c("x", "y"))
IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
setkeyv(IDr_ss, c("WTH", "Zone"))
# IDr_ss[, plot(x, y)]
# points(zmet_dt, pch = 16, cex = 0.5)
# tst <- extract(idgr, zmet_dt)
# all(tst[order(tst[, 1]), 2] == zmetsamps[order(zmetsamps[, 4]), 5])

# alternative for yield data selection points:
# Creatre stratified, representative random sample of 50 pts from ID/zone grid
# ztab <- freq(idgr$zoneid)[-5, ]
# ztabs <- round((ztab[, 2] / sum(ztab[, 2])) * 50)
# set.seed(34)
# gsamps <- do.call(rbind, lapply(1:4, function(i) {
#   gsamp <- sampleRandom(idgr$zoneid == i, size = ztabs[i], cells = TRUE)
#   sampxy <- xyFromCell(idgr, gsamp[, "cell"])
# }))
# plot(idgr$ID)  # unique(values(idgr$ID))
# length(which(!is.na(values(idgr$ID))))
# length(which(!is.na(unique(values(idgr$ID)))))
# points(gsamps, pch = 16, cex = 0.5)
# 
# # merge those points with the main ID grid
# gsamps_dt <- data.table(gsamps)
# setkeyv(gsamps_dt, c("x", "y"))
# setkeyv(IDr, c("x", "y"))
# IDr_ss <- IDr[gsamps_dt]  # pick up IDs, soils, wth variables
# setkeyv(IDr_ss, c("WTH", "SOL", "Zone"))
# IDr_ss[, plot(x, y)]
# points(gsamps, pch = 16, cex = 0.5)

# sgrid <- fread(fp(p_cropmod, "Soil_clim_match_ID.csv"), 
#                stringsAsFactors = FALSE)
# setnames(sgrid, old = c("V1", "X", "Y", "WeatherGrid", "SoilType"), 
#          new = c("ID", "x", "y", "WTH", "SOL"))
# wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
# sr <- dt_to_raster(sgrid[, list(ID, x, y)], CRSobj = wgs)
# setkeyv(sgrid, c("x", "y"))
# sgrid[SOL == "WI_ARBW438"]
```

# Full DSSAT yield grids
```{r, eval = FALSE}
crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
fylds <- lapply(crops, function(X) {  # X <- "maize"
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(c("S", "R"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # idt <- copy(IDr)
  # key(idt)
  idt <- copy(ID)
  idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
  setkeyv(idt, c("WTH", "Zone"))
  for(y in 1:length(fl)) { # y <- 1
    # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("WTH", "Zone"))

    crm <- merge(cr, idt, by = c("WTH", "Zone"))

    # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]

    # Soil averaged yields
    solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
                 by = .(WTH, Zone, YR)]
    snm <- gsub("*.*_|\\.csv", "", fl[y])

    # Yield means
    v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
    setnames(v, "V1", paste0("Y", snm))
    idt <- idt[v]
    setkeyv(idt, c("WTH", "Zone"))
    idt
  }
  idt
})
names(fylds) <- crops

rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
mzs <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
mzr <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$soybean[, c(1:2, grep("YS", names(fylds$soybean))), with = FALSE]
soys <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$soybean[, c(1:2, grep("YR", names(fylds$soybean))), with = FALSE]
soyr <- dt_to_raster(rdt, raster()@crs)

# par(mar = rep(0, 4))
# plot(calc(mzs, mean))
# plot(zmetf, pch = 16, cex = 0.5, add = TRUE)

```

# Extract subset of DSSAT yields, soils, climate data
## Yield
```{r, eval=FALSE}
# merge function for lists of data.tables
mymerge <- function(x, y) merge(x, y, all = TRUE)

# Yield data for those points
mannms <- c("S", "R")
ylds <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(mannms, function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # fl <- fl[1:3]
  crml <- lapply(1:length(fl), function(y) { # y <- 1
    idt <- copy(IDr_ss)
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("WTH", "Zone", "SOL"))

    # merge IDs with subset location
    # crm <- cr[idt]
    crm <- merge(cr, idt)
    # crm[, SharePct := NULL]
    nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
    setcolorder(crm, nms)
    mid <- 1:nrow(crm)
    crm <- cbind(mid, crm)
    keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
    setkeyv(crm, keys)
    # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
    newnm <- gsub(".*._||\\.csv", "", fl[y])
    setnames(crm, "HWAH", newnm)
    crm
  })
  yldmerge <- Reduce(mymerge, crml)
  ynms <- names(yldmerge)
  # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
    # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
  # }))))
  # calculate 
  for(i in mannms) {
    onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
    yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
    yldmerge[, c(onms[[2]]) := NULL]
  }

  # Internal check on yield means
  if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
  if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
  for(ii in sample(1:nrow(yldmerge), size = 20)) {
    for(jj in 1:length(mannms)) {  # jj = 2
      a <- yldmerge[ii, ]
      b <- sapply(crind[[jj]], function(x) {
        crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
      })
      if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
        print(paste("Yield mismatch at", ii, mannms[jj]))
      }
    }
  }

  # swap in averaged yields
  # yldmerge[, mid := NULL]
  # yldmerge <- cbind(yldmerge, yldred)
  yldmerge
})
names(ylds) <- crops

```

## Soil

Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
pts <- IDr_ss[, .(x, y)]
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_dssat_prof(solfile = full_path(p_soil, "WI.SOL"), 
                           profiles = profs$SoilProfile) 
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

# max depth in profile
mdepth <- rowMaxs(as.matrix(profdat[, .SD, .SDcols = grep("slb", lnms)]), 
                  na.rm = TRUE)

# use depths as weights to calculate mean values of soil parameters to 1 m
# but need to fill to stop at 1 m, and also correct for soils that don't get to 
# 1 m.  Assume values for deepest horizon where deepest < 1 m are the values to 
# 1 m. Truncate horizon that spans 1 m at 1 m, makes this deepest horizon in 
# calc. Horizon depths become weights for weighted mean. 
tst <- profdat[, .SD , .SDcols = grep("slb", lnms)]
dtest <- function(x) {
  o <- x
  o[is.na(o)] <- -10000
  if(all(o < 100)) {
    o[which.max(o)] <- 100
  } else if(any(o > 100)) {
    o[which(o > 100)[1]] <- 100
  }
  o[o < 0] <- NA
  o
}
depvec <- t(apply(as.matrix(tst), 1, dtest))  # fix depths

infill <- function(x, ind) {
  o <- x
  o0 <- which(o < 0)
  if(length(o0) > 0) {
    for(i in 1:length(o0)) {
      if(o0[i] < max(ind) & o0[i] > min(ind)) {
        o[o0[i]] <- mean(o[o0[i] - 1], o[o0[i] + 1])
      } else if(o0[i] == max(ind)) {
        o[o0[i]] <- o[o0[i] - 1]
      } else if(o0[i] == min(ind)) {
        o[o0[i]] <- o[o0[i] + 1]
      }
    }
  }  
  o
}

# calculate mean values to 1 m depth  
solvarsr <- solvars[solvars %in% 
                     c("sbdm", "sloc", "slcl", "slsi", "slhw")]
solvals <- do.call(cbind, lapply(solvarsr, function(i) {  # i <- annvars[2]
  mat <- as.matrix(profdat[, .SD , .SDcols = grep(i, lnms)])
  oval <- sapply(1:nrow(mat), function(j) {  # j <- 6
    ind <- which(!is.na(depvec[j, ]))
    matv <- mat[j, ind]
    matv <- infill(x = matv, ind)
    weighted.mean(matv, depvec[j, ind], na.rm = TRUE)
  })
  oval
}))
colnames(solvals) <- solvarsr

# finalize
sol1m <- cbind("pro" = profdat[, pro], data.table(solvals))
```

## Climate
### Indices
```{r, eval = FALSE}
sdate <- "1948/01/01"
edate <- "2010/12/31"
dts <- seq.Date(as.Date(sdate), as.Date(edate), "days")
yrs <- as.numeric(substr(dts, 1, 4))
ano <- unique(yrs)
mos <- as.numeric(substr(dts, 6, 7))
dys <- as.numeric(substr(dts, 9, 10))

# variabes to derive: MAP, N raindays, rainfall depth, n rain days, 
# n dry days in Jan and Feb, mean max temp, N days > 30, annual mean temp, 
# GDD > 30, average daily SRAD
# set up indices for selecting dates
ystrt <- which(mos == 7 & dys == 15) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 7 & dys == 14)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
janind <- rep(0, length(dts))  # january of each year
for(i in sindu) janind[which(sind == i & mos == 1)] <- i
febind <- rep(0, length(dts))  # feb of each year
for(i in sindu) febind[which(sind == i & mos == 2)] <- i
ndjfind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) ndjfind[which(sind == i & mos %in% c(1, 2, 11, 12))] <- i
octind <- rep(0, length(dts))  # feb of each year
for(i in sindu) octind[which(sind == i & mos == 10)] <- i

sindu2 <- sindu[sindu %in% which(ano %in% 1979:2010)]

```

### Rainfall
```{r, eval=FALSE}
prec <- brick(fp(p_pgf, "prec.Zambia.nc"))

# functions for raindays
nrday <- function(x) length(which(x > 0))  # raindays
ndry <- function(x) length(which(x == 0))
sdepth <- function(x) mean(x[x > 0])  # storm depth

# storm depth in NDJF
# sdeps <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(ndjfind == x)]]
#   r <- calc(ryr, sdepth)
#   names(r) <- as.character(ano[x])
#   r
# })
# sdeps <- stack(sdeps)
# 
# # storm depth in NDJF
# rdays <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(djfind == x)]]
#   r <- calc(ryr, nrday)
#   names(r) <- as.character(ano[x])
#   r
# })
# rdays <- stack(rdays)
# 
# # dry days in January
# drydaysj <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(janind == x)]]
#   r <- calc(ryr, ndry)
#   names(r) <- as.character(ano[x])
#   r
# })
# drydaysj <- stack(drydaysj)
# 
# # dry days in February
# drydaysf <- lapply(sindu2, function(x) { # x <- sindu2[1]
#   print(x)
#   ryr <- prec[[which(febind == x)]]
#   r <- calc(ryr, ndry)
#   names(r) <- as.character(ano[x])
#   r
# })
# drydaysf <- stack(drydaysf)

registerDoMC(7)

# total season rainfall
rftot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(rftot, filename = fp(p_clim, "rftot.tif"), overwrite = TRUE)

# total season rainfall
ndjftot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(ndjftot, filename = fp(p_clim, "ndjftot.tif"), overwrite = TRUE)

# january rainfall
jantot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(jantot, filename = fp(p_clim, "jantot.tif"), overwrite = TRUE)

# february total
febtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(febtot, filename = fp(p_clim, "febtot.tif"), overwrite = TRUE)

# february total
octtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(octind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(octtot, filename = fp(p_clim, "octtot.tif"), overwrite = TRUE)

# par(mfrow = c(6, 6), mar = c(0, 0, 0, 0))
# for(i in 1:nlayers(octtot)) {
#   plot(octtot[[i]], axes = FALSE, box = FALSE)
#   plot(prov, add = TRUE, zlim = c(0, 1000))
# }
# which(1980:2010 %in% ) 
# pre2005 <- calc(octtot[[1:26]], mean)
# post2005 <- calc(octtot[[27:31]], mean)
# 
# par(mfrow = c(1, 2), mar = c(0, 0, 0, 0))
# plot(pre2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
# plot(prov, add = TRUE)
# plot(post2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
# plot(prov, add = TRUE)
# 
# par(mfrow = c(1, 1), mar = c(0, 0, 0, 0))
# plot((pre2005 - post2005) / pre2005 * 100, zlim = c(-100, 100))
# plot(prov, add = TRUE)

# plot(rftot[[1]] - ndjftot[[1]])
# plot(ndjftot[[1]]); plot(rftot[[1]])

# plot out to show artifacts at 1 degree boundaries
# p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
#                  "rda_dump/Final Shape files RDA/")
# prov <- readOGR(fp(p_prov, "prov.shp"), layer = "prov")
# plot(rftot[[1]])
# plot(calc((ryr > 0), sum))
# plot(sdeps[[1]])
# plot(rdays[[23]])
```

### Temperature
```{r, eval = FALSE}
tmax <- brick(fp(p_pgf, "Tmax.Zambia.nc")) 
tmin <- brick(fp(p_pgf, "Tmin.Zambia.nc"))

# Create tmean variable
tmu <- foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  tmu <- (tmax[[ind]] + tmin[[ind]]) / 2 - 273.15
}

# Functions
# GDD between 8-30 degrees
gdd830 <- function(x) {
  a <- x[x < 30] - 8
  sum(ifelse(a < 0, 0, a))
}
# x <- sample(4:40, size = 100, replace = T)
# gdd830(x)

# killing degree days
kdd <- function(x) {
  a <- x - 30
  sum(ifelse(a < 0, 0, a))
}
# plot(calc(tmu[[1]], kdd))

# frequency of days with T > 30
ftmax30 <- function(x) length(x[x > 30])

# growing degree day and killing degree days
kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
writeRaster(kdd, filename = fp(p_clim, "kdd.tif"))
gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
writeRaster(gdd, filename = fp(p_clim, "gdd.tif"))

# tmax greater than 30, Jan and Feb
nkilljan <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkilljan, filename = fp(p_clim, "nkilljan.tif"))
nkillfeb <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkillfeb, filename = fp(p_clim, "nkillfeb.tif"))

# mean tmax in NDJF
tmax_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_gs, filename = fp(p_clim, "tmax_gs.tif"))

# mean tmax whole year
tmax_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_ano, filename = fp(p_clim, "tmax_ano.tif"))

```

### Solar radiation
```{r, eval = FALSE}
srad <- brick(fp(p_pgf, "Rs.Zambia.nc"))

# convert to MJ m^2 
mjday <- function(x) x / 1000000 * (3600 * 24)

# total and mean solar radiation for the growing season
srtot_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, sum)
})
writeRaster(srtot_gs, filename = fp(p_clim, "srtot_gs.tif"))
srmu_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_gs, filename = fp(p_clim, "srmu_gs.tif"))

# total and mean gs for whole year
srtot_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, sum)
})
writeRaster(srtot_ano, filename = fp(p_clim, "srtot_ano.tif"))
srmu_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  mj <- mjday(srad[[ind]])
  calc(mj, mean)
})
writeRaster(srmu_ano, filename = fp(p_clim, "srmu_ano.tif"))
```

## Assemble datasets

1. Predictors: 
    + Weather/climate: 
        +`rftot`, `ndjftot`, `jantot`, `febtot` for rainfall; `kdays`, `gdd`, `f30jan`, `f30feb`, `tmax_gs`, `tmax_ano` for temperature; `srtot_gs`, `srmu_gs`, `srtot_ano`, and `srmu_ano` for solar radiation
    + Soil: mean soil properties within top 1 m
    + Management variables: Codes S, R, I
2. Predictands: mean yields across cultivars within each management type

```{r, eval = FALSE}
# Organize rainfall
climnms <- c("rftot", "ndjftot", "jantot", "febtot", "octtot", "kdd", "gdd", 
             "nkilljan", "nkillfeb", "tmax_ano", "tmax_gs", "srtot_ano",
             "srtot_mu", "srtot_gs", "srmu_gs")
bnms <- sapply(climnms, function(x) dir(p_clim, pattern = x, full.names = TRUE))

blist <- lapply(bnms, function(x) {
  b <- brick(x)
  names(b) <- paste0("Y", 1:nlayers(b))
  b
})

onms <- c("MAP", "MSP", "JP", "FP", "OP", "KDD", "GDD", "NKJ", "NKF", "ATMX",
          "STMX", "SSRT", "SSRM", "ASRT", "ASRM")
climl <- lapply(1:length(blist), function(x) {
  r <- blist[[x]]
  names(r) <- paste0(onms[x], 1:nlayers(r))
  r
})
names(climl) <- onms

# rfnms <- c("MAP", "MSP", "JP", "FP", "OP")
# rfl <- list(rftot, ndjftot, jantot, febtot, octtot)
# rfl <- lapply(1:length(rfl), function(x) {
#   r <- rfl[[x]]
#   names(r) <- paste0(rfnms[x], 1:nlayers(r))
#   r
# })
# names(rfl) <- rfnms 
# 
# # Organize temp
# tnms <- c("KDD", "GDD", "NKJ", "NKF", "ATMX", "STMX")
# tfl <- list(kdd, gdd, nkilljan, nkillfeb, tmax_ano, tmax_gs)
# tfl <- lapply(1:length(tfl), function(x) {
#   r <- tfl[[x]]
#   names(r) <- paste0(tnms[x], 1:nlayers(r))
#   r
# })
# names(tfl) <- tnms 
# 
# # Organize srad
# snms <- c("SSRT", "SSRM", "ASRT", "ASRM")
# sfl <- list(srtot_gs, srmu_gs, srtot_ano, srmu_ano)
# sfl <- lapply(1:length(sfl), function(x) {
#   r <- sfl[[x]]
#   names(r) <- paste0(snms[x], 1:nlayers(r))
#   r
# })
# names(sfl) <- snms

# melt predictors
xy <- unique(ylds$maize[, .(x, y)])
# xy[, ind := 1:.N]
cols <- c("x", "y")
xy[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]

# climate predictors
climvals <- lapply(1:length(climl), function(i) extract(x = climl[[i]], xy))
climxy <- lapply(climvals, function(x) {  # x <- rvals[[1]]
  rmelt <- melt(cbind(xy, x), id = 1:2)
  setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
  rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
  setnames(rmelt, "variable", "YR")
  rmelt[, ind := 1:.N]
  setkeyv(rmelt, c("ind", "x", "y", "YR"))
  rmelt
})
climpreds <- Reduce(mymerge, climxy)
climpreds[, YR := as.integer(YR)]
setkeyv(climpreds, c("x", "y", "YR"))

# # rainfall
# rvals <- lapply(1:length(rfl), function(i) extract(x = rfl[[i]], xy))
# rvxy <- lapply(rvals, function(x) {  # x <- rvals[[1]]
#   rmelt <- melt(cbind(xy, x), id = 1:2)
#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
#   setnames(rmelt, "variable", "YR")
#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
#   rmelt[, ind := 1:.N]
#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
#   rmelt
# })
# rfpreds <- Reduce(mymerge, rvxy)
# 
# # temperature
# tvals <- lapply(1:length(tfl), function(i) extract(x = tfl[[i]], xy))
# tvxy <- lapply(tvals, function(x) {  # x <- tvals[[3]]
#   rmelt <- melt(cbind(xy, x), id = 1:2)
#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
#   setnames(rmelt, "variable", "YR")
#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
#   rmelt[, ind := 1:.N]
#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
#   rmelt
# })
# tfpreds <- Reduce(mymerge, tvxy)
# 
# 
# # srad
# svals <- lapply(1:length(sfl), function(i) extract(x = sfl[[i]], xy))
# svxy <- lapply(svals, function(x) {  # x <- rvals[[1]]
#   rmelt <- melt(cbind(xy, x), id = 1:2)
#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
#   setnames(rmelt, "variable", "YR")
#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
#   rmelt[, ind := 1:.N]
#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
#   rmelt
# })
# sfpreds <- Reduce(mymerge, svxy)

# Join all climate predictors
# climpreds <- Reduce(mymerge, list(rfpreds, tfpreds, sfpreds))
# climpreds[, YR := as.integer(YR)]
# setkeyv(climpreds, c("x", "y", "YR"))

setnames(sol1m, "pro", "SOL")
setkey(sol1m, "SOL")
sols1mu <- unique(sol1m)

# merge together all the data
regdat <- lapply(ylds, function(i) {  # i <- ylds[[1]]
  i[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]
  setkeyv(i, c("x", "y", "SOL"))
  m1 <- merge(i, sols1mu)
  m1[, YR := as.integer(YR)]
  setkeyv(m1, c("x", "y", "YR"))
  m2 <- merge(m1, climpreds)
})

# Checking merges
set.seed(234)
samp <- sample(1:nrow(regdat$maize), size = 20)
for(ii in samp) {  # ii <- samp[5]
  a <- regdat$maize[ii, ]; b <- regdat$soybean[ii, ]
  d <- ID[WTH == a$WTH & zoneid == a$zoneid & SOL == a$SOL]
  e <- ylds$maize[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
                  YR == b$YR][1, ]
  f <- ylds$soybean[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
                    YR == b$YR][1, ]
  # SOIL climate match
  for(i in list(b, d, e, f)) {
    # i <- d
    for(j in c("zoneid", "WTH", "SOL")) {
      # print(j)
      if(a[[j]] != i[[j]]) print(paste("mismatch at", j, "row", ii))
    }
    for(j in c("x", "y")) {
      # print(j)
      if(round(a[[j]], 5) != round(i[[j]], 5)) {  # round because IDr wasn't
        print(paste("mismatch at", j, "row", ii))
      }
    }
  }
  for(i in c("YLDS", "YLDR")) { # i <- "YLDS"
    if(a[[i]] != e[[i]]) print(paste("maize mismatch at", i, "row", ii))
    if(b[[i]] != f[[i]]) print(paste("maize mismatch at", i, "row", ii))
  }
}

# ylds[[1]][ID == 6968 & zoneid == 1]
regdat <- lapply(regdat, function(i) {
  setcolorder(i, c(5, 1:4, 6:ncol(i)))
  i <- setkey(i, "mid")
  i
})

regdt <- rbindlist(lapply(names(regdat), function(x) {
  CROP <- x
  cbind(CROP, regdat[[x]])
}))
save(regdt, file = fp(p_yield, "empirical/regdat.rda"))

```

### 2016-02-10 Rmd/DSSAT-yields.Rmd
Updated on DSSAT and empirical modeling side: 1. Created DSSAT-yields.Rmd, for running DSSAT for both maize and soybean. 2. Results of these go into yield-predictors and empirical-yields to create modeled crop yields (using a GAM). 3. zmd-gsod-stations.Rmd creates the station points used in the analysis.

---
title: "DSSAT soybean yields"
author: "Lyndon Estes"
date: "February 9, 2016"
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# DSSAT soybean runs for Zambia

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(rCropMod)
library(doMC)
library(dismo)

setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path("zambiaToff")
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_dssat <- "~/DSSAT45"  # DSSAT main folder
p_wth <- "~/DSSAT45/Weather/"
p_soil <- "~/DSSAT45/SOIL/"
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"

load(fp(p_yield, "empirical/stations.rda"))

# Read in original soil-climate data from rCropMod
soilgrid <- raster(fp(p_cropmod, "af_soil_grid.tif"))  # soil grid
soilshare <- fread(fp(p_cropmod, "af_soil_share.csv"))  # share of profiles

```

## Weather data
```{r, eval = FALSE}
sdate <- "1948/01/01"
edate <- "2010/12/31"
dts <- seq.Date(as.Date(sdate), as.Date(edate), "days")
yrs <- as.numeric(substr(dts, 1, 4))
ano <- unique(yrs)
mos <- as.numeric(substr(dts, 6, 7))
dys <- as.numeric(substr(dts, 9, 10))

# set up indices for selecting dates
ystrt <- which(mos == 1 & dys == 1) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 12 & dys == 31)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
sindu2 <- sindu[sindu %in% which(ano %in% 1979:2010)]
ind <- which(sind %in% sindu2)

prec <- brick(fp(p_pgf, "prec.Zambia.nc"))
tmax <- brick(fp(p_pgf, "Tmax.Zambia.nc")) 
tmin <- brick(fp(p_pgf, "Tmin.Zambia.nc"))
srad <- brick(fp(p_pgf, "Rs.Zambia.nc")) 
wind <- brick(fp(p_pgf, "wnd10m.Zambia.nc"))

# extract weather station data
registerDoMC(cores = 7)
wdat <- foreach(i = 1:nrow(zmetf)) %dopar% {  # i = 1 
  print(i)
  rain <- unname(extract(prec, zmetf[i, ])[1, ])
  tmx <- unname(extract(tmax, zmetf[i, ])[1, ])
  tmn <- unname(extract(tmin, zmetf[i, ])[1, ])
  sr <- unname(extract(srad, zmetf[i, ])[1, ])
  wnd <- unname(extract(wind, zmetf[i, ])[1, ])
  cbind(rain, tmx, tmn, sr, wnd)
}

# reduce down and modify
wdat2 <- lapply(wdat, function(x) {
  y <- x[ind, ]
  y[, 2] <- y[, 2] - 273.15
  y[, 3] <- y[, 3] - 273.15
  y[, 4] <- y[, 4] / 1000000 * (3600 * 24)
  y
})
# tstwsta <- cbind.data.frame(dts[ind], wdat2[[1]])
# head(tstwsta); tail(tstwsta)

# get elevations for the weather stations
# raster::getData('ISO3')  # raster is a marvelous package
elev <- raster::getData("alt", country = "ZMB")
zmet_elev <- extract(elev, zmetf)
# plot(elev)

# Write WTH files to climate folder
sdate <- "19790101" 
edate <- "20101231"
zmetxy <- as.data.frame(zmetf)
zmetxy[1:6, ]
wthnms <- substr(zmetxy$nms, 1, 4)
ginds <- grep("g", wthnms)
combis <- expand.grid(data.frame(cbind(LETTERS, LETTERS, LETTERS)))[, 3:1]
combis <- paste0(combis[, 1], combis[, 2], combis[, 3])
wthnms[ginds] <- paste0("G", combis[1:length(ginds)])
for(i in 1:length(wdat2)) {  # i <- 1
  xy <- c("x" = zmetxy[i, "x"], "y" = zmetxy[i, "y"]) 
  elev <- zmet_elev[i]
  name <- wthnms[i]
  d <- wdat2[[i]]  # head(d)
  weather(xy, elev = elev, srad = d[, "sr"], tmax = d[, "tmx"], 
          tmin = d[, "tmn"], prec = d[, "rain"], 
          sdate = sdate, edate = edate, name = name, 
          outdir = fp(p_clim, "DSSAT"))
}
file.copy(dir(fp(p_clim, "DSSAT"), full.names = TRUE), 
          fp(p_wth, dir(fp(p_clim, "DSSAT"))), overwrite = TRUE)

```

## Pull the soil profiles
```{r, eval = FALSE}
soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
wthsoil <- cbind(wthnms, zmetxy, soilid)
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile)

# checks
set.seed(234)
for(i in sample(1:nrow(profdat), 4)) {
  print(profdat[prof == profdat$prof[i], c(1:5, 11:16), with = FALSE])
}

# uniprofs <- unique(profs$SoilProfile)
# profdat2 <- do.call(rbind, lapply(1:length(uniprofs), function(i) {
#   print(i)
#   profdat <- read_SOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
#                            profiles = uniprofs[i]) 
# }))
# profdat$slb3 == profdat2$SLB3

# profdat[pro == "WI_VRBW509"]
profdat[, prof := as.character(prof)]
setkey(profdat, "prof")
wthprofs <- data.table(merge(wthsoil, profs, by.x = "soilid", by.y = "CELL5M"))
setnames(wthprofs, "SoilProfile", "prof")
wthsoilm <- merge(wthprofs, profdat, by.x = "prof", by.y = "prof")
wthsoilmr <- wthsoilm[, 1:37, with = FALSE]
names(wthsoilm)
# profdat[, 1:4, with = FALSE]      
# head(wthprofs)
# wthsoilm[1:10, 1:7, with = FALSE]
# wthsoilm[prof == "WI_VRBW509", 1:20, with = FALSE]
```

## Soybean X-files and runs
```{r, eval = FALSE}

# read in field parameters
field <- readxl::read_excel(fp(p_yield, "DSSAT/fields.xls"))  
field <- cbind(wthsoilmr, field)  # join the soils data
field[, WTH := paste0(wthnms, "7932")]
field[, c("ID_FIELD", "XNAME") := list(fid(field$WTH), xname(field$WTH))]

# Create X table
xtab <- x_tab(fids = field)  # table created, with all fixed field parameters
cult <- c("990008", "990009", "990010")
CLNUM <- 1:length(cult)
pdates <- strftime(as.Date("1979-12-31"), "%y%j")
tcomb <- expand.grid(list("PDATE" = pdates, "INGENO" = cult), 
                     stringsAsFactors = FALSE)
ttab <- cbind("N" = 1:nrow(tcomb), tcomb, 
              t_tab(tvars = c("PDATE", "INGENO"), topts = c("MP", "CU"), 
                    ttab = tcomb))

# Speed this up
xtabl <- lapply(1:nrow(xtab), function(x) { # x <- 1
# xtabl <- lapply(1:10, function(x) { # x <- 1
  print(x)
  d <- xtab[x, ]
  d2 <- do.call(rbind, lapply(1:nrow(ttab), function(x) d))
  upd_col <- colnames(d2)[which(colnames(d2) %in% colnames(ttab))]
  d2[, c(upd_col) := ttab[, upd_col]]
  xt <- cbind(data.table(ttab), d2[, !colnames(d2) %in% upd_col, with = FALSE])
})

xrun <- lapply(xtabl, function(x) {  # x <- xtabl[[1]]
  print(x$XNAME[1])
  xf <- copy(x)
  xf <- as.data.frame(xf)
  xfnm <- x_file(xf, outdir = fp(p_dssat, "Soybean"), z = "01", xtype = ".SNX")
  bname <- batch_file(xl = list(xf), xfiles = xfnm, 
                    outdir = full_path(p_dssat, "Soybean"), btype = "SOY")
  exec_csm(projdir = fp(p_yield, "DSSAT"), csmdir = p_dssat, 
         rundir = full_path(p_dssat, "Soybean"), bname = bname)
  sdat <- read_CSM_outfile(rundir = full_path(p_dssat, "Soybean"), 
                           type = "summary", 
                           vars = c("RUNNO", "TRNO", "FNAM", "SOIL_ID...", 
                                    "PDAT", "MDAT", "HWAH", "PRCM"))
})
# save(xrun, file = fp(p_yield, "DSSAT/soyyields.rda"))

soy_ylds <- rbindlist(lapply(1:length(xrun), function(i) { # i <- 1
  xbinder <- xtabl[[i]][1, list(XNAME, wthnms, x, y, prof)]
  setnames(xbinder, c("wthnms", "prof"), c("WTH", "PROF"))
  cbind(xbinder, xrun[[i]])
}))  
# soy_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, plot(V2, V1)]
# soy_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, mean(V1)]
# 3074 mean with current settings

write.table(soy_ylds, file = fp(p_yield, "DSSAT/soy_ylds.csv"), 
            row.names = FALSE, sep = ",")
# soy_ylds[, unique(XNAME)]

# check on soil configuration
tst <- unique(soy_ylds[, .(x, y, WTH, PROF)])
tst[WTH == "LUSA"]
unixy <- unique(wthprofs[, .(x, y)])
for(i in 1:nrow(unixy)) {  # i = 17
  p <- wthsoilmr[x == unixy$x[i] & y == unixy$y[i], .(wthnms, prof)]
  # p <- wthprofs[x == unixy$x[i] & y == unixy$y[i], .(wthnms, prof)]
  tstp <- tst[x == unixy$x[i] & y == unixy$y[i], .(WTH, PROF)]
  if(!all(tstp == p)) print(paste(i, "has a problem"))
#   if(!all(tstp[order(WTH, PROF)] == p[order(wthnms, prof)])) {
#     print(paste(i, "has a problem"))
#   }
}

```

## Maize yields
```{r, eval = FALSE}
# read in field parameters
field <- readxl::read_excel(fp(p_yield, "DSSAT/mzfields.xls"))  
field <- cbind(wthsoilmr, field)  # join the soils data
field[, WTH := paste0(wthnms, "7932")]
field[, c("ID_FIELD", "XNAME") := list(fid(field$WTH), xname(field$WTH))]

# 
# gdd <- brick("external/input_devel/climate/gdd.tif")
# gddmu <- calc(gdd, mean)
# plot(gddmu)

# Create X table
xtab <- x_tab(fids = field)  # table created, with all fixed field parameters
cult <- c("OP9991", "OP9996", "HY9991", "HY9996")
CLNUM <- 1:length(cult)
pdates <- strftime(as.Date("1979-12-01"), "%y%j")
tcomb <- expand.grid(list("PDATE" = pdates, "INGENO" = cult), 
                     stringsAsFactors = FALSE)
ttab <- cbind("N" = 1:nrow(tcomb), tcomb, 
              t_tab(tvars = c("PDATE", "INGENO"), topts = c("MP", "CU"), 
                    ttab = tcomb))

# Speed this up
xtabl <- lapply(1:nrow(xtab), function(x) { # x <- 1
# xtabl <- lapply(1:10, function(x) { # x <- 1
  print(x)
  d <- xtab[x, ]
  d2 <- do.call(rbind, lapply(1:nrow(ttab), function(x) d))
  upd_col <- colnames(d2)[which(colnames(d2) %in% colnames(ttab))]
  d2[, c(upd_col) := ttab[, upd_col]]
  xt <- cbind(data.table(ttab), d2[, !colnames(d2) %in% upd_col, with = FALSE])
})

cdt <- c("Maize", ".MZX")
mxrun <- lapply(xtabl, function(x) {  # x <- xtabl[[1]]
  print(x$XNAME[1])
  xf <- copy(x)
  xf <- as.data.frame(xf)
  xfnm <- x_file(xf, outdir = fp(p_dssat, cdt[1]), z = "01", xtype = cdt[2])
  bname <- batch_file(xl = list(xf), xfiles = xfnm, 
                    outdir = full_path(p_dssat, cdt[1]), btype = cdt[1])
  exec_csm(projdir = fp(p_yield, "DSSAT"), csmdir = p_dssat, 
         rundir = full_path(p_dssat, cdt[1]), bname = bname)
  sdat <- read_CSM_outfile(rundir = full_path(p_dssat, cdt[1]), 
                           type = "summary", 
                           vars = c("RUNNO", "TRNO", "FNAM", "SOIL_ID...", 
                                    "PDAT", "MDAT", "HWAH", "PRCM"))
})
# save(xrun, file = fp(p_yield, "DSSAT/soyyields.rda"))

mz_ylds <- rbindlist(lapply(1:length(mxrun), function(i) { # i <- 1
  xbinder <- xtabl[[i]][1, list(XNAME, wthnms, x, y, prof)]
  setnames(xbinder, c("wthnms", "prof"), c("WTH", "PROF"))
  cbind(xbinder, mxrun[[i]])
}))  
# mz_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, plot(V2, V1)]
# mz_ylds[, lapply(.(HWAH, PRCM), mean), by = XNAME][, mean(V1)]
# 9024 mean with 1/11 start date (SYMBI no difference of course for maize)
# 8904 with 15/11 start date
# 8518 over whole country

write.table(mz_ylds, file = fp(p_yield, "DSSAT/maize_ylds.csv"), 
            row.names = FALSE, sep = ",")
# mz_ylds[, unique(XNAME)]

# check on soil configuration
tst <- unique(mz_ylds[, .(x, y, WTH, PROF)])
tst[WTH == "LUSA"]
unixy <- unique(wthprofs[, .(x, y)])
for(i in 1:nrow(unixy)) {  # i = 17
  p <- wthsoilmr[x == unixy$x[i] & y == unixy$y[i], .(wthnms, prof)]
  tstp <- tst[x == unixy$x[i] & y == unixy$y[i], .(WTH, PROF)]
  if(!all(tstp == p)) print(paste(i, "has a problem"))
}

```

### 2016-02-10 Rmd/empirical-yields.Rmd
Updated on DSSAT and empirical modeling side: 1. Created DSSAT-yields.Rmd, for running DSSAT for both maize and soybean. 2. Results of these go into yield-predictors and empirical-yields to create modeled crop yields (using a GAM). 3. zmd-gsod-stations.Rmd creates the station points used in the analysis.

@@
 p_yield <- fp(p_root, "external/input_devel/yield")
+p_sgrids <- fp(p_root, "external/input_devel/soil")
 
@@
 regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
-regdt[CROP == "maize" & YLDS > 8000, 
+# regdt[CROP == "maize", plot(MAP, YLDS)]
+regdt[CROP == "maize" & YLDS > 9000, 
       points(x, y, pch = 16, cex = 0.5, col = "red")]
-regdt[CROP == "maize" & YLDS > 8000, 
-      points(x, y, pch = 16, cex = 0.5, col = "red")]
-regdt[CROP == "soybean" & YLDS > 1700, 
-      points(x, y, pch = 16, cex = 0.5, col = "blue")]
+# regdt[CROP == "soybean", plot(MAP, YLDR)]
+# regdt[CROP == "soybean" & YLDS > 1700, 
+#       points(x, y, pch = 16, cex = 0.5, col = "blue")]
 
 # Prune out yields > 8000 maize > 1700 soybean in subsistence case
-mzfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
-                .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
-soyfilt <- regdt[CROP == "maize" & YLDS > 8000, unique(.SD), 
-                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
-regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
-# regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]
+# soyfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
+#                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
+mzfilt <- regdt[CROP == "maize" & YLDS > 9000, unique(.SD), 
+                .SDcols = c("x", "y", "SOL", "WTH")]
+# regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
+regnew <- regdt[!x %in% mzfilt$x | !y %in% mzfilt$y]
+regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]
 
@@
 regnew[CROP == "maize", plot(MSP, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sloc, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sloc, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slcl, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slcl, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slsi, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slsi, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sbdm, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sbdm, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slhw, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
-regnew[CROP == "maize", plot(slhw, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
+regnew[CROP == "maize", plot(SLOC, YLDS, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SLOC, YLDR, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SLCL, YLDS, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SLCL, YLDR, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SLSI, YLDS, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SLSI, YLDR, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SBDM, YLDS, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SBDM, YLDR, pch = 16, cex = 0.5)]
+regnew[CROP == "maize", plot(SLHW, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
+regnew[CROP == "maize", plot(SLHW, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
 regnew[CROP == "maize", plot(GDD, YLDS, pch = 16, cex = 0.5)]
@@
 
-summary(lm(YLDS ~ poly(sbdm, 2) + poly(slhw, 2), 
+summary(lm(YLDS ~ poly(SBDM, 2) + poly(SLHW, 2), 
            data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ poly(slhw, 2), data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ poly(sbdm, 2), data = regnew[CROP == "maize", ]))
+summary(lm(YLDS ~ poly(SLHW, 2), data = regnew[CROP == "maize", ]))
+summary(lm(YLDS ~ poly(SBDM, 2), data = regnew[CROP == "maize", ]))
 summary(lm(YLDS ~ poly(KDD, 2), data = regnew[CROP == "maize", ]))
@@
 
-# p <- "sbdm"; d <- "YLDS"
+# p <- "SBDM"; d <- "YLDS"
 # regnew[CROP == "maize", {
@@
 # regnew[CROP == "maize", reg_plotter(YLDR, GDD)]  # 
-regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.045
-regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
-regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.012
+regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.058
+regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.050
+regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.0186
 # regnew[CROP == "maize", reg_plotter(YLDS, JP - mean(JP))]  # P 0.012
-regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.005
-regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.003
+regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.009
+regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.007
 
-regnew[CROP == "maize", reg_plotter(YLDS, slhw)]  # P 0.17
-regnew[CROP == "maize", reg_plotter(YLDS, sloc)]  # P 0.266
-# regnew[CROP == "maize", reg_plotter(YLDR, sloc)]  # P 
-regnew[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.185
-regnew[CROP == "maize", reg_plotter(YLDS, sbdm)]  # 0.181
+regnew[CROP == "maize", reg_plotter(YLDS, SLHW)]  # P 0.16
+regnew[CROP == "maize", reg_plotter(YLDS, SLOC)]  # P 0.43
+# regnew[CROP == "maize", reg_plotter(YLDR, SLOC)]  # P 
+regnew[CROP == "maize", reg_plotter(YLDS, SLCL)]  # P 0.191
+regnew[CROP == "maize", reg_plotter(YLDS, SBDM)]  # 0.20
 # mtax pairings
 regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # P 0.07
-regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.056 
+regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.044 
 # NK
-regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.054
-regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.027 
+regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.056
+regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.024 
 # GDD/KDD
-regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.04
-regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.027
+regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.03
+regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.026
 # srad
-regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.012 
-regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # Same, picked up wrong one
-regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.0288 
-regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.0299
+regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.015 
+regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # 0.0155
+regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.038 
+regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.039
 
 # best predictors look to be
-# MAP, SLOC (but polynomial relationship odd), SLHW, slcl, sbdm, all good
+# MAP, SLOC (but polynomial relationship odd), SLHW, SLCL, SBDM, all good
 # STMX (linear KDD as a possible alternate), GDD, ASRM
@@
 regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4279
+}]  # 0.2941
 
 regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4199
+}]  # 0.2926
 
 regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(sloc, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLOC, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4044 # SLOC doesn't give as much value, NS on linear
+}]  # 0.4244 # SLOC doesn't give as much value, NS on linear
 
 regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4293  # Doesn't add much
+}]  # 0.2961  # Doesn't add much
 
 regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4293  # Oct precip add a tiny bit, significant
+}]  # 0.2972  # Jan precip adds more R2, but makes MAP linear NS
 
 regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4312  # Jan precip adds more R2, but makes MAP linear NS
+}]  # 0.303  # Feb precip even more R2, but makes MAP linear NS
 
 regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4374  # Feb precip even more R2, but makes MAP linear NS
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
   summary(lm1)
@@
 # regnew[CROP == "maize", {
-#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-#               poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
+#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
+#               poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
 #              family = gaussian(link = "log"), method = "ML")
@@
 # regred[, c("Y10S", "Y90S") := as.integer(0)]
-
-regmu <- regnew[CROP == "maize", lapply(.SD, mean), 
-                by = list(CROP, WTH, Zone, SOL), 
-                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
+regmu <- regred[CROP == "maize", lapply(.SD, mean), 
+                by = list(CROP, x, y, WTH, SOL), 
+                .SDcols = names(regred)[c(9:ncol(regred))]]
 regmu[, YST := "mu"]
@@
 
-a <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
-            SOL == "WI_RGTZ054", YLDS]
-b <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
-            SOL == "WI_RGTZ054", mid]
+a <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", YLDS]
+b <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", mid]
 qs(a, b, 0.9)
@@
 m10 <- regred[,
-  qs(YLDS, mid2, 0.1), by = list(CROP, WTH, Zone, SOL)
+  qs(YLDS, mid2, 0.1), by = list(CROP, x, y, WTH, SOL)
 ][, V1]
 m90 <- regred[,
-  qs(YLDS, mid2, 0.9), by = list(CROP, WTH, Zone, SOL) 
+  qs(YLDS, mid2, 0.9), by = list(CROP, x, y, WTH, SOL)
 ][, V1]
@@
 
-regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.071
-regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
-regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.020
-regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.017
+regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.1235
+regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.1097
+regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.036
+regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.026
 regred2[CROP == "maize",  summary(lm(YLDS ~ poly(FP, 2)))]
-regred2[CROP == "maize", reg_plotter(YLDS, 100 - slcl + slsi)]  # P 0.102
-regred2[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.1879
-regred2[CROP == "maize", reg_plotter(YLDS, slsi)]  # P 0.03
+regred2[CROP == "maize", reg_plotter(YLDS, 100 - SLCL + SLSI)]  # P 0.1243
+regred2[CROP == "maize", reg_plotter(YLDS, SLCL)]  # P 0.19
+regred2[CROP == "maize", reg_plotter(YLDS, SLSI)]  # P 0.049
 
 regred2[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
+}]  # 0.3324  # Oct precip add a tiny bit, significant
 
 regred2[CROP == "maize" & YST == "mu", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
+}]  # 0.3271  # Oct precip add a tiny bit, significant
 
 regred2[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
+}]  # 0.34  # Oct precip add a tiny bit, significant
 
 regred2[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(sloc, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
+  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + poly(SLOC, 2) +
+              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
   summary(lm1)
   # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
+}]  # 0.466  # Oct precip add a tiny bit, significant
 
-regred2[CROP == "maize", {
-  gam1 <- gam(YLDS ~ s(MAP) + s(JP) + s(slhw) + s(slcl) + (sbdm) + s(GDD) +
-                KDD + s(ASRM), 
-            family = gaussian(link = "log"), method = "ML")
-}]
-
-gam1 <- gam(YLDS ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
-              s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
+# Fit gams
+gam1 <- gam(YLDS ~ s(MAP, k = 3) + s(JP, k = 3) + s(SLHW, k = 3) + 
+              s(SLCL, k = 3) + s(SBDM, k = 3) + s(GDD, k = 3) + KDD + 
               s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
             data = regred2[CROP == "maize"])
-gam2 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
+gam2 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
               poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam3 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
+gam3 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
               poly(ASRM, 2) + s(x, y), 
@@
             data = regred2[CROP == "maize"])
-gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
+gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
               s(x, y), family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD, 
+gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam6 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(GDD, 2) + KDD, 
+gam6 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam7 <- gam(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(GDD, 2) + KDD, 
+gam7 <- gam(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam8 <- gam(YLDS ~ poly(MAP, 2) + slhw +  poly(slcl, 2) + poly(GDD, 2) + KDD, 
+gam8 <- gam(YLDS ~ poly(MAP, 2) + SLHW +  poly(SLCL, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam9 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-              poly(100 - (slcl + slsi), 2) + poly(GDD, 2) + KDD, 
+gam9 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
+              poly(100 - (SLCL + SLSI), 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam10 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(sloc, 2) + poly(GDD, 2) + KDD, 
+gam10 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
+               poly(SLOC, 2) + poly(GDD, 2) + KDD, 
              family = gaussian(link = "log"), method = "REML", 
              data = regred2[CROP == "maize"])
-gam11 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               sloc + poly(GDD, 2) + KDD, 
+gam11 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
+               SLOC + poly(GDD, 2) + KDD, 
              family = gaussian(link = "log"), method = "REML", 
              data = regred2[CROP == "maize"])
-gam12 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               sloc + poly(GDD, 2) + KDD + s(x, y), 
+gam12 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
+               SLOC + poly(GDD, 2) + KDD + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam13 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + 
+               poly(SLOC, 2) + poly(GDD, 2) + KDD, 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam14 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + 
+               poly(SLOC, 2) + poly(GDD, 2) + KDD + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam15 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
+               poly(GDD, 2) + KDD + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam16 <- gam(YLDS ~ poly(JFMP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
+               poly(GDD, 2) + KDD + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam17 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + SLOC + 
+               poly(GDD, 2) + KDD + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam18 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLHW, 2) + SLOC + 
+               poly(GDD, 2) + s(x, y), family = gaussian(link = "log"), 
+             method = "REML", data = regred2[CROP == "maize"])
+gam19 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + SLHW + SLOC + 
+               poly(GDD, 2) + KDD + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam20 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + 
+               poly(SLOC, 2) + poly(GDD, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam21 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
+               poly(GDD, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam22 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + #(SLCL, 2) + 
+               s(SLOC, k = 3) + poly(GDD, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam23 <- gam(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + 
+               s(SLOC, k = 3) + poly(GDD, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+             data = regred2[CROP == "maize"])
+gam24 <- gam(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + s(SLCL, k = 3) + 
+               s(SLOC, k = 3) + poly(GDD, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", 
@@
 
+summary(gam1)
+plot.gam(gam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam2)
+plot.gam(gam2, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(gam3)
@@
 plot.gam(gam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam11)  # This one or
+summary(gam11)  
 plot.gam(gam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam12)  # of this one, with smooth x, y terms
+summary(gam12)  
 plot.gam(gam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam13)  # this one, without smooth x, y terms
+plot.gam(gam13, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam14)  # of this one, with smooth x, y terms
+plot.gam(gam14, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam15)  # this one, with MSP, is best, with smooth x, y terms
+plot.gam(gam15, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam16)  # JFMP doesn't add as much
+plot.gam(gam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam17)  # Turn SLOC into linear predictor, which makes more sense
+plot.gam(gam17, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam18)  # add polynomial on SLHW, drop KDD. More R2
+plot.gam(gam18, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam19)  # SLHW linear, keep KDD as NS linear 
+plot.gam(gam19, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam20)  # SLHW linear, keep KDD as NS linear 
+plot.gam(gam20, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam21)  # SLHW linear, keep KDD as NS linear 
+plot.gam(gam21, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam22)  ### SLOC smooth, SLCL poly
+plot.gam(gam22, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam23)  ### SLOC smooth and SLHW poly
+plot.gam(gam23, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam24)
+plot.gam(gam24, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+
+plot(predst[["SLCL"]])
 
 par(mfrow = c(2, 2))
-plot(regred2[CROP == "maize", YLDS], predict(gam11, type = "response"))
-plot(regred2[CROP == "maize", YLDS], predict(gam12, type = "response"))
-plot(regred2[CROP == "maize", YLDS], predict(gam4, type = "response"))
+regred2[CROP == "maize", plot(YLDS, predict(gam15, type = "response"))]
+lines(c(2000, 8000), c(2000, 8000), col = "red")
+regred2[CROP == "maize", plot(YLDS, predict(gam17, type = "response"))]
+lines(c(2000, 8000), c(2000, 8000), col = "red")
+regred2[CROP == "maize", plot(YLDS, predict(gam18, type = "response"))]
+lines(c(2000, 8000), c(2000, 8000), col = "red")
+regred2[CROP == "maize", plot(YLDS, predict(gam19, type = "response"))]
+lines(c(2000, 8000), c(2000, 8000), col = "red")
 
 # Reduce down to model for mean, 10th, 90th)
-gam12mu <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + s(x, y), 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[YST == "mu"])
-summary(gam12mu)
-plot.gam(gam12mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-gam1210 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + s(x, y), 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[YST == "10"])
-summary(gam1210)
-plot.gam(gam1210, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-gam1290 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + s(x, y), 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[YST == "90"])
-summary(gam1290)
-plot.gam(gam1290, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+gam18mu <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLHW, 2) + SLOC + 
+               poly(GDD, 2) + s(x, y), family = gaussian(link = "log"), 
+               method = "REML", data = regred2[YST == "mu"])
+summary(gam18mu)  # Rainfall drops out, but still retains shape
+plot.gam(gam18mu, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+gam19mu <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + SLHW + SLOC + 
+               poly(GDD, 2) + KDD + s(x, y), family = gaussian(link = "log"), 
+               method = "REML", data = regred2[YST == "mu"])
+summary(gam19mu)  # Rainfall drops out, but still retains shape
+plot.gam(gam19mu, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+
+# gam1510 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
+#                poly(GDD, 2) + s(x, y),
+#              family = gaussian(link = "log"), method = "REML", 
+#              data = regred2[YST == "10"])
+# summary(gam1510)
+# plot.gam(gam1510, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+# gam1590 <- gam(YLDS ~ poly(MSP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
+#                poly(GDD, 2) + s(x, y),
+#              family = gaussian(link = "log"), method = "REML", 
+#              data = regred2[YST == "90"])
+# summary(gam1590)
+# plot.gam(gam1590, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot(regred2[YST == "mu", YLDS], predict(gam18mu, type = "response"))
+plot(regred2[YST == "mu", YLDS], predict(gam19mu, type = "response"))
+# plot(regred2[YST == "10", YLDS], predict(gam1510, type = "response"))
+# plot(regred2[YST == "90", YLDS], predict(gam1590, type = "response"))
 
@@
 
-# regred[, c("Y10S", "Y90S") := as.integer(0)]
-
-regmus <- regreds[, lapply(.SD, mean), by = list(CROP, WTH, Zone, SOL), 
-                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
+regmus <- regreds[, lapply(.SD, mean), by = list(CROP, x, y, WTH, SOL), 
+                  .SDcols = names(regred)[c(9:ncol(regred))]]
 regmus[, YST := "mu"]
@@
 m10 <- regreds[,
-  qs(YLDR, mid2, 0.1), by = list(CROP, WTH, Zone, SOL)
+  qs(YLDR, mid2, 0.1), by = list(CROP, x, y, WTH, SOL)
 ][, V1]
 m90 <- regreds[,
-  qs(YLDR, mid2, 0.9), by = list(CROP, WTH, Zone, SOL) 
+  qs(YLDR, mid2, 0.9), by = list(CROP, x, y, WTH, SOL)
 ][, V1]
@@
 regreds2 <- rbind(regmus, y1090s[, cnames, with = FALSE])
-regreds2[, slcf := 100 - (slcl + slsi)]
+regreds2[, slcf := 100 - (SLCL + SLSI)]
+regreds2[, ONDP := OP + (MSP - (JP + FP))]
+regreds2[, NDP := (MSP - (JP + FP))]
 
@@
 
-regreds2[, reg_plotter(YLDR, MAP)]  # L 0.065
-regreds2[, reg_plotter(YLDR, MSP)]  # L 0.1105
-regreds2[, reg_plotter(YLDR, JP)]  # L 0.08
-regreds2[, reg_plotter(YLDR, FP)]  # P 0.016
-regreds2[, reg_plotter(YLDR, 100 - slcl + slsi)]  # P 0.12
-regreds2[, reg_plotter(YLDR, slcl)]  # P 0.31
-regreds2[, reg_plotter(YLDR, slsi)]  # P 0.003
-regreds2[, reg_plotter(YLDR, sloc)]  # P 0.31
-regreds2[, reg_plotter(YLDR, slhw)]  # P 0.27
-regreds2[, reg_plotter(YLDR, GDD)]  # P 0.006
-regreds2[, reg_plotter(YLDR, KDD)]  # P 0.006
-regreds2[, reg_plotter(YLDR, STMX)]  # L 0.029
-regreds2[, reg_plotter(YLDR, NKF)]  # L 0.032
-regreds2[, reg_plotter(YLDR, NKJ)]  # P 0.029
-regreds2[, reg_plotter(YLDR, SSRT)]  # L 0.023
-regreds2[, reg_plotter(YLDR, ASRM)]  # L 0.055
+regreds2[, reg_plotter(YLDR, MAP)]  # L 0.19
+regreds2[, reg_plotter(YLDR, MSP)]  # L 0.15
+regreds2[, reg_plotter(YLDR, JFMP)]  # L 0.23
+regreds2[, reg_plotter(YLDR, JP)]  # L 0.04
+regreds2[, reg_plotter(YLDR, FP)]  # P 0.11
+regreds2[, reg_plotter(YLDR, NDP)]  # 
+regreds2[, reg_plotter(YLDR, ONDP)] 
+regreds2[, reg_plotter(YLDR, OP)]
 
-sgam1 <- gam(YLDR ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
-               s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
+regreds2[, reg_plotter(YLDR, 100 - SLCL + SLSI)]  # P 0.16
+regreds2[, reg_plotter(YLDR, SLCL)]  # P 0.15
+regreds2[, reg_plotter(YLDR, SLSI)]  # P 0.01
+regreds2[, reg_plotter(YLDR, SLOC)]  # L 0.16 (decline)
+regreds2[, reg_plotter(YLDR, SLHW)]  # P 0.18
+regreds2[, reg_plotter(YLDR, GDD)]  # NS
+regreds2[, reg_plotter(YLDR, KDD)]  # NS
+regreds2[, reg_plotter(YLDR, STMX)]  # L 0.04
+regreds2[, reg_plotter(YLDR, ATMX)]  # L 0.04
+regreds2[, reg_plotter(YLDR, NKF)]  # L 0.06
+regreds2[, reg_plotter(YLDR, NKJ)]  # P 0.06
+regreds2[, reg_plotter(YLDR, SSRT)]  # P 0.11
+regreds2[, reg_plotter(YLDR, SSRM)]  # P 0.11
+regreds2[, reg_plotter(YLDR, ASRM)]  # L 0.13
+
+sgam1 <- gam(YLDR ~ s(MAP, k = 3) + s(JP, k = 3) + s(SLHW, k = 3) + 
+               s(SLCL, k = 3) + s(SBDM, k = 3) + s(GDD, k = 3) + KDD + 
                s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
             data = regreds2)
-
-sgam2 <- gam(YLDR ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
+sgam2 <- gam(YLDR ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
               poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam3 <- gam(YLDR ~ poly(MAP, 2) + poly(FP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
+sgam2a <- gam(YLDR ~ poly(JFMP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
               poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam4 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
+summary(sgam2); summary(sgam2a)  # MAP better
+sgam3 <- gam(YLDR ~ poly(MAP, 2) + poly(FP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
               poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam5 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sloc, 2) + poly(GDD, 2) + KDD + 
+sgam4 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
               poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam6 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + sloc + poly(GDD, 2) + KDD + 
+sgam5 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SLOC, 2) + poly(GDD, 2) + KDD + 
               poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam7 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(GDD, 2) + KDD + 
+sgam6 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + SLOC + poly(GDD, 2) + KDD + 
               poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam8 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
+sgam7 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(GDD, 2) + KDD + 
+              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam8 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
                poly(GDD, 2) + KDD + poly(ASRM, 2), 
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
+sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
                poly(GDD, 2) + poly(ASRM, 2),
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam10 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(ASRM, 2), 
+sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
+               poly(, 2) + poly(ASRM, 2),
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam11 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(sbdm, 2) + 
-               poly(GDD, 2) + poly(ASRM, 2), 
+sgam10 <- gam(YLDR ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
+               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam12 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2), 
+summary(sgam10)
+sgam11 <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
+               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam13 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + sloc +  
-               poly(GDD, 2) + poly(SSRM, 2), 
+sgam12 <- gam(YLDR ~ poly(JFMP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
+               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam14 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2), 
+sgam13 <- gam(YLDR ~ poly(JFMP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
+                poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam15 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + poly(SSRM, 2), 
+sgam14 <- gam(YLDR ~ poly(MAP, 2) + poly(SLCL, 2) + poly(SLOC, 2) + 
+                poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam16 <- gam(YLDR ~ poly(MSP, 2) + poly(OP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam17 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam18 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + NKJ + poly(SSRM, 2) + s(x, y), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam19 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + NKF + poly(SSRM, 2) + s(x, y), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
+summary(sgam14)
 
 summary(sgam1)
-plot.gam(sgam1, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot.gam(sgam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(sgam2)
 plot.gam(sgam2, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam2a)
+plot.gam(sgam2a, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(sgam3)
@@
 summary(sgam4)
-plot.gam(sgam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot.gam(sgam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(sgam5)
@@
 plot.gam(sgam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam9)  # MSP stronger than MAP
+summary(sgam9)  
 plot.gam(sgam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam10)  # MSP stronger than MAP
+summary(sgam10)  
 plot.gam(sgam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam11)  # SBDM much weaker SLCL
+summary(sgam11)  
 plot.gam(sgam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam12)  # SSRM a tad better than 
+summary(sgam12)  
 plot.gam(sgam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam13)  # SLOC goes in the wrong direction
+summary(sgam13)  
 plot.gam(sgam13, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam14)  # OP adds a bit
+summary(sgam14)  
 plot.gam(sgam14, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam15)  # KDD NS
-plot.gam(sgam15, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam16)  # Polynomial term on October precip doesn't add much
-plot.gam(sgam16, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam17)  # xy term added to model 15, all remain signif, seems way to go
-plot.gam(sgam17, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam18)  # NKJ NS
-plot.gam(sgam18, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam19)  # NKF NS
-plot.gam(sgam19, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+
+regreds2[, plot(MAP, YLDR)]
+regreds2[, plot(JFMP, YLDR)]
+
+plot(regreds2[, YLDR], predict(sgam12, type = "response"))
+lines(c(0, 5000), c(0, 5000), col = "red")
 
 # Now see about taking same model and fitting separately to 10th/90th/mean years
-sgam17mu <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-                  poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
+sgam12mu <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
+               poly(GDD, 2) + poly(ASRM, 2) + s(x, y),
                 family = gaussian(link = "log"), method = "REML", 
                 data = regreds2[YST == "mu"])
-summary(sgam17mu)  # OP drops out, SSRM drops out
-plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-sgam17mu <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + 
-                  poly(slcl, 2) + poly(GDD, 2) + s(x, y), 
-                family = gaussian(link = "log"), method = "REML", 
-                data = regreds2[YST == "mu"])
-summary(sgam17mu)  # OP drops out, SSRM drops out
-plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-sgam1710 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
-                + poly(GDD, 2) + s(x, y), 
-                family = gaussian(link = "log"), method = "REML", 
-                data = regreds2[YST == "10"])
-summary(sgam1710)  # OP drops out, SSRM drops out
-plot.gam(sgam1710, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-sgam1790 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
-                + poly(GDD, 2) + s(x, y), 
-                family = gaussian(link = "log"), method = "REML", 
-                data = regreds2[YST == "90"])
-summary(sgam1790)  # OP drops out, SSRM drops out
-plot.gam(sgam1790, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam12mu)  # OP drops out, SSRM drops out
+# plot.gam(sgam12mu, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+# sgam1810 <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
+#                   SLOC + poly(GDD, 2) + s(x, y),
+#                 family = gaussian(link = "log"), method = "REML", 
+#                 data = regreds2[YST == "10"])
+# summary(sgam1810)  # OP drops out, SSRM drops out
+# plot.gam(sgam1810, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+# sgam1890 <- gam(YLDR ~ poly(MSP, 2) + poly(SLHW, 2) + poly(SLCL, 2) + 
+#                   SLOC + poly(GDD, 2) + s(x, y),
+#                 family = gaussian(link = "log"), method = "REML", 
+#                 data = regreds2[YST == "90"])
+# summary(sgam1890)  # OP drops out, SSRM drops out
+# plot.gam(sgam1890, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+
+par(mfrow = c(2, 2))
+plot(regreds2[, YLDR], predict(sgam12, type = "response"))
+lines(c(0, 5000), c(0, 5000), col = "red")
+plot(regreds2[YST == "mu", YLDR], predict(sgam12mu, type = "response"))
+lines(c(0, 5000), c(0, 5000), col = "red")
+# plot(regreds2[YST == "10", YLDR], predict(sgam1810, type = "response"))
+# plot(regreds2[YST == "90", YLDR], predict(sgam1890, type = "response"))
 
@@
 
-### Save GAMs outputs
+## Load in grid predictors
 ```{r, eval = FALSE}
-save(gam12, gam12mu, gam1210, gam1290, sgam17, sgam17mu, sgam1710, sgam1790, 
-     file = fp(p_yield, "gams.rda"))
+soilgrids <- brick(fp(p_sgrids, "soilgrids.tif"))
+climvars <- c("rftot", "ndjftot", "jfmtot", "gdd", "kdd", "srmu_ano")
+climgrids <- sapply(climvars, function(x) {  # x <- climvars[3]
+  print(x)
+  brick(dir(p_clim, full.names = TRUE)[grep(x, dir(p_clim))])
+})
+climgrids_mu <- sapply(climgrids, function(x) calc(x, mean))
+climst <- stack(climgrids_mu)
+names(climst) <- c("MAP", "MSP", "JFMP", "GDD", "KDD", "ASRM")
+
+# disaggregate down to 1 km
+climstbase <- disaggregate(climst[[1]], fact = 30)
+climststep <- disaggregate(climst, fact = 20)
+climst1km <- resample(climst, climstbase)
+
+# need to warp soilgrids back to that also
+ext <- bbox(climst1km)[1:4]  # extent   
+onm <- fp(p_sgrids, "soilgridsgcs.tif")
+gdalwarp(srcfile = soilgrids@file@name, t_srs = projection(climst1km), 
+         dstfile = onm, r = "average", ot = "Float32", te = ext, 
+         srcnodata = -32768, dstnodata = -99, tr = res(climst1km), 
+         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
+soilgcs <- brick(onm)
+names(soilgcs) <- c("SLOC", "SLCL", "SLHW")
+plot(soilgcs)
+
+# need also a grid of x,y values
+x <- soilgcs[[1]]
+x[] <- xFromCell(x, 1:ncell(x))
+y <- soilgcs[[1]]
+y[] <- yFromCell(y, 1:ncell(y))
+plot(y)
+xy <- stack(x, y)
+names(xy) <- c("x", "y")
+predst <- stack(xy, climst1km, soilgcs)
 
 ```
+
+## Predicted yields surfaces
+```{r, eval = FALSE}
+mzfull1 <- raster::predict(predst, gam18, type = "response")
+mzfull2 <- raster::predict(predst, gam19, type = "response")
+mzfull3 <- raster::predict(predst, gam22, type = "response")
+mzfull4 <- raster::predict(predst, gam23, type = "response")
+soyfull <- raster::predict(predst, sgam10, type = "response")
+soyfull2 <- raster::predict(predst, sgam13, type = "response")
+soyfull3 <- raster::predict(predst, sgam14, type = "response")
+summary(sgam10)
+predict(gam15)
+# mzfull1[mzfull1 > 8000] <- 8000
+# mzfull2[mzfull2 > 8000] <- 8000
+# soyfull[soyfull > 5000] <- 5000
+par(mfrow = c(2, 2))
+plot(mzfull1, zlim = c(0, 9000))
+plot(mzfull2, zlim = c(0, 9000))
+plot(mzfull3, zlim = c(0, 9000))
+plot(mzfull4, zlim = c(0, 9000))
+
+prov@proj4string <- soyfull@crs
+soymsk <- crop(soyfull, prov)
+soymsk2 <- crop(soyfull2, prov)
+soymsk3 <- crop(soyfull3, prov)
+plot(soymsk > 5000)
+plot(prov, add = TRUE)
+plot(soymsk2)
+plot(prov, add = TRUE)
+plot(soymsk3 > 5000)
+plot(prov, add = TRUE)
+
+# Save bricks and models
+save(gam18, gam19, gam22, gam23, sgam10, sgam13, sgam14, 
+     file = fp(p_yield, "empirical/gams1.rda"))
+yldrs <- brick(stack(list(mzfull1, mzfull2, mzfull3, mzfull4, soyfull,
+                          soyfull2, soyfull3)))
+writeRaster(yldrs, filename = fp(p_yield, "empirical/yield_surfaces1.tif"))
+
+```

### 2016-02-10 Rmd/yield-predictors.rmd
Updated on DSSAT and empirical modeling side: 1. Created DSSAT-yields.Rmd, for running DSSAT for both maize and soybean. 2. Results of these go into yield-predictors and empirical-yields to create modeled crop yields (using a GAM). 3. zmd-gsod-stations.Rmd creates the station points used in the analysis.

@@
 library(agroEcoTradeoff)
+library(gdalUtils)
 library(rCropMod)
@@
 p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
+p_sgrids <- "~/Dropbox/data/soil/"
 
@@
 # Yield data selection points
-# GSOD points
-gsod <- fread(fp(p_clim, "gsod.txt"), stringsAsFactors = FALSE)
-setnames(gsod, names(gsod), c("nms", "x", "y", "n"))
-gsod <- gsod[, 1:3, with = FALSE]
-gsod[, nms := paste0("g", nms)]
-# plot(idgr)
-# points(zamgsod)  # not representative enough
+load(fp(p_yield, "empirical/stations.rda"))
+zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
+zmetxy <- data.frame(zmetf)[, c("x", "y")]
+# xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
+# yFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$y
+# zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
+zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy
 
-# ZMD rainfall stations
-nms <- c('CHIPAT01', 'CHIPEP01', 'CHOMA001', 'ISOKA001', 'KABOMP01', 'KABWE001',
-         'KAFIRO01', 'KAFUE001', 'KALABO01', 'KAOMA001', 'KASAMA01', 'KASEMP01',          'KAWAMB01', 'LIVING01', 'LUNDAZ01', 'LUSAKA01', 'LUSAKA02', 'LUSITU01',          'MAGOYE01', 'MANSA001', 'MBALA001', 'MFUWE001', 'MISAMF01', 'MKUSHI01',          'MONGU002', 'MPIKA001')
-x <- c(32.58, 28.38, 27.07, 32.63, 24.20, 28.47, 28.12, 27.92, 22.70, 24.80,
-       31.13, 25.85, 29.08, 25.82, 33.20, 28.32, 28.45, 29.37, 27.60, 28.85, 
-       31.33, 31.93, 31.25, 29.33, 23.15, 31.43)  
-y <- c(-13.55, -17.33, -16.83, -10.12, -13.60, -14.45, -12.60, -15.77, -14.95,
-       -14.80, -10.22, -13.53, -9.80, -17.82, -12.28, -15.42, -15.32, -16.30,
-       -16.00, -11.10, -8.85, -13.27, -10.10, -14.00, -15.25, -11.90)
-zmd <- cbind.data.frame(nms = as.character(nms), x, y, 
-                        stringsAsFactors = FALSE)
-write.csv(zmd, file = fp(p_climo, "zmd-stations.csv"))
+# zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
+zmet_dt <- data.table(data.frame(zmetsamps))
+# setkeyv(zmet_dt, c("x", "y"))
+setkeyv(zmet_dt, c("ID", "zoneid"))
 
-# bring together GSOD and ZMD stations
-zmd_gsod <- rbind(zmd, gsod)
-coordinates(zmd_gsod) <- c("x", "y")
+# setkeyv(IDr, c("x", "y"))
+setkeyv(IDr, c("ID", "zoneid"))
+IDr_ss <- merge(IDr, zmet_dt)
+# IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
+IDr_ss[, c("x.x", "y.x", "cells") := NULL]
+setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
+setkeyv(IDr_ss, c("WTH", "Zone"))
 
-# Select those intersecting Zambia
-gint <- rgeos::gIntersects(prov, zmd_gsod, byid = TRUE)
-sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
-zmet <- zmd_gsod[sel, ]
-# str(zmet)
+# Collect soils up here also 
+pts <- IDr_ss
+# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
+coordinates(pts) <- ~x+y
+soilid <- extract(soilgrid, pts)  # grid id corresponding to point
+IDr_ss[, soilid := soilid]
+profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
+profs[, SharePct := NULL]
+IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
+IDr_ss[, soilid := NULL]
+setnames(IDr_ss, "SoilProfile", "SOL")
+setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))
 
-# buffer the ZMD stations by ~20 km to see if any GSOD stations fall within that
-zbuffer <- rgeos::gBuffer(zmet[1:26, ], byid = TRUE, width = 0.3)
-zbuffer <- spChFIDs(zbuffer, as.character(1:nrow(zbuffer)))
-gint <- rgeos::gIntersects(zbuffer, zmet[27:nrow(zmet), ], byid = TRUE)
-sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
-desel <- c(27:nrow(zmet))[sel]
-# plot(zmet[-desel, ], pch = 16, cex = 0.5)
-# points(zmet[desel, ], pch = 16, cex = 0.5, col = "red")
-# plot(prov, add = TRUE)
-zmetf <- zmet[-desel, ]
-
-zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
-zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
-
-zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
-setkeyv(zmet_dt, c("x", "y"))
-setkeyv(IDr, c("x", "y"))
-IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
-setkeyv(IDr_ss, c("WTH", "Zone"))
 # IDr_ss[, plot(x, y)]
@@
 ```{r, eval = FALSE}
-crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
+# crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
+crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
 fylds <- lapply(crops, function(X) {  # X <- "maize"
@@
 mzr <- dt_to_raster(rdt, raster()@crs)
-rdt <- fylds$soybean[, c(1:2, grep("YS", names(fylds$soybean))), with = FALSE]
-soys <- dt_to_raster(rdt, raster()@crs)
-rdt <- fylds$soybean[, c(1:2, grep("YR", names(fylds$soybean))), with = FALSE]
-soyr <- dt_to_raster(rdt, raster()@crs)
 
 # par(mar = rep(0, 4))
-# plot(calc(mzs, mean))
+plot(calc(mzs, max))
+plot(calc(mzr, mean))
+# plot(calc(soyr, mean))
 # plot(zmetf, pch = 16, cex = 0.5, add = TRUE)
@@
 
+crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
 # Yield data for those points
@@
     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
-    setkeyv(cr, c("WTH", "Zone", "SOL"))
+    setkeyv(cr, c("Zone", "WTH", "SOL"))
 
@@
     onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
-    yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
+    # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
+    # Selecting rowMax here
+    yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
     yldmerge[, c(onms[[2]]) := NULL]
@@
       })
-      if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
+      #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
+      if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
         print(paste("Yield mismatch at", ii, mannms[jj]))
@@
 
+soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
+# soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
+soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
+soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
+setnames(soy_ylds, "SOIL_ID...", "SOL")
+
+# calculate mean and max as YLDS and YLDR
+soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
+setnames(soy_ylds2, "V1", "YLDS")
+soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
+soy_ylds2[, XNAME := NULL]
+
+ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
+soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
+setnames(soy_ylds2, "V1", "mid")
+
+ylds <- list(ylds$maize, soy_ylds2)
+names(ylds) <- c("maize", "soybean")
+
 ```
@@
 ## Soil
-
+### Profile data
 Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
@@
 profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
-profdat <- read_dssat_prof(solfile = full_path(p_soil, "WI.SOL"), 
+profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                              profiles = profs$SoilProfile) 
@@
 # max depth in profile
-mdepth <- rowMaxs(as.matrix(profdat[, .SD, .SDcols = grep("slb", lnms)]), 
+mdepth <- rowMaxs(as.matrix(profdat[, .SD, .SDcols = grep("SLB", lnms)]), 
                   na.rm = TRUE)
@@
 # calc. Horizon depths become weights for weighted mean. 
-tst <- profdat[, .SD , .SDcols = grep("slb", lnms)]
+tst <- profdat[, .SD , .SDcols = grep("SLB", lnms)]
 dtest <- function(x) {
@@
 solvarsr <- solvars[solvars %in% 
-                     c("sbdm", "sloc", "slcl", "slsi", "slhw")]
-solvals <- do.call(cbind, lapply(solvarsr, function(i) {  # i <- annvars[2]
+                     c("SBDM", "SLOC", "SLCL", "SLSI", "SLHW")]
+solvals <- do.call(cbind, lapply(solvarsr, function(i) {  # i <- solvarsr[2]
   mat <- as.matrix(profdat[, .SD , .SDcols = grep(i, lnms)])
@@
     matv <- mat[j, ind]
-    matv <- infill(x = matv, ind)
-    weighted.mean(matv, depvec[j, ind], na.rm = TRUE)
+    matv <- infill(x = matv, ind)  # fill in NA values within profiles
+    w <- depvec[j, ind] - c(0, depvec[j, ind[-length(ind)]])  # depth weights
+    # weighted.mean(matv, depvec[j, ind], na.rm = TRUE)  # this was wrong
+    weighted.mean(matv, w, na.rm = TRUE)
   })
@@
 # finalize
-sol1m <- cbind("pro" = profdat[, pro], data.table(solvals))
+sol1m <- cbind("prof" = profdat[, prof], data.table(solvals))
+# unique(ylds$soybean$SOL) %in% sol1m$prof
+```
+
+### ISRIC grids
+```{r, eval = FALSE}
+p_spreds <- fp(p_root, "external/input_devel/soil")
+mgrid <- raster("agroEcoTradeoff/external/data/ZA/ZA-mask.tif")
+ext <- bbox(mgrid)[1:4]  # extent   
+crpshp <- as(extent(mgrid), "SpatialPolygons")
+crpshp@proj4string <- mgrid@crs
+
+depths <- c(0, 5, 15, 30, 60, 100) 
+thickness <- depths[-1] - depths[-length(depths)]
+
+p_soc <- fp(p_sgrids, "carbon/SoilGrids1km")
+fnms <- dir(p_soc, pattern = "sd[1-5]", full.names = TRUE)
+socr <- stack(lapply(fnms, function(x) {
+  r <- raster(x)
+  r <- crop(r, spTransform(crpshp, r@crs)) / 10 # crop, convert permille to %
+}))
+# plot(socr[[1]])
+
+# wmu <- function(x, na.rm = na.rm) weighted.mean(x, w = thickness, na.rm = na.rm) 
+# soc1m <- calc(socr, fun = wmu, na.rm = TRUE)
+soc1m <- weighted.mean(socr, w = thickness)
+soc1m <- writeRaster(soc1m, filename = fp(p_spreds, "soctemp.tif"), 
+                     overwrite = TRUE)
+# soc1m <- raster(fp(p_spreds, "soctemp.tif"))
+# weighted.mean(socr[1], thickness) == soc1m[1]
+
+onm <- fp(p_spreds, "soctempalb.tif")
+gdalwarp(srcfile = soc1m@file@name, t_srs = projection(mgrid), 
+         dstfile = onm, r = "average", ot = "Float32", te = ext, 
+         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
+         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
+soc1m <- raster(onm)
+
+# clay grids
+p_cl <- fp(p_sgrids, "clay/SoilGrids1km")
+fnms <- dir(p_cl, pattern = "sd[1-5]", full.names = TRUE)
+sclr <- stack(lapply(fnms, function(x) {
+  r <- raster(x)
+  r <- crop(r, spTransform(crpshp, r@crs))
+  # r[r > 100] <- 100
+  r
+}))
+scl1m <- weighted.mean(sclr, w = thickness)
+scl1m <- writeRaster(scl1m, filename = fp(p_spreds, "claytemp.tif"), 
+                     overwrite = TRUE)
+# scl1m <- raster(fp(p_spreds, "claytemp.tif"))
+# weighted.mean(sclr[2], thickness) == scl1m[2]
+onm <- fp(p_spreds, "claytempalb.tif")
+gdalwarp(srcfile = scl1m@file@name, t_srs = projection(mgrid), 
+         dstfile = onm, r = "average", ot = "Float32", te = ext, 
+         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
+         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
+scl1m <- raster(onm)
+
+# ph
+p_ph <- fp(p_sgrids, "pH/SoilGrids1km")
+fnms <- dir(p_ph, pattern = "sd[1-5]", full.names = TRUE)
+sphr <- stack(lapply(fnms, function(x) {
+  r <- raster(x)
+  r <- crop(r, spTransform(crpshp, r@crs)) / 10
+  r
+}))
+sph1m <- weighted.mean(sphr, w = thickness)
+sph1m <- writeRaster(sph1m, filename = fp(p_spreds, "phtemp.tif"), 
+                     overwrite = TRUE)
+# sph1m <- raster(fp(p_spreds, "phtemp.tif"))
+# weighted.mean(sphr[2], thickness) == sph1m[2]
+
+onm <- fp(p_spreds, "phtempalb.tif")
+gdalwarp(srcfile = sph1m@file@name, t_srs = projection(mgrid), 
+         dstfile = onm, r = "average", ot = "Float32", te = ext, 
+         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
+         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
+sph1m <- raster(onm)
+# plot(sph1m)
+
+soilgrids <- brick(stack(list("SLOC" = soc1m, "SLCL" = scl1m, "SLHW" = sph1m)))
+soilgrids <- writeRaster(soilgrids, filename = fp(p_spreds, "soilgrids.tif"), 
+                         overwrite = TRUE)
+# plot(soilgrids)
 ```
@@
 for(i in sindu) febind[which(sind == i & mos == 2)] <- i
+jfmind <- rep(0, length(dts))  # djf of each growing season
+for(i in sindu) jfmind[which(sind == i & mos %in% c(1, 2, 3))] <- i
 ndjfind <- rep(0, length(dts))  # djf of each growing season
@@
 writeRaster(ndjftot, filename = fp(p_clim, "ndjftot.tif"), overwrite = TRUE)
+
+# JFM season rainfall
+jfmtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
+  ind <- which(jfmind == sindu2[i])
+  calc(prec[[ind]], sum)
+})
+writeRaster(jfmtot, filename = fp(p_clim, "jfmtot.tif"), overwrite = TRUE)
 
@@
 kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
-writeRaster(kdd, filename = fp(p_clim, "kdd.tif"))
+writeRaster(kdd, filename = fp(p_clim, "kdd.tif"), overwrite = TRUE)
 gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
-writeRaster(gdd, filename = fp(p_clim, "gdd.tif"))
+writeRaster(gdd, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)
 
@@
 })
-writeRaster(nkilljan, filename = fp(p_clim, "nkilljan.tif"))
+writeRaster(nkilljan, filename = fp(p_clim, "nkilljan.tif"), overwrite = TRUE)
 nkillfeb <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
@@
 })
-writeRaster(nkillfeb, filename = fp(p_clim, "nkillfeb.tif"))
+writeRaster(nkillfeb, filename = fp(p_clim, "nkillfeb.tif"), overwrite = TRUE)
 
@@
 })
-writeRaster(tmax_gs, filename = fp(p_clim, "tmax_gs.tif"))
+writeRaster(tmax_gs, filename = fp(p_clim, "tmax_gs.tif"), overwrite = TRUE)
 
@@
 })
-writeRaster(tmax_ano, filename = fp(p_clim, "tmax_ano.tif"))
+writeRaster(tmax_ano, filename = fp(p_clim, "tmax_ano.tif"), overwrite = TRUE)
 
@@
 })
-writeRaster(srtot_gs, filename = fp(p_clim, "srtot_gs.tif"))
+writeRaster(srtot_gs, filename = fp(p_clim, "srtot_gs.tif"), overwrite = TRUE)
 srmu_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
@@
 })
-writeRaster(srmu_gs, filename = fp(p_clim, "srmu_gs.tif"))
+writeRaster(srmu_gs, filename = fp(p_clim, "srmu_gs.tif"), overwrite = TRUE)
 
@@
 })
-writeRaster(srtot_ano, filename = fp(p_clim, "srtot_ano.tif"))
+writeRaster(srtot_ano, filename = fp(p_clim, "srtot_ano.tif"), 
+            overwrite = TRUE)
 srmu_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
@@
 })
-writeRaster(srmu_ano, filename = fp(p_clim, "srmu_ano.tif"))
+writeRaster(srmu_ano, filename = fp(p_clim, "srmu_ano.tif"), overwrite = TRUE)
 ```
@@
 # Organize rainfall
-climnms <- c("rftot", "ndjftot", "jantot", "febtot", "octtot", "kdd", "gdd", 
-             "nkilljan", "nkillfeb", "tmax_ano", "tmax_gs", "srtot_ano",
-             "srtot_mu", "srtot_gs", "srmu_gs")
+climnms <- c("rftot", "ndjftot", "jfmtot",  "jantot", "febtot", "octtot", "kdd",
+             "gdd", "nkilljan", "nkillfeb", "tmax_ano", "tmax_gs", "srtot_ano",
+             "srmu_ano", "srtot_gs", "srmu_gs")
 bnms <- sapply(climnms, function(x) dir(p_clim, pattern = x, full.names = TRUE))
@@
 
-onms <- c("MAP", "MSP", "JP", "FP", "OP", "KDD", "GDD", "NKJ", "NKF", "ATMX",
-          "STMX", "SSRT", "SSRM", "ASRT", "ASRM")
+onms <- c("MAP", "MSP", "JFMP", "JP", "FP", "OP", "KDD", "GDD", "NKJ", "NKF", 
+          "ATMX", "STMX", "SSRT", "SSRM", "ASRT", "ASRM")
 climl <- lapply(1:length(blist), function(x) {
@@
 names(climl) <- onms
-
-# rfnms <- c("MAP", "MSP", "JP", "FP", "OP")
-# rfl <- list(rftot, ndjftot, jantot, febtot, octtot)
-# rfl <- lapply(1:length(rfl), function(x) {
-#   r <- rfl[[x]]
-#   names(r) <- paste0(rfnms[x], 1:nlayers(r))
-#   r
-# })
-# names(rfl) <- rfnms 
-# 
-# # Organize temp
-# tnms <- c("KDD", "GDD", "NKJ", "NKF", "ATMX", "STMX")
-# tfl <- list(kdd, gdd, nkilljan, nkillfeb, tmax_ano, tmax_gs)
-# tfl <- lapply(1:length(tfl), function(x) {
-#   r <- tfl[[x]]
-#   names(r) <- paste0(tnms[x], 1:nlayers(r))
-#   r
-# })
-# names(tfl) <- tnms 
-# 
-# # Organize srad
-# snms <- c("SSRT", "SSRM", "ASRT", "ASRM")
-# sfl <- list(srtot_gs, srmu_gs, srtot_ano, srmu_ano)
-# sfl <- lapply(1:length(sfl), function(x) {
-#   r <- sfl[[x]]
-#   names(r) <- paste0(snms[x], 1:nlayers(r))
-#   r
-# })
-# names(sfl) <- snms
 
@@
 
-setnames(sol1m, "pro", "SOL")
+setnames(sol1m, "prof", "SOL")
 setkey(sol1m, "SOL")
@@
 
-# Checking merges
-set.seed(234)
-samp <- sample(1:nrow(regdat$maize), size = 20)
-for(ii in samp) {  # ii <- samp[5]
-  a <- regdat$maize[ii, ]; b <- regdat$soybean[ii, ]
-  d <- ID[WTH == a$WTH & zoneid == a$zoneid & SOL == a$SOL]
-  e <- ylds$maize[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
-                  YR == b$YR][1, ]
-  f <- ylds$soybean[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
-                    YR == b$YR][1, ]
-  # SOIL climate match
-  for(i in list(b, d, e, f)) {
-    # i <- d
-    for(j in c("zoneid", "WTH", "SOL")) {
-      # print(j)
-      if(a[[j]] != i[[j]]) print(paste("mismatch at", j, "row", ii))
-    }
-    for(j in c("x", "y")) {
-      # print(j)
-      if(round(a[[j]], 5) != round(i[[j]], 5)) {  # round because IDr wasn't
-        print(paste("mismatch at", j, "row", ii))
-      }
-    }
-  }
-  for(i in c("YLDS", "YLDR")) { # i <- "YLDS"
-    if(a[[i]] != e[[i]]) print(paste("maize mismatch at", i, "row", ii))
-    if(b[[i]] != f[[i]]) print(paste("maize mismatch at", i, "row", ii))
-  }
-}
+all(regdat$maize$SOL == regdat$soybean$SOL)
+all(regdat$maize$x == regdat$soybean$x)
+all(regdat$maize$y == regdat$soybean$y)
+all(regdat$maize$OP == regdat$soybean$OP)
+all(regdat$maize$SLOC == regdat$soybean$SLOC)
+
+# Checking merges (disabled at present)
+# set.seed(234)
+# samp <- sample(1:nrow(regdat$maize), size = 20)
+# for(ii in samp) {  # ii <- samp[5]
+#   a <- regdat$maize[ii, ]; b <- regdat$soybean[ii, ]
+#   d <- ID[WTH == a$WTH & zoneid == a$zoneid & SOL == a$SOL]
+#   e <- ylds$maize[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
+#                   YR == b$YR][1, ]
+#   f <- ylds$soybean[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
+#                     YR == b$YR][1, ]
+#   # SOIL climate match
+#   for(i in list(b, d, e, f)) {
+#     # i <- d
+#     for(j in c("zoneid", "WTH", "SOL")) {
+#       # print(j)
+#       if(a[[j]] != i[[j]]) print(paste("mismatch at", j, "row", ii))
+#     }
+#     for(j in c("x", "y")) {
+#       # print(j)
+#       if(round(a[[j]], 5) != round(i[[j]], 5)) {  # round because IDr wasn't
+#         print(paste("mismatch at", j, "row", ii))
+#       }
+#     }
+#   }
+#   for(i in c("YLDS", "YLDR")) { # i <- "YLDS"
+#     if(a[[i]] != e[[i]]) print(paste("maize mismatch at", i, "row", ii))
+#     if(b[[i]] != f[[i]]) print(paste("maize mismatch at", i, "row", ii))
+#   }
+# }
 
@@
   setcolorder(i, c(5, 1:4, 6:ncol(i)))
+  setnames(i, "mid", "midold")
+  i <- cbind(data.table("mid" = 1:nrow(i)), i)
   i <- setkey(i, "mid")
@@
 })
+
+all(regdat$maize$SOL == regdat$soybean$SOL)
+all(regdat$maize$x == regdat$soybean$x)
+all(regdat$maize$y == regdat$soybean$y)
+all(regdat$maize$OP == regdat$soybean$OP)
+all(regdat$maize$SLOC == regdat$soybean$SLOC)
 

### 2016-02-10 Rmd/zmd-gsod-stations.Rmd
Updated on DSSAT and empirical modeling side: 1. Created DSSAT-yields.Rmd, for running DSSAT for both maize and soybean. 2. Results of these go into yield-predictors and empirical-yields to create modeled crop yields (using a GAM). 3. zmd-gsod-stations.Rmd creates the station points used in the analysis.

---
title: "ZMD GSOD stations"
author: "Lyndon Estes"
date: "February 9, 2016"
output: html_document
---

# ZMD and GSOD stations

```{r, eval = FALSE}
library(data.table)
setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"

# provinces
p_prov <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                 "rda_dump/Final Shape files RDA/")
prov <- rgdal::readOGR(fp(p_prov, "prov.shp"), layer = "prov")

# Yield data selection points
# GSOD points
gsod <- fread(fp(p_clim, "gsod.txt"), stringsAsFactors = FALSE)
setnames(gsod, names(gsod), c("nms", "x", "y", "n"))
gsod <- gsod[, 1:3, with = FALSE]
gsod[, nms := paste0("g", nms)]

# ZMD rainfall stations
nms <- c('CHIPAT01', 'CHIPEP01', 'CHOMA001', 'ISOKA001', 'KABOMP01', 'KABWE001',
         'KAFIRO01', 'KAFUE001', 'KALABO01', 'KAOMA001', 'KASAMA01', 'KASEMP01',          'KAWAMB01', 'LIVING01', 'LUNDAZ01', 'LUSAKA01', 'LUSAKA02', 'LUSITU01',          'MAGOYE01', 'MANSA001', 'MBALA001', 'MFUWE001', 'MISAMF01', 'MKUSHI01',          'MONGU002', 'MPIKA001') 
x <- c(32.58, 28.38, 27.07, 32.63, 24.20, 28.47, 28.12, 27.92, 22.70, 24.80,
       31.13, 25.85, 29.08, 25.82, 33.20, 28.32, 28.45, 29.37, 27.60, 28.85, 
       31.33, 31.93, 31.25, 29.33, 23.15, 31.43)  
y <- c(-13.55, -17.33, -16.83, -10.12, -13.60, -14.45, -12.60, -15.77, -14.95,
       -14.80, -10.22, -13.53, -9.80, -17.82, -12.28, -15.42, -15.32, -16.30,
       -16.00, -11.10, -8.85, -13.27, -10.10, -14.00, -15.25, -11.90)
zmd <- cbind.data.frame(nms = as.character(nms), x, y, 
                        stringsAsFactors = FALSE)
write.csv(zmd, file = fp(p_climo, "zmd-stations.csv"))

# bring together GSOD and ZMD stations
zmd_gsod <- rbind(zmd, gsod)
coordinates(zmd_gsod) <- c("x", "y")

# Select those intersecting Zambia
gint <- rgeos::gIntersects(prov, zmd_gsod, byid = TRUE)
sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
zmet <- zmd_gsod[sel, ]
# str(zmet)

# buffer the ZMD stations by ~20 km to see if any GSOD stations fall within that
zbuffer <- rgeos::gBuffer(zmet[1:26, ], byid = TRUE, width = 0.3)
zbuffer <- spChFIDs(zbuffer, as.character(1:nrow(zbuffer)))
gint <- rgeos::gIntersects(zbuffer, zmet[27:nrow(zmet), ], byid = TRUE)
sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
desel <- c(27:nrow(zmet))[sel]
# plot(zmet[-desel, ], pch = 16, cex = 0.5)
# points(zmet[desel, ], pch = 16, cex = 0.5, col = "red")
# plot(prov, add = TRUE)
zmetf <- zmet[-desel, ]

# save to yields folder because these are primarily used in empirical modeling
save(zmet, zmetf, file = fp(p_yield, "empirical/stations.rda"))
```

### 2016-02-10 Rmd/yield-predictors.rmd
Redid yield-predictors.Rmd with DSSAT-modeled maize, so greatly reduced now--no complicated merges.

@@
 
-# Predict DSSAT yields onto finer soil map
+# Develop predictors for two-stage yield model
 
@@
 
+setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
 p_root <- lmisc::set_base_path()
@@
 p_sgrids <- "~/Dropbox/data/soil/"
+p_dssat <- fp(p_yield, "DSSAT")
 
@@
 # Read in soil-climate grid from Di's work
-ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
-ID[, V1 := NULL]
-setkeyv(ID, c("WTH", "SOL", "Zone"))
-ID[, SharePct := as.numeric(SharePct)]
-zone <- sort(unique(ID$Zone))
-for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
-setkeyv(ID, c("WTH", "Zone"))
-# ID[, FID := 1:.N]
-IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
-setkeyv(IDr, c("WTH", "Zone"))
-# ID[SOL == "WI_ARBW438"]
-gcs <- "+proj=longlat +datum=WGS84"
-idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
+# ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
+# ID[, V1 := NULL]
+# setkeyv(ID, c("WTH", "SOL", "Zone"))
+# ID[, SharePct := as.numeric(SharePct)]
+# zone <- sort(unique(ID$Zone))
+# for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
+# setkeyv(ID, c("WTH", "Zone"))
+# # ID[, FID := 1:.N]
+# IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
+# setkeyv(IDr, c("WTH", "Zone"))
+# # ID[SOL == "WI_ARBW438"]
+# idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
 # plot(idgr)
@@
 # Yield data selection points
+gcs <- "+proj=longlat +datum=WGS84"
 load(fp(p_yield, "empirical/stations.rda"))
-zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
-zmetxy <- data.frame(zmetf)[, c("x", "y")]
+zmetf@proj4string <- CRS(gcs)
+
+# zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
+# zmetxy <- data.frame(zmetf)[, c("x", "y")]
+zmetxy <- data.table(data.frame(zmetf)[, 1:3])
+wthnms <- substr(zmetxy$nms, 1, 4)
+ginds <- grep("g", wthnms)
+combis <- expand.grid(data.frame(cbind(LETTERS, LETTERS, LETTERS)))[, 3:1]
+combis <- paste0(combis[, 1], combis[, 2], combis[, 3])
+wthnms[ginds] <- paste0("G", combis[1:length(ginds)])
+zmetxy <- cbind(data.table("WTH" = wthnms), zmetxy)
+
 # xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
@@
 # zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
-zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy
+# zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy
 
 # zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
-zmet_dt <- data.table(data.frame(zmetsamps))
-# setkeyv(zmet_dt, c("x", "y"))
-setkeyv(zmet_dt, c("ID", "zoneid"))
-
-# setkeyv(IDr, c("x", "y"))
-setkeyv(IDr, c("ID", "zoneid"))
-IDr_ss <- merge(IDr, zmet_dt)
-# IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
-IDr_ss[, c("x.x", "y.x", "cells") := NULL]
-setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
-setkeyv(IDr_ss, c("WTH", "Zone"))
+# zmet_dt <- data.table(data.frame(zmetsamps))
+# # setkeyv(zmet_dt, c("x", "y"))
+# setkeyv(zmet_dt, c("ID", "zoneid"))
+# 
+# # setkeyv(IDr, c("x", "y"))
+# setkeyv(IDr, c("ID", "zoneid"))
+# IDr_ss <- merge(IDr, zmet_dt)
+# # IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
+# IDr_ss[, c("x.x", "y.x", "cells") := NULL]
+# setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
+# setkeyv(IDr_ss, c("WTH", "Zone"))
 
 # Collect soils up here also 
-pts <- IDr_ss
+# pts <- IDr_ss
 # pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
-coordinates(pts) <- ~x+y
-soilid <- extract(soilgrid, pts)  # grid id corresponding to point
-IDr_ss[, soilid := soilid]
-profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
-profs[, SharePct := NULL]
-IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
-IDr_ss[, soilid := NULL]
-setnames(IDr_ss, "SoilProfile", "SOL")
-setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))
+# coordinates(pts) <- ~x+y
+# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
+# soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
+# IDr_ss[, soilid := soilid]
+# profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
+# profs[, SharePct := NULL]
+# IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
+# IDr_ss[, soilid := NULL]
+# setnames(IDr_ss, "SoilProfile", "SOL")
+# setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))
 
@@
 ```{r, eval = FALSE}
-# crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
-crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
-fylds <- lapply(crops, function(X) {  # X <- "maize"
-  fnms <- dir(full_path(p_crop, X), pattern = X)
-  fl <- unlist(lapply(c("S", "R"), function(x) {
-    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
-  }))  # reorder to read subsistence first
-  # idt <- copy(IDr)
-  # key(idt)
-  idt <- copy(ID)
-  idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
-  setkeyv(idt, c("WTH", "Zone"))
-  for(y in 1:length(fl)) { # y <- 1
-    # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
-    print(paste("Processing", fl[y]))
-    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
-    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
-    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
-    cr[, WTH := substr(WTH, 1, 4)]
-    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
-    setkeyv(cr, c("WTH", "Zone"))
-
-    crm <- merge(cr, idt, by = c("WTH", "Zone"))
-
-    # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]
-
-    # Soil averaged yields
-    solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
-                 by = .(WTH, Zone, YR)]
-    snm <- gsub("*.*_|\\.csv", "", fl[y])
-
-    # Yield means
-    v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
-    setnames(v, "V1", paste0("Y", snm))
-    idt <- idt[v]
-    setkeyv(idt, c("WTH", "Zone"))
-    idt
-  }
-  idt
-})
-names(fylds) <- crops
-
-rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
-mzs <- dt_to_raster(rdt, raster()@crs)
-rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
-mzr <- dt_to_raster(rdt, raster()@crs)
-
-# par(mar = rep(0, 4))
-plot(calc(mzs, max))
-plot(calc(mzr, mean))
-# plot(calc(soyr, mean))
-# plot(zmetf, pch = 16, cex = 0.5, add = TRUE)
+# # crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
+# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
+# fylds <- lapply(crops, function(X) {  # X <- "maize"
+#   fnms <- dir(full_path(p_crop, X), pattern = X)
+#   fl <- unlist(lapply(c("S", "R"), function(x) {
+#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
+#   }))  # reorder to read subsistence first
+#   # idt <- copy(IDr)
+#   # key(idt)
+#   idt <- copy(ID)
+#   idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
+#   setkeyv(idt, c("WTH", "Zone"))
+#   for(y in 1:length(fl)) { # y <- 1
+#     # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
+#     print(paste("Processing", fl[y]))
+#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
+#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
+#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
+#     cr[, WTH := substr(WTH, 1, 4)]
+#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
+#     setkeyv(cr, c("WTH", "Zone"))
+#     
+#     crm <- merge(cr, idt, by = c("WTH", "Zone"))
+# 
+#     # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]
+#     
+#     # Soil averaged yields
+#     solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
+#                  by = .(WTH, Zone, YR)]
+#     snm <- gsub("*.*_|\\.csv", "", fl[y])
+#     
+#     # Yield means
+#     v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
+#     setnames(v, "V1", paste0("Y", snm))
+#     idt <- idt[v]
+#     setkeyv(idt, c("WTH", "Zone"))
+#     idt
+#   }
+#   idt
+# })
+# names(fylds) <- crops
+# 
+# rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
+# mzs <- dt_to_raster(rdt, raster()@crs)
+# rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
+# mzr <- dt_to_raster(rdt, raster()@crs)
+# 
+# # par(mar = rep(0, 4))
+# plot(calc(mzs, max))
+# plot(calc(mzr, mean))
+# # plot(calc(soyr, mean))
+# # plot(zmetf, pch = 16, cex = 0.5, add = TRUE)
 
@@
 ```{r, eval=FALSE}
-# merge function for lists of data.tables
-mymerge <- function(x, y) merge(x, y, all = TRUE)
+# mymerge <- function(x, y) merge(x, y, all = TRUE)
+crops <- c("maize", "soy")
+ylds <- lapply(crops, function(X) {  # X <- crops[2]
+  fnms <- dir(p_dssat, pattern = paste0(X, "_"), full.names = TRUE)
+  cr <- fread(fnms, stringsAsFactors = FALSE)
+  cr[, YR := RUNNO - min(RUNNO) + 1, by = .(XNAME, TRNO)]  # unique year  
+  cr[, c("SOIL_ID...", "RUNNO") := NULL]
 
-crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
-# Yield data for those points
-mannms <- c("S", "R")
-ylds <- lapply(crops, function(X) {
-  # X <- crops[1]
-  fnms <- dir(full_path(p_crop, X), pattern = X)
-  fl <- unlist(lapply(mannms, function(x) {
-    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
-  }))  # reorder to read subsistence first
-  # fl <- fl[1:3]
-  crml <- lapply(1:length(fl), function(y) { # y <- 1
-    idt <- copy(IDr_ss)
-    print(paste("Processing", fl[y]))
-    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
-    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
-    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
-    cr[, WTH := substr(WTH, 1, 4)]
-    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
-    setkeyv(cr, c("Zone", "WTH", "SOL"))
-
-    # merge IDs with subset location
-    # crm <- cr[idt]
-    crm <- merge(cr, idt)
-    # crm[, SharePct := NULL]
-    nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
-    setcolorder(crm, nms)
-    mid <- 1:nrow(crm)
-    crm <- cbind(mid, crm)
-    keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
-    setkeyv(crm, keys)
-    # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
-    newnm <- gsub(".*._||\\.csv", "", fl[y])
-    setnames(crm, "HWAH", newnm)
-    crm
-  })
-  yldmerge <- Reduce(mymerge, crml)
-  ynms <- names(yldmerge)
-  # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
-    # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
-  # }))))
-  # calculate 
-  for(i in mannms) {
-    onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
-    # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
-    # Selecting rowMax here
-    yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
-    yldmerge[, c(onms[[2]]) := NULL]
-  }
-
-  # Internal check on yield means
-  if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
-  if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
-  for(ii in sample(1:nrow(yldmerge), size = 20)) {
-    for(jj in 1:length(mannms)) {  # jj = 2
-      a <- yldmerge[ii, ]
-      b <- sapply(crind[[jj]], function(x) {
-        crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
-      })
-      #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
-      if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
-        print(paste("Yield mismatch at", ii, mannms[jj]))
-      }
-    }
-  }
-
-  # swap in averaged yields
-  # yldmerge[, mid := NULL]
-  # yldmerge <- cbind(yldmerge, yldred)
-  yldmerge
+  # Yields. S becomes mean across cultivars, R becomes max of cultivars
+  cr2 <- cr[, round(mean(HWAH)), by = .(XNAME, x, y, WTH, PROF, YR)]
+  setnames(cr2, c("V1", "PROF"), c("YLDS", "SOL"))
+  cr2[, YLDR := cr[, round(max(HWAH)), by = .(XNAME, YR)][, V1]]
+  cr2
 })
@@
 
-soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
-# soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
-soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
-soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
-setnames(soy_ylds, "SOIL_ID...", "SOL")
-
-# calculate mean and max as YLDS and YLDR
-soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
-setnames(soy_ylds2, "V1", "YLDS")
-soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
-soy_ylds2[, XNAME := NULL]
-
-ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
-soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
-setnames(soy_ylds2, "V1", "mid")
-
-ylds <- list(ylds$maize, soy_ylds2)
-names(ylds) <- c("maize", "soybean")
+# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
+# # Yield data for those points
+# mannms <- c("S", "R")
+# ylds <- lapply(crops, function(X) {
+#   # X <- crops[1]
+#   fnms <- dir(full_path(p_crop, X), pattern = X)
+#   fl <- unlist(lapply(mannms, function(x) {
+#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
+#   }))  # reorder to read subsistence first
+#   # fl <- fl[1:3]
+#   crml <- lapply(1:length(fl), function(y) { # y <- 1
+#     idt <- copy(IDr_ss)
+#     print(paste("Processing", fl[y]))
+#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
+#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
+#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
+#     cr[, WTH := substr(WTH, 1, 4)]
+#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
+#     setkeyv(cr, c("Zone", "WTH", "SOL"))
+# 
+#     # merge IDs with subset location
+#     # crm <- cr[idt]
+#     crm <- merge(cr, idt)
+#     # crm[, SharePct := NULL]
+#     nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
+#     setcolorder(crm, nms)
+#     mid <- 1:nrow(crm)
+#     crm <- cbind(mid, crm)
+#     keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
+#     setkeyv(crm, keys)
+#     # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
+#     newnm <- gsub(".*._||\\.csv", "", fl[y])
+#     setnames(crm, "HWAH", newnm)
+#     crm
+#   })
+#   yldmerge <- Reduce(mymerge, crml)
+#   ynms <- names(yldmerge)
+#   # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
+#     # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
+#   # }))))
+#   # calculate 
+#   for(i in mannms) {
+#     onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
+#     # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
+#     # Selecting rowMax here
+#     yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
+#     yldmerge[, c(onms[[2]]) := NULL]
+#   }
+#   
+#   # Internal check on yield means
+#   if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
+#   if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
+#   for(ii in sample(1:nrow(yldmerge), size = 20)) {
+#     for(jj in 1:length(mannms)) {  # jj = 2
+#       a <- yldmerge[ii, ]
+#       b <- sapply(crind[[jj]], function(x) {
+#         crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
+#       })
+#       #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
+#       if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
+#         print(paste("Yield mismatch at", ii, mannms[jj]))
+#       }
+#     }
+#   }
+#   
+#   # swap in averaged yields
+#   # yldmerge[, mid := NULL]
+#   # yldmerge <- cbind(yldmerge, yldred)
+#   yldmerge
+# })
+# names(ylds) <- crops
+# 
+# soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
+# # soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
+# soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
+# soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
+# setnames(soy_ylds, "SOIL_ID...", "SOL")
+# 
+# # calculate mean and max as YLDS and YLDR
+# soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
+# setnames(soy_ylds2, "V1", "YLDS")
+# soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
+# soy_ylds2[, XNAME := NULL]
+# 
+# ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
+# soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
+# setnames(soy_ylds2, "V1", "mid")
+# 
+# ylds <- list(ylds$maize, soy_ylds2)
+# names(ylds) <- c("maize", "soybean")
 
@@
 ```{r, eval = FALSE}
-pts <- IDr_ss[, .(x, y)]
+# pts <- IDr_ss[, .(x, y)]
 # pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
-coordinates(pts) <- ~x+y
-soilid <- extract(soilgrid, pts)  # grid id corresponding to point
+# coordinates(pts) <- ~x+y
+soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
+# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
 profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
@@
                              profiles = profs$SoilProfile)
+setnames(profdat, "prof", "SOL")
 solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
@@
 # finalize
-sol1m <- cbind("prof" = profdat[, prof], data.table(solvals))
+sol1m <- cbind("SOL" = profdat[, SOL], data.table(solvals))
 # unique(ylds$soybean$SOL) %in% sol1m$prof
@@
 for(i in sindu) octind[which(sind == i & mos == 10)] <- i
+djfmind <- rep(0, length(dts))  # feb of each year
+for(i in sindu) djfmind[which(sind == i & mos  %in% c(12, 1:3))] <- i
+# cbind(yrs[sind == 1], mos[sind == 1], djfmind[sind == 1])
 
@@
 #
-# # storm depth in NDJF
-# rdays <- lapply(sindu2, function(x) { # x <- sindu2[1]
-#   print(x)
-#   ryr <- prec[[which(djfind == x)]]
-#   r <- calc(ryr, nrday)
-#   names(r) <- as.character(ano[x])
-#   r
-# })
-# rdays <- stack(rdays)
-# 
-# # dry days in January
-# drydaysj <- lapply(sindu2, function(x) { # x <- sindu2[1]
-#   print(x)
-#   ryr <- prec[[which(janind == x)]]
-#   r <- calc(ryr, ndry)
-#   names(r) <- as.character(ano[x])
-#   r
-# })
-# drydaysj <- stack(drydaysj)
-# 
-# # dry days in February
-# drydaysf <- lapply(sindu2, function(x) { # x <- sindu2[1]
-#   print(x)
-#   ryr <- prec[[which(febind == x)]]
-#   r <- calc(ryr, ndry)
-#   names(r) <- as.character(ano[x])
-#   r
-# })
-# drydaysf <- stack(drydaysf)
 
@@
 writeRaster(ndjftot, filename = fp(p_clim, "ndjftot.tif"), overwrite = TRUE)
+
+# DJFM season rainfall
+djfmtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
+  ind <- which(djfmind == sindu2[i])
+  calc(prec[[ind]], sum)
+})
+writeRaster(djfmtot, filename = fp(p_clim, "djfmtot.tif"), overwrite = TRUE)
 
@@
 
-# par(mfrow = c(6, 6), mar = c(0, 0, 0, 0))
-# for(i in 1:nlayers(octtot)) {
-#   plot(octtot[[i]], axes = FALSE, box = FALSE)
-#   plot(prov, add = TRUE, zlim = c(0, 1000))
-# }
-# which(1980:2010 %in% ) 
-# pre2005 <- calc(octtot[[1:26]], mean)
-# post2005 <- calc(octtot[[27:31]], mean)
-# 
-# par(mfrow = c(1, 2), mar = c(0, 0, 0, 0))
-# plot(pre2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
-# plot(prov, add = TRUE)
-# plot(post2005, axes = FALSE, box = FALSE, zlim = c(0, 400))
-# plot(prov, add = TRUE)
-# 
-# par(mfrow = c(1, 1), mar = c(0, 0, 0, 0))
-# plot((pre2005 - post2005) / pre2005 * 100, zlim = c(-100, 100))
-# plot(prov, add = TRUE)
-
 # plot(rftot[[1]] - ndjftot[[1]])
@@
 }
+writeRaster(stack(tmu), filename = fp(p_clim, "tmu.tif"), overwrite = TRUE)
 
 # Functions
+gddf <- function(x) {
+  a <- x - 8
+  sum(ifelse(a < 0, 0, a))
+}
+
 # GDD between 8-30 degrees
@@
 writeRaster(kdd, filename = fp(p_clim, "kdd.tif"), overwrite = TRUE)
-gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
-writeRaster(gdd, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)
+gdd30 <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
+writeRaster(gdd30, filename = fp(p_clim, "gdd30.tif"), overwrite = TRUE)
+gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gddf))
+writeRaster(gdd30, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)
 
@@
 writeRaster(srtot_gs, filename = fp(p_clim, "srtot_gs.tif"), overwrite = TRUE)
-srmu_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
+srmu_gs1 <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
   ind <- which(ndjfind == sindu2[i])
@@
 })
-writeRaster(srmu_gs, filename = fp(p_clim, "srmu_gs.tif"), overwrite = TRUE)
+writeRaster(srmu_gs1, filename = fp(p_clim, "srmu_gs1.tif"), overwrite = TRUE)
+
+# DJFM mean solar radiation
+srmu_gs2 <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
+  ind <- which(djfmind == sindu2[i])
+  mj <- mjday(srad[[ind]])
+  calc(mj, mean)
+})
+writeRaster(srmu_gs2, filename = fp(p_clim, "srmu_gs2.tif"), overwrite = TRUE)
+
+# JFM mean solar radiation
+srmu_gs3 <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
+  ind <- which(jfmind == sindu2[i])
+  mj <- mjday(srad[[ind]])
+  calc(mj, mean)
+})
+writeRaster(srmu_gs3, filename = fp(p_clim, "srmu_gs3.tif"), overwrite = TRUE)
 
@@
     + Weather/climate: 
-        +`rftot`, `ndjftot`, `jantot`, `febtot` for rainfall; `kdays`, `gdd`, `f30jan`, `f30feb`, `tmax_gs`, `tmax_ano` for temperature; `srtot_gs`, `srmu_gs`, `srtot_ano`, and `srmu_ano` for solar radiation
+        +`rftot`, `ndjftot`, `jantot`, `febtot` for rainfall; `kdays`, `gdd`, `f30jan`, `f30feb`, `tmax_gs`, `tmax_ano` for temperature; `srtot_gs`, `srmu_gs`, `srtot_ano`, and `srmu_ano` for solar radiation (plus others added on the way)
     + Soil: mean soil properties within top 1 m
-    + Management variables: Codes S, R, I
+    + Management parameters: Codes S, R, 
 2. Predictands: mean yields across cultivars within each management type
@@
 # Organize rainfall
-climnms <- c("rftot", "ndjftot", "jfmtot",  "jantot", "febtot", "octtot", "kdd",
-             "gdd", "nkilljan", "nkillfeb", "tmax_ano", "tmax_gs", "srtot_ano",
-             "srmu_ano", "srtot_gs", "srmu_gs")
+climnms <- c("rftot", "ndjftot", "^djfmtot", "^jfmtot", "jantot", "febtot",
+             "octtot", "kdd", "gdd\\.", "gdd30", "nkilljan", "nkillfeb",
+             "tmax_ano", "tmax_gs", "srtot_ano", "srmu_ano", "srtot_gs",
+             "srmu_gs1", "srmu_gs2", "srmu_gs3")
 bnms <- sapply(climnms, function(x) dir(p_clim, pattern = x, full.names = TRUE))
@@
 
-onms <- c("MAP", "MSP", "JFMP", "JP", "FP", "OP", "KDD", "GDD", "NKJ", "NKF", 
-          "ATMX", "STMX", "SSRT", "SSRM", "ASRT", "ASRM")
+onms <- c("MAP", "MSP", "DJFMP", "JFMP", "JP", "FP", "OP", "KDD", "GDD", 
+          "GDDL", "NKJ", "NKF", "ATMX", "STMX", "ASRT", "ASRM", "SSRT", "SSRMI",
+          "SSRMII", "SSRMIII")
 climl <- lapply(1:length(blist), function(x) {
@@
 # xy[, ind := 1:.N]
-cols <- c("x", "y")
-xy[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]
+# cols <- c("x", "y")
+# xy[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]
+# coordinates(xy) <- ~x+y
 
 # climate predictors
+mymerge <- function(x, y) merge(x, y, all = TRUE)
 climvals <- lapply(1:length(climl), function(i) extract(x = climl[[i]], xy))
-climxy <- lapply(climvals, function(x) {  # x <- rvals[[1]]
+climxy <- lapply(climvals, function(x) {  # x <- climvals[[18]]
   rmelt <- melt(cbind(xy, x), id = 1:2)
@@
 
-# # rainfall
-# rvals <- lapply(1:length(rfl), function(i) extract(x = rfl[[i]], xy))
-# rvxy <- lapply(rvals, function(x) {  # x <- rvals[[1]]
-#   rmelt <- melt(cbind(xy, x), id = 1:2)
-#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
-#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
-#   setnames(rmelt, "variable", "YR")
-#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
-#   rmelt[, ind := 1:.N]
-#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
-#   rmelt
-# })
-# rfpreds <- Reduce(mymerge, rvxy)
-# 
-# # temperature
-# tvals <- lapply(1:length(tfl), function(i) extract(x = tfl[[i]], xy))
-# tvxy <- lapply(tvals, function(x) {  # x <- tvals[[3]]
-#   rmelt <- melt(cbind(xy, x), id = 1:2)
-#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
-#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
-#   setnames(rmelt, "variable", "YR")
-#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
-#   rmelt[, ind := 1:.N]
-#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
-#   rmelt
-# })
-# tfpreds <- Reduce(mymerge, tvxy)
-# 
-# 
-# # srad
-# svals <- lapply(1:length(sfl), function(i) extract(x = sfl[[i]], xy))
-# svxy <- lapply(svals, function(x) {  # x <- rvals[[1]]
-#   rmelt <- melt(cbind(xy, x), id = 1:2)
-#   setnames(rmelt, "value", unique(gsub("[0-9]", "", rmelt$variable)))
-#   rmelt[, variable := gsub("[[:alpha:]]", "", variable)]
-#   setnames(rmelt, "variable", "YR")
-#   # rmelt[, (cols) := lapply(.SD, as.character), .SDcols = cols]
-#   rmelt[, ind := 1:.N]
-#   setkeyv(rmelt, c("ind", "x", "y", "YR"))
-#   rmelt
-# })
-# sfpreds <- Reduce(mymerge, svxy)
-
-# Join all climate predictors
-# climpreds <- Reduce(mymerge, list(rfpreds, tfpreds, sfpreds))
-# climpreds[, YR := as.integer(YR)]
-# setkeyv(climpreds, c("x", "y", "YR"))
-
-setnames(sol1m, "prof", "SOL")
+# soil predictors
 setkey(sol1m, "SOL")
@@
 # merge together all the data
+cols <- c("x", "y")
 regdat <- lapply(ylds, function(i) {  # i <- ylds[[1]]
@@
 
-all(regdat$maize$SOL == regdat$soybean$SOL)
-all(regdat$maize$x == regdat$soybean$x)
-all(regdat$maize$y == regdat$soybean$y)
-all(regdat$maize$OP == regdat$soybean$OP)
-all(regdat$maize$SLOC == regdat$soybean$SLOC)
+all(regdat$maize$SOL == regdat$soy$SOL)
+all(regdat$maize$x == regdat$soy$x)
+all(regdat$maize$y == regdat$soy$y)
+all(regdat$maize$OP == regdat$soy$OP)
+all(regdat$maize$SLOC == regdat$soy$SLOC)
 
@@
 # ylds[[1]][ID == 6968 & zoneid == 1]
-regdat <- lapply(regdat, function(i) {
-  setcolorder(i, c(5, 1:4, 6:ncol(i)))
-  setnames(i, "mid", "midold")
-  i <- cbind(data.table("mid" = 1:nrow(i)), i)
-  i <- setkey(i, "mid")
-  i
-})
-
-all(regdat$maize$SOL == regdat$soybean$SOL)
-all(regdat$maize$x == regdat$soybean$x)
-all(regdat$maize$y == regdat$soybean$y)
-all(regdat$maize$OP == regdat$soybean$OP)
-all(regdat$maize$SLOC == regdat$soybean$SLOC)
+# regdat <- lapply(regdat, function(i) {
+#   setcolorder(i, c(5, 1:4, 6:ncol(i)))
+#   setnames(i, "mid", "midold")
+#   i <- cbind(data.table("mid" = 1:nrow(i)), i)
+#   i <- setkey(i, "mid")
+#   i
+# })
+# 
+# all(regdat$maize$SOL == regdat$soybean$SOL)
+# all(regdat$maize$x == regdat$soybean$x)
+# all(regdat$maize$y == regdat$soybean$y)
+# all(regdat$maize$OP == regdat$soybean$OP)
+# all(regdat$maize$SLOC == regdat$soybean$SLOC)
 

### 2016-02-10 Rmd/yield-predictors.rmd
Deleted prior code for collecting yields from gridded yield datasets and merging with predictors

@@
 
-# Read in soil-climate grid from Di's work
-# ID <- fread(full_path(p_crop, "Soil_clim_match_ID_master.csv"))
-# ID[, V1 := NULL]
-# setkeyv(ID, c("WTH", "SOL", "Zone"))
-# ID[, SharePct := as.numeric(SharePct)]
-# zone <- sort(unique(ID$Zone))
-# for(i in 1:length(zone)) ID[Zone == zone[i], zoneid := i]
-# setkeyv(ID, c("WTH", "Zone"))
-# # ID[, FID := 1:.N]
-# IDr <- subset(unique(ID[, .(x, y, ID, WTH, Zone, zoneid)]))
-# setkeyv(IDr, c("WTH", "Zone"))
-# # ID[SOL == "WI_ARBW438"]
-# idgr <- dt_to_raster(IDr[, .(x, y, ID, zoneid)], CRS(gcs))
-# plot(idgr)
-
 # Yield data selection points
@@
 
-# zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
-# zmetxy <- data.frame(zmetf)[, c("x", "y")]
 zmetxy <- data.table(data.frame(zmetf)[, 1:3])
@@
 zmetxy <- cbind(data.table("WTH" = wthnms), zmetxy)
-
-# xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
-# yFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$y
-# zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
-# zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy
-
-# zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
-# zmet_dt <- data.table(data.frame(zmetsamps))
-# # setkeyv(zmet_dt, c("x", "y"))
-# setkeyv(zmet_dt, c("ID", "zoneid"))
-# 
-# # setkeyv(IDr, c("x", "y"))
-# setkeyv(IDr, c("ID", "zoneid"))
-# IDr_ss <- merge(IDr, zmet_dt)
-# # IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
-# IDr_ss[, c("x.x", "y.x", "cells") := NULL]
-# setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
-# setkeyv(IDr_ss, c("WTH", "Zone"))
-
-# Collect soils up here also 
-# pts <- IDr_ss
-# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
-# coordinates(pts) <- ~x+y
-# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
-# soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
-# IDr_ss[, soilid := soilid]
-# profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
-# profs[, SharePct := NULL]
-# IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
-# IDr_ss[, soilid := NULL]
-# setnames(IDr_ss, "SoilProfile", "SOL")
-# setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))
-
-# IDr_ss[, plot(x, y)]
-# points(zmet_dt, pch = 16, cex = 0.5)
-# tst <- extract(idgr, zmet_dt)
-# all(tst[order(tst[, 1]), 2] == zmetsamps[order(zmetsamps[, 4]), 5])
-
-# alternative for yield data selection points:
-# Creatre stratified, representative random sample of 50 pts from ID/zone grid
-# ztab <- freq(idgr$zoneid)[-5, ]
-# ztabs <- round((ztab[, 2] / sum(ztab[, 2])) * 50)
-# set.seed(34)
-# gsamps <- do.call(rbind, lapply(1:4, function(i) {
-#   gsamp <- sampleRandom(idgr$zoneid == i, size = ztabs[i], cells = TRUE)
-#   sampxy <- xyFromCell(idgr, gsamp[, "cell"])
-# }))
-# plot(idgr$ID)  # unique(values(idgr$ID))
-# length(which(!is.na(values(idgr$ID))))
-# length(which(!is.na(unique(values(idgr$ID)))))
-# points(gsamps, pch = 16, cex = 0.5)
-# 
-# # merge those points with the main ID grid
-# gsamps_dt <- data.table(gsamps)
-# setkeyv(gsamps_dt, c("x", "y"))
-# setkeyv(IDr, c("x", "y"))
-# IDr_ss <- IDr[gsamps_dt]  # pick up IDs, soils, wth variables
-# setkeyv(IDr_ss, c("WTH", "SOL", "Zone"))
-# IDr_ss[, plot(x, y)]
-# points(gsamps, pch = 16, cex = 0.5)
-
-# sgrid <- fread(fp(p_cropmod, "Soil_clim_match_ID.csv"), 
-#                stringsAsFactors = FALSE)
-# setnames(sgrid, old = c("V1", "X", "Y", "WeatherGrid", "SoilType"), 
-#          new = c("ID", "x", "y", "WTH", "SOL"))
-# wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
-# sr <- dt_to_raster(sgrid[, list(ID, x, y)], CRSobj = wgs)
-# setkeyv(sgrid, c("x", "y"))
-# sgrid[SOL == "WI_ARBW438"]
-```
-
-# Full DSSAT yield grids
-```{r, eval = FALSE}
-# # crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
-# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
-# fylds <- lapply(crops, function(X) {  # X <- "maize"
-#   fnms <- dir(full_path(p_crop, X), pattern = X)
-#   fl <- unlist(lapply(c("S", "R"), function(x) {
-#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
-#   }))  # reorder to read subsistence first
-#   # idt <- copy(IDr)
-#   # key(idt)
-#   idt <- copy(ID)
-#   idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
-#   setkeyv(idt, c("WTH", "Zone"))
-#   for(y in 1:length(fl)) { # y <- 1
-#     # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
-#     print(paste("Processing", fl[y]))
-#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
-#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
-#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
-#     cr[, WTH := substr(WTH, 1, 4)]
-#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
-#     setkeyv(cr, c("WTH", "Zone"))
-#     
-#     crm <- merge(cr, idt, by = c("WTH", "Zone"))
-# 
-#     # nms <- names(crm)[c(6:7, 8, 1:2, 4, 9, 5, 3)]
-#     
-#     # Soil averaged yields
-#     solmu <- crm[, round(weighted.mean(HWAH, w = SharePct)), 
-#                  by = .(WTH, Zone, YR)]
-#     snm <- gsub("*.*_|\\.csv", "", fl[y])
-#     
-#     # Yield means
-#     v <- solmu[, round(mean(V1)), by = .(WTH, Zone)]
-#     setnames(v, "V1", paste0("Y", snm))
-#     idt <- idt[v]
-#     setkeyv(idt, c("WTH", "Zone"))
-#     idt
-#   }
-#   idt
-# })
-# names(fylds) <- crops
-# 
-# rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
-# mzs <- dt_to_raster(rdt, raster()@crs)
-# rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
-# mzr <- dt_to_raster(rdt, raster()@crs)
-# 
-# # par(mar = rep(0, 4))
-# plot(calc(mzs, max))
-# plot(calc(mzr, mean))
-# # plot(calc(soyr, mean))
-# # plot(zmetf, pch = 16, cex = 0.5, add = TRUE)
 
@@
 
-# crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
-# # Yield data for those points
-# mannms <- c("S", "R")
-# ylds <- lapply(crops, function(X) {
-#   # X <- crops[1]
-#   fnms <- dir(full_path(p_crop, X), pattern = X)
-#   fl <- unlist(lapply(mannms, function(x) {
-#     fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
-#   }))  # reorder to read subsistence first
-#   # fl <- fl[1:3]
-#   crml <- lapply(1:length(fl), function(y) { # y <- 1
-#     idt <- copy(IDr_ss)
-#     print(paste("Processing", fl[y]))
-#     cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
-#     cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
-#     setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
-#     cr[, WTH := substr(WTH, 1, 4)]
-#     cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
-#     setkeyv(cr, c("Zone", "WTH", "SOL"))
-# 
-#     # merge IDs with subset location
-#     # crm <- cr[idt]
-#     crm <- merge(cr, idt)
-#     # crm[, SharePct := NULL]
-#     nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
-#     setcolorder(crm, nms)
-#     mid <- 1:nrow(crm)
-#     crm <- cbind(mid, crm)
-#     keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
-#     setkeyv(crm, keys)
-#     # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
-#     newnm <- gsub(".*._||\\.csv", "", fl[y])
-#     setnames(crm, "HWAH", newnm)
-#     crm
-#   })
-#   yldmerge <- Reduce(mymerge, crml)
-#   ynms <- names(yldmerge)
-#   # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
-#     # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
-#   # }))))
-#   # calculate 
-#   for(i in mannms) {
-#     onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
-#     # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
-#     # Selecting rowMax here
-#     yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
-#     yldmerge[, c(onms[[2]]) := NULL]
-#   }
-#   
-#   # Internal check on yield means
-#   if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
-#   if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
-#   for(ii in sample(1:nrow(yldmerge), size = 20)) {
-#     for(jj in 1:length(mannms)) {  # jj = 2
-#       a <- yldmerge[ii, ]
-#       b <- sapply(crind[[jj]], function(x) {
-#         crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
-#       })
-#       #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
-#       if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
-#         print(paste("Yield mismatch at", ii, mannms[jj]))
-#       }
-#     }
-#   }
-#   
-#   # swap in averaged yields
-#   # yldmerge[, mid := NULL]
-#   # yldmerge <- cbind(yldmerge, yldred)
-#   yldmerge
-# })
-# names(ylds) <- crops
-# 
-# soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
-# # soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
-# soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
-# soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
-# setnames(soy_ylds, "SOIL_ID...", "SOL")
-# 
-# # calculate mean and max as YLDS and YLDR
-# soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
-# setnames(soy_ylds2, "V1", "YLDS")
-# soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
-# soy_ylds2[, XNAME := NULL]
-# 
-# ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
-# soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
-# setnames(soy_ylds2, "V1", "mid")
-# 
-# ylds <- list(ylds$maize, soy_ylds2)
-# names(ylds) <- c("maize", "soybean")
-
 ```
@@
 ```{r, eval = FALSE}
-# pts <- IDr_ss[, .(x, y)]
-# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
-# coordinates(pts) <- ~x+y
 soilid <- extract(soilgrid, zmetf)  # grid id corresponding to point
-# soilid <- extract(soilgrid, pts)  # grid id corresponding to point
 profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
@@
 xy <- unique(ylds$maize[, .(x, y)])
-# xy[, ind := 1:.N]
-# cols <- c("x", "y")
-# xy[, (cols) := lapply(.SD, function(x) round(x, 5)), .SDcols = cols]
-# coordinates(xy) <- ~x+y
 
@@
 
-# Checking merges (disabled at present)
-# set.seed(234)
-# samp <- sample(1:nrow(regdat$maize), size = 20)
-# for(ii in samp) {  # ii <- samp[5]
-#   a <- regdat$maize[ii, ]; b <- regdat$soybean[ii, ]
-#   d <- ID[WTH == a$WTH & zoneid == a$zoneid & SOL == a$SOL]
-#   e <- ylds$maize[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
-#                   YR == b$YR][1, ]
-#   f <- ylds$soybean[WTH == b$WTH & zoneid == b$zoneid & SOL == b$SOL[1] & 
-#                     YR == b$YR][1, ]
-#   # SOIL climate match
-#   for(i in list(b, d, e, f)) {
-#     # i <- d
-#     for(j in c("zoneid", "WTH", "SOL")) {
-#       # print(j)
-#       if(a[[j]] != i[[j]]) print(paste("mismatch at", j, "row", ii))
-#     }
-#     for(j in c("x", "y")) {
-#       # print(j)
-#       if(round(a[[j]], 5) != round(i[[j]], 5)) {  # round because IDr wasn't
-#         print(paste("mismatch at", j, "row", ii))
-#       }
-#     }
-#   }
-#   for(i in c("YLDS", "YLDR")) { # i <- "YLDS"
-#     if(a[[i]] != e[[i]]) print(paste("maize mismatch at", i, "row", ii))
-#     if(b[[i]] != f[[i]]) print(paste("maize mismatch at", i, "row", ii))
-#   }
-# }
-
-# ylds[[1]][ID == 6968 & zoneid == 1]
-# regdat <- lapply(regdat, function(i) {
-#   setcolorder(i, c(5, 1:4, 6:ncol(i)))
-#   setnames(i, "mid", "midold")
-#   i <- cbind(data.table("mid" = 1:nrow(i)), i)
-#   i <- setkey(i, "mid")
-#   i
-# })
-# 
-# all(regdat$maize$SOL == regdat$soybean$SOL)
-# all(regdat$maize$x == regdat$soybean$x)
-# all(regdat$maize$y == regdat$soybean$y)
-# all(regdat$maize$OP == regdat$soybean$OP)
-# all(regdat$maize$SLOC == regdat$soybean$SLOC)
-
 regdt <- rbindlist(lapply(names(regdat), function(x) {

### 2016-02-11 Rmd/all-inputs.Rmd
Developed several flavors of GAM-derived soy and maize grids, and adjusted them to FAO mean projected yields.

--- 2016-01-20 Rmd/all-inputs.Rmd
@@
 p_cost <- full_path(p_proj, "external/input_devel/roads/cost")
+p_yield <- full_path(p_proj, "external/input_devel/yield/empirical")
+p_h20 <- full_path(p_proj, "external/input_devel/water")
+p_terr <- full_path(p_proj, "external/input_devel/terrain")
+p_carb <- full_path(p_proj, "external/input_devel/carbon")
+p_production <- full_path(p_proj, "external/data/production")
 
@@
 cropland <- raster(full_path(p_crop, fname = "cropland-1k.tif"))
-bd <- raster(full_path(p_bd, fname = "bdprior.tif"))
-vegr <- raster(full_path(p_bd, fname = "ZA-vegremain.tif"))
-cost <- raster(full_path(p_cost, fname = "zamcost.tif"))
+cons <- raster(full_path(p_bd, fname = "cons-priorities.tif"))
+bd <- brick(full_path(p_bd, fname = "ZA-vegremain.tif"))
+load(full_path(p_bd, fname = "ZA-vegremain.rda"))
+names(bd) <- vegnames
+carbon <- brick(full_path(p_carb, fname = "ZA-carbon.tif"))
+names(carbon) <- c("veg", "soil")
+cost <- raster(full_path(p_cost, fname = "ZA-cost.tif"))
+pas <- raster(full_path(p_bd, "ZA-pas.tif"))
+intact <- raster(full_path(p_bd, "intactness.tif"))
+# ynms <- dir(p_yield, pattern = "FAO.tif")  # FAO adjusted now
+# ynms <- sapply(c("S", "R", "I"), function(x) ynms[grep(x, ynms)])
+# ynms <- ynms[c(1, 3, 5, 2, 4, 6)]
+# ylds <- lapply(ynms, function(x) {
+#   b <- brick(full_path(p_yield, x))
+#   ylnms <- gsub("\\.tif", "\\.rda", x)
+#   load(full_path(p_yield, ylnms))
+#   names(b) <- layernms
+#   b
+# })
+# names(ylds) <- gsub("_10K|\\.tif|_FAO", "", ynms)
+
+crops <- c("maize", "soy")
+ynms <- dir(p_yield, pattern = "fao")  # FAO adjusted now
+ylds <- lapply(ynms, function(x) brick(full_path(p_yield, x)))
+names(ylds) <- crops
+ylds <- stack(list(ylds[[1]][[5]], ylds[[2]][[4]]))  # take the last two of each
+names(ylds) <- crops
+ylds <- round(ylds, 2)
+
+# mask materials
+remfrac <- raster(full_path(p_bd, fname = "ZA-remfrac.tif"))
+water <- raster(full_path(p_h20, fname = "water-1k.tif"))
+farmable <- raster(full_path(p_terr, fname = "farmable-1k.tif"))
 
 ```
+
+## Create farmable area
+
+```{r, eval=FALSE}
+# merge together the three farmland available masks
+avail <- remfrac - water
+avail[avail < 0] <- 0
+avail <- avail - (1 - farmable)
+avail[avail < 0] <- 0
+avail <- avail * (cons < 0.962)  # remove PAs entirely (1 if nps, 0.962 if GMAs)
+names(avail) <- "convertible"
+# plot(avail)
+```
+
+That creates the farmland availability mask that is needed.
+
+## Stack and mask all grids
+```{r, eval = FALSE}
+# deal with yields
+# potmaize <- ylds$maize_R  # just taking rainfed commercial potential to start
+# w <- matrix(nrow = 5, ncol = 5, rep(1, 5 * 5))  # have to filter b/c artifacts
+# potmaizef <- focal(potmaize[[1]], w = w, fun = mean, na.rm = TRUE)
+# potsoy <- ylds$soybean_R  # just taking rainfed commercial potential to start
+# potsoy[potsoy > 5] <- 5 #round(cellStats(potsoy, mean), 1)  # really high fix
+# potsoyf <- focal(potsoy[[1]], w = w, fun = mean, na.rm = TRUE)
+# # potsoy <- mask(potsoyf, potsoy)  # crop yield
+# 
+yield_fix <- function(x, w, na.rm = TRUE) {
+  r <- focal(x, w = w, fun = mean, na.rm = na.rm)
+  rmsk <- (!is.na(r) & is.na(x)) * r
+  x[is.na(x)] <- 0
+  rfix <- x + rmsk  # add back in edge bits
+}
+w <- matrix(nrow = 9, ncol = 9, rep(1, 81))
+mzf <- yield_fix(ylds[[1]], w)
+soyf <- yield_fix(ylds[[2]], w)
+ylds_s <- stack(sapply(list(mzf, soyf), function(x) {
+  x[is.na(x)] <- 0
+  x
+}))
+ylds_s <- raster::mask(ylds_s, mgrid)
+names(ylds_s) <- crops
+# plot(ylds_s[[1]])
+# potmaizef2 <- yield_fix(potmaizef[[1]], w)
+# plot(potmaizef2)
+# # first get rid of all NA area in all grids
+# # start with yield grids
+# # yldmask <- calc(stack(ylds), sum)
+# 
+# # Function to disaggregate and reproject yield rasters
+# yld_disagg <- function(x, msk, fact = 10) {
+#   r <- disaggregate(x, fact = 10)
+#   r2 <- resample(x, r)
+#   r3 <- projectRaster(r2, msk)
+#   r4 <- mask(r3, msk)
+#   r4
+# }
+# 
+# ylds_1k <- lapply(list(potmaizef2, potsoyf2), function(x) {
+#   yld_disagg(x, msk = mgrid)
+# })
+# ylds_s <- stack(ylds_1k) 
+# names(ylds_s) <- c("maize", "soy")
+# ylds_s <- round(ylds_s, 1)  
+# plot(ylds_s)
+
+# yldmask <- calc(stack(potmaizef2, potsoyf2), sum)
+# yldmask2 <- disaggregate(yldmask, fact = 10)
+# yldmask3 <- resample(yldmask, yldmask2)
+# yldmask4 <- projectRaster(yldmask3, mgrid)
+# yldmask5 <- mask(yldmask4, mgrid)
+# plot(yldmask5)
+
+# now with other
+carbonf <- stack(lapply(1:nlayers(carbon), function(x) {
+  r <- carbon[[x]]
+  r[is.na(r)] <- 0
+  r
+}))
+
+# pas and intactness
+intpa <- stack(intact, pas)
+
+mgridc <- crop(mgrid, avail)
+inlist <- list("convertible" = avail, "bd" = bd, "carbon" = carbonf, 
+               "cost" = cost, "cons" = cons, "cropland" = cropland, 
+               "intpa" = intpa, "p_yield" = ylds_s)
+inlist <- lapply(inlist, function(x) crop(x, mgridc))
+inlist2 <- lapply(inlist, function(x) calc(x, mean))
+inlist3 <- calc(stack(inlist2), mean)
+# inlist4 <- calc(stack(inlist3, crop(ylds_s, mgridc)), mean)
+
+# filter NA values
+# nafilt <- calc(stack(!is.na(inlist3), !is.na(crop(yldmask5, mgridc))), sum)
+nafilt <- !is.na(inlist3)
+# plot(nafilt)
+# plot(!is.na(calc(vegr, mean)))
+# plot(!is.na(crop(yldmask5, mgridc)), col = c("transparent", "red"), add = TRUE)
+msk <- nafilt == maxValue(nafilt)  # freq(nafilt)
+msk[msk == 0] <- NA
+# plot(msk)
+
+# updated mask
+mgridf <- raster::mask(mgridc, msk)
+valinds <- which(!is.na(values(mgridf)))
+outlist <- inlist[names(inlist) != "cropland"]
+
+# write out new data.tables
+### make sure convertible writes out column header name properly
+cnames <- "ZA"  
+onms <- names(outlist)
+onms[onms == "cons"] <- "cons-priorities"
+onms[onms == "p_yield"] <- "potential-yields"
+
+for(i in 1:length(outlist)) { # i <- 1
+  fnm <- full_path(p_production, paste0(cnames, "-", onms[i], ".csv"))
+  DT <- as.data.table(outlist[[i]], xy = FALSE)[valinds, ]
+  write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+}
+
+# write out mask raster
+base <- as.data.table(msk, xy = TRUE)
+setnames(base, old = names(msk), new = "val")
+base[, ind := 1:nrow(base)]
+base <- base[!is.na(val)][, val := NULL]
+fnm <- full_path(p_production, paste0(cnames, "-mask.csv"))
+write.table(base, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+# plot(dt_to_raster(base, mgridf@crs))
+
+# # data.table them
+# cnames <- "ZA"  
+# fnm <- full_path(p_production, paste0(cnames, "-potential-yields.csv"))
+# DT <- as.data.table(ylds_s, xy = FALSE)[valinds, ]
+# # DT[which(is.na(DT))]
+# write.table(DT, file = fnm, sep = ",", col.names = TRUE, row.names = FALSE)
+```
+
+# Copy over files to the correct directory
+```{r, eval}
+p_mod <- full_path(p_proj, "agroEcoTradeoff/external/data/ZA/")
+p_master <- "~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/"
+fnms <- dir(p_production)
+file.copy(dir(p_production, full.names = TRUE), full_path(p_mod, fnms),
+          overwrite = TRUE)
+file.copy(dir(p_production, full.names = TRUE), full_path(p_master, fnms),
+          overwrite = TRUE)
+
+```

### 2016-02-11 Rmd/empirical-yields.Rmd
Developed several flavors of GAM-derived soy and maize grids, and adjusted them to FAO mean projected yields.

--- 2016-02-08 Rmd/empirical-yields.Rmd
@@
 p_yield <- fp(p_root, "external/input_devel/yield")
+p_sgrids <- fp(p_root, "external/input_devel/soil") 
+p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
+                    "base_data/")
+p_cropland <- full_path(p_root, "external/input_devel/cropland")
+
+# Read in Zambia grid
+mgrid <- raster(full_path(p_basedat, "ZA-grid.tif"))
 
@@
 plot(prov)
-regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
-regdt[CROP == "maize" & YLDS > 8000, 
-      points(x, y, pch = 16, cex = 0.5, col = "red")]
-regdt[CROP == "maize" & YLDS > 8000, 
-      points(x, y, pch = 16, cex = 0.5, col = "red")]
-regdt[CROP == "soybean" & YLDS > 1700, 
-      points(x, y, pch = 16, cex = 0.5, col = "blue")]
+regdt[CROP == "maize", plot(DJFMP, YLDR)]
+regdt[CROP == "maize", plot(DJFMP, YLDS)]
+regdt[CROP == "soy", plot(DJFMP, YLDR)]
+regdt[CROP == "soy", plot(DJFMP, YLDS)]
+regdt[CROP == "maize", mean(YLDR)]
+regdt[CROP == "maize", mean(YLDS)]
+regdt[CROP == "soy", mean(YLDR)]
+regdt[CROP == "soy", mean(YLDS)]
 
-# Prune out yields > 8000 maize > 1700 soybean in subsistence case
-mzfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
-                .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
-soyfilt <- regdt[CROP == "maize" & YLDS > 8000, unique(.SD), 
-                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
-regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
+regnew <- copy(regdt)
+# regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
+# regdt[CROP == "maize" & YLDS > 15000, 
+#       points(x, y, pch = 16, cex = 0.5, col = "red")]
+# regdt[CROP == "soy", plot(MAP, YLDR)]
+# regdt[CROP == "soy" & YLDS > 1700, 
+#       points(x, y, pch = 16, cex = 0.5, col = "blue")]
+
+# Prune out yields > 8000 maize > 1700 soy in subsistence case
+# soyfilt <- regdt[CROP == "soy" & YLDS > 1700, unique(.SD), 
+#                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
+# mzfilt <- regdt[CROP == "maize" & YLDS > 9000, unique(.SD), 
+#                 .SDcols = c("x", "y", "SOL", "WTH")]
+# # regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
+# regnew <- regdt[!x %in% mzfilt$x | !y %in% mzfilt$y]
 # regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]
@@
 # summary(lm(YLDR ~ poly(MSP, 2), data = regnew[CROP == "maize", ]))
-regnew[CROP == "maize", plot(MAP, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "soybean", plot(MAP, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "soybean", plot(MAP, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "soybean", plot(MSP, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "soybean", plot(MSP, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(FP, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(MSP, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(MSP, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sloc, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sloc, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slcl, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slcl, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slsi, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slsi, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sbdm, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(sbdm, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(slhw, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
-regnew[CROP == "maize", plot(slhw, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
-regnew[CROP == "maize", plot(GDD, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(KDD, YLDS, pch = 16, cex = 0.5)]
-# regnew[CROP == "maize", plot(log(KDD), YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(NKJ, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(NKF, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(ATMX, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(STMX, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(STMX, YLDR, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(SSRT, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(SSRM, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(ASRT, YLDS, pch = 16, cex = 0.5)]
-regnew[CROP == "maize", plot(ASRM, YLDS, pch = 16, cex = 0.5)]
-
-regnew[, round(cor(.SD), 2), .SDcols = names(regnew)[-c(1:13)]]
-
-summary(lm(YLDS ~ poly(sbdm, 2) + poly(slhw, 2), 
-           data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ poly(slhw, 2), data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ poly(sbdm, 2), data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ poly(KDD, 2), data = regnew[CROP == "maize", ]))
-summary(lm(YLDR ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ KDD, data = regnew[CROP == "maize", ]))
-summary(lm(YLDS ~ poly(GDD, 2), data = regnew[CROP == "maize", ]))
-
-# p <- "sbdm"; d <- "YLDS"
-# regnew[CROP == "maize", {
-#   lm1 <- lm(get(d) ~ get(p))
-#   lm2 <- lm(get(d) ~ poly(get(p), 2))
-#   plot(get(p), get(d), pch = 16, cex = 0.1)
-#   abline(lm1)
-#   dum <- data.frame(floor(min(get(p))):ceiling(max(get(p))))
-#   colnames(dum) <- p
-#   lines(dum[, 1], predict(lm2, newdata = dum), col = "red")
-#   summary(lm1)
-#   # summary(lm2)
-# }]
 
@@
 }
-# regnew[CROP == "maize", reg_plotter(YLDR, GDD)]  # 
-regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.045
-regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
-regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.012
-# regnew[CROP == "maize", reg_plotter(YLDS, JP - mean(JP))]  # P 0.012
-regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.005
-regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.003
 
-regnew[CROP == "maize", reg_plotter(YLDS, slhw)]  # P 0.17
-regnew[CROP == "maize", reg_plotter(YLDS, sloc)]  # P 0.266
-# regnew[CROP == "maize", reg_plotter(YLDR, sloc)]  # P 
-regnew[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.185
-regnew[CROP == "maize", reg_plotter(YLDS, sbdm)]  # 0.181
-# mtax pairings
-regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # P 0.07
-regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.056 
-# NK
-regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.054
-regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.027 
-# GDD/KDD
-regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.04
-regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.027
+# Rainfall
+regnew[CROP == "maize", reg_plotter(YLDS, MAP)] ##
+regnew[CROP == "maize", reg_plotter(YLDS, MSP)] #
+regnew[CROP == "maize", reg_plotter(YLDS, DJFMP)] ###
+regnew[CROP == "maize", reg_plotter(YLDS, JFMP)] #
+regnew[CROP == "maize", reg_plotter(YLDS, FP)]  
+regnew[CROP == "maize", reg_plotter(YLDS, OP)]
+
+# soil
+regnew[CROP == "maize", reg_plotter(YLDS, SLHW)]  ###
+regnew[CROP == "maize", reg_plotter(YLDS, SLOC)]  # 
+regnew[CROP == "maize", reg_plotter(YLDS, SLCL)]  ## 
+regnew[CROP == "maize", reg_plotter(YLDS, SBDM)]  ##
+
+# TEMP
+regnew[CROP == "maize", reg_plotter(YLDS, TMU)]  ## 
+regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # 
+regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  #
+regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # 
+regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  
+regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  ##
+regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  ##
+
 # srad
-regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.012 
-regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # Same, picked up wrong one
-regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.0288 
-regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.0299
-
-# best predictors look to be
-# MAP, SLOC (but polynomial relationship odd), SLHW, slcl, sbdm, all good
-# STMX (linear KDD as a possible alternate), GDD, ASRM
-
-#
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4279
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4199
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(sloc, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4044 # SLOC doesn't give as much value, NS on linear
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4293  # Doesn't add much
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4293  # Oct precip add a tiny bit, significant
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4312  # Jan precip adds more R2, but makes MAP linear NS
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4374  # Feb precip even more R2, but makes MAP linear NS
-
-regnew[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
-  summary(lm1)
-  plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
-
-# regnew[CROP == "maize", {
-#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-#               poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
-#              family = gaussian(link = "log"), method = "ML")
-#   summary(gam1)
-#   # plot(YLDS, predict(lm1))
-# }]  # 0.4301  # Oct precip add a tiny bit, significant
+regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  
+regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  
+regnew[CROP == "maize", reg_plotter(YLDS, SSRMI)]  ##
+regnew[CROP == "maize", reg_plotter(YLDS, SSRMII)]  # 
+regnew[CROP == "maize", reg_plotter(YLDS, SSRMIII)] #
 
@@
 # regred[, c("Y10S", "Y90S") := as.integer(0)]
-
-regmu <- regnew[CROP == "maize", lapply(.SD, mean), 
-                by = list(CROP, WTH, Zone, SOL), 
-                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
+# regred[CROP == "maize" & SOL == "HC_GEN0010" & WTH == "KALA", mean(GDD)]
+regmu <- regred[CROP == "maize", lapply(.SD, mean), 
+                by = list(CROP, x, y, WTH, SOL), 
+                .SDcols = names(regred)[c(8:ncol(regred))]]
 regmu[, YST := "mu"]
@@
 
-a <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
-            SOL == "WI_RGTZ054", YLDS]
-b <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
-            SOL == "WI_RGTZ054", mid]
-qs(a, b, 0.9)
-sort(a)
+# a <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", YLDS]
+# b <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", mid]
+# qs(a, b, 0.9)
+# sort(a)
 
@@
 
-regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.071
-regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
-regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.020
-regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.017
-regred2[CROP == "maize",  summary(lm(YLDS ~ poly(FP, 2)))]
-regred2[CROP == "maize", reg_plotter(YLDS, 100 - slcl + slsi)]  # P 0.102
-regred2[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.1879
-regred2[CROP == "maize", reg_plotter(YLDS, slsi)]  # P 0.03
+# Fit gams
+regred2[, round(cor(.SD), 2), 
+        .SDcols = c("MAP", "MSP", "DJFMP", "JFMP", "JP", "FP", "OP")]  # 0.7
+#        MAP  MSP DJFMP  JFMP    JP    FP    OP
+# MAP   1.00 0.91  0.91  0.88  0.59  0.57  0.23
+# MSP   0.91 1.00  0.90  0.83  0.71  0.63  0.06
+# DJFMP 0.91 0.90  1.00  0.96  0.73  0.66 -0.02
+# JFMP  0.88 0.83  0.96  1.00  0.73  0.73 -0.03
+# JP    0.59 0.71  0.73  0.73  1.00  0.31 -0.07
+# FP    0.57 0.63  0.66  0.73  0.31  1.00 -0.09
+# OP    0.23 0.06 -0.02 -0.03 -0.07 -0.09  1.00
+regred2[, round(cor(.SD), 2),  
+        .SDcols = c("TMU", "GDD", "KDD", "NKF", "NKJ", "STMX", "ATMX")]
+#       TMU  GDD  KDD  NKF  NKJ STMX ATMX
+# TMU  1.00 0.87 0.45 0.80 0.87 0.95 0.90
+# GDD  0.87 1.00 0.27 0.65 0.71 0.80 0.89
+# KDD  0.45 0.27 1.00 0.31 0.48 0.42 0.34
+# NKF  0.80 0.65 0.31 1.00 0.87 0.81 0.78
+# NKJ  0.87 0.71 0.48 0.87 1.00 0.86 0.80
+# STMX 0.95 0.80 0.42 0.81 0.86 1.00 0.90
+# ATMX 0.90 0.89 0.34 0.78 0.80 0.90 1.00
+regred2[, round(cor(.SD), 2),  
+        .SDcols = c("TMU", "SSRMI", "SSRMIII", "MAP", "DJFMP", "JFMP", "GDD")]
+#           TMU SSRMI SSRMIII   MAP DJFMP  JFMP   GDD
+# TMU      1.00  0.70    0.69 -0.65 -0.62 -0.60  0.87
+# SSRMI    0.70  1.00    0.91 -0.69 -0.60 -0.58  0.40
+# SSRMIII  0.69  0.91    1.00 -0.64 -0.59 -0.61  0.45
+# MAP     -0.65 -0.69   -0.64  1.00  0.91  0.88 -0.45
+# DJFMP   -0.62 -0.60   -0.59  0.91  1.00  0.96 -0.47
+# JFMP    -0.60 -0.58   -0.61  0.88  0.96  1.00 -0.43
+# GDD      0.87  0.40    0.45 -0.45 -0.47 -0.43  1.00
 
-regred2[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
+regred2[, round(cor(.SD), 2), 
+        .SDcols = c("SLOC", "SBDM", "SLCL", "SLHW", "SLSI")]  
+#       SLOC  SBDM  SLCL  SLHW  SLSI
+# SLOC  1.00 -0.67  0.53 -0.30  0.03
+# SBDM -0.67  1.00 -0.77  0.34 -0.26
+# SLCL  0.53 -0.77  1.00 -0.64 -0.09
+# SLHW -0.30  0.34 -0.64  1.00  0.36
+# SLSI  0.03 -0.26 -0.09  0.36  1.00
 
-regred2[CROP == "maize" & YST == "mu", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
+gam1 <- gam(YLDS ~ s(MAP, k = 3)  + s(TMU, k = 3) + s(SLHW, k = 3) + 
+              s(SLCL, k = 3) + s(SLOC, k = 3) + 
+              s(SSRMI, k = 3), family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
 
-regred2[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(slcl, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
+gam2 <- gam(YLDS ~ s(MAP, k = 3) + s(TMU, k = 3) + s(SLHW, k = 3) + 
+              s(SLCL, k = 3) + s(SLOC, k = 3), 
+            family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
 
-regred2[CROP == "maize", {
-  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(sloc, 2) +
-              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
-  summary(lm1)
-  # plot(YLDS, predict(lm1))
-}]  # 0.4301  # Oct precip add a tiny bit, significant
-
-regred2[CROP == "maize", {
-  gam1 <- gam(YLDS ~ s(MAP) + s(JP) + s(slhw) + s(slcl) + (sbdm) + s(GDD) +
-                KDD + s(ASRM), 
-            family = gaussian(link = "log"), method = "ML")
-}]
-
-gam1 <- gam(YLDS ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
-              s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
-              s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
+gam3 <- gam(YLDS ~ poly(MAP, 2) + s(TMU, k = 3) + s(SLHW, k = 3) + 
+              s(SLCL, k = 3) + s(SLOC, k = 3), 
+            family = gaussian(link = "log"), method = "ML", 
             data = regred2[CROP == "maize"])
-gam2 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2) + s(SLHW, k = 3) + 
+              s(SLCL, k = 3) + s(SLOC, k = 3), 
+            family = gaussian(link = "log"), method = "ML", 
             data = regred2[CROP == "maize"])
-gam3 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2) + s(x, y), 
+gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2) + s(SLHW, k = 3) + 
+              s(SLCL, k = 3), 
+            family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam6 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + s(SLHW, k = 3) + 
+              s(SLCL, k = 3), 
+            family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam7 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+              s(SLCL, k = 3), 
+            family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam8 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2), 
+            family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam9 <- gam(YLDS ~ poly(MSP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2), 
+            family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam10 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2), family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam11 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam12 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+              poly(SLCL, 2) + poly(SLOC, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam13 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam14 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "ML", 
+            data = regnew[CROP == "maize"])
+gam15 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "ML", 
+            data = regred2[YST == "mu"])
+gam16 <- gam(YLDS ~ s(JFMP, k = 3)  + s(TMU, k = 3) + s(SLHW, k = 3) + 
+               s(SLCL, k = 3) + s(SSRMI, k = 3), 
+             family = gaussian(link = "log"), method = "ML", 
+             data = regred2[CROP == "maize"])
+gam17 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "ML", 
+            data = regred2[CROP == "maize"])
+gam18 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "ML", 
+            data = regnew[CROP == "maize"])
+gam19 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
-gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
-              s(x, y), family = gaussian(link = "log"), method = "REML", 
-            data = regred2[CROP == "maize"])
-gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD, 
+gam20 <- gam(YLDS ~ poly(DJFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", 
-            data = regred2[CROP == "maize"])
-gam6 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(GDD, 2) + KDD, 
+            data = regred2)
+gam21 <- gam(YLDS ~ poly(DJFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", 
-            data = regred2[CROP == "maize"])
-gam7 <- gam(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(GDD, 2) + KDD, 
+            data = regnew[CROP == "maize"])
+gam22 <- gam(YLDS ~ poly(MAP, 2) + poly(GDD, 2) + KDD + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", 
-            data = regred2[CROP == "maize"])
-gam8 <- gam(YLDS ~ poly(MAP, 2) + slhw +  poly(slcl, 2) + poly(GDD, 2) + KDD, 
+            data = regnew[CROP == "maize"])
+gam23 <- gam(YLDS ~ poly(MAP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + 
+               poly(SLHW, 2) + poly(SLCL, 2) + s(x, y), 
              family = gaussian(link = "log"), method = "REML", 
-            data = regred2[CROP == "maize"])
-gam9 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-              poly(100 - (slcl + slsi), 2) + poly(GDD, 2) + KDD, 
-            family = gaussian(link = "log"), method = "REML", 
-            data = regred2[CROP == "maize"])
-gam10 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(sloc, 2) + poly(GDD, 2) + KDD, 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[CROP == "maize"])
-gam11 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               sloc + poly(GDD, 2) + KDD, 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[CROP == "maize"])
-gam12 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               sloc + poly(GDD, 2) + KDD + s(x, y), 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[CROP == "maize"])
+            data = regnew[CROP == "maize"])
 
+summary(gam1)
+plot.gam(gam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam2)
+plot.gam(gam2, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(gam3)
-plot.gam(gam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot.gam(gam3, residuals = FALSE, all.terms = TRUE, pages = 1)
 summary(gam4)
-plot.gam(gam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam5)
-plot.gam(gam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam6)
-plot.gam(gam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot.gam(gam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam5)  #
+plot.gam(gam5, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam6)  #+  MAP -> DJMF
+plot.gam(gam6, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(gam7)
-plot.gam(gam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot.gam(gam7, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(gam8)
-plot.gam(gam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam9)
-plot.gam(gam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam10)
-plot.gam(gam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam11)  # This one or
-plot.gam(gam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(gam12)  # of this one, with smooth x, y terms
-plot.gam(gam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot.gam(gam8, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam9)  #- DJFM -> MSP
+plot.gam(gam9, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam10) #+ MSP -> JFM ~ DJMF
+plot.gam(gam10, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam11)  
+plot.gam(gam11, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam12)  #<-> SLOC
+plot.gam(gam12, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam13)  #+ SLOC out SSRMI in
+plot.gam(gam13, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam14)  #+ SLOC out SSRMI in, full dataset
+plot.gam(gam14, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam15)  #+ SLOC out SSRMI in, means only
+plot.gam(gam15, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam16)  #- Smooth version of 13, explains less 
+plot.gam(gam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam17)  # GDD Subs in exactly for TMU, less correlated with SSRMI
+plot.gam(gam17, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam18)  # 18 on full dataset
+plot.gam(gam18, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam19)  #+ DJFMP + TMU 
+plot.gam(gam19, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam20)  #+ DJFMP GDD
+plot.gam(gam20, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam21)  #+ 20 with full dataset
+plot.gam(gam21, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam22)  #+ MAP and KDD full dataset
+plot.gam(gam22, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(gam23)  #+ MAP full dataset
+plot.gam(gam23, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 
 par(mfrow = c(2, 2))
-plot(regred2[CROP == "maize", YLDS], predict(gam11, type = "response"))
-plot(regred2[CROP == "maize", YLDS], predict(gam12, type = "response"))
-plot(regred2[CROP == "maize", YLDS], predict(gam4, type = "response"))
+regred2[CROP == "maize", plot(YLDS, predict(gam13, type = "response"))]
+lines(c(2000, 12000), c(2000, 12000), col = "red")
+regnew[CROP == "maize", plot(YLDS, predict(gam14, type = "response"))]
+lines(c(2000, 12000), c(2000, 12000), col = "red")
+regred2[YST == "mu", plot(YLDS, predict(gam15, type = "response"))]
+lines(c(2000, 12000), c(2000, 12000), col = "red")
+regred2[, plot(YLDS, predict(gam20, type = "response"))]
+lines(c(2000, 12000), c(2000, 12000), col = "red")
+regnew[CROP == "maize", plot(YLDS, predict(gam21, type = "response"))]
+lines(c(2000, 12000), c(2000, 12000), col = "red")
 
-# Reduce down to model for mean, 10th, 90th)
-gam12mu <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + s(x, y), 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[YST == "mu"])
-summary(gam12mu)
-plot.gam(gam12mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-gam1210 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + s(x, y), 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[YST == "10"])
-summary(gam1210)
-plot.gam(gam1210, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-gam1290 <- gam(YLDS ~ poly(MAP, 2) + slhw + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + s(x, y), 
-             family = gaussian(link = "log"), method = "REML", 
-             data = regred2[YST == "90"])
-summary(gam1290)
-plot.gam(gam1290, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+# Predict on 17 and 18
 
@@
 
-### Soybean
+### soy
 
 ```{r, eval = FALSE}
-regreds <- copy(regnew[CROP == "soybean"])
+regreds <- copy(regnew[CROP == "soy"])
 regreds[, mid2 := 1:.N]
 
-# regred[, c("Y10S", "Y90S") := as.integer(0)]
-
-regmus <- regreds[, lapply(.SD, mean), by = list(CROP, WTH, Zone, SOL), 
-                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
+regmus <- regreds[, lapply(.SD, mean), by = list(CROP, x, y, WTH, SOL), 
+                  .SDcols = names(regred)[c(8:ncol(regred))]]
 regmus[, YST := "mu"]
@@
 regreds2 <- rbind(regmus, y1090s[, cnames, with = FALSE])
-regreds2[, slcf := 100 - (slcl + slsi)]
 
@@
 for(i in c("10", "mu", "90")) {
-  regreds2[CROP == "soybean" & YST == i, hist(YLDR)]
+  regreds2[YST == i, hist(YLDR)]
 }
-regreds2[CROP == "soybean", mean(YLDR), by = YST]
+regreds2[ , mean(YLDR), by = YST]
 
-regreds2[, reg_plotter(YLDR, MAP)]  # L 0.065
-regreds2[, reg_plotter(YLDR, MSP)]  # L 0.1105
-regreds2[, reg_plotter(YLDR, JP)]  # L 0.08
-regreds2[, reg_plotter(YLDR, FP)]  # P 0.016
-regreds2[, reg_plotter(YLDR, 100 - slcl + slsi)]  # P 0.12
-regreds2[, reg_plotter(YLDR, slcl)]  # P 0.31
-regreds2[, reg_plotter(YLDR, slsi)]  # P 0.003
-regreds2[, reg_plotter(YLDR, sloc)]  # P 0.31
-regreds2[, reg_plotter(YLDR, slhw)]  # P 0.27
-regreds2[, reg_plotter(YLDR, GDD)]  # P 0.006
-regreds2[, reg_plotter(YLDR, KDD)]  # P 0.006
-regreds2[, reg_plotter(YLDR, STMX)]  # L 0.029
-regreds2[, reg_plotter(YLDR, NKF)]  # L 0.032
-regreds2[, reg_plotter(YLDR, NKJ)]  # P 0.029
-regreds2[, reg_plotter(YLDR, SSRT)]  # L 0.023
-regreds2[, reg_plotter(YLDR, ASRM)]  # L 0.055
+# Rainfall
+regnew[CROP == "soy", reg_plotter(YLDS, MAP)] ###
+regnew[CROP == "soy", reg_plotter(YLDS, MSP)] ##
+regnew[CROP == "soy", reg_plotter(YLDS, DJFMP)] ###
+regnew[CROP == "soy", reg_plotter(YLDS, JFMP)] ###
+regnew[CROP == "soy", reg_plotter(YLDS, JP)]   
+regnew[CROP == "soy", reg_plotter(YLDS, FP)] #  
+regnew[CROP == "soy", reg_plotter(YLDS, OP)]
 
-sgam1 <- gam(YLDR ~ s(MAP, k = 3) + s(JP, k = 3) + s(slhw, k = 3) + 
-               s(slcl, k = 3) + s(sbdm, k = 3) + s(GDD, k = 3) + KDD + 
-               s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
+# soil
+regnew[CROP == "soy", reg_plotter(YLDS, SLHW)]  ###
+regnew[CROP == "soy", reg_plotter(YLDS, SLOC)]  ## 
+regnew[CROP == "soy", reg_plotter(YLDS, SLCL)]  ### 
+regnew[CROP == "soy", reg_plotter(YLDS, SBDM)]  ##
+
+# TEMP
+regnew[CROP == "soy", reg_plotter(YLDS, TMU)]  ### 
+regnew[CROP == "soy", reg_plotter(YLDS, STMX)]  ### 
+regnew[CROP == "soy", reg_plotter(YLDS, ATMX)]  ##
+regnew[CROP == "soy", reg_plotter(YLDS, GDD)]  # 
+regnew[CROP == "soy", reg_plotter(YLDS, KDD)]  
+regnew[CROP == "soy", reg_plotter(YLDS, NKJ)]  #L
+regnew[CROP == "soy", reg_plotter(YLDS, NKF)]  ##
+
+# srad
+regnew[CROP == "soy", reg_plotter(YLDS, ASRT)]  ##
+regnew[CROP == "soy", reg_plotter(YLDS, ASRM)]  ##
+regnew[CROP == "soy", reg_plotter(YLDS, SSRMI)]  ## L
+regnew[CROP == "soy", reg_plotter(YLDS, SSRMII)]  ## L 
+regnew[CROP == "soy", reg_plotter(YLDS, SSRMIII)] ### L
+
+sgam1 <- gam(YLDS ~ s(MAP, k = 3) + s(TMU, k = 3)  + s(SLHW, k = 3) + 
+               s(SLCL, k = 3) + s(SLOC, k = 3) + s(SSRMIII, k = 3), 
+             family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-
-sgam2 <- gam(YLDR ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+sgam2 <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
+             family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam3 <- gam(YLDR ~ poly(MAP, 2) + poly(FP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+sgam3 <- gam(YLDS ~ poly(DJFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
+             family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam4 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sbdm, 2) + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+sgam4 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2), 
+             family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam5 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(sloc, 2) + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+sgam5 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SLOC, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam6 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + sloc + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+sgam6 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2)  + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam7 <- gam(YLDR ~ poly(MAP, 2) + poly(OP, 2) + poly(slhw, 2) + 
-              poly(slcl, 2) + poly(GDD, 2) + KDD + 
-              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
+sgam7 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
             data = regreds2)
-sgam8 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + poly(ASRM, 2), 
-             family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam9 <- gam(YLDR ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(ASRM, 2), 
-             family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam10 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(ASRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam11 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(sbdm, 2) + 
-               poly(GDD, 2) + poly(ASRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam12 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam13 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2) + sloc +  
-               poly(GDD, 2) + poly(SSRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam14 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam15 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + KDD + poly(SSRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam16 <- gam(YLDR ~ poly(MSP, 2) + poly(OP, 2) + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam17 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam18 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + NKJ + poly(SSRM, 2) + s(x, y), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
-sgam19 <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-               poly(GDD, 2) + NKF + poly(SSRM, 2) + s(x, y), 
-              family = gaussian(link = "log"), method = "REML", data = regreds2)
+sgam8 <- gam(YLDS ~ poly(JFMP, 2) + poly(STMX, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam9 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKJ, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam10 <- gam(YLDS ~ poly(JFMP, 2) + poly(NKF, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam11 <- gam(YLDS ~ poly(JFMP, 2) + poly(ATMX, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam12 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam13 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + KDD + poly(SLHW, 2) + 
+               poly(SLCL, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam14 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regreds2)
+sgam15 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regnew[CROP == "soy"])
+sgam16 <- gam(YLDS ~ poly(JFMP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+               poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+             family = gaussian(link = "log"), method = "REML", 
+            data = regnew[CROP == "soy"])
 
 summary(sgam1)
-plot.gam(sgam1, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam2)
-plot.gam(sgam2, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam3)
-plot.gam(sgam3, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam4)
-plot.gam(sgam4, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam5)
-plot.gam(sgam5, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam6)
-plot.gam(sgam6, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam7)
-plot.gam(sgam7, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam8)
-plot.gam(sgam8, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam9)  # MSP stronger than MAP
-plot.gam(sgam9, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam10)  # MSP stronger than MAP
-plot.gam(sgam10, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam11)  # SBDM much weaker SLCL
-plot.gam(sgam11, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam12)  # SSRM a tad better than 
-plot.gam(sgam12, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam13)  # SLOC goes in the wrong direction
-plot.gam(sgam13, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam14)  # OP adds a bit
-plot.gam(sgam14, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam15)  # KDD NS
-plot.gam(sgam15, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam16)  # Polynomial term on October precip doesn't add much
-plot.gam(sgam16, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam17)  # xy term added to model 15, all remain signif, seems way to go
-plot.gam(sgam17, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam18)  # NKJ NS
-plot.gam(sgam18, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam19)  # NKF NS
-plot.gam(sgam19, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+plot.gam(sgam1, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam2)  #+ poly version
+plot.gam(sgam2, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam3) #+ DJFMP 
+plot.gam(sgam3, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam4) #+ Weakly higher for JFMP
+plot.gam(sgam4, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam5) #+ s(x,y)
+plot.gam(sgam5, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam6) #- dropping SLOC doesn't lose too much, wasn't sensible
+plot.gam(sgam6, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam7) # <-> swapping in GDD doesn't change much
+plot.gam(sgam7, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam8) #+ STMX 
+plot.gam(sgam8, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam9) #- NKJ loses a bit, but looks more sensible
+plot.gam(sgam9, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam10) #<-> NKF loses a bit, but wrong direction
+plot.gam(sgam10, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam11) #<-> Annual TMx, same deal
+plot.gam(sgam11, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam12) #<-> TMU still increasing, even with SSRMIII dropped
+plot.gam(sgam12, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam13) #<-> KDD added, going in wrong direction
+plot.gam(sgam13, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam14) #<-> same as 6
+plot.gam(sgam14, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam15) #<-> Full data version of 7 (with GDD), becomes significant
+plot.gam(sgam15, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam16) #<-> Full data version of 14 (Tmu)
+plot.gam(sgam16, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 
-# Now see about taking same model and fitting separately to 10th/90th/mean years
-sgam17mu <- gam(YLDR ~ poly(MSP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) + 
-                  poly(GDD, 2) + poly(SSRM, 2) + s(x, y), 
-                family = gaussian(link = "log"), method = "REML", 
-                data = regreds2[YST == "mu"])
-summary(sgam17mu)  # OP drops out, SSRM drops out
-plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-sgam17mu <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + 
-                  poly(slcl, 2) + poly(GDD, 2) + s(x, y), 
-                family = gaussian(link = "log"), method = "REML", 
-                data = regreds2[YST == "mu"])
-summary(sgam17mu)  # OP drops out, SSRM drops out
-plot.gam(sgam17mu, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-sgam1710 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
-                + poly(GDD, 2) + s(x, y), 
-                family = gaussian(link = "log"), method = "REML", 
-                data = regreds2[YST == "10"])
-summary(sgam1710)  # OP drops out, SSRM drops out
-plot.gam(sgam1710, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
-sgam1790 <- gam(YLDR ~ poly(MSP, 2) + poly(slhw, 2) + poly(slcl, 2)
-                + poly(GDD, 2) + s(x, y), 
-                family = gaussian(link = "log"), method = "REML", 
-                data = regreds2[YST == "90"])
-summary(sgam1790)  # OP drops out, SSRM drops out
-plot.gam(sgam1790, residuals = TRUE, se = TRUE, all.terms = TRUE, pages = 1)
+par(mfrow = c(2, 2))
+plot(regreds2[, YLDS], predict(sgam7, type = "response"))
+lines(c(0, 8000), c(0, 8000), col = "red")
+plot(regnew[CROP == "soy", YLDS], predict(sgam15, type = "response"))
+lines(c(0, 8000), c(0, 8000), col = "red")
+plot(regreds2[, YLDS], predict(sgam14, type = "response"))
+lines(c(0, 8000), c(0, 8000), col = "red")
+
+# gams to keep 
+save(gam13, gam14, gam20, gam21, gam23, sgam7, sgam14, sgam15, sgam16, 
+     file = fp(p_yield, "empirical/gams1.rda"))
 
@@
 
-### Save GAMs outputs
+## Load in grid predictors
 ```{r, eval = FALSE}
-save(gam12, gam12mu, gam1210, gam1290, sgam17, sgam17mu, sgam1710, sgam1790, 
-     file = fp(p_yield, "gams.rda"))
+soilgrids <- brick(fp(p_sgrids, "soilgrids.tif"))
+climvars <- c("rftot", "djfmtot", "^jfmtot", "gdd\\.", "tmu", "srmu_gs1", 
+              "srmu_gs3")
+climgrids <- sapply(climvars, function(x) {  # x <- climvars[2]
+  print(x)
+  brick(dir(p_clim, full.names = TRUE)[grep(x, dir(p_clim))])
+})
+climgrids_mu <- sapply(climgrids, function(x) calc(x, mean))
+climst <- stack(climgrids_mu)
+names(climst) <- c("MAP", "DJFMP", "JFMP", "GDD", "TMU", "SSRMI", "SSRMIII")
+
+# disaggregate down to 1 km
+climstbase <- disaggregate(climst[[1]], fact = 30)
+climststep <- disaggregate(climst, fact = 20)
+climst1km <- resample(climst, climstbase)
+
+# need to warp soilgrids back to that also
+ext <- bbox(climst1km)[1:4]  # extent   
+onm <- fp(p_sgrids, "soilgridsgcs.tif")
+gdalwarp(srcfile = soilgrids@file@name, t_srs = projection(climst1km), 
+         dstfile = onm, r = "average", ot = "Float32", te = ext, 
+         srcnodata = -32768, dstnodata = -99, tr = res(climst1km), 
+         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
+soilgcs <- brick(onm)
+names(soilgcs) <- c("SLOC", "SLCL", "SLHW")
+plot(soilgcs)
+
+# need also a grid of x,y values
+x <- soilgcs[[1]]
+x[] <- xFromCell(x, 1:ncell(x))
+y <- soilgcs[[1]]
+y[] <- yFromCell(y, 1:ncell(y))
+plot(y)
+xy <- stack(x, y)
+names(xy) <- c("x", "y")
+predst <- stack(xy, climst1km, soilgcs)
+# plot(predst[["TMU"]])
 
 ```
+
+## Predicted yields surfaces
+```{r, eval = FALSE}
+mzred <- raster::predict(predst, gam20, type = "response")
+mzred2 <- raster::predict(predst, gam13, type = "response")
+mzfull <- raster::predict(predst, gam21, type = "response")
+mzfull2 <- raster::predict(predst, gam14, type = "response")
+mzfull3 <- raster::predict(predst, gam23, type = "response")
+soyred <- raster::predict(predst, sgam7, type = "response")
+soyred2 <- raster::predict(predst, sgam14, type = "response")
+soyfull <- raster::predict(predst, sgam15, type = "response")
+soyfull2 <- raster::predict(predst, sgam16, type = "response")
+
+# Not unplausible looking, maize not sure
+par(mfrow = c(2, 2), mar = c(1, 1, 1, 1))
+plot(mzred)
+plot(mzred2)
+plot(mzfull)
+plot(mzfull2)
+plot(mzfull3)
+plot(soyred)
+plot(soyred2)
+plot(soyfull)
+plot(soyfull2)
+plot(prov, add = TRUE)
+
+# prov@proj4string <- soyfull@crs
+# soymsk <- crop(soyfull, prov)
+# soymsk2 <- crop(soyfull2, prov)
+# soymsk3 <- crop(soyfull3, prov)
+
+# Save bricks and models
+yldrs <- brick(stack(list(mzred, mzred2, mzfull, mzfull2, mzfull3, soyred, 
+                          soyred2, soyfull, soyfull2)))
+writeRaster(yldrs, filename = fp(p_yield, "empirical/yield_surfaces2.tif"), 
+            overwrite = TRUE)
+yldrs <- brick(fp(p_yield, "empirical/yield_surfaces2.tif"))
+
+onm <- fp(p_yield, "empirical/yield_surfaces2_alb.tif")
+gdalwarp(srcfile = yldrs@file@name, t_srs = projection(mgrid), 
+         dstfile = onm, r = "average", ot = "Float32", te = bbox(mgrid)[1:4], 
+         srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), 
+         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
+yldralb <- brick(onm)
+names(yldralb) <- c("maizer1", "maizer2", "maizef1", "maizef2", "maizef3",
+                    "soyr1", "soyr2", "soyf1", "soyf2")
+
+yldralbmsk <- raster::mask(yldralb, mgrid)
+writeRaster(yldralbmsk, 
+            filename = fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))
+
+prov@proj4string <- raster()@crs
+png(fp(p_yield, "empirical/yield_maps.png"), height = 800, width = 800)
+par(mfrow = c(3, 3), mar = rep(2, 4), oma = c(0, 0, 0, 2))
+for(i in 1:9) {
+  plot(yldralbmsk[[i]], axes = FALSE, box = FALSE)
+  plot(spTransform(prov, mgrid@crs), add = TRUE)
+  mtext(names(yldralb)[[i]], side = 3, line = 0)
+}
+dev.off()
+
+```
+
+## Adjust to FAO yields
+
+Per `maize-soy-grid.Rmd`
+```{r, eval = FALSE}
+load(full_path(p_cropland, "currprod.rda"))
+load(full_path(p_cropland, "usda-yields.rda"))
+load(full_path(p_cropland, "fao-yields.rda"))
+cols <- names(usda)[4:9]
+usda_ylds <- usda[var == "prod", cols, with = FALSE] / 
+  usda[var == "harvarea", cols, with = FALSE]
+usda_mu <- round(rowMeans(usda_ylds), 1)
+
+muprod <- lapply(c("Maize", "Soybeans"), function(x) {
+  fao[Crop == x & var == "prod" & Year %in% c(2008:2015), sum(Value), 
+      by = Year]
+})
+names(muprod) <- c("maize", "soy")
+mupa <- lapply(c("Maize", "Soybeans"), function(x) {
+  fao[Crop == x & var == "pa" & Year %in% c(2008:2015), sum(Value), 
+      by = Year]
+})
+names(mupa) <- c("maize", "soy")
+fao_mu <- c("maize" = mean(muprod$maize[, V1] / mupa$maize[, V1]), 
+            "soy" = mean(muprod$soy[, V1] / mupa$soy[, V1]))
+
+# current yields, mean of USDA and FAO
+curr_mu <- round(colMeans(rbind(usda_mu, fao_mu)), 1)  # 2.3 maize, 2.0 soy
+
+# Calculate the mean yields for the country in each management scenario (across 
+# cultivar types)
+cropland <- raster(fp(p_cropland, "cropland-1k.tif"))
+plot(cropland)
+crops <- c("maize", "soy")
+ydiffs <- lapply(crops, function(x) { # x <- "maize"
+  rl <- yldralbmsk[[grep(x, names(yldralbmsk))]]
+  yr <- lapply(1:nlayers(rl), function(y) {  # y <- rl[[1]]
+    ysum <- cellStats(rl[[y]] * cropland, sum)
+    clsum <- cellStats(cropland, sum)
+    wmu <- ysum / clsum
+    rmu <- cellStats(rl[[y]], mean)
+    rvar <- rl[[y]] / rmu
+    list("ratio" = rvar, "mu" = rmu, "croplandmu" = wmu)
+  })
+  names(yr) <- names(rl)
+  yr
+})
+names(ydiffs) <- crops
+
+# Take the DSSAT means and figure out how much they differ relative to their 
+# overall mean across management types - use as management specific ratio
+dssat_mus <- sapply(ydiffs, function(x) sapply(x, function(y) y$mu))
+dssat_cmus <- sapply(ydiffs, function(x) sapply(x, function(y) y$croplandmu))
+# dssat_adj <- sweep(dssat_mus, MARGIN = 2,  colMeans(dssat_mus), FUN = `/`)
+
+# Bruinsma 2007-2050 growth
+bruinsma <- c(maize = 3224 / 1674, soy = 2124 / 1257)
+
+# adjusted DSSAT factors
+potmus <- curr_mu * c(bruinsma)  # 4.43 maize, 3.4 soy
+potmu_mz <- potmus[1] / (dssat_cmus[[1]] / 1000)  # maize on cropland mean
+potmu_soy <- potmus[2] / (dssat_mus[[2]] / 1000)  # soy on overall mean
+
+# adjust yields
+mzfao <- lapply(1:length(potmu_mz), function(x) {
+  (yldralbmsk[[x]] * potmu_mz[x]) / 1000
+})
+names(mzfao) <- names(yldralbmsk)[1:5]
+
+soyfao <- lapply(1:length(potmu_soy), function(x) {
+  (yldralbmsk[[x + 5]] * potmu_soy[x]) / 1000
+})
+names(soyfao) <- names(yldralbmsk)[6:9]
+
+# plot(soyfao[[1]])
+mzfao <- brick(stack(mzfao))
+soyfao <- brick(stack(soyfao))
+writeRaster(mzfao, filename = fp(p_yield, "empirical/maize-faoadj.tif"))
+writeRaster(soyfao, filename = fp(p_yield, "empirical/soy-faoadj.tif"))
+
+```

### 2016-02-11 Rmd/to-do.Rmd
Developed several flavors of GAM-derived soy and maize grids, and adjusted them to FAO mean projected yields.

@@
               "Exact method for variability and production zones, including inability to reach potential yields", 
-              "Run with improved soils maps", 
+              "Run with improved soils maps, using two stage regression", 
               "With ILUA database (and Nate's help)", 
@@
               "Upgrade all maps to SERVIR landcover, and rebase yield estimates on that", "Baccini vegetation carbon + ISRIC grids"),
-  "Doing" = c("D", "F*", "U", "U", "D", "D", "D", "L", "D", "F**", "F")
+  "Doing" = c("D", "F*", "F", "U", "F", "F", "F***", "L", "D", "F**", "F")
 )
@@
 **Just using total production estimates, which will not be mapped for now. These will be basis for estimating production targets.
+
+***This is what the model effectively does, and issue has been resolved.

### 2016-02-11 Rmd/yield-predictors.rmd
Developed several flavors of GAM-derived soy and maize grids, and adjusted them to FAO mean projected yields.

@@
   tmu <- (tmax[[ind]] + tmin[[ind]]) / 2 - 273.15
+  round(tmu, 2)
 }
-writeRaster(stack(tmu), filename = fp(p_clim, "tmu.tif"), overwrite = TRUE)
+tmu <- stack(tmu)
+
+# tmu for the DJFM
+# sindu <- unique(sind)[-1]
+djfmind2 <- djfmind[which(sind %in% sindu2)]
+for(i in sindu2) djfmind2[which(djfmind2 == i)] <- i - 31 
+sindu3 <- unique(djfmind2)[-1]
+tmuseas <- stack(foreach(i = 1:length(sindu3)) %dopar% { #  i <- 1
+  ind <- which(djfmind2 == sindu3[i])
+  calc(tmu[[ind]], mean)
+})
+writeRaster(tmuseas, filename = fp(p_clim, "tmu.tif"), overwrite = TRUE)
 
@@
 gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gddf))
-writeRaster(gdd30, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)
+# writeRaster(gdd, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)
 
@@
 climnms <- c("rftot", "ndjftot", "^djfmtot", "^jfmtot", "jantot", "febtot",
-             "octtot", "kdd", "gdd\\.", "gdd30", "nkilljan", "nkillfeb",
+             "octtot", "tmu", "kdd", "gdd\\.", "gdd30", "nkilljan", "nkillfeb",
              "tmax_ano", "tmax_gs", "srtot_ano", "srmu_ano", "srtot_gs",
@@
 
-onms <- c("MAP", "MSP", "DJFMP", "JFMP", "JP", "FP", "OP", "KDD", "GDD", 
+onms <- c("MAP", "MSP", "DJFMP", "JFMP", "JP", "FP", "OP", "TMU", "KDD", "GDD", 
           "GDDL", "NKJ", "NKF", "ATMX", "STMX", "ASRT", "ASRM", "SSRT", "SSRMI",

### 2016-02-14 Rmd/all-inputs.Rmd
Many updates. Up through new analysis.

@@
 names(ylds) <- crops
-ylds <- stack(list(ylds[[1]][[5]], ylds[[2]][[4]]))  # take the last two of each
+ylds <- stack(list(ylds[[1]][[1]], ylds[[2]][[1]]))  # take first two of each
 names(ylds) <- crops
@@
   rfix <- x + rmsk  # add back in edge bits
+  # rfix[rfix == 0] <- NA
+  # rfix
 }
@@
 soyf <- yield_fix(ylds[[2]], w)
-ylds_s <- stack(sapply(list(mzf, soyf), function(x) {
-  x[is.na(x)] <- 0
-  x
-}))
-ylds_s <- raster::mask(ylds_s, mgrid)
-names(ylds_s) <- crops
-# plot(ylds_s[[1]])
+yld_s <- stack(list(mzf, soyf))
+# ylds_s <- stack(sapply(list(mzf, soyf), function(x) {
+#   x[is.na(x)] <- 0
+#   x
+# }))
+yld_s <- raster::mask(yld_s, mgrid)
+names(yld_s) <- crops
+# plot(yld_s[[1]])
 # potmaizef2 <- yield_fix(potmaizef[[1]], w)
@@
 intpa <- stack(intact, pas)
-
 mgridc <- crop(mgrid, avail)
@@
                "cost" = cost, "cons" = cons, "cropland" = cropland, 
-               "intpa" = intpa, "p_yield" = ylds_s)
+               "intpa" = intpa, "p_yield" = yld_s)
 inlist <- lapply(inlist, function(x) crop(x, mgridc))
@@
 # plot(dt_to_raster(base, mgridf@crs))
+# plot(dt_to_raster(cbind(base[, .(x, y)], DT), mgridf@crs))
 
@@
 p_master <- "~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/"
+p_downloads <- "~/Dropbox/projects/agroEcoTradeoff/external/data/downloads/"
 fnms <- dir(p_production)
@@
           overwrite = TRUE)
+setwd(p_production)
+zip("demodat.zip", dir())
+file.copy("demodat.zip", fp(p_downloads, "demodat.zip"), overwrite = TRUE)
+file.remove("demodat.zip")
+setwd(p_proj)
 

### 2016-02-14 Rmd/biodiversity-layer.Rmd
Many updates. Up through new analysis.

@@
 s <- stack(lapply(fnms, function(x) raster(x)))
-sm <- mask(s, mgrid)
+sm <- raster::mask(s, mgrid)
 

### 2016-02-14 Rmd/cost-distance.Rmd
Many updates. Up through new analysis.

@@
 library(gdalUtils)
+setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
 p_root <- lmisc::set_base_path("zambiaToff")
@@
 setwd(p_roads)
-ztowns <- readOGR(dsn = "district_towns.shp", layer = "district_towns")
-ntowns <- readOGR("zambia-neighbor-cities.shp", 
-                  layer = "zambia-neighbor-cities")
-ntowns <- ntowns[ntowns$ES00POP > 10000, ]  # filter down to 10,000 or greater
-ztowns$cid <- 1:nrow(ztowns)
-ztowns@data <- ztowns@data[, "cid", drop = FALSE]  # reduce for binding
-ntowns$cid <- (max(ztowns$cid) + 1):(nrow(ntowns) + (max(ztowns$cid)))
-ntowns@data <- ntowns@data[, "cid", drop = FALSE]  # reduce for binding
-ntowns@proj4string <- ztowns@proj4string
-alltowns <- spTransform(rbind(ztowns, ntowns), CRSobj = projection(zfr))
+# ztowns <- readOGR(dsn = "district_towns.shp", layer = "district_towns")
+# ntowns <- readOGR("zambia-neighbor-cities.shp", 
+#                   layer = "zambia-neighbor-cities")
+# ntowns <- ntowns[ntowns$ES00POP > 10000, ]  # filter down to 10,000 or greater
+# ztowns$cid <- 1:nrow(ztowns)
+# ztowns@data <- ztowns@data[, "cid", drop = FALSE]  # reduce for binding
+# ntowns$cid <- (max(ztowns$cid) + 1):(nrow(ntowns) + (max(ztowns$cid)))
+# ntowns@data <- ntowns@data[, "cid", drop = FALSE]  # reduce for binding
+# ntowns@proj4string <- ztowns@proj4string
+# alltowns <- spTransform(rbind(ztowns, ntowns), CRSobj = projection(zfr))
 
 # write this out--needed for determining nodes in cost distance analysis
-writeOGR(alltowns, dsn = "alltowns.shp", layer = "alltowns", 
-         driver = "ESRI Shapefile", overwrite = TRUE)
+# writeOGR(alltowns, dsn = "alltowns.shp", layer = "alltowns", 
+#          driver = "ESRI Shapefile", overwrite = TRUE)
+
+alltowns <- readOGR("alltowns.shp", layer = "alltowns")
 
@@
 rdrcl <- c(1, 2, 4, 8)
-roads <- lapply(1:length(roads), function(x) {
+roads <- lapply(1:length(roadnms), function(x) {
   nm <- unname(roadnms)[x]
@@
 
+tmd2 <- spChFIDs(tmd, as.character(1:nrow(tmd)))
+pfr2 <- spChFIDs(pfr, as.character((nrow(tmd) + 1):(nrow(tmd) + nrow(pfr))))
+tmd2@data <- tmd2@data[, "roadid", drop = FALSE]
+pfr2@data <- pfr2@data[, "roadid", drop = FALSE]
+
+zroads <- rbind(tmd2, pfr2)
+zroads <- spTransform(zroads, mgrid@crs)
+# save(zroads, file = fp(p_roads, "tmdpfroads.rda"))
 # plot(zframe)
@@
 `ZA-cost.tif` - for masking, etc
+
+## Later fix
+
+Attempt to fix roads so they display properly in model output maps.
+Essentially combined OSM roads with Zambia roads, rasterized them to 30 m, added them, reduced all to values of 1, and then within ArcGIS (a bit of it earlier done in QGIS) I converted to polylines.
+Results used in main-analysis.Rmd.
+```{r, eval = FALSE}
+# roads
+setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
+p_root <- getwd()
+p_demand <- fp(p_root, "external/input_devel/demand/")
+p_mod <- fp(p_root, "agroEcoTradeoff")
+p_moddat <- fp(p_mod, "external/data/ZA")
+p_figures <- full_path(p_root, "paper/figures/")
+p_roads <- fp(p_root, "external/input_devel/roads")
+p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
+                   "rda_dump/Final Shape files RDA/")
+
+mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
+osmroads <- readOGR(fp(p_roads, "osmroads/zamroads.shp"), layer = "zamroads")
+
+osmroads$val <- 1
+tmd <- readOGR(fp(p_roads2, "TMDLinks_Merged.shp"), layer = "TMDLinks_Merged")
+pfr <- readOGR(fp(p_roads2, "PFRLinks.shp"), layer = "PFRLinks")
+pfrf <- pfr[-which(duplicated(pfr@data$RoadId)), ]
+tmdf <- tmd[-which(duplicated(tmd@data$RoadId)), ]
+tmdf <- spChFIDs(tmdf, as.character(1:nrow(tmdf)))
+pfrf <- spChFIDs(pfrf, as.character((nrow(tmdf) + 1):(nrow(tmdf) + nrow(pfrf))))
+tmdf@data <- tmdf@data[, "RoadId", drop = FALSE]
+pfrf@data <- pfrf@data[, "RoadId", drop = FALSE]
+roads <- rbind(tmdf, pfrf)
+roads <- spTransform(roads, pas@proj4string)
+osmroads <- spTransform(osmroads, pas@proj4string)
+roads$val <- 1
+plot(roads)
+plot(tmd)
+plot(pfr, add = TRUE, col = "red")
+plot(osmroads)
+plot(tmdf)
+writeOGR(roads, dsn = fp(p_roads, "osmroads/tmdpfrf.shp"), layer = "tmpdpfr", 
+         driver = "ESRI Shapefile", overwrite_layer = TRUE)
+writeOGR(osmroads, dsn = fp(p_roads, "osmroads/zamroadsalb.shp"), 
+         layer = "zamroadsalb", driver = "ESRI Shapefile")
+writeOGR(spTransform(roads, osmroads@proj4string),
+         dsn = fp(p_roads, "tmdpfrf_gcs.shp"), layer = "tmpdpfr", 
+         driver = "ESRI Shapefile")
+inm <- fp(p_roads, "osmroads/zamroadsalb.shp")
+onm <- fp(p_roads, "osmroads/zamroadsalb.tif")
+gdal_rasterize(src_datasource = inm, l = "zamroadsalb", ot = "Byte", 
+               a = "val", tr = c(30, 30), at = TRUE, co = "COMPRESS=LZW",
+               dst_filename = onm, te = bbox(mgrid)[1:4], of = "GTiff", 
+               verbose = TRUE)
+inm <- fp(p_roads, "osmroads/zamroadsalb.shp")
+onm <- fp(p_roads, "osmroads/zamroadsalb.tif")
+gdal_rasterize(src_datasource = inm, l = "zamroadsalb", ot = "Byte", 
+               a = "val", tr = c(30, 30), at = TRUE, co = "COMPRESS=LZW",
+               dst_filename = onm, te = bbox(mgrid)[1:4], of = "GTiff", 
+               verbose = TRUE)
+inm <- fp(p_roads, "osmroads/tmdpfrf.shp")
+onm <- fp(p_roads, "osmroads/tmdpfrf.tif")
+gdal_rasterize(src_datasource = inm, l = "tmdpfrf", ot = "Byte", 
+               a = "val", tr = c(30, 30), at = TRUE, co = "COMPRESS=LZW",
+               dst_filename = onm, te = bbox(mgrid)[1:4], of = "GTiff", 
+               verbose = TRUE)
+r <- raster(fp(p_roads, "osmroads/tmdpfr_plus_osm.tif"))
+r2 <- rast_math(r > 0, tmpdir = fp(p_roads, "osmroads"))
+
+```

### 2016-02-14 Rmd/empirical-yields.Rmd
Many updates. Up through new analysis.

@@
 library(mgcv)
+library(rworldxtra)
 
+setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
 p_root <- lmisc::set_base_path()
@@
 p_cropland <- full_path(p_root, "external/input_devel/cropland")
+p_figures <- full_path(p_root, "paper/figures/")
 
@@
             data = regreds2)
+# sgam14a <- gam(YLDS ~ poly(MAP, 2) + poly(TMU, 2) + poly(SLHW, 2) + 
+#                poly(SLCL, 2) + poly(SSRMIII, 2) + s(x, y), 
+#              family = gaussian(link = "log"), method = "REML", 
+#             data = regreds2[YST == "mu"])
+# summary(sgam14a)
+# plot.gam(sgam14a, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 sgam15 <- gam(YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + 
@@
 plot.gam(sgam13, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
-summary(sgam14) #<-> same as 6
-plot.gam(sgam14, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
+summary(sgam14a) #<-> same as 6
+plot.gam(sgam14a, residuals = FALSE, se = TRUE, all.terms = TRUE, pages = 1)
 summary(sgam15) #<-> Full data version of 7 (with GDD), becomes significant
@@
      file = fp(p_yield, "empirical/gams1.rda"))
+# load(fp(p_yield, "empirical/gams1.rda"))
+summary(gam13)
+summary(sgam14)
 
@@
             filename = fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))
+# yldralbmsk <- brick(fp(p_yield, "empirical/yield_surfaces2_albmsk.tif"))
+# names(yldralbmsk) <- c("maizer1", "maizer2", "maizef1", "maizef2", "maizef3",
+#                        "soyr1", "soyr2", "soyf1", "soyf2")
 
@@
 bruinsma <- c(maize = 3224 / 1674, soy = 2124 / 1257)
+fao <- fread(fp(p_yield, "empirical/zambia-maize-soy-yield.csv"))
+fao2005 <- fao[year > 2004,]
+yrates <- sapply(c("Maize", "Soybeans"), function(i) {
+  fao[crop == i, coef(lm(yield / yield[1] ~ year))[2]]
+})
+ymult <- 1 + yrates * (2050-2014)  # fao yield multipliers
+
+# take average of Bruinsma and FAO projection to account for maize
+ymults <- c("maize" = sum(ymult[1] * 0.25, bruinsma[1] * 0.75), bruinsma[2])
 
 # adjusted DSSAT factors
-potmus <- curr_mu * c(bruinsma)  # 4.43 maize, 3.4 soy
+potmus <- curr_mu * c(ymults)  # 4.43 maize, 3.4 soy
 potmu_mz <- potmus[1] / (dssat_cmus[[1]] / 1000)  # maize on cropland mean
@@
 soyfao <- brick(stack(soyfao))
-writeRaster(mzfao, filename = fp(p_yield, "empirical/maize-faoadj.tif"))
-writeRaster(soyfao, filename = fp(p_yield, "empirical/soy-faoadj.tif"))
+writeRaster(mzfao, filename = fp(p_yield, "empirical/maize-faoadj.tif"), 
+            overwrite = TRUE)
+writeRaster(soyfao, filename = fp(p_yield, "empirical/soy-faoadj.tif"), 
+            overwrite = TRUE)
+# mzfao <- brick(fp(p_yield, "empirical/maize-faoadj.tif"))
+# soyfao <- brick(fp(p_yield, "empirical/soy-faoadj.tif"))
 
+data("countriesHigh")
+zambia <- countriesHigh[countriesHigh@data$SOVEREIGNT == "Zambia", ]
+zambia <- spTransform(zambia, mzfao@crs)
+
+png(fp(p_figures, "yields.png"), width = 800, height = 400)
+par(mfrow = c(1, 2), mar = c(0, 0, 0, 4))
+plot(zambia, col = "grey80", border = "transparent")
+plot(mzfao[[1]], add = TRUE)
+plot(zambia, col = "grey80", border = "transparent")
+plot(soyfao[[1]], add = TRUE)
+dev.off()
 ```

### 2016-02-14 Rmd/main-analysis.Rmd
Many updates. Up through new analysis.

---
title: "Main Analyses"
author: "Lyndon Estes"
date: "February 11, 2016"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: yes
---

# Main Analyses

```{r, warning=FALSE, message=FALSE}
library(agroEcoTradeoff)
library(readxl)
library(gdalUtils)
library(gisutils)
setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
p_root <- getwd()
p_demand <- fp(p_root, "external/input_devel/demand/")
p_mod <- fp(p_root, "agroEcoTradeoff")
p_moddat <- fp(p_mod, "external/data/ZA")
p_figures <- full_path(p_root, "paper/figures/")
p_roads <- fp(p_root, "external/input_devel/roads")
p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
                   "rda_dump/Final Shape files RDA/")
p_out <- fp(p_mod, "external/output")
```

```{r, eval = FALSE}
# roads
mgrid <- raster(fp(p_moddat, "ZA-mask.tif"))
load(fp(p_mod, "external/data/ZA/parks_roads.rda"))
zamroads <- readOGR(fp(p_roads, "osmroads/zamroadsfp.shp"), 
                    layer = "zamroadsfp")
roads <- zamroads
save(pas, roads, zambia, 
     file = full_path(p_mod, "external/data/ZA/parks_roads.rda"))

```

## Demand
Following the example of Gasparri et al (2015), who report demand and area planted trends for soybean in Southern Africa (9 countries) between 2000-2013, we use the production trends for both maize and soybean for the periods for 2000-2014 to calculate scenario for production increases by extrapolating linearly from recent trends in production for each crop between 2014 and 2050. We took an average of both the 2000-2014 as well as the 2005-2014 trends, which show larger rates of gain, which may reflect the more recent agricultural development interests (cite).

```{r, eval = FALSE}
# fao stats for maize area harvested
ha <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-ha.xlsx")))
prod <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-prod.xlsx")))

# countries included in Gasparri et al (2015)
gasparri <- unique(ha$AreaName)[c(1:2, 7:8, 13, 15, 17, 19, 22)]
har <- ha[AreaName %in% gasparri]
setkey(har, "AreaName")
prodr <- prod[AreaName %in% gasparri]
setkey(prodr, "AreaName")
# prodr[AreaName == "Zambia" & Year %in% 2012:2014, mean(Value), by = ItemName]
# har[, unique(AreaName)]; prodr[, unique(AreaName)]

haregion <- har[, mean(Value), by = .(ItemName, Year)]
ha_rates <- sapply(c("Maize", "Soybeans"), function(i) {
  haregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
})
prodregion <- prodr[, mean(Value), by = .(ItemName, Year)]
prod_rates <- sapply(c("Maize", "Soybeans"), function(i) {
  prodregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
})
# prodregion[ItemName == "Maize", plot(Year, V1)]
# prodregion[ItemName == "Soybeans", plot(Year, V1)]
# haregion[ItemName == "Maize", plot(Year, V1)]
# haregion[ItemName == "Soybeans", plot(Year, V1)]

ha_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
  haregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
})
prod_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
  prodregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
})
# (1 + prod_rates05[2] * 13) * prodregion[ItemName == "Soybeans", V1]

allrates <- rbind(1 + ha_rates * (2050-2014), 1 + prod_rates * (2050-2014), 
                  1 + ha_rates05 * (2050-2014), 1 + prod_rates05 * (2050-2014))

# a <- 2
# b <- 10
# (a + sum((b - a) / (2010 - 2000) * rep(1, 10))) / a
# ((b - a) / (2010 - 2000))
# 5 * 100
# targets(c("maize" = 2), 10, 2)

round(colMeans(allrates[c(2, 4), ]), 1)  ## by production
round(colMeans(allrates[c(1, 3), ]), 1)  ## by area harvested
round(colMeans(allrates), 1)  ## averaged across production and area harvested
```

### Yield gap closure on existing land

We use the production based target as our potential scenario for Zambia.  Next is to factor in an assumption for how much demand will be met by closing current yield gaps
```{r}
setwd(p_mod)
path <- "external/data/ZA"
# load(fp(path, "currprod.rda"))
# load("~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/currprod.rda")
# save(currprod, file = fp(path, "currprod.rda"))
cp <- c("maize" = 2912004, "soy" = 203732)

# gap closing function
gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp

# Current gaps
mzgap <- 4.43 / 2.3 
soygap <- 3.4 / 2

# assume 50% closed on current land
g <- gapper(cp, c(mzgap, soygap), 0.25)

# targets given g
targ <- c("maize" = 3, "soy" = 9)
# targ <- c("maize" = 3.2, "soy" = 8.7)
targs <- targets(targ, cp, g)

```

## Tradeoff runs

### Best areas to meet targets for each crop

Running the model individually for each crop. This currently entails updating and the `currprod.rda` file, and then reinvoking `input_handler`  
```{r, eval = FALSE}
# currprod_switch <- function(cropind, path = p_moddat) {
#   master_prod <- c("maize" = 2912004, "soy" = 203732)
#   currprod <- master_prod[cropind]
#   save(currprod, file = fp(path, "currprod.rda"))
#   return(currprod)
# }

# Separate soy and maize impacts
cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
mzto <- tradeoff_mod(prod_targ = targ["maize"], cbetas = cb, currprodmod = g[1])
soyto <- tradeoff_mod(prod_targ = targ["soy"], cbetas = cb, currprodmod = g[2])
mzsoyto <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)

# mzto$conv[, sum(maize)]
# soyto$conv[, sum(soy)]

CRSobj <- CRS(mzto$inputs$sp$crs)
mztor <- dt_to_raster(mzto$conv, CRSobj)
soytor <- dt_to_raster(soyto$conv, CRSobj)
mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results

# rasters
mzsoyr <- stack(list(mztor, soytor * 2))
mzsoyr <- calc(mzsoyr, sum)
plot(mzsoyr)
# plot(roads)

# Plot
i1 <- rbind(mzto$impacts, soyto$impacts)
crops <- c("maize", "soy", "maize/soy")
# plot(1:10, pch = 20, cex = 4, col = "gold2")
# plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
imp_plot2(mzsoyr, i1, p_figures, "ideal_yield2", "", crops, TRUE, 
         c("gold2", "green3", "orange"))

```

### Best Areas for Other Three Factors
```{r, eval = FALSE}
basedt <- soyto$inputs$mask
ybl <- list(yb1 <- c(1, 1))
step <- 0.05
cnames <- c("C", "BD", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2); head(ps)
Yv = "conv_area"; Cv = "tot_C"; BDv = "int_mean"; COSTv = "mu_cost"
Yst = sum; Cst = sum; BDst = mean; COSTst = mean
crops <- list("maize", "soy", c("maize", "soy"))
croplnms <- list("maize", "soy", "both")
# lapply(crops, function(x) targ[x])
# lapply(crops, function(x) g[x])
cbdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                 yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                 Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                 Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                 silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(cbdc) <- croplnms

# check for datasets where certain weights were removed from contention
clength <- t(sapply(cbdc, function(x) c(nrow(x$params), nrow(x$optitab))))
cbdcmz <- heat2(cbdc$maize, fp(p_out, cbdc$maize$bcode), basedt, todisk = TRUE)
cbdc$maize$heat <- cbdcmz

# load(fp(p_out, "ZA_20160213_172518_13217_37779/impacts_tab.rda"))
# save(cbdc, file = fp(p_out, "results.rda"))
# par(mfrow = c(2, 2), mar = c(0, 0, 0, 4))
# plot(cbdc$both$heat, axes = FALSE, box = FALSE)
# plot(cbdc$maize$heat, axes = FALSE, box = FALSE)
# plot(cbdc$soy$heat, axes = FALSE, box = FALSE)
# a <- Sys.time()
# print(a)
# cbdc_mz <- pareto(cnames, step = step, prod_targ = targ["maize"], 
#                   yblist = ybl, currprodmod = g["maize"], input_key = "ZA", 
#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                   silent = FALSE, ncl = 8)
# Sys.time() - a
# load(fp(p_out, "bcode.rda"))
# bcodes <- bcode
# a <- Sys.time()
# cbdc_soy <- pareto(cnames, step = step, prod_targ = targ["soy"], 
#                    yblist = ybl, currprodmod = g["soy"], input_key = "ZA", 
#                    Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                    Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                    silent = FALSE, ncl = 8)
# Sys.time() - a
# load(fp(p_out, "bcode.rda"))
# bcodes[2] <- bcode
# implist <- list(cbdc_mz, cbdc_soy)
# names(implist) <- bcodes

# tstmp <- function() {
#   paste0(gsub(" ", "", gsub(":|-", "", as.character(Sys.time()))))
# }
# save(implist, file = fp(p_out, paste0("imptabs_", tstmp(), ".rda")))
# load(fp(p_out, "imptabs_20160212165418.rda"))
# implist

# Create overlap maps
# CRSobj <- CRS(soyto$inputs$sp$crs)
# cbdc_mz_ht <- heat_func(fp(p_out, names(implist)[1]), basedt) 
# cbdc_soy_ht <- heat_func(fp(p_out, names(implist)[2]), basedt) 
# plot(cbdc_soy_ht + cbdc_mz_ht)
# plot(cbdc_soy_ht)
# Might need to run it for both crops simultaneously

# Do RGB plot comparing overlaps between two

```

### Conservation interests versus agricultural interests
```{r, eval = FALSE}
ybl <- list(yb1 <- c(1, 1))
step <- 0.02
cnames <- c("Y", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
# ps[which(rowSums(ps[, c(1, 3)]) == 0), ]

ycost <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(ycost) <- croplnms

clength <- t(sapply(ycost, function(x) c(nrow(x$params), nrow(x$optitab))))

cnames <- c("C", "BD")
bdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab, "heat" = heatr)
})
names(bdc) <- croplnms
# bdc3 <- c(bdc, bdc2)  # vestiges to show I forgot to do maize soy case
# bdc <- bdc3

bdcfix <- lapply(bdc, function(x) {  # x <- bdc$maize
  if(nrow(x$params) > nrow(x$optitab)) {
    print(paste("fixing heat map for", x$bcode))
    heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE)
  } else {
    heatfix <- NULL
  }
  heatfix
})
for(i in 1:length(bdc)) bdc[[i]]$heat <- bdcfix[[i]]

save(cbdc, ycost, bdc, file = fp(p_out, "mainresults1302.rda"))
```

### All 4 together
```{r, eval = FALSE}
step <- 0.05
cnames <- c("Y", "C", "BD", "COST")
ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
ybdc <- lapply(crops, function(x) {
  a <- Sys.time()
  print(paste("Starting run for", x, "at", a))
  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
                  silent = FALSE, ncl = 7)
  load(fp(p_out, "bcode.rda"))
  # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
  print(Sys.time() - a)
  c("bcode" = bcode, otab) #, "heat" = heatr)
})
names(ybdc) <- croplnms

# ran because of fail on iteration for both (croplnms not properly compiled)
# bnames <- c("ZA_20160213_231906_17404_28678", "ZA_20160213_234959_17404_39818")
# ybdc <- lapply(bnames, function(x) {
#   bpath <- fp(p_out, x)
#   load(fp(bpath, "optitab.rda"))
#   load(fp(bpath, "parms.rda"))
#   list("bcode" = x, "optitab" = optimal, "params" = parms)
# })
# names(ybdc) <- croplnms[1:2]

# ybdc2 <- lapply(crops[3], function(x) {
#   a <- Sys.time()
#   print(paste("Starting run for", x, "at", a))
#   otab <- pareto(cnames, step = step, prod_targ = targ[x], 
#                   yblist = ybl, currprodmod = g[x], input_key = "ZA", 
#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
#                   silent = FALSE, ncl = 7)
#   load(fp(p_out, "bcode.rda"))
#   # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
#   print(Sys.time() - a)
#   c("bcode" = bcode, otab) #, "heat" = heatr)
# })
# names(ybdc2) <- croplnms[[3]]
# ybdc3 <- c(ybdc, ybdc2)  # vestiges to show I forgot to do maize soy case
# ybdc <- ybdc3

# heat maps
ybdcheat <- lapply(ybdc, function(x) {  # x <- bdc$maize
  print(paste("N params =", nrow(x$params), ": N optimal =",  nrow(x$optitab)))
  heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE, 
                   overwrite = TRUE)
  heatfix
})
for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcheat[[i]]

plot(ybdc$maize$heat)
par(mfrow = c(1, 2))
plot(ycost$maize$heat)
plot(ycost$soy$heat)
plot(ycost$both$heat)
plot(bdc$both$heat)

```

### 2016-02-14 Rmd/paper-todo.Rmd
Many updates. Up through new analysis.

@@
 
-+ Near-term: See if we can bias-correct using NPP.  
-+ Longer-term: DSSAT yields -> machine learning (Nate)
++ Near-term: See if we can bias-correct using NPP.  __Not workable__
++ Longer-term: DSSAT yields -> machine learning (Nate) __Done, with GAMs__
 
@@
 
-+ Near-term: Code in impact based on intactness + protected weight
-+ Longer-term: Code in Newbold model (Laura and Tobias)
++ Near-term: Code in impact based on intactness + protected weight __Done__
++ Longer-term: Code in Newbold model (Laura and Tobias) __Pending__
 
@@
 ------------- | -------------  | -------------
-NPP-crop correct  | 29 Jan |
-Biodiversity fix  | 31 Jan |
+NPP-crop correct  | 29 Jan | Not pursued
+Biodiversity fix  | 31 Jan | 11 Feb
 Crop overlaps | ??? | 
-Yield variability |  ??? |
+Yield variability |  ??? | Not pursued
 Combined optimization, back out weights? |  ???

### 2016-02-18 Rmd/carbon.Rmd
Many updates as getting ready to assemble and mask inputs
Redid Carbon stocks to correct fact that I was using veg biomass, not C, adding below ground, and re-ran through most of main-analysis up until first plots, using updated model to account for faulty grid ordering.

@@
 ```{r, eval=FALSE}
-onm <- full_path(p_devel, "veg-carbon.tif")
+onm <- full_path(p_devel, "veg-biomass.tif")
 ext <- bbox(mgrid)[1:4]  # extent   
@@
          of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
-zvcarb <- raster(onm)  # vegetation carbon
-plot(zvcarb)
+zvbiom <- raster(onm)  # vegetation carbon
+
+# convert to carbon and then above and below ground carbon
+zvcarb <- zvbiom * 0.47  # IPCC 2006
+
+# mean root shoot ratio, average of Mokany et al (2006) for savannas (Table 2)
+# and Mugasha et al (2013), Tanzanian miombo root-shoot ratio
+rooshoo <- 1 + mean(c(0.642, 0.4))
+zvcarbf <- zvcarb * rooshoo
+onm2 <- fp(p_devel, "veg-carbon.tif")
+writeRaster(zvcarbf, filename = onm2, overwrite = TRUE)
+# plot(zvcarbf - ocarb[[1]])
 
@@
 soct_1ma <- projectRaster(soct_1m, mgrid)
-writeRaster(soct_1ma, filename = full_path(p_devel, "soil-C-1m.tif"))
+writeRaster(soct_1ma, filename = full_path(p_devel, "soil-C-1m.tif"), 
+            overwrite = TRUE)
 
@@
 
-zamcarb <- brick(stack(zvcarb, soct_1ma), 
-                 filename = full_path(p_devel, "ZA-carbon.tif"))
+zamcarb <- brick(stack(zvcarbf, soct_1ma), 
+                 filename = full_path(p_devel, "ZA-carbon.tif"), 
+                 overwrite = TRUE)
 
 ```
+
+## Resulting model inputs
+
+`ZA-carbon.tif` - grid of carbon densities, for impact and prioritization

### 2016-02-18 Rmd/main-analysis.Rmd
Redid Carbon stocks to correct fact that I was using veg biomass, not C, adding below ground, and re-ran through most of main-analysis up until first plots, using updated model to account for faulty grid ordering.

@@
 library(gisutils)
+
 setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
@@
 mzsoyr <- calc(mzsoyr, sum)
-plot(mzsoyr)
+# plot(mzsoyr)
 # plot(roads)
@@
 # plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
-imp_plot2(mzsoyr, i1, p_figures, "ideal_yield2", "", crops, TRUE, 
+imp_plot2(mzsoyr, i1, p_figures, "ideal_yield", "", crops, 1, TRUE, 
          c("gold2", "green3", "orange"))
 
+basedt <- soyto$inputs$mask
 ```
@@
 ### Best Areas for Other Three Factors
+
+Switch off for now
 ```{r, eval = FALSE}
-basedt <- soyto$inputs$mask
-ybl <- list(yb1 <- c(1, 1))
-step <- 0.05
-cnames <- c("C", "BD", "COST")
-ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2); head(ps)
-Yv = "conv_area"; Cv = "tot_C"; BDv = "int_mean"; COSTv = "mu_cost"
-Yst = sum; Cst = sum; BDst = mean; COSTst = mean
-crops <- list("maize", "soy", c("maize", "soy"))
-croplnms <- list("maize", "soy", "both")
-# lapply(crops, function(x) targ[x])
-# lapply(crops, function(x) g[x])
-cbdc <- lapply(crops, function(x) {
-  a <- Sys.time()
-  print(paste("Starting run for", x, "at", a))
-  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
-                 yblist = ybl, currprodmod = g[x], input_key = "ZA", 
-                 Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-                 Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-                 silent = FALSE, ncl = 7)
-  load(fp(p_out, "bcode.rda"))
-  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
-  print(Sys.time() - a)
-  c("bcode" = bcode, otab, "heat" = heatr)
-})
-names(cbdc) <- croplnms
-
-# check for datasets where certain weights were removed from contention
-clength <- t(sapply(cbdc, function(x) c(nrow(x$params), nrow(x$optitab))))
-cbdcmz <- heat2(cbdc$maize, fp(p_out, cbdc$maize$bcode), basedt, todisk = TRUE)
-cbdc$maize$heat <- cbdcmz
-
-# load(fp(p_out, "ZA_20160213_172518_13217_37779/impacts_tab.rda"))
-# save(cbdc, file = fp(p_out, "results.rda"))
-# par(mfrow = c(2, 2), mar = c(0, 0, 0, 4))
-# plot(cbdc$both$heat, axes = FALSE, box = FALSE)
-# plot(cbdc$maize$heat, axes = FALSE, box = FALSE)
-# plot(cbdc$soy$heat, axes = FALSE, box = FALSE)
+# ybl <- list(yb1 <- c(1, 1))
+# step <- 0.05
+# cnames <- c("C", "BD", "COST")
+# ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2); head(ps)
+# Yv = "conv_area"; Cv = "tot_C"; #BDv = "int_prior"; COSTv = "mu_cost"
+# BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
+# wmu <- function(x, na.rm = na.rm) {
+#   weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
+# }
+# # Yst = sum; Cst = sum; BDst = mean; COSTst = mean
+# Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
+# 
+# # crops <- list("maize", "soy", c("maize", "soy"))
+# crops <- list(c("maize", "soy"))
+# # croplnms <- list("maize", "soy", "both")
+# croplnms <- list("both")
+# # lapply(crops, function(x) targ[x])
+# # lapply(crops, function(x) g[x])
+# cbdc <- lapply(crops, function(x) {
 #   a <- Sys.time()
-# print(a)
-# cbdc_mz <- pareto(cnames, step = step, prod_targ = targ["maize"], 
-#                   yblist = ybl, currprodmod = g["maize"], input_key = "ZA", 
+#   print(paste("Starting run for", x, "at", a))
+#   otab <- pareto(cnames, step = step, prod_targ = targ[x], 
+#                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
 #                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
 #                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-#                   silent = FALSE, ncl = 8)
-# Sys.time() - a
+#                  silent = FALSE, ncl = 7)
 #   load(fp(p_out, "bcode.rda"))
-# bcodes <- bcode
-# a <- Sys.time()
-# cbdc_soy <- pareto(cnames, step = step, prod_targ = targ["soy"], 
-#                    yblist = ybl, currprodmod = g["soy"], input_key = "ZA", 
-#                    Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-#                    Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-#                    silent = FALSE, ncl = 8)
-# Sys.time() - a
-# load(fp(p_out, "bcode.rda"))
-# bcodes[2] <- bcode
-# implist <- list(cbdc_mz, cbdc_soy)
-# names(implist) <- bcodes
-
-# tstmp <- function() {
-#   paste0(gsub(" ", "", gsub(":|-", "", as.character(Sys.time()))))
-# }
-# save(implist, file = fp(p_out, paste0("imptabs_", tstmp(), ".rda")))
-# load(fp(p_out, "imptabs_20160212165418.rda"))
-# implist
-
-# Create overlap maps
-# CRSobj <- CRS(soyto$inputs$sp$crs)
-# cbdc_mz_ht <- heat_func(fp(p_out, names(implist)[1]), basedt) 
-# cbdc_soy_ht <- heat_func(fp(p_out, names(implist)[2]), basedt) 
-# plot(cbdc_soy_ht + cbdc_mz_ht)
-# plot(cbdc_soy_ht)
-# Might need to run it for both crops simultaneously
-
-# Do RGB plot comparing overlaps between two
+#   heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
+#   print(Sys.time() - a)
+#   c("bcode" = bcode, otab, "heat" = heatr)
+# })
+# names(cbdc) <- croplnms
+# 
+# # cbdc <- lapply("ZA_20160215_071436_1558_16149", function(x) {
+# #   bpath <- fp(p_out, x)
+# #   load(fp(bpath, "impacts_tab.rda"))
+# #   inlist <- lapply(unique(outtab$iter), function(x) {
+# #     outtab[outtab$iter == x, ]
+# #   })
+# #   # load(fp(bpath, "optitab.rda"))
+# #   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
+# #                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
+# #   save(optimal, file = fp(bpath, "optitab.rda"))
+# #   load(fp(bpath, "parms.rda"))
+# #   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
+# #   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
+# # })
+# # names(cbdc) <- croplnms
+# 
+# # check for datasets where certain weights were removed from contention
+# # clength <- t(sapply(cbdc, function(x) c(nrow(x$params), nrow(x$optitab))))
+# # cbdcmz <- heat2(cbdc$maize, fp(p_out, cbdc$maize$bcode), basedt, todisk = TRUE)
+# # cbdc$maize$heat <- cbdcmz
 
@@
 # ps[which(rowSums(ps[, c(1, 3)]) == 0), ]
+Yv = "conv_area"; Cv = "tot_C"
+BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
+wmu <- function(x, na.rm = na.rm) {
+  weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
+}
+Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
+
+crops <- list(c("maize", "soy"))
+croplnms <- list("both")
 
@@
   load(fp(p_out, "bcode.rda"))
-  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
+  heatr <- heat2(fp(p_out, bcode), basedt, todisk = TRUE) 
   print(Sys.time() - a)
@@
 names(ycost) <- croplnms
+# plot(ycost$both$optitab$Y, ycost$both$optitab$COST)
 
-clength <- t(sapply(ycost, function(x) c(nrow(x$params), nrow(x$optitab))))
+# fix because ran with heat2 not yet adjusted for improperly ordered csv files
+ycostfix <- lapply(ycost, function(x) {  # x <- ycost$both
+  if(nrow(x$params) > nrow(x$optitab)) {
+    print(paste("fixing heat map for", x$bcode))
+    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
+  } else {
+    heatfix <- NULL
+  }
+  heatfix
+})
+for(i in 1:length(ycost)) ycost[[i]]$heat <- ycostfix[[i]]
+
+# ycost <- lapply("ZA_20160215_072118_1558_04246", function(x) {
+#   bpath <- fp(p_out, x)
+#   load(fp(bpath, "impacts_tab.rda"))
+#   inlist <- lapply(unique(outtab$iter), function(x) {
+#     outtab[outtab$iter == x, ]
+#   })
+#   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
+#                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
+#   save(optimal, file = fp(bpath, "optitab.rda"))
+#   load(fp(bpath, "parms.rda"))
+#   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
+#   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
+# })
+# names(ycost) <- croplnms
+# clength <- t(sapply(ycost, function(x) c(nrow(x$params), nrow(x$optitab))))
 
@@
   load(fp(p_out, "bcode.rda"))
-  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
+  # heatr <- heat2(fp(p_out, bcode), basedt, todisk = TRUE) 
   print(Sys.time() - a)
@@
 names(bdc) <- croplnms
-# bdc3 <- c(bdc, bdc2)  # vestiges to show I forgot to do maize soy case
-# bdc <- bdc3
 
-bdcfix <- lapply(bdc, function(x) {  # x <- bdc$maize
+# fix because ran with heat2 not yet adjusted for improperly ordered csv files
+bdcfix <- lapply(bdc, function(x) {  # x <- ycost$both
   if(nrow(x$params) > nrow(x$optitab)) {
     print(paste("fixing heat map for", x$bcode))
-    heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE)
+    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
   } else {
@@
 
-save(cbdc, ycost, bdc, file = fp(p_out, "mainresults1302.rda"))
+# plot(bdc$both$optitab$C, bdc$both$optitab$BD)
+# bdc3 <- c(bdc, bdc2)  # vestiges to show I forgot to do maize soy case
+# bdc <- bdc3
+
+# # Fix pareto tab
+# load(fp(fp(p_out, bdc$both$bcode), "impacts_tab.rda"))
+# inlist <- lapply(unique(outtab$iter), function(x) outtab[outtab$iter == x, ])
+# outtabf <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
+#                       Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
+
+# bdcfix <- lapply(bdc, function(x) {  # x <- bdc$maize
+#   if(nrow(x$params) > nrow(x$optitab)) {
+#     print(paste("fixing heat map for", x$bcode))
+#     heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE)
+#   } else {
+#     heatfix <- NULL
+#   }
+#   heatfix
+# })
+# for(i in 1:length(bdc)) bdc[[i]]$heat <- bdcfix[[i]]
+
+# bdc <- lapply("ZA_20160215_072332_1558_54465", function(x) {
+#   bpath <- fp(p_out, x)
+#   load(fp(bpath, "impacts_tab.rda"))
+#   inlist <- lapply(unique(outtab$iter), function(x) {
+#     outtab[outtab$iter == x, ]
+#   })
+#   # load(fp(bpath, "optitab.rda"))
+#   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
+#                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
+#   save(optimal, file = fp(bpath, "optitab.rda"))
+#   load(fp(bpath, "parms.rda"))
+#   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
+#   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
+# })
+# names(bdc) <- croplnms
+
+# save(cbdc, ycost, bdc, file = fp(p_out, "mainresults1402.rda"))
+save(ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
 ```
@@
 
+# fix because ran with heat2 not yet adjusted for improperly ordered csv files
+ybdcfix <- lapply(ybdc, function(x) {  # x <- ycost$both
+  if(nrow(x$params) > nrow(x$optitab)) {
+    print(paste("fixing heat map for", x$bcode))
+    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
+  } else {
+    heatfix <- NULL
+  }
+  heatfix
+})
+for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcfix[[i]]
+
 # ran because of fail on iteration for both (croplnms not properly compiled)
 # bnames <- c("ZA_20160213_231906_17404_28678", "ZA_20160213_234959_17404_39818")
-# ybdc <- lapply(bnames, function(x) {
+# ybdc <- lapply("ZA_20160215_072541_1558_17131", function(x) {
 #   bpath <- fp(p_out, x)
-#   load(fp(bpath, "optitab.rda"))
+#   load(fp(bpath, "impacts_tab.rda"))
+#   inlist <- lapply(unique(outtab$iter), function(x) {
+#     outtab[outtab$iter == x, ]
+#   })
+#   # load(fp(bpath, "optitab.rda"))
+#   optimal <- batch_stat(inlist, Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, 
+#                         Yst = Yst, Cst = Cst, BDst = BDst, COSTst = COSTst)
+#   save(optimal, file = fp(bpath, "optitab.rda"))
 #   load(fp(bpath, "parms.rda"))
-#   list("bcode" = x, "optitab" = optimal, "params" = parms)
+#   heatr <- heat2(fp(p_out, x), basedt, todisk = TRUE, overwrite = TRUE) 
+#   list("bcode" = x, "optimal" = optimal, "params" = parms, "heat" = heatr)
 # })
-# names(ybdc) <- croplnms[1:2]
+# names(ybdc) <- croplnms
 
@@
 # heat maps
-ybdcheat <- lapply(ybdc, function(x) {  # x <- bdc$maize
-  print(paste("N params =", nrow(x$params), ": N optimal =",  nrow(x$optitab)))
-  heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE, 
-                   overwrite = TRUE)
-  heatfix
+# ybdcheat <- lapply(ybdc, function(x) {  # x <- bdc$maize
+#   print(paste("N params =", nrow(x$params), ": N optimal =",  nrow(x$optitab)))
+#   heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE, 
+#                    overwrite = TRUE)
+#   heatfix
+# })
+# for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcheat[[i]]
+
+# plot(ybdc$maize$heat)
+# par(mfrow = c(1, 2))
+# plot(ycost$maize$heat)
+# plot(ycost$soy$heat)
+# plot(ycost$both$heat)
+# plot(bdc$both$heat)
+# save(ybdc, cbdc, ycost, bdc, file = fp(p_out, "mainresults1502.rda"))
+save(ybdc, ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
+# load(fp(p_out, "mainresults1402.rda"))
+```
+
+## Figures
+
+How reconcilable are the different land use interests, and what would be the impacts of each?
+
+I am using the two crops combined approach because there is little overlap between the two.
+
+### Figure 1
+```{r, eval = FALSE}
+# select out results for "pure" cases (weight = 1)
+pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))
+
+fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = pure, fullnames = TRUE)
+pure_r <- lapply(fnames, function(x) {  # x <- pure[1]
+  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
+  DTr <- dt_to_raster(DT, CRSobj)
 })
-for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcheat[[i]]
+# sapply(1:4, function(x) which(ps[, x] == 1))
+# sum across crops and multiply by convertible areas
+rs <- lapply(pure_r, function(x) calc(x[[1:2]], sum) * x[[3]])  # actual area
+rs2 <- lapply(pure_r, function(x) calc(x[[1:2]], sum))  # conv pixels
+# check
+# sapply(rs, function(x) cellStats(x, sum))
+# ybdc$both$optitab[ind %in% pure, Y]  # same, but C optimized out
+# sapply(rs2, function(x) cellStats(x, sum))
+# plot(pure_r[[3]])
 
-plot(ybdc$maize$heat)
-par(mfrow = c(1, 2))
-plot(ycost$maize$heat)
-plot(ycost$soy$heat)
-plot(ycost$both$heat)
-plot(bdc$both$heat)
+# figure out overlaps between different pure objectives
+pms <- expand.grid(rep(list(0:1), 4))[-1, ]  # logic grid
+rownames(pms) <- 1:15
+rs_over <- lapply(1:nrow(pms), function(x) {  # apply logic tests
+  (rs2[[1]] == pms[x, 1]) & (rs2[[2]] == pms[x, 2]) & (rs2[[3]] == pms[x, 3]) &     (rs2[[4]] == pms[x, 4])
+})
+rs_overs <- stack(rs_over)
+
+# legend to show class of overlap
+pms$leg <- c("Y", "C", "YC", "Bd", "YBd", "BdC", "YBdC", "Ct", 
+             "YCt", "CCt", "YCCt", "BdCt", "YBdCt", "CBdCt", 
+             "YCBdCt")
+
+# frequency of overlaps in different categories
+# sum(pms[c(1, 3, 5, 7, 9, 11, 13, 15), "hist"])
+pms$hist <- sapply(1:nlayers(rs_overs), function(x) {
+  round(cellStats(rs_overs[[x]] * pure_r[[1]]$convertible, sum), 1)
+})
+
+# legend color
+pms$col <- c("dodgerblue3", "goldenrod2", "tomato1", "yellow", 
+             rep("tomato1", 2), "red3", "deepskyblue1", rep("tomato1", 2),
+             rep("red3", 5))
+
+# recode with unique number for overlap class, collapse to RasterLayer, 
+# simplify classes for plotting
+rs_over_rc <- stack(lapply(1:length(rs_over), function(x) rs_over[[x]] * x))
+rs_over_rc <- calc(rs_over_rc, sum)
+# plot(rs_over_rc == 6)
+rclmat <- cbind(0:15, c(0, 1, 3, 5, 4, 5, 5, 6, 2, 5, 5, 6, 5, 6, 6, 6))
+leg2 <- c("Yield", "Cost", "Carbon", "Biodiversity", "2 overlap", "3+ overlap")
+rs_over_rc2 <- raster::reclassify(rs_over_rc, rclmat)
+
+# plot(1:2, pch = 20, cex = 3, col = c("yellow", "yellow2"))
+
+# For Panel B - impact statistics 
+bpath <- fp(p_out, ybdc$both$bcode)
+load(fp(bpath, "impacts_tab.rda"))  # get impacts table
+
+# select out corresponding impact statistics, summarize
+pure2 <- unlist(lapply(pure, function(x) c(x, x+1)))
+vars <- c("iter", "conv_area", "tot_C", "fpa_conv", "mu_cost", "rarity",
+          "priority",  "int_prior", "int_mean")
+conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
+# conv_stats <- outtab[outtab$iter %in% pure, vars]
+# conv_stats <- data.frame(t(sapply(pure, function(x) {  # x <- pure[1]
+#   DF <- conv_stats[conv_stats$iter == x, ]
+#   c(colSums(DF[, 2:4]), 
+#     sapply(5:ncol(DF), function(y) weighted.mean(DF[, y], DF$conv_area)))
+# }))) 
+# colnames(conv_stats) <- vars[-1]
+
+# rescale for common plot
+svec <- c(100000, 10000000, 1000, 1, 0.1, 0.1, 0.1, 0.1)  # rescale factor
+conv_stats_n <- sweep(as.matrix(conv_stats)[, -1], MARGIN = 2, svec, "/")
+conv_stats_n <- conv_stats_n[, c(1, 4, 2, 6:ncol(conv_stats_n), 3)]
+
+# Plots
+# Panel A
+# cols <- RColorBrewer::brewer.pal(11, "RdYlBu")
+# cols <- colorspace::diverge_hsv(n = 11)
+# colsi <- sapply(c(0.5, 0.75, 1), function(x) rgb(1, 1, 0.9, alpha = x))
+cols <- c("dodgerblue3", "deepskyblue1", "goldenrod2", "yellow", "tomato1",
+          "red3")
+plot(1:6, pch = 20, cex = 3, col = cols)
+
+tiff(fp(p_figures, "single_priorities.tif"), height = 3, width = 8, 
+     units = "in", res = 300)
+par(mfrow = c(1, 3))
+imp_plot2(dtr = rs_over_rc2, impacts = i1, fpath = p_figures, 
+          fnm = "pure_overlap", cexscale = 0.6, ftitle = "", crops = leg2, 
+          stats = FALSE, cols = cols, hold = TRUE, marg = c(0, 0, 0, 0))
+
+# Panel B
+# totf <- cellStats(rs_over_rc > 0, sum)
+par(mar = c(7, 4, 2, 2))
+bp <- barplot(pms$hist / sum(pms$hist) * 100, ylim = c(0, 25), las = 2, 
+              col = pms$col, mgp = c(2, 0.25, 0), tcl = -0.2, 
+              ylab = "% converted area")
+axis(side = 1, at = bp, labels = pms$leg, las = 2, mgp = c(2, 0.25, 0), 
+     tcl = -0.2)
+# TeachingDemos::subplot( 
+#   bp <- barplot(pms$hist / sum(pms$hist) * 100, ylim = c(0, 25), las = 2, 
+#                 col = pms$col, mgp = c(2, 0.25, 0), tcl = -0.2, 
+#                 ylab = "% converted pixels", names.arg = pms$leg), 
+#   x=grconvertX(c(0.1,0.45), from='npc'),
+#   y=grconvertY(c(0.6,1), from='npc'),
+#   type='fig', pars=list(mar=c(3,3,0,0) + 0.1))
+
+# Panel C
+# barplot(conv_stats_n)
+blabs <- c("Area", "Time", "Carbon", "Priority", "Priority/intactness", 
+           "Intactness", "Forest reserves")
+# blabs <- sub(':', '\n', blabs)
+tlabs <- c(expression(10^5~"km"^2), "hours", expression(10^7~"tons"), 
+           rep("X10", 3), expression(10^3~"km"^2))
+bp <- barplot(conv_stats_n, beside = TRUE, ylim = c(0, 14), las = 2, 
+              col = pms$col[c(1, 2, 4, 8)], names.arg = rep("", length(blabs)), 
+              ylab = "Impact*", mgp = c(2, 0.25, 0), tcl = -0.2)
+text(x = bp[2, ] - 0.4, y = rep(-0.3, length(bp[1, ])), adj = 1, 
+     labels = blabs, xpd = NA, srt = 90)
+text(x = bp[3, ] + 0.4, y = rep(-0.3, length(bp[3, ])), adj = 1, 
+     labels = tlabs, xpd = NA, srt = 90, cex = 0.8)
+dev.off()
+
+#### Need to redo pareto to pass in weighted mean function for all average 
+#### costs (DONE)
 
 ```
+
+## Figure 2 & 3 (probably supplemental)
+```{r, eval = FALSE}
+# yc_bdc_st <- as.matrix(stack(ycost$both$heat, bdc$both$heat) / 100)
+# yc_bdc_st <- stack(ycost$both$heat, ycost$both$heat * 0, bdc$both$heat) / 100
+# plot(ycost$both$heat > 1)
+# plotRGB(yc_bdc_st, r = 1, g = 2, b = 3, colNA = "transparent", scale = 1)
+# rgb(red = yc_bdc_st[, 1], rep(0, nrow(yc_bdc_st)), yc_bdc_st[, 2])
+# plot(ycost$both$heat)
+
+# individual maps
+png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
+     units = "in", res = 300)
+par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
+ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
+legp <- c(TRUE, rep(FALSE, 3))
+for(i in 1:4) {
+  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
+  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
+            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
+            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
+}
+dev.off()
+
+# heat maps
+implist <- list(ycost$both, bdc$both)
+load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
+pacols <- c("transparent", "grey85", "grey75")
+png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
+par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
+for(i in 1:2) {
+  par(mar = c(1, 1, 1, 4))
+  plot(zambia, col = "grey90", border = "transparent")
+  ir <- implist[[i]]$heat
+  ir[ir == 0] <- NA
+  plot(ir, col = rev(heat.colors(20)), add = TRUE, )
+  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+       border = FALSE)
+  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+       border = FALSE)
+  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
+}
+dev.off()
+
+```
+
+## Figure 4
+
+Value of just a bit of compromise
+
+```{r, eval = FALSE}
+p <- data.frame(ybdc$both$params)
+outtab[1, ]
+p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
+# scens <- list(c(0.9, 0.05, 0.05, 0), 
+#               c(0.85, 0.05, 0.05, 0.05), 
+#               c(0.1, 0.45, 0.45, 0.0), 
+#               c(0.05, 0.45, 0.45, 0.05))
+scens <- lapply(c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), function(x) {
+  c(1 - x, rep(x / 2, 2), 0)
+})
+# scens <- lapply(c(0, 0.15, 0.3, 0.45, 0.6, 0.75), function(x) {
+#   c(1 - x, rep(x / 3, 3))
+# })
+pind <- sapply(scens, function(x) {
+  x <- round(x, 2)
+  which(p$Y == x[1] & p$C == x[2] & p$BD == x[3] & p$COST == x[4])
+})
+
+# select out impacts corresponding to those runs
+cstats <- selected_impacts(pind, p_out, ybdc$both$bcode)
+
+# check on selection
+# il <- input_handler(p_moddat, crops[[1]])
+# chk <- tradeoff_mod(targ, ybdc$both$params[pind[5], 7:10], 
+#                     currprodmod = g, exist_list = il)
+# DF <- data.frame(chk$impacts)
+# outtab[outtab$iter == pind[5], -2]
+# round(c(colSums(DF[, 2:4]),
+#         sapply(5:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))), 
+#       2)
+# cstats[5, ]
+outtab[outtab$iter == pind[5], ]
+outtab[, 1:4][1:100,]
+
+fnames <- dir(bpath, pattern = "csv")
+fnames[pind[5]]
+
+# calculate max impacts across both crops, and figure out the impact range, 
+# and what percent that is relative to the worst case
+allstats <- selected_impacts(1:nrow(p), p_out, ybdc$both$bcode)
+minmaxstats <- allstats[, lapply(.SD, range)]
+imp_range <- minmaxstats[, lapply(.SD, diff)]  # difference between best/worst
+pct_of_tot <- minmaxstats[, lapply(.SD, function(x) diff(x) / max(x))]
+
+dfunc <- function(x, y) 100 - y / x * 100
+stnms <- names(cstats)[-1]
+imp_pct <- cstats[, lapply(stnms, function(x) {
+  dfunc(minmaxstats[[x]][2], get(x))
+})]
+setnames(imp_pct, stnms)
+
+plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
+     axes = FALSE)        # xaxs = "r", yaxs = "r")
+
+sctab <- do.call(rbind, scens) * 100
+cx <- 0.7
+lcols <- c("black", "red", "blue")
+imp_pct[, {
+  png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
+      units = "in", res = 300)
+  par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
+  plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
+       axes = FALSE, ylim = c(0, 65), xlim = c(2, 12))#, xaxs = "r", yaxs = "r")
+  lines(conv_area, int_prior, col = "red")
+  lines(conv_area, mu_cost, col = "blue")
+  axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
+       tcl = -0.2, mgp = c(1, 0.25, 0))
+  axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
+       tcl = -0.15, mgp = c(1, 0.25, 0))
+  mtext("% land savings", side = 1, line = 1, cex = cx)
+  mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
+  inms <- c("tot_C", "int_prior", "mu_cost")
+  for(i in 1:length(inms)) {
+    text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
+         cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
+    text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
+         offset = i * 0.15 - 0.2, col = lcols[i])
+  }
+  legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
+         text.col = lcols, col = lcols, bty = "n", lty = 1)
+  dev.off()
+}]
+
+scens_r <- lapply(pind, function(x) {  # x <- pind[[1]]
+  print(x)
+  inm <- paste0(outtab[outtab$iter == x, "rc"][1], ".csv")
+  fpath <- fp(fp(p_out, ybdc$both$bcode), inm)
+  # fname <- dir(bpath, pattern = "csv", full.names = TRUE)[x]
+  DT <- cbind(fread(fpath), basedt[, .(x, y, convertible)])
+  DTr <- dt_to_raster(DT, CRSobj)
+})
+
+scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
+# freq((rs2$Y + scens_rs2[[1]]) * scens_r[[1]]$convertible
+# plot(scens_rs2[[1]] + scens_rs2[[2]])
+
+# round(sapply(scens_rs2, function(x) {
+#   cellStats(x * scens_r[[1]]$convertible * 100, sum)
+# }))
+# cstats$conv_area
+
+# pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
+# pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid
+
+# rasters showing overlap as weight is progressively put on C and BD
+scens_over <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)
+
+# overlap areas
+scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
+  sapply(1:3, function(y) {  # y <- 3
+    r <- scens_over[[x]] == y
+    # plot(r)
+    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+  })
+})
+sum(scens_over_a[c(2, 3), 4])
+# over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
+# par(mar = c(0, 0, 0, 0))
+# plot(1:3, pch = 20, col = rgb(1, 0, 1))
+# plot(zambia)
+
+cols <- c("grey", c(rgb(0, 1, 0), rgb(0, 0, 1), rgb(1, 0, 1)))
+ftits <- c("Y=0.9")
+par(mfrow = c(1, 4))
+for(i in 1:4) {
+  plot(zambia)
+  imp_plot2(scens_over[[i]], impacts = i1, fpath = p_figures, fnm = "bob", 
+            col = cols, add = TRUE, 
+            legend = FALSE)
+}
+plot(scens_over[[1]], col = c("transparent", over_cols[[5]]))
+plot(scens_rs2[[1]])
+plot(scens_over[[4]])
+plot(calc(stack(scens_over), sum))
+
+plot(calc(stack(scens_rs2[2:4]), sum))
+
+ybdc$both$optimal[, {
+  par(mfrow = c(2, 2), mar = rep(3, 4))
+  hist(.SD, axes = FALSE, labels = FALSE)
+}, .SDcol = 1:4]
+
+frange <- function(x) diff(range(x)) / max(x)
+fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
+ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
+ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
+
+hist(ybdc$both$optimal$Y)
+
+```

### 2016-02-21 Rmd/main-analysis.Rmd
Analyses done through objective weights figure, which uses nearest neighbors and Jaccard Index to assess spatial differences between conversion maps.

@@
 library(gisutils)
+# URL <- paste0("https://cran.r-project.org/src/contrib/Archive/spatstat/",
+#               "spatstat_1.42-2.tar.gz")
+# install.packages(URL, repos = NULL, type="source", dependencies = TRUE)
+library(spatstat)
 
@@
 p_figures <- full_path(p_root, "paper/figures/")
+p_other <- fp(p_root, "external/input_devel/other/")
 p_roads <- fp(p_root, "external/input_devel/roads")
@@
 
+# farmblocks
+fbp <- "/Users/lestes/Dropbox/data/zari/misc/NEW4GEO/Fblocks.sqlite"
+# fb <- readOGR(fbp, layer = "fblocks")
+# fbalb <- spTransform(fb, mgrid@crs)
+# fbalb$ID <- 1:nrow(fbalb)
+# fbalb@data <- fbalb@data[, "ID", drop = FALSE]
+fnm <- fp(p_other, "fblocks_alb.sqlite")
+# writeOGR(fbalb, dsn = fnm, layer = "fblocks_alb", driver = "SQLite")
+
+# read in farmblocks
+fbalb <- readOGR(fnm, layer = "fblocks_alb")
+fbr <- rasterize(fbalb, mgrid)
+fbr[is.na(fbr)] <- 0
+fbr[fbr > 0] <- 1
+fbr <- raster::mask(fbr, mgrid)
+# plot(fbr)
+
 ```
+
+<a href="#top">Back to top</a>
 
@@
 
+<a href="#top">Back to top</a>
+
 ### Yield gap closure on existing land
@@
 ```
+
+<a href="#top">Back to top</a>
 
@@
 ```
+<a href="#top">Back to top</a>
 
@@
 
+<a href="#top">Back to top</a>
+
 ### All 4 together
@@
 save(ybdc, ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
-# load(fp(p_out, "mainresults1402.rda"))
+# load(fp(p_out, "mainresults1702.rda"))
 ```
+
+<a href="#top">Back to top</a>
 
@@
 # select out results for "pure" cases (weight = 1)
+cnames <- c("Y", "C", "BD", "COST")
 pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))
@@
 # For Panel B - impact statistics 
-bpath <- fp(p_out, ybdc$both$bcode)
-load(fp(bpath, "impacts_tab.rda"))  # get impacts table
+# bpath <- fp(p_out, ybdc$both$bcode)
+# load(fp(bpath, "impacts_tab.rda"))  # get impacts table
 
 # select out corresponding impact statistics, summarize
-pure2 <- unlist(lapply(pure, function(x) c(x, x+1)))
+pure2 <- unlist(lapply(pure, function(x) c(x, x + 1)))
 vars <- c("iter", "conv_area", "tot_C", "fpa_conv", "mu_cost", "rarity",
@@
 conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
-# conv_stats <- outtab[outtab$iter %in% pure, vars]
-# conv_stats <- data.frame(t(sapply(pure, function(x) {  # x <- pure[1]
-#   DF <- conv_stats[conv_stats$iter == x, ]
-#   c(colSums(DF[, 2:4]), 
-#     sapply(5:ncol(DF), function(y) weighted.mean(DF[, y], DF$conv_area)))
-# }))) 
-# colnames(conv_stats) <- vars[-1]
 
@@
           "red3")
-plot(1:6, pch = 20, cex = 3, col = cols)
+# plot(1:6, pch = 20, cex = 3, col = cols)
 
@@
 imp_plot2(dtr = rs_over_rc2, impacts = i1, fpath = p_figures, 
-          fnm = "pure_overlap", cexscale = 0.6, ftitle = "", crops = leg2, 
+          fnm = "pure_overlap", cexscale = 0.5, ftitle = "", crops = leg2, 
           stats = FALSE, cols = cols, hold = TRUE, marg = c(0, 0, 0, 0))
@@
 
-#### Need to redo pareto to pass in weighted mean function for all average 
-#### costs (DONE)
-
 ```
 
-## Figure 2 & 3 (probably supplemental)
+<a href="#top">Back to top</a>
+
+### Figures 2 & 3 (probably supplemental)
 ```{r, eval = FALSE}
@@
 
-## Figure 4
+<a href="#top">Back to top</a>
 
-Value of just a bit of compromise
+### Figures 4 (main) and 5 (probably supplemental)
+
+What happens when weight is progressively added
 
@@
 p <- data.frame(ybdc$both$params)
-outtab[1, ]
 p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
@@
 # select out impacts corresponding to those runs
-cstats <- selected_impacts(pind, p_out, ybdc$both$bcode)
+cstats <- selected_impacts(pind, fp(p_out, ybdc$both$bcode))
 
@@
 # il <- input_handler(p_moddat, crops[[1]])
-# chk <- tradeoff_mod(targ, ybdc$both$params[pind[5], 7:10], 
+# chk <- tradeoff_mod(prod_targ = targ, cbetas = ybdc$both$params[pind[5], 7:10], 
 #                     currprodmod = g, exist_list = il)
@@
 # outtab[outtab$iter == pind[5], -2]
-# round(c(colSums(DF[, 2:4]),
-#         sapply(5:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))), 
-#       2)
+# round(c(colSums(DF[, 1:2]),
+#         sapply(4:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))),2)
 # cstats[5, ]
-outtab[outtab$iter == pind[5], ]
-outtab[, 1:4][1:100,]
-
-fnames <- dir(bpath, pattern = "csv")
-fnames[pind[5]]
+# outtab[outtab$iter == pind[5], ]
+# fnames <- dir(bpath, pattern = "csv")
+# fnames[pind[5]]
 
@@
 # and what percent that is relative to the worst case
-allstats <- selected_impacts(1:nrow(p), p_out, ybdc$both$bcode)
+allstats <- selected_impacts(1:nrow(p), fp(p_out, ybdc$both$bcode))
 minmaxstats <- allstats[, lapply(.SD, range)]
@@
 
-plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
-     axes = FALSE)        # xaxs = "r", yaxs = "r")
+# plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
+#      axes = FALSE)        # xaxs = "r", yaxs = "r")
 
@@
 
-scens_r <- lapply(pind, function(x) {  # x <- pind[[1]]
+fnames <- fp(fp(p_out, ybdc$both$bcode), 
+             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
+scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
   print(x)
-  inm <- paste0(outtab[outtab$iter == x, "rc"][1], ".csv")
-  fpath <- fp(fp(p_out, ybdc$both$bcode), inm)
-  # fname <- dir(bpath, pattern = "csv", full.names = TRUE)[x]
-  DT <- cbind(fread(fpath), basedt[, .(x, y, convertible)])
+  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
   DTr <- dt_to_raster(DT, CRSobj)
@@
 scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
-# freq((rs2$Y + scens_rs2[[1]]) * scens_r[[1]]$convertible
 # plot(scens_rs2[[1]] + scens_rs2[[2]])
-
+# check
 # round(sapply(scens_rs2, function(x) {
@@
 })
-sum(scens_over_a[c(2, 3), 4])
+# colSums(scens_over_a[c(2, 3), ])
 # over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
@@
 
-cols <- c("grey", c(rgb(0, 1, 0), rgb(0, 0, 1), rgb(1, 0, 1)))
-ftits <- c("Y=0.9")
-par(mfrow = c(1, 4))
-for(i in 1:4) {
-  plot(zambia)
+# cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
+cols <- c("red", "blue", "gold")
+plot(1:3, pch = 20, col = cols, cex = 20)
+ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
+              "Y 80% C 10% Biod 10% Cost 0%",
+              "Y 70% C 15% Biod 15% Cost 0%",
+              "Y 60% C 20% Biod 20% Cost 0%")
+leg <- c("Y", "YCBd", "Y + YCBd")
+
+tiff(fp(p_figures, "changing_weights.tif"), height = 4, width = 7, 
+     units = "in", res = 300)
+par(mfrow = c(2, 4), oma = c(0, 0, 1, 0))
+for(i in 1:4) { # i <- 1
+  # map
   imp_plot2(scens_over[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
-            col = cols, add = TRUE, 
-            legend = FALSE)
+            crops = leg[c(1, 3, 2)], ftitle = ftits[[i]], col = cols, 
+            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+            marg = rep(0, 4))
+  # barplot
+  par(mar = c(5, 4, 3, 3))
+  overl <- scens_over_a[1:3, i][c(1, 3, 2)]
+  bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
+                col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
+                ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
+  axis(side = 1, at = bp, labels = leg, las = 2, 
+       mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
 }
-plot(scens_over[[1]], col = c("transparent", over_cols[[5]]))
-plot(scens_rs2[[1]])
-plot(scens_over[[4]])
-plot(calc(stack(scens_over), sum))
-
-plot(calc(stack(scens_rs2[2:4]), sum))
-
-ybdc$both$optimal[, {
-  par(mfrow = c(2, 2), mar = rep(3, 4))
-  hist(.SD, axes = FALSE, labels = FALSE)
-}, .SDcol = 1:4]
-
-frange <- function(x) diff(range(x)) / max(x)
-fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
-ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
-ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
-
-hist(ybdc$both$optimal$Y)
+dev.off()
 
 ```
+<a href="#top">Back to top</a>
+
+### Figure 6
+
+Objective weighting of outcomes, based on range of potential impacts
+
+```{r, eval = FALSE}
+# objective weights
+obj_wgts <- round(pct_of_tot[, {
+  c(conv_area, tot_C, int_mean, mu_cost) / 
+    sum(c(conv_area, tot_C, int_mean, mu_cost))
+}], 2)
+# round(obj_wgts, 2)
+names(obj_wgts) <- cnames
+
+# Objective weights
+toobj <- tradeoff_mod(targ, cbetas = obj_wgts, currprodmod = g, it = "OBJ")
+
+# ybdc$both$params[1, ]
+# yp <- ybdc$both$params[, 7:10]
+# which(round(yp[, 1], 2) == 0.05 & round(yp[, 2], 2) == 0.30 & 
+#         round(yp[, 3], 2) == 0.25 & round(yp[, 4], 2) == 0.4)
+
+# objective weights on ag only 
+# considering that these might be two sides of the same coin
+ag_wgts <- round(obj_wgts[c(1, 4)] / sum(obj_wgts[c(1, 4)]), 2)
+ag_wgts <- c(ag_wgts[1], "C" = 0, "BD" = 0, ag_wgts[2])
+ag_wgts2 <- ag_wgts[c(4, 2:3, 1)]
+names(ag_wgts2) <- names(ag_wgts)
+
+toag <- tradeoff_mod(targ, cbetas = ag_wgts, currprodmod = g, it = "AGOBJ")
+toag2 <- tradeoff_mod(targ, cbetas = ag_wgts2, currprodmod = g, it = "AGOBJ2")
+
+# subjective weights, where we want to treat everyone equally
+subj_wgts <- rep(0.25, 4)
+names(subj_wgts) <- cnames  # rename to make sure model runs properly
+tosubj <- tradeoff_mod(targ, cbetas = subj_wgts, currprodmod = g, it = "SUBJ")
+
+# extract impacts of these, and add in BAU case
+tores <- lapply(list(toag, toag2, toobj, tosubj), function(x) {  # x <- toobj
+  DTsum <- data.table(x$conv[, rowSums(.SD), .SDcols = c(crops[[1]])])
+  DTspat <- cbind(basedt[, .(x, y)], DTsum)
+  DTr <- dt_to_raster(DTspat, CRSobj)
+  DTr <- DTr + (fbr * 2)  # overlap with farmblocks, for output
+  list("map" = DTr, "imp" = x$impacts)
+})
+names(tores) <- c("ag", "ag2", "obj", "subj")
+
+Calculate area of overlaps between farmblocks and 
+yfbr <- calc(mzsoytor, sum) + fbr * 2
+plot(yfbr)
+fb_over_a <- sapply(1:length(tores), function(x) { # x <- 4
+  sapply(1:3, function(y) {  # y <- 3
+    r <- tores[[x]]$map == y
+    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+  })
+})
+fb_sim <- fb_over_a[3, ] / colSums(fb_over_a)
+
+# grid for calculating overlaps
+pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
+pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
+pms2b$cl <- c("AgAg2", "AgObj", "Ag2Obj", "AgSubj", "Ag2Subj", "ObjSubj")
+
+toresr <- lapply(tores, function(x) {
+  r <- x$map
+  r[r == 2] <- 0
+  r[r == 3] <- 1
+  r
+})
+
+# tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+#   (toresr[[1]] == pms2b[x, 1]) & (toresr[[2]] == pms2b[x, 2]) & 
+#     (toresr[[3]] == pms2b[x, 3]) & (toresr[[4]] == pms2b[x, 4])
+# })
+# names(tores_over_r) <- pms2b$cl
+# tores_overs <- stack(tores_over)
+tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
+})
+names(tores_over_r) <- pms2b$cl
+
+# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
+tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
+  sapply(1:2, function(y) {  # y <- 3
+    r <- tores_over_r[[x]] == y
+    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+  })
+})
+pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)
+
+# calculate nearest neighbor distance between each version, using spatstat 
+# nncross
+# plot(zam_owin)
+zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+                 unitname = "meter")
+which1 <- function(x) x == 1  # selector function for r to xy
+tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
+  ind <- which(pms2b[x, 1:4] == 1)
+  rl <- toresr[ind]
+  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+  pps <- lapply(rxy, function(x) {
+    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+  })
+  # reorder to find which has most points, to put that in as X
+  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+})
+pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
+# with(pms2b, plot(NND, JS))
+
+# tores_over_a <- lapply(1:nrow(pms2a), function(x) {  # apply logic tests
+#   (toresr[[1]] == pms2a[x, 1]) & (toresr[[2]] == pms2a[x, 2]) & 
+#     (toresr[[3]] == pms2a[x, 3]) & (toresr[[4]] == pms2a[x, 4])
+# })
+# names(tores_over_a) <- pms2a$cl
+
+# colst <- c(rgb(1, 0, 0),   # Ag
+#            rgb(1, 0.5, 0),  # Ag 2
+#            rgb(1, 1, 0),   # Ag1 and 2
+#            rgb(0, 0, 1),  # Obj
+#            rgb(1, 0, 1),  # Ag1 and Obj
+#            rgb(1, 0.5, 1),  #  Ag2 and Obj
+#            rgb(1, 1, 0.5),  #  Ag, Ag2 and Obj
+#            rgb(0, 0.5, 1),  # Subj
+#            rgb(0, 1, 1),  # Ag and Subj
+#            rgb(0, 1, 1),  # Ag2 and Subj
+#            rgb(0, 1, 1),  # Ag, Ag2, Subj
+#            rgb(0, 1, 1),  # Obj and Subj
+#            rgb(0, 1, 1),  # Ag,  Obj and Subj
+#            rgb(0, 1, 1),  # Ag2,  Obj and Subj
+#            rgb(0, 1, 1))  # All
+# plot(1:length(colst), pch = 20, cex = 3, col = colst)
+
+# colst <- c(rgb(0, 1, 1), 
+#            rgb(0, 0.5, 1), 
+#            rgb(0, 0, 1))
+# colst <- c(rgb(1, 1, 0), 
+#            rgb(1, 0.5, 0), 
+#            rgb(1, 0, 0))
+
+# do sum factoring in overlaps
+
+# sapply(tores_over_r, function(x) {
+#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
+# })
+
+# toressr <- calc(stack(toresr), sum)
+# plot(toressr)
+# freq(toressr)
+
+# plot(toresr$ag2)
+# plot(tores$obj$map - tores$subj$map)
+
+# cols_tr <- c(rgb(1, 0.5, 0, alpha = 0.5), rgb(1, 0, 0, alpha = 0.5), 
+#              rgb(0, 0.5, 1, alpha = 0.5), rgb(0, 0, 1, alpha = 0.5))
+cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
+plot(1:4, cex = 3, pch = 20, col = cols_tr)
+
+# par(mfrow = c(2, 2), mar = rep(0, 4))
+# # plot(tores$ag$map, col = c("grey", cols_tr[1]), axes = FALSE, legend = FALSE)
+# for(i in 1:4) {
+#   plot(zambia, col = "grey")
+#   plot(toresr[[i]], col = c("transparent", cols_tr[i]), add = TRUE) 
+# }
+                   # add = TRUE, legend = FALSE)
+# plot(toresr[[4]])
+
+# impact table
+imp_list <- list(toag$impacts, toag2$impacts, toobj$impacts, tosubj$impacts)
+imps <- rbindlist(lapply(imp_list, function(x) selected_impacts2(x)))
+impsr <- lapply(c(1, 4, 2, 6:8, 3), function(x) imps[[x]])
+
+# layout matrix
+m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
+m2 <- matrix(rep(5, 20), 4, 5)
+m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
+lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
+
+# plot(1:10, ylab = ylabs[2])
+# ylabs <- c("Jaccard Index", 
+ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
+           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
+           "Intactness", expression(paste("FR converted (ha"^2, ")")))
+# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
+yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
+ftits <- list("Y 11% C 0% Biod 0% Cost 89%",
+              "Y 89% C 0% Biod 0% Cost 11%",
+              "Y 5% C 29% Biod 25% Cost 41%",
+              "Y 25% C 25% Biod 25% Cost 25%")
+leg <- c("Converted")
+
+# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
+yaxes <- list(seq(0, 14, 2), seq(0, 2.5, 0.5),               
+              seq(0, 1000, 200), seq(0, 0.2, 0.05), seq(0, 0.5, 0.1), 
+              seq(0, 0.8, 0.2), seq(0, 3500, 500))
+# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
+bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
+                cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
+# lcols <- c(list(c(cols_tr[2], cols_tr[3], cols_tr[3], cols_tr[4], cols_tr[4] 
+# lcols <- lapply(1:7, function(x) cols_tr)
+
+tiff(fp(p_figures, "objective_weights.tif"), height = 3.5, width = 7, 
+     units = "in", res = 300)
+l <- layout(mat = lmat)
+# layout.show(l)
+
+# map plots
+for(i in 1:4) { # i <- 1
+  # map
+  # plot(zambia)
+  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
+            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+            marg = rep(0, 4))
+}
+
+# impact plots
+mgp <- c(1, 0.25, 0)
+par(mar = c(4, 3, 0, 0))
+with(pms2b, {
+     plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
+          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 1))
+     points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
+     axis(1, at = seq(0, 130, 10), cex.axis = 0.6, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
+     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
+})
+for(i in 1:7) {
+  v <- impsr[[i]] / yscale[i]
+  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
+               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
+               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
+               cex.lab = 0.7)
+  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
+       tcl = -0.2)
+}
+dev.off()
+
+# ybdc$both$optitab[, {
+#   par(mfrow = c(2, 2), mar = rep(3, 4))
+#   hist(.SD, axes = FALSE, labels = FALSE)
+# }, .SDcol = 1:4]
+# 
+# frange <- function(x) diff(range(x)) / max(x)
+# fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
+# ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
+# ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
+# 
+# hist(ybdc$both$optimal$Y)
+
+```

### 2016-02-23 Rmd/figures.Rmd
Separated figures onto new Rmd, and completely analysis up through weight selection as function of range in data relative to worst case.

--- 2016-02-14 Rmd/main-analysis.Rmd
@@
 library(gisutils)
+# URL <- paste0("https://cran.r-project.org/src/contrib/Archive/spatstat/",
+#               "spatstat_1.42-2.tar.gz")
+# install.packages(URL, repos = NULL, type="source", dependencies = TRUE)
+library(spatstat)
+
 setwd("~/Dropbox/publications/zambiaToff/private/zambiaToff/")
@@
 p_figures <- full_path(p_root, "paper/figures/")
-p_roads <- fp(p_root, "external/input_devel/roads")
-p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
-                   "rda_dump/Final Shape files RDA/")
+p_other <- fp(p_root, "external/input_devel/other/")
+# p_roads <- fp(p_root, "external/input_devel/roads")
+# p_roads2 <- paste0("/Users/lestes/Dropbox/data/zari/infrastructure/zambia/", 
+#                    "rda_dump/Final Shape files RDA/")
 p_out <- fp(p_mod, "external/output")
@@
 load(fp(p_mod, "external/data/ZA/parks_roads.rda"))
-zamroads <- readOGR(fp(p_roads, "osmroads/zamroadsfp.shp"), 
-                    layer = "zamroadsfp")
-roads <- zamroads
-save(pas, roads, zambia, 
-     file = full_path(p_mod, "external/data/ZA/parks_roads.rda"))
+
+# farmblocks
+fbp <- "/Users/lestes/Dropbox/data/zari/misc/NEW4GEO/Fblocks.sqlite"
+# fb <- readOGR(fbp, layer = "fblocks")
+# fbalb <- spTransform(fb, mgrid@crs)
+# fbalb$ID <- 1:nrow(fbalb)
+# fbalb@data <- fbalb@data[, "ID", drop = FALSE]
+fnm <- fp(p_other, "fblocks_alb.sqlite")
+# writeOGR(fbalb, dsn = fnm, layer = "fblocks_alb", driver = "SQLite")
+
+# read in farmblocks
+fbalb <- readOGR(fnm, layer = "fblocks_alb")
+fbr <- rasterize(fbalb, mgrid)
+fbr[is.na(fbr)] <- 0
+fbr[fbr > 0] <- 1
+fbr <- raster::mask(fbr, mgrid)
+# plot(fbr)
 
@@
 
-## Demand
-Following the example of Gasparri et al (2015), who report demand and area planted trends for soybean in Southern Africa (9 countries) between 2000-2013, we use the production trends for both maize and soybean for the periods for 2000-2014 to calculate scenario for production increases by extrapolating linearly from recent trends in production for each crop between 2014 and 2050.
-We took an average of both the 2000-2014 as well as the 2005-2014 trends, which show larger rates of gain, which may reflect the more recent agricultural development interests (cite).
-
-```{r, eval = FALSE}
-# fao stats for maize area harvested
-ha <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-ha.xlsx")))
-prod <- data.table(read_excel(fp(p_demand, "fao-saf-maize-soybean-prod.xlsx")))
-
-# countries included in Gasparri et al (2015)
-gasparri <- unique(ha$AreaName)[c(1:2, 7:8, 13, 15, 17, 19, 22)]
-har <- ha[AreaName %in% gasparri]
-setkey(har, "AreaName")
-prodr <- prod[AreaName %in% gasparri]
-setkey(prodr, "AreaName")
-# prodr[AreaName == "Zambia" & Year %in% 2012:2014, mean(Value), by = ItemName]
-# har[, unique(AreaName)]; prodr[, unique(AreaName)]
-
-haregion <- har[, mean(Value), by = .(ItemName, Year)]
-ha_rates <- sapply(c("Maize", "Soybeans"), function(i) {
-  haregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
-})
-prodregion <- prodr[, mean(Value), by = .(ItemName, Year)]
-prod_rates <- sapply(c("Maize", "Soybeans"), function(i) {
-  prodregion[ItemName == i, coef(lm(V1 / V1[1] ~ Year))[2]]
-})
-# prodregion[ItemName == "Maize", plot(Year, V1)]
-# prodregion[ItemName == "Soybeans", plot(Year, V1)]
-# haregion[ItemName == "Maize", plot(Year, V1)]
-# haregion[ItemName == "Soybeans", plot(Year, V1)]
-
-ha_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
-  haregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
-})
-prod_rates05 <- sapply(c("Maize", "Soybeans"), function(i) {
-  prodregion[ItemName == i & Year > 2004, coef(lm(V1 / V1[1] ~ Year))[2]]
-})
-# (1 + prod_rates05[2] * 13) * prodregion[ItemName == "Soybeans", V1]
-
-allrates <- rbind(1 + ha_rates * (2050-2014), 1 + prod_rates * (2050-2014), 
-                  1 + ha_rates05 * (2050-2014), 1 + prod_rates05 * (2050-2014))
-
-# a <- 2
-# b <- 10
-# (a + sum((b - a) / (2010 - 2000) * rep(1, 10))) / a
-# ((b - a) / (2010 - 2000))
-# 5 * 100
-# targets(c("maize" = 2), 10, 2)
-
-round(colMeans(allrates[c(2, 4), ]), 1)  ## by production
-round(colMeans(allrates[c(1, 3), ]), 1)  ## by area harvested
-round(colMeans(allrates), 1)  ## averaged across production and area harvested
-```
-
-### Yield gap closure on existing land
-
-We use the production based target as our potential scenario for Zambia.  Next is to factor in an assumption for how much demand will be met by closing current yield gaps
-```{r}
-setwd(p_mod)
-path <- "external/data/ZA"
-# load(fp(path, "currprod.rda"))
-# load("~/Dropbox/projects/agroEcoTradeoff/external/data/ZA/currprod.rda")
-# save(currprod, file = fp(path, "currprod.rda"))
-cp <- c("maize" = 2912004, "soy" = 203732)
-
-# gap closing function
-gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp
-
-# Current gaps
-mzgap <- 4.43 / 2.3 
-soygap <- 3.4 / 2
-
-# assume 50% closed on current land
-g <- gapper(cp, c(mzgap, soygap), 0.25)
-
-# targets given g
-targ <- c("maize" = 3, "soy" = 9)
-# targ <- c("maize" = 3.2, "soy" = 8.7)
-targs <- targets(targ, cp, g)
-
-```
+<a href="#top">Back to top</a>
 
@@
 
-Running the model individually for each crop. This currently entails updating and the `currprod.rda` file, and then reinvoking `input_handler`  
+Running the model individually for each crop and for both crops.
 ```{r, eval = FALSE}
-# currprod_switch <- function(cropind, path = p_moddat) {
-#   master_prod <- c("maize" = 2912004, "soy" = 203732)
-#   currprod <- master_prod[cropind]
-#   save(currprod, file = fp(path, "currprod.rda"))
-#   return(currprod)
-# }
+setwd(p_mod)
+path <- "external/data/ZA"
+
+# gaps 
+cp <- c("maize" = 2912004, "soy" = 203732)  # current production
+gapper <- function(cp, gaps, pct) ((((cp * gaps) - cp) * pct) + cp) / cp
+mzgap <- 4.43 / 2.3 
+soygap <- 3.4 / 2
+g <- gapper(cp, c(mzgap, soygap), 0.25)  # assume 50% closed on current land
+targ <- c("maize" = 3, "soy" = 9)  # targets given g
+targs <- targets(targ, cp, g)
 
@@
 
-# mzto$conv[, sum(maize)]
-# soyto$conv[, sum(soy)]
-
+# rasters
 CRSobj <- CRS(mzto$inputs$sp$crs)
@@
 # plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
-imp_plot2(mzsoyr, i1, p_figures, "ideal_yield2", "", crops, TRUE, 
+imp_plot2(mzsoyr, i1, p_figures, "ideal_yield", "", crops, 1, TRUE, 
          c("gold2", "green3", "orange"))
+
+basedt <- soyto$inputs$mask
+crops <- list(c("maize", "soy"))
+
+```
+<a href="#top">Back to top</a>
+
+```{r, eval = FALSE}
+load(fp(p_out, "mainresults1702.rda"))
+```
+
+<a href="#top">Back to top</a>
+
+## Figures
+
+How reconcilable are the different land use interests, and what would be the impacts of each?
+
+I am using the two crops combined approach because there is little overlap between the two.
+
+### Figure 1
+```{r, eval = FALSE}
+# select out results for "pure" cases (weight = 1)
+cnames <- c("Y", "C", "BD", "COST")
+pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))
+
+fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = pure, fullnames = TRUE)
+pure_r <- lapply(fnames, function(x) {  # x <- pure[1]
+  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
+  DTr <- dt_to_raster(DT, CRSobj)
+})
+
+# sum across crops and multiply by convertible areas
+rs <- lapply(pure_r, function(x) calc(x[[1:2]], sum) * x[[3]])  # actual area
+rs2 <- lapply(pure_r, function(x) calc(x[[1:2]], sum))  # conv pixels
+names(rs2) <- cnames
+# check
+# sapply(rs, function(x) cellStats(x, sum))
+# ybdc$both$optitab[ind %in% pure, Y]  # same, but C optimized out
+# sapply(rs2, function(x) cellStats(x, sum))
+# plot(pure_r[[3]])
+
+# figure out overlaps between different pure objectives
+pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
+pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
+pms2b$cl <- c("YC", "YBd", "CBd", "YCt", "CCt", "BdCt")
+rownames(pms2b) <- 1:6
+
+# pairwise overlaps
+rs_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+  calc(stack(rs2[which(pms2b[x, ] == 1)]), sum)
+})
+names(rs_over) <- pms2b$cl
+
+# Jaccard Index of overlaps
+rs_over_a <- sapply(1:length(rs_over), function(x) { # x <- 4
+  sapply(1:2, function(y) {  # y <- 3
+    r <- rs_over[[x]] == y
+    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
+  })
+})
+pms2b$JS <- round(rs_over_a[2, ] / colSums(rs_over_a), 3)
+
+# calculate nearest neighbor distance between each version, using spatstat 
+# nncross
+zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+                 unitname = "meter")
+which1 <- function(x) x == 1  # selector function for r to xy
+rsover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
+  ind <- which(pms2b[x, 1:4] == 1)
+  rl <- rs2[ind]
+  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+  pps <- lapply(rxy, function(x) {
+    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+  })
+  # reorder to find which has most points, to put that in as X
+  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+})
+pms2b$NND <- round(sapply(rsover_nn, function(x) mean(x$dist))) / 1000
+
+cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
+plot(1:4, cex = 3, pch = 20, col = cols_tr)
+
+# impact table
+
+# For Panel B - impact statistics 
+# bpath <- fp(p_out, ybdc$both$bcode)
+# load(fp(bpath, "impacts_tab.rda"))  # get impacts table
+
+# select out corresponding impact statistics, summarize
+conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
+impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) conv_stats[[x]])
+
+# Plot setup
+m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
+m2 <- matrix(rep(5, 20), 4, 5)
+m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
+lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
+
+# plot(1:10, ylab = ylabs[2])
+# ylabs <- c("Jaccard Index", 
+ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
+           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
+           "Intactness", expression(paste("FR converted (ha"^2, ")")))
+# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
+yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
+ftits <- list("Y 100%", "C 100%", "Biod 100%", "Cost 100%")
+leg <- c("Converted")
+
+yaxes <- list(seq(0, 14, 2), seq(0, 7, 0.5),               
+              seq(0, 1100, 100), seq(0, 0.12, 0.02), seq(0, 0.6, 0.1), 
+              seq(0, 1, 0.1), seq(0, 6500, 500))
+bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
+                  cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
+
+# plot
+tiff(fp(p_figures, "single_priorities2.tif"), height = 3.5, width = 7, 
+     units = "in", res = 300)
+l <- layout(mat = lmat)
+# layout.show(l)
+
+# map plots
+for(i in 1:4) { # i <- 1
+  imp_plot2(rs2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
+            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+            marg = rep(0, 4))
+}
+
+# impact plots
+mgp <- c(1.6, 0.25, 0)
+par(mar = c(4, 3, 0, 0))
+with(pms2b, {
+  plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
+       axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.3))
+  points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
+  axis(1, at = seq(0, 230, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
+       tcl = -0.2)
+  axis(2, at = seq(0, 0.3, 0.05), cex.axis = 0.6, las = 2, mgp = mgp, 
+       tcl = -0.2)
+  mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
+  mtext(text = "Jaccard Index", side = 2, line = 1.25, cex = 0.45)
+})
+mgp <- c(1, 0.25, 0)
+for(i in 1:7) {
+  v <- impsr[[i]] / yscale[i]
+  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
+               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
+               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
+               cex.lab = 0.7)
+  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
+       tcl = -0.2)
+}
+dev.off()
 
@@
 
-### Best Areas for Other Three Factors
+<a href="#top">Back to top</a>
+
+### Figures 2 & 3 (probably supplemental)
 ```{r, eval = FALSE}
-basedt <- soyto$inputs$mask
-ybl <- list(yb1 <- c(1, 1))
-step <- 0.05
-cnames <- c("C", "BD", "COST")
-ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2); head(ps)
-Yv = "conv_area"; Cv = "tot_C"; BDv = "int_mean"; COSTv = "mu_cost"
-Yst = sum; Cst = sum; BDst = mean; COSTst = mean
-crops <- list("maize", "soy", c("maize", "soy"))
-croplnms <- list("maize", "soy", "both")
-# lapply(crops, function(x) targ[x])
-# lapply(crops, function(x) g[x])
-cbdc <- lapply(crops, function(x) {
-  a <- Sys.time()
-  print(paste("Starting run for", x, "at", a))
-  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
-                 yblist = ybl, currprodmod = g[x], input_key = "ZA", 
-                 Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-                 Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-                 silent = FALSE, ncl = 7)
-  load(fp(p_out, "bcode.rda"))
-  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
-  print(Sys.time() - a)
-  c("bcode" = bcode, otab, "heat" = heatr)
-})
-names(cbdc) <- croplnms
+# individual maps
+png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
+     units = "in", res = 300)
+par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
+ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
+legp <- c(TRUE, rep(FALSE, 3))
+for(i in 1:4) {
+  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
+  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
+            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
+            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
+}
+dev.off()
 
-# check for datasets where certain weights were removed from contention
-clength <- t(sapply(cbdc, function(x) c(nrow(x$params), nrow(x$optitab))))
-cbdcmz <- heat2(cbdc$maize, fp(p_out, cbdc$maize$bcode), basedt, todisk = TRUE)
-cbdc$maize$heat <- cbdcmz
-
-# load(fp(p_out, "ZA_20160213_172518_13217_37779/impacts_tab.rda"))
-# save(cbdc, file = fp(p_out, "results.rda"))
-# par(mfrow = c(2, 2), mar = c(0, 0, 0, 4))
-# plot(cbdc$both$heat, axes = FALSE, box = FALSE)
-# plot(cbdc$maize$heat, axes = FALSE, box = FALSE)
-# plot(cbdc$soy$heat, axes = FALSE, box = FALSE)
-# a <- Sys.time()
-# print(a)
-# cbdc_mz <- pareto(cnames, step = step, prod_targ = targ["maize"], 
-#                   yblist = ybl, currprodmod = g["maize"], input_key = "ZA", 
-#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-#                   silent = FALSE, ncl = 8)
-# Sys.time() - a
-# load(fp(p_out, "bcode.rda"))
-# bcodes <- bcode
-# a <- Sys.time()
-# cbdc_soy <- pareto(cnames, step = step, prod_targ = targ["soy"], 
-#                    yblist = ybl, currprodmod = g["soy"], input_key = "ZA", 
-#                    Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-#                    Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-#                    silent = FALSE, ncl = 8)
-# Sys.time() - a
-# load(fp(p_out, "bcode.rda"))
-# bcodes[2] <- bcode
-# implist <- list(cbdc_mz, cbdc_soy)
-# names(implist) <- bcodes
-
-# tstmp <- function() {
-#   paste0(gsub(" ", "", gsub(":|-", "", as.character(Sys.time()))))
-# }
-# save(implist, file = fp(p_out, paste0("imptabs_", tstmp(), ".rda")))
-# load(fp(p_out, "imptabs_20160212165418.rda"))
-# implist
-
-# Create overlap maps
-# CRSobj <- CRS(soyto$inputs$sp$crs)
-# cbdc_mz_ht <- heat_func(fp(p_out, names(implist)[1]), basedt) 
-# cbdc_soy_ht <- heat_func(fp(p_out, names(implist)[2]), basedt) 
-# plot(cbdc_soy_ht + cbdc_mz_ht)
-# plot(cbdc_soy_ht)
-# Might need to run it for both crops simultaneously
-
-# Do RGB plot comparing overlaps between two
+# heat maps
+implist <- list(ycost$both, bdc$both)
+load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
+pacols <- c("transparent", "grey85", "grey75")
+png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
+par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
+for(i in 1:2) {
+  par(mar = c(1, 1, 1, 4))
+  plot(zambia, col = "grey90", border = "transparent")
+  ir <- implist[[i]]$heat
+  ir[ir == 0] <- NA
+  plot(ir, col = rev(heat.colors(20)), add = TRUE, )
+  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+       border = FALSE)
+  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+       border = FALSE)
+  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
+}
+dev.off()
 
@@
 
-### Conservation interests versus agricultural interests
+<a href="#top">Back to top</a>
+
+### Figures 4 (main) and 5 (probably supplemental)
+
+What happens when weight is progressively added
+
 ```{r, eval = FALSE}
-ybl <- list(yb1 <- c(1, 1))
-step <- 0.02
-cnames <- c("Y", "COST")
-ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
-# ps[which(rowSums(ps[, c(1, 3)]) == 0), ]
+p <- data.frame(ybdc$both$params)
+p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
+scens <- lapply(c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), function(x) {
+  c(1 - x, rep(x / 2, 2), 0)
+})
+pind <- sapply(scens, function(x) {
+  x <- round(x, 2)
+  which(p$Y == x[1] & p$C == x[2] & p$BD == x[3] & p$COST == x[4])
+})
 
-ycost <- lapply(crops, function(x) {
-  a <- Sys.time()
-  print(paste("Starting run for", x, "at", a))
-  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
-                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
-                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-                  silent = FALSE, ncl = 7)
-  load(fp(p_out, "bcode.rda"))
-  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
-  print(Sys.time() - a)
-  c("bcode" = bcode, otab, "heat" = heatr)
+# select out impacts corresponding to those runs
+cstats <- selected_impacts(pind, fp(p_out, ybdc$both$bcode))
+
+# check on selection
+# il <- input_handler(p_moddat, crops[[1]])
+# chk <- tradeoff_mod(prod_targ = targ, cbetas = ybdc$both$params[pind[5], 7:10], 
+#                     currprodmod = g, exist_list = il)
+# DF <- data.frame(chk$impacts)
+# round(c(colSums(DF[, 1:2]),
+#         sapply(4:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))),2)
+# cstats[5, ]
+# fnames <- dir(bpath, pattern = "csv")
+# fnames[pind[5]]
+
+# calculate max impacts across both crops, and figure out the impact range, 
+# and what percent that is relative to the worst case
+allstats <- selected_impacts(1:nrow(p), fp(p_out, ybdc$both$bcode))
+minmaxstats <- allstats[, lapply(.SD, range)]
+imp_range <- minmaxstats[, lapply(.SD, diff)]  # difference between best/worst
+pct_of_tot <- minmaxstats[, lapply(.SD, function(x) diff(x) / max(x))]
+
+dfunc <- function(x, y) 100 - y / x * 100
+stnms <- names(cstats)[-1]
+imp_pct <- cstats[, lapply(stnms, function(x) {
+  dfunc(minmaxstats[[x]][2], get(x))
+})]
+setnames(imp_pct, stnms)
+
+# plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
+#      axes = FALSE)        # xaxs = "r", yaxs = "r")
+
+sctab <- do.call(rbind, scens) * 100
+cx <- 0.7
+lcols <- c("black", "red", "blue")
+imp_pct[, {
+  png(fp(p_figures, "weight_impact2.png"), height = 4, width = 4, 
+      units = "in", res = 300)
+  par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
+  plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
+       axes = FALSE, ylim = c(0, 65), xlim = c(2, 12))#, xaxs = "r", yaxs = "r")
+  lines(conv_area, int_prior, col = "red")
+  lines(conv_area, mu_cost, col = "blue")
+  axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
+       tcl = -0.2, mgp = c(1, 0.25, 0))
+  axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
+       tcl = -0.15, mgp = c(1, 0.25, 0))
+  mtext("% land savings", side = 1, line = 1, cex = cx)
+  mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
+  inms <- c("tot_C", "int_prior", "mu_cost")
+  for(i in 1:length(inms)) {
+    text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
+         cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
+    text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
+         offset = i * 0.15 - 0.2, col = lcols[i])
+  }
+  legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
+         text.col = lcols, col = lcols, bty = "n", lty = 1)
+  dev.off()
+}]
+
+fnames <- fp(fp(p_out, ybdc$both$bcode), 
+             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
+scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
+  print(x)
+  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
+  DTr <- dt_to_raster(DT, CRSobj)
 })
-names(ycost) <- croplnms
 
-clength <- t(sapply(ycost, function(x) c(nrow(x$params), nrow(x$optitab))))
+scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
+names(scens_rs2) <- paste0("w", 1:length(scens_rs2))
 
-cnames <- c("C", "BD")
-bdc <- lapply(crops, function(x) {
-  a <- Sys.time()
-  print(paste("Starting run for", x, "at", a))
-  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
-                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
-                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-                  silent = FALSE, ncl = 7)
-  load(fp(p_out, "bcode.rda"))
-  heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
-  print(Sys.time() - a)
-  c("bcode" = bcode, otab, "heat" = heatr)
+pms2b <- do.call(rbind.data.frame, lapply(2:6, function(x) {
+  a <- rep(0, 6)
+  a[c(1, x)] <- 1
+  a
+}))
+colnames(pms2b) <- paste0("Var", 1:ncol(pms2b))
+pms2b$cl <- c("W010", "W020", "W030", "W040", "W050")
+rownames(pms2b) <- 1:5
+
+# pairwise overlaps
+scens_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+  calc(stack(scens_rs2[which(pms2b[x, ] == 1)]), sum)
 })
-names(bdc) <- croplnms
-# bdc3 <- c(bdc, bdc2)  # vestiges to show I forgot to do maize soy case
-# bdc <- bdc3
+names(scens_over) <- pms2b$cl
 
-bdcfix <- lapply(bdc, function(x) {  # x <- bdc$maize
-  if(nrow(x$params) > nrow(x$optitab)) {
-    print(paste("fixing heat map for", x$bcode))
-    heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE)
-  } else {
-    heatfix <- NULL
+# Jaccard Index of overlaps
+scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
+  sapply(1:2, function(y) {  # y <- 3
+    r <- scens_over[[x]] == y
+    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
+  })
+})
+pms2b$JS <- round(scens_over_a[2, ] / colSums(scens_over_a), 3)
+
+# calculate nearest neighbor distance between each version, using spatstat 
+# nncross
+zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+                 unitname = "meter")
+which1 <- function(x) x == 1  # selector function for r to xy
+scenover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 4
+  print(x)
+  ind <- which(pms2b[x, 1:6] == 1)
+  rl <- scens_rs2[ind]
+  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+  pps <- lapply(rxy, function(x) {
+    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+  })
+  # reorder to find which has most points, to put that in as X
+  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+})
+pms2b$NND <- round(sapply(scenover_nn, function(x) mean(x$dist))) / 1000
+
+cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
+plot(1:4, cex = 3, pch = 20, col = cols_tr)
+
+# plot(scens_rs2[[1]] + scens_rs2[[2]])
+# check
+# round(sapply(scens_rs2, function(x) {
+#   cellStats(x * scens_r[[1]]$convertible * 100, sum)
+# }))
+# cstats$conv_area
+
+# pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
+# pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid
+
+# rasters showing overlap as weight is progressively put on C and BD
+scens_over2 <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)
+
+# # overlap areas
+# scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
+#   sapply(1:3, function(y) {  # y <- 3
+#     r <- scens_over[[x]] == y
+#     # plot(r)
+#     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+#   })
+# })
+# colSums(scens_over_a[c(2, 3), ])
+# over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
+# par(mar = c(0, 0, 0, 0))
+# plot(1:3, pch = 20, col = rgb(1, 0, 1))
+# plot(zambia)
+
+# cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
+# cols <- c("red", "blue", "gold")
+bluecols <- sapply(c(1, 0.75, 0.5, 0.25, 0), function(x) rgb(0, x, 1))
+bluecols[5] <- rgb(0.4, 0, 1)
+cols <- sapply(1:length(bluecols), function(x) c("red", bluecols[x], "gold"))
+plot(1:ncol(cols), pch = 20, col = cols[2, ], cex = 20)
+ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
+              "Y 80% C 10% Biod 10% Cost 0%",
+              "Y 70% C 15% Biod 15% Cost 0%",
+              "Y 60% C 20% Biod 20% Cost 0%", 
+              "Y 50% C 25% Biod 25% Cost 0%")
+leg <- c("Y", "YCBd", "Y + YCBd")
+
+tiff(fp(p_figures, "changing_weights2.tif"), height = 3.5, width = 7, 
+     units = "in", res = 300)
+par(mfrow = c(2, 3), oma = c(0, 0, 0, 1))
+for(i in 1:5) { # i <- 1
+  # map
+  imp_plot2(scens_over2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+            crops = leg, ftitle = ftits[[i]], col = cols[, i], 
+            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+            marg = c(0, 0, 1, 0), titline = 0)
 }
-  heatfix
+mgp <- c(1, 0.25, 0)
+par(mar = c(3, 3, 1, 0))
+with(pms2b, {
+     plot(NND, JS, col = cols[2, ], pch = 16, cex = 1.5, 
+          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.4), 
+          xlim = c(0, 180))
+     points(NND, JS, col = cols[[1]][1], pch = 20, cex = 1)
+     axis(1, at = seq(0, 180, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
+     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
 })
-for(i in 1:length(bdc)) bdc[[i]]$heat <- bdcfix[[i]]
+dev.off()
 
-save(cbdc, ycost, bdc, file = fp(p_out, "mainresults1302.rda"))
-```
-
-### All 4 together
-```{r, eval = FALSE}
-step <- 0.05
-cnames <- c("Y", "C", "BD", "COST")
-ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
-ybdc <- lapply(crops, function(x) {
-  a <- Sys.time()
-  print(paste("Starting run for", x, "at", a))
-  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
-                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
-                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-                  silent = FALSE, ncl = 7)
-  load(fp(p_out, "bcode.rda"))
-  # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
-  print(Sys.time() - a)
-  c("bcode" = bcode, otab) #, "heat" = heatr)
-})
-names(ybdc) <- croplnms
-
-# ran because of fail on iteration for both (croplnms not properly compiled)
-# bnames <- c("ZA_20160213_231906_17404_28678", "ZA_20160213_234959_17404_39818")
-# ybdc <- lapply(bnames, function(x) {
-#   bpath <- fp(p_out, x)
-#   load(fp(bpath, "optitab.rda"))
-#   load(fp(bpath, "parms.rda"))
-#   list("bcode" = x, "optitab" = optimal, "params" = parms)
-# })
-# names(ybdc) <- croplnms[1:2]
-
-# ybdc2 <- lapply(crops[3], function(x) {
-#   a <- Sys.time()
-#   print(paste("Starting run for", x, "at", a))
-#   otab <- pareto(cnames, step = step, prod_targ = targ[x], 
-#                   yblist = ybl, currprodmod = g[x], input_key = "ZA", 
-#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-#                   silent = FALSE, ncl = 7)
-#   load(fp(p_out, "bcode.rda"))
-#   # heatr <- heat(fp(p_out, bcode), basedt, todisk = TRUE) 
-#   print(Sys.time() - a)
-#   c("bcode" = bcode, otab) #, "heat" = heatr)
-# })
-# names(ybdc2) <- croplnms[[3]]
-# ybdc3 <- c(ybdc, ybdc2)  # vestiges to show I forgot to do maize soy case
-# ybdc <- ybdc3
-
-# heat maps
-ybdcheat <- lapply(ybdc, function(x) {  # x <- bdc$maize
-  print(paste("N params =", nrow(x$params), ": N optimal =",  nrow(x$optitab)))
-  heatfix <- heat2(x, fp(p_out, x$bcode), basedt, todisk = TRUE, 
-                   overwrite = TRUE)
-  heatfix
-})
-for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcheat[[i]]
-
-plot(ybdc$maize$heat)
-par(mfrow = c(1, 2))
-plot(ycost$maize$heat)
-plot(ycost$soy$heat)
-plot(ycost$both$heat)
-plot(bdc$both$heat)
+#   # barplot
+#   par(mar = c(5, 4, 3, 3))
+#   overl <- scens_over_a[1:3, i][c(1, 3, 2)]
+#   bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
+#                 col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
+#                 ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
+#   axis(side = 1, at = bp, labels = leg, las = 2, 
+#        mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
+# }
+# dev.off()
 
 ```
+<a href="#top">Back to top</a>
+
+### Figure 6
+
+Objective weighting of outcomes, based on range of potential impacts
+
+```{r, eval = FALSE}
+# objective weights
+obj_wgts <- round(pct_of_tot[, {
+  c(conv_area, tot_C, int_mean, mu_cost) / 
+    sum(c(conv_area, tot_C, int_mean, mu_cost))
+}], 2)
+# round(obj_wgts, 2)
+names(obj_wgts) <- cnames
+
+# Objective weights
+toobj <- tradeoff_mod(targ, cbetas = obj_wgts, currprodmod = g, it = "OBJ")
+
+# ybdc$both$params[1, ]
+# yp <- ybdc$both$params[, 7:10]
+# which(round(yp[, 1], 2) == 0.05 & round(yp[, 2], 2) == 0.30 & 
+#         round(yp[, 3], 2) == 0.25 & round(yp[, 4], 2) == 0.4)
+
+# objective weights on ag only 
+# considering that these might be two sides of the same coin
+ag_wgts <- round(obj_wgts[c(1, 4)] / sum(obj_wgts[c(1, 4)]), 2)
+ag_wgts <- c(ag_wgts[1], "C" = 0, "BD" = 0, ag_wgts[2])
+ag_wgts2 <- ag_wgts[c(4, 2:3, 1)]
+names(ag_wgts2) <- names(ag_wgts)
+
+toag <- tradeoff_mod(targ, cbetas = ag_wgts, currprodmod = g, it = "AGOBJ")
+toag2 <- tradeoff_mod(targ, cbetas = ag_wgts2, currprodmod = g, it = "AGOBJ2")
+
+# subjective weights, where we want to treat everyone equally
+subj_wgts <- rep(0.25, 4)
+names(subj_wgts) <- cnames  # rename to make sure model runs properly
+tosubj <- tradeoff_mod(targ, cbetas = subj_wgts, currprodmod = g, it = "SUBJ")
+
+# extract impacts of these, and add in BAU case
+tores <- lapply(list(toag, toag2, toobj, tosubj), function(x) {  # x <- toobj
+  DTsum <- data.table(x$conv[, rowSums(.SD), .SDcols = c(crops[[1]])])
+  DTspat <- cbind(basedt[, .(x, y)], DTsum)
+  DTr <- dt_to_raster(DTspat, CRSobj)
+  DTr <- DTr + (fbr * 2)  # overlap with farmblocks, for output
+  list("map" = DTr, "imp" = x$impacts)
+})
+names(tores) <- c("ag", "ag2", "obj", "subj")
+
+# Calculate area of overlaps between farmblocks and 
+yfbr <- calc(mzsoytor, sum) + fbr * 2
+plot(yfbr)
+fb_over_a <- sapply(1:length(tores), function(x) { # x <- 4
+  sapply(1:3, function(y) {  # y <- 3
+    r <- tores[[x]]$map == y
+    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+  })
+})
+fb_sim <- fb_over_a[3, ] / colSums(fb_over_a)
+
+# grid for calculating overlaps
+pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
+pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
+pms2b$cl <- c("AgAg2", "AgObj", "Ag2Obj", "AgSubj", "Ag2Subj", "ObjSubj")
+
+toresr <- lapply(tores, function(x) {
+  r <- x$map
+  r[r == 2] <- 0
+  r[r == 3] <- 1
+  r
+})
+
+# tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+#   (toresr[[1]] == pms2b[x, 1]) & (toresr[[2]] == pms2b[x, 2]) & 
+#     (toresr[[3]] == pms2b[x, 3]) & (toresr[[4]] == pms2b[x, 4])
+# })
+# names(tores_over_r) <- pms2b$cl
+# tores_overs <- stack(tores_over)
+tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
+})
+names(tores_over_r) <- pms2b$cl
+
+# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
+tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
+  sapply(1:2, function(y) {  # y <- 3
+    r <- tores_over_r[[x]] == y
+    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+  })
+})
+pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)
+
+# calculate nearest neighbor distance between each version, using spatstat 
+# nncross
+# plot(zam_owin)
+zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+                 unitname = "meter")
+which1 <- function(x) x == 1  # selector function for r to xy
+tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
+  ind <- which(pms2b[x, 1:4] == 1)
+  rl <- toresr[ind]
+  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+  pps <- lapply(rxy, function(x) {
+    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+  })
+  # reorder to find which has most points, to put that in as X
+  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+})
+pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
+# with(pms2b, plot(NND, JS))
+
+# do sum factoring in overlaps 
+# sapply(tores_over_r, function(x) {
+#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
+# })
+
+cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
+plot(1:4, cex = 3, pch = 20, col = cols_tr)
+
+# impact table
+imp_list <- list(toag$impacts, toag2$impacts, toobj$impacts, tosubj$impacts)
+imps <- rbindlist(lapply(imp_list, function(x) selected_impacts2(x)))
+impsr <- lapply(c(1, 4, 2, 6:8, 3), function(x) imps[[x]])
+
+# layout matrix
+m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
+m2 <- matrix(rep(5, 20), 4, 5)
+m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
+lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
+
+# plot(1:10, ylab = ylabs[2])
+# ylabs <- c("Jaccard Index", 
+ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
+           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
+           "Intactness", expression(paste("FR converted (ha"^2, ")")))
+# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
+yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
+ftits <- list("Y 11% C 0% Biod 0% Cost 89%",
+              "Y 89% C 0% Biod 0% Cost 11%",
+              "Y 5% C 29% Biod 25% Cost 41%",
+              "Y 25% C 25% Biod 25% Cost 25%")
+leg <- c("Converted")
+
+# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
+yaxes <- list(seq(0, 14, 2), seq(0, 2.5, 0.5),               
+              seq(0, 1000, 200), seq(0, 0.2, 0.05), seq(0, 0.5, 0.1), 
+              seq(0, 0.8, 0.2), seq(0, 3500, 500))
+# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
+bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
+                cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
+# lcols <- c(list(c(cols_tr[2], cols_tr[3], cols_tr[3], cols_tr[4], cols_tr[4] 
+# lcols <- lapply(1:7, function(x) cols_tr)
+
+tiff(fp(p_figures, "objective_weights.tif"), height = 3.5, width = 7, 
+     units = "in", res = 300)
+l <- layout(mat = lmat)
+# layout.show(l)
+
+# map plots
+for(i in 1:4) { # i <- 1
+  # map
+  # plot(zambia)
+  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
+            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+            marg = rep(0, 4))
+}
+
+# impact plots
+mgp <- c(1, 0.25, 0)
+par(mar = c(4, 3, 0, 0))
+with(pms2b, {
+     plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
+          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 1))
+     points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
+     axis(1, at = seq(0, 130, 10), cex.axis = 0.6, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
+     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
+})
+for(i in 1:7) {
+  v <- impsr[[i]] / yscale[i]
+  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
+               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
+               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
+               cex.lab = 0.7)
+  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
+       tcl = -0.2)
+}
+dev.off()
+
+# ybdc$both$optitab[, {
+#   par(mfrow = c(2, 2), mar = rep(3, 4))
+#   hist(.SD, axes = FALSE, labels = FALSE)
+# }, .SDcol = 1:4]
+# 
+# frange <- function(x) diff(range(x)) / max(x)
+# fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
+# ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
+# ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
+# 
+# hist(ybdc$both$optimal$Y)
+
+```

### 2016-02-23 Rmd/main-analysis.Rmd
Separated figures onto new Rmd, and completely analysis up through weight selection as function of range in data relative to worst case.

--- 2016-02-18 Rmd/main-analysis.Rmd
@@
      file = full_path(p_mod, "external/data/ZA/parks_roads.rda"))
+```
 
-```
+<a href="#top">Back to top</a>
 
@@
 save(ybdc, ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
-# load(fp(p_out, "mainresults1402.rda"))
+# load(fp(p_out, "mainresults1702.rda"))
 ```
-
-## Figures
-
-How reconcilable are the different land use interests, and what would be the impacts of each?
-
-I am using the two crops combined approach because there is little overlap between the two.
-
-### Figure 1
-```{r, eval = FALSE}
-# select out results for "pure" cases (weight = 1)
-pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))
-
-fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = pure, fullnames = TRUE)
-pure_r <- lapply(fnames, function(x) {  # x <- pure[1]
-  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
-  DTr <- dt_to_raster(DT, CRSobj)
-})
-# sapply(1:4, function(x) which(ps[, x] == 1))
-# sum across crops and multiply by convertible areas
-rs <- lapply(pure_r, function(x) calc(x[[1:2]], sum) * x[[3]])  # actual area
-rs2 <- lapply(pure_r, function(x) calc(x[[1:2]], sum))  # conv pixels
-# check
-# sapply(rs, function(x) cellStats(x, sum))
-# ybdc$both$optitab[ind %in% pure, Y]  # same, but C optimized out
-# sapply(rs2, function(x) cellStats(x, sum))
-# plot(pure_r[[3]])
-
-# figure out overlaps between different pure objectives
-pms <- expand.grid(rep(list(0:1), 4))[-1, ]  # logic grid
-rownames(pms) <- 1:15
-rs_over <- lapply(1:nrow(pms), function(x) {  # apply logic tests
-  (rs2[[1]] == pms[x, 1]) & (rs2[[2]] == pms[x, 2]) & (rs2[[3]] == pms[x, 3]) &     (rs2[[4]] == pms[x, 4])
-})
-rs_overs <- stack(rs_over)
-
-# legend to show class of overlap
-pms$leg <- c("Y", "C", "YC", "Bd", "YBd", "BdC", "YBdC", "Ct", 
-             "YCt", "CCt", "YCCt", "BdCt", "YBdCt", "CBdCt", 
-             "YCBdCt")
-
-# frequency of overlaps in different categories
-# sum(pms[c(1, 3, 5, 7, 9, 11, 13, 15), "hist"])
-pms$hist <- sapply(1:nlayers(rs_overs), function(x) {
-  round(cellStats(rs_overs[[x]] * pure_r[[1]]$convertible, sum), 1)
-})
-
-# legend color
-pms$col <- c("dodgerblue3", "goldenrod2", "tomato1", "yellow", 
-             rep("tomato1", 2), "red3", "deepskyblue1", rep("tomato1", 2),
-             rep("red3", 5))
-
-# recode with unique number for overlap class, collapse to RasterLayer, 
-# simplify classes for plotting
-rs_over_rc <- stack(lapply(1:length(rs_over), function(x) rs_over[[x]] * x))
-rs_over_rc <- calc(rs_over_rc, sum)
-# plot(rs_over_rc == 6)
-rclmat <- cbind(0:15, c(0, 1, 3, 5, 4, 5, 5, 6, 2, 5, 5, 6, 5, 6, 6, 6))
-leg2 <- c("Yield", "Cost", "Carbon", "Biodiversity", "2 overlap", "3+ overlap")
-rs_over_rc2 <- raster::reclassify(rs_over_rc, rclmat)
-
-# plot(1:2, pch = 20, cex = 3, col = c("yellow", "yellow2"))
-
-# For Panel B - impact statistics 
-bpath <- fp(p_out, ybdc$both$bcode)
-load(fp(bpath, "impacts_tab.rda"))  # get impacts table
-
-# select out corresponding impact statistics, summarize
-pure2 <- unlist(lapply(pure, function(x) c(x, x+1)))
-vars <- c("iter", "conv_area", "tot_C", "fpa_conv", "mu_cost", "rarity",
-          "priority",  "int_prior", "int_mean")
-conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
-# conv_stats <- outtab[outtab$iter %in% pure, vars]
-# conv_stats <- data.frame(t(sapply(pure, function(x) {  # x <- pure[1]
-#   DF <- conv_stats[conv_stats$iter == x, ]
-#   c(colSums(DF[, 2:4]), 
-#     sapply(5:ncol(DF), function(y) weighted.mean(DF[, y], DF$conv_area)))
-# }))) 
-# colnames(conv_stats) <- vars[-1]
-
-# rescale for common plot
-svec <- c(100000, 10000000, 1000, 1, 0.1, 0.1, 0.1, 0.1)  # rescale factor
-conv_stats_n <- sweep(as.matrix(conv_stats)[, -1], MARGIN = 2, svec, "/")
-conv_stats_n <- conv_stats_n[, c(1, 4, 2, 6:ncol(conv_stats_n), 3)]
-
-# Plots
-# Panel A
-# cols <- RColorBrewer::brewer.pal(11, "RdYlBu")
-# cols <- colorspace::diverge_hsv(n = 11)
-# colsi <- sapply(c(0.5, 0.75, 1), function(x) rgb(1, 1, 0.9, alpha = x))
-cols <- c("dodgerblue3", "deepskyblue1", "goldenrod2", "yellow", "tomato1",
-          "red3")
-plot(1:6, pch = 20, cex = 3, col = cols)
-
-tiff(fp(p_figures, "single_priorities.tif"), height = 3, width = 8, 
-     units = "in", res = 300)
-par(mfrow = c(1, 3))
-imp_plot2(dtr = rs_over_rc2, impacts = i1, fpath = p_figures, 
-          fnm = "pure_overlap", cexscale = 0.6, ftitle = "", crops = leg2, 
-          stats = FALSE, cols = cols, hold = TRUE, marg = c(0, 0, 0, 0))
-
-# Panel B
-# totf <- cellStats(rs_over_rc > 0, sum)
-par(mar = c(7, 4, 2, 2))
-bp <- barplot(pms$hist / sum(pms$hist) * 100, ylim = c(0, 25), las = 2, 
-              col = pms$col, mgp = c(2, 0.25, 0), tcl = -0.2, 
-              ylab = "% converted area")
-axis(side = 1, at = bp, labels = pms$leg, las = 2, mgp = c(2, 0.25, 0), 
-     tcl = -0.2)
-# TeachingDemos::subplot( 
-#   bp <- barplot(pms$hist / sum(pms$hist) * 100, ylim = c(0, 25), las = 2, 
-#                 col = pms$col, mgp = c(2, 0.25, 0), tcl = -0.2, 
-#                 ylab = "% converted pixels", names.arg = pms$leg), 
-#   x=grconvertX(c(0.1,0.45), from='npc'),
-#   y=grconvertY(c(0.6,1), from='npc'),
-#   type='fig', pars=list(mar=c(3,3,0,0) + 0.1))
-
-# Panel C
-# barplot(conv_stats_n)
-blabs <- c("Area", "Time", "Carbon", "Priority", "Priority/intactness", 
-           "Intactness", "Forest reserves")
-# blabs <- sub(':', '\n', blabs)
-tlabs <- c(expression(10^5~"km"^2), "hours", expression(10^7~"tons"), 
-           rep("X10", 3), expression(10^3~"km"^2))
-bp <- barplot(conv_stats_n, beside = TRUE, ylim = c(0, 14), las = 2, 
-              col = pms$col[c(1, 2, 4, 8)], names.arg = rep("", length(blabs)), 
-              ylab = "Impact*", mgp = c(2, 0.25, 0), tcl = -0.2)
-text(x = bp[2, ] - 0.4, y = rep(-0.3, length(bp[1, ])), adj = 1, 
-     labels = blabs, xpd = NA, srt = 90)
-text(x = bp[3, ] + 0.4, y = rep(-0.3, length(bp[3, ])), adj = 1, 
-     labels = tlabs, xpd = NA, srt = 90, cex = 0.8)
-dev.off()
-
-#### Need to redo pareto to pass in weighted mean function for all average 
-#### costs (DONE)
-
-```
-
-## Figure 2 & 3 (probably supplemental)
-```{r, eval = FALSE}
-# yc_bdc_st <- as.matrix(stack(ycost$both$heat, bdc$both$heat) / 100)
-# yc_bdc_st <- stack(ycost$both$heat, ycost$both$heat * 0, bdc$both$heat) / 100
-# plot(ycost$both$heat > 1)
-# plotRGB(yc_bdc_st, r = 1, g = 2, b = 3, colNA = "transparent", scale = 1)
-# rgb(red = yc_bdc_st[, 1], rep(0, nrow(yc_bdc_st)), yc_bdc_st[, 2])
-# plot(ycost$both$heat)
-
-# individual maps
-png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
-     units = "in", res = 300)
-par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
-ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
-legp <- c(TRUE, rep(FALSE, 3))
-for(i in 1:4) {
-  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
-  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
-            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
-            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
-}
-dev.off()
-
-# heat maps
-implist <- list(ycost$both, bdc$both)
-load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
-pacols <- c("transparent", "grey85", "grey75")
-png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
-par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
-for(i in 1:2) {
-  par(mar = c(1, 1, 1, 4))
-  plot(zambia, col = "grey90", border = "transparent")
-  ir <- implist[[i]]$heat
-  ir[ir == 0] <- NA
-  plot(ir, col = rev(heat.colors(20)), add = TRUE, )
-  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
-       border = FALSE)
-  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
-       border = FALSE)
-  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
-}
-dev.off()
-
-```
-
-## Figure 4
-
-Value of just a bit of compromise
-
-```{r, eval = FALSE}
-p <- data.frame(ybdc$both$params)
-outtab[1, ]
-p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
-# scens <- list(c(0.9, 0.05, 0.05, 0), 
-#               c(0.85, 0.05, 0.05, 0.05), 
-#               c(0.1, 0.45, 0.45, 0.0), 
-#               c(0.05, 0.45, 0.45, 0.05))
-scens <- lapply(c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9), function(x) {
-  c(1 - x, rep(x / 2, 2), 0)
-})
-# scens <- lapply(c(0, 0.15, 0.3, 0.45, 0.6, 0.75), function(x) {
-#   c(1 - x, rep(x / 3, 3))
-# })
-pind <- sapply(scens, function(x) {
-  x <- round(x, 2)
-  which(p$Y == x[1] & p$C == x[2] & p$BD == x[3] & p$COST == x[4])
-})
-
-# select out impacts corresponding to those runs
-cstats <- selected_impacts(pind, p_out, ybdc$both$bcode)
-
-# check on selection
-# il <- input_handler(p_moddat, crops[[1]])
-# chk <- tradeoff_mod(targ, ybdc$both$params[pind[5], 7:10], 
-#                     currprodmod = g, exist_list = il)
-# DF <- data.frame(chk$impacts)
-# outtab[outtab$iter == pind[5], -2]
-# round(c(colSums(DF[, 2:4]),
-#         sapply(5:ncol(DF), function(z) weighted.mean(DF[, z], DF$conv_area))), 
-#       2)
-# cstats[5, ]
-outtab[outtab$iter == pind[5], ]
-outtab[, 1:4][1:100,]
-
-fnames <- dir(bpath, pattern = "csv")
-fnames[pind[5]]
-
-# calculate max impacts across both crops, and figure out the impact range, 
-# and what percent that is relative to the worst case
-allstats <- selected_impacts(1:nrow(p), p_out, ybdc$both$bcode)
-minmaxstats <- allstats[, lapply(.SD, range)]
-imp_range <- minmaxstats[, lapply(.SD, diff)]  # difference between best/worst
-pct_of_tot <- minmaxstats[, lapply(.SD, function(x) diff(x) / max(x))]
-
-dfunc <- function(x, y) 100 - y / x * 100
-stnms <- names(cstats)[-1]
-imp_pct <- cstats[, lapply(stnms, function(x) {
-  dfunc(minmaxstats[[x]][2], get(x))
-})]
-setnames(imp_pct, stnms)
-
-plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
-     axes = FALSE)        # xaxs = "r", yaxs = "r")
-
-sctab <- do.call(rbind, scens) * 100
-cx <- 0.7
-lcols <- c("black", "red", "blue")
-imp_pct[, {
-  png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
-      units = "in", res = 300)
-  par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
-  plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
-       axes = FALSE, ylim = c(0, 65), xlim = c(2, 12))#, xaxs = "r", yaxs = "r")
-  lines(conv_area, int_prior, col = "red")
-  lines(conv_area, mu_cost, col = "blue")
-  axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
-       tcl = -0.2, mgp = c(1, 0.25, 0))
-  axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
-       tcl = -0.15, mgp = c(1, 0.25, 0))
-  mtext("% land savings", side = 1, line = 1, cex = cx)
-  mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
-  inms <- c("tot_C", "int_prior", "mu_cost")
-  for(i in 1:length(inms)) {
-    text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
-         cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
-    text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
-         offset = i * 0.15 - 0.2, col = lcols[i])
-  }
-  legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
-         text.col = lcols, col = lcols, bty = "n", lty = 1)
-  dev.off()
-}]
-
-scens_r <- lapply(pind, function(x) {  # x <- pind[[1]]
-  print(x)
-  inm <- paste0(outtab[outtab$iter == x, "rc"][1], ".csv")
-  fpath <- fp(fp(p_out, ybdc$both$bcode), inm)
-  # fname <- dir(bpath, pattern = "csv", full.names = TRUE)[x]
-  DT <- cbind(fread(fpath), basedt[, .(x, y, convertible)])
-  DTr <- dt_to_raster(DT, CRSobj)
-})
-
-scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
-# freq((rs2$Y + scens_rs2[[1]]) * scens_r[[1]]$convertible
-# plot(scens_rs2[[1]] + scens_rs2[[2]])
-
-# round(sapply(scens_rs2, function(x) {
-#   cellStats(x * scens_r[[1]]$convertible * 100, sum)
-# }))
-# cstats$conv_area
-
-# pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
-# pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid
-
-# rasters showing overlap as weight is progressively put on C and BD
-scens_over <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)
-
-# overlap areas
-scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
-  sapply(1:3, function(y) {  # y <- 3
-    r <- scens_over[[x]] == y
-    # plot(r)
-    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
-  })
-})
-sum(scens_over_a[c(2, 3), 4])
-# over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
-# par(mar = c(0, 0, 0, 0))
-# plot(1:3, pch = 20, col = rgb(1, 0, 1))
-# plot(zambia)
-
-cols <- c("grey", c(rgb(0, 1, 0), rgb(0, 0, 1), rgb(1, 0, 1)))
-ftits <- c("Y=0.9")
-par(mfrow = c(1, 4))
-for(i in 1:4) {
-  plot(zambia)
-  imp_plot2(scens_over[[i]], impacts = i1, fpath = p_figures, fnm = "bob", 
-            col = cols, add = TRUE, 
-            legend = FALSE)
-}
-plot(scens_over[[1]], col = c("transparent", over_cols[[5]]))
-plot(scens_rs2[[1]])
-plot(scens_over[[4]])
-plot(calc(stack(scens_over), sum))
-
-plot(calc(stack(scens_rs2[2:4]), sum))
-
-ybdc$both$optimal[, {
-  par(mfrow = c(2, 2), mar = rep(3, 4))
-  hist(.SD, axes = FALSE, labels = FALSE)
-}, .SDcol = 1:4]
-
-frange <- function(x) diff(range(x)) / max(x)
-fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
-ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
-ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
-
-hist(ybdc$both$optimal$Y)
-
-```
+<a href="#top">Back to top</a>

### 2016-02-24 Rmd/figures.Rmd
Updated results with new figure 3. Fixed small mistake in Figure 2, which was to select impact range from optimized results, not all results (very tiny difference).

@@
 p <- data.frame(ybdc$both$params)
+pare <- ybdc$both$optitab
 p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
@@
 # and what percent that is relative to the worst case
-allstats <- selected_impacts(1:nrow(p), fp(p_out, ybdc$both$bcode))
+allstats <- selected_impacts(pare$ind, fp(p_out, ybdc$both$bcode))
 minmaxstats <- allstats[, lapply(.SD, range)]
@@
 imp_pct[, {
-  png(fp(p_figures, "weight_impact2.png"), height = 4, width = 4, 
+  png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
       units = "in", res = 300)
@@
 
-Objective weighting of outcomes, based on range of potential impacts
+"Objective" subjective weighting of outcomes, based on different preferences.
+The examples here are intended to understand the impacts that will happen when different actors with different land use interests are willing to make some compromise, e.g.
+they are willing to give up 5-10% of the value of their best case cost, whereas the other parties must be willing to pay more.
+Do that for each of the 4, and then compare to the case where each gets equal weight.
+
+```{r, eval = FALSE}
+# use minmax stats from optimal solution set
+# pare <- ybdc$both$optitab
+optirange <- pare[, lapply(.SD, range), .SDcols = 1:4]
+# optirange[1, lapply(.SD, function(x) x * 1.05)]
+
+# y_comp <- sweep(as.matrix(pare)[, 1:4], MARGIN = 2, swvec, `<=`)
+cvar <- "C"
+compromise <- function(pare, optirange, cvar, cost1 = 1.05, cost2) {  
+  sval <- optirange[1, (eval(parse(text = cvar))) * cost1]
+  pare_ss <- as.matrix(pare[(eval(parse(text = cvar))) <= sval])
+  vars <- names(pare)[!names(pare) %in% c(cvar, "ind")]
+  swvec <- as.matrix(optirange[1, vars, with = FALSE])[1, ]
+  ocosts <- sweep(pare_ss[, vars], MARGIN = 2, swvec, "/")
+  # cost2 <- seq(1, 2, 0.05)
+  # pare_ss[2, vars][3] / swvec[3]
+  # ocosts[2, ]
+
+  chks <- sapply(cost2, function(x) {  # x <- cost2[13]
+    chkvals <- ifelse(ocosts < x, 1, 0)
+    chkvec <- which(rowSums(chkvals) == 3)
+    if(length(chkvec) == 1) {
+      ochks <- chkvec
+    } else if(length(chkvec) > 1) {
+      ochks <- chkvec[which.min(rowSums(ocosts[chkvec, ]))]  # select minimum
+    } else {
+      ochks <- 0
+    }
+    ochks
+  })
+  # pare_ss[269, ]
+  # ocosts[269, ]
+  # pare_ss[269, "COST"] / optirange[1, COST]
+  # optirange[1, Y] * 1.05
+  c("cost" = cost2[chks > 0][1], pare_ss[chks[chks > 0][1], "ind"])
+}
+
+# indices of costs in other three variables
+cost2 <- seq(1, 5, 0.05)
+y_comp <- compromise(pare, optirange, cvar = "Y", cost1 = 1.05, cost2 = cost2)
+c_comp <- compromise(pare, optirange, cvar = "C", cost1 = 1.05, cost2 = cost2)
+bd_comp <- compromise(pare, optirange, cvar = "BD", cost1 = 1.05, cost2 = cost2)
+ct_comp <- compromise(pare, optirange, cvar = "COST", cost1 = 1.05, cost2=cost2)
+# pare[ind == y_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
+# pare[ind == c_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
+# pare[ind == bd_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
+# pare[ind == ct_comp["ind"], .(Y, C, BD, COST)] / optirange[1, ]
+
+# select out those values from the results tables and grab conversion areas
+parms <- round(ybdc$both$params[, 7:10], 2)
+eqind <- which(apply(parms, 1, function(x) all(x == 0.25))) # iter of eq wt run
+# pare[ind == eqind]
+
+pind <- c(y_comp["ind"], c_comp["ind"], bd_comp["ind"], ct_comp["ind"], eqind)
+# round(ybdc$both$params[pind, 7:10], 2)
+fnames <- fp(fp(p_out, ybdc$both$bcode), 
+             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
+objsubj_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
+  print(x)
+  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
+  DTr <- dt_to_raster(DT, CRSobj)
+})
+names(objsubj_r) <- c("ycomp", "ccomp", "bdcomp", "ctcomp", "eqwt")
+
+# run for subjective weights (could have pull this out also) as check
+# run for subjective weights, where we want to treat everyone equally
+# eq_wgts <- rep(0.25, 4)
+# names(eq_wgts) <- cnames  # rename to make sure model runs properly
+# tosubj <- tradeoff_mod(targ, cbetas = eq_wgts, currprodmod = g, it = "SUBJ")
+# tosubj_imp <- selected_impacts2(tosubj$impacts)
+
+# tosubj_imp <- lapply(c(1, 4, 2, 6:8, 3), function(x) tosubj_imp[[x]])
+
+# pull out the impact tables (brief check shows they are the same values)
+all_stats <- selected_impacts(unname(pind), fp(p_out, ybdc$both$bcode))
+impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) all_stats[[x]])
+# pare[ind == y_comp["ind"],]; pare[ind == c_comp["ind"],]; 
+# pare[ind == bd_comp["ind"],]; pare[ind == ct_comp["ind"],]
+# tosubj_imp
+
+# grid for calculating overlaps
+pms2a <- expand.grid(rep(list(0:1), 5))[-1, ] 
+pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
+pms2b$cl <- c("YC", "YBD", "CBd", "YCt", "CCt", "BDCt", "YEq", "CEq", 
+              "BdEq", "CtEq")
+
+toresr <- lapply(objsubj_r, function(x) calc(x[[1:2]], sum))
+tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
+})
+names(tores_over_r) <- pms2b$cl
+
+# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
+tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
+  sapply(1:2, function(y) {  # y <- 3
+    r <- tores_over_r[[x]] == y
+    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+  })
+})
+pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)
+
+# calculate nearest neighbor distance between each version, using spatstat 
+# nncross
+# plot(zam_owin)
+zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+                 unitname = "meter")
+which1 <- function(x) x == 1  # selector function for r to xy
+tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
+  ind <- which(pms2b[x, 1:5] == 1)  ### check col number
+  rl <- toresr[ind]
+  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+  pps <- lapply(rxy, function(x) {
+    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+  })
+  # reorder to find which has most points, to put that in as X
+  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+})
+pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
+# with(pms2b, plot(NND, JS))
+
+# do sum factoring in overlaps 
+# sapply(tores_over_r, function(x) {
+#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
+# })
+
+# cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
+cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.5, 0), 
+             "gold")#rgb(1, 1, 0))
+plot(1:length(cols_tr), cex = 3, pch = 20, col = cols_tr)
+
+# layout matrix
+m1 <- do.call(cbind, lapply(1:5, function(x) matrix(rep(x, 45), 6, 9)))
+m2 <- matrix(rep(6, 36), 4, 9)
+m3 <- do.call(cbind, lapply(7:13, function(x) matrix(rep(x, 20), 4, 5)))
+f1 <- lapply(14:15, function(x) cbind(rep(x, 4), rep(x, 4)))
+lmat <- rbind(m1, cbind(m2, m3, rep(14, 4)))
+# lmat <- rbind(m1, m3)
+# cbind(m2, m3)
+
+# plot(1:10, ylab = ylabs[2])
+# ylabs <- c("Jaccard Index", 
+ylabs <- c(expression(paste("Conv. area (", 10^5, " ", km^2, ")")), 
+           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
+           "Intactness", expression(paste("FR converted (ha"^2, ")")))
+# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
+yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
+
+# figure titles
+ftits <- lapply(pind, function(x) {
+  paste(sapply(1:4, function(y) {
+    paste(paste(cnames[y], paste0(parms[x, y] * 100, "%")))
+  }), collapse = " ")
+})
+leg <- c("Converted")
+
+# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
+maxes <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][2])
+mins <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][1])
+ymaxes <- maxes / yscale
+ymins <- mins / yscale
+yaxes <- list(seq(0, 15, 3), 0:7, seq(0, 1100, 100), seq(0, 0.15, 0.05), 
+              seq(0, 0.6, 0.1), seq(0, 1, 0.1), seq(0, 7000, 1000))
+
+# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
+pms2b$icol <- cols_tr[c(1, 1, 2, 1, 2, 3, 1, 2, 3, 4)]  # circle inner
+pms2b$ocol <- cols_tr[c(2, 3, 3, 4, 4, 4, 5, 5, 5, 5)]  # circle outer
+
+tiff(fp(p_figures, "objective-subjective_weights.tif"), height = 2.5, width = 7, 
+     units = "in", res = 300)
+l <- layout(mat = lmat)
+# layout.show(l)
+
+# map plots
+for(i in 1:length(toresr)) { # i <- 1
+  # map
+#   par(mar = rep(0, 4))
+#   plot(zambia, col = cols_tr[i])
+  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
+            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
+            marg = rep(0, 4), inset = c(0.1, 0.1, 0, 0))
+}
+
+# impact plots
+mgp <- c(0.9, 0.25, 0)
+par(mar = c(3, 3, 0, 3))
+with(pms2b, {
+     plot(NND, JS, col = pms2b$icol, pch = 16, cex = 0.8, 
+          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.6), 
+          xlim = c(0, 25))
+     points(NND, JS, col = pms2b$ocol, pch = 20, cex = 0.5)
+     axis(1, at = seq(0, 25, 5), cex.axis = 0.5, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     axis(2, at = seq(0, 0.6, 0.1), cex.axis = 0.5, las = 2, mgp = mgp, 
+          tcl = -0.2)
+     mtext(text = "NN Distance (km)", side = 1, line = 0.5, cex = 0.4)
+     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.4)
+})
+par(mar = c(3, 3, 0, 0))
+mgp <- c(1.3, 0.25, 0)
+for(i in 1:7) { # i <- 1
+  v <- impsr[[i]] / yscale[i]
+  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
+               mgp = mgp, tcl = -0.2, yaxt = "n", 
+               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.4,
+               cex.lab = 0.55)
+  lines(c(-1, 7), rep(ymins[i], 2), lwd = 1, col = "grey50")
+  lines(c(-1, 7), rep(ymaxes[i], 2), lwd = 1)
+  axis(2, at = yaxes[[i]], cex.axis = 0.5, las = 2, mgp = mgp, 
+       tcl = -0.2)
+}
+dev.off()
+
+```
+
+# Earlier version of figure 3
 

### 2016-02-26 Rmd/figures.Rmd
Near final figures for main text done after revision to Fig 1 and 3, and new Figure 4 made. Model rerun for 100% yield gap closure on existing land.

@@
 
-# Main Analyses
+# Figures
 
@@
 soygap <- 3.4 / 2
-g <- gapper(cp, c(mzgap, soygap), 0.25)  # assume 50% closed on current land
-targ <- c("maize" = 3, "soy" = 9)  # targets given g
+g <- gapper(cp, c(mzgap, soygap), 1)  # assume 50% closed on current land
+targ <- c("maize" = 4, "soy" = 10)  # targets given g
 targs <- targets(targ, cp, g)
@@
 # plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
-imp_plot2(mzsoyr, i1, p_figures, "ideal_yield", "", crops, 1, TRUE, 
-         c("gold2", "green3", "orange"))
+imp_plot2(mzsoyr, i1, p_figures, "ideal_yield2", "", crops, 1.2, TRUE, 
+         c("red", "blue1", "gold"))
 
@@
 
-```{r, eval = FALSE}
-load(fp(p_out, "mainresults1702.rda"))
-```
-
 <a href="#top">Back to top</a>
@@
 ```{r, eval = FALSE}
+load(fp(p_out, "mainresults2502.rda"))
+
 # select out results for "pure" cases (weight = 1)
@@
 pure <- sapply(cnames, function(x) which(round(ybdc$both$params[, x], 4) == 1))
+
+# check on dominated status
+# ybdc$both$optitab[ind %in% 20:22]  # pure C not in optimum table
+sic <- selected_impacts(20:22, fp(p_out, ybdc$both$bcode))
+# (1 - sic$tot_C[1] / sic$tot_C[2]) * 100  # C is .07 more (dominated)
+# probably because of over-shooting target
+# ctst <- tradeoff_mod(prod_targ = targ, 
+#                      cbetas = c("Y" = 0, "C" = 1, "BD" = 0, "COST" = 0),
+#                      currprodmod = g)  
+# selected_impacts2(ctst$impacts)  # correct one selected
+# ybdc$both$params[20:22, ]
 
@@
 })
-pms2b$JS <- round(rs_over_a[2, ] / colSums(rs_over_a), 3)
+pms2b$JS <- round(rs_over_a[2, ] / colSums(rs_over_a) * 100, 2)
 
@@
 conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode))
-impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) conv_stats[[x]])
+# impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) conv_stats[[x]])
+impsr <- lapply(c(2, 3, 8, 5), function(x) conv_stats[[x]])
 
@@
 m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
-m2 <- matrix(rep(5, 20), 4, 5)
-m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
-lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
+m2 <- do.call(cbind, lapply(c(5, 8, 11, 14), function(x) {
+  a <- matrix(rep(x, 16), 4, 4)
+  b <- matrix(rep(x + 1, 6), 2, 3)
+  cbind(a, rbind(b, b + 1))
+}))
+
+# m2 <- matrix(rep(5, 20), 4, 5)
+# m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
+# lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
+lmat <- rbind(m1, m2)
 
@@
 ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
-           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
-           "Intactness", expression(paste("FR converted (ha"^2, ")")))
+           "Carbon loss (Mt)", #"Priority", 
+           "Biodiversity Index", "Travel time (hours)")#,
+           # "Intactness", expression(paste("FR converted (ha"^2, ")")))
 # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
-yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
-ftits <- list("Y 100%", "C 100%", "Biod 100%", "Cost 100%")
-leg <- c("Converted")
-
-yaxes <- list(seq(0, 14, 2), seq(0, 7, 0.5),               
-              seq(0, 1100, 100), seq(0, 0.12, 0.02), seq(0, 0.6, 0.1), 
-              seq(0, 1, 0.1), seq(0, 6500, 500))
-bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
-                  cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
+yscale <- c(10^5, 100000, 1, 1)#, 1, 1, 1)
+# ftits <- list("Y 100%", "C 100%", "Biod 100%", "Cost 100%")
+ftits <- list("Maximize Yields", "Minimize Carbon Loss", 
+              "Minimize Biodiversity Loss", "Minimize Travel Time")
+# leg <- c("Converted")
+
+yaxes <- list(seq(0, 16, 2), seq(0, 1200, 100), seq(0, 0.6, 0.1), 
+              seq(0, 7, 1))#,               
+              # seq(0, 1200, 100), #seq(0, 0.12, 0.02)), 
+              #seq(0, 1, 0.1), seq(0, 6500, 500))
+# bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
+#                   cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
 
 # plot
-tiff(fp(p_figures, "single_priorities2.tif"), height = 3.5, width = 7, 
+# tiff(fp(p_figures, "single_priorities2.tif"), height = 3.5, width = 7, 
+png(fp(p_figures, "single_priorities2.png"), height = 3.5, width = 7, 
     units = "in", res = 300)
+# postscript(fp(p_figures, "single_priorities2.eps"), horizontal = FALSE, 
+#            height = 3.5, width = 7)
 l <- layout(mat = lmat)
@@
 # map plots
+legtf <- c(FALSE, FALSE, FALSE, FALSE)
 for(i in 1:4) { # i <- 1
+  # par(mar = rep(0, 4))
+  # plot(zambia, col = cols_tr[i])
   imp_plot2(rs2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
-            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
-            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+            crops = "", ftitle = ftits[[i]], col = cols_tr[[i]], 
+            legend = legtf[i], hold = TRUE, stats = FALSE, cexscale = 0.4, 
             marg = rep(0, 4))
 }  
-
 # impact plots
-mgp <- c(1.6, 0.25, 0)
-par(mar = c(4, 3, 0, 0))
-with(pms2b, {
-  plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
-       axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.3))
-  points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
-  axis(1, at = seq(0, 230, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
-       tcl = -0.2)
-  axis(2, at = seq(0, 0.3, 0.05), cex.axis = 0.6, las = 2, mgp = mgp, 
-       tcl = -0.2)
-  mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
-  mtext(text = "Jaccard Index", side = 2, line = 1.25, cex = 0.45)
-})
+# mgp <- c(1.6, 0.25, 0)
+# par(mar = c(4, 3, 0, 0))
+# with(pms2b, {
+#   plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
+#        axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.3))
+#   points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
+#   axis(1, at = seq(0, 230, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
+#        tcl = -0.2)
+#   axis(2, at = seq(0, 0.3, 0.05), cex.axis = 0.6, las = 2, mgp = mgp, 
+#        tcl = -0.2)
+#   mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
+#   mtext(text = "Jaccard Index", side = 2, line = 1.25, cex = 0.45)
+# })
 mgp <- c(1, 0.25, 0)
-for(i in 1:7) {
+# for(i in 1:7) {
+cx <- 0.6
+for(i in 1:4) {
   v <- impsr[[i]] / yscale[i]
+  par(mar = c(2, 3.5, 2, 0))
   b <- barplot(impsr[[i]] / yscale[i], las = 2, col = cols_tr, 
-               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
-               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
+               mgp = c(1, 0.25, 0), tcl = -0.2, yaxt = "n", 
+               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = cx,
                cex.lab = 0.7)
-  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
-       tcl = -0.2)
+  axis(2, at = yaxes[[i]], cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
+  # mtext("Impacts", side = 3, cex = 1)
+  ind <- which(pms2b[, paste0("Var", i)] == 1)
+  with(pms2b[ind, ], {
+    par(mar = c(0.5, 3, 2, 1), mgp = mgp)
+    barplot(JS, col = cols_tr[-i], ylim = c(0, 30), yaxt = "n", 
+            ylab = "% Overlap", cex.lab = cx)
+    axis(2, at=seq(0, 30, 10), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
+    # mtext("Promixity", side = 3, cex = 1)
+    par(mar = c(2, 3, 0.5, 1), mgp = mgp)
+    barplot(NND, col = cols_tr[-i], ylim = c(0, 250), las = 2, yaxt = "n", 
+            ylab = "", cex.lab = cx)
+    axis(2, at=seq(0, 250, 50), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
+    mtext("Average distance", side = 2, line = 1.72, cex = 0.4)
+    mtext("to neighbor (km)", side = 2, line = 1.2, cex = 0.4)
+  })
 }
+par(xpd = NA)
+legx <- grconvertX(0.4, from = "ndc", to = "user")
+legy <- grconvertY(0.5, from = "ndc", to = "user")
+legend(x = legx, y = legy, legend = "Protected Areas", pch = 15, col = "grey75", 
+       bty = "n", pt.cex = 3 * 0.6, cex = 1.5 * 0.6)
 dev.off()
@@
 # heat maps
-implist <- list(ycost$both, bdc$both)
-load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
-pacols <- c("transparent", "grey85", "grey75")
-png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
-par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
-for(i in 1:2) {
-  par(mar = c(1, 1, 1, 4))
-  plot(zambia, col = "grey90", border = "transparent")
-  ir <- implist[[i]]$heat
-  ir[ir == 0] <- NA
-  plot(ir, col = rev(heat.colors(20)), add = TRUE, )
-  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
-       border = FALSE)
-  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
-       border = FALSE)
-  plot(roads, add = TRUE, lwd = 0.1, col = "grey")
-}
-dev.off()
+# implist <- list(ycost$both, bdc$both)
+# load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
+# pacols <- c("transparent", "grey85", "grey75")
+# png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
+# par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
+# for(i in 1:2) {
+#   par(mar = c(1, 1, 1, 4))
+#   plot(zambia, col = "grey90", border = "transparent")
+#   ir <- implist[[i]]$heat
+#   ir[ir == 0] <- NA
+#   plot(ir, col = rev(heat.colors(20)), add = TRUE, )
+#   plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+#        border = FALSE)
+#   plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+#        border = FALSE)
+#   plot(roads, add = TRUE, lwd = 0.1, col = "grey")
+# }
+# dev.off()
 
@@
 cstats <- selected_impacts(pind, fp(p_out, ybdc$both$bcode))
+cstats <- cstats[, .(iter, conv_area, tot_C, mu_cost, int_prior)]
 
@@
 
-dfunc <- function(x, y) 100 - y / x * 100
+# cost is how much you pay relative to the best case
+# savings is how much you gain relative to the worst case
+dfunc <- function(x, y) (1 - y / x) * 100
+# dfunc <- function(x, y) (x / y - 1) * 100
 stnms <- names(cstats)[-1]
 imp_pct <- cstats[, lapply(stnms, function(x) {
-  dfunc(minmaxstats[[x]][2], get(x))
+  mx <- minmaxstats[[x]][2]
+  dfunc(mx, get(x))
 })]
-setnames(imp_pct, stnms)
-
-# plot(imp_pct$tot_C, imp_pct$conv_area, type = "l", xlab = "", ylab = "", 
-#      axes = FALSE)        # xaxs = "r", yaxs = "r")
-
-sctab <- do.call(rbind, scens) * 100
+dfunc(minmaxstats[2, ], minmaxstats[1, ])  # maximum avoided cost
+
+setnames(imp_pct, paste0(stnms, "p"))
+imp_pct <- cbind("conv_km" = cstats[, conv_area] / 10^5, imp_pct)
+ywgts <- do.call(rbind, scens)[, 1]
+xwgts <- rowSums(do.call(rbind, scens)[, 2:3])
+imp_pct <- cbind(cbind(ywgts, xwgts), cstats, imp_pct)
+
+cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
+# sctab <- do.call(rbind, scens) * 100
 cx <- 0.7
-lcols <- c("black", "red", "blue")
+# lcols <- c("black", "red", "blue")
+# imp_pct[, {
+#   png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
+#       units = "in", res = 300)
+#   par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
+#   plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
+#   lines(conv_area, int_prior, col = "red")
+#   lines(conv_area, mu_cost, col = "blue")
+#   axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
+#        tcl = -0.2, mgp = c(1, 0.25, 0))
+#   axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
+#        tcl = -0.15, mgp = c(1, 0.25, 0))
+#   mtext("% land savings", side = 1, line = 1, cex = cx)
+#   mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
+#   inms <- c("tot_C", "int_prior", "mu_cost")
+#   for(i in 1:length(inms)) {
+#     text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
+#          cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
+#     text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
+#          offset = i * 0.15 - 0.2, col = lcols[i])
+#   }
+#   legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
+#          text.col = lcols, col = lcols, bty = "n", lty = 1)
 imp_pct[, {
-  png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
+  png(fp(p_figures, "weight_impact3.png"), height = 4, width = 4, 
       units = "in", res = 300)
-  par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
-  plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
-       axes = FALSE, ylim = c(0, 65), xlim = c(2, 12))#, xaxs = "r", yaxs = "r")
-  lines(conv_area, int_prior, col = "red")
-  lines(conv_area, mu_cost, col = "blue")
-  axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
-       tcl = -0.2, mgp = c(1, 0.25, 0))
-  axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
+  plot(xwgts, conv_areap, type = "l", xlab = "", ylab = "", axes = FALSE, 
+       ylim = c(0, 55), col = cols_tr[1])
+  lines(xwgts, tot_Cp, col = cols_tr[2])
+  lines(xwgts, int_priorp, col = cols_tr[3])
+  lines(xwgts, mu_costp, col = cols_tr[4])
+  axis(1, at = xwgts, labels = seq(0, 90, 10), cex = cx, cex.axis = cx, 
+       line = -0.25, tcl = -0.2, mgp = c(1, 0.25, 0))
+  axis(2, at = seq(0, 55, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
        tcl = -0.15, mgp = c(1, 0.25, 0))
-  mtext("% land savings", side = 1, line = 1, cex = cx)
-  mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
-  inms <- c("tot_C", "int_prior", "mu_cost")
-  for(i in 1:length(inms)) {
-    text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
-         cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
-    text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
-         offset = i * 0.15 - 0.2, col = lcols[i])
-  }
-  legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
-         text.col = lcols, col = lcols, bty = "n", lty = 1)
+  mtext("% Weight taken from Yield and divided", side = 1, line = 1, cex = cx)
+  mtext("equally between Carbon & Biodiversity", side = 1, line = 1.75, 
+        cex = cx)
+  mtext("% Avoided Cost", side = 2, line = 1, cex = cx)
+  legend("bottomright", 
+         legend = c("Yield", "Carbon", "Biodiversity", "Travel time"), 
+         cex = cx, text.col = cols_tr, col = cols_tr, bty = "n", lty = 1, 
+         inset = c(0, 0.25, 0, 0))
   dev.off()
@@
 
-fnames <- fp(fp(p_out, ybdc$both$bcode), 
-             imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
-scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
-  print(x)
-  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
-  DTr <- dt_to_raster(DT, CRSobj)
-})
-
-scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
-names(scens_rs2) <- paste0("w", 1:length(scens_rs2))
-
-pms2b <- do.call(rbind.data.frame, lapply(2:6, function(x) {
-  a <- rep(0, 6)
-  a[c(1, x)] <- 1
-  a
-}))
-colnames(pms2b) <- paste0("Var", 1:ncol(pms2b))
-pms2b$cl <- c("W010", "W020", "W030", "W040", "W050")
-rownames(pms2b) <- 1:5
-
-# pairwise overlaps
-scens_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
-  calc(stack(scens_rs2[which(pms2b[x, ] == 1)]), sum)
-})
-names(scens_over) <- pms2b$cl
-
-# Jaccard Index of overlaps
-scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
-  sapply(1:2, function(y) {  # y <- 3
-    r <- scens_over[[x]] == y
-    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
-  })
-})
-pms2b$JS <- round(scens_over_a[2, ] / colSums(scens_over_a), 3)
-
-# calculate nearest neighbor distance between each version, using spatstat 
-# nncross
-zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
-                 unitname = "meter")
-which1 <- function(x) x == 1  # selector function for r to xy
-scenover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 4
-  print(x)
-  ind <- which(pms2b[x, 1:6] == 1)
-  rl <- scens_rs2[ind]
-  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
-  pps <- lapply(rxy, function(x) {
-    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
-  })
-  # reorder to find which has most points, to put that in as X
-  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
-  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
-})
-pms2b$NND <- round(sapply(scenover_nn, function(x) mean(x$dist))) / 1000
-
-cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
-plot(1:4, cex = 3, pch = 20, col = cols_tr)
-
-# plot(scens_rs2[[1]] + scens_rs2[[2]])
-# check
-# round(sapply(scens_rs2, function(x) {
-#   cellStats(x * scens_r[[1]]$convertible * 100, sum)
+# fnames <- fp(fp(p_out, ybdc$both$bcode), 
+#              imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
+# scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
+#   print(x)
+#   DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
+#   DTr <- dt_to_raster(DT, CRSobj)
+# })
+# 
+# scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
+# names(scens_rs2) <- paste0("w", 1:length(scens_rs2))
+# 
+# pms2b <- do.call(rbind.data.frame, lapply(2:6, function(x) {
+#   a <- rep(0, 6)
+#   a[c(1, x)] <- 1
+#   a
 # }))
-# cstats$conv_area
-
-# pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
-# pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid
-
-# rasters showing overlap as weight is progressively put on C and BD
-scens_over2 <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)
-
-# # overlap areas
+# colnames(pms2b) <- paste0("Var", 1:ncol(pms2b))
+# pms2b$cl <- c("W010", "W020", "W030", "W040", "W050")
+# rownames(pms2b) <- 1:5
+# 
+# # pairwise overlaps
+# scens_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+#   calc(stack(scens_rs2[which(pms2b[x, ] == 1)]), sum)
+# })
+# names(scens_over) <- pms2b$cl
+# 
+# # Jaccard Index of overlaps
 # scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
-#   sapply(1:3, function(y) {  # y <- 3
+#   sapply(1:2, function(y) {  # y <- 3
 #     r <- scens_over[[x]] == y
-#     # plot(r)
-#     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
-#   })
-# })
-# colSums(scens_over_a[c(2, 3), ])
-# over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
-# par(mar = c(0, 0, 0, 0))
-# plot(1:3, pch = 20, col = rgb(1, 0, 1))
-# plot(zambia)
-
-# cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
-# cols <- c("red", "blue", "gold")
-bluecols <- sapply(c(1, 0.75, 0.5, 0.25, 0), function(x) rgb(0, x, 1))
-bluecols[5] <- rgb(0.4, 0, 1)
-cols <- sapply(1:length(bluecols), function(x) c("red", bluecols[x], "gold"))
-plot(1:ncol(cols), pch = 20, col = cols[2, ], cex = 20)
-ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
-              "Y 80% C 10% Biod 10% Cost 0%",
-              "Y 70% C 15% Biod 15% Cost 0%",
-              "Y 60% C 20% Biod 20% Cost 0%", 
-              "Y 50% C 25% Biod 25% Cost 0%")
-leg <- c("Y", "YCBd", "Y + YCBd")
-
-tiff(fp(p_figures, "changing_weights2.tif"), height = 3.5, width = 7, 
-     units = "in", res = 300)
-par(mfrow = c(2, 3), oma = c(0, 0, 0, 1))
-for(i in 1:5) { # i <- 1
-  # map
-  imp_plot2(scens_over2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
-            crops = leg, ftitle = ftits[[i]], col = cols[, i], 
-            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
-            marg = c(0, 0, 1, 0), titline = 0)
-}
-mgp <- c(1, 0.25, 0)
-par(mar = c(3, 3, 1, 0))
-with(pms2b, {
-     plot(NND, JS, col = cols[2, ], pch = 16, cex = 1.5, 
-          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.4), 
-          xlim = c(0, 180))
-     points(NND, JS, col = cols[[1]][1], pch = 20, cex = 1)
-     axis(1, at = seq(0, 180, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
-          tcl = -0.2)
-     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
-          tcl = -0.2)
-     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
-     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
-})
-dev.off()
-
-#   # barplot
-#   par(mar = c(5, 4, 3, 3))
-#   overl <- scens_over_a[1:3, i][c(1, 3, 2)]
-#   bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
-#                 col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
-#                 ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
-#   axis(side = 1, at = bp, labels = leg, las = 2, 
-#        mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
+#     round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
+#   })
+# })
+# pms2b$JS <- round(scens_over_a[2, ] / colSums(scens_over_a), 3)
+# 
+# # calculate nearest neighbor distance between each version, using spatstat 
+# # nncross
+# zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+#                  unitname = "meter")
+# which1 <- function(x) x == 1  # selector function for r to xy
+# scenover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 4
+#   print(x)
+#   ind <- which(pms2b[x, 1:6] == 1)
+#   rl <- scens_rs2[ind]
+#   rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+#   pps <- lapply(rxy, function(x) {
+#     ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+#   })
+#   # reorder to find which has most points, to put that in as X
+#   reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+#   ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+# })
+# pms2b$NND <- round(sapply(scenover_nn, function(x) mean(x$dist))) / 1000
+# 
+# cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
+# plot(1:4, cex = 3, pch = 20, col = cols_tr)
+# 
+# # plot(scens_rs2[[1]] + scens_rs2[[2]])
+# # check
+# # round(sapply(scens_rs2, function(x) {
+# #   cellStats(x * scens_r[[1]]$convertible * 100, sum)
+# # }))
+# # cstats$conv_area
+# 
+# # pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
+# # pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid
+# 
+# # rasters showing overlap as weight is progressively put on C and BD
+# scens_over2 <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)
+# 
+# # # overlap areas
+# # scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
+# #   sapply(1:3, function(y) {  # y <- 3
+# #     r <- scens_over[[x]] == y
+# #     # plot(r)
+# #     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+# #   })
+# # })
+# # colSums(scens_over_a[c(2, 3), ])
+# # over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
+# # par(mar = c(0, 0, 0, 0))
+# # plot(1:3, pch = 20, col = rgb(1, 0, 1))
+# # plot(zambia)
+# 
+# # cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
+# # cols <- c("red", "blue", "gold")
+# bluecols <- sapply(c(1, 0.75, 0.5, 0.25, 0), function(x) rgb(0, x, 1))
+# bluecols[5] <- rgb(0.4, 0, 1)
+# cols <- sapply(1:length(bluecols), function(x) c("red", bluecols[x], "gold"))
+# plot(1:ncol(cols), pch = 20, col = cols[2, ], cex = 20)
+# ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
+#               "Y 80% C 10% Biod 10% Cost 0%",
+#               "Y 70% C 15% Biod 15% Cost 0%",
+#               "Y 60% C 20% Biod 20% Cost 0%", 
+#               "Y 50% C 25% Biod 25% Cost 0%")
+# leg <- c("Y", "YCBd", "Y + YCBd")
+#               
+# tiff(fp(p_figures, "changing_weights2.tif"), height = 3.5, width = 7, 
+#      units = "in", res = 300)
+# par(mfrow = c(2, 3), oma = c(0, 0, 0, 1))
+# for(i in 1:5) { # i <- 1
+#   # map
+#   imp_plot2(scens_over2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+#             crops = leg, ftitle = ftits[[i]], col = cols[, i], 
+#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+#             marg = c(0, 0, 1, 0), titline = 0)
 # }
+# mgp <- c(1, 0.25, 0)
+# par(mar = c(3, 3, 1, 0))
+# with(pms2b, {
+#      plot(NND, JS, col = cols[2, ], pch = 16, cex = 1.5, 
+#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.4), 
+#           xlim = c(0, 180))
+#      points(NND, JS, col = cols[[1]][1], pch = 20, cex = 1)
+#      axis(1, at = seq(0, 180, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
+#           tcl = -0.2)
+#      axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
+#           tcl = -0.2)
+#      mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
+#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
+# })
 # dev.off()
+# 
+# 
+# #   # barplot
+# #   par(mar = c(5, 4, 3, 3))
+# #   overl <- scens_over_a[1:3, i][c(1, 3, 2)]
+# #   bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
+# #                 col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
+# #                 ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
+# #   axis(side = 1, at = bp, labels = leg, las = 2, 
+# #        mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
+# # }
+# # dev.off()
+#
 
@@
 optirange <- pare[, lapply(.SD, range), .SDcols = 1:4]
-# optirange[1, lapply(.SD, function(x) x * 1.05)]
+optirange[1, lapply(.SD, function(x) x * 1.05)] - optirange[1, ]
 
@@
 all_stats <- selected_impacts(unname(pind), fp(p_out, ybdc$both$bcode))
-impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) all_stats[[x]])
+# impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) all_stats[[x]])
+impsr <- lapply(c(2, 3, 8, 5), function(x) all_stats[[x]])
+
 # pare[ind == y_comp["ind"],]; pare[ind == c_comp["ind"],]; 
@@
     r <- tores_over_r[[x]] == y
-    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
-  })
-})
-pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)
+    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
+  })
+})
+pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a) * 100, 2)
 
@@
 pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
+pms2bb <- pms2b[pms2b$Var5 != 1, -5]
 # with(pms2b, plot(NND, JS))
@@
 # cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
-cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.5, 0), 
-             "gold")#rgb(1, 1, 0))
-plot(1:length(cols_tr), cex = 3, pch = 20, col = cols_tr)
+# cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.5, 0), 
+#              "gold")#rgb(1, 1, 0))
+# plot(1:length(cols_tr), cex = 3, pch = 20, col = cols_tr)
 
 # layout matrix
-m1 <- do.call(cbind, lapply(1:5, function(x) matrix(rep(x, 45), 6, 9)))
-m2 <- matrix(rep(6, 36), 4, 9)
-m3 <- do.call(cbind, lapply(7:13, function(x) matrix(rep(x, 20), 4, 5)))
-f1 <- lapply(14:15, function(x) cbind(rep(x, 4), rep(x, 4)))
-lmat <- rbind(m1, cbind(m2, m3, rep(14, 4)))
+# m1 <- do.call(cbind, lapply(1:5, function(x) matrix(rep(x, 45), 6, 9)))
+# m2 <- matrix(rep(6, 36), 4, 9)
+# m3 <- do.call(cbind, lapply(7:13, function(x) matrix(rep(x, 20), 4, 5)))
+# f1 <- lapply(14:15, function(x) cbind(rep(x, 4), rep(x, 4)))
+# lmat <- rbind(m1, cbind(m2, m3, rep(14, 4)))
 # lmat <- rbind(m1, m3)
@@
 
-# plot(1:10, ylab = ylabs[2])
-# ylabs <- c("Jaccard Index", 
-ylabs <- c(expression(paste("Conv. area (", 10^5, " ", km^2, ")")), 
-           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
-           "Intactness", expression(paste("FR converted (ha"^2, ")")))
-# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
-yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
-
-# figure titles
-ftits <- lapply(pind, function(x) {
-  paste(sapply(1:4, function(y) {
-    paste(paste(cnames[y], paste0(parms[x, y] * 100, "%")))
-  }), collapse = " ")
-})
-leg <- c("Converted")
-
-# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
-maxes <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][2])
-mins <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][1])
+# # plot(1:10, ylab = ylabs[2])
+# # ylabs <- c("Jaccard Index", 
+# ylabs <- c(expression(paste("Conv. area (", 10^5, " ", km^2, ")")), 
+#            "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
+#            "Intactness", expression(paste("FR converted (ha"^2, ")")))
+# # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
+# yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
+
+# # figure titles
+# ftits <- lapply(pind, function(x) {
+#   paste(sapply(1:4, function(y) {
+#     paste(paste(cnames[y], paste0(parms[x, y] * 100, "%")))
+#   }), collapse = " ")
+# })
+# leg <- c("Converted")
+
+# # yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
+# maxes <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][2])
+# mins <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][1])
+# ymaxes <- maxes / yscale
+# ymins <- mins / yscale
+# yaxes <- list(seq(0, 15, 3), 0:7, seq(0, 1100, 100), seq(0, 0.15, 0.05), 
+#               seq(0, 0.6, 0.1), seq(0, 1, 0.1), seq(0, 7000, 1000))
+#
+
+maxes <- sapply(c(2:3, 8, 5), function(j) minmaxstats[[j]][2])
+mins <- sapply(c(2:3, 8, 5), function(j) minmaxstats[[j]][1])
 ymaxes <- maxes / yscale
 ymins <- mins / yscale
-yaxes <- list(seq(0, 15, 3), 0:7, seq(0, 1100, 100), seq(0, 0.15, 0.05), 
-              seq(0, 0.6, 0.1), seq(0, 1, 0.1), seq(0, 7000, 1000))
-
-# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
-pms2b$icol <- cols_tr[c(1, 1, 2, 1, 2, 3, 1, 2, 3, 4)]  # circle inner
-pms2b$ocol <- cols_tr[c(2, 3, 3, 4, 4, 4, 5, 5, 5, 5)]  # circle outer
-
-tiff(fp(p_figures, "objective-subjective_weights.tif"), height = 2.5, width = 7, 
+pctlines <- rbind(ymins, 
+                  t(sapply(c(0.25, 0.5, 0.75), function(x) {
+                    (ymaxes - ymins) * x + ymins
+                  })), ymaxes)
+
+# # bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
+# pms2b$icol <- cols_tr[c(1, 1, 2, 1, 2, 3, 1, 2, 3, 4)]  # circle inner
+# pms2b$ocol <- cols_tr[c(2, 3, 3, 4, 4, 4, 5, 5, 5, 5)]  # circle outer
+# 
+# tiff(fp(p_figures, "objective-subjective_weights.tif"), height = 2.5, width = 7,      units = "in", res = 300)
+# l <- layout(mat = lmat)
+# layout.show(l)
+
+# # map plots
+# for(i in 1:length(toresr)) { # i <- 1
+#   # map
+# #   par(mar = rep(0, 4))
+# #   plot(zambia, col = cols_tr[i])
+#   imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+#             crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
+#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
+#             marg = rep(0, 4), inset = c(0.1, 0.1, 0, 0))
+# }  
+# 
+# # impact plots
+# mgp <- c(0.9, 0.25, 0)
+# par(mar = c(3, 3, 0, 3))
+# with(pms2b, {
+#      plot(NND, JS, col = pms2b$icol, pch = 16, cex = 0.8, 
+#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.6), 
+#           xlim = c(0, 25))
+#      points(NND, JS, col = pms2b$ocol, pch = 20, cex = 0.5)
+#      axis(1, at = seq(0, 25, 5), cex.axis = 0.5, las = 2, mgp = mgp, 
+#           tcl = -0.2)
+#      axis(2, at = seq(0, 0.6, 0.1), cex.axis = 0.5, las = 2, mgp = mgp, 
+#           tcl = -0.2)
+#      mtext(text = "NN Distance (km)", side = 1, line = 0.5, cex = 0.4)
+#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.4)
+# })
+# par(mar = c(3, 3, 0, 0))
+# mgp <- c(1.3, 0.25, 0)
+# for(i in 1:7) { # i <- 1
+#   v <- impsr[[i]] / yscale[i]
+#   b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
+#                mgp = mgp, tcl = -0.2, yaxt = "n", 
+#                ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.4,
+#                cex.lab = 0.55)
+#   lines(c(-1, 7), rep(ymins[i], 2), lwd = 1, col = "grey50")
+#   lines(c(-1, 7), rep(ymaxes[i], 2), lwd = 1)
+#   axis(2, at = yaxes[[i]], cex.axis = 0.5, las = 2, mgp = mgp, 
+#        tcl = -0.2)
+# }
+# dev.off()
+
+# Plot setup
+ftits <- list("Compromise 5% on Yield", "Compromise 5% on Carbon", 
+              "Compromise 5% on Biodiversity", "Compromise 5% on Travel Time")
+yaxes <- list(seq(0, 16, 4), seq(0, 1200, 100), seq(0, 0.6, 0.1), 
+              seq(0, 7, 1))
+pcttxt <- c("best", "50%", "worst")
+
+# plot
+png(fp(p_figures, "compromise_5pct.png"), height = 3.5, 
+    width = 7, units = "in", res = 300)
+l <- layout(mat = lmat)
+# layout.show(l)
+
+# map plots
+legtf <- c(FALSE, FALSE, FALSE, FALSE)
+for(i in 1:4) { # i <- 1
+  # plot(zambia)
+  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+            crops = "", ftitle = ftits[[i]], col = cols_tr[[i]], 
+            legend = legtf[i], hold = TRUE, stats = FALSE, cexscale = 0.4, 
+            marg = rep(0, 4))
+}  
+# impact plots
+mgp <- c(0.8, 0.25, 0)
+cx <- 0.6
+for(i in 1:4) { 
+  v <- impsr[[i]][1:4] / yscale[i]
+  par(mar = c(2, 3.5, 2, 0))
+  b <- barplot(v, las = 2, col = cols_tr, mgp = c(1, 0.25, 0), tcl = -0.2, 
+               yaxt = "n", ylim = range(yaxes[[i]]), ylab = ylabs[i], 
+               cex.axis = cx, cex.lab = 0.7)
+  for(j in 1:3) {
+    lines(c(-1, 7), rep(pctlines[j + (j - 1), i], 2), lwd = 1, col = "grey50")
+    text(4.5, y = rep(pctlines[j + (j - 1), i], 2), labels = pcttxt[j], 
+         col = "grey50", xpd = NA, pos = 4, cex = cx * 0.9)
+  }
+  axis(2, at = yaxes[[i]], cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
+  with(pms2bb[ind, ], {
+    par(mar = c(0.5, 3, 2, 1), mgp = mgp)
+    barplot(JS, col = cols_tr[-i], ylim = c(0, 70), yaxt = "n", 
+            ylab = "% Overlap", cex.lab = cx)
+    axis(2, at=seq(0, 70, 10), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
+    par(mar = c(2, 3, 0.5, 1), mgp = mgp)
+    barplot(NND, col = cols_tr[-i], ylim = c(0, 30), las = 2, yaxt = "n", 
+            ylab = "", cex.lab = cx)
+    axis(2, at=seq(0, 30, 10), cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
+    mtext("Average distance", side = 2, line = 1.3, cex = 0.4)
+    mtext("to neighbor (km)", side = 2, line = 0.8, cex = 0.4)
+  })
+}
+par(xpd = NA)
+legx <- grconvertX(0.4, from = "ndc", to = "user")
+legy <- grconvertY(0.5, from = "ndc", to = "user")
+legend(x = legx, y = legy, legend = "Protected Areas", pch = 15, col = "grey75",        bty = "n", pt.cex = 3 * 0.6, cex = 1.5 * 0.6)
+dev.off()
+
+# 25-25-25-25 Compromise
+m1 <- matrix(rep(1, 24), nrow = 4, ncol = 6)
+m2 <- do.call(cbind, lapply(2:3, function(x) matrix(rep(x, 4), 2, 2)))
+m3 <- do.call(cbind, lapply(4:5, function(x) matrix(rep(x, 4), 2, 2)))
+lmat <- cbind(m1, rbind(m2, m3))
+
+png(fp(p_figures, "compromise_25pct.png"), height = 2, width = 4, 
     units = "in", res = 300)
@@
 
-# map plots
-for(i in 1:length(toresr)) { # i <- 1
-  # map
-#   par(mar = rep(0, 4))
-#   plot(zambia, col = cols_tr[i])
+par(oma = c(0, 0, 0, 1))
+# plot(zambia)
 imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
-            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
-            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
-            marg = rep(0, 4), inset = c(0.1, 0.1, 0, 0))
+          crops = "", ftitle = "Compromise equally", col = "purple", 
+          legend = FALSE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+          marg = c(0, 0, 1, 0), titline = 0)
+legend("bottomright", legend = "Protected Areas", pch = 15, col = "grey75",
+       bty = "n", pt.cex = 3 * 0.5, cex = 1.5 * 0.5, inset = c(0, 0.1, 0, 0))
+
+mgp <- c(1.25, 0.25, 0)
+cx <- 0.5
+for(i in 1:4) { 
+  v <- impsr[[i]][c(5, 1:4)] / yscale[i]
+  if(i %in% 1:2) par(mar = c(1, 3.5, 2, 0.2), lwd = 0.5)
+  if(i %in% 3:4) par(mar = c(2, 3.5, 1, 0.2), lwd = 0.5)
+  b <- barplot(v, las = 2, col = c("purple", cols_tr), mgp = c(1, 0.25, 0), 
+               tcl = -0.2, yaxt = "n", ylim = range(yaxes[[i]]), 
+               ylab = ylabs[i], cex.axis = cx, cex.lab = 0.5)
+  for(j in 1:3) {
+    lines(c(-1, 7), rep(pctlines[j + (j - 1), i], 2), lwd = 1, col = "grey50")
+    text(5.5, y = rep(pctlines[j + (j - 1), i], 2), labels = pcttxt[j], 
+         col = "grey50", xpd = NA, pos = 4, cex = cx * 0.7)
   }
-
-# impact plots
-mgp <- c(0.9, 0.25, 0)
-par(mar = c(3, 3, 0, 3))
-with(pms2b, {
-     plot(NND, JS, col = pms2b$icol, pch = 16, cex = 0.8, 
-          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.6), 
-          xlim = c(0, 25))
-     points(NND, JS, col = pms2b$ocol, pch = 20, cex = 0.5)
-     axis(1, at = seq(0, 25, 5), cex.axis = 0.5, las = 2, mgp = mgp, 
-          tcl = -0.2)
-     axis(2, at = seq(0, 0.6, 0.1), cex.axis = 0.5, las = 2, mgp = mgp, 
-          tcl = -0.2)
-     mtext(text = "NN Distance (km)", side = 1, line = 0.5, cex = 0.4)
-     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.4)
-})
-par(mar = c(3, 3, 0, 0))
-mgp <- c(1.3, 0.25, 0)
-for(i in 1:7) { # i <- 1
-  v <- impsr[[i]] / yscale[i]
-  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
-               mgp = mgp, tcl = -0.2, yaxt = "n", 
-               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.4,
-               cex.lab = 0.55)
-  lines(c(-1, 7), rep(ymins[i], 2), lwd = 1, col = "grey50")
-  lines(c(-1, 7), rep(ymaxes[i], 2), lwd = 1)
-  axis(2, at = yaxes[[i]], cex.axis = 0.5, las = 2, mgp = mgp, 
-       tcl = -0.2)
+  axis(2, at = yaxes[[i]], cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
 }
@@
 
-# Earlier version of figure 3
-
+<a href="#top">Back to top</a>
+
+### Overlap with Farm Blocks
 ```{r, eval = FALSE}
-# objective weights
-obj_wgts <- round(pct_of_tot[, {
-  c(conv_area, tot_C, int_mean, mu_cost) / 
-    sum(c(conv_area, tot_C, int_mean, mu_cost))
-}], 2)
-# round(obj_wgts, 2)
-names(obj_wgts) <- cnames
-
-# Objective weights
-toobj <- tradeoff_mod(targ, cbetas = obj_wgts, currprodmod = g, it = "OBJ")
-
-# ybdc$both$params[1, ]
-# yp <- ybdc$both$params[, 7:10]
-# which(round(yp[, 1], 2) == 0.05 & round(yp[, 2], 2) == 0.30 & 
-#         round(yp[, 3], 2) == 0.25 & round(yp[, 4], 2) == 0.4)
-
-# objective weights on ag only 
-# considering that these might be two sides of the same coin
-ag_wgts <- round(obj_wgts[c(1, 4)] / sum(obj_wgts[c(1, 4)]), 2)
-ag_wgts <- c(ag_wgts[1], "C" = 0, "BD" = 0, ag_wgts[2])
-ag_wgts2 <- ag_wgts[c(4, 2:3, 1)]
-names(ag_wgts2) <- names(ag_wgts)
-
-toag <- tradeoff_mod(targ, cbetas = ag_wgts, currprodmod = g, it = "AGOBJ")
-toag2 <- tradeoff_mod(targ, cbetas = ag_wgts2, currprodmod = g, it = "AGOBJ2")
-
-# subjective weights, where we want to treat everyone equally
-subj_wgts <- rep(0.25, 4)
-names(subj_wgts) <- cnames  # rename to make sure model runs properly
-tosubj <- tradeoff_mod(targ, cbetas = subj_wgts, currprodmod = g, it = "SUBJ")
-
-# extract impacts of these, and add in BAU case
-tores <- lapply(list(toag, toag2, toobj, tosubj), function(x) {  # x <- toobj
-  DTsum <- data.table(x$conv[, rowSums(.SD), .SDcols = c(crops[[1]])])
-  DTspat <- cbind(basedt[, .(x, y)], DTsum)
-  DTr <- dt_to_raster(DTspat, CRSobj)
-  DTr <- DTr + (fbr * 2)  # overlap with farmblocks, for output
-  list("map" = DTr, "imp" = x$impacts)
-})
-names(tores) <- c("ag", "ag2", "obj", "subj")
-
 # Calculate area of overlaps between farmblocks and 
-yfbr <- calc(mzsoytor, sum) + fbr * 2
-plot(yfbr)
-fb_over_a <- sapply(1:length(tores), function(x) { # x <- 4
+Y <- calc(mzsoytor, sum)
+yfbr <- Y + fbr * 2
+eqfbr <- toresr$eqwt + fbr * 2
+fbrlist <- list(yfbr, eqfbr)
+
+fb_over_a <- sapply(1:2, function(x) { # x <- 4
   sapply(1:3, function(y) {  # y <- 3
-    r <- tores[[x]]$map == y
-    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+    r <- fbrlist[[x]] == y
+    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
   })
@@
 
-# grid for calculating overlaps
-pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
-pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
-pms2b$cl <- c("AgAg2", "AgObj", "Ag2Obj", "AgSubj", "Ag2Subj", "ObjSubj")
-
-toresr <- lapply(tores, function(x) {
-  r <- x$map
-  r[r == 2] <- 0
-  r[r == 3] <- 1
-  r
-})
-
-# tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
-#   (toresr[[1]] == pms2b[x, 1]) & (toresr[[2]] == pms2b[x, 2]) & 
-#     (toresr[[3]] == pms2b[x, 3]) & (toresr[[4]] == pms2b[x, 4])
-# })
-# names(tores_over_r) <- pms2b$cl
-# tores_overs <- stack(tores_over)
-tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
-  calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
-})
-names(tores_over_r) <- pms2b$cl
-
-# jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
-tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
-  sapply(1:2, function(y) {  # y <- 3
-    r <- tores_over_r[[x]] == y
-    round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
-  })
-})
-pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)
-
-# calculate nearest neighbor distance between each version, using spatstat 
-# nncross
-# plot(zam_owin)
-zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
-                 unitname = "meter")
-which1 <- function(x) x == 1  # selector function for r to xy
-tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
-  ind <- which(pms2b[x, 1:4] == 1)
-  rl <- toresr[ind]
+fbrna <- fbr > 0
+fbrna[fbrna == 0] <- NA
+
+# calculate nearest neighbor dist between each version, using spatstat nncross
+inlist <- list("Y" = Y, "eqwt" = toresr$eqwt)
+fb_nn <- lapply(1:length(inlist), function(x) {  # x <- 1
+  rl <- list(inlist[[x]], fbrna)
+  names(rl) <- c(names(inlist)[x], "fbr")
   rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
@@
 })
-pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
-# with(pms2b, plot(NND, JS))
-
-# do sum factoring in overlaps 
-# sapply(tores_over_r, function(x) {
-#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
-# })
-
-cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
-plot(1:4, cex = 3, pch = 20, col = cols_tr)
-
-# impact table
-imp_list <- list(toag$impacts, toag2$impacts, toobj$impacts, tosubj$impacts)
-imps <- rbindlist(lapply(imp_list, function(x) selected_impacts2(x)))
-impsr <- lapply(c(1, 4, 2, 6:8, 3), function(x) imps[[x]])
-
-# layout matrix
-m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
-m2 <- matrix(rep(5, 20), 4, 5)
-m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
-lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
-
-# plot(1:10, ylab = ylabs[2])
-# ylabs <- c("Jaccard Index", 
-ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
-           "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
-           "Intactness", expression(paste("FR converted (ha"^2, ")")))
-# yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
-yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
-ftits <- list("Y 11% C 0% Biod 0% Cost 89%",
-              "Y 89% C 0% Biod 0% Cost 11%",
-              "Y 5% C 29% Biod 25% Cost 41%",
-              "Y 25% C 25% Biod 25% Cost 25%")
-leg <- c("Converted")
-
-# yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
-yaxes <- list(seq(0, 14, 2), seq(0, 2.5, 0.5),               
-              seq(0, 1000, 200), seq(0, 0.2, 0.05), seq(0, 0.5, 0.1), 
-              seq(0, 0.8, 0.2), seq(0, 3500, 500))
-# bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
-bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
-                cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
-# lcols <- c(list(c(cols_tr[2], cols_tr[3], cols_tr[3], cols_tr[4], cols_tr[4] 
-# lcols <- lapply(1:7, function(x) cols_tr)
-
-tiff(fp(p_figures, "objective_weights.tif"), height = 3.5, width = 7, 
-     units = "in", res = 300)
-l <- layout(mat = lmat)
-# layout.show(l)
-
-# map plots
-for(i in 1:4) { # i <- 1
-  # map
-  # plot(zambia)
-  imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
-            crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
-            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
-            marg = rep(0, 4))
-}
-
-# impact plots
-mgp <- c(1, 0.25, 0)
-par(mar = c(4, 3, 0, 0))
-with(pms2b, {
-     plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
-          axes = FALSE, xlab = "", ylab = "", ylim = c(0, 1))
-     points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
-     axis(1, at = seq(0, 130, 10), cex.axis = 0.6, las = 2, mgp = mgp, 
-          tcl = -0.2)
-     axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
-          tcl = -0.2)
-     mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
-     mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
-})
-for(i in 1:7) {
-  v <- impsr[[i]] / yscale[i]
-  b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
-               mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
-               ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
-               cex.lab = 0.7)
-  axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
-       tcl = -0.2)
+fb_nnd <- round(sapply(fb_nn, function(x) mean(x$dist))) / 1000
+
+ftits <- c("Maximize Yields", "Compromise Equally")
+leg <- c("converted", "farmblock", "converted in farmblock")
+inlist <- list("Y" = yfbr, "eqwt" = eqfbr)
+png(fp(p_figures, "fb_overlaps.png"), height = 2.5, width = 5, units = "in", 
+     res = 300)
+par(mfrow = c(1, 2))
+for(i in 1:2) {
+  imp_plot2(inlist[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+            crops = leg, ftitle = ftits[i], 
+            col = c("red", "orange", rgb(0, 1, 1)),
+            legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
+            marg = c(0, 0, 1, 0), inset = c(0, 0, 0, 0), titline = 0)
+  mtext(paste("Jaccard Index =", round(fb_sim[i], 2)), side = 3, line = -2, 
+        cex = 0.4, adj = 0.2)
+  mtext(paste("NN Distance =", round(fb_nnd[i]), "km"), side = 3, line = -2.5, 
+        cex = 0.4, adj = 0.2)
 }
@@
 
-# ybdc$both$optitab[, {
-#   par(mfrow = c(2, 2), mar = rep(3, 4))
-#   hist(.SD, axes = FALSE, labels = FALSE)
-# }, .SDcol = 1:4]
-# 
-# frange <- function(x) diff(range(x)) / max(x)
-# fqtile <- function(x) diff(quantile(x, probs = c(0.025, 0.975))) / max(x) 
-# ybdc$both$optimal[, lapply(.SD, frange), .SDcol = 1:4]
-# ybdc$both$optimal[, lapply(.SD, fqtile), .SDcol = 1:4]
-# 
-# hist(ybdc$both$optimal$Y)
-
 ```
+
+### Impact of closing yield gaps
+```{r, eval = FALSE}
+# Separate soy and maize impacts
+cb <- c("Y" = 0.25, "C" = 0.25, "BD" = 0.25, "COST" = 0.25)
+g <- gapper(cp, c(mzgap, soygap), 0.5)  # assume 50% closed on current land
+lowgap <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)
+g <- gapper(cp, c(mzgap, soygap), 1.5)  # assume 50% closed on current land
+highgap <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)
+
+lowi <- selected_impacts2(lowgap$impacts)
+highi <- selected_impacts2(highgap$impacts)
+gaptab <- rbind(lowi, all_stats[5, names(all_stats)[-1], with = FALSE], highi)
+gaptab[1, ] / gaptab[2, ]
+gaptab[, tot_C / 100000]
+
+```
+
+### Additional supplementals
+```{r, eval = FALSE}
+il <- mzsoyto$inputs
+ylds <- dt_to_raster(cbind(basedt[, .(x, y)], il$p_yield), CRSobj)
+carb <- dt_to_raster(cbind(basedt[, .(x, y)], il$carbon), CRSobj)
+cost <- dt_to_raster(cbind(basedt[, .(x, y)], il$cost), CRSobj)
+closs <- carb[[1]] + carb[[2]] * 0.25
+pacols <- c("transparent", "grey85", "grey75")
+
+png(fp(p_figures, "yields.png"), height = 3.5, width = 8, unit = "in", 
+     res = 300)
+par(mar = c(0, 0, 0, 4), mfrow = c(1, 2))
+for(i in 1:2) {
+  plot(zambia, col = "white", border = "transparent")
+  plot(ylds[[i]], legend = TRUE, add = TRUE)
+  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+       border = pacols[2])
+  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+       border = pacols[3])
+  legend("bottomright", legend = c("GMAs", "NatParks"),
+         pch = 15, col = c(pacols[2:3]), bty = "n",
+         pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))
+
+}  
+dev.off()
+
+png(fp(p_figures, "carbon.png"), height = 3.5, width = 4.5, unit = "in", 
+     res = 300)
+par(mar = c(0, 0, 0, 4))
+plot(zambia, col = "white", border = "transparent")
+plot(closs, legend = TRUE, add = TRUE)
+plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+     border = pacols[2])
+plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+     border = pacols[3])
+legend("bottomright", legend = c("GMAs", "NatParks"),
+       pch = 15, col = c(pacols[2:3]), bty = "n",
+       pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))
+dev.off()
+
+png(fp(p_figures, "cost.png"), height = 3.5, width = 4.5, unit = "in", 
+     res = 300)
+par(mar = c(0, 0, 0, 4))
+plot(zambia, col = "white", border = "transparent")
+plot(cost, legend = TRUE, add = TRUE)
+plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+     border = pacols[2])
+plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+     border = pacols[3])
+legend("bottomright", legend = c("GMAs", "NatParks"),
+       pch = 15, col = c(pacols[2:3]), bty = "n",
+       pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))
+dev.off()
+
+```
+
+<a href="#top">Back to top</a>
+
+### Deprecated version of Figure 3
+
+```{r, eval = FALSE}
+# objective weights
+# obj_wgts <- round(pct_of_tot[, {
+#   c(conv_area, tot_C, int_mean, mu_cost) / 
+#     sum(c(conv_area, tot_C, int_mean, mu_cost))
+# }], 2)
+# # round(obj_wgts, 2)
+# names(obj_wgts) <- cnames
+# 
+# # Objective weights
+# toobj <- tradeoff_mod(targ, cbetas = obj_wgts, currprodmod = g, it = "OBJ")
+# 
+# # ybdc$both$params[1, ]
+# # yp <- ybdc$both$params[, 7:10]
+# # which(round(yp[, 1], 2) == 0.05 & round(yp[, 2], 2) == 0.30 & 
+# #         round(yp[, 3], 2) == 0.25 & round(yp[, 4], 2) == 0.4)
+# 
+# # objective weights on ag only 
+# # considering that these might be two sides of the same coin
+# ag_wgts <- round(obj_wgts[c(1, 4)] / sum(obj_wgts[c(1, 4)]), 2)
+# ag_wgts <- c(ag_wgts[1], "C" = 0, "BD" = 0, ag_wgts[2])
+# ag_wgts2 <- ag_wgts[c(4, 2:3, 1)]
+# names(ag_wgts2) <- names(ag_wgts)
+# 
+# toag <- tradeoff_mod(targ, cbetas = ag_wgts, currprodmod = g, it = "AGOBJ")
+# toag2 <- tradeoff_mod(targ, cbetas = ag_wgts2, currprodmod = g, it = "AGOBJ2")
+# 
+# # subjective weights, where we want to treat everyone equally
+# subj_wgts <- rep(0.25, 4)
+# names(subj_wgts) <- cnames  # rename to make sure model runs properly
+# tosubj <- tradeoff_mod(targ, cbetas = subj_wgts, currprodmod = g, it = "SUBJ")
+# 
+# # extract impacts of these, and add in BAU case
+# tores <- lapply(list(toag, toag2, toobj, tosubj), function(x) {  # x <- toobj
+#   DTsum <- data.table(x$conv[, rowSums(.SD), .SDcols = c(crops[[1]])])
+#   DTspat <- cbind(basedt[, .(x, y)], DTsum)
+#   DTr <- dt_to_raster(DTspat, CRSobj)
+#   DTr <- DTr + (fbr * 2)  # overlap with farmblocks, for output
+#   list("map" = DTr, "imp" = x$impacts)
+# })
+# names(tores) <- c("ag", "ag2", "obj", "subj")
+# 
+# # grid for calculating overlaps
+# pms2a <- expand.grid(rep(list(0:1), 4))[-1, ] 
+# pms2b <- pms2a[which(rowSums(pms2a) == 2), ]
+# pms2b$cl <- c("AgAg2", "AgObj", "Ag2Obj", "AgSubj", "Ag2Subj", "ObjSubj")
+# 
+# toresr <- lapply(tores, function(x) {
+#   r <- x$map
+#   r[r == 2] <- 0
+#   r[r == 3] <- 1
+#   r
+# })
+# 
+# # tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+# #   (toresr[[1]] == pms2b[x, 1]) & (toresr[[2]] == pms2b[x, 2]) & 
+# #     (toresr[[3]] == pms2b[x, 3]) & (toresr[[4]] == pms2b[x, 4])
+# # })
+# # names(tores_over_r) <- pms2b$cl
+# # tores_overs <- stack(tores_over)
+# tores_over_r <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+#   calc(stack(toresr[which(pms2b[x, ] == 1)]), sum)
+# })
+# names(tores_over_r) <- pms2b$cl
+# 
+# # jaccard's similarity, per https://gist.github.com/jlehtoma/3369793
+# tores_over_a <- sapply(1:length(tores_over_r), function(x) { # x <- 4
+#   sapply(1:2, function(y) {  # y <- 3
+#     r <- tores_over_r[[x]] == y
+#     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
+#   })
+# })
+# pms2b$JS <- round(tores_over_a[2, ] / colSums(tores_over_a), 3)
+# 
+# # calculate nearest neighbor distance between each version, using spatstat 
+# # nncross
+# # plot(zam_owin)
+# zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+#                  unitname = "meter")
+# which1 <- function(x) x == 1  # selector function for r to xy
+# tores_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
+#   ind <- which(pms2b[x, 1:4] == 1)
+#   rl <- toresr[ind]
+#   rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+#   pps <- lapply(rxy, function(x) {
+#     ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+#   })
+#   # reorder to find which has most points, to put that in as X
+#   reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+#   ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+# })
+# pms2b$NND <- round(sapply(tores_nn, function(x) mean(x$dist))) / 1000
+# # with(pms2b, plot(NND, JS))
+# 
+# # do sum factoring in overlaps 
+# # sapply(tores_over_r, function(x) {
+# #   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
+# # })
+# 
+# cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
+# plot(1:4, cex = 3, pch = 20, col = cols_tr)
+# 
+# # impact table
+# imp_list <- list(toag$impacts, toag2$impacts, toobj$impacts, tosubj$impacts)
+# imps <- rbindlist(lapply(imp_list, function(x) selected_impacts2(x)))
+# impsr <- lapply(c(1, 4, 2, 6:8, 3), function(x) imps[[x]])
+# 
+# # layout matrix
+# m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
+# m2 <- matrix(rep(5, 20), 4, 5)
+# m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
+# lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
+# 
+# # plot(1:10, ylab = ylabs[2])
+# # ylabs <- c("Jaccard Index", 
+# ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
+#            "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
+#            "Intactness", expression(paste("FR converted (ha"^2, ")")))
+# # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
+# yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
+# ftits <- list("Y 11% C 0% Biod 0% Cost 89%",
+#               "Y 89% C 0% Biod 0% Cost 11%",
+#               "Y 5% C 29% Biod 25% Cost 41%",
+#               "Y 25% C 25% Biod 25% Cost 25%")
+# leg <- c("Converted")
+# 
+# # yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
+# yaxes <- list(seq(0, 14, 2), seq(0, 2.5, 0.5),               
+#               seq(0, 1000, 200), seq(0, 0.2, 0.05), seq(0, 0.5, 0.1), 
+#               seq(0, 0.8, 0.2), seq(0, 3500, 500))
+# # bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
+# bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
+#                 cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
+# # lcols <- c(list(c(cols_tr[2], cols_tr[3], cols_tr[3], cols_tr[4], cols_tr[4] 
+# # lcols <- lapply(1:7, function(x) cols_tr)
+# 
+# tiff(fp(p_figures, "objective_weights.tif"), height = 3.5, width = 7, 
+#      units = "in", res = 300)
+# l <- layout(mat = lmat)
+# # layout.show(l)
+# 
+# # map plots
+# for(i in 1:4) { # i <- 1
+#   # map
+#   # plot(zambia)
+#   imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+#             crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
+#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
+#             marg = rep(0, 4))
+# }  
+# 
+# # impact plots
+# mgp <- c(1, 0.25, 0)
+# par(mar = c(4, 3, 0, 0))
+# with(pms2b, {
+#      plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
+#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 1))
+#      points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
+#      axis(1, at = seq(0, 130, 10), cex.axis = 0.6, las = 2, mgp = mgp, 
+#           tcl = -0.2)
+#      axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
+#           tcl = -0.2)
+#      mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
+#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
+# })
+# for(i in 1:7) {
+#   v <- impsr[[i]] / yscale[i]
+#   b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
+#                mgp = c(1.5, 0.25, 0), tcl = -0.2, yaxt = "n", 
+#                ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.5,
+#                cex.lab = 0.7)
+#   axis(2, at = yaxes[[i]], cex.axis = 0.6, las = 2, mgp = mgp, 
+#        tcl = -0.2)
+# }
+# dev.off()
+#
+
+```
+
+<a href="#top">Back to top</a>

### 2016-02-26 Rmd/main-analysis.Rmd
Near final figures for main text done after revision to Fig 1 and 3, and new Figure 4 made. Model rerun for 100% yield gap closure on existing land.

@@
 # assume 50% closed on current land
-g <- gapper(cp, c(mzgap, soygap), 0.25)
+g <- gapper(cp, c(mzgap, soygap), 1)
 
 # targets given g
-targ <- c("maize" = 3, "soy" = 9)
+targ <- c("maize" = 4, "soy" = 10)
 # targ <- c("maize" = 3.2, "soy" = 8.7)
@@
 cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
-mzto <- tradeoff_mod(prod_targ = targ["maize"], cbetas = cb, currprodmod = g[1])
-soyto <- tradeoff_mod(prod_targ = targ["soy"], cbetas = cb, currprodmod = g[2])
 mzsoyto <- tradeoff_mod(prod_targ = targ, cbetas = cb, currprodmod = g)
@@
 CRSobj <- CRS(mzto$inputs$sp$crs)
-mztor <- dt_to_raster(mzto$conv, CRSobj)
-soytor <- dt_to_raster(soyto$conv, CRSobj)
 mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results
-
-# rasters
-mzsoyr <- stack(list(mztor, soytor * 2))
-mzsoyr <- calc(mzsoyr, sum)
-# plot(mzsoyr)
-# plot(roads)
 
@@
 i1 <- rbind(mzto$impacts, soyto$impacts)
-crops <- c("maize", "soy", "maize/soy")
-# plot(1:10, pch = 20, cex = 4, col = "gold2")
-# plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
-imp_plot2(mzsoyr, i1, p_figures, "ideal_yield", "", crops, 1, TRUE, 
-         c("gold2", "green3", "orange"))
-
-basedt <- soyto$inputs$mask
+basedt <- mzsoyto$inputs$mask
 ```
@@
 # Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
+
 # 
@@
 ### Conservation interests versus agricultural interests
+
+Switched off for now
 ```{r, eval = FALSE}
-ybl <- list(yb1 <- c(1, 1))
-step <- 0.02
-cnames <- c("Y", "COST")
-ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
-# ps[which(rowSums(ps[, c(1, 3)]) == 0), ]
-Yv = "conv_area"; Cv = "tot_C"
-BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
-wmu <- function(x, na.rm = na.rm) {
-  weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
-}
-Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
-
-crops <- list(c("maize", "soy"))
-croplnms <- list("both")
-
-ycost <- lapply(crops, function(x) {
-  a <- Sys.time()
-  print(paste("Starting run for", x, "at", a))
-  otab <- pareto(cnames, step = step, prod_targ = targ[x], 
-                  yblist = ybl, currprodmod = g[x], input_key = "ZA", 
-                  Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
-                  Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
-                  silent = FALSE, ncl = 7)
-  load(fp(p_out, "bcode.rda"))
-  heatr <- heat2(fp(p_out, bcode), basedt, todisk = TRUE) 
-  print(Sys.time() - a)
-  c("bcode" = bcode, otab, "heat" = heatr)
-})
-names(ycost) <- croplnms
-# plot(ycost$both$optitab$Y, ycost$both$optitab$COST)
-
-# fix because ran with heat2 not yet adjusted for improperly ordered csv files
-ycostfix <- lapply(ycost, function(x) {  # x <- ycost$both
-  if(nrow(x$params) > nrow(x$optitab)) {
-    print(paste("fixing heat map for", x$bcode))
-    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
-  } else {
-    heatfix <- NULL
-  }
-  heatfix
-})
-for(i in 1:length(ycost)) ycost[[i]]$heat <- ycostfix[[i]]
+# ybl <- list(yb1 <- c(1, 1))
+# step <- 0.02
+# cnames <- c("Y", "COST")
+# ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2)
+# # ps[which(rowSums(ps[, c(1, 3)]) == 0), ]
+# Yv = "conv_area"; Cv = "tot_C"
+# BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
+# wmu <- function(x, na.rm = na.rm) {
+#   weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
+# }
+# Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
+# 
+# crops <- list(c("maize", "soy"))
+# croplnms <- list("both")
+# 
+# ycost <- lapply(crops, function(x) {
+#   a <- Sys.time()
+#   print(paste("Starting run for", x, "at", a))
+#   otab <- pareto(cnames, step = step, prod_targ = targ[x], 
+#                   yblist = ybl, currprodmod = g[x], input_key = "ZA", 
+#                   Yv = Yv, Cv = Cv, BDv = BDv, COSTv = COSTv, Yst = Yst, 
+#                   Cst = Cst, BDst = BDst, COSTst = COSTst, todisk = TRUE, 
+#                   silent = FALSE, ncl = 7)
+#   load(fp(p_out, "bcode.rda"))
+#   heatr <- heat2(fp(p_out, bcode), basedt, todisk = TRUE) 
+#   print(Sys.time() - a)
+#   c("bcode" = bcode, otab, "heat" = heatr)
+# })
+# names(ycost) <- croplnms
+# # plot(ycost$both$optitab$Y, ycost$both$optitab$COST)
+# 
+# # fix because ran with heat2 not yet adjusted for improperly ordered csv files
+# ycostfix <- lapply(ycost, function(x) {  # x <- ycost$both
+#   if(nrow(x$params) > nrow(x$optitab)) {
+#     print(paste("fixing heat map for", x$bcode))
+#     heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
+#   } else {
+#     heatfix <- NULL
+#   }
+#   heatfix
+# })
+# for(i in 1:length(ycost)) ycost[[i]]$heat <- ycostfix[[i]]
 
@@
 ```{r, eval = FALSE}
+ybl <- list(yb1 <- c(1, 1))
+step <- 0.05
+cnames <- c("C", "BD", "COST")
+ps <- round(do.call(rbind, pareto_steps(cnames, step)), 2); head(ps)
+Yv = "conv_area"; Cv = "tot_C"; #BDv = "int_prior"; COSTv = "mu_cost"
+BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
+wmu <- function(x, na.rm = na.rm) {
+  weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
+}
+Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
+crops <- list(c("maize", "soy"))
+croplnms <- list("both")
+
 step <- 0.05
@@
 # fix because ran with heat2 not yet adjusted for improperly ordered csv files
-ybdcfix <- lapply(ybdc, function(x) {  # x <- ycost$both
-  if(nrow(x$params) > nrow(x$optitab)) {
-    print(paste("fixing heat map for", x$bcode))
-    heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
-  } else {
-    heatfix <- NULL
-  }
-  heatfix
-})
-for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcfix[[i]]
+# ybdcfix <- lapply(ybdc, function(x) {  # x <- ycost$both
+#   if(nrow(x$params) > nrow(x$optitab)) {
+#     print(paste("fixing heat map for", x$bcode))
+#     heatfix <- heat2(fp(p_out, x$bcode), basedt, TRUE, TRUE)
+#   } else {
+#     heatfix <- NULL
+#   }
+#   heatfix
+# })
+# for(i in 1:length(ybdc)) ybdc[[i]]$heat <- ybdcfix[[i]]
 
@@
 # save(ybdc, cbdc, ycost, bdc, file = fp(p_out, "mainresults1502.rda"))
-save(ybdc, ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
+# save(ybdc, ycost, bdc, file = fp(p_out, "mainresults1702.rda"))
+save(ybdc, file = fp(p_out, "mainresults2502.rda"))
 # load(fp(p_out, "mainresults1702.rda"))

### 2016-02-26 paper/main.tex
Near final figures for main text done after revision to Fig 1 and 3, and new Figure 4 made. Model rerun for 100% yield gap closure on existing land.

%%%% added from https://groups.google.com/forum/#!topic/latexusersgroup/CDlEjgNnF80

%\documentclass[onecolumn]{rsauthor}    
\documentclass[a4paper]{article}
\usepackage[affil-it]{authblk}

\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{endfloat}
\usepackage[numbers,sort,compress]{natbib}
%\bibpunct{[}{]}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
\usepackage{endnotes}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
%\usepackage{mathtime}
\usepackage{bm}
\usepackage{url}
%\usepackage{babel}
\usepackage{dcolumn}
\usepackage{multirow}
%%%%%%

\usepackage[
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,anchorcolor=blue,
  citecolor=blue,filecolor=blue,
  menucolor=blue,pagecolor=blue,
  urlcolor=blue]{hyperref}

% Let's add todonotes and comments:
%\usepackage[draft]{todonotes}
%
%% Select what to do with command \comment:  
%% \newcommand{\comment}[1]{}  %comment not showed
%\newcommand{\comment}[1]
%{{[\bfseries \color{blue} #1]}} %comment showed

\usepackage{lineno}
\usepackage{float}
\usepackage[anythingbreaks]{breakurl}

\title{Reconciling agriculture, carbon, and biodiversity in a savanna transformation frontier}
\author[1,2]{Estes, L.D.}
\author[2]{Searchinger, T.}
\author[2]{Spiegel, M.}
\author[2]{Tian, D.}
\author[3]{Kehoe, L.}
\author[3]{Kuemmerle, T.}
\author[1]{Caylor, K.K.}
\affil[1]{Civil and Environmental Engineering, Princeton University, Princeton, NJ, 08544 USA}
\affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ, 08544 USA}
\affil[3]{Humboldt University, Berlin}
\date{}

\begin{document}
\maketitle
%\begin{frontmatter}

\begin{abstract}
Blah blah
\end{abstract}
%% Text of abstract

%\begin{keyword}
%keywords
%\end{keyword}

%\end{frontmatter}

%\linenumbers

%% main text
\section*{\large Introduction}

   Meeting growing food demands while minimizing loss of native ecosystems and their carbon and biodiversity presents one of the world’s great challenges for food and the environment.  Many global sustainability studies and climate mitigation pathways call for eliminating or rapidly phasing out emissions from land use change.  CITATIONS IPCC AR V; UK SCIENCE OFFICE.  Unfortunately, these ambitions do not appear realistically attainable in sub-Saharan Africa due to a population likely to more than double to 2.1 billion by 2050. For example, even with healthy linear yield growth from 2006 to 2050 at 3\% to 5\% per year projected by the UN Food and Agriculture Organization (FAO), the region would likely need to add more than 100 million hectares of cropland by 2050.  Searchinger et al. 2013. Much and probably the great bulk of that new cropland is likely to occur in the ~500 million hectares of savannas and shrublands of the region that receive sufficient rainfall to permit crop production.  Searchinger et al. 2013.  This conversion would release tens of millions of tons of carbon and have large impacts on biodiversity.

[I would add in here something about the fact that this is a major challenge, it is one that needs careful planning to achieve, but that this planning has to be done at the country level as this is where most land use policy decisions are made.  Here we look at Zambia, which is perhaps best representative of the challenge.  [this ties us back to the call we made in the NCC paper for finer-level planning, and also inoculates us from the parochiality problem, as we are looking at a representative country].]

Zambia presents an acute part of the challenge.  According to the medium estimate of the United Nations, Zambia’s population is likely to grow 265\% from 16.2 million in 2015 to 43 million in 2050.  (United Nations Department of Economic and Social Affairs Population Division.  2015.  World Population Prospects:  The 2015 Revision, DVD Edition).   As per capita calorie food availability in Zambia is also low, hunger indices high (with child stunting rates at  40\%), https://www.wfp.org/stories/10-facts-about-hunger-zambia GET BETTER CITATION) , and the percentage of the population dependent on agriculture at X\%, Zambia has great need to boost its agricultural production.  Zambia has also been expanding its agricultural exports, and has agronomic and situational characteristics that put it in a strong position to expand exports to other parts of the region.[1]

At the same time, two thirds of Zambia’s 75 milloin hectares is dominated by Miombo Woodland considered dense to be categorized as forest by the UN REDD program, or X\%.  (Vinya, R., Syampungani, S., Kasumu, E.C., Monde, C. \& Kasubika, R. (2011). Preliminary Study on the Drivers of Deforestation and Potential for REDD+ in Zambia. A consultancy report prepared for Forestry Department and FAO under the national UN-REDD+ Programme Ministry of Lands \& Natural Resources. Lusaka, Zambia).  Using this measure of forest, Zambia’s deforestation rate has been estimated at 250,000 to 300,000 hectares per year, among the highest rates of deforestation of any country in the world. Ibid. Fifty percent of these X million hectares have carbon stocks in excess of X tC/h, and an additional X\% have carbon stocks in excess of X.

The biodiversity of these woody savannas is also high.  Overall, the wetter woodland-savannas and shrublands of Africa have a similar average richness for birds and mammals as the world’s wetter tropical forests, and 75\% of these habitats have higher overall vertebrate diversity than 75\% of the rest of the world.   (Searchiner et al. 2015).   Most of Zambia is covered by what the World Wildlife Fund has categorized as Central Zambian Miombo Woodland.  It contains mature miombo woodland trees typically 15 to 20 meters tall, around a quarter of which are evergreen, with a broadleaf shrub and grass understory beneath.[2]  It is also interspersed with a large quantity of wetlands, up to 30\% of many areas.   Its large expanse of habitat make it a center of large mammals, including elephant, black rhino and African buffalo.  Although endemism is low, birdlife is rich with X species, and reptile and amphibian endemism is high, with 19 endemic reptile species 13 endemic amphibians.  LYNDON, CAN WE ADD SOME NUMBERS OF PLANT SPECIES, AND VERTEBRATE SPECIES.   Zambia contains eight other ecoregions according to WWF classifications.[3]   Its extensive wetlands are particular centers of bird and animal life and under extensive pressures.  The value of this habitat is based in part on its remaining relative abundance, which makes it possible for wide-ranging mammals, including many predator species, to survive.

These multiple trends imply challenging trade-offs between agricultural development and environmental concerns.  They raise the question how could decision-makers best target agricultural expansion so that it balances economic, carbon and biodiversity concerns.  A related question is how could a system for targeting well illuminate the trade-offs to support decision-making and a proper role for influence by stakeholders.  And an important question is whether if so what benefits could be achieved by targeting land expansion in certain areas from the standpoint of achieving areas with a desirable mix of objectives.

[We need a somewhat clearer phrasing of our objectives here, or we need to state the objective below the next paragraph. What is that objective: two-fold I would say. 1) Is this sort of modeling approach useful and/or does it have realistic potential for finding compromises in a more general sense (i.e. can it be applied beyond just Zambia), 2) how much potential does Zambia appear to have for reducing the impacts of its expansion. Something along these lines.]

There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs subject or while limited land use budget for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}. Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.  We adopt a similar approach here, using a newly developed model in which varying weights can be placed on the different objectives, thereby allowing optimal spatial allocations to be found that reflect different degrees of compromise between competing land use interests. To enable this functionality, the model is structured to evaluate each objective in terms of efficiency, or the ratio of cost to yield benefit that results from converting land to agriculture in order to achieve a production target.

We apply this model to answer a basic but important question: can agricultural expansion in Zambia be achieved for substantially lower environmental cost than a business as usual approach, where expansion occurs without consideration of its ecological consequences? This question is particularly important for tropical savanna shrubland, and grassland biomes, which are commonly assumed to provide a low environmental cost reserve for agricultural development \cite{searchinger_high_2015}.

\section*{\large Materials and Methods}
The model we developed, called \emph{AgroEcoTradeoff}, is designed to find the best land use configuration to satisfy the production targets for multiple crops while minimizing four costs: the total land area required, which is inversely related to productivity; the potential losses of carbon and biodiversity; and transportation costs, which are closely correlated with agricultural production costs (ref).  The model takes the basic form:

\begin{equation}
B_{ijk} = \frac{J_{jk}}{P_{ij}} \quad i \in m, \ j = 1, \ldots,n, \ k = 1, \ldots,q 
\end{equation}

Where J$_{k}$ is one of the four costs to be minimized, within a landscape composed of $n$ grid cells containing values for each of the $q$ cost objectives and potential productivity values $P_i$, for each of $m$ crops. $B_{ijk}$ therefore is an efficiency that describes the ecological cost per unit of production, which is normalized so that values are between 0 and 1, so that the different objectives can be compared in common units on a common scale. The normalized values are subtracted from 1 to equate high values with high probability of conversion. Each of these probabilities is then multiplied by a weight (range 0-1, constrained to sum to 1 across all four weights), and the four resulting products are summed, producing a scalarized single objective that represents the landscape conversion probability for each individual crop in each grid cell. An objective function seeks to maximize the sum of these conversion probabilities, based on the requirement that the sum of the potential production of the cells converted for each crop must meet that crop's production target. A second constraint restricts each cell so that it is converted to no more than one crop, as most crops are grown during the same season, and producing multiple crops on the same land would introduce further complexity in setting assumptions about the share of each crop allocated to shared pixels. We therefore formulated a greedy approximation algorithm, based on solving the knapsack problem  \citep{dantzig_discrete-variable_1957} and an algorithm for solving the generalized assignment problem \citep{cohen_efficient_2006}, which allocates each crop to its highest value cells.

This formulation allows quantitatively optimal solutions to be found across a wide range of weight combinations, which is valuable because it is rare to find a single solution that simultaneously obtains the lowest cost for each objective. By finding the optimal solutions across a range of weights, the model can identify the Pareto front, or the set of solutions where the cost in one objective cannot be further minimized without increasing the cost in another. This set then allows model users to choose the most appealing combination of weights based on their land use preferences, and their own assessments of what are and are not acceptable costs within each of the four competing objectives.

This weighting system thus has two basic components.  One reflects the pure value judgment for one output versus another, e.g., how much more a person cares about yield over carbon, or carbon over biodiversity.  The other component expresses each output in quantitative terms so that it reasonably expresses preferences for that output.  For example, units for the carbon, biodiversity and yield outputs are scaled so that someone seeking to weight yield at twice the value each of carbon and biodiversity can do so and have the model find an outcome that reasonably reflects this preference. This approach is necessary to support public decision-making processes, in which different individuals and stakeholders might assign different values to different interests, and where compromise solutions must be sought.

\subsection*{Model inputs} 
For conducting our analysis of potential land use tradeoffs in Zambia, we developed the following spatial datasets at a resolution of 1 km$^2$.

\subsubsection*{\emph{Agricultural potential and production scenarios}}
There are many ways of evaluating agricultural potential factoring in agronomic and economic characteristics.  Here we focused on two crops, maize and soybeans, because of their existing and potential significance for Zambia’s agricultural sector, and because they are considered to be the two dominant commodities that will be grown in Africa's broader savanna regions \citep{searchinger_high_2015}.  Maize is the largest crop by area and consumption in Zambia \citep{fao_fao_2012}, while soybeans are expanding rapidly as a commodity crop to meet rapidly growing regional and global demands for animal feed and direct human consumption \citep{gasparri_emerging_2015}.

We mapped potential yields for each crop using a three-step approach. In the first step, we used the decision support system for agrotechnology transfer \cite[DSSAT, ][]{jones_modelling_2003} to simulate maize and soybean yields at the sites of 40 weather station distributed across Zambia, which represent the points where a 31-year, gridded daily meteorological dataset has received the greatest amount of bias-correction and infilling by weather observations \citep{sheffield_development_2006,estes_changing_2014}. We used soil profiles corresponding to those locations drawn from the WISE v1.1 gridded soil profile database \citep{harvestchoice_converting_2010}, and simulated for each location maize yields over all 31 years under commercial input practices for three-four cultivars representative of short and medium season length according to the Zambian crop calendar \cite{grassini_how_2015}.  This step allowed us to characterize crop responses to Zambia’s inherent climatic variation, and to improved management practices, which was otherwise impossible given the coarse resolution of available observed yield data and the fact that much of Zambia’s current production occurs on low input, smallholder farms (cite). In the second step, we used an empirical model to map the DSSAT-modeled yields onto a 1 km grid, taking advantage of a finer resolution soil map that provides a common frame for estimating both yield potential and soil carbon stocks (see below), but has insufficient detail in several parameters (particularly effective rooting depth) required by DSSAT.  We used a generalized additive model \citep{wood_mgcv:_2001} to predict the values of the DSSAT-simulated yields, using best-fitting subsets of the following weather and soil variables: growing season precipitation; growing degree days; mean growing season shortwave solar radiation; percent organic carbon; pH; percent clay content. The soil predictors represented mean values in the top 1 m of soil, and were extracted from a new 1 km global database of soil properties \cite{hengl_soilgrids1km_2014}.  These first two steps were necessary to 1) capture crop responses to Zambia's climatic variability and potential management practices, and 2) to base potential yields on the same dataset used to estimate carbon costs (see below). This resulted in a map of potential yield under relatively high inputs and good management, and assuming no losses due to other limitations such as diseases. These yield values therefore overestimated what farmers are truly likely to achieve. To correct this, we therefore rescaled the yield maps to match the mean yields (4.5 t/ha for maize and 3.6 t/ha for soybean) that are projected to be achieved in Zambia by 2050, using FAO-projections to derive an annual growth factor, which we then applied to the 2009-2014 reported mean yields for each crop. Overall, the distribution of yields is most important to targeting land use, but the average yield determines the total quantity of land needed to meet production targets.

As a second measure related to agricultural potential, we estimated the travel time to the nearest city of 10,000 people or more (check this), which serves as a proxy for production costs, such as fertilizer prices and transport costs (citation), and is strongly correlated with the suitability of land for farming (cites from ConBio paper).  We used a cost-distance analysis drawing on spatial data for roads, rivers and water bodies, urban centers inside and neighboring Zambia, and national boundaries to calculate travel time in hours. Further details on the cost distance and yield estimation methods can be found in the online electronic supplement.

\subsubsection*{\emph{Carbon}} 
To estimate potential carbon costs from cropland conversions, we developed maps of vegetative and soil carbon stocks for Zambia. Vegetative carbon was calculated from a map of above-ground biomass developed by Baccini et al. \citep{baccini_estimated_2012}, from which we estimated below-ground biomass using savanna and miombo woodland-specific root-shoot ratios \cite{mokany_critical_2006,mugasha_allometric_2013}, and then converted the sum to carbon using a ratio of 0.47 (IPCC 2006 cite). We used the same soil carbon dataset used for crop yield prediction together with an accompanying map of bulk density to calculate soil carbon stocks in the top 1 m. The potential carbon loss due to agricultural conversion was calculated as 100\% of vegetative carbon and 25\% of soil carbon \cite{searchinger_high_2015}.

\subsubsection*{\emph{Biodiversity}}
There are many ways of evaluating biodiversity value and potential impacts, including species richness of different taxa in absolute terms or based on level of threat or endemism, and including factors designed to address the significance of contiguity or the impact of different types of development \citep[e.g.][]{newbold_global_2015,searchinger_high_2015, gasparri_emerging_2015,koh_spatially_2010}. Such measures should reflect both conservation priorities and the quality of available data.  For a fine-scaled, country-level analysis such as this one, the species range map data that form the basis for many impact assessments (Kehoe et al, 2015) have an effective resolution that is too coarse (Laura's cite). We therefore developed a composite measure of biodiversity value based on several datasets. First, we used the potential vegetation maps of East Africa \citep[][]{van_breugel_potential_2011}, a 2010 landcover dataset for Zambia (SERVIR cite), and a shapefile of Zambian protected areas to calculate how much of Zambia's 20 unique vegetation types remains after conversion to croplands and settlements, and what proportion of that remainder falls within national parks. From this we calculated a hybrid measure of rarity and threat, wherein we assigned a weight to each vegetation type based on its relative remaining area in proportion to Zambia's extent, log-transformed to account for a highly skewed distribution, and then re-weighted that weight by each vegetation types protected fraction, following \citep{van_breugel_environmental_2015}. We then used the remaining vegetation map to calculate a second measure, the proportion of untransformed vegetation within a 11X11 km neighborhood centered on each 1 km$^2$, providing a measure of vegetation intactness. By adding this measure to the first index and normalizing, we created a unitless biodiversity score (range 0-1) that gives equal weight to how rare or threatened a habitat is, and how undisturbed it is.

As a final modification, we masked out the areas of existing national parks and game management areas, on the assumption that they would not be used for expanding farmland, but retained a third protected category, forest reserves, as these tend to be far less stringently protected, having had 8.3\% of their area converted to cropland or settlements according to the landcover data. We assigned these areas a high biodiversity score of 0.917 (1 - the already converted proportion), indicating the value of keeping such areas protected, but recognizing that they have some chance of being converted, which can be estimated from their current degree of transformation.

\subsubsection*{\emph{Potential farmland and production targets}}
Our tradeoff analysis is concerned with cropland expansion into areas that are currently not farmed but could be. To identify such areas, we calculated the proportion of currently cropped and settled grid cells within each 1 km$^2$ grid cell, and used a 30 m digital elevation model to calculate slope percent over the entire country, and calculated the proportion of cells greater than 20\%. Slopes above this percent are considered unsuitable for farming (USDA cite), and slope-based measures are strong predictors of cropland presence or absence \citep{estes_using_2014}. Adding the two resulting maps together, subtracting their value from 1, and removing protected areas provided the fractional potential cropland in each cell. In addition to defining cropland availability, this map also provided the area term necessary for calculating production from potential yield.

To develop the production targets used by the model, we calculated the rate of change in annual production of maize for Zambia between 2000-2014, collected from the FAOStat database. For soybean we did the same, but used the mean production trend across the 9 southern African countries examined by Gasparri et al \citep{gasparri_emerging_2015}, given that soybean production is very underdeveloped in Zambia compared to countries such as South Africa that are currently experiencing rapid growth. Extrapolating linearly from 2015 until 2050, maize production will triple and soybean production will increase 9-fold. We used these linear extrapolations as the basis for our targets, adding an additional unit to each factor to account for the fact that Zambia, which has one of the highest agricultural potentials in sub-Saharan Africa, is likely to be subject to larger demands for exports to many of its neighbors, thus our final targets were a four-fold increase for maize (from 2009-2014 average production of 2,912,000 tons) and a 10-fold increase for soybean (from 203,700 tons).

Given that current Zambian maize and soybean yields are well below their current potentials (Waddington et al, 2010), it is an open question how much of future production needs will be satisfied by boosting production on existing croplands. However, it seems likely that some portion of this target will come from current land, given recent Zambian yield growth trends (FAOStat).  In our production scenario, we therefore assumed that an amount equal to a quarter of the gap (redo) between current and projected yields would be met on current farmland before significant expansion occurs.

\subsection*{Simulations}
Using the selected production targets, we ran the model across all combinations of weights for the four parameters, varying the weights in 5\% increments. This allowed us to evaluate the impact of different land use preferences on the location of conversion impacts, the range and magnitude of potential conversion impacts, including for each objective, and to test how sensitive costs in each objective were to changes in weights.

\section*{\large Results}
The first order determinant of land use conversion costs is unrelated to any weights applied to the four objectives, as it is based on how much of future production needs will be met on existing croplands, and therefore how much non-agricultural land will be spared from conversion. In our main scenario, we assumed that the gap between current and our projected future potential yields would be fully closed, but to illustrate the impact of this assumption on land conversion costs, we also ran the model (with equal weights across all objectives) for cases when 50\% and 150\% of the gap was closed. In the former case, the total land area required to meet the target increased by 18\% (14,634 to 17,202 km$^2$), average travel time increased 3\% (1.96 to 2.01 hours), carbon losses increased by 19\% (597 to 711 mt), and the average biodiversity index of converted areas increased by 2\% (0.297 to 0.303). In the latter case, the impacts were reversed, with impacts declining by almost exactly the same percentage.

To identify the minimum possible costs for each individual objective, and the degree to which their areas of new cropland correspond, we examined the simulations were 100\% of the weight was given to each objective. The yield maximizing objective (100\% weight to yield) describes both a cost in terms of the land area required, and a benefit that is based on the productivity gained from converting that land. Scenarios that involve the lowest land use costs also provide farmers the highest potential yields, and therefore have economic advantages.  An examination of overlaps shows two interesting results. First, converted areas that would maximize yield have little correspondence with the areas that minimize carbon costs, biodiversity impacts or travel time (Figure 1), as measured along two dimensions: actual shared area, as measured by the percentage overlap between the two maps, and their adjacency to one another, calculated as the average distance between each pixel of one map and its nearest neighbor in the other. Second, the areas selected for minimizing carbon, biodiversity, and transportation costs have high adjacency, albeit little overlap. All three criteria tend to identify lands along existing road networks and population and agricultural centers. The reason for this correlation is that carbon stocks and biodiversity values are usually lowest in areas of high human density, because of high levels of charcoal and wood harvesting (cite) and fragmentation caused by settlements, existing croplands, and road networks, which exist primarily to connect agricultural and urban centers.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/single_priorities2.png}}
    %\includegraphics[scale=0.5]{fig2.eps}
    \caption{Areas converted to new maize and soybean croplands and their associated costs when each individual objective (yield maximization, and carbon cost, biodiversity cost, and transport time minimization) is given 100\% weight.  The impact categories are color-coded to the specific conversion maps.}
    \label{default}
  \end{center}
\end{figure}

One of the goals of this analysis is to determine whether opportunities exist to achieve a reasonable accommodation between different objectives, and here the results are also interesting. While the best case scenario for maximizing yield would convert  million hectares, the worst case scenario (which would minimize carbon costs) would convert Y million hectares, only a Z\% increase, which is far lower than the variability in potential crop yields (3.7-6.1 t ha$^{-1}$ for maize and 0.8 to 5.1 ha$^{-1}$ for soybean). This result reflects our model's use, and the value, of an efficiency approach. Because yield is in the denominator of each objective (e.g. carbon loss per ton of yield, or biodiversity value per ton of yield), land with better yield potential tend to be selected--even when the yield maximization priority is given no weight--so that the crop production can be achieved with the least cost.

By contrast to the limited differences for yields, the differences in transportation costs using existing transportation networks range from an average of 0.6 hours for each hectare to 6.9 hours, an increase of 1050\%.  Carbon costs from minimum to maximum vary by 188\%, and average biodiversity impacts vary by 100\%.  These differences suggest that there is high potential to reduce transportation, carbon and biodiversity costs for relatively sacrifice of yield potential.

This potential is illustrated in Figure 2, which shows how much cost is avoided in each objective when weights are transferred between them. Here avoided cost is calculated as the percent difference between the cost paid under a given weight combination and the largest cost that is incurred in all combinations (the worst case). For carbon, biodiversity, and travel time, the maximum cost occurs when yield is given 100\% weight, which is where the yield objective has the lowest possible cost, as measured by the percent of land saved relative to the largest area converted, because the cropland that is converted is the highest yielding. However, by sequentially adding weights in 5\% increments to both biodiversity and carbon conservation objectives, large reductions in not only those costs but also in average travel time (because of the correlations between these three measures). Weighting carbon and biodiversity by 15\% each reduces the average area of new lands (i.e. new crop area is increased) by just 3\%, in exchange for avoided costs of 34\% for biodiversity, 48\% for carbon, and 44\% for travel time, which respectively represent 68\%, 77\%, and 50\% of the maximum possible avoided cost for each objective.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/weight_impact3.png}}
    %\includegraphics[scale=0.5]{fig2.eps}
    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.}
    \label{default}
  \end{center}
\end{figure}

%These substantial cost savings suggest that a user  achieved for relatively little transfer of weight between objectives indicates that a user primarily concerned with acquiring the highest-yielding land for agriculture would have to make relatively little sacrifice to give enormous gains to conservation interests, or to substantially reduce the potential costs of their farming operation. value of compromise in terms that users might be more likely

These results show that transferring relatively little weight between objectives can have large savings. However, people involved in making land use decisions are more likely to evaluate their willingness to compromise in actual cost units, and from the perspective of how much more they might be willing to pay to compromise with other interests. We therefore evaluated four further scenarios, in which a hypothetical land user with interests in a single objective willingly pays up to 5\% more than the lowest cost case (i.e. the 100\% weight outcome) in order to compromise with other interests. The yield maximizing user is thus willing to ``pay'' for an extra 68,400 km$^2$ of new cropland, the carbon and biodiversity conservationists would respectively sacrifice an additional 2,312,000 tons of carbon and 0.013 biodiversity index units, and the production cost-minded agricultural developer would pay just 4.1 minutes of travel time. Given these sacrifices, we identified how much more the other three interests would have to pay relative to their lowest cost (100\% weight) scenarios, treating each of the three equally, i.e. the amount each has to pay extra is the same percentage of their lowest cost case.

The impacts of this assessment show that in for all four cases, the carbon, biodiversity, and travel time objectives typically pay less than 50\% more than their lowest cost when another party is willing to pay 5\% more cost (Figure 3), and all four associated conversion maps show substantial convergence around road networks and population and agricultural centers.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_5pct.png}}
    %\includegraphics[scale=0.5]{fig2.eps}
    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.}
    \label{default}
  \end{center}
\end{figure}

Figure 3 also maps and shows the results for the equal weighting option and how it contrasts with the minimum costs for each factor.  It shows increased savings for each factor of X to X for transportation costs, carbon and biodiversity compared to their maximum costs.  It also holds these costs to no more than X above their minimum costs.

\begin{figure}[!ht]
  \begin{center}
       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_25pct.png}}
    \caption{25\% compromise}
    \label{default}
  \end{center}
\end{figure}

Another interesting feature of these results is that there is little overlap between any targeting areas selected by any of the criteria we have calculated and designated farm development blocks.  In 2005, the Zambian Ministry of Agriculture designated nine farm blocks with roughly one million hectares for assisted agricultural development and spread around the country.  (Zambia Ministry of Agriculture. 2005. Farm Block Development Plan 2005-07).  Although these areas have in fact as a whole attracted little development, they remain “on the books,” and there are at least some reports of continued intent to develop them.  Undeveloped farm blocks worry Agriculture Minister, Zambia Daily Mall: https://www.daily-mail.co.zm/?p=29505

%\begin{equation}
%  \textrm{I} = \left(\frac{\textrm{A}}{\textrm{A + C}} + \frac{\textrm{D}}{\textrm{B + D}}\right)0.5
%\end{equation}

\section*{\large Discussion}

\section*{\large Data accessibility}
\section*{\large Authors' contributions}
\section*{\large Competing interests}
\section*{\large Funding}
\section*{\large Acknowledgements}
This work was supported by funds from: the Norwegian Aid Agency (NORAD) under the Agricultural Synergies Project; the Princeton Environmental Institute Grand Challenges program; the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463).

\section*{\large Supplementary materials}

%\label{}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%\section*{\large References}
\bibliographystyle{prsb} 
\bibliography{/Users/lestes/Dropbox/publications/full}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

%\begin{thebibliography}{00}
%
%%% \bibitem[Author(year)]{label}
%%% Text of bibliographic item
%
%\bibitem[ ()]{}
%
%\end{thebibliography}
\end{document}

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.

### 2016-02-26 paper/supplementary_methods.tex
Near final figures for main text done after revision to Fig 1 and 3, and new Figure 4 made. Model rerun for 100% yield gap closure on existing land.

\documentclass[a4paper]{article}
\usepackage[affil-it]{authblk}  
%\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}  % apostrophe's in cites
\def\linenumberfont{\normalfont\small\sffamily}
\usepackage{graphicx}
\usepackage{epstopdf}
% From link below: next two lines allow caption to fill the full figure box, and caption formats to be changed (e.g. font)
% http://tex.stackexchange.com/questions/107350/caption-below-the-figure-and-aligned-with-left-side-of-figure
% http://ctan.mackichan.com/macros/latex/contrib/caption/caption-eng.pdf
\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{caption}
\captionsetup[figure]{slc=off, font=footnotesize}
\captionsetup[table]{margin={4cm, 4cm}, slc=off, font=footnotesize}
\usepackage[numbers,sort,compress]{natbib}

\begin{document}
\section*{\Large SI Methods}

\section*{\large Model overview}

The agroEcoTradeoff model was designed to identify optimal land use configurations that allow a certain level of agricultural production to be achieved while minimizing associated ecological and economic costs. The model is based on that developed by Koh and Ghazoul \citep{koh_spatially_2010}, which was designed to select areas with promising potential for oil palm production but whose conversion would result in less loss of carbon, avian diversity, and land with high rice production potential. In this case, each unit of land was ranked based on its relative value within each measure, which was potential productivity for oil palm and the inverse of its potential rice productivity, carbon density, or bird diversity. Normalizing these values converted each into an oil palm conversion probability \emph{c} that was conditioned on the individual land value, such that the highest ranking areas were converted until a specified oil palm production target was reached. For example, in the case of carbon conservation being the priority, lands were sequentially selected from the least carbon dense to the most carbon dense, without concern for any other objective. By multiplying the four \emph{c} values together into a joint conversion probability \emph{C}, each land value was given equal weight, achieving a satisfactory tradeoff between the four different land use objectives.

We retained the basic rationale underlying this model in developing our own, wherein tradeoffs between yield maximizing and ecological cost minimizing objectives can be made in achieving a crop production target, with several amendments.  First, the model has the capacity to simulate more than one crop. Second, the ranking of land units is based on the cost/benefit ratio of converting it to meet the target, thus areathat maximize benefits while minimizing costs along a particular cost dimension.

We adapt this approach by recasting the model in terms of efficiency, where land units are ranked in terms of its potential productivity relative to the cost that will be incurred from its conversion.

%\begin{figure}[!ht]
%       \begin{center}
%      \includegraphics[scale=1]{figures/figS1.pdf} 
%        \end{center}
%     %\vspace{-1cm}
%      \caption{The cumulative number of workers seeking qualification during the South Africa cropland mapping trial period. Successful qualification seekers are indicated by blue circles, unsuccessful by open circles.}
%      \label{fig:default}
%\end{figure}
%\section*{\large References}
\bibliographystyle{prsb} 
\bibliography{/Users/lestes/Dropbox/publications/full}

\end{document}

### 2016-02-26 paper/supplementary_results.tex
Near final figures for main text done after revision to Fig 1 and 3, and new Figure 4 made. Model rerun for 100% yield gap closure on existing land.

--- 2016-02-26 paper/supplementary_methods.tex
@@
 \begin{document}
-\section*{\Large SI Methods}
+\section*{\Large SI Results}
 
-\section*{\large Model overview}
+\section*{\large Crop yields}
 
-The agroEcoTradeoff model was designed to identify optimal land use configurations that allow a certain level of agricultural production to be achieved while minimizing associated ecological and economic costs.
-The model is based on that developed by Koh and Ghazoul \citep{koh_spatially_2010}, which was designed to select areas with promising potential for oil palm production but whose conversion would result in less loss of carbon, avian diversity, and land with high rice production potential.
-In this case, each unit of land was ranked based on its relative value within each measure, which was potential productivity for oil palm and the inverse of its potential rice productivity, carbon density, or bird diversity.
-Normalizing these values converted each into an oil palm conversion probability \emph{c} that was conditioned on the individual land value, such that the highest ranking areas were converted until a specified oil palm production target was reached.
-For example, in the case of carbon conservation being the priority, lands were sequentially selected from the least carbon dense to the most carbon dense, without concern for any other objective.
-By multiplying the four \emph{c} values together into a joint conversion probability \emph{C}, each land value was given equal weight, achieving a satisfactory tradeoff between the four different land use objectives.
+\begin{figure}[!ht]
+       \begin{center}
+       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/ideal_yield2.png}}
+        \end{center}
+      \caption{The distribution of converted areas when the model is run separately for maize and soybean when 100\% is placed on yield maximization.
+The map is color-coded according to which crop was allocated to each location (red for maize, blue for soybean, and yellow for the places where both crops were allocated).}
+      \label{fig:default}
+\end{figure}
 
-We retained the basic rationale underlying this model in developing our own, wherein tradeoffs between yield maximizing and ecological cost minimizing objectives can be made in achieving a crop production target, with several amendments.
-First, the model has the capacity to simulate more than one crop.
-Second, the ranking of land units is based on the cost/benefit ratio of converting it to meet the target, thus areathat maximize benefits while minimizing costs along a particular cost dimension.
-
-We adapt this approach by recasting the model in terms of efficiency, where land units are ranked in terms of its potential productivity relative to the cost that will be incurred from its conversion.
-
-%\begin{figure}[!ht]
-%       \begin{center}
-%      \includegraphics[scale=1]{figures/figS1.pdf} 
-%        \end{center}
-%     %\vspace{-1cm}
-%      \caption{The cumulative number of workers seeking qualification during the South Africa cropland mapping trial period.
-Successful qualification seekers are indicated by blue circles, unsuccessful by open circles.}
-%      \label{fig:default}
-%\end{figure}
-%\section*{\large References}
+\section*{\large References}
 \bibliographystyle{prsb} 

### 2016-02-27 Rmd/figures.Rmd
First full (very rough) draft assembled

@@
 imp_pct <- cbind(cbind(ywgts, xwgts), cstats, imp_pct)
+
+((minmaxstats[2, c(2:3, 8, 5), with = FALSE]) - 
+  (imp_pct[4, c(4:5, 7, 6), with = FALSE])) /
+  imp_range[, c(2:3, 8, 5), with = FALSE]
 
@@
 # # figure titles
-# ftits <- lapply(pind, function(x) {
-#   paste(sapply(1:4, function(y) {
-#     paste(paste(cnames[y], paste0(parms[x, y] * 100, "%")))
-#   }), collapse = " ")
-# })
+ftits <- lapply(pind, function(x) {
+  paste(sapply(1:4, function(y) {
+    paste(paste(cnames[y], paste0(parms[x, y] * 100, "%")))
+  }), collapse = " ")
+})
 # leg <- c("Converted")
@@
 # Plot setup
+# Plot setup
+m1 <- do.call(cbind, lapply(1:4, function(x) matrix(rep(x, 35), 5, 7)))
+m2 <- do.call(cbind, lapply(c(5, 8, 11, 14), function(x) {
+  a <- matrix(rep(x, 16), 4, 4)
+  b <- matrix(rep(x + 1, 6), 2, 3)
+  cbind(a, rbind(b, b + 1))
+}))
+lmat <- rbind(m1, m2)
+
 ftits <- list("Compromise 5% on Yield", "Compromise 5% on Carbon", 
@@
   axis(2, at = yaxes[[i]], cex.axis = cx, las = 2, mgp = mgp, tcl = -0.2)
+  ind <- which(pms2bb[, paste0("Var", i)] == 1)
   with(pms2bb[ind, ], {
@@
 dev.off()
+
+for(i in 1:4) {
+  1 - print(all_stats[5, c(2:3, 8, 5), with = FALSE] / 
+              all_stats[i, c(2:3, 8, 5), with = FALSE])
+}
+
+for(i in 1:5) {
+  print(((minmaxstats[2, c(2:3, 8, 5), with = FALSE]) - 
+     (all_stats[i, c(2:3, 8, 5), with = FALSE])) /
+    imp_range[, c(2:3, 8, 5), with = FALSE])
+}
 

### 2016-02-27 paper/main.tex
First full (very rough) draft assembled

@@
 \usepackage{graphicx}
+\usepackage{tabularx}
 \usepackage{epstopdf}
@@
 
-%\linenumbers
+\linenumbers
 
@@
 
-These multiple trends imply challenging trade-offs between agricultural development and environmental concerns.
-They raise the question how could decision-makers best target agricultural expansion so that it balances economic, carbon and biodiversity concerns.
-A related question is how could a system for targeting well illuminate the trade-offs to support decision-making and a proper role for influence by stakeholders.
-And an important question is whether if so what benefits could be achieved by targeting land expansion in certain areas from the standpoint of achieving areas with a desirable mix of objectives.
+These multiple trends imply challenging trade-offs between agricultural development and environmental concerns, and raise questions about what modeling approaches might best enable a society of multiple stakeholders to find optimal solutions.
+There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs or fitting a specific land use budgets for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}.
+Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.
 
-[We need a somewhat clearer phrasing of our objectives here, or we need to state the objective below the next paragraph.
-What is that objective: two-fold I would say.
-1) Is this sort of modeling approach useful and/or does it have realistic potential for finding compromises in a more general sense (i.e.
-can it be applied beyond just Zambia), 2) how much potential does Zambia appear to have for reducing the impacts of its expansion.
-Something along these lines.]
-
-There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs subject or while limited land use budget for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}.
-Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.
-We adopt a similar approach here, using a newly developed model in which varying weights can be placed on the different objectives, thereby allowing optimal spatial allocations to be found that reflect different degrees of compromise between competing land use interests.
-To enable this functionality, the model is structured to evaluate each objective in terms of efficiency, or the ratio of cost to yield benefit that results from converting land to agriculture in order to achieve a production target.
-
-We apply this model to answer a basic but important question: can agricultural expansion in Zambia be achieved for substantially lower environmental cost than a business as usual approach, where expansion occurs without consideration of its ecological consequences?
+Focusing on the same categories of objectives, we here present a new model a new model in which varying weights can be placed on the different objectives in an iterative approach to allow decisionmakers and stakeholders to find optimal allocations of new cropland.
+This kind of approach can enable different stakeholders to find compromises between competing land use interests while still meeting production targets.
+We apply this model to Zambia to answer a basic but important question: is it possible to achieve a reasonable balance of different interests, and to reduce the carbon and biodiversity costs of expansion with limited impact on agricultural potential?
 This question is particularly important for tropical savanna shrubland, and grassland biomes, which are commonly assumed to provide a low environmental cost reserve for agricultural development \cite{searchinger_high_2015}.
@@
 One of the goals of this analysis is to determine whether opportunities exist to achieve a reasonable accommodation between different objectives, and here the results are also interesting.
-While the best case scenario for maximizing yield would convert  million hectares, the worst case scenario (which would minimize carbon costs) would convert Y million hectares, only a Z\% increase, which is far lower than the variability in potential crop yields (3.7-6.1 t ha$^{-1}$ for maize and 0.8 to 5.1 ha$^{-1}$ for soybean).
-This result reflects our model's use, and the value, of an efficiency approach.
-Because yield is in the denominator of each objective (e.g.
-carbon loss per ton of yield, or biodiversity value per ton of yield), land with better yield potential tend to be selected--even when the yield maximization priority is given no weight--so that the crop production can be achieved with the least cost.
+While the best case scenario for maximizing yield would convert 13,668 km$^2$, the worst case scenario (which would minimize carbon costs) would convert 15,302 km$^2$, only an 12\% increase, a range in outcomes that is substantially lower than range in potential crop yields (3.7-6.1 t ha$^{-1}$ for maize and 0.8 to 5.1 ha$^{-1}$ for soybean).
+This result reflects our model's use, and the value, of an efficiency approach; because yield is in the denominator of each objective (e.g.
+carbon loss per ton of yield), land with better yield potential tends to be selected--even when the yield maximization priority is given no weight.
 
-By contrast to the limited differences for yields, the differences in transportation costs using existing transportation networks range from an average of 0.6 hours for each hectare to 6.9 hours, an increase of 1050\%.
-Carbon costs from minimum to maximum vary by 188\%, and average biodiversity impacts vary by 100\%.
+In contrast to the limited range in total converted area, the differences in travel times range from an average of 0.8 hours for each hectare to 6.7 hours, an increase of 710\%.
+Carbon costs increased by 156\% (462 to 1184 Mt), while the worst case for biodiversity was 93\% higher, with the mean biodiversity index of converted areas increasing from 0.26 to 0.5.
 These differences suggest that there is high potential to reduce transportation, carbon and biodiversity costs for relatively sacrifice of yield potential.
 
-This potential is illustrated in Figure 2, which shows how much cost is avoided in each objective when weights are transferred between them.
-Here avoided cost is calculated as the percent difference between the cost paid under a given weight combination and the largest cost that is incurred in all combinations (the worst case).
-For carbon, biodiversity, and travel time, the maximum cost occurs when yield is given 100\% weight, which is where the yield objective has the lowest possible cost, as measured by the percent of land saved relative to the largest area converted, because the cropland that is converted is the highest yielding.
-However, by sequentially adding weights in 5\% increments to both biodiversity and carbon conservation objectives, large reductions in not only those costs but also in average travel time (because of the correlations between these three measures).
-Weighting carbon and biodiversity by 15\% each reduces the average area of new lands (i.e.
-new crop area is increased) by just 3\%, in exchange for avoided costs of 34\% for biodiversity, 48\% for carbon, and 44\% for travel time, which respectively represent 68\%, 77\%, and 50\% of the maximum possible avoided cost for each objective.
+This potential is illustrated in Figure 2, which shows how much cost can be avoided within different objective by transferring increasing amounts of weight to them from the yield maximizing objective.
+Avoided cost is the percent difference between the cost paid under a given weight combination and the most expensive case.
+For carbon, biodiversity, and travel time, the maximum cost occurs when yield is given 100\% weight, which is where the yield objective has the lowest possible cost, as measured by the percent of land saved relative to the largest area converted (because the cropland that is selected is the highest yielding).
+However, by sequentially adding weights in 5\% increments to both biodiversity and carbon conservation objectives, large reductions in these costs as well as travel time are rapidly realized (because of the correlations between these three measures), but for little cost to the yield objective.
+For example, adding just 15\% weight each to carbon and biodiversity increases the average area of new lands (i.e.
+the average yield is decreased) by just 2.7\% (avoided area costs drop to 8\% from 10.7\%), in exchange for avoided costs of 33\% for biodiversity, 47\% for carbon, and 43\% for travel time, which respectively represent 75\%, 78\%, and 49\% of the maximum avoided cost for each objective.
 
@@
 
-These results show that transferring relatively little weight between objectives can have large savings.
-However, people involved in making land use decisions are more likely to evaluate their willingness to compromise in actual cost units, and from the perspective of how much more they might be willing to pay to compromise with other interests.
+Transferring relatively little weight between objectives can therefore result in large savings.
+However, people involved in making land use decisions are more likely to evaluate actual cost units with respect to how much more they would be willing to pay to compromise with other interests.
 We therefore evaluated four further scenarios, in which a hypothetical land user with interests in a single objective willingly pays up to 5\% more than the lowest cost case (i.e.
@@
 
-The impacts of this assessment show that in for all four cases, the carbon, biodiversity, and travel time objectives typically pay less than 50\% more than their lowest cost when another party is willing to pay 5\% more cost (Figure 3), and all four associated conversion maps show substantial convergence around road networks and population and agricultural centers.
+The impacts of this assessment show that in for all four cases, the carbon and biodiversity, and travel time objectives typically avoid more than 50\% of potential cost when any of the other objectives is willing to pay 5\% more cost (Figure 3), while the travel time objective avoids slightly less than 50\% of potential cost.
+Because of the small range in converted areas, a willingness to pay 5\% by the yield objective translates to just 50\% cost avoidance, while the 5\% compromise by carbon and biodiversity results in nearly the maximum potential cost in land area.
+The same compromise by the cost objective results in a bit less than 50\% cost avoidance by the yield objective.
+
+The location of the conversions associated with each of these compromises show substantial convergence along roads and population and agricultural centers, with overlap between each of the resulting cropland allocations exceeding 30\% between the yield and biodiversity and cost compromises, and 25\% between the biodiversity and carbon compromises.
+The carbon and cost compromises showed the lowest overlaps ($<$10\%) followed by carbon and yield ($\sim$15\%).
+The adjacency of each map was high, with all compromise allocations having an average nearest neighbor distance of $<$20km.
 
@@
     %\includegraphics[scale=0.5]{fig2.eps}
-    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.}
+    \caption{The impact of a willingness to pay 5\% more cost to compromise with other land use interests by each of the individual interests.}
     \label{default}
@@
 
-Figure 3 also maps and shows the results for the equal weighting option and how it contrasts with the minimum costs for each factor.
-It shows increased savings for each factor of X to X for transportation costs, carbon and biodiversity compared to their maximum costs.
-It also holds these costs to no more than X above their minimum costs.
+These weighting scenarios that corresponded to these selected compromises are shown in Table 1.
+The weight on the objective subject to the 5\% compromise is proportional to the size of its potential avoided cost, which is why yield requires a weight of only 40\% to pay 5\% or less of its lowest cost, whereas cost and carbon, which have the largest range of costs, require 85\% weight.
+
+\begin{table}[h]
+\begin{center}
+\caption{Weights corresponding to the scenarios in which each objective was a) willing to pay 5\% more than its lowest possible cost, and b) the costs for the other three objectives were minimized equitably.}
+\begin{tabular}{ccccc}
+\hline
+& \multicolumn{4}{c}{Weights}\\
+Objective making 5\% compromise & Yield & Carbon & Biodiversity & Travel time \\
+\hline\hline
+Yield & 40 & 10 & 5 & 45 \\
+Carbon & 0 & 85 & 5 & 10 \\
+Biodiversity & 0 & 10 & 55 & 35 \\
+Travel time & 15 & 0 & 0 & 85 \\
+\hline
+\end{tabular}
+\end{center}
+\label{default}
+\end{table}
+
+We also compared the degree to which these relatively small compromises by individual objectives differed from a situation in which all compromised equally (Figure 4).
+Equal compromise avoids over 80\% of the potential cost in the carbon, biodiversity, and travel time objectives, and 41\% of the cost in the yield objective.
+This is lower than, but compares favorably to, the $\sim$95\% savings achieved by the first three objectives within their individual cost categories, under their smaller levels of compromise, and to the 58\% savings by the yield objective.
 
@@
 
-Another interesting feature of these results is that there is little overlap between any targeting areas selected by any of the criteria we have calculated and designated farm development blocks.
-In 2005, the Zambian Ministry of Agriculture designated nine farm blocks with roughly one million hectares for assisted agricultural development and spread around the country.
+Another interesting feature of these results is that there is little overlap between the land allocated by the model under these scenarios and areas that are designated farm development blocks (<2\% overlap compared to the yield maximization and equal compromise scenarios; online electronic supplement).
+Zambia has a large number of these area, and as recently as 2005 the Ministry of Agriculture designated a further nine farm blocks with roughly one million hectares for assisted agricultural development.
 (Zambia Ministry of Agriculture.
@@
 Farm Block Development Plan 2005-07).
-Although these areas have in fact as a whole attracted little development, they remain “on the books,” and there are at least some reports of continued intent to develop them.
+Although these areas have in fact as a whole attracted little development, they remain ``on the books'', and there are at least some reports of continued intent to develop them.
 Undeveloped farm blocks worry Agriculture Minister, Zambia Daily Mall: https://www.daily-mail.co.zm/?p=29505
@@
 \section*{\large Discussion}
+
+Our results for Zambia suggest at least the potential to plan agricultural expansion such that it limits carbon and biodiversity costs with limited sacrifice of yield potential and with general consistency with existing transportation corridors.
+
+The fact that this potential mostly tracks existing transportation corridors would also appear to have several  advantages.
+Road construction has been a primary driver of the location of agricultural expansion.
+Laurance, Gossam \& Laurance 2009.
+Zambia is already engaged in a program to improve existing major roads, but there is evidence that agriculture responds in particular to improvements in the quality of smaller, feeder rural roads to these paved roadways.
+Focusing on these expansion areas would therefore appear to be one means of  focusing Zambia’s agriculture development.
+Because agricultural development tends to build on the private and public infrastructure that accompanies initial development, the focus on existing areas would also seem likely to be a more robust, long-term strategy for directing agricultural development \cite{laurance_estimating_2015}.
+Kassali, Rabirou, Ayanwale, Adeolu, Idowu, Ezekiel, \& Williams, Stella.
+“Effect of rural transportation system on agricultural productivity in Oyo State, Nigeria”.
+2012.
+Vol 113.
+Journal of Agricultural and Rural Development in the Tropics and Subtropics.
+
+This model appears to be the right kind of tool for assisting the planning of agricultural development.
+Optimization models that require ex ante specification of preferences in mathematical terms work less well for decision-making that must weigh very different objectives.
+Our model makes possible an iterative optimization process that allows people to realize their preferences in the face of real information about options.
+It is particularly suitable to decisions that will inevitably have strong political elements, and to consensus-building among stakeholders  with different preferences.
+
+For the tool to be truly legitimate and useful on a practical level, we believe several elements are required.
+First, it must be further developed and applied in an iterative way with the government and other stakeholders.
+Optimization decisions have many degrees of freedom, and closing them requires informed decision-making.
+Second, development planning must reflect and incorporate other real world factors beyond these simple four factors.
+For example, many factors beyond those incorporated into our biodiversity analysis have importance for biodiversity.
+For example, large, multi-lane roads can lead to high rates of mortality and serve as effective barriers to migration for large mammals and others regardless of other land uses (Dobson modeling paper about Serengeti highway).
+Special areas may be entitled to high protection regardless of our criteria, as our ranking system already excludes protected areas from potential development.
+Other infrastructure development will also influence the practicalities of agricultural development.
+
+Third, data quality is critical.
+This kind of tool focuses on particular hectares, not broad averages, so errors in data have little or no chance to ``balance out''.
+Crop modeling predictions and most of our other scores are highly sensitive to a range of data parameters.
+These facts mean that long-term use of the tool should be accompanied by steady data improvement.
+It also means that the tool should mostly be used as a guide to likely values.
+Before governments will make important decisions ``for keeps'', there should be site-specific analyses of soil properties, tests of soil fertility and validation of potential yield predictions, as well as confirmation of biodiversity and carbon characteristics.
+
+We also believe some elements of the model should be further developed although all depend on significant additional data gathering.
+Yield potential, and existing transportation costs, are only two indicators of the economic and practical prospects of agricultural development.
+Different areas of land development will require different quantities of inputs, including labor.
+An agricultural potential indicator based on a spatially differentiable economic model that better estimates both likely net returns to agriculture and agricultural labor would improve the model.
+
+Because the input use and therefore economics of commercial and small-scale agricultural development greatly differ, it would be useful to develop the model to differentiate their different potentials.
+The carbon calculations would benefit from site-specific root-shoot ratios.
+The biodiversity measures would benefit from incorporation of data on endemism, other wildlife use and contiguity needs of important animals.
+
+The model does not now incorporate analysis of areas with significant irrigation potential but should.
 
@@
 
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
-%\begin{thebibliography}{00}
-%
-%%% \bibitem[Author(year)]{label}
-%%% Text of bibliographic item
-%
-%\bibitem[ ()]{}
-%
-%\end{thebibliography}
 \end{document}

### 2016-02-28 Rmd/biodiversity-layer.Rmd
Supplementary methods draft complete. Required changes to biodiversity-layer.Rmd. Fixed bug in Figure 4 for main text--wrong map selected.

--- 2016-01-28 Rmd/biodiversity-layer.Rmd
@@
 s <- stack(lapply(fnms, function(x) raster(x)))
-sm <- mask(s, mgrid)
+# sm <- raster::mask(s, mgrid)
 
 # keep for assessing impacts
-writeRaster(sm, filename = full_path(p_bd, "ZA-vegremain.tif"), 
-            overwrite = TRUE)
+# writeRaster(sm, filename = full_path(p_bd, "ZA-vegremain.tif"), 
+#             overwrite = TRUE)
+sm <- brick(full_path(p_bd, "ZA-vegremain.tif"))
 vegnames <- gsub("\\.tif", "", gsub("veg", "v", vnms))
 save(vegnames, file = full_path(p_bd, "ZA-vegremain.rda"))
+names(sm) <- vegnames
 
 # Sum for total - the inverse of this will be the farmable areas
-remfrac <- calc(s, sum)
-remfrac <- mask(remfrac, mgrid)
-writeRaster(remfrac, filename = full_path(p_bd, "ZA-remfrac.tif"), 
-            overwrite = TRUE)
+# remfrac <- calc(s, sum)
+# remfrac <- mask(remfrac, mgrid)
+# writeRaster(remfrac, filename = full_path(p_bd, "ZA-remfrac.tif"), 
+#             overwrite = TRUE)
+remfrac <- raster(full_path(p_bd, "ZA-remfrac.tif"))
 
 # Calculate intactness
-w <- matrix(nrow = 11, ncol = 11)
-w[] <- 1
-intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
-writeRaster(intact, filename = full_path(p_bd, "intactness.tif"), 
-            overwrite = TRUE)
-
-# intact <- mask(intact, mgrid)
+# w <- matrix(nrow = 11, ncol = 11)
+# w[] <- 1
+# intact <- focal(remfrac, w = w, fun = mean, na.rm = TRUE)
+# writeRaster(intact, filename = full_path(p_bd, "intactness.tif"), 
+#             overwrite = TRUE)
+intact <- raster(full_path(p_bd, "intactness.tif"))
+intact <- raster::mask(intact, mgrid)
 # plot(intact)
@@
 wgti <- round(1 / (vtots / nrow(veg_dt)), 2)  # bd weights (1 = rarest)
-# wgt <- round(vtots / nrow(veg_dt), 4)  # biodiversity weights (1 = most common)
+wgts <- round((vtots / 100) / nrow(veg_dt), 4)  # veg prop (1 = most common)
 # plot(unlist(wgti), unlist(wgt))
@@
 # Combine protected areas (NPs and GMAs, forest reserves)
-nps$code <- ifelse(nps$TYPE == "gma", 2, 1) 
-npsr <- rasterize(nps, mgrid, field = "code")  # rasterize NPs/GMAs
-npsr[is.na(npsr)] <- 0  
-freserves$code <- 3  
-freservesr <- rasterize(freserves, mgrid, field = "code")  # forest reserves
-freservesr[is.na(freservesr)] <- 0
-pas <- npsr + freservesr  # sum them
-reclmat <- cbind(0:5, c(0:3, 1, 2))
-pas <- reclassify(pas, reclmat)
-pas <- mask(pas, mgrid)
-writeRaster(pas, filename = full_path(p_bd, "ZA-pas.tif"))
+# nps$code <- ifelse(nps$TYPE == "gma", 2, 1) 
+# npsr <- rasterize(nps, mgrid, field = "code")  # rasterize NPs/GMAs
+# npsr[is.na(npsr)] <- 0  
+# freserves$code <- 3  
+# freservesr <- rasterize(freserves, mgrid, field = "code")  # forest reserves
+# freservesr[is.na(freservesr)] <- 0
+# pas <- npsr + freservesr  # sum them
+# reclmat <- cbind(0:5, c(0:3, 1, 2))
+# pas <- reclassify(pas, reclmat)
+# pas <- mask(pas, mgrid)
+# writeRaster(pas, filename = full_path(p_bd, "ZA-pas.tif"))
+pas <- raster(full_path(p_bd, "ZA-pas.tif"))
 # plot(pas)
@@
 
-# Let's look at how much veg is protected, and maybe use that as basis for 
+# Let's look at how much veg is protected, and use that as basis for 
 # weights - rarity + proportion protected
-smpa <- mask(stack(s, pas), mgrid)
+names(s) <- vnms
+smpa <- raster::mask(stack(s, pas), mgrid)
 vegpa_dt <- as.data.table(smpa)
 vegpa_dt <- na.omit(vegpa_dt)
-setnames(vegpa_dt, "layer", "PA")
+setnames(vegpa_dt, "ZA.pas", "PA")
 protected <- vegpa_dt[, lapply(.SD, sum(x * 100)), by = PA, .SDcols = vnms]
@@
 # stack PAS with intact area
-pasconvs <- mask(stack(pas, remfrac), mgrid)
+pasconvs <- raster::mask(stack(pas, remfrac), mgrid)
 pa_dt <- as.data.table(pasconvs)
 pa_dt <- na.omit(pa_dt)
-setnames(pa_dt, c("layer.1", "layer.2"), c("pas", "nveg"))
+setnames(pa_dt, c("ZA.pas", "ZA.remfrac"), c("pas", "nveg"))
 pa_dt[, (1 - sum(nveg) / .N) * 100 , by = "pas"]
@@
 
+# save vegetation stats
+vegstats <- rbind(cbind("PA" = NA, wgts), protected)
+save(vegstats, file = fp(p_bd, "vegstats.rda"))
+
 # plots to examine them
-labs <- gsub("veg", "", names(wgtsn))
+labs <- gsub("v", "", names(wgtsn))
 # par(mfrow = c(2, 1), mar = c(2, 2, 2, 2))
@@
 # load(full_path(p_bd, "ZA-bdprops.rda"))
-
+# bdprops
+# bdprops[var == "area", rowSums(.SD), .SDcols = 2:ncol(bdprops)]
 ```

### 2016-02-28 Rmd/empirical-yields.Rmd
Supplementary methods draft complete. Required changes to biodiversity-layer.Rmd. Fixed bug in Figure 4 for main text--wrong map selected.

@@
 # load(fp(p_yield, "empirical/gams1.rda"))
-summary(gam13)
-summary(sgam14)
-
 ```
@@
 plot(prov, add = TRUE)
+
+summary(gam20)
+summary(sgam7)
 

### 2016-02-28 Rmd/figures.Rmd
Supplementary methods draft complete. Required changes to biodiversity-layer.Rmd. Fixed bug in Figure 4 for main text--wrong map selected.

@@
 p_other <- fp(p_root, "external/input_devel/other/")
+p_bd <- fp(p_root, "external/input_devel/biodiversity")
 # p_roads <- fp(p_root, "external/input_devel/roads")
@@
 imp_pct[, {
-  png(fp(p_figures, "weight_impact3.png"), height = 4, width = 4, 
+  png(fp(p_figures, "weight_impact4.png"), height = 4, width = 4, 
       units = "in", res = 300)
@@
 # use minmax stats from optimal solution set
-# pare <- ybdc$both$optitab
+pare <- ybdc$both$optitab
 optirange <- pare[, lapply(.SD, range), .SDcols = 1:4]
@@
 eqind <- which(apply(parms, 1, function(x) all(x == 0.25))) # iter of eq wt run
+parms[1031, ]
 # pare[ind == eqind]
@@
 # plot
-png(fp(p_figures, "compromise_5pct.png"), height = 3.5, 
+png(fp(p_figures, "compromise_5pct2.png"), height = 3.5, 
     width = 7, units = "in", res = 300)
@@
 # plot(zambia)
-imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
+imp_plot2(toresr$eqwt, impacts = i1, fpath = p_figures, fnm = "bob",
           crops = "", ftitle = "Compromise equally", col = "purple", 
@@
 
+# vegetation types
+load(fp(p_bd, "vegstats.rda"))
+vegstats <- round(vegstats * 100, 2)
+nms <- c("Afromontane rain forest", "Zambezian dry evergreen forest", 
+         "Zambezian dry deciduous forest", "Palm wooded grassland",
+         "Riverine wooded vegetation", "Zambezian Kalahari woodland",
+         "Drier miombo woodland", "Miombo woodland on hills/outcrops",
+         "Wetter mimbo woodland", "N. Zambezian undifferentiated woodland",
+         "Catena of N. Zambezian undifferentiated woodland/edaphic grassland", 
+         "Mopane woodland/scrub woodland", "Zambezian chipya woodland",
+         "Itigi thicket", "Swamp forest",
+         "Edaphic grassland on drainage impeded/seasonally flooded soils", 
+         "Bush groups, typically around termitaria",
+         "Edaphic grassland w/Zambezian Kalahari woodland",
+         "Edaphic grassland w/freshwater swamp", 
+         "Edaphic woodland on drainage impeded/seasonally flooded soils")
+print(xtable::xtable(cbind(c("", nms), unname(t(vegstats)[, c(1:2, 4, 3, 5)])), 
+                           digits = c(2, 2, 2, 1, 1, 1, 1), ))
+
 ```

### 2016-02-28 Rmd/yield-predictors.rmd
Supplementary methods draft complete. Required changes to biodiversity-layer.Rmd. Fixed bug in Figure 4 for main text--wrong map selected.

@@
 names(ylds) <- crops
+# ylds$maize[, mean(YLDS)]
+# ylds$soy[, mean(YLDS)]
+# ylds$maize[, unique(WTH)]
+# ylds$maize[, unique(SOL)]
 

### 2016-02-28 paper/supplementary_methods.tex
Supplementary methods draft complete. Required changes to biodiversity-layer.Rmd. Fixed bug in Figure 4 for main text--wrong map selected.

@@
 \documentclass[a4paper]{article}
-\usepackage[affil-it]{authblk}  
+%\usepackage[affil-it]{authblk}  
 %\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}  % apostrophe's in cites
@@
 \usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
-\usepackage{caption}
+\usepackage[skip=2pt,font=normalsize]{caption}
 \captionsetup[figure]{slc=off, font=footnotesize}
@@
 \usepackage[numbers,sort,compress]{natbib}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+%\usepackage{listings}
+%\usepackage{endfloat}
+\usepackage{float}
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 The agroEcoTradeoff model was designed to identify optimal land use configurations that allow a certain level of agricultural production to be achieved while minimizing associated ecological and economic costs.
-The model is based on that developed by Koh and Ghazoul \citep{koh_spatially_2010}, which was designed to select areas with promising potential for oil palm production but whose conversion would result in less loss of carbon, avian diversity, and land with high rice production potential.
-In this case, each unit of land was ranked based on its relative value within each measure, which was potential productivity for oil palm and the inverse of its potential rice productivity, carbon density, or bird diversity.
+The model follows the rationale of Koh and Ghazoul \citep{koh_spatially_2010}, which was designed to select areas with promising potential for oil palm production but whose conversion would result in less loss of carbon, avian diversity, and land with high rice production potential.
+In this case, each unit of land was ranked based on its relative value within each measure.
 Normalizing these values converted each into an oil palm conversion probability \emph{c} that was conditioned on the individual land value, such that the highest ranking areas were converted until a specified oil palm production target was reached.
@@
 
-We retained the basic rationale underlying this model in developing our own, wherein tradeoffs between yield maximizing and ecological cost minimizing objectives can be made in achieving a crop production target, with several amendments.
-First, the model has the capacity to simulate more than one crop.
-Second, the ranking of land units is based on the cost/benefit ratio of converting it to meet the target, thus areathat maximize benefits while minimizing costs along a particular cost dimension.
+Our model is formulated as multi-objective optimization problem that seeks to find acceptable tradeoffs between four potentially competing land use objectives--the desire to select the most productive lands for growing crops, versus the desire to minimize carbon and biodiversity loss, as well as production costs (represented in this study by a proxy variable, travel time).
+The model starts with the assumption that some level of crop production must be reached, for multiple crops.
+To achieve optimal tradeoffs, each objective is treated as an efficiency, wherein the lowest cost solution for meeting each production target is sought, using the following formulation:
 
-We adapt this approach by recasting the model in terms of efficiency, where land units are ranked in terms of its potential productivity relative to the cost that will be incurred from its conversion.
+\begin{equation}
+B_{ijk} = \frac{J_{jk}}{P_{ij}} \quad i \in m, \  j = 1, \ldots,n, \ k = 1, \ldots,q 
+\end{equation}
 
-%\begin{figure}[!ht]
-%       \begin{center}
-%      \includegraphics[scale=1]{figures/figS1.pdf} 
-%        \end{center}
-%     %\vspace{-1cm}
-%      \caption{The cumulative number of workers seeking qualification during the South Africa cropland mapping trial period.
-Successful qualification seekers are indicated by blue circles, unsuccessful by open circles.}
-%      \label{fig:default}
-%\end{figure}
+Where J$_{k}$ is one of the $q$ costs (e.g.
+carbon loss) to be minimized, within a landscape composed of $n$ grid cells containing values for each cost and potential productivity values $P_i$ for each of $m$ crops.
+We then normalize the values $B_{ijk}$ (range 0-1) so that  objectives measured in different units can be directly compared:
+
+\begin{equation}
+B_{ijk}^{prob} = 1 - \frac{B_{ijk} - min(B_k)}{range(B_k)}
+\end{equation}
+
+The normalized values are subtracted from 1 so that areas are ranking in descending order of their efficiency, $B_{ijk}^{rank}$, which is the order in which they will be converted to meet the target.
+
+We then use the weighted-sum approach to finding Pareto optimal solutions by scalarizing the set of objectives into a single objective.
+Each normalized objective is multiplied by a weight, $\lambda_k$, between 0 and 1, with $\sum_k \lambda_k = 1$, giving each cell a conversion probability to each crop:
+
+\begin{equation}
+C_{ij} = \sum_k \lambda_k B_{ijk}^{rank}
+\end{equation}
+
+Using the conversion probability metric, the problem is stated formally as a discrete optimization problem:
+
+\begin{align}
+maximize \sum_{i \in m} \sum_{j=1}^{n} C_{ij}x_{ij} \\
+s.t. \ \sum_{j=1}^n P_{ij}x_{ij} \geq t_i \quad \forall i \in m \tag{4.1}\label{eq:4.1} \\
+\sum_{i \in m} x_{ij} \leq 1 \quad j = 1, \ldots, n \tag{4.2}\label{eq:4.2} \\
+%x_{ij} \in \{ 0, 1\} \tag{4.3}\label{eq:4.3}
+\end{align}
+
+Where the decision variables are the $x_{ij}$'s.
+$x_{ij} = 1$ if cell $j$ is converted to crop $i$ and $x_{ij} = 0$ otherwise.
+The objective function seeks to maximize the sum of the conversion probabilities of allocations.
+Constraint \ref{eq:4.1} requires that the sum of the potential production of the cells converted for each crop must meet that crop's target production $t_i$.
+Constraint \ref{eq:4.2} restricts each cell to only one crop or no crop at all.
+
+\subsection*{Optimizing for multiple crops}
+Because the model is used for more than one crop, we also had to solve the problem of allocating land between crops.
+To do this, we developed a greedy approximation algorithm by adapting algorithms for solving the knapsack problem  \citep{dantzig_discrete-variable_1957} and the generalized assignment problem \citep{cohen_efficient_2006}, which allocates each crop to its highest value cells.
+
+The algorithm begins by initializing all $x_{ij}$ to 0. While all targets have not been met:
+
+\begin{enumerate}
+\item Eliminate from current consideration all crops whose targets have been met and all cells already allocated.
+\item Initialize all elements to 0 of a vector $r$ of length equal to the number of remaining cells, $n'$ ($n' = n$ in the first iteration).
+\item For each remaining crop $i$:
+\begin{enumerate}
+\item Set $C'_i = C_i - r$. $C'_i$ is a vector of length $n'$.
+\item Sort $C'_i$. Choose the cell $j$ with the largest $C'_{ij}$ for conversion to crop $i$ until the target $t_i$ is reached or exceeded.
+\item Set $\hat{C_i}$ equal to the largest $C'_{ij}$ from the cells not chosen for conversion.
+\item For each cell selected for conversion, set $r_j = r_j + (C'_{ij} - \hat{C_i})$.
+\end{enumerate}
+\end{enumerate}
+
+The algorithm considers crops one at a time and selects the cells with the highest conversion probability until the target is reached.
+Since two crops may not claim the same cell, the algorithm enables cells to be reallocated to crops in later iterations for which they are found to be more valuable towards maximizing the objective function.
+Consider each crop as a knapsack that needs to be filled to meet a target capacity.
+The vector $r$ represents the additional conversion probability that each cell brings to its current knapsack than the next available cell, $\hat{C_i}$, would bring.
+A cell $x_{ij}$ is reallocated to a different crop only if it is still of high enough rank among all remaining cells in terms of conversion probability after reassessing the conversion probability as $C'_{ij} - r_{j}$.
+(Note that $r$ does not take into account that when more than one cell is reallocated, multiple cells are often required to replace them).
+Due to this reallocation, some targets may not be satisfied after completing one pass through all crops.
+Subsequent passes are completed after removing from consideration the crops whose targets have been met and the cells already allocated.
+All targets are guaranteed to be satisfied after $m$ passes because the last crop considered is always satisfied (none of its selections are reallocated).
+By running the algorithm with the $C_{ij}$'s calculated with varying $\lambda_k$'s, we can find approximately optimal, non-dominated scenarios and plot the Pareto front.
+
+\subsection*{Model Assumptions and Limitations}
+Our model is limited to finding solutions in a convex solution space, and lacks the optimality guarantees possessed by more sophisticated algorithms.
+More sophisticated approaches, such as genetic approaches \citep[e.g.][]{stewart_genetic_2004,cao_spatial_2011}, which are adept at finding diverse sets of solutions, including those in non-convex spaces, their computational complexity causes model runs to be very time consuming \citep{geertman_participatory_2002}.
+Algorithms continue to grow more efficient using more sophisticated algorithms, but even these new routines compromise the number of outputs they are able to generate within a given time, which is problematic especially as the size of the inputs grows.
+The advantage of our approach is that it enables quick determination of land use changes, by applying a hybrid set of weights to multiple objectives in order to search for qualitatively appealing land use change scenarios.
+
+With respect to the greedy approximation algorithm, Dantzig \citep{dantzig_discrete-variable_1957} acknowledges that it is mathematically imperfect, as it is often optimal to eschew more efficient items in favor of selecting a combination of items that fill the ``knapsack'' closer to the target.
+However, Dantzig \citep{dantzig_discrete-variable_1957} also remarks that, in practice, if the target is not strict, the greedy algorithm always leads to a reasonable solution.
+Moreover, the difference between the greedy approximation and the optimal solution is particularly trivial if the weight of each individual item is small relative to the total weight limit.
+In the application for which this variant was designed, the potential production of each pixel tends to be orders of magnitude smaller than the target production for a crop.
+Also, the target is not so strict, as variability is implicit in predicting agricultural productivity.
+This flexibility leads the greedy algorithm to produce reasonable scenarios.
+
+\section*{Further Details on Development of Model Inputs}
+\section*{Yield Potential}
+In order to estimate yields using DSSAT, we took four stock cultivars for maize representing generic short and medium growing season length open-pollinated and hybrid cultivars, using coefficients developed by Elliott et al \citep{elliott_parallel_2014}.
+We adapted these using the approach of Grassini et al \citep{grassini_how_2015}, where the phenological coefficients were tuned so that the average maturity date simulated by DSSAT matched typical maturity times for short and medium length cultivars in Zambia.
+For soybean, we used three cultivars (for zones 7-9) used in a previous study to simulate soybean potential across the African savanna regions \citep{searchinger_high_2015}.
+
+Both maize and soybean were set to automatically plant when soil water reached 90\% of capacity, with the first possible plant date for maize set to November 11 and soybean to January 1 (approximately each crop's average first dates for planting in Zambia).
+Maize and soybean were respectively given 100 and 50 kg ha$^{-1}$ of N fertilizer applied at planting, with maize planted with a density of 4.5 plants m$^{-2}$ and a row spacing of 90 cm, with soybean a density of 40 plants m$^{-2}$ and row spacing of 65 cm.
+
+We ran the model at 40 weather stations for each of the dominant soils associated in the 10X10 km area \cite{romero_reanalysis_2012} around those stations, for each growing season between 1979 and 2010, resulting in a total of 5,859 climate X soil X weather year combinations.
+This was repeated for each cultivar, after which we averaged the results at each station and soil types and for each year across cultivars.
+The average maize yield across all permutations was 8518 t ha$^{-1}$ and for soybean it was 3074 t ha$^{-1}$.
+
+We fit generalized additive models \citep{hastie_generalized_1990,wood_mgcv:_2001} to the cultivar averaged yields, which we further reduced to a mean, 10th percentile, and 90th percentile yield value for each weather station-soil combination, using the predictors described in the main text.
+We adopted the GAM approach because it makes no assumptions about the distribution of the data, and can fit complex non-linear relationships.
+We took advantage of this feature to control for spatial correlations in the data, using a two-dimensional spline on the x and y coordinates of the station locations \citep[e.g.][]{estes_projected_2013,estes_comparing_2013}.
+For all other predictors we used parametric, and primarily polynomial terms, accounting for the well-known polynomial relationship between most of the predictors and crop yields \citep{lobell_nonlinear_2011,lobell_use_2010}.
+We selected the best-fitting, least-correlated subset of predictors, further screening the selected model to make sure it made biophysical sense (e.g.
+we rejected models where the relationship between soil organic carbon, pH, or temperature showed a concave, as opposed, to convex shape).
+The following is the summary of the GAM fit for maize:
+
+\begin{verbatim}
+Family: gaussian 
+Link function: log
+
+Formula:
+YLDS ~ poly(DJFMP, 2) + poly(GDD, 2) + poly(SSRMI, 2) + poly(SLHW, 
+    2) + poly(SLCL, 2) + s(x, y)
+
+Parametric coefficients:
+                Estimate Std. Error t value Pr(>|t|)    
+(Intercept)      8.98378    0.01116 804.700  < 2e-16 ***
+poly(DJFMP, 2)1  0.55483    0.39486   1.405  0.16054    
+poly(DJFMP, 2)2 -3.06686    0.37129  -8.260 1.08e-15 ***
+poly(GDD, 2)1   -0.71357    0.32256  -2.212  0.02736 *  
+poly(GDD, 2)2   -0.31694    0.28074  -1.129  0.25941    
+poly(SSRMI, 2)1  1.10670    0.53159   2.082  0.03781 *  
+poly(SSRMI, 2)2 -1.05899    0.33248  -3.185  0.00153 ** 
+poly(SLHW, 2)1   0.22406    0.39604   0.566  0.57179    
+poly(SLHW, 2)2  -2.45059    0.39453  -6.211 1.03e-09 ***
+poly(SLCL, 2)1  -0.52933    0.41173  -1.286  0.19911    
+poly(SLCL, 2)2  -1.85528    0.37560  -4.939 1.04e-06 ***
+---
+Signif. codes:  0 Ô***Õ 0.001 Ô**Õ 0.01 Ô*Õ 0.05 Ô.Õ 0.1 Ô Õ 1
+
+Approximate significance of smooth terms:
+         edf Ref.df     F p-value  
+s(x,y) 2.013  2.025 3.877  0.0211 *
+---
+Signif. codes:  0 Ô***Õ 0.001 Ô**Õ 0.01 Ô*Õ 0.05 Ô.Õ 0.1 Ô Õ 1
+
+R-sq.(adj) =  0.388   Deviance explained = 40.1%
+-REML = 5123.9  Scale est. = 4.0555e+06  n = 567
+\end{verbatim}
+
+And for soybean:
+
+\begin{verbatim}
+Family: gaussian 
+Link function: log
+
+Formula:
+YLDS ~ poly(JFMP, 2) + poly(GDD, 2) + poly(SLHW, 2) + poly(SLCL, 
+    2) + poly(SSRMIII, 2) + s(x, y)
+
+Parametric coefficients:
+                  Estimate Std. Error t value Pr(>|t|)    
+(Intercept)        7.86979    0.02608 301.806  < 2e-16 ***
+poly(JFMP, 2)1     5.29558    0.98371   5.383 1.09e-07 ***
+poly(JFMP, 2)2    -6.33762    0.91534  -6.924 1.24e-11 ***
+poly(GDD, 2)1      1.03137    0.64155   1.608 0.108496    
+poly(GDD, 2)2     -0.06524    0.52657  -0.124 0.901440    
+poly(SLHW, 2)1    -2.36290    0.65132  -3.628 0.000313 ***
+poly(SLHW, 2)2    -5.99456    0.70424  -8.512  < 2e-16 ***
+poly(SLCL, 2)1    -0.64178    0.64034  -1.002 0.316672    
+poly(SLCL, 2)2    -1.61319    0.61927  -2.605 0.009439 ** 
+poly(SSRMIII, 2)1 -7.03157    1.06314  -6.614 8.95e-11 ***
+poly(SSRMIII, 2)2 -4.06347    0.74592  -5.448 7.74e-08 ***
+---
+Signif. codes:  0 Ô***Õ 0.001 Ô**Õ 0.01 Ô*Õ 0.05 Ô.Õ 0.1 Ô Õ 1
+
+Approximate significance of smooth terms:
+         edf Ref.df     F p-value   
+s(x,y) 10.92  14.85 2.564 0.00102 **
+---
+Signif. codes:  0 Ô***Õ 0.001 Ô**Õ 0.01 Ô*Õ 0.05 Ô.Õ 0.1 Ô Õ 1
+
+R-sq.(adj) =  0.545   Deviance explained = 56.1%
+-REML = 4832.3  Scale est. = 1.445e+06  n = 567
+\end{verbatim}
+
+The maize model explained 40\% of the deviance in the DSSAT-simulated yields, and the soybean model explained 56\% of deviance.
+The maize model used predictors for December-March cumulative precipitation (DJFMP), growing degree days (GDD), mean November-March solar radiation (SSRMI), soil pH (SLHW), and soil clay content, with polynomial terms fit to each.
+The soybean used the same predictors, except the precipitation was the sum of the January-March period (JFMP) and solar radiation was averaged over the same time period (SSRMIII).
+With the exception of GDD in soybean (which was near significant at the p<0.1 level and was retained for consistency with the maize model), all predictors were significant on one or both of the linear and squared term of the predictor, while the geographic terms was significant in both models.
+
+Using the gridded predictor variables, we mapped predicted yield values for all of Zambia.
+We then adjusted these data to match the FAO-projected yields as described in the main text.
+We used two sources to calculate the likely future yield improvement trend for Zambia.
+We used the Food and Agricutural Organization's (FAO) historical and projected yield dataset \citep{alexandratos_world_2012} for sub-Saharan Africa to calculate a projected annual rate of change between 2007 and 2050, and we also examined 2000-2014 yield trends for Zambia, collected from FAOStat \citep{fao_fao_2012}, from which we calculated an annual rate of change.
+We then took a weighted average of these two factors, giving 25\% weight to the Zambia projections and 75\% to the Bruinsma-derived value to calculate a maize growth rate, to given Zambia some credit for the larger yield gains it has recently shown relative to most of the rest of the region.
+For soybean we used the Bruinsma rate.
+
+We then applied these rates to calculate a projected 2050 yield for each crop in Zambia as follows:
+
+\begin{equation}
+y_{2050} = (1 + \beta(2050-2014))y_{current}
+\end{equation}
+
+Where $\beta$ is the yield gain rate, and $y_{current}$ is the 2009-2014 mean yield for each crop (2.3 t ha$^{-1}$ for maize and 2 t ha$^{-1}$ for soybean).
+Applying this formula resulted in projected yields of 4.43 and 3.4 tons t ha$^{-1}$ for maize and soybean, respectively.
+
+The then adjusted the yields in the GAM-derived maps so that their means matched these projected means.
+However, since the yield rates and average values on which they were based were obtained from existing cropland, we extracted the gridded yield values that corresponded with existing cropland (determined using the landcover dataset, see main text).
+We calculated the mean GAM-predicted yield on existing cropland, which we divided into the FAO-projected yield to determine a correction factor, which we then applied to the entirety of each map.
+The resulting maps are shown in Figure 1.
+
+\begin{figure}[ht]
+\captionsetup{width=0.95\linewidth, font=small}
+    \begin{center}
+       \makebox[\textwidth][c]{\includegraphics[width=1\textwidth]{figures/yields.png} }
+      \caption{Predicted yield maps for maize (left) and soybean (right), shown in tons ha$^{-1}$, produced by fitting a generalized additive model to DSSAT-predicted yields at 40 points throughout Zambia where weather stations are located.
+The GAM-predicted yields where adjusted so that their mean values matched 2050 projected mean yields for Zambia, derived from FAO data.
+Yields are shown for locations outside of protected areas (GMAs = game management areas; NatParks = national parks).
+}
+      \label{fig:default}
+\end{center}
+\end{figure}
+
+\subsection*{Landcover data}
+The choice of landcover product can have a critical impact on the results of any spatial analysis.
+This is particularly true with respect to trying to locate the current extent of existing cropland, where existing landcover datasets are particularly discrepant and in wide disagreement \cite{fritz_comparison_2010} because the average cropland extent is small and, from a nadir perspective, are often difficult to distinguish from the surrounding savanna vegetation \cite{estes_platform_2016,debats_generalized_2015}.
+The most accurate landcover datasets tend to be country-specific Landsat-based maps \cite[][and per a forthcoming South Africa-focused landcover intercomparison]{fritz_comparison_2010}, therefore we selected the 30 m resolution type II landcover map for Zambia for the year 2010, developed under the NASA SERVIR program\footnote{downloaded from http://apps.rcmrd.org/landcoverviewer/}.
+These products have undergone extensive validation and provide the most current and best available landcover maps for Zambia and several other eastern and southern African countries, and their consistency across the region is another advantage in that it facilitates the transferability of our model.
+
+We used this landcover map as the basis for determining how much land in each 1 km$^2$ grid cell was unconverted to cropland or urban areas.
+This information formed the basis for determining what land was potentially available for new cropland (although it does not factor in ownership or other factors that cannot be seen from space, which determine true availability), and also the degree to which habitats remained intact and un-fragmented.
+To calculate both measures, we first removed cropland and urban cover types from the 30 m resolution landcover map, and then aggregated the remainder to 1 km$^2$, producing a fractional map expressing how much untransformed land was available in each grid cell.
+To calculate the intactness measure, we passed a simple 11X11 focal mean filter over this map, with the result describing the proportion of intact habitat remaining in a broader 121 km$^2$ neighborhood around each grid cell.
+We selected this neighborhood size because 1) it is roughly equivalent to the size of a large provincial nature in South Africa \cite{estes_using_2014}, which has been more heavily transformed to agriculture, and 2) in a situation where a country is facing large-scale agricultural transformation, it is useful to highlight areas that have both a viable amount of remaining habitat, yet are not so big that they would be heavily conflicted with agricultural interests.
+
+\subsection*{Biodiversity}
+The rarity/threat measure incorporated into our biodiversity value criterion is a simplified version of the measure developed by van Breugel et al \citep{van_breugel_environmental_2015}.
+Our metric differs in that we define the rarity of each vegetation type with respect to how much of its remains untransformed to cropland or urban areas within Zambia, whereas van Breugel et al define this regionally.
+We adopted this national-level criterion because 1) the other land use criteria in our analysis are scaled to this level, and 2) conservation decisions about what to protect or not to protect are typically made based on representativeness at national to sub-national scales \citep[e.g.][]{cowling_conservation_2003}.
+
+We also did not calculate the same threat index for non-protected parts of these vegetation types, opting instead to use the measure of intactness described in the previous section and in the main text.
+We wanted to give equal weight to intactness because the overwhelmingly dominant habitat types in Zambia are miombo woodlands (Table 1), which to date have remained relatively intact.
+Applying just the rarity/threat index, would penalize these habitats, yet their relative intactness, together with high mammal species richness, is one of the factors that make them of conservation interest \cite{olson_terrestrial_2001,searchinger_high_2015,brooks_global_2006}.
+Furthermore, a unique feature of miombo woodlands is that, due to the digestibility of forage, megaherbivores (buffalo, elephant) are over-represented in the assemblage of large mammalian herbivores, and correspondingly require large range sizes to maintain viable populations \cite{mcnaughton_ecology_1986}.
+
+\begin{table}[ht]
+\centering
+\captionsetup{width=0.95\linewidth, font=small}
+\caption{The vegetation types of Zambia \citep[per][]{van_breugel_potential_2011}, as a percent of Zambia's total area covered by each of the type's remaining untransformed area, and what percetange of that remaining untransformed area is inside and outside of protected areas.
+Protected areas are distinguished by class: NPs = national parks; GMAs = game management areas; FRs = forest reserves.}
+\begin{tabular}{R{7cm}rrrrr}
+  \hline
+  & & \multicolumn{4}{c}{Percent in Protected Areas}\\
+Vegetation type & \% Total & Outside & in NPs & in GMAs & in FRs \\ 
+  \hline
+Afromontane rain forest & 0.01 & 46.1 & 28.7 & 0 & 25.2 \\ 
+Zambezian dry evergreen forest & 2.36 & 67.2 & 6.1 & 17.4 & 9.3 \\ 
+Zambezian dry deciduous forest & 0.97 & 41.7 & 11.7 & 28.9 & 17.7 \\ 
+Palm wooded grassland & 0.02 & 100 & 0 & 0 & 0 \\ 
+Riverine wooded vegetation & 0.10 & 22.7 & 33.9 & 42.5 & 1 \\ 
+Zambezian Kalahari woodland & 4.45 & 36 & 6.1 & 51.1 & 6.9 \\ 
+Drier miombo woodland & 15.10 & 65.5 & 8.8 & 19.8 & 5.9 \\ 
+Miombo woodland on hills/outcrops & 0.15 & 69.2 & 6.7 & 13 & 11.2 \\ 
+ Wetter mimbo woodland & 31.73 & 68.9 & 6.1 & 13.2 & 11.8 \\ 
+ N. Zambezian undifferentiated woodland & 3.8 & 56.7 & 9.2 & 29.2 & 4.9 \\ 
+ Catena of N. Zambezian undifferentiated woodland/edaphic grassland & $<$0.01 & 100 & 0 & 0 & 0 \\ 
+ Mopane woodland/scrub woodland & 5.90 & 22.9 & 29.8 & 46.3 & 1 \\ 
+ Zambezian chipya woodland & 2.08 & 73.9 & 2.6 & 11.9 & 11.6 \\ 
+ Itigi thicket & 0.16 & 20.9 & 41 & 34.3 & 3.8 \\ 
+ Swamp forest & 0.02 & 89.6 & 0 & 0 & 10.4 \\ 
+ Edaphic grassland on drainage impeded/seasonally flooded soils & $<$0.01 & 69.3 & 4.9 & 21.3 & 4.6 \\ 
+ Bush groups, typically around termitaria & 2.34 & 57 & 17.8 & 23.4 & 1.8 \\ 
+ Edaphic grassland w/Zambezian Kalahari woodland & 1.14 & 24.7 & 5 & 69.7 & 0.6 \\ 
+ Edaphic grassland w/freshwater swamp & 16.68 & 56.3 & 8.2 & 31.3 & 4.3 \\ 
+ Edaphic woodland on drainage impeded/seasonally flooded soils & 0.01 & 92.7 & 1.8 & 2.6 & 2.9 \\ \hline
+\end{tabular}
+\end{table}
+
+\subsection*{Estimating Travel Time}
+
+To calculate the time to travel to the nearest major town, we considered not just Zambia but also portions of the neighboring countries, under the assumption that for certain areas within Zambia near its borders, the closest market center might be in a neighboring country.
+We therefore calculated travel time within a region defined by longitudes 19.5 to 34.5 degrees, and latitudes -19.5 to -8 degrees.
+
+We obtained a vector dataset of roads from the Zambian government, and downloaded an additional dataset available for broader region from OpenStreetMap.
+We merged these two sets into a single dataset in order to capture the road network in and around Zambia, and classified roads into trunk roads, primary roads, secondary roads, and tertiary roads.
+
+We obtained vector datasets for lakes and major towns in Zambia from the government.
+For towns outside Zambia, we used the Gridded Population of the World Version 3 \citep{ciesin_gridded_2005} town vectors.
+We used the USGS Hydrosheds database to indicate the position and size of rivers, based on their contributing area.
+
+The roads, rivers, and lakes were converted to a raster of 1 km resolution (the same as our analysis), together with Zambia's borders.
+Travel times were then assigned to each of the features, based on how long it would take to cross a 1 km pixel:
+
+\begin{enumerate}
+\item Roads
+\subitem Trunk roads: 1 minute (i.e. 60 km hr$^{-1}$)
+\subitem Primary roads: 1.3333 minutes
+\subitem Secondary roads: 2 minutes
+\subitem Tertiary roads: 4 minutes
+\item Lakes: 6 minutes to traverse (assuming ferry or boat)
+\item Rivers: 
+\subitem Accumulating area $>$600: 90 minutes where no bridge exists
+\subitem Accumulating area $>$400: 60 minutes
+\subitem Accumulating area $>$200: 30 minutes
+\item Country border: 60 minutes
+\item No road: 20 minutes
+\end{enumerate}
+
+The travel time rasters where then merged into a single friction surface.
+Except at borders, where two different levels of cost intersected, the fastest time was assigned, i.e.
+where a road and river intersected, it was assumed a bridge was present and therefore the road's travel time was assigned.
+We then used the costdistance function of ArcGIS to calculate travel time, using the town locations as the origins, which resulting in a map representing travel time in hours to the nearest town (Figure 2).
+
+\begin{figure}[ht]
+\captionsetup{width=0.95\linewidth, font=small}
+    \begin{center}
+       \makebox[\textwidth][c]{\includegraphics[width=1\textwidth]{figures/cost.png} }
+      \caption{Cost distance surface for Zambia, showing travel time in hours to the nearest major town.
+Travel times are shown for locations outside of protected areas (GMAs = game management areas; NatParks = national parks).
+}
+      \label{fig:default}
+\end{center}
+\end{figure}
+
 %\section*{\large References}

### 2016-03-01 Rmd/figures.Rmd
Submitted version 29/2/2016

@@
 # plot(mzsoyr, col = c("grey", "goldenrod", "green4", "orange"))
-imp_plot2(mzsoyr, i1, p_figures, "ideal_yield2", "", crops, 1.2, TRUE, 
+imp_plot2(mzsoyr, i1, p_figures, "ideal_yield3", "", crops, 1.2, TRUE, 
          c("red", "blue1", "gold"))
@@
 ```{r, eval = FALSE}
-load(fp(p_out, "mainresults2502.rda"))
+load(fp(p_out, "mainresults2802.rda"))
 
@@
 conv_stats <- selected_impacts(pure, fp(p_out, ybdc$both$bcode)) 
+# selected_impacts(pure, fp(p_out, ybdc$both$bcode)) / 100000
+conv_stats[1, ] / conv_stats[3, ]
+
 # impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) conv_stats[[x]])
-impsr <- lapply(c(2, 3, 8, 5), function(x) conv_stats[[x]])
+impsr <- lapply(c(2, 3, 5, 4), function(x) conv_stats[[x]])
+# impacts[ind %in% pure]
 
@@
 
-# plot(1:10, ylab = ylabs[2])
-# ylabs <- c("Jaccard Index", 
-ylabs <- c(expression(paste("Converted area (", 10^5, " ", km^2, ")")), 
+ylabs <- c(expression(paste("Converted area (", km^2, " X ", 10^-5, ")")), 
            "Carbon loss (Mt)", #"Priority", 
-           "Biodiversity Index", "Travel time (hours)")#,
+           expression(paste("Biodiversity Index X ", 10^-5)),
+           expression(paste("Truck hours X ", 10^-5)))#,
            # "Intactness", expression(paste("FR converted (ha"^2, ")")))
 # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
-yscale <- c(10^5, 100000, 1, 1)#, 1, 1, 1)
-# ftits <- list("Y 100%", "C 100%", "Biod 100%", "Cost 100%")
+yscale <- rep(100000, 4)#, 10000, 10000)#, 1, 1, 1)
 ftits <- list("Maximize Yields", "Minimize Carbon Loss", 
-              "Minimize Biodiversity Loss", "Minimize Travel Time")
+              "Minimize Biodiversity Loss", "Minimize Travel Cost")
 # leg <- c("Converted")
-
-yaxes <- list(seq(0, 16, 2), seq(0, 1200, 100), seq(0, 0.6, 0.1), 
-              seq(0, 7, 1))#,               
+yaxes <- list(seq(0, 16, 2), seq(0, 1200, 100), seq(0, 9, 1), 
+              seq(0, 35, 5))#,               
               # seq(0, 1200, 100), #seq(0, 0.12, 0.02)), 
@@
 # tiff(fp(p_figures, "single_priorities2.tif"), height = 3.5, width = 7, 
-png(fp(p_figures, "single_priorities2.png"), height = 3.5, width = 7, 
+png(fp(p_figures, "single_priorities3.png"), height = 3.5, width = 7, 
     units = "in", res = 300)
@@
 }  
-# impact plots
-# mgp <- c(1.6, 0.25, 0)
-# par(mar = c(4, 3, 0, 0))
-# with(pms2b, {
-#   plot(NND, JS, col = cols_tr[c(2, 3, 3, 4, 4, 4)], pch = 16, cex = 1.5, 
-#        axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.3))
-#   points(NND, JS, col = cols_tr[c(1, 1, 2, 1, 2, 3)], pch = 20, cex = 1)
-#   axis(1, at = seq(0, 230, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
-#        tcl = -0.2)
-#   axis(2, at = seq(0, 0.3, 0.05), cex.axis = 0.6, las = 2, mgp = mgp, 
-#        tcl = -0.2)
-#   mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
-#   mtext(text = "Jaccard Index", side = 2, line = 1.25, cex = 0.45)
-# })
+# overlap/NND impact plots (recover from commit before 28/2 if needed)
 mgp <- c(1, 0.25, 0)
@@
 cx <- 0.6
-for(i in 1:4) {
+for(i in 1:4) { # i <- 3
   v <- impsr[[i]] / yscale[i]
@@
     mtext("Average distance", side = 2, line = 1.72, cex = 0.4)
-    mtext("to neighbor (km)", side = 2, line = 1.2, cex = 0.4)
+    mtext("to neighbour (km)", side = 2, line = 1.2, cex = 0.4)
   })
@@
 ### Figures 2 & 3 (probably supplemental)
+Deprecated
 ```{r, eval = FALSE}
 # individual maps
-png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
-     units = "in", res = 300)
-par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
-ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
-legp <- c(TRUE, rep(FALSE, 3))
-for(i in 1:4) {
-  dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
-  imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
-            cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
-            stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
-}
-dev.off()
+# png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
+#      units = "in", res = 300)
+# par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
+# ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
+# legp <- c(TRUE, rep(FALSE, 3))
+# for(i in 1:4) {
+#   dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
+#   imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
+#             cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
+#             stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
+# }
+# dev.off()
 
@@
 cstats <- selected_impacts(pind, fp(p_out, ybdc$both$bcode))
-cstats <- cstats[, .(iter, conv_area, tot_C, mu_cost, int_prior)]
+cstats <- cstats[, .(iter, conv_area, tot_C, tot_cost, cons_tot)]
+# colSums(outtab[1:2, c("conv_area", "tot_C", "tot_cost", "cons_tot")])
+# pare[ind %in% pind]
 
@@
 
-((minmaxstats[2, c(2:3, 8, 5), with = FALSE]) - 
-  (imp_pct[4, c(4:5, 7, 6), with = FALSE])) /
-  imp_range[, c(2:3, 8, 5), with = FALSE]
+# percent of total possible savings by the 15\% weight to both C and BD
+((minmaxstats[2, .(conv_area, tot_C, cons_tot, tot_cost)]) - 
+  (imp_pct[4, .(conv_area, tot_C, cons_tot, tot_cost)])) /
+  imp_range[, .(conv_area, tot_C, cons_tot, tot_cost)]
 
@@
 imp_pct[, {
-  png(fp(p_figures, "weight_impact4.png"), height = 4, width = 4, 
+  png(fp(p_figures, "weight_impact5.png"), height = 4, width = 4, 
       units = "in", res = 300)
@@
   lines(xwgts, tot_Cp, col = cols_tr[2])
-  lines(xwgts, int_priorp, col = cols_tr[3])
-  lines(xwgts, mu_costp, col = cols_tr[4])
+  lines(xwgts, cons_totp, col = cols_tr[3])
+  lines(xwgts, tot_costp, col = cols_tr[4])
   axis(1, at = xwgts, labels = seq(0, 90, 10), cex = cx, cex.axis = cx, 
@@
   legend("bottomright", 
-         legend = c("Yield", "Carbon", "Biodiversity", "Travel time"), 
+         legend = c("Yield", "Carbon", "Biodiversity", "Travel cost"), 
          cex = cx, text.col = cols_tr, col = cols_tr, bty = "n", lty = 1, 
@@
 
+minmaxstats[1, ] / 10000 * 1.05 - minmaxstats[1, ] / 10000
+
 # indices of costs in other three variables
-cost2 <- seq(1, 5, 0.05)
+cost2 <- seq(1, 6, 0.05)
 y_comp <- compromise(pare, optirange, cvar = "Y", cost1 = 1.05, cost2 = cost2)
@@
 # impsr <- lapply(c(2, 5, 3, 7:9, 4), function(x) all_stats[[x]])
-impsr <- lapply(c(2, 3, 8, 5), function(x) all_stats[[x]])
+impsr <- lapply(c(2, 3, 5, 4), function(x) all_stats[[x]])
 
@@
 
-maxes <- sapply(c(2:3, 8, 5), function(j) minmaxstats[[j]][2])
-mins <- sapply(c(2:3, 8, 5), function(j) minmaxstats[[j]][1])
+maxes <- sapply(c(2:3, 5, 4), function(j) minmaxstats[[j]][2])
+mins <- sapply(c(2:3, 5, 4), function(j) minmaxstats[[j]][1])
 ymaxes <- maxes / yscale
@@
 ftits <- list("Compromise 5% on Yield", "Compromise 5% on Carbon", 
-              "Compromise 5% on Biodiversity", "Compromise 5% on Travel Time")
-yaxes <- list(seq(0, 16, 4), seq(0, 1200, 100), seq(0, 0.6, 0.1), 
-              seq(0, 7, 1))
+              "Compromise 5% on Biodiversity", "Compromise 5% on Travel Cost")
+yaxes <- list(seq(0, 16, 4), seq(0, 1200, 100), seq(0, 9, 1), 
+              seq(0, 35, 5))
 pcttxt <- c("best", "50%", "worst")
+impsr
 
 # plot
-png(fp(p_figures, "compromise_5pct2.png"), height = 3.5, 
+png(fp(p_figures, "compromise_5pct3.png"), height = 3.5, 
     width = 7, units = "in", res = 300)
@@
     mtext("Average distance", side = 2, line = 1.3, cex = 0.4)
-    mtext("to neighbor (km)", side = 2, line = 0.8, cex = 0.4)
+    mtext("to neighbour (km)", side = 2, line = 0.8, cex = 0.4)
   })
@@
 
-png(fp(p_figures, "compromise_25pct.png"), height = 2, width = 4, 
+png(fp(p_figures, "compromise_25pct2.png"), height = 2, width = 4, 
     units = "in", res = 300)
@@
 for(i in 1:4) {
-  1 - print(all_stats[5, c(2:3, 8, 5), with = FALSE] / 
-              all_stats[i, c(2:3, 8, 5), with = FALSE])
-}
-
+  1 - print(all_stats[5, c(2:3, 5, 4), with = FALSE] / 
+              all_stats[i, c(2:3, 5, 4), with = FALSE])
+}
+
+# equal compromise avoided cost
 for(i in 1:5) {
-  print(((minmaxstats[2, c(2:3, 8, 5), with = FALSE]) - 
-     (all_stats[i, c(2:3, 8, 5), with = FALSE])) /
-    imp_range[, c(2:3, 8, 5), with = FALSE])
+  print(((minmaxstats[2, c(2:3, 5, 4), with = FALSE]) - 
+     (all_stats[i, c(2:3, 5, 4), with = FALSE])) /
+    imp_range[, c(2:3, 5, 4), with = FALSE])
+}
+
+# equal cost difference relative to best case
+for(i in 1:5) {
+  print(((all_stats[i, c(2:3, 5, 4), with = FALSE]) - 
+     (minmaxstats[1, c(2:3, 5, 4), with = FALSE])) /
+    imp_range[, c(2:3, 5, 4), with = FALSE])
 }
@@
             marg = c(0, 0, 1, 0), inset = c(0, 0, 0, 0), titline = 0)
-  mtext(paste("Jaccard Index =", round(fb_sim[i], 2)), side = 3, line = -2, 
+  mtext(paste("% Overlap =", round(fb_sim[i], 2)), side = 3, line = -2, 
         cex = 0.4, adj = 0.2)
-  mtext(paste("NN Distance =", round(fb_nnd[i]), "km"), side = 3, line = -2.5, 
-        cex = 0.4, adj = 0.2)
+  mtext(paste("Mean dist. to neighbour =", round(fb_nnd[i]), "km"), side = 3,
+        line = -2.5, cex = 0.4, adj = 0.2)
 }
@@
 gaptab <- rbind(lowi, all_stats[5, names(all_stats)[-1], with = FALSE], highi)
-gaptab[1, ] / gaptab[2, ]
+gaptab[1, ] / gaptab[2, ] - 1
+gaptab[3, ] / gaptab[2, ] - 1
 gaptab[, tot_C / 100000]
@@
 cost <- dt_to_raster(cbind(basedt[, .(x, y)], il$cost), CRSobj)
+bd <- dt_to_raster(cbind(basedt[, .(x, y)], il$cons), CRSobj)
 closs <- carb[[1]] + carb[[2]] * 0.25
@@
   plot(zambia, col = "white", border = "transparent")
-  plot(ylds[[i]], legend = TRUE, add = TRUE)
-  plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
-       border = pacols[2])
-  plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+  plot(ylds[[i]], legend = TRUE, add = TRUE, axis.args = list(cex.axis = 0.7),
+       legend.args = list(text = expression("tons ha"^-1), cex = 0.8))
+  plot(pas[pas@data$type %in% c("gma", "npark"), ], add = TRUE, col = pacols[3],
        border = pacols[3])
-  legend("bottomright", legend = c("GMAs", "NatParks"),
-         pch = 15, col = c(pacols[2:3]), bty = "n",
-         pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))
-
+  if(i == 1) {
+    legend("bottomright", legend = c("Protected Areas"),
+           pch = 15, col = c(pacols[2]), bty = "n",
+           pt.cex = 2, cex = 0.8, inset = c(0, 0.1, 0, 0))
+  }  
 }
@@
      res = 300)
-par(mar = c(0, 0, 0, 4))
+par(mar = c(0, 0, 0, 5))
 plot(zambia, col = "white", border = "transparent")
-plot(closs, legend = TRUE, add = TRUE)
-plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+plot(closs, legend = TRUE, add = TRUE, axis.args = list(cex.axis = 0.7),
+       legend.args = list(text = expression("tons ha"^-1), cex = 0.8))
+plot(pas[pas@data$type %in% c("gma", "npark"), ], add = TRUE, col = pacols[2],
      border = pacols[2])
-plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
-     border = pacols[3])
-legend("bottomright", legend = c("GMAs", "NatParks"),
-       pch = 15, col = c(pacols[2:3]), bty = "n",
-       pt.cex = 2, cex = 1, inset = c(0.1, 0.1, 0, 0))
+legend("bottomright", legend = c("Protected Areas"),
+       pch = 15, col = c(pacols[2]), bty = "n",
+       pt.cex = 2, cex = 0.8, inset = c(0, 0.1, 0, 0))
 dev.off()
@@
      res = 300)
-par(mar = c(0, 0, 0, 4))
+par(mar = c(0, 0, 0, 5))
 plot(zambia, col = "white", border = "transparent")
-plot(cost, legend = TRUE, add = TRUE)
+plot(cost, legend = TRUE, add = TRUE, 
+     axis.args = list(cex.axis = 0.7),
+     legend.args = list(text = "hours", cex = 0.8))
+plot(pas[pas@data$type %in% c("gma", "npark"), ], add = TRUE, col = pacols[2],
+     border = pacols[2])
+legend("bottomright", legend = c("Protected Areas"),
+       pch = 15, col = c(pacols[2]), bty = "n",
+       pt.cex = 2, cex = 0.8, inset = c(0, 0.1, 0, 0))
+dev.off()
+
+png(fp(p_figures, "bd.png"), height = 3.5, width = 4.5, unit = "in", 
+     res = 300)
+par(mar = c(0, 0, 0, 5))
+plot(zambia, col = "white", border = "transparent")
+plot(bd, legend = TRUE, add = TRUE, axis.args = list(cex.axis = 0.7))
 plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
@@
 
+### Carbon stocks
+```{r, eval = FALSE}
+
+carbon <- brick(fp(p_root, "external/input_devel/carbon/ZA-carbon.tif"))
+carbont <- calc(carbon, sum)
+carbont <- raster::mask(carbont, mgrid, maskvalue = 0)
+freq(!is.na(carbont))  # 738442 km2
+quantile(carbont, prob = seq(0, 1, 0.1))  # 50% 130 t/ha, 80 t/ha
+
+```
+
 <a href="#top">Back to top</a>

### 2016-03-01 Rmd/main-analysis.Rmd
Submitted version 29/2/2016

@@
 Yv = "conv_area"; Cv = "tot_C"; #BDv = "int_prior"; COSTv = "mu_cost"
-BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
-wmu <- function(x, na.rm = na.rm) {
-  weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
-}
-Yst = sum; Cst = sum; BDst = wmu; COSTst = wmu
+# BDv = c(Yv, "int_prior"); COSTv = c(Yv, "mu_cost")
+# wmu <- function(x, na.rm = na.rm) {
+#   weighted.mean(x = x[, 2], w = x[, 1], na.rm = na.rm)
+# }
+Yst = sum; Cst = sum; # BDst = wmu; COSTst = wmu
 crops <- list(c("maize", "soy"))
@@
 # load(fp(p_out, "mainresults1702.rda"))
+
+# 28/2 Rebuilding impacts with new travel time and bd cost metrics (totals)
+load(fp(p_out, "mainresults2502.rda"))
+crops <- list(c("maize", "soy"))
+fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), inds = 1:nrow(ybdc$both$params),                      fullnames = TRUE)
+il <- input_handler("agroEcoTradeoff/external/data/ZA", crops = crops[[1]])
+registerDoMC(7)
+imptab <- foreach(i = 1:length(fnames)) %dopar% {  # x <- fnames[1]
+  conv <- fread(fnames[i])
+  impact(conv, il)
+}
+outtab <- do.call(rbind, lapply(1:length(imptab), function(x) {
+  cbind.data.frame("iter" = x, imptab[[x]])
+}))
+head(outtab)
+
+Yv = "conv_area"; Cv = "tot_C"; BDv = "cons_tot"; COSTv = "tot_cost"
+Yst = sum; Cst = sum; BDst = sum; COSTst = sum
+impacts <- batch_stat(outtab, Yv, Cv, BDv, COSTv, Yst, Cst, BDst, COSTst)
+optitab <- non_dominator(impacts, cnames, ybdc$both$bcode)
+optimal <- optitab
+bpath <- "agroEcoTradeoff/external/output"
+save(optimal, file = fp(bpath, paste0(ybdc$both$bcode, "/optitab2.rda")))
+save(outtab, file = fp(bpath, paste0(ybdc$both$bcode, "/impacts_tab2.rda")))
+# impacts[, lapply(.SD, range)]
+# impacts[1765:1771, ]
+# round(ybdc$both$params[1765:1771, 7:10], 2)
+
+ybdc$both$optitab <- optimal
+save(ybdc, outtab, impacts, file = fp(bpath, "mainresults2802.rda"))
+
 ```

### 2016-03-01 paper/main.tex
Supplementary methods draft complete. Required changes to biodiversity-layer.Rmd. Fixed bug in Figure 4 for main text--wrong map selected.
Submitted version 29/2/2016

--- 2016-02-26 paper/main.tex
@@
 \author[2]{Searchinger, T.}
-\author[2]{Spiegel, M.}
-\author[2]{Tian, D.}
-\author[3]{Kehoe, L.}
-\author[3]{Kuemmerle, T.}
+\author[1]{Spiegel, M.}
+\author[1]{Tian, D.}
+\author[3]{Sichinga, S.}
+\author[3]{Mwale, M.}
+\author[4]{Kehoe, L.}
+\author[4]{Kuemmerle, T.}
+\author[1]{Berven, A.}
+\author[5]{Chaney, N.}
+\author[1]{Sheffield, J.}
+\author[1]{Wood, E.F.}
 \author[1]{Caylor, K.K.}
@@
 \affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ, 08544 USA}
-\affil[3]{Humboldt University, Berlin}
+\affil[3]{Zambia Agricultural Research Institute, Mt. Makhulu Research Station, Chilanga, Zambia}
+\affil[4]{Geography Department, Humboldt University, 10099 Berlin, Germany}
+\affil[5]{Program in Atmospheric and Ocean Sciences, Princeton University, Princeton, NJ USA}
 \date{}
@@
 \begin{abstract}
-Blah blah
+Rapidly rising populations and likely increases in incomes in sub-Saharan Africa make tens of millions of hectares of cropland expansion nearly inevitable, even with large increases in crop yields.
+Much of that expansion is likely to occur in higher rainfall savannas, with substantial costs to biodiversity and carbon storage.
+Zambia presents an acute example of this challenge, with an expected tripling of population by 2050, good potential to expand maize and soybean production, and large areas of relatively undisturbed miombo woodland and associated habitat types of high biodiversity value.
+Here we present a new model designed to explore the potential for targeting agricultural expansion in ways that achieve quantitatively optimal trade-offs between competing economic and environmental objectives: total converted land area (the reciprocal of potential yield); carbon loss, biodiversity loss, and transportation costs.
+To allow different interests to find potential compromises, users can apply varying weights to examine the effects of their subjective preferences on the spatial allocation of new croplands and its costs.
+We find that small compromises from the objective to convert the highest yielding areas permit large savings in transportation costs, and the carbon and biodiversity impacts resulting from savanna conversion.
+For example, transferring just 30\% of weight from a yield maximizing objective equally between carbon and biodiversity protection objectives would increase total cropland area by just 2.7\%, but result in avoided costs of 27\%-47\% for carbon, biodiversity, and transportation.
+Compromise solutions tend to focus agricultural expansion along existing transportation corridors and in already disturbed areas.
+Used appropriately, this type of model could help countries find agricultural expansion alternatives and related infrastructure and land use policies that help achieve production targets while helping to conserve Africa's rapidly transforming savannas.
+
+%Rapidly rising populations and likely increases in incomes in sub-Saharan Africa make tens of millions of hectares of cropland expansion nearly inevitable even with large increases in crop yields.
+Much of that expansion is likely to occur in higher rainfall wetter savannas, at substantial costs to biodiversity and carbon storage.
+Zambia presents an acute example of this challenge with an expected tripling of population by 2050, good agronomic potential to expand maize and soybean production, and large swaths of relatively undisturbed Miombo woodland and associated habitat types of high value to mammals and birds.
+Here we present a new model designed to explore the potential to target agricultural expansion in ways that achieve reasonable trade-offs among different economic and environmental objectives.
+To allow different interests to find potential compromises, the model allows users to apply different weights to different objectives and examine how they alter outcomes.
+We find that modest compromises in yield targets permit large savings in transportation costs, and carbon and biodiversity impacts from savanna conversion.
+For example, transferring just 30\% of weight from a yield maximizing objective equally to carbon and biodiversity would increase total cropland area by just 2.7\%, but result in avoided cost of 27\%-47\% for the carbon, biodiversity, and transportation cost objectives.
+Compromise solutions tend to focus agricultural expansion along existing transportation corridors and in already disturbed areas.
+Used appropriately, this type of model could offer a practical means by which countries could explore agricultural expansion alternatives and target transportation investments and land use policies to achieve production targets while helping to conserve the values of Africa's savannas.
 \end{abstract}
@@
 \section*{\large Introduction}
+Meeting growing food demands while minimizing ecological losses presents one of the 21st Century's major challenges.
+Many global sustainability studies and climate mitigation pathways call for eliminating or rapidly phasing out emissions from land use change \cite{searchinger_high_2015,field_summary_2014}.
+Unfortunately, these ambitions do not appear realistically attainable in sub-Saharan Africa, due to a projected doubling of population by 2050 combined with rapidly growing  economies, which will combine to treble food demand \citep{tilman_global_2011,searchinger_high_2015}.
+For example, even with the substantial yield growth (3-5\% per year) projected by the UN Food and Agriculture Organization (FAO), the region would likely need to add more than 100 million hectares of cropland by 2050 \citep{searchinger_high_2015}.
+Much and probably the great bulk of that new cropland is likely to occur in the ~500 million hectares of savannas and shrublands of the region that receive sufficient rainfall to permit crop production \citep{searchinger_high_2015}.
 
-   Meeting growing food demands while minimizing loss of native ecosystems and their carbon and biodiversity presents one of the world’s great challenges for food and the environment.
-Many global sustainability studies and climate mitigation pathways call for eliminating or rapidly phasing out emissions from land use change.
-CITATIONS IPCC AR V; UK SCIENCE OFFICE.
-Unfortunately, these ambitions do not appear realistically attainable in sub-Saharan Africa due to a population likely to more than double to 2.1 billion by 2050.
-For example, even with healthy linear yield growth from 2006 to 2050 at 3\% to 5\% per year projected by the UN Food and Agriculture Organization (FAO), the region would likely need to add more than 100 million hectares of cropland by 2050.
-Searchinger et al.
-2013.
-Much and probably the great bulk of that new cropland is likely to occur in the ~500 million hectares of savannas and shrublands of the region that receive sufficient rainfall to permit crop production.
-Searchinger et al.
-2013.
-This conversion would release tens of millions of tons of carbon and have large impacts on biodiversity.
+Savannas sometimes receive limited respect in global land use assessments and have been a major focus of agricultural expansion, not only in Africa but in places such as the Cerrado of Brazil and northern Thailand \citep{morris_awakening_2009}.
+Their carbon contents and biodiversity values may be completely ignored on the grounds that they are not forests \citep{searchinger_high_2015}.
+Yet, the wetter woodland-savannas and shrublands of Africa have a similar average richness, although not density, for birds and mammals as the world’s wetter tropical forests, and 75\% of these habitats have higher overall vertebrate diversity than 75\% of the rest of the world \citep{searchinger_high_2015}.
+Their carbon contents are also substantial, particularly relative to their likely yields, so that their use for food crops is unlikely to result in fewer carbon emissions per ton of crop \citep{searchinger_high_2015}.
+Conversion of over 100 million hectares of these lands would release tens of millions of tons of carbon.
 
-[I would add in here something about the fact that this is a major challenge, it is one that needs careful planning to achieve, but that this planning has to be done at the country level as this is where most land use policy decisions are made.
-Here we look at Zambia, which is perhaps best representative of the challenge.
-[this ties us back to the call we made in the NCC paper for finer-level planning, and also inoculates us from the parochiality problem, as we are looking at a representative country].]
+The high likelihood of some cropland expansion in sub-Saharan Africa raises the prospect of whether this expansion could be located in ways that allow agricultural potential to be achieved, but for substantially lower loss of carbon and biodiversity than conventional agricultural development pathways.
+Analyzing this question must be done at national or sub-national scales, because that is where key decisions are made, and it requires data that is both in resolution and quality to provide confidence for decision-makers.
 
-Zambia presents an acute part of the challenge.
-According to the medium estimate of the United Nations, Zambia’s population is likely to grow 265\% from 16.2 million in 2015 to 43 million in 2050.
-(United Nations Department of Economic and Social Affairs Population Division.
-2015.
-World Population Prospects:  The 2015 Revision, DVD Edition).
-As per capita calorie food availability in Zambia is also low, hunger indices high (with child stunting rates at  40\%), https://www.wfp.org/stories/10-facts-about-hunger-zambia GET BETTER CITATION) , and the percentage of the population dependent on agriculture at X\%, Zambia has great need to boost its agricultural production.
-Zambia has also been expanding its agricultural exports, and has agronomic and situational characteristics that put it in a strong position to expand exports to other parts of the region.[1]
+In this paper we look at Zambia, which is a bellwether for this land use challenge in sub-Saharan Africa.
+On the one hand, Zambia has great need to boost its food production.
+Sixty percent of Zambia's population is rural {\citep{the_world_bank_world_2016}, per capita calorie food availability is just 2100 kcal \citep{fao_faostat:_2016}, and child stunting rates are 40\% \citep{haddad_global_2015}.
+According to the medium estimate of the United Nations, Zambia’s population is projected to grow 265\% from 16.2 million in 2015 to 43 million in 2050 \citep{united_nations_world_2015}, which will combine with rapid economic growth \citep[$>$6\%,][]{the_world_bank_world_2016} to substantially increase per capita calorie demands \citep{tilman_global_2011}.
+Zambia has also been expanding its agricultural exports, and has agronomic and situational characteristics that put it in a strong position to continue to do so to other parts of the region.
 
-At the same time, two thirds of Zambia’s 75 milloin hectares is dominated by Miombo Woodland considered dense to be categorized as forest by the UN REDD program, or X\%.
-(Vinya, R., Syampungani, S., Kasumu, E.C., Monde, C.
-\& Kasubika, R.
-(2011).
-Preliminary Study on the Drivers of Deforestation and Potential for REDD+ in Zambia.
-A consultancy report prepared for Forestry Department and FAO under the national UN-REDD+ Programme Ministry of Lands \& Natural Resources.
-Lusaka, Zambia).
-Using this measure of forest, Zambia’s deforestation rate has been estimated at 250,000 to 300,000 hectares per year, among the highest rates of deforestation of any country in the world.
-Ibid.
-Fifty percent of these X million hectares have carbon stocks in excess of X tC/h, and an additional X\% have carbon stocks in excess of X.
+At the same time, two thirds of Zambia’s 738,400 km$^2$ land area is dominated by woodland savannas, which have sufficiently dense tree cover to be classified as forest under either of the United Nations Framework Convention on Climate Change thresholds ($>$10 or $>$30\% tree cover) for defining this vegetation type \citep{sexton_conservation_2015}.
+These ecosystems store substantial amounts of carbon, with 50\% of Zambia having a carbon density $\geq$130 t ha$^{-1}$ (vegetative carbon and soil carbon in the top 1 m), while an additional 40\% exceeds 80 t ha$^{-1}$.
+Meanwhile, Zambia’s deforestation rate of 2,500--3,000 km$^2$ year$^{-1}$ is one of the highest in the world \citep{vinya_preliminary_2011}.
 
 The biodiversity of these woody savannas is also high.
-Overall, the wetter woodland-savannas and shrublands of Africa have a similar average richness for birds and mammals as the world’s wetter tropical forests, and 75\% of these habitats have higher overall vertebrate diversity than 75\% of the rest of the world.
-(Searchiner et al.
-2015).
-Most of Zambia is covered by what the World Wildlife Fund has categorized as Central Zambian Miombo Woodland.
-It contains mature miombo woodland trees typically 15 to 20 meters tall, around a quarter of which are evergreen, with a broadleaf shrub and grass understory beneath.[2]  It is also interspersed with a large quantity of wetlands, up to 30\% of many areas.
-Its large expanse of habitat make it a center of large mammals, including elephant, black rhino and African buffalo.
-Although endemism is low, birdlife is rich with X species, and reptile and amphibian endemism is high, with 19 endemic reptile species 13 endemic amphibians.
-LYNDON, CAN WE ADD SOME NUMBERS OF PLANT SPECIES, AND VERTEBRATE SPECIES.
-Zambia contains eight other ecoregions according to WWF classifications.[3]   Its extensive wetlands are particular centers of bird and animal life and under extensive pressures.
-The value of this habitat is based in part on its remaining relative abundance, which makes it possible for wide-ranging mammals, including many predator species, to survive.
+Most of Zambia is covered by what the World Wildlife Fund has categorized as Central Zambian Miombo Woodland \citep{olson_terrestrial_2001,hogan_central_2004}.
+It contains mature miombo woodland trees typically 15 to 20 meters tall, around a quarter of which are evergreen, with a broadleaf shrub and grass understory beneath, and is interspersed with wetlands that comprise up to 30\% of the landscape \citep{hogan_central_2004}.
+In addition to hosting one of the world's three richest mammal assemblages \cite{olson_terrestrial_2001}, the large expanses of relatively intact habitat make these woodlands, and several of Zambia's eight other ecoregions \citep{fund_ecoregions_2011, estes_zambezian_????}, important centers of large mammal populations, including black rhino, African buffalo, and highly migratory or wide-ranging species, such as elephants and major predators \citep{hogan_central_2004}.
+Zambia in general has a high bird diversity \citep{searchinger_high_2015}.
+Plant diversity in much of the country ranges between 3,000-5,000 species, a richness that is exceeded only in tropical forests \citep{kier_global_2005}.
+Although there are relatively few endemic bird and mammal species, reptile and amphibian endemism is high, with 19 endemic reptile species and 13 endemic amphibians \citep{fund_ecoregions_2011}.
 
-These multiple trends imply challenging trade-offs between agricultural development and environmental concerns.
-They raise the question how could decision-makers best target agricultural expansion so that it balances economic, carbon and biodiversity concerns.
-A related question is how could a system for targeting well illuminate the trade-offs to support decision-making and a proper role for influence by stakeholders.
-And an important question is whether if so what benefits could be achieved by targeting land expansion in certain areas from the standpoint of achieving areas with a desirable mix of objectives.
+There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs or within a constrained budget of land area for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}.
+Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.
+One of the challenges for land use planning in practice is that decisions involve trade-offs among different objectives and values that most people find difficult to weight against each other \emph{ex ante}, because the values are often abstract, making them harder to evaluate without first viewing the consequential trade-offs.
+Here we present a new model in which varying weights can be placed on the competing land use objectives.
+This set of options then allows model users to choose the most appealing combination of weights based on their land use preferences, and their own assessments of what are and are not acceptable costs within those objectives.
+The objective of this is to assist different stakeholders in finding compromises between competing land use interests while still meeting agricultural production targets.
+We apply this model to Zambia so that we can explore its strengths and limitations as a decision-making tool, and also to answer a basic but important question: is it possible to achieve a reasonable balance of different interests by reducing the carbon and biodiversity costs of expansion while still focusing on land use ares with high agricultural potential?
 
-[We need a somewhat clearer phrasing of our objectives here, or we need to state the objective below the next paragraph.
-What is that objective: two-fold I would say.
-1) Is this sort of modeling approach useful and/or does it have realistic potential for finding compromises in a more general sense (i.e.
-can it be applied beyond just Zambia), 2) how much potential does Zambia appear to have for reducing the impacts of its expansion.
-Something along these lines.]
-
-There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs subject or while limited land use budget for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}.
-Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.
-We adopt a similar approach here, using a newly developed model in which varying weights can be placed on the different objectives, thereby allowing optimal spatial allocations to be found that reflect different degrees of compromise between competing land use interests.
-To enable this functionality, the model is structured to evaluate each objective in terms of efficiency, or the ratio of cost to yield benefit that results from converting land to agriculture in order to achieve a production target.
-
-We apply this model to answer a basic but important question: can agricultural expansion in Zambia be achieved for substantially lower environmental cost than a business as usual approach, where expansion occurs without consideration of its ecological consequences?
-This question is particularly important for tropical savanna shrubland, and grassland biomes, which are commonly assumed to provide a low environmental cost reserve for agricultural development \cite{searchinger_high_2015}.
+Answering this question is becoming increasingly important as savannas and other grassy biomes receive increasing agricultural development pressure, and may become increasingly targeted as an indirect result of the laudable efforts to conserve tropical forests \citep{searchinger_high_2015}.
 
 \section*{\large Materials and Methods}
-The model we developed, called \emph{AgroEcoTradeoff}, is designed to find the best land use configuration to satisfy the production targets for multiple crops while minimizing four costs: the total land area required, which is inversely related to productivity; the potential losses of carbon and biodiversity; and transportation costs, which are closely correlated with agricultural production costs (ref).
-The model takes the basic form:
 
-\begin{equation}
-B_{ijk} = \frac{J_{jk}}{P_{ij}} \quad i \in m, \ j = 1, \ldots,n, \ k = 1, \ldots,q 
-\end{equation}
+The model we developed seeks to find optimal land use configurations that satisfy the production targets for multiple crops while minimizing four costs: the total land area required, transportation costs, carbon released from land conversion, and impacts on biodiversity.
+Although total land area is treated as a cost, this cost is minimized by selecting areas with the highest likely yields.
+By minimizing total land required as well as transportation costs, the model finds areas that reflect two important indicators of agricultural potential.
+By minimizing carbon and biodiversity costs, the model finds areas that minimize two important dimensions of environmental impact.
+The model chooses final areas depending on the weights that a user assigns to each of these four objectives.
 
-Where J$_{k}$ is one of the four costs to be minimized, within a landscape composed of $n$ grid cells containing values for each of the $q$ cost objectives and potential productivity values $P_i$, for each of $m$ crops.
-$B_{ijk}$ therefore is an efficiency that describes the ecological cost per unit of production, which is normalized so that values are between 0 and 1, so that the different objectives can be compared in common units on a common scale.
-The normalized values are subtracted from 1 to equate high values with high probability of conversion.
-Each of these probabilities is then multiplied by a weight (range 0-1, constrained to sum to 1 across all four weights), and the four resulting products are summed, producing a scalarized single objective that represents the landscape conversion probability for each individual crop in each grid cell.
-An objective function seeks to maximize the sum of these conversion probabilities, based on the requirement that the sum of the potential production of the cells converted for each crop must meet that crop's production target.
-A second constraint restricts each cell so that it is converted to no more than one crop, as most crops are grown during the same season, and producing multiple crops on the same land would introduce further complexity in setting assumptions about the share of each crop allocated to shared pixels.
-We therefore formulated a greedy approximation algorithm, based on solving the knapsack problem  \citep{dantzig_discrete-variable_1957} and an algorithm for solving the generalized assignment problem \citep{cohen_efficient_2006}, which allocates each crop to its highest value cells.
+To enable this functionality, and to identify optimal solutions, the model is structured to evaluate each objective in terms of production efficiency, or the ratio of cost per hectare to crop production per hectare (yield) that results from converting land to agriculture.
+For land area costs, output is measured in terms of the number of km$^2$ required to meet the crop production target.
+For the other objectives, the indicators are transportation costs, carbon loss, or biodiversity impacts per ton of crop yield.
 
-This formulation allows quantitatively optimal solutions to be found across a wide range of weight combinations, which is valuable because it is rare to find a single solution that simultaneously obtains the lowest cost for each objective.
-By finding the optimal solutions across a range of weights, the model can identify the Pareto front, or the set of solutions where the cost in one objective cannot be further minimized without increasing the cost in another.
-This set then allows model users to choose the most appealing combination of weights based on their land use preferences, and their own assessments of what are and are not acceptable costs within each of the four competing objectives.
+We focus our model initially on two crops, maize and soybeans, because of their existing and potential significance for Zambia's agricultural sector, and their likely dominance of in the future growth of agriculture in Africa's broader savanna regions \citep{searchinger_high_2015}.
+Maize is the largest crop by area and consumption in Zambia \citep{fao_faostat:_2016}, while soybeans are expanding rapidly as a commodity crop to meet regional and global demands for animal feed and direct human consumption \citep{gasparri_emerging_2015}.
+Because the model is designed to meet production targets for multiple crops, we apply algorithms for allocating each crop to its highest value cells.
 
-This weighting system thus has two basic components.
-One reflects the pure value judgment for one output versus another, e.g., how much more a person cares about yield over carbon, or carbon over biodiversity.
-The other component expresses each output in quantitative terms so that it reasonably expresses preferences for that output.
-For example, units for the carbon, biodiversity and yield outputs are scaled so that someone seeking to weight yield at twice the value each of carbon and biodiversity can do so and have the model find an outcome that reasonably reflects this preference.
-This approach is necessary to support public decision-making processes, in which different individuals and stakeholders might assign different values to different interests, and where compromise solutions must be sought.
+Further details on model structure and all model input calculations can be found in the online electronic supplement.
 
 \subsection*{Model inputs}
-For conducting our analysis of potential land use tradeoffs in Zambia, we developed the following spatial datasets at a resolution of 1 km$^2$.
 
-\subsubsection*{\emph{Agricultural potential and production scenarios}}
-There are many ways of evaluating agricultural potential factoring in agronomic and economic characteristics.
-Here we focused on two crops, maize and soybeans, because of their existing and potential significance for Zambia’s agricultural sector, and because they are considered to be the two dominant commodities that will be grown in Africa's broader savanna regions \citep{searchinger_high_2015}.
-Maize is the largest crop by area and consumption in Zambia \citep{fao_fao_2012}, while soybeans are expanding rapidly as a commodity crop to meet rapidly growing regional and global demands for animal feed and direct human consumption \citep{gasparri_emerging_2015}.
+Our model focuses on the costs of cropland expansion into areas that are currently not farmed but could be.
+To identify such areas, we calculated the proportion of currently cropped and settled grid cells within each 1 km$^2$ grid cell.
+We excluded the proportion of cells that are either protected or with slope greater than 20\% slope, which are both generally considered unsuitable for farming and strong actual predictors of cropland presence or absence \citep{estes_using_2014}.
+We estimated the costs of converting each cell as described in the following sections.
 
-We mapped potential yields for each crop using a three-step approach.
-In the first step, we used the decision support system for agrotechnology transfer \cite[DSSAT, ][]{jones_modelling_2003} to simulate maize and soybean yields at the sites of 40 weather station distributed across Zambia, which represent the points where a 31-year, gridded daily meteorological dataset has received the greatest amount of bias-correction and infilling by weather observations \citep{sheffield_development_2006,estes_changing_2014}.
-We used soil profiles corresponding to those locations drawn from the WISE v1.1 gridded soil profile database \citep{harvestchoice_converting_2010}, and simulated for each location maize yields over all 31 years under commercial input practices for three-four cultivars representative of short and medium season length according to the Zambian crop calendar \cite{grassini_how_2015}.
-This step allowed us to characterize crop responses to Zambia’s inherent climatic variation, and to improved management practices, which was otherwise impossible given the coarse resolution of available observed yield data and the fact that much of Zambia’s current production occurs on low input, smallholder farms (cite).
-In the second step, we used an empirical model to map the DSSAT-modeled yields onto a 1 km grid, taking advantage of a finer resolution soil map that provides a common frame for estimating both yield potential and soil carbon stocks (see below), but has insufficient detail in several parameters (particularly effective rooting depth) required by DSSAT.
+\subsubsection*{\emph{Likely yields}}
+
+We spatially estimated likely yields for maize and soybeans using a three-step approach.
+In the first step, we used the decision support system for agrotechnology transfer \cite[DSSAT, ][]{jones_modelling_2003} to simulate maize and soybean yields at the sites of 40 weather stations distributed across Zambia.
+They are the points where a 31-year, gridded daily meteorological dataset has received the greatest amount of bias-correction and infilling by weather observations \citep{sheffield_development_2006,chaney_spatial_2014, estes_changing_2014}.
+We used soil profiles corresponding to those locations drawn from the WISE v1.1 gridded soil profile database \citep{harvestchoice_converting_2010}, and simulated for each location maize yields over all 31 years under commercial input practices for three-four cultivars representative of short and medium season length according to the Zambian crop calendar \citep{grassini_how_2015}.
+
+In the second step, we used an empirical model to map the DSSAT-modeled yields onto a 1 km grid, taking advantage of a finer resolution soil map that provides a common frame for estimating both yield potential and soil carbon stocks (see below), but has insufficient detail in several parameters (particularly effective rooting depth) required by DSSAT to provide a sound empirical basis for running the model directly.
 We used a generalized additive model \citep{wood_mgcv:_2001} to predict the values of the DSSAT-simulated yields, using best-fitting subsets of the following weather and soil variables: growing season precipitation; growing degree days; mean growing season shortwave solar radiation; percent organic carbon; pH; percent clay content.
 The soil predictors represented mean values in the top 1 m of soil, and were extracted from a new 1 km global database of soil properties \cite{hengl_soilgrids1km_2014}.
-These first two steps were necessary to 1) capture crop responses to Zambia's climatic variability and potential management practices, and 2) to base potential yields on the same dataset used to estimate carbon costs (see below).
-This resulted in a map of potential yield under relatively high inputs and good management, and assuming no losses due to other limitations such as diseases.
-These yield values therefore overestimated what farmers are truly likely to achieve.
-To correct this, we therefore rescaled the yield maps to match the mean yields (4.5 t/ha for maize and 3.6 t/ha for soybean) that are projected to be achieved in Zambia by 2050, using FAO-projections to derive an annual growth factor, which we then applied to the 2009-2014 reported mean yields for each crop.
-Overall, the distribution of yields is most important to targeting land use, but the average yield determines the total quantity of land needed to meet production targets.
+This two step approach allowed us to more soundly capture crop responses to Zambia's climatic variability and potential management practices, and to base potential yields on the same dataset used to estimate carbon costs (see below).
 
-As a second measure related to agricultural potential, we estimated the travel time to the nearest city of 10,000 people or more (check this), which serves as a proxy for production costs, such as fertilizer prices and transport costs (citation), and is strongly correlated with the suitability of land for farming (cites from ConBio paper).
+The first two steps resulted in a map of potential yield under high inputs and good management, and assuming no losses due to other limitations such as diseases.
+These yield values therefore overestimate what farmers are truly likely to achieve.
+To reflect more realistic estimates of future yields, we therefore rescaled the yield maps to match the mean yields that are projected to be achieved in Zambia by 2050 (4.4 t ha$^{-1}$ for maize and 3.6 t ha$^{-1}$ for soybean).
+We projected these yields by using FAO-projections for sub-Saharan Africa as a whole to derive an annual growth factor and applying that factor to the average 2009-2014 yields for each crop reported in Zambia.
+Overall, the distribution of yields derived from our crop modeling is most important to targeting land use, while this choice of average yield determines the total quantity of land needed to meet production targets.
+
+\subsubsection*{\emph{Transportation Costs}}
+
+As a second measure related to agricultural potential, we estimated the travel time to the nearest major town.
+Travel time serves here as a proxy for differences in output prices and production costs, such as fertilizer prices and transport costs \cite{stifel_isolation_2008}, and, via the density of road networks, is strongly correlated with the suitability of land for farming \citep{laurance_estimating_2015,estes_using_2014}.
 We used a cost-distance analysis drawing on spatial data for roads, rivers and water bodies, urban centers inside and neighboring Zambia, and national boundaries to calculate travel time in hours.
-Further details on the cost distance and yield estimation methods can be found in the online electronic supplement.
+Total travel costs for any land use scenario equal the total time needed to transport the tonnage of crops produced in each converted cell to the nearest town, which we divide by 20 to produce "truck hours" on the assumption of a 20-ton truck (although the division  of all cell results by this constant does not alter the mapping outcomes).
 
 \subsubsection*{\emph{Carbon}}
+
 To estimate potential carbon costs from cropland conversions, we developed maps of vegetative and soil carbon stocks for Zambia.
-Vegetative carbon was calculated from a map of above-ground biomass developed by Baccini et al.
-\citep{baccini_estimated_2012}, from which we estimated below-ground biomass using savanna and miombo woodland-specific root-shoot ratios \cite{mokany_critical_2006,mugasha_allometric_2013}, and then converted the sum to carbon using a ratio of 0.47 (IPCC 2006 cite).
+Vegetative carbon was calculated from a recent map of above-ground biomass developed by Baccini et al.
+\citep{baccini_estimated_2012}, from which we estimated below-ground biomass using savanna and miombo woodland-specific root-shoot ratios \cite{mokany_critical_2006, mugasha_allometric_2013}, and then converted the sum to carbon using a ratio of 0.47 \citep{ipcc_land_1996}.
 We used the same soil carbon dataset used for crop yield prediction together with an accompanying map of bulk density to calculate soil carbon stocks in the top 1 m.
@@
 \subsubsection*{\emph{Biodiversity}}
+
 There are many ways of evaluating biodiversity value and potential impacts, including species richness of different taxa in absolute terms or based on level of threat or endemism, and including factors designed to address the significance of contiguity or the impact of different types of development \citep[e.g.][]{newbold_global_2015,searchinger_high_2015, gasparri_emerging_2015,koh_spatially_2010}.
-Such measures should reflect both conservation priorities and the quality of available data.
-For a fine-scaled, country-level analysis such as this one, the species range map data that form the basis for many impact assessments (Kehoe et al, 2015) have an effective resolution that is too coarse (Laura's cite).
+The actual measures selected should reflect both conservation priorities and the quality of available data.
+For a fine-scaled, country-level analysis such as this one, the species range map data that form the basis for broader impact assessments \citep[e.g.][]{kehoe_global_2015} have an effective resolution that is too coarse.
 We therefore developed a composite measure of biodiversity value based on several datasets.
-First, we used the potential vegetation maps of East Africa \citep[][]{van_breugel_potential_2011}, a 2010 landcover dataset for Zambia (SERVIR cite), and a shapefile of Zambian protected areas to calculate how much of Zambia's 20 unique vegetation types remains after conversion to croplands and settlements, and what proportion of that remainder falls within national parks.
-From this we calculated a hybrid measure of rarity and threat, wherein we assigned a weight to each vegetation type based on its relative remaining area in proportion to Zambia's extent, log-transformed to account for a highly skewed distribution, and then re-weighted that weight by each vegetation types protected fraction, following \citep{van_breugel_environmental_2015}.
+
+First, we used the potential vegetation maps of East Africa \citep[][]{van_breugel_potential_2011}, a 2010 landcover dataset for Zambia\footnote{downloaded from http://apps.rcmrd.org/landcoverviewer/}, and a shapefile of protected areas to calculate how much of Zambia's 20 unique vegetation types remains after conversion to croplands and settlements, and what proportion of that remainder falls within national parks.
+From this estimate we calculated a hybrid measure of rarity and threat, wherein we assigned a weight to each vegetation type based on its relative remaining area in proportion to Zambia's extent, log-transformed to account for a highly skewed distribution.
+We then re-weighted that weight by the protected fraction of each vegetation type's protected fraction, following \citep{van_breugel_environmental_2015}.
 We then used the remaining vegetation map to calculate a second measure, the proportion of untransformed vegetation within a 11X11 km neighborhood centered on each 1 km$^2$, providing a measure of vegetation intactness.
-By adding this measure to the first index and normalizing, we created a unitless biodiversity score (range 0-1) that gives equal weight to how rare or threatened a habitat is, and how undisturbed it is.
+By adding this measure to the first index and normalizing, we created a biodiversity score (range 0-1 per ha) that gives equal weight to how rare or threatened a habitat is, and how undisturbed its local area.
+When the biodiversity score for each converted grid cell converted is summed, it generates a total cost of biodiversity impacts.
 
-As a final modification, we masked out the areas of existing national parks and game management areas, on the assumption that they would not be used for expanding farmland, but retained a third protected category, forest reserves, as these tend to be far less stringently protected, having had 8.3\% of their area converted to cropland or settlements according to the landcover data.
-We assigned these areas a high biodiversity score of 0.917 (1 - the already converted proportion), indicating the value of keeping such areas protected, but recognizing that they have some chance of being converted, which can be estimated from their current degree of transformation.
+As a final modification, we masked out the areas of existing national parks and game management areas to respect the national judgment that they should not be used for expanding farmland.
+We retained a third protected category, forest reserves, as these tend to be far less stringently protected, having already had 8.3\% of their area converted to cropland or settlements, but assigned these areas a high biodiversity score.
 
 \subsubsection*{\emph{Potential farmland and production targets}}
-Our tradeoff analysis is concerned with cropland expansion into areas that are currently not farmed but could be.
-To identify such areas, we calculated the proportion of currently cropped and settled grid cells within each 1 km$^2$ grid cell, and used a 30 m digital elevation model to calculate slope percent over the entire country, and calculated the proportion of cells greater than 20\%.
-Slopes above this percent are considered unsuitable for farming (USDA cite), and slope-based measures are strong predictors of cropland presence or absence \citep{estes_using_2014}.
-Adding the two resulting maps together, subtracting their value from 1, and removing protected areas provided the fractional potential cropland in each cell.
-In addition to defining cropland availability, this map also provided the area term necessary for calculating production from potential yield.
 
-To develop the production targets used by the model, we calculated the rate of change in annual production of maize for Zambia between 2000-2014, collected from the FAOStat database.
-For soybean we did the same, but used the mean production trend across the 9 southern African countries examined by Gasparri et al \citep{gasparri_emerging_2015}, given that soybean production is very underdeveloped in Zambia compared to countries such as South Africa that are currently experiencing rapid growth.
-Extrapolating linearly from 2015 until 2050, maize production will triple and soybean production will increase 9-fold.
-We used these linear extrapolations as the basis for our targets, adding an additional unit to each factor to account for the fact that Zambia, which has one of the highest agricultural potentials in sub-Saharan Africa, is likely to be subject to larger demands for exports to many of its neighbors, thus our final targets were a four-fold increase for maize (from 2009-2014 average production of 2,912,000 tons) and a 10-fold increase for soybean (from 203,700 tons).
+Likely future demand for maize and soybeans produced in Zambia is difficult to estimate, and our model can easily be run with different production targets.
+For this paper, we adopted targets of a four-fold increase for maize for 2050 compared to 2009-2014 average, and a 10-fold increase for soybeans.
+These increases reflect the rate of growth in Zambia from 2000-2014, and a mean production trend for soybean growth in southern Africa estimated by Gasparri et al \citep{gasparri_emerging_2015}, with small additional adjustments to to account for the fact that Zambia has one of the highest agricultural potentials in sub-Saharan Africa and has high potential to expand exports.
 
-Given that current Zambian maize and soybean yields are well below their current potentials (Waddington et al, 2010), it is an open question how much of future production needs will be satisfied by boosting production on existing croplands.
-However, it seems likely that some portion of this target will come from current land, given recent Zambian yield growth trends (FAOStat).
-In our production scenario, we therefore assumed that an amount equal to a quarter of the gap (redo) between current and projected yields would be met on current farmland before significant expansion occurs.
+\subsubsection*{\emph{Yields on current land}}
 
-\subsection*{Simulations}
-Using the selected production targets, we ran the model across all combinations of weights for the four parameters, varying the weights in 5\% increments.
-This allowed us to evaluate the impact of different land use preferences on the location of conversion impacts, the range and magnitude of potential conversion impacts, including for each objective, and to test how sensitive costs in each objective were to changes in weights.
+The amount of land conversion to meet a production target will depend on the growth of yields on existing croplands, which are well short of their current potentials \citep{waddington_getting_2010}.
+For the principal scenarios we present in this paper, we assume that yields on current cropland will achieve the average likely yields for Zambia's \emph{potential} farmland.
+However, to examine the possible significance of yield gains on existing cropland, we also examined scenarios in which the yield gains on existing cropland 1) only manage to close 50\% of the gap between current and our projected future yields, and 2) close 150\% of the gap.
+
+\subsection*{Normalization and Weighting}
+
+Weighting different objectives is to some extent based purely on preferences, but it is important that these be expressed in a numerical form and processed so that a quantitatively optimal solution for that set of preferences is identified.
+To achieve this result, we first express the different objectives as efficiencies (e.g.
+carbon loss per ton of crop yield, as described above), and then normalize the units of each objective based on the range of its scores.
+We then multiplying each normalized objective by its corresponding weight (which fall between 0 and 1 and must sum to 1, with values closer to 1 indicating stronger preferences) and then adding the resulting products.
+The result is a single score that ranks each grid cell according to its ability to meet the production target for the lowest cost, given those weights.
+The model then selects cells in descending rank order until each the cumulative production of converted cells reaches each crop's corresponding target (see online electronic supplement for details of land allocation between crops).
+
+Although this weighting system treats choices as preferences, our model calculates the actual costs of each objective in absolute terms:  area of land converted (which is increasingly minimized as weight is added to the preference to maximize crop yields), total transportation costs from new croplands, total carbon released from conversions, and total impacts on biodiversity.
+After reviewing these costs, model users can then adjust the preferences to reflect desirable or politically acceptable trade-offs.
 
 \section*{\large Results}
-The first order determinant of land use conversion costs is unrelated to any weights applied to the four objectives, as it is based on how much of future production needs will be met on existing croplands, and therefore how much non-agricultural land will be spared from conversion.
-In our main scenario, we assumed that the gap between current and our projected future potential yields would be fully closed, but to illustrate the impact of this assumption on land conversion costs, we also ran the model (with equal weights across all objectives) for cases when 50\% and 150\% of the gap was closed.
-In the former case, the total land area required to meet the target increased by 18\% (14,634 to 17,202 km$^2$), average travel time increased 3\% (1.96 to 2.01 hours), carbon losses increased by 19\% (597 to 711 mt), and the average biodiversity index of converted areas increased by 2\% (0.297 to 0.303).
-In the latter case, the impacts were reversed, with impacts declining by almost exactly the same percentage.
 
-To identify the minimum possible costs for each individual objective, and the degree to which their areas of new cropland correspond, we examined the simulations were 100\% of the weight was given to each objective.
-The yield maximizing objective (100\% weight to yield) describes both a cost in terms of the land area required, and a benefit that is based on the productivity gained from converting that land.
-Scenarios that involve the lowest land use costs also provide farmers the highest potential yields, and therefore have economic advantages.
-An examination of overlaps shows two interesting results.
-First, converted areas that would maximize yield have little correspondence with the areas that minimize carbon costs, biodiversity impacts or travel time (Figure 1), as measured along two dimensions: actual shared area, as measured by the percentage overlap between the two maps, and their adjacency to one another, calculated as the average distance between each pixel of one map and its nearest neighbor in the other.
-Second, the areas selected for minimizing carbon, biodiversity, and transportation costs have high adjacency, albeit little overlap.
-All three criteria tend to identify lands along existing road networks and population and agricultural centers.
-The reason for this correlation is that carbon stocks and biodiversity values are usually lowest in areas of high human density, because of high levels of charcoal and wood harvesting (cite) and fragmentation caused by settlements, existing croplands, and road networks, which exist primarily to connect agricultural and urban centers.
+\subsection*{Significance of yield gains on existing cropland}
+
+In our main scenario, we assumed that the gap between current and projected future yields would be fully closed on existing cropland, but we also ran the model (with equal weights across all objectives) for cases when only 50\% or 150\% of the gap was closed.
+In the former case, the total land area required to meet the target increased by 18\% (14,634 to 17,202 km$^2$), transport costs increased 22\% (5.52 X 10$^{-5}$ to 6.61 X 10$^{-5}$ truck hours), carbon losses increased by 19\% (597 to 711 Mt), and the total biodiversity index of converted areas increased by 20\% (5.53 X 10$^{-5}$ to 6.61 X 10$^{-5}$).
+In the 150\% gap closure scenario, the impacts were reversed, with impacts declining by almost exactly the same percentage.
+
+\subsection*{Overlap and adjacency from optimal individual solutions}
+
+To identify the minimum possible costs for each individual objective, and the degree to which their areas of new cropland correspond, we examined simulations giving 100\% of the weight to each of the four different objectives.
+Figure 1 maps the results, and shows the costs for each objective in each simulation, as well as the extent of overlap among objectives.
+
+The scenario is which the sole preference is to maximize yield (100\% weight) converts land with the most promising yield potential for farmers, which should have economic advantages.
+These areas have little direct overlap with the areas picked by scenarios that minimize carbon costs, biodiversity impacts or travel time (Figure 1).
+They also have little adjacency, as measured by the average distance between each pixel of one map and its nearest neighbor in the other (Figure 1).
+
+However, the areas selected solely for minimizing carbon, biodiversity, or transportation costs also have little direct overlap, but the map reveals that they have high adjacency, confirmed by direct estimates of distance (Figure 1).
+All three criteria tend to identify lands along existing road networks and population and agricultural centers, with the carbon minimizing objective converting a larger area of land in the western half of the country.
+The reasons for this correlation are probably that carbon stocks and biodiversity values are lowest in areas of high human density, because of high levels of wood harvesting for charcoal and fuelwood, and because human uses similarly cause fragmentation that leads to low biodiversity scores.
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/single_priorities2.png}}
-    %\includegraphics[scale=0.5]{fig2.eps}
+       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/single_priorities3.png}}
     \caption{Areas converted to new maize and soybean croplands and their associated costs when each individual objective (yield maximization, and carbon cost, biodiversity cost, and transport time minimization) is given 100\% weight.
-The impact categories are color-coded to the specific conversion maps.}
+Cropland converted under each scenario is shown in the maps, colour-coded according to which objective receives 100\% weight.
+Large bar charts, colour-coded to the maps, display the absolute value of costs resulting from each scenario.
+Small bar charts show the percentage of overlap (top) and adjacency (bottom; nearest neighbor distance) between the map directly above it, and each of the other three maps.}
     \label{default}
@@
 
-One of the goals of this analysis is to determine whether opportunities exist to achieve a reasonable accommodation between different objectives, and here the results are also interesting.
-While the best case scenario for maximizing yield would convert  million hectares, the worst case scenario (which would minimize carbon costs) would convert Y million hectares, only a Z\% increase, which is far lower than the variability in potential crop yields (3.7-6.1 t ha$^{-1}$ for maize and 0.8 to 5.1 ha$^{-1}$ for soybean).
-This result reflects our model's use, and the value, of an efficiency approach.
-Because yield is in the denominator of each objective (e.g.
-carbon loss per ton of yield, or biodiversity value per ton of yield), land with better yield potential tend to be selected--even when the yield maximization priority is given no weight--so that the crop production can be achieved with the least cost.
+\subsection*{Potential to harmonize different interests}
 
-By contrast to the limited differences for yields, the differences in transportation costs using existing transportation networks range from an average of 0.6 hours for each hectare to 6.9 hours, an increase of 1050\%.
-Carbon costs from minimum to maximum vary by 188\%, and average biodiversity impacts vary by 100\%.
-These differences suggest that there is high potential to reduce transportation, carbon and biodiversity costs for relatively sacrifice of yield potential.
+Although the ideal solutions to maximize each objective do not lead to much direct overlap, areas selected by less ideal but still good solutions for each objective may still overlap.
+The reasons start with the different potential ranges of costs among the scenarios that maximize each objective.
 
-This potential is illustrated in Figure 2, which shows how much cost is avoided in each objective when weights are transferred between them.
-Here avoided cost is calculated as the percent difference between the cost paid under a given weight combination and the largest cost that is incurred in all combinations (the worst case).
-For carbon, biodiversity, and travel time, the maximum cost occurs when yield is given 100\% weight, which is where the yield objective has the lowest possible cost, as measured by the percent of land saved relative to the largest area converted, because the cropland that is converted is the highest yielding.
-However, by sequentially adding weights in 5\% increments to both biodiversity and carbon conservation objectives, large reductions in not only those costs but also in average travel time (because of the correlations between these three measures).
-Weighting carbon and biodiversity by 15\% each reduces the average area of new lands (i.e.
-new crop area is increased) by just 3\%, in exchange for avoided costs of 34\% for biodiversity, 48\% for carbon, and 44\% for travel time, which respectively represent 68\%, 77\%, and 50\% of the maximum possible avoided cost for each objective.
+While the best case scenario for maximizing yield would convert 13,668 km$^2$, the worst case scenario among our single-purpose optimization scenarios (the scenario minimizing carbon costs) would convert 15,302 km$^2$.
+This range in converted areas is only 12\%, which is far less than the range in yields across the country’s potential cropland (3.7-6.1 t/ha/y ha$^{-1}$ for maize and 0.8 to 5.1 t/ha/y ha$^{-1}$ for soybean).
+One reason is the significance of yield to minimizing each of the costs because of the need to meet a fixed production target; because yield is in the denominator of each objective (e.g.
+carbon loss per ton of yield), land with better yield potential tends to be selected even when the yield maximization priority is given no weight.
+By contrast, the differences in travel times among single-objective scenarios range from a total of 3.4 X 10$^{-5}$ to 32.3 X 10$^{-5}$ truck hours, a difference of 872\%.
+Carbon costs differ by 156\% (462 to 1184 Mt), and biodiversity of 68\% (4.96 X 10$^{-5}$ to 8.33 X 10$^{-5}$).
+The low range in the yield-related cost (converted area), and the high ranges in the other costs, suggest that there is strong potential to reduce transportation, carbon and biodiversity costs for relatively little sacrifice of yield.
+
+This potential is illustrated in Figure 2, which shows how much carbon, biodiversity, and transport cost can be avoided by transferring increasing amounts of weight to these objectives from the yield maximization objectives.
+Avoided cost is the percent difference between the cost paid under a given weight combination and the maximum cost found across all weighting scenarios (we ran all possible weight permutations varying in 5\% increments).
+Sequentially adding weights in 5\% increments to both the biodiversity and carbon objectives results in little cost to yield, but rapid, large reductions in both biodiversity and carbon costs as well as transport costs, despite no weight given to transport costs because of its correlations with the other two objectives).
+For example, adding just 15\% weight each to carbon and biodiversity increases the area of new lands (i.e.
+reduces the average yield) by just 2.7\% (avoided area costs drop to 8\% from 10.7\%), but avoids costs equal to 27\% for biodiversity, 47\% for carbon, and 44\% for transportation costs.
+For comparison, in the best case (100\% weight) scenario for each objective, 61\%, 41\%, and 90\% of carbon, biodiversity, and transportation cost would be avoided.
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/weight_impact3.png}}
+       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/weight_impact5.png}}
     %\includegraphics[scale=0.5]{fig2.eps}
-    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.}
+    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.
+In these results, no weight was assigned to the transport cost objective.}
     \label{default}
@@
 
-%These substantial cost savings suggest that a user  achieved for relatively little transfer of weight between objectives indicates that a user primarily concerned with acquiring the highest-yielding land for agriculture would have to make relatively little sacrifice to give enormous gains to conservation interests, or to substantially reduce the potential costs of their farming operation.
-value of compromise in terms that users might be more likely
+Figure 3 shows the potential for compromise in which a hypothetical land user only interested in a single objective is willing to ``pay'' 5\% more costs than the lowest possible costs for that objective.
+For example, the yield maximizing user would be willing to pay for an extra 683 km$^2$ of new cropland, or the person interested in minimizing carbon costs would pay an additional 231 Mt of carbon.
+Lines in the bar charts show the minimum and maximum costs across all possible weighting scenarios.
+Each of the four different ``compromise'' scenarios, represented by a different colour, permits solutions that allow the other scenarios to save substantially relative to maximum potential costs.
+The carbon, biodiversity, and transport cost objectives in all but two cases avoid more than 50\% of their maximum potential costs when any of the other objectives is willing to pay 5\% more cost (Figure 3).
 
-These results show that transferring relatively little weight between objectives can have large savings.
-However, people involved in making land use decisions are more likely to evaluate their willingness to compromise in actual cost units, and from the perspective of how much more they might be willing to pay to compromise with other interests.
-We therefore evaluated four further scenarios, in which a hypothetical land user with interests in a single objective willingly pays up to 5\% more than the lowest cost case (i.e.
-the 100\% weight outcome) in order to compromise with other interests.
-The yield maximizing user is thus willing to ``pay'' for an extra 68,400 km$^2$ of new cropland, the carbon and biodiversity conservationists would respectively sacrifice an additional 2,312,000 tons of carbon and 0.013 biodiversity index units, and the production cost-minded agricultural developer would pay just 4.1 minutes of travel time.
-Given these sacrifices, we identified how much more the other three interests would have to pay relative to their lowest cost (100\% weight) scenarios, treating each of the three equally, i.e.
-the amount each has to pay extra is the same percentage of their lowest cost case.
-
-The impacts of this assessment show that in for all four cases, the carbon, biodiversity, and travel time objectives typically pay less than 50\% more than their lowest cost when another party is willing to pay 5\% more cost (Figure 3), and all four associated conversion maps show substantial convergence around road networks and population and agricultural centers.
+The location of the conversions associated with each of these compromises show substantial convergence along roads and population and agricultural centers.
+Overlap between the resulting cropland allocations exceeds 30\% between the yield and biodiversity and cost compromises, and 25\% between the biodiversity and carbon compromises.
+The carbon and cost compromises overlap least ($<$10\%), followed by carbon and yield ($\sim$15\%), but each pair of maps has high adjacency, with all compromise allocations having an average nearest neighbor distance of $<$20km.
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_5pct.png}}
-    %\includegraphics[scale=0.5]{fig2.eps}
-    \caption{The impact of transferring weights from the yield maximization objective to the carbon and biodiversity objectives in 5\% increments, measured in terms of the percent cost that is avoided in each objective relative to the highest cost paid across all weight combinations.}
+       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_5pct3.png}}
+    \caption{The results of scenarios in which each land use objective is willing to pay 5\% more cost to compromise with the other three objectives.
+Cropland converted under each compromise scenario is shown in the maps, colour-coded according to which objective is making the 5\% compromise.
+Large bar charts, colour-coded to the maps, display the absolute value of costs resulting from each compromise scenario, together with the maximum, minimum, and median costs across all weighting permutations (grey horizontal lines in bar charts).
+Small bar charts show the percentage of overlap (top) and adjacency (bottom; nearest neighbor distance) between the map directly above it, and each of the other three maps, with the colour denoting which map is being compared.}
     \label{default}
@@
 
-Figure 3 also maps and shows the results for the equal weighting option and how it contrasts with the minimum costs for each factor.
-It shows increased savings for each factor of X to X for transportation costs, carbon and biodiversity compared to their maximum costs.
-It also holds these costs to no more than X above their minimum costs.
+The weight on the objective subject to the 5\% compromise is proportional to the size of its potential avoided cost (see Table 1 in supplementary results), thus yield requires a weight of only 40\% to pay 5\% or less of its lowest cost, whereas transportation cost and carbon, which have the largest cost ranges, require 85\% weight each, whereas biodiversity requires 50\% weight.
+
+We also analyzed an ``equal compromise''  scenario in which each objective receives equal weight, as shown in Figure 4.
+The bar charts show how this scenario, represented in purple, compares with each of the 5\% compromise scenarios shown in Figure 3, plus the best and worst case options among all weighting scenarios.
+Relative to the worst case, equal compromise avoids 80\% or more of the cost for carbon, biodiversity, and transportation, and 41\% of the cost for land area.
+With respect to the 5\% compromises, the equal weighting scenario is lower than, but compares favorably to, the $\sim$95\% savings achieved by carbon, biodiversity, and transportation costs within their individual cost categories, and to the 58\% savings by the yield objective.
+
+As a general rule, the areas selected for conversion under equal compromise lie along existing major transportation corridors, and generally near areas of existing cropland.
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_25pct.png}}
-    \caption{25\% compromise}
+       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_25pct2.png}}
+    \caption{Areas converted to new maize and soybean croplands and their associated costs when when all objectives (yield maximization, and carbon, biodiversity, and transport cost minimization) compromise equally (i.e.
+receive 25\% weight).
+Cropland converted under this scenario is denoted by purple areas in the map, with its associated cost impacts shown alongside those from the 5\% compromise scenarios (Figure 3), together with the maximum, minimum, and median costs across all weighting permutations (grey horizontal lines in bar charts).}
     \label{default}
@@
 
-Another interesting feature of these results is that there is little overlap between any targeting areas selected by any of the criteria we have calculated and designated farm development blocks.
-In 2005, the Zambian Ministry of Agriculture designated nine farm blocks with roughly one million hectares for assisted agricultural development and spread around the country.
-(Zambia Ministry of Agriculture.
-2005.
-Farm Block Development Plan 2005-07).
-Although these areas have in fact as a whole attracted little development, they remain “on the books,” and there are at least some reports of continued intent to develop them.
-Undeveloped farm blocks worry Agriculture Minister, Zambia Daily Mall: https://www.daily-mail.co.zm/?p=29505
-
-%\begin{equation}
-%  \textrm{I} = \left(\frac{\textrm{A}}{\textrm{A + C}} + \frac{\textrm{D}}{\textrm{B + D}}\right)0.5
-%\end{equation}
+All of these scenarios also suggest the value of deliberate and optimized land use planning.
+Zambia has designated a large number of farm blocks, and added nine more blocks with roughly one million hectares as recently as 2005 \citep{anonymous_farm_2005}.
+Although these areas have in fact as a whole attracted little development, they remain ``on the books'', and there are at least some reports of continued intent to develop them \citep{anonymous_undeveloped_2015}.
+But there is less than 2\% overlap between designated farm blocks and either the yield maximization scenario and the equal weighting scenario, and very little adjacency (see supplementary results).
 
@@
 
+Our results for Zambia suggest at least the potential to plan agricultural expansion such that it limits carbon and biodiversity costs with limited sacrifice of yield potential.
+In fact, if agricultural projection goals alone are best represented by a combination of agronomic potential and existing road access, then solutions that provide a good balance of these objectives also correlate strongly with solutions that limit carbon and biodiversity impacts.
+
+This finding demonstrates the importance of targeting land uses based on their ability to deliver a particular benefit for the lowest possible cost.
+The value of such efficiency-based approach, which we apply here to agricultural land use, is in line with findings from the field of conservation planning, where it has been shown that, by considering both the price of land and its ecological value, conservation objectives can be met for less than half the cost of approaches that target land solely based on ecological value \citep{naidoo_integrating_2006}.
+
+Another result, if more obvious, is the potential benefits of increasing yields on existing cropland.
+Actions that boost yields on existing cropland also have the potential to have feedback effects that lead to more local expansion, for example, by boosting the competitiveness and therefore quantity of exports \cite{ewers_increases_2009,angelsen_policies_2010}.
+Our analysis ignores those potential feedback effects because all options analyzed meet the same production targets, but our analyses quantities the large potential benefits if these yield gains are used to limit environmental impacts.
+
+The fact that the most advantageous compromises occur along existing transportation corridors also appears to have several  advantages.
+Road construction has been a primary driver of the location of agricultural expansion \citep{laurance_impacts_2009}.
+Zambia is already engaged in a program to improve existing major roads, but there is evidence that agriculture responds in particular to improvements in the quality of smaller, feeder rural roads to larger paved roadways \citep{bourguignon_rural_2008,blimpo_public_2013}.
+Improving smaller rural roads would be one way of focusing Zambia's agriculture development.
+Because agricultural development tends to build on the private and public infrastructure that accompanies initial development, the focus on existing areas would also seem likely to be a more robust, long-term strategy for directing agricultural development \cite{laurance_estimating_2015}.
+A potential impediment to this, however, might be that land closer to roads is already ``spoken for'', and thus less likely to be available for new farms.
+
+This model appears to be the right kind of tool for assisting the planning of agricultural development.
+Optimization models that require ex ante specification of preferences in mathematical terms work less well for decision-making that must weigh very different objectives in unclear ways.
+Our model makes possible an iterative optimization process that allows people to realize their preferences in the face of real information about options.
+It is particularly suitable to decisions that will inevitably have strong political elements, and to consensus-building among stakeholders with different preferences.
+
+For this kind of tool to be truly legitimate and useful on a practical level, we believe several elements are required.
+First, it must be further developed and applied in an iterative way with the government and other stakeholders so that the impact of different weighting decisions, in terms of their costs and benefits, can be fully understood, thereby facilitating informed land use planning.
+
+Second, development planning must reflect and incorporate other real world factors beyond those incorporated in our model.
+For example, large, multi-lane roads can lead to high rates of wildlife mortality, and serve as effective barriers to migration for large mammals, regardless of other land uses \citep{holdo_predicted_2011}.
+Although our ``equal weight'' scenario identifies some target areas in the heavily wooded northeastern part of the country, these concerns would be one reason to explore whether areas closer to existing cropland concentrations in the south might be reasonable substitutes.
+Such consolidation might also make it easier to support agricultural development with lower attendant infrastructure development costs.
+And this biodiversity concern is only one example of other issues, many purely practical, which should, and inevitably will, influence where agriculture develops.
+
+Third, data quality is critical.
+This kind of tool focuses on particular hectares, not broad averages, so data errors have little or no chance of averaging out through aggregation.
+Crop modeling predictions and most of our other scores can be highly sensitive to a range of data parameters.
+These facts mean that long-term use of the tool should be accompanied by steady data improvement.
+It also means that the tool should mostly serve as a guide to likely values.
+Before governments will make important decisions ``for keeps'', there should be site-specific ground-truthing of soil properties and yield potential, as well as confirmation of biodiversity and carbon characteristics.
+
+Some elements of the model should also be further developed, depending on the ability to gather significant additional data.
+Analysis of agricultural potential could incorporate more economic factors, differentiate between smallholder and commercial farming practices, and incorporate irrigation potential.
+Additional information on land ownership and title would allow a finer delineation of potential agricultural land, while the carbon estimates would be improved by root-shoot specific to each vegetation type.
+Biodiversity impacts could be better assessed by incorporation data on endemism and the degree to which they are impacted by habitat fragmentation.
+One potentially promising approach would be to incorporate the biodiversity impact model developed by Newbold et al.
+\citep{newbold_global_2015}, provided the land use intensity maps on which it is conditioned can be developed for this resolution.
+
+Overall, however, our results are promising, and suggest that efficiency-based land use planning approaches such as ours can help to substantially minimize the ecological cost of cropland expansion in sub-Saharan Africa's savannas, the 21st Century's emerging hotspot of agricultural land use change.
+
 \section*{\large Data accessibility}
+The model source and inputs data are available online at https://github.com/PrincetonUniversity/agroEcoTradeoff.
+
 \section*{\large Authors' contributions}
+L.E.
+and T.S.
+conceived and designed the research, together with K.C.
+L.E.
+and M.S.
+designed and coded the model.
+L.E.
+developed the input datasets, and ran and analyzed simulations, together with T.S.
+D.T.
+and L.E.
+developed the crop yield potential surfaces.
+L.K.
+and T.K.
+helped design the biodiversity impact metric.
+S.S.
+and M.M.
+collected and processed spatial datasets used in developing inputs.
+A.B.
+processed infrastructure data and helped develop the travel cost surface.
+J.S., N.C., and E.W.
+developed the bias-corrected weather data needed to estimate yield potential.
+L.E.
+and T.S.
+wrote the manuscript, with input from K.C., T.K., and L.K.
+
 \section*{\large Competing interests}
+We declare we have no competing interests.
+
 \section*{\large Funding}
+This work was supported by funds from: the Norwegian Aid Agency (NORAD) under the Agricultural Synergies Project; the Princeton Environmental Institute Grand Challenges program; the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463).
+
 \section*{\large Acknowledgements}
-This work was supported by funds from: the Norwegian Aid Agency (NORAD) under the Agricultural Synergies Project; the Princeton Environmental Institute Grand Challenges program; the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463).
+We thank Paulo van Breugel for providing the potential vegetation map of East Africa, and Philip Thornton for advice regarding modeling of yield potential.
 
@@
 
-%\label{}
-
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
-%% \appendix
-
-%% \section{}
-%% \label{}
-
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%
 %\section*{\large References}

### 2016-03-01 paper/supplementary_methods.tex
Submitted version 29/2/2016

@@
 
-The agroEcoTradeoff model was designed to identify optimal land use configurations that allow a certain level of agricultural production to be achieved while minimizing associated ecological and economic costs.
+Our model was designed to identify optimal land use configurations that allow a certain level of agricultural production to be achieved while minimizing associated ecological and economic costs.
 The model follows the rationale of Koh and Ghazoul \citep{koh_spatially_2010}, which was designed to select areas with promising potential for oil palm production but whose conversion would result in less loss of carbon, avian diversity, and land with high rice production potential.
@@
 The GAM-predicted yields where adjusted so that their mean values matched 2050 projected mean yields for Zambia, derived from FAO data.
-Yields are shown for locations outside of protected areas (GMAs = game management areas; NatParks = national parks).
+Yields are shown for locations outside of protected areas.
+}
+      \label{fig:default}
+\end{center}
+\end{figure}
+
+\subsection*{Carbon}
+
+Figure 2 shows the carbon densities (100\% of vegetative and 25\% of soil carbon in the top 1 m) used by the model to calculate total carbon loss upon conversion to cropland.
+
+\begin{figure}[ht]
+\captionsetup{width=0.95\linewidth, font=small}
+    \begin{center}
+       \makebox[\textwidth][c]{\includegraphics[width=1\textwidth]{figures/carbon.png} }
+      \caption{The density of carbon loss (in tons ha$^{-1]}$) per grid cell used by the model to calculate total carbon loss when cells are converted to cropland, as shown for locations outside of existing protected areas.
 }
@@
 
+The final biodiversity index map is shown in Figure 2, with 0 being the lowest value areas and 1 the highest.
+
+\begin{figure}[ht]
+\captionsetup{width=0.95\linewidth, font=small}
+    \begin{center}
+       \makebox[\textwidth][c]{\includegraphics[width=1\textwidth]{figures/bd.png} }
+      \caption{Biodiversity index values for Zambia, shown for locations outside of existing protected areas (GMAs = game management areas; NatParks = national parks), excluding forest reserves (visible as the darkest green areas on the map).
+}
+      \label{fig:default}
+\end{center}
+\end{figure}
+
 \subsection*{Estimating Travel Time}
@@
        \makebox[\textwidth][c]{\includegraphics[width=1\textwidth]{figures/cost.png} }
-      \caption{Cost distance surface for Zambia, showing travel time in hours to the nearest major town.
-Travel times are shown for locations outside of protected areas (GMAs = game management areas; NatParks = national parks).
-}
+      \caption{Cost distance surface for Zambia, showing travel time in hours to the nearest major town. Travel times are shown for locations outside of protected areas. }
       \label{fig:default}
@@
 
+\clearpage
 %\section*{\large References}

### 2016-03-01 paper/supplementary_results.tex
Submitted version 29/2/2016

@@
 \usepackage[numbers,sort,compress]{natbib}
+\usepackage[skip=2pt,font=normalsize]{caption}
 
@@
 
-\section*{\large Crop yields}
+\section*{\large Allocation of land between crops}
+
+To compare the degree to which the cropland allocation routine of \emph{agroEcoTradeoff} causes land that is best for maize to be allocated to soybean and vice versa, we ran the model for each crop separately (with 100\% weight on yield maximization) and compared the overlaps between the resulting conversion maps (Figure 1).
+The two crops were allocated to the same areas in $<$1\% of converted pixels.
 
 \begin{figure}[!ht]
+\captionsetup{width=0.95\linewidth, font=small}
        \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/ideal_yield2.png}}
+       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/ideal_yield3.png}}
         \end{center}
@@
 
-\section*{\large References}
+\section*{\large Overlaps with farm blocks}
+
+\begin{figure}[!ht]
+\captionsetup{width=0.95\linewidth, font=small}
+       \begin{center}
+       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/fb_overlaps.png}}
+        \end{center}
+      \caption{Simulations run under the 100\% weight to yield maximization and the equal weights scenarios (25\% weight to all objectives), showing the areas of overlap with designated farm development blocks.
+}
+      \label{fig:default}
+\end{figure}
+
+\clearpage
+\section*{\large Weights associated with 5\% cost compromises}
+
+\begin{table}[ht]
+\begin{center}
+\captionsetup{width=0.85\linewidth, font=small}
+\caption{Weights corresponding to the scenarios in which each objective was a) willing to pay 5\% more than its lowest possible cost, and b) the costs for the other three objectives were minimized equitably.}
+\begin{tabular}{ccccc}
+\hline
+& \multicolumn{4}{c}{Weights}\\
+Objective making 5\% compromise & Yield & Carbon & Biodiversity & Transport Cost \\
+\hline\hline
+Yield & 40 & 10 & 5 & 45 \\
+Carbon & 15 & 75 & 0 & 10 \\
+Biodiversity & 15 & 5 & 50 & 30 \\
+Travel time & 5 & 5 & 0 & 90 \\
+\hline
+\end{tabular}
+\end{center}
+\label{default}
+\end{table}
+
+%\section*{\large References}
 \bibliographystyle{prsb} 

### 2016-05-05 Rmd/figures.Rmd
Full paper revisions: main manuscript, fully revised, with edits and additions to SI methods and results.  New figure created (figures.Rmd), slight edit of loose code on main-analysis.Rmd

@@
 p_out <- fp(p_mod, "external/output")
+
+for(i in c("selected_impacts.R", "imp_sorter.R", "imp_plot2.R")) {
+  source(fp(p_root, paste0("R/", i)))
+}
 ```
@@
 ## Figures
-
-How reconcilable are the different land use interests, and what would be the impacts of each?
-
-I am using the two crops combined approach because there is little overlap between the two.
-
-### Figure 1
+### Figure 1 - pure objectives
 ```{r, eval = FALSE}
@@
 }))
-
-# m2 <- matrix(rep(5, 20), 4, 5)
-# m3 <- do.call(cbind, lapply(6:12, function(x) matrix(rep(x, 12), 4, 3)))
-# lmat <- rbind(m1, cbind(rep(5, 4), m2, m3, rep(13, 4)))
 lmat <- rbind(m1, m2)
@@
            expression(paste("Biodiversity Index X ", 10^-5)),
-           expression(paste("Truck hours X ", 10^-5)))#,
-           # "Intactness", expression(paste("FR converted (ha"^2, ")")))
+           expression(paste("Truck hours X ", 10^-5)))
+
 # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
@@
 yaxes <- list(seq(0, 16, 2), seq(0, 1200, 100), seq(0, 9, 1), 
-              seq(0, 35, 5))#,               
-              # seq(0, 1200, 100), #seq(0, 0.12, 0.02)), 
-              #seq(0, 1, 0.1), seq(0, 6500, 500))
-# bcols <- c(list(c(cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2], 
-#                   cols_tr[3])), lapply(1:7, function(x) rep("black", 4)))
+              seq(0, 35, 5))
 
@@
 legy <- grconvertY(0.5, from = "ndc", to = "user")
-legend(x = legx, y = legy, legend = "Protected Areas", pch = 15, col = "grey75", 
-       bty = "n", pt.cex = 3 * 0.6, cex = 1.5 * 0.6)
+legend(x = legx, y = legy, legend = "Protected Areas", pch = 15, col = "grey75",        bty = "n", pt.cex = 3 * 0.6, cex = 1.5 * 0.6)
 dev.off()
@@
 
-### Figures 2 & 3 (probably supplemental)
-Deprecated
+### Figure 2 - yield-carbon efficiency frontier
 ```{r, eval = FALSE}
-# individual maps
-# png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
-#      units = "in", res = 300)
-# par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
-# ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
-# legp <- c(TRUE, rep(FALSE, 3))
-# for(i in 1:4) {
-#   dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
-#   imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
-#             cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
-#             stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
-# }
-# dev.off()
-
-# heat maps
-# implist <- list(ycost$both, bdc$both)
-# load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
-# pacols <- c("transparent", "grey85", "grey75")
-# png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
-# par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
-# for(i in 1:2) {
-#   par(mar = c(1, 1, 1, 4))
-#   plot(zambia, col = "grey90", border = "transparent")
-#   ir <- implist[[i]]$heat
-#   ir[ir == 0] <- NA
-#   plot(ir, col = rev(heat.colors(20)), add = TRUE, )
-#   plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
-#        border = FALSE)
-#   plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
-#        border = FALSE)
-#   plot(roads, add = TRUE, lwd = 0.1, col = "grey")
-# }
-# dev.off()
+p <- data.frame(ybdc$both$params)
+pare <- ybdc$both$optitab
+p[, c("Y", "C", "BD" ,"COST")] <- round(p[, c("Y", "C", "BD" ,"COST")], 2)
+
+# Y versus C tradeoffs only
+yconly <- which(p[, "BD"] == 0 & p[, "COST"] == 0)
+
+# p[yconly, ]
+ycstats <- selected_impacts(yconly, fp(p_out, ybdc$both$bcode))
+ycstats <- ycstats[, .(iter, conv_area, tot_C)]
+
+# subset of sites to map
+yc_ss <- c(1, 2, 6, 11, 21)
+ycmins <- sapply(2:3, function(x) range(ycstats[[x]]))
+(ycmins[2, 2] - ycstats[yc_ss, get("tot_C")]) / diff(ycmins[, 2]) 
+round(ycstats[yc_ss, get("tot_C")] / ycmins[2, 2], 2)
+# 1.00 0.69 0.46 0.41 0.39
+round(ycstats[yc_ss, get("conv_area")] / ycmins[1, 1], 3)
+# 1.000 1.006 1.034 1.057 1.129
+
+# plot colors
+colsyc <- colorRampPalette(c("red", "gold", "blue"))
+colsyc <- colsyc(5)
+colsyc[4] <- "skyblue"
+
+png(fp(p_figures, "pareto_yc.png"), height = 7, width = 7, 
+    units = "in", res = 300)
+par(mar = c(6, 6, 3, 1))
+plot(ycstats$conv_area / yscale[1], ycstats$tot_C / yscale[2], axes = FALSE, 
+     xlab = "", ylab = "", pch = 16, col = "grey40", xlim = c(13.5, 15.5), 
+     ylim = c(425, 1200), xaxs = "i", yaxs = "i", cex = 1.5)
+points(ycstats$conv_area[yc_ss] / yscale[1], cex = 2,
+       ycstats$tot_C[yc_ss] / yscale[2], pch = 16, col = colsyc)
+# lines(ycstats$conv_area / yscale[1], ycstats$tot_C / yscale[2])
+axis(1, at = seq(13.5, 15.5, 0.25), mgp = c(3, 0.5, 0), tcl = -0.2, 
+     labels = sprintf("%.2f", seq(13.5, 15.5, 0.25)), las = 2, cex.axis = 1.5)
+ys <- seq(400, 1200, 50)
+# ys[-seq(2, 30, 3)]
+yslab <- ys
+# yslab[-seq(2, 32, 3)] <- ""
+axis(2, at = ys, labels = yslab, las = 2, mgp = c(3, 0.5, 0), tcl = -0.2, 
+     cex.axis = 1.5)
+mtext(ylabs[1], side = 1, line = 4.75, cex = 1.5)
+mtext(ylabs[2], side = 2, line = 3.5, cex = 1.5)
+dev.off()
+
+# maps for those sites, just for 0.25 increments
+fnames <- imp_sorter(fp(p_out, ybdc$both$bcode), 
+                     inds = yc_ss, fullnames = TRUE)
+yc_r <- lapply(fnames, function(x) {  # x <- pure[1]
+  DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
+  DTr <- dt_to_raster(DT, CRSobj)
+})
+yc_r2 <- lapply(yc_r, function(x) calc(x[[1:2]], sum))  # conv pixels
+names(yc_r2) <- paste0("YC", 1:5)
+
+# Plot maps
+# plot(1:5, pch = 16, col = colsyc(5))
+legtf <- rep(FALSE, 5)
+for(i in 1:5) { # i <- 1
+  par(bg = "transparent")
+  imp_plot2(yc_r2[[i]], impacts = i1, fpath = p_figures, 
+            fnm = paste0("ycplot_", i), crops = "", ftitle = "", 
+            col = colsyc[i], legend = legtf[i], hold = FALSE, stats = FALSE, 
+            cexscale = 0.4, marg = rep(0, 4))
+}
+
+# figure out overlaps between different pure objectives
+pms2a <- expand.grid(rep(list(0:1), 5))[-1, ] 
+pms2b <- pms2a[which(rowSums(pms2a) == 2 & pms2a$Var1 == 1), ]
+pms2b$cl <- c("YC1", "YC2", "YC3", "YC4")
+rownames(pms2b) <- 1:4
+
+## proximity statistics
+# pairwise overlaps
+yc_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
+  calc(stack(yc_r2[which(pms2b[x, ] == 1)]), sum)
+})
+names(yc_over) <- pms2b$cl
+
+# Jaccard Index of overlaps
+yc_over_a <- sapply(1:length(yc_over), function(x) { # x <- 4
+  sapply(1:2, function(y) {  # y <- 3
+    r <- yc_over[[x]] == y
+    round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
+  })
+})
+pms2b$JS <- round(yc_over_a[2, ] / colSums(yc_over_a) * 100, 2)
+
+# calculate nearest neighbor distance between each version, using spatstat 
+# nncross
+zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
+                 unitname = "meter")
+which1 <- function(x) x == 1  # selector function for r to xy
+rsover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 2
+  ind <- which(pms2b[x, 1:5] == 1)
+  rl <- yc_r2[ind]
+  rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
+  pps <- lapply(rxy, function(x) {
+    ppp(x = x[, 1], y = x[, 2], window = zam_owin)
+  })
+  # reorder to find which has most points, to put that in as X
+  reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
+  ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+})
+pms2b$NND <- round(sapply(rsover_nn, function(x) mean(x$dist))) / 1000
 
@@
 
-### Figures 4 (main) and 5 (probably supplemental)
+### Figure 3
 
@@
 cx <- 0.7
-# lcols <- c("black", "red", "blue")
-# imp_pct[, {
-#   png(fp(p_figures, "weight_impact.png"), height = 4, width = 4, 
-#       units = "in", res = 300)
-#   par(mgp = c(2, 0.75, 0), mar = c(3, 3, 1, 1))
-#   plot(conv_area, tot_C, type = "l", xlab = "", ylab = "", #xaxt = "n", 
-#   lines(conv_area, int_prior, col = "red")
-#   lines(conv_area, mu_cost, col = "blue")
-#   axis(1, at = seq(2, 12, 1), cex = cx, cex.axis = cx, line = -0.25, 
-#        tcl = -0.2, mgp = c(1, 0.25, 0))
-#   axis(2, at = seq(0, 60, 5), las = 2, cex = cx, cex.axis = cx, line = -0.25, 
-#        tcl = -0.15, mgp = c(1, 0.25, 0))
-#   mtext("% land savings", side = 1, line = 1, cex = cx)
-#   mtext("% Carbon/Biodiv/Cost savings", side = 2, line = 1, cex = cx)
-#   inms <- c("tot_C", "int_prior", "mu_cost")
-#   for(i in 1:length(inms)) {
-#     text(conv_area[-1], get(inms[i])[-1], labels = sctab[-1, i + 1], 
-#          cex = cx * 0.8, pos = 3, offset = 0.2, col = lcols[i])
-#     text(conv_area[1] + 0.1, i, labels = 0, cex = cx * 0.8, pos = 3, 
-#          offset = i * 0.15 - 0.2, col = lcols[i])
-#   }
-#   legend("bottomleft", legend = c("Carbon", "Biodiv", "Cost"), cex = cx,
-#          text.col = lcols, col = lcols, bty = "n", lty = 1)
 imp_pct[, {
@@
 
-# fnames <- fp(fp(p_out, ybdc$both$bcode), 
-#              imp_sorter(fp(p_out, ybdc$both$bcode), inds = pind))
-# scens_r <- lapply(fnames, function(x) {  # x <- pind[[1]]
-#   print(x)
-#   DT <- cbind(fread(x), basedt[, .(x, y, convertible)])
-#   DTr <- dt_to_raster(DT, CRSobj)
-# })
-# 
-# scens_rs2 <- lapply(scens_r, function(x) calc(stack(x[[1:2]]), sum))
-# names(scens_rs2) <- paste0("w", 1:length(scens_rs2))
-# 
-# pms2b <- do.call(rbind.data.frame, lapply(2:6, function(x) {
-#   a <- rep(0, 6)
-#   a[c(1, x)] <- 1
-#   a
-# }))
-# colnames(pms2b) <- paste0("Var", 1:ncol(pms2b))
-# pms2b$cl <- c("W010", "W020", "W030", "W040", "W050")
-# rownames(pms2b) <- 1:5
-# 
-# # pairwise overlaps
-# scens_over <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests
-#   calc(stack(scens_rs2[which(pms2b[x, ] == 1)]), sum)
-# })
-# names(scens_over) <- pms2b$cl
-# 
-# # Jaccard Index of overlaps
-# scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
-#   sapply(1:2, function(y) {  # y <- 3
-#     r <- scens_over[[x]] == y
-#     round(cellStats(r * pure_r[[1]]$convertible, sum), 1)
-#   })
-# })
-# pms2b$JS <- round(scens_over_a[2, ] / colSums(scens_over_a), 3)
-# 
-# # calculate nearest neighbor distance between each version, using spatstat 
-# # nncross
-# zam_owin <- owin(xrange = bbox(zambia)[1, ], yrange = bbox(zambia)[2, ],
-#                  unitname = "meter")
-# which1 <- function(x) x == 1  # selector function for r to xy
-# scenover_nn <- lapply(1:nrow(pms2b), function(x) {  # apply logic tests; x = 4
-#   print(x)
-#   ind <- which(pms2b[x, 1:6] == 1)
-#   rl <- scens_rs2[ind]
-#   rxy <- lapply(rl, function(x) rasterToPoints(x, fun = which1))
-#   pps <- lapply(rxy, function(x) {
-#     ppp(x = x[, 1], y = x[, 2], window = zam_owin)
-#   })
-#   # reorder to find which has most points, to put that in as X
-#   reord <- sort(sapply(pps, function(x) x$n), decreasing = TRUE)
-#   ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
-# })
-# pms2b$NND <- round(sapply(scenover_nn, function(x) mean(x$dist))) / 1000
-# 
-# cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.75, 0))
-# plot(1:4, cex = 3, pch = 20, col = cols_tr)
-# 
-# # plot(scens_rs2[[1]] + scens_rs2[[2]])
-# # check
-# # round(sapply(scens_rs2, function(x) {
-# #   cellStats(x * scens_r[[1]]$convertible * 100, sum)
-# # }))
-# # cstats$conv_area
-# 
-# # pms <- list(c(1, 2), c(1, 3), c(1, 4), c(1, 5))
-# # pms <- expand.grid(rep(list(0:1), 5))[-1, ]  # logic grid
-# 
-# # rasters showing overlap as weight is progressively put on C and BD
-# scens_over2 <- lapply(2:6, function(x) scens_rs2[[1]] + scens_rs2[[x]] * 2)
-# 
-# # # overlap areas
-# # scens_over_a <- sapply(1:length(scens_over), function(x) { # x <- 4
-# #   sapply(1:3, function(y) {  # y <- 3
-# #     r <- scens_over[[x]] == y
-# #     # plot(r)
-# #     round(cellStats(r * scens_r[[1]]$convertible, sum), 1)
-# #   })
-# # })
-# # colSums(scens_over_a[c(2, 3), ])
-# # over_cols <- sapply(seq(0, 0.4, 0.1), function(x) rgb(1 - x, 0, x))
-# # par(mar = c(0, 0, 0, 0))
-# # plot(1:3, pch = 20, col = rgb(1, 0, 1))
-# # plot(zambia)
-# 
-# # cols <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0.75, 0, 0.75))
-# # cols <- c("red", "blue", "gold")
-# bluecols <- sapply(c(1, 0.75, 0.5, 0.25, 0), function(x) rgb(0, x, 1))
-# bluecols[5] <- rgb(0.4, 0, 1)
-# cols <- sapply(1:length(bluecols), function(x) c("red", bluecols[x], "gold"))
-# plot(1:ncol(cols), pch = 20, col = cols[2, ], cex = 20)
-# ftits <- list("Y 90% C 5% Biod 5% Cost 0%",
-#               "Y 80% C 10% Biod 10% Cost 0%",
-#               "Y 70% C 15% Biod 15% Cost 0%",
-#               "Y 60% C 20% Biod 20% Cost 0%", 
-#               "Y 50% C 25% Biod 25% Cost 0%")
-# leg <- c("Y", "YCBd", "Y + YCBd")
-#               
-# tiff(fp(p_figures, "changing_weights2.tif"), height = 3.5, width = 7, 
-#      units = "in", res = 300)
-# par(mfrow = c(2, 3), oma = c(0, 0, 0, 1))
-# for(i in 1:5) { # i <- 1
-#   # map
-#   imp_plot2(scens_over2[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
-#             crops = leg, ftitle = ftits[[i]], col = cols[, i], 
-#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.4, 
-#             marg = c(0, 0, 1, 0), titline = 0)
-# }
-# mgp <- c(1, 0.25, 0)
-# par(mar = c(3, 3, 1, 0))
-# with(pms2b, {
-#      plot(NND, JS, col = cols[2, ], pch = 16, cex = 1.5, 
-#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.4), 
-#           xlim = c(0, 180))
-#      points(NND, JS, col = cols[[1]][1], pch = 20, cex = 1)
-#      axis(1, at = seq(0, 180, 20), cex.axis = 0.6, las = 2, mgp = mgp, 
-#           tcl = -0.2)
-#      axis(2, at = seq(0, 1, 0.1), cex.axis = 0.6, las = 2, mgp = mgp, 
-#           tcl = -0.2)
-#      mtext(text = "NN Distance (km)", side = 1, line = 1, cex = 0.45)
-#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.45)
-# })
-# dev.off()
-# 
-# 
-# #   # barplot
-# #   par(mar = c(5, 4, 3, 3))
-# #   overl <- scens_over_a[1:3, i][c(1, 3, 2)]
-# #   bp <- barplot(overl, las = 2, ylim = c(0, 12000),  
-# #                 col = cols[c(1, 3, 2)], mgp = c(1.5, 0.25, 0), tcl = -0.2, 
-# #                 ylab = expression(km^2), cex.axis = 0.6, cex.lab = 0.7)
-# #   axis(side = 1, at = bp, labels = leg, las = 2, 
-# #        mgp = c(1, 0.25, 0), tcl = -0.2, cex.axis = 0.6)
-# # }
-# # dev.off()
-#
-
 ```
@@
 
-### Figure 6
+### Figures 4 and 5
 
@@
 
-# do sum factoring in overlaps 
-# sapply(tores_over_r, function(x) {
-#   round(cellStats(x * scens_r[[1]]$convertible, sum), 1)
-# })
-
-# cols_tr <- c(rgb(1, 0, 0), rgb(1, 0.75, 0), rgb(0, 0, 1), rgb(0, 0.75, 1))
-# cols_tr <- c(rgb(1, 0, 0), rgb(0, 0, 1), rgb(0, 0.75, 1), rgb(1, 0.5, 0), 
-#              "gold")#rgb(1, 1, 0))
-# plot(1:length(cols_tr), cex = 3, pch = 20, col = cols_tr)
-
-# layout matrix
-# m1 <- do.call(cbind, lapply(1:5, function(x) matrix(rep(x, 45), 6, 9)))
-# m2 <- matrix(rep(6, 36), 4, 9)
-# m3 <- do.call(cbind, lapply(7:13, function(x) matrix(rep(x, 20), 4, 5)))
-# f1 <- lapply(14:15, function(x) cbind(rep(x, 4), rep(x, 4)))
-# lmat <- rbind(m1, cbind(m2, m3, rep(14, 4)))
-# lmat <- rbind(m1, m3)
-# cbind(m2, m3)
-
-# # plot(1:10, ylab = ylabs[2])
-# # ylabs <- c("Jaccard Index", 
-# ylabs <- c(expression(paste("Conv. area (", 10^5, " ", km^2, ")")), 
-#            "Hours", "C loss (Mt)", "Priority", "Intactness/priority",
-#            "Intactness", expression(paste("FR converted (ha"^2, ")")))
-# # yscale <- c(1, 10^5, 1, 100000, 1, 1, 1, 1)
-# yscale <- c(10^5, 1, 100000, 1, 1, 1, 1)
-
 # # figure titles
@@
 # leg <- c("Converted")
-
-# # yaxes <- list(seq(0, 0.8, 0.1), seq(0, 14, 2), seq(0, 2.5, 0.5), 
-# maxes <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][2])
-# mins <- sapply(c(2, 5, 3, 7:9, 4), function(j) minmaxstats[[j]][1])
-# ymaxes <- maxes / yscale
-# ymins <- mins / yscale
-# yaxes <- list(seq(0, 15, 3), 0:7, seq(0, 1100, 100), seq(0, 0.15, 0.05), 
-#               seq(0, 0.6, 0.1), seq(0, 1, 0.1), seq(0, 7000, 1000))
-#
 
@@
                   })), ymaxes)
-
-# # bcols <- c(list(c(cols_tr[1], cols_tr[1], cols_tr[2], cols_tr[1], cols_tr[2],
-# pms2b$icol <- cols_tr[c(1, 1, 2, 1, 2, 3, 1, 2, 3, 4)]  # circle inner
-# pms2b$ocol <- cols_tr[c(2, 3, 3, 4, 4, 4, 5, 5, 5, 5)]  # circle outer
-# 
-# tiff(fp(p_figures, "objective-subjective_weights.tif"), height = 2.5, width = 7,      units = "in", res = 300)
-# l <- layout(mat = lmat)
-# layout.show(l)
-
-# # map plots
-# for(i in 1:length(toresr)) { # i <- 1
-#   # map
-# #   par(mar = rep(0, 4))
-# #   plot(zambia, col = cols_tr[i])
-#   imp_plot2(toresr[[i]], impacts = i1, fpath = p_figures, fnm = "bob",
-#             crops = leg, ftitle = ftits[[i]], col = cols_tr[[i]], 
-#             legend = TRUE, hold = TRUE, stats = FALSE, cexscale = 0.3, 
-#             marg = rep(0, 4), inset = c(0.1, 0.1, 0, 0))
-# }  
-# 
-# # impact plots
-# mgp <- c(0.9, 0.25, 0)
-# par(mar = c(3, 3, 0, 3))
-# with(pms2b, {
-#      plot(NND, JS, col = pms2b$icol, pch = 16, cex = 0.8, 
-#           axes = FALSE, xlab = "", ylab = "", ylim = c(0, 0.6), 
-#           xlim = c(0, 25))
-#      points(NND, JS, col = pms2b$ocol, pch = 20, cex = 0.5)
-#      axis(1, at = seq(0, 25, 5), cex.axis = 0.5, las = 2, mgp = mgp, 
-#           tcl = -0.2)
-#      axis(2, at = seq(0, 0.6, 0.1), cex.axis = 0.5, las = 2, mgp = mgp, 
-#           tcl = -0.2)
-#      mtext(text = "NN Distance (km)", side = 1, line = 0.5, cex = 0.4)
-#      mtext(text = "Jaccard Index", side = 2, line = 1, cex = 0.4)
-# })
-# par(mar = c(3, 3, 0, 0))
-# mgp <- c(1.3, 0.25, 0)
-# for(i in 1:7) { # i <- 1
-#   v <- impsr[[i]] / yscale[i]
-#   b <- barplot(impsr[[i]] / yscale[i], las = 2,  col = cols_tr, 
-#                mgp = mgp, tcl = -0.2, yaxt = "n", 
-#                ylim = range(yaxes[[i]]), ylab = ylabs[i], cex.axis = 0.4,
-#                cex.lab = 0.55)
-#   lines(c(-1, 7), rep(ymins[i], 2), lwd = 1, col = "grey50")
-#   lines(c(-1, 7), rep(ymaxes[i], 2), lwd = 1)
-#   axis(2, at = yaxes[[i]], cex.axis = 0.5, las = 2, mgp = mgp, 
-#        tcl = -0.2)
-# }
-# dev.off()
 
@@
 
+## Supplemental figures
 ### Overlap with Farm Blocks
@@
 })
-fb_sim <- fb_over_a[3, ] / colSums(fb_over_a)
+fb_sim <- fb_over_a[3, ] / colSums(fb_over_a) * 100
 
@@
 
+```{r, eval = FALSE, echo=FALSE}
 ### Deprecated version of Figure 3
-
-```{r, eval = FALSE}
 # objective weights
@@
 
-<a href="#top">Back to top</a>
+```{r, eval = FALSE, echo = FALSE}
+### Figures 2 & 3 (probably supplemental)
+# Deprecated
+
+# individual maps
+# png(full_path(p_figures, "all-4-maps.png"), height = 7, width = 7, 
+#      units = "in", res = 300)
+# par(mfrow = c(2, 2))#, oma = c(0, 0, 0, 2))
+# ftitles <- c("Yield", "Carbon", "Biodiversity", "Cost")
+# legp <- c(TRUE, rep(FALSE, 3))
+# for(i in 1:4) {
+#   dtr <- calc(pure_r[[i]][[1:2]] * c(1, 2), sum)
+#   imp_plot2(dtr, impacts = i1, fpath = p_figures, fnm = "all-4-maps", 
+#             cexscale = 0.6, ftitle = ftitles[i], crops = c("maize", "soy"), 
+#             stats = FALSE, hold = TRUE, legend = legp[i], marg = c(0, 0, 2, 0))
+# }
+# dev.off()
+
+# heat maps
+# implist <- list(ycost$both, bdc$both)
+# load(full_path(set_base_path(), paste0("external/data/ZA/parks_roads.rda")))
+# pacols <- c("transparent", "grey85", "grey75")
+# png(full_path(p_figures, "ycost-bdc-heat.png"), height = 400, width = 900)
+# par(mfrow = c(1, 2), oma = c(0, 0, 0, 2))
+# for(i in 1:2) {
+#   par(mar = c(1, 1, 1, 4))
+#   plot(zambia, col = "grey90", border = "transparent")
+#   ir <- implist[[i]]$heat
+#   ir[ir == 0] <- NA
+#   plot(ir, col = rev(heat.colors(20)), add = TRUE, )
+#   plot(pas[pas@data$type == "gma", ], add = TRUE, col = pacols[2],
+#        border = FALSE)
+#   plot(pas[pas@data$type == "npark", ], add = TRUE, col = pacols[3],
+#        border = FALSE)
+#   plot(roads, add = TRUE, lwd = 0.1, col = "grey")
+# }
+# dev.off()
+
+```

### 2016-05-05 Rmd/main-analysis.Rmd
Full paper revisions: main manuscript, fully revised, with edits and additions to SI methods and results.  New figure created (figures.Rmd), slight edit of loose code on main-analysis.Rmd

@@
 # Separate soy and maize impacts
+
+# together
 cb <- c("Y" = 1, "C" = 0, "BD" = 0, "COST" = 0)
@@
 
-CRSobj <- CRS(mzto$inputs$sp$crs)
+CRSobj <- CRS(mzsoyto$inputs$sp$crs)
 mzsoytor <- dt_to_raster(mzsoyto$conv, CRSobj)  # converges on same results
@@
 # Plot
-i1 <- rbind(mzto$impacts, soyto$impacts)
+# i1 <- rbind(mzto$impacts, soyto$impacts)
 basedt <- mzsoyto$inputs$mask

### 2016-05-05 paper/main.tex
Full paper revisions: main manuscript, fully revised, with edits and additions to SI methods and results.  New figure created (figures.Rmd), slight edit of loose code on main-analysis.Rmd

@@
 Their carbon contents and biodiversity values may be completely ignored on the grounds that they are not forests \citep{searchinger_high_2015}.
-Yet, the wetter woodland-savannas and shrublands of Africa have a similar average richness, although not density, for birds and mammals as the world’s wetter tropical forests, and 75\% of these habitats have higher overall vertebrate diversity than 75\% of the rest of the world \citep{searchinger_high_2015}.
+Yet, the wetter woodland-savannas and shrublands of Africa have a similar average richness, although not density, for birds and mammals as the world's wetter tropical forests, and 75\% of these habitats have higher overall vertebrate diversity than 75\% of the rest of the world \citep{searchinger_high_2015}.
 Their carbon contents are also substantial, particularly relative to their likely yields, so that their use for food crops is unlikely to result in fewer carbon emissions per ton of crop \citep{searchinger_high_2015}.
-Conversion of over 100 million hectares of these lands would release tens of millions of tons of carbon.
+Conversion of over 100 million hectares of these lands would release tens of millions of tons of carbon \citep{searchinger_high_2015}.
 
-The high likelihood of some cropland expansion in sub-Saharan Africa raises the prospect of whether this expansion could be located in ways that allow agricultural potential to be achieved, but for substantially lower loss of carbon and biodiversity than conventional agricultural development pathways.
-Analyzing this question must be done at national or sub-national scales, because that is where key decisions are made, and it requires data that is both in resolution and quality to provide confidence for decision-makers.
+The high likelihood of some cropland expansion in sub-Saharan Africa raises the prospect of whether this expansion could be located in ways that meet agricultural production goals, but for substantially lower loss of carbon and biodiversity than conventional agricultural development pathways.
+Analyzing this question must be done at national or sub-national scales, because that is where key decisions are made, and accurate results require data that are high in both resolution and quality.
 
@@
 On the one hand, Zambia has great need to boost its food production.
-Sixty percent of Zambia's population is rural {\citep{the_world_bank_world_2016}, per capita calorie food availability is just 2100 kcal \citep{fao_faostat:_2016}, and child stunting rates are 40\% \citep{haddad_global_2015}.
-According to the medium estimate of the United Nations, Zambia’s population is projected to grow 265\% from 16.2 million in 2015 to 43 million in 2050 \citep{united_nations_world_2015}, which will combine with rapid economic growth \citep[$>$6\%,][]{the_world_bank_world_2016} to substantially increase per capita calorie demands \citep{tilman_global_2011}.
+Sixty percent of Zambia's population is rural {\citep{the_world_bank_world_2016}, per capita calorie food availability is just 2100 kcal \citep{fao_faostat_2016}, and child stunting rates are 40\% \citep{haddad_global_2015}.
+According to the medium estimate of the United Nations, Zambia's population is projected to grow 265\% from 16.2 million in 2015 to 43 million in 2050 \citep{united_nations_world_2015}, which will combine with rapid economic growth \citep[$>$6\%,][]{the_world_bank_world_2016} to substantially increase per capita calorie demands \citep{tilman_global_2011}.
 Zambia has also been expanding its agricultural exports, and has agronomic and situational characteristics that put it in a strong position to continue to do so to other parts of the region.
 
-At the same time, two thirds of Zambia’s 738,400 km$^2$ land area is dominated by woodland savannas, which have sufficiently dense tree cover to be classified as forest under either of the United Nations Framework Convention on Climate Change thresholds ($>$10 or $>$30\% tree cover) for defining this vegetation type \citep{sexton_conservation_2015}.
-These ecosystems store substantial amounts of carbon, with 50\% of Zambia having a carbon density $\geq$130 t ha$^{-1}$ (vegetative carbon and soil carbon in the top 1 m), while an additional 40\% exceeds 80 t ha$^{-1}$.
-Meanwhile, Zambia’s deforestation rate of 2,500--3,000 km$^2$ year$^{-1}$ is one of the highest in the world \citep{vinya_preliminary_2011}.
+At the same time, two thirds of Zambia's 738,400 km$^2$ land area is dominated by woodland savannas, which have at sufficiently dense tree cover to be classified as forest under either of the definitions ($>$10 or $>$30\% tree cover) used by the United Nations Framework Convention on Climate Change \citep{sexton_conservation_2015}.
+These ecosystems store substantial amounts of carbon, with 50\% of Zambia having a carbon density of at least 130 t ha$^{-1}$ (vegetative carbon and soil carbon in the top 1 m), while an additional 40\% exceeds 80 t ha$^{-1}$.
+Meanwhile, Zambia's deforestation rate of 2,500--3,000 km$^2$ year$^{-1}$ is one of the highest in the world \citep{vinya_preliminary_2011}.
 
 The biodiversity of these woody savannas is also high.
-Most of Zambia is covered by what the World Wildlife Fund has categorized as Central Zambian Miombo Woodland \citep{olson_terrestrial_2001,hogan_central_2004}.
-It contains mature miombo woodland trees typically 15 to 20 meters tall, around a quarter of which are evergreen, with a broadleaf shrub and grass understory beneath, and is interspersed with wetlands that comprise up to 30\% of the landscape \citep{hogan_central_2004}.
-In addition to hosting one of the world's three richest mammal assemblages \cite{olson_terrestrial_2001}, the large expanses of relatively intact habitat make these woodlands, and several of Zambia's eight other ecoregions \citep{fund_ecoregions_2011, estes_zambezian_????}, important centers of large mammal populations, including black rhino, African buffalo, and highly migratory or wide-ranging species, such as elephants and major predators \citep{hogan_central_2004}.
-Zambia in general has a high bird diversity \citep{searchinger_high_2015}.
-Plant diversity in much of the country ranges between 3,000-5,000 species, a richness that is exceeded only in tropical forests \citep{kier_global_2005}.
-Although there are relatively few endemic bird and mammal species, reptile and amphibian endemism is high, with 19 endemic reptile species and 13 endemic amphibians \citep{fund_ecoregions_2011}.
+Most of Zambia is covered by what the World Wildlife Fund has categorized as the Central Zambian Miombo Woodland ecoregion \citep{olson_terrestrial_2001,hogan_central_2004}, which contains some 3800 plant species, making it the 17th richest ecoregion in the world in floral diversity (out of 867) \citep{kier_global_2005}.
+Vegetation is characterized by a mix of tall, frequently evergreen, trees (15-20 m), broadleaf shrubs, and grasses, and is heavily interspersed with wetlands \citep{hogan_central_2004}.
+The large expanses of relatively intact habitat give this woodland the third highest richness for mammal species of the worlds ecoregions  \citep{olson_terrestrial_2001,fund_ecoregions_2011, estes_zambezian_????}.
+It supports rare species including the  black rhino,  major predators, and highly migratory, wide-ranging species, such as elephants \citep{hogan_central_2004}.
+Zambia overall also has a high bird diversity, with 753 recorded species \citep{dowsett_birds_2008}.
+Although few bird and mammal species are endemic, reptile and amphibian endemism is high, with 19 endemic reptile species and 13 endemic amphibians \citep{fund_ecoregions_2011}.
 
-There is a long history of conservation planning that attempts to target conservation areas or objectives while minimizing economic costs or within a constrained budget of land area for conservation \cite[e.g.][]{koh_spatially_2010, naidoo_integrating_2006,margules_systematic_2000}.
-Koh and Ghazoul \citep{koh_spatially_2010} provide one example of an effort to examine trade-offs among production, carbon and biodiversity for oil palm development in Indonesia by giving each land use priority equal weighting.
-One of the challenges for land use planning in practice is that decisions involve trade-offs among different objectives and values that most people find difficult to weight against each other \emph{ex ante}, because the values are often abstract, making them harder to evaluate without first viewing the consequential trade-offs.
-Here we present a new model in which varying weights can be placed on the competing land use objectives.
-This set of options then allows model users to choose the most appealing combination of weights based on their land use preferences, and their own assessments of what are and are not acceptable costs within those objectives.
-The objective of this is to assist different stakeholders in finding compromises between competing land use interests while still meeting agricultural production targets.
-We apply this model to Zambia so that we can explore its strengths and limitations as a decision-making tool, and also to answer a basic but important question: is it possible to achieve a reasonable balance of different interests by reducing the carbon and biodiversity costs of expansion while still focusing on land use ares with high agricultural potential?
+There is a long history of using models to achieve conservation objectives in the face of cost and other constraints \cite[e.g.][]{naidoo_integrating_2006,margules_systematic_2000}.
+MARXAN, perhaps the most widely used of these, is a decision-support program that is designed to efficiently select conservation area networks \citep{ball_marxan_2009,watts_marxan_2009}.
+Within the more general land use planning field, an even broader range of models exists for finding optimal spatial configurations that satisfy multiple land use objectives (see Cao et al.
+\citep{cao_spatial_2011} for an example and review of alternatives).
+In the present case, the need is not to select protected areas but to identify areas for cropland expansion that meet the objectives of agricultural development and environmental conservation.
+Koh and Ghazoul \citep{koh_spatially_2010} developed one model for examining trade-offs among oil palm development, rice production, and carbon and biodiversity production in Indonesia, and a related approach was used to identify sustainable agricultural intensification solutions in southern Tanzania \citep{nijbroek_regional_2016}.
+In both examples, the model assigned equal weight to each land use objective in the compromise scenarios that were analyzed.
+However, one of the great challenges for land use planning is that different objectives can be difficult to weigh against each other \emph{ex ante}, because the values are often abstract, making them harder to evaluate without first viewing the consequential trade-offs \citep{cao_spatial_2011,nijbroek_regional_2016}.
+Governments and other stakeholders also tend to have varying, and often conflicting, priorities, and their willingness to compromise may depend on the degree to which their priorities are impacted.
 
+To overcome these problems, we developed a new, publicly available model that allows varying levels of compromise between competing agricultural and environmental interests to be evaluated while still meeting agricultural production targets.
+The model adopts different methods than the two aforementioned models \citep{koh_spatially_2010,nijbroek_regional_2016} to estimate agricultural potential and conservation objectives, and can allocate land for multiple crops simultaneously.
+More fundamentally, our model is based on linear programming, which allows users to place varying weights on the different land use objectives \citep{cao_spatial_2011}.
+This feature allows different stakeholders to work together to choose the most appealing combination of weights based on their land use preferences, and their own assessments of what are and are not acceptable costs within those objectives.
+
+We apply this model to Zambia so that we can explore its strengths and limitations as a decision-making tool, and also to answer a basic but important question: is it possible to achieve a reasonable balance of different interests by reducing the carbon and biodiversity costs of expansion while still focusing on land use areas with high agricultural potential?
 Answering this question is becoming increasingly important as savannas and other grassy biomes receive increasing agricultural development pressure, and may become increasingly targeted as an indirect result of the laudable efforts to conserve tropical forests \citep{searchinger_high_2015}.
@@
 The model we developed seeks to find optimal land use configurations that satisfy the production targets for multiple crops while minimizing four costs: the total land area required, transportation costs, carbon released from land conversion, and impacts on biodiversity.
-Although total land area is treated as a cost, this cost is minimized by selecting areas with the highest likely yields.
-By minimizing total land required as well as transportation costs, the model finds areas that reflect two important indicators of agricultural potential.
-By minimizing carbon and biodiversity costs, the model finds areas that minimize two important dimensions of environmental impact.
+Although total land area is treated as a cost, minimizing this cost identifies lands with the highest likely yields.
+By minimizing total land required as well as transportation costs, the model therefore finds areas that reflect two important indicators of agricultural potential.
+By minimizing carbon and biodiversity costs, the model also finds areas that minimize two important dimensions of environmental impact.
 The model chooses final areas depending on the weights that a user assigns to each of these four objectives.
 
-To enable this functionality, and to identify optimal solutions, the model is structured to evaluate each objective in terms of production efficiency, or the ratio of cost per hectare to crop production per hectare (yield) that results from converting land to agriculture.
-For land area costs, output is measured in terms of the number of km$^2$ required to meet the crop production target.
-For the other objectives, the indicators are transportation costs, carbon loss, or biodiversity impacts per ton of crop yield.
+To enable this functionality, and to identify optimal solutions, the model is structured to evaluate each objective in terms of production efficiency, or the ratio of each cost per hectare to crop production per hectare (yield) that results from converting land to agriculture.
+For land area, cost is measured as the number of km$^2$ required to meet the crop production target.
+For the other objectives, the costs indicators are transportation time, carbon loss, or biodiversity impacts.
 
 We focus our model initially on two crops, maize and soybeans, because of their existing and potential significance for Zambia's agricultural sector, and their likely dominance of in the future growth of agriculture in Africa's broader savanna regions \citep{searchinger_high_2015}.
-Maize is the largest crop by area and consumption in Zambia \citep{fao_faostat:_2016}, while soybeans are expanding rapidly as a commodity crop to meet regional and global demands for animal feed and direct human consumption \citep{gasparri_emerging_2015}.
-Because the model is designed to meet production targets for multiple crops, we apply algorithms for allocating each crop to its highest value cells.
+Maize is the largest crop by area and consumption in Zambia \citep{fao_faostat_2016}, while soybeans are expanding rapidly as a commodity crop to meet regional and global demands for animal feed and direct human consumption \citep{gasparri_emerging_2015}.
+Because the model is designed to meet production targets for multiple crops, we apply algorithms for allocating each crop to its highest value cells, given the particular combination of objectives being assessed.
 
@@
 In the second step, we used an empirical model to map the DSSAT-modeled yields onto a 1 km grid, taking advantage of a finer resolution soil map that provides a common frame for estimating both yield potential and soil carbon stocks (see below), but has insufficient detail in several parameters (particularly effective rooting depth) required by DSSAT to provide a sound empirical basis for running the model directly.
-We used a generalized additive model \citep{wood_mgcv:_2001} to predict the values of the DSSAT-simulated yields, using best-fitting subsets of the following weather and soil variables: growing season precipitation; growing degree days; mean growing season shortwave solar radiation; percent organic carbon; pH; percent clay content.
+We used a generalized additive model \citep{wood_mgcv_2001} to predict the values of the DSSAT-simulated yields, using best-fitting subsets of the following weather and soil variables: growing season precipitation; growing degree days; mean growing season shortwave solar radiation; percent organic carbon; pH; percent clay content.
 The soil predictors represented mean values in the top 1 m of soil, and were extracted from a new 1 km global database of soil properties \cite{hengl_soilgrids1km_2014}.
@@
 These yield values therefore overestimate what farmers are truly likely to achieve.
-To reflect more realistic estimates of future yields, we therefore rescaled the yield maps to match the mean yields that are projected to be achieved in Zambia by 2050 (4.4 t ha$^{-1}$ for maize and 3.6 t ha$^{-1}$ for soybean).
-We projected these yields by using FAO-projections for sub-Saharan Africa as a whole to derive an annual growth factor and applying that factor to the average 2009-2014 yields for each crop reported in Zambia.
+To reflect more realistic estimates of future yields, we therefore rescaled the yield maps to match FAO-projected Zambian yields in 2050 (4.4 t ha$^{-1}$ for maize and 3.6 t ha$^{-1}$ for soybean).
+We used the FAO-projections for sub-Saharan Africa as a whole to derive an annual growth factor and applied that factor to the average 2009-2014 yields for each crop in Zambia.
 Overall, the distribution of yields derived from our crop modeling is most important to targeting land use, while this choice of average yield determines the total quantity of land needed to meet production targets.
@@
 
-As a second measure related to agricultural potential, we estimated the travel time to the nearest major town.
-Travel time serves here as a proxy for differences in output prices and production costs, such as fertilizer prices and transport costs \cite{stifel_isolation_2008}, and, via the density of road networks, is strongly correlated with the suitability of land for farming \citep{laurance_estimating_2015,estes_using_2014}.
-We used a cost-distance analysis drawing on spatial data for roads, rivers and water bodies, urban centers inside and neighboring Zambia, and national boundaries to calculate travel time in hours.
-Total travel costs for any land use scenario equal the total time needed to transport the tonnage of crops produced in each converted cell to the nearest town, which we divide by 20 to produce "truck hours" on the assumption of a 20-ton truck (although the division  of all cell results by this constant does not alter the mapping outcomes).
+As a second measure related to agricultural potential, we estimated the travel time to the nearest major market town.
+Travel time serves here as a proxy for differences in output prices and production costs, such as fertilizer and transport costs \cite{stifel_isolation_2008}, and, via the density of road networks, is strongly correlated with the suitability of land for farming \citep{laurance_estimating_2015,estes_using_2014}.
+For this analysis, we defined market towns as the administrative capitals of Zambia's districts, as well as any town outside of Zambia having a population of $\geq$10,000 within a region bounded by 19.5$^{\circ}$ to 34.5$^{\circ}$ longitude and -19.5$^{\circ}$ to -8$^{\circ}$ latitude.
+These neighboring towns were included because they may be more easily accessible to Zambians living in border regions than the nearest district capital.
+
+We calculated travel time in hours using cost-distance analysis drawing on several spatial datasets.
+Vector data for Zambian roads obtained from the Government of Zambia (GOZ) and for the broader region from OpenStreetMap\footnote{http://openstreetmap.org} were merged to create a single dataset of regional roads that distinguished between trunk, primary, secondary, and tertiary roads.
+For market towns, we obtained district capital towns locations from the GOZ and for towns outside Zambia we used data from the Gridded Population of the World Version 3 \citep{ciesin_gridded_2005} database.
+For waterbodies, we used the USGS Hydrosheds database \citep{lehner_new_2008} to indicate the position and size of rivers, based on their contributing area, and used a lakes vector provided by the GOZ.
+The roads, rivers, and lakes data, as well as Zambia's border, were converted to 1 km grids wherein each feature was assigned a travel time value based on how long it would take to traverse 1 km.
+The resulting grids were merged into a single ``friction'' surface, which was used to compute the number of travel hours along the fastest route to the nearest market town (see online electronic supplement for further details).
+
+Total travel costs for any land use scenario equal the total time needed to transport the tonnage of crops produced in each converted cell to the nearest town, We divided these costs by 20 to produce "truck hours" on the assumption of a 20-ton truck.
+Doing so produces a more intuitive unit, but does not alter mapping outcomes.
 
@@
 Vegetative carbon was calculated from a recent map of above-ground biomass developed by Baccini et al.
-\citep{baccini_estimated_2012}, from which we estimated below-ground biomass using savanna and miombo woodland-specific root-shoot ratios \cite{mokany_critical_2006, mugasha_allometric_2013}, and then converted the sum to carbon using a ratio of 0.47 \citep{ipcc_land_1996}.
-We used the same soil carbon dataset used for crop yield prediction together with an accompanying map of bulk density to calculate soil carbon stocks in the top 1 m.
-The potential carbon loss due to agricultural conversion was calculated as 100\% of vegetative carbon and 25\% of soil carbon \cite{searchinger_high_2015}.
+\citep{baccini_estimated_2012}.
+We estimated below-ground biomass using savanna and miombo woodland-specific root-shoot ratios \cite{mokany_critical_2006, mugasha_allometric_2013}, and then converted the total biomass to carbon using a ratio of 0.47 \citep{ipcc_land_1996}.
+The same soil carbon dataset used for crop yield prediction, together with an accompanying map of bulk density, was used to calculate soil carbon stocks in the top 1 m.
+We calculated the potential carbon loss due to agricultural conversion was calculated as 100\% of vegetative carbon and 25\% of soil carbon \cite{searchinger_high_2015}.
 
@@
 
-There are many ways of evaluating biodiversity value and potential impacts, including species richness of different taxa in absolute terms or based on level of threat or endemism, and including factors designed to address the significance of contiguity or the impact of different types of development \citep[e.g.][]{newbold_global_2015,searchinger_high_2015, gasparri_emerging_2015,koh_spatially_2010}.
+There are many ways of evaluating biodiversity value and potential impacts, including species richness of different taxa in absolute terms or based on level of threat or endemism \citep[e.g.][]{searchinger_high_2015, gasparri_emerging_2015}.
+Other possibilities include factors designed to address the significance of contiguity or the impact of different types of development on diversity \citep[e.g.][]{newbold_global_2015,koh_spatially_2010}.
 The actual measures selected should reflect both conservation priorities and the quality of available data.
-For a fine-scaled, country-level analysis such as this one, the species range map data that form the basis for broader impact assessments \citep[e.g.][]{kehoe_global_2015} have an effective resolution that is too coarse.
+For a fine-scaled, country-level analysis such as this one, the species range map data that form the basis for many broader impact assessments \citep[e.g.][]{kehoe_global_2015} have an effective resolution that is too coarse.
 We therefore developed a composite measure of biodiversity value based on several datasets.
 
-First, we used the potential vegetation maps of East Africa \citep[][]{van_breugel_potential_2011}, a 2010 landcover dataset for Zambia\footnote{downloaded from http://apps.rcmrd.org/landcoverviewer/}, and a shapefile of protected areas to calculate how much of Zambia's 20 unique vegetation types remains after conversion to croplands and settlements, and what proportion of that remainder falls within national parks.
-From this estimate we calculated a hybrid measure of rarity and threat, wherein we assigned a weight to each vegetation type based on its relative remaining area in proportion to Zambia's extent, log-transformed to account for a highly skewed distribution.
-We then re-weighted that weight by the protected fraction of each vegetation type's protected fraction, following \citep{van_breugel_environmental_2015}.
-We then used the remaining vegetation map to calculate a second measure, the proportion of untransformed vegetation within a 11X11 km neighborhood centered on each 1 km$^2$, providing a measure of vegetation intactness.
-By adding this measure to the first index and normalizing, we created a biodiversity score (range 0-1 per ha) that gives equal weight to how rare or threatened a habitat is, and how undisturbed its local area.
-When the biodiversity score for each converted grid cell converted is summed, it generates a total cost of biodiversity impacts.
+First, we used the potential vegetation maps of East Africa \citep[][]{van_breugel_potential_2011}, a 2010 landcover dataset for Zambia\footnote{downloaded from http://apps.rcmrd.org/landcoverviewer/}, and a shapefile of protected areas to calculate how much of Zambia's 20 unique vegetation types remained after conversion to croplands and settlements as of 2010, and what proportion of that remainder fell within national parks.
+We used these values to calculate an index of rarity and threat.
+We calculated rarity as the proportion of Zambia occupied by each vegetation type's remaining area, which we then log-transformed to account for the highly skewed distribution.
+For threat, we calculated the proportion of each vegetation type falling within protected areas, and then multiplied the threat and rarity values to create the index \citep{van_breugel_environmental_2015}.
+We then used the remaining vegetation map to calculate a second measure, the proportion of untransformed vegetation within a 11x11 km neighborhood centered on each 1 km$^2$.
+By adding this measure to the first index and normalizing, we created a biodiversity score with a range of 0-1 per ha, which gives equal weight to a habitat based on a) how rare or threatened and b) how undisturbed it is.
+Summing the biodiversity score for each converted grid cell generates a total cost of biodiversity impacts for each scenario.
 
 As a final modification, we masked out the areas of existing national parks and game management areas to respect the national judgment that they should not be used for expanding farmland.
-We retained a third protected category, forest reserves, as these tend to be far less stringently protected, having already had 8.3\% of their area converted to cropland or settlements, but assigned these areas a high biodiversity score.
+We retained a third protected category, forest reserves, as these tend to be far less stringently protected, having already lost 8.3\% of their area to cropland or settlements, but assigned these areas a high biodiversity score (0.927, the inverse of their converted proportion).
 
@@
 For this paper, we adopted targets of a four-fold increase for maize for 2050 compared to 2009-2014 average, and a 10-fold increase for soybeans.
-These increases reflect the rate of growth in Zambia from 2000-2014, and a mean production trend for soybean growth in southern Africa estimated by Gasparri et al \citep{gasparri_emerging_2015}, with small additional adjustments to to account for the fact that Zambia has one of the highest agricultural potentials in sub-Saharan Africa and has high potential to expand exports.
+These increases reflect the rate of growth in Zambia from 2000-2014, and a mean production trend for soybean growth in southern Africa estimated by Gasparri et al \citep{gasparri_emerging_2015}, with small additional adjustments to account for the fact that Zambia has one of the highest agricultural potentials in sub-Saharan Africa and has high potential to expand exports.
 
@@
 For the principal scenarios we present in this paper, we assume that yields on current cropland will achieve the average likely yields for Zambia's \emph{potential} farmland.
-However, to examine the possible significance of yield gains on existing cropland, we also examined scenarios in which the yield gains on existing cropland 1) only manage to close 50\% of the gap between current and our projected future yields, and 2) close 150\% of the gap.
+However, to examine the possible significance of yield gains on existing cropland, we also examined scenarios in which the yield gains on existing cropland 1) only manage to close 50\% of the gap between current and our projected future yields, and 2) exceed the gap by 50\% (i.e.
+achieve 150\% of the expected yield gain).
 
@@
 
-Weighting different objectives is to some extent based purely on preferences, but it is important that these be expressed in a numerical form and processed so that a quantitatively optimal solution for that set of preferences is identified.
-To achieve this result, we first express the different objectives as efficiencies (e.g.
-carbon loss per ton of crop yield, as described above), and then normalize the units of each objective based on the range of its scores.
-We then multiplying each normalized objective by its corresponding weight (which fall between 0 and 1 and must sum to 1, with values closer to 1 indicating stronger preferences) and then adding the resulting products.
-The result is a single score that ranks each grid cell according to its ability to meet the production target for the lowest cost, given those weights.
-The model then selects cells in descending rank order until each the cumulative production of converted cells reaches each crop's corresponding target (see online electronic supplement for details of land allocation between crops).
+Weighting different objectives to some extent depends purely on preferences, but it is important that these preferences be expressed in a numerical form that reasonably reflects them so that a quantitative solution can be computed.
+To achieve this result, we first express the different objectives as efficiencies, e.g.
+carbon loss per ton of crop yield, and then normalize the units of each objective based on the range of its scores.
+Weights representing the percentage of importance are placed on each objective, expressed as a decimal between 0 and 1, with values closer to 1 indicating stronger preferences, and with all weights summing to 1.
+The final score is the sum of the resulting products for each objective for each grid cell, which is used to rank each grid cell according to its ability to meet the production target most for the lowest total cost, given those weights.
+The model then selects cells in descending rank order until the cumulative production of converted cells reaches each crop's corresponding production target.
 
-Although this weighting system treats choices as preferences, our model calculates the actual costs of each objective in absolute terms:  area of land converted (which is increasingly minimized as weight is added to the preference to maximize crop yields), total transportation costs from new croplands, total carbon released from conversions, and total impacts on biodiversity.
+Although this weighting system treats choices as relative preferences, our model calculates the actual costs of each objective in absolute terms:  total area converted, total transportation costs from new croplands, total carbon released from land conversions, and total impacts on biodiversity.
 After reviewing these costs, model users can then adjust the preferences to reflect desirable or politically acceptable trade-offs.
@@
 
-In our main scenario, we assumed that the gap between current and projected future yields would be fully closed on existing cropland, but we also ran the model (with equal weights across all objectives) for cases when only 50\% or 150\% of the gap was closed.
-In the former case, the total land area required to meet the target increased by 18\% (14,634 to 17,202 km$^2$), transport costs increased 22\% (5.52 X 10$^{-5}$ to 6.61 X 10$^{-5}$ truck hours), carbon losses increased by 19\% (597 to 711 Mt), and the total biodiversity index of converted areas increased by 20\% (5.53 X 10$^{-5}$ to 6.61 X 10$^{-5}$).
+In our main scenario, we assumed that the gap between current and projected future yields would be fully closed on existing cropland, but we also ran the model for cases when only 50\% or 150\% of the gap was closed (with equal weights across all objectives).
+In the former case, the total land area required to meet the target increased by 18\% (14,634 to 17,202 km$^2$), transport costs increased 22\% (5.52 x 10$^{-5}$ to 6.61 x 10$^{-5}$ truck hours), carbon losses increased by 19\% (597 to 711 Mt), and the total biodiversity index of converted areas increased by 20\% (5.53 x 10$^{-5}$ to 6.61 x 10$^{-5}$).
 In the 150\% gap closure scenario, the impacts were reversed, with impacts declining by almost exactly the same percentage.
@@
 
-The scenario is which the sole preference is to maximize yield (100\% weight) converts land with the most promising yield potential for farmers, which should have economic advantages.
+The scenario designed solely to maximize yield (100\% weight) converts land with the most promising yield potential for farmers, which should have economic advantages.
 These areas have little direct overlap with the areas picked by scenarios that minimize carbon costs, biodiversity impacts or travel time (Figure 1).
@@
 
-However, the areas selected solely for minimizing carbon, biodiversity, or transportation costs also have little direct overlap, but the map reveals that they have high adjacency, confirmed by direct estimates of distance (Figure 1).
+The areas selected solely for minimizing carbon, biodiversity, or transportation costs also have little direct overlap, but the map reveals that they have high adjacency, confirmed by direct estimates of distance (Figure 1).
 All three criteria tend to identify lands along existing road networks and population and agricultural centers, with the carbon minimizing objective converting a larger area of land in the western half of the country.
-The reasons for this correlation are probably that carbon stocks and biodiversity values are lowest in areas of high human density, because of high levels of wood harvesting for charcoal and fuelwood, and because human uses similarly cause fragmentation that leads to low biodiversity scores.
+The reasons for this correlation are probably that carbon stocks and biodiversity values are lowest in areas of high human densities, which have correspondingly high levels of wood harvesting for charcoal and fuelwood, and higher levels of habitat loss and fragmentation leading to lower biodiversity scores.
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/single_priorities3.png}}
+       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/figure1.pdf}}
     \caption{Areas converted to new maize and soybean croplands and their associated costs when each individual objective (yield maximization, and carbon cost, biodiversity cost, and transport time minimization) is given 100\% weight.
@@
 
-While the best case scenario for maximizing yield would convert 13,668 km$^2$, the worst case scenario among our single-purpose optimization scenarios (the scenario minimizing carbon costs) would convert 15,302 km$^2$.
-This range in converted areas is only 12\%, which is far less than the range in yields across the country’s potential cropland (3.7-6.1 t/ha/y ha$^{-1}$ for maize and 0.8 to 5.1 t/ha/y ha$^{-1}$ for soybean).
-One reason is the significance of yield to minimizing each of the costs because of the need to meet a fixed production target; because yield is in the denominator of each objective (e.g.
-carbon loss per ton of yield), land with better yield potential tends to be selected even when the yield maximization priority is given no weight.
-By contrast, the differences in travel times among single-objective scenarios range from a total of 3.4 X 10$^{-5}$ to 32.3 X 10$^{-5}$ truck hours, a difference of 872\%.
-Carbon costs differ by 156\% (462 to 1184 Mt), and biodiversity of 68\% (4.96 X 10$^{-5}$ to 8.33 X 10$^{-5}$).
-The low range in the yield-related cost (converted area), and the high ranges in the other costs, suggest that there is strong potential to reduce transportation, carbon and biodiversity costs for relatively little sacrifice of yield.
+While the best case scenario for maximizing yield would convert 13,668 km$^2$, the worst case scenario for yields among our single-purpose optimization scenarios would convert 15,302 km$^2$ (the carbon-optimizing scenario).
+This range in converted areas is only 12\%.
+This difference is far lower than the range in yields across the country's potential cropland (3.7-6.1 t/ha/y ha$^{-1}$ for maize and 0.8 to 5.1 t/ha/y ha$^{-1}$ for soybean), but yield plays such a significant role for each objective (because it is always in the denominator) that the model selects land with better yield potential even when the yield maximization priority is given no weight.
+By contrast, the differences in travel times among single-objective scenarios range from a total of 3.4 x 10$^{-5}$ to 32.3 x 10$^{-5}$ truck hours, a difference of 872\%.
+Carbon costs differ by 156\% (462 to 1184 Mt), and biodiversity by 68\% (4.96 x 10$^{-5}$ to 8.33 x 10$^{-5}$).
+These results indicate that there is a strong potential for reducing transportation, carbon and biodiversity costs with only a small increase in additional land area needed to meet production targets.
+In other words, these cost savings can be achieved for relatively little sacrifice of potential yield, which is inversely related to converted area.
 
-This potential is illustrated in Figure 2, which shows how much carbon, biodiversity, and transport cost can be avoided by transferring increasing amounts of weight to these objectives from the yield maximization objectives.
-Avoided cost is the percent difference between the cost paid under a given weight combination and the maximum cost found across all weighting scenarios (we ran all possible weight permutations varying in 5\% increments).
-Sequentially adding weights in 5\% increments to both the biodiversity and carbon objectives results in little cost to yield, but rapid, large reductions in both biodiversity and carbon costs as well as transport costs, despite no weight given to transport costs because of its correlations with the other two objectives).
-For example, adding just 15\% weight each to carbon and biodiversity increases the area of new lands (i.e.
-reduces the average yield) by just 2.7\% (avoided area costs drop to 8\% from 10.7\%), but avoids costs equal to 27\% for biodiversity, 47\% for carbon, and 44\% for transportation costs.
-For comparison, in the best case (100\% weight) scenario for each objective, 61\%, 41\%, and 90\% of carbon, biodiversity, and transportation cost would be avoided.
+This potential is illustrated in Figure 2, which plots the ``efficiency frontier'' \citep{polasky_where_2008} resulting from the range of possible tradeoffs between the yield maximization and carbon protection objectives.
+Each point represents the most efficient solution, in terms of lowest possible carbon loss and land area converted, for meeting the crop production targets under the given combination of weights.
+Color-coded maps indicate the associated cropland conversions for 5 of the 21 weighting combinations.
+Adding just 5\% weight to carbon protection (orange scenario) results in 31\% less carbon loss but $<$1\% additional converted area relative to the pure yield maximization objective (red scenario), without significantly altering the location of the new cropland (adjacency = 7.4 km).
+Transferring 25\% weight to carbon (yellow scenario) substantially shifts the conversion map (adjacency to pure yield maximization map = 65 km), but results in 54\% less carbon loss.
+These savings nearly equal the 61\% reduction in carbon loss achieved under the pure carbon protection scenario (dark blue).
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/weight_impact5.png}}
+       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/figure2.pdf}}
+    %\includegraphics[scale=0.5]{fig2.eps}
+    \caption{The costs of the optimal (most efficient) tradeoffs between the yield maximization and carbon protection objectives under varying  combinations of weights (varied in 5\% increments).
+Costs are measured in total carbon loss and converted area.
+The conversion areas resulting from a subset of the weight combinations are illustrated in the accompanying maps, and are color-coded to their corresponding costs on the scatter plot, which represents the efficiency frontier.
+The weightings (expressed as proportions) associated with this subset of scenarios are also provided: Y=yield maximization objective; C=carbon protection objective.}
+    \label{default}
+  \end{center}
+\end{figure}
+
+These last examples illustrate that tradeoffs can be usefully assessed in terms of avoided cost, or the percent difference between the cost paid under a given weight combination and the maximum cost generated by all possible weighting scenarios.
+By expressing avoided cost as a percentage, we can directly compare the impacts of tradeoffs between more than two objectives.
+Figure 3 shows how much carbon, biodiversity, and transport cost can be avoided by transferring increasing amounts of weight away from the yield maximization objective.
+Sequentially adding weights in 5\% increments to both the biodiversity and carbon objectives results in little cost in land area, but rapid, large reductions in both biodiversity and carbon costs.
+It also results in substantial avoided transport costs, even without assigning any weight to this objective, because of correlations between savings in transport, carbon, and biodiversity.
+For example, adding just 15\% weight each to carbon and biodiversity increases the area of new lands (i.e.
+reduces the average yield) by just 2.7\% (avoided area costs drop to 8\% from 10.7\%), but avoids potential costs equal to 27\% for biodiversity, 47\% for carbon, and 44\% for transportation.
+For comparison, in the best case  scenario for each objective, 61\%, 41\%, and 90\% of carbon, biodiversity, and transportation cost would be avoided.
+
+\begin{figure}[!ht]
+  \begin{center}
+       \makebox[\textwidth][c]{\includegraphics[width=0.8\textwidth]{figures/figure3.pdf}}
     %\includegraphics[scale=0.5]{fig2.eps}
@@
 
-Figure 3 shows the potential for compromise in which a hypothetical land user only interested in a single objective is willing to ``pay'' 5\% more costs than the lowest possible costs for that objective.
-For example, the yield maximizing user would be willing to pay for an extra 683 km$^2$ of new cropland, or the person interested in minimizing carbon costs would pay an additional 231 Mt of carbon.
+Figure 4 shows the potential for compromise in which a hypothetical land user only interested in a single objective is willing to ``pay'' 5\% more costs than the lowest possible costs for that objective.
+For example, the yield maximizing user would be willing to pay for an extra 683 km$^2$ of new cropland, while the person interested in minimizing carbon costs would pay an additional 231 Mt of carbon.
 Lines in the bar charts show the minimum and maximum costs across all possible weighting scenarios.
 Each of the four different ``compromise'' scenarios, represented by a different colour, permits solutions that allow the other scenarios to save substantially relative to maximum potential costs.
-The carbon, biodiversity, and transport cost objectives in all but two cases avoid more than 50\% of their maximum potential costs when any of the other objectives is willing to pay 5\% more cost (Figure 3).
+The carbon, biodiversity, and transport cost objectives in all but two cases avoid more than 50\% of their maximum potential costs when any of the other objectives is willing to pay 5\% more cost (Figure 4).
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_5pct3.png}}
+       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/figure4.pdf}}
     \caption{The results of scenarios in which each land use objective is willing to pay 5\% more cost to compromise with the other three objectives.
@@
 
-The weight on the objective subject to the 5\% compromise is proportional to the size of its potential avoided cost (see Table 1 in supplementary results), thus yield requires a weight of only 40\% to pay 5\% or less of its lowest cost, whereas transportation cost and carbon, which have the largest cost ranges, require 85\% weight each, whereas biodiversity requires 50\% weight.
-
-We also analyzed an ``equal compromise''  scenario in which each objective receives equal weight, as shown in Figure 4.
-The bar charts show how this scenario, represented in purple, compares with each of the 5\% compromise scenarios shown in Figure 3, plus the best and worst case options among all weighting scenarios.
+We also analyzed an ``equal compromise''  scenario in which each objective receives equal weight, as shown in Figure 5.
+The bar charts show how this scenario, represented in purple, compares with each of the 5\% compromise scenarios shown in Figure 4, plus the best and worst case options among all weighting scenarios.
 Relative to the worst case, equal compromise avoids 80\% or more of the cost for carbon, biodiversity, and transportation, and 41\% of the cost for land area.
-With respect to the 5\% compromises, the equal weighting scenario is lower than, but compares favorably to, the $\sim$95\% savings achieved by carbon, biodiversity, and transportation costs within their individual cost categories, and to the 58\% savings by the yield objective.
+The equal weighting scenario compares favorably to the $\sim$95\% savings achieved by carbon, biodiversity, and transportation costs within their individual cost categories, and to the 58\% savings by the yield objective.
 
@@
   \begin{center}
-       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/compromise_25pct2.png}}
+       \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{figures/figure5.pdf}}
     \caption{Areas converted to new maize and soybean croplands and their associated costs when when all objectives (yield maximization, and carbon, biodiversity, and transport cost minimization) compromise equally (i.e.
 receive 25\% weight).
-Cropland converted under this scenario is denoted by purple areas in the map, with its associated cost impacts shown alongside those from the 5\% compromise scenarios (Figure 3), together with the maximum, minimum, and median costs across all weighting permutations (grey horizontal lines in bar charts).}
+Cropland converted under this scenario is denoted by purple areas in the map, with its associated cost impacts shown alongside those from the 5\% compromise scenarios (Figure 4), together with the maximum, minimum, and median costs across all weighting permutations (grey horizontal lines in bar charts).}
     \label{default}
@@
 Zambia has designated a large number of farm blocks, and added nine more blocks with roughly one million hectares as recently as 2005 \citep{anonymous_farm_2005}.
-Although these areas have in fact as a whole attracted little development, they remain ``on the books'', and there are at least some reports of continued intent to develop them \citep{anonymous_undeveloped_2015}.
+Although these areas have as a whole attracted relatively little development, there is continued interest in developing them \citep{anonymous_undeveloped_2015}.
 But there is less than 2\% overlap between designated farm blocks and either the yield maximization scenario and the equal weighting scenario, and very little adjacency (see supplementary results).
@@
 Our results for Zambia suggest at least the potential to plan agricultural expansion such that it limits carbon and biodiversity costs with limited sacrifice of yield potential.
-In fact, if agricultural projection goals alone are best represented by a combination of agronomic potential and existing road access, then solutions that provide a good balance of these objectives also correlate strongly with solutions that limit carbon and biodiversity impacts.
+In fact, if agricultural production goals alone are best represented by a combination of agronomic potential and existing road access, then solutions that provide a good balance of these objectives alone also correlate strongly with solutions that limit carbon and biodiversity impacts.
 
-This finding demonstrates the importance of targeting land uses based on their ability to deliver a particular benefit for the lowest possible cost.
-The value of such efficiency-based approach, which we apply here to agricultural land use, is in line with findings from the field of conservation planning, where it has been shown that, by considering both the price of land and its ecological value, conservation objectives can be met for less than half the cost of approaches that target land solely based on ecological value \citep{naidoo_integrating_2006}.
+This finding demonstrates the potential benefits of targeting land uses based on their ability to deliver a particular benefit for the lowest possible cost.
+The value of such an efficiency-based approach, which we apply here to agricultural land use, is in line with findings from the field of conservation planning, where it has been shown that conservation outcomes can be achieved for substantially lower cost when both the economic and ecological value of land is considered \citep[e.g.][]{naidoo_integrating_2006, polasky_where_2008}.
 
-Another result, if more obvious, is the potential benefits of increasing yields on existing cropland.
+Another result, if more obvious, is the potential benefits of increasing yields on existing cropland if used to spare land.
 Actions that boost yields on existing cropland also have the potential to have feedback effects that lead to more local expansion, for example, by boosting the competitiveness and therefore quantity of exports \cite{ewers_increases_2009,angelsen_policies_2010}.
-Our analysis ignores those potential feedback effects because all options analyzed meet the same production targets, but our analyses quantities the large potential benefits if these yield gains are used to limit environmental impacts.
+Our analysis ignores those potential feedback effects because all options analyzed meet the same production targets, but shows how yield gains could be used to limit environmental impacts.
 
@@
 Road construction has been a primary driver of the location of agricultural expansion \citep{laurance_impacts_2009}.
-Zambia is already engaged in a program to improve existing major roads, but there is evidence that agriculture responds in particular to improvements in the quality of smaller, feeder rural roads to larger paved roadways \citep{bourguignon_rural_2008,blimpo_public_2013}.
-Improving smaller rural roads would be one way of focusing Zambia's agriculture development.
-Because agricultural development tends to build on the private and public infrastructure that accompanies initial development, the focus on existing areas would also seem likely to be a more robust, long-term strategy for directing agricultural development \cite{laurance_estimating_2015}.
-A potential impediment to this, however, might be that land closer to roads is already ``spoken for'', and thus less likely to be available for new farms.
+Zambia is already engaged in a program to upgrade existing major roads, but there is some evidence in other parts of Africa that agriculture responds even more to improvements in the quality of smaller, rural roads than to improvements in larger paved roadways \citep{bourguignon_rural_2008,blimpo_public_2013}.
+Improving rural roads would be one way of focusing Zambia's agriculture development on existing road networks.
+Because agricultural development tends to build on the private and public infrastructure that accompanies initial development, the focus on existing areas would also seem likely to be a more robust, strategy for avoiding carbon and biodiversity loss over the long-term \cite{laurance_estimating_2015}.
 
-This model appears to be the right kind of tool for assisting the planning of agricultural development.
+Beyond road infrastructure, our results also indicate the importance of assessing current agricultural development plans because of the low overlap between Zambia's agricultural development blocks and the areas selected for any objective.
+
+We believe this type of the model would be particularly useful for planning agricultural development in any politically heterogeneous environment.
 Optimization models that require ex ante specification of preferences in mathematical terms work less well for decision-making that must weigh very different objectives in unclear ways.
@@
 For this kind of tool to be truly legitimate and useful on a practical level, we believe several elements are required.
-First, it must be further developed and applied in an iterative way with the government and other stakeholders so that the impact of different weighting decisions, in terms of their costs and benefits, can be fully understood, thereby facilitating informed land use planning.
+First, it should be further developed and applied in an iterative way with the government and other stakeholders so that the costs and benefits of different weighting decisions are fully explored.
+We are developing plans to work with governments, various stakeholders, and international collaborators to further develop and use this model as an operational decision-support tool in Zambia and other countries in the region.
 
-Second, development planning must reflect and incorporate other real world factors beyond those incorporated in our model.
+Second, development planning must reflect and incorporate other real world factors beyond those included in our model.
 For example, large, multi-lane roads can lead to high rates of wildlife mortality, and serve as effective barriers to migration for large mammals, regardless of other land uses \citep{holdo_predicted_2011}.
@@
 Third, data quality is critical.
-This kind of tool focuses on particular hectares, not broad averages, so data errors have little or no chance of averaging out through aggregation.
-Crop modeling predictions and most of our other scores can be highly sensitive to a range of data parameters.
-These facts mean that long-term use of the tool should be accompanied by steady data improvement.
-It also means that the tool should mostly serve as a guide to likely values.
-Before governments will make important decisions ``for keeps'', there should be site-specific ground-truthing of soil properties and yield potential, as well as confirmation of biodiversity and carbon characteristics.
+This kind of model focuses on particular hectares, not broad averages, so data errors have little or no chance of averaging out through aggregation.
+Predicted potential yields and most of our other scores can be highly sensitive to errors in the inputs used to calculate them.
+These facts mean that long-term use of the model should be accompanied by steady data improvement.
+It also means that the model should mostly serve as a guide to likely values.
+Before governments make any important and potentially irreversible decisions, there should be site-specific ground-truthing of soil properties and yield potential, as well as confirmation of biodiversity and carbon characteristics.
 
-Some elements of the model should also be further developed, depending on the ability to gather significant additional data.
-Analysis of agricultural potential could incorporate more economic factors, differentiate between smallholder and commercial farming practices, and incorporate irrigation potential.
+Fourth, some elements of the model should also be further developed.
+Analysis of agricultural potential could incorporate more economic factors, differentiate between smallholder and commercial farming practices, and incorporate irrigation potential \citep[e.g.][]{nijbroek_regional_2016}.
 Additional information on land ownership and title would allow a finer delineation of potential agricultural land, while the carbon estimates would be improved by root-shoot specific to each vegetation type.
-Biodiversity impacts could be better assessed by incorporation data on endemism and the degree to which they are impacted by habitat fragmentation.
+For biodiversity, defining rarity/threat in relation to national borders may provide a biased view of conservation importance--habitats that are rare in Zambia might be common across the border, while a relatively common habitat in Zambia might be regionally rare.
+A regional approach to defining this measure might therefore be more appropriate \citep[e.g.][]{van_breugel_environmental_2015}.
+This index could also be adjusted to factor in additional indicators of biodiversity value, such as Important Bird Areas \citep{butchart_protecting_2012} or Key Biodiversity Areas \citep{eken_key_2004}.
+
+Analysis of biodiversity impacts could also improved by more direct quantification of biodiversity responses to land use.
 One potentially promising approach would be to incorporate the biodiversity impact model developed by Newbold et al.
-\citep{newbold_global_2015}, provided the land use intensity maps on which it is conditioned can be developed for this resolution.
+\citep{newbold_global_2015}, which estimates changes in species diversity and abundance in response to landcover and land-use intensity.
+With sufficient data, this approach could also allow comparison of the biodiversity impacts of different agricultural practices, including land-sparing approaches that maintain trees, vegetated field borders, or other habitats within the agricultural landscape \citep{kremen_reframing_2015}.
+The model can be easily adjusted to simulate varying intensities of agricultural management.
 
-Overall, however, our results are promising, and suggest that efficiency-based land use planning approaches such as ours can help to substantially minimize the ecological cost of cropland expansion in sub-Saharan Africa's savannas, the 21st Century's emerging hotspot of agricultural land use change.
+The model could also be used to evaluate the temporal persistence of tradeoffs in response to key uncertainties, such as the future effects of climate change on agricultural potential.
+Robustness to such uncertainty should be a key factor in determining new agricultural development areas.
+Assessment of robustness can be achieved by identifying stable areas of conversion that emerge over many iterations within the bounds of uncertainty, which is a feature of the model that we intend to develop.
+
+Despite the current limitations of the model and this initial analysis, our results are promising.
+They suggest that efficiency-based land use planning approaches such as ours, if operationalized and incorporated into decision-making processes, can help to substantially minimize the ecological cost of cropland expansion in sub-Saharan Africa's savannas, the 21st Century's emerging hotspot of agricultural land use change.
 
@@
 \section*{\large Funding}
-This work was supported by funds from: the Norwegian Aid Agency (NORAD) under the Agricultural Synergies Project; the Princeton Environmental Institute Grand Challenges program; the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463).
+This work was supported by funds from: the Norwegian Aid Agency (NORAD) under the Agricultural Synergies Project; the Princeton Environmental Institute Grand Challenges program; the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463 and SES-1534544).
 
@@
 \bibliographystyle{prsb} 
-\bibliography{/Users/lestes/Dropbox/publications/full}
+\bibliography{/Users/lestes/Dropbox/publications/fullbib}
 

### 2016-05-05 paper/supplementary_methods.tex
Full paper revisions: main manuscript, fully revised, with edits and additions to SI methods and results.  New figure created (figures.Rmd), slight edit of loose code on main-analysis.Rmd

@@
 \documentclass[a4paper]{article}
-%\usepackage[affil-it]{authblk}  
+\usepackage[affil-it]{authblk}  
 %\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}  % apostrophe's in cites
@@
 % http://ctan.mackichan.com/macros/latex/contrib/caption/caption-eng.pdf
+\usepackage[numbers,sort,compress]{natbib}
+
 \usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
@@
 \captionsetup[figure]{slc=off, font=footnotesize}
-\captionsetup[table]{margin={4cm, 4cm}, slc=off, font=footnotesize}
+\captionsetup[table]{margin={4cm, 4cm}, slc=off, font=small}
 \usepackage[numbers,sort,compress]{natbib}
@@
 \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+\usepackage{multirow}
 
 \begin{document}
-\section*{\Large SI Methods}
+\section*{SI Methods}
 
@@
 
+To assess the effectiveness of this routine, and the degree to which it might cause the ideal allocation area for one crop to be displaced by that of another, we ran the model separately for each crop under a yield maximization scenario and compared the degree of overlap (see SI Results).
+
 \subsection*{Model Assumptions and Limitations}
 Our model is limited to finding solutions in a convex solution space, and lacks the optimality guarantees possessed by more sophisticated algorithms.
-More sophisticated approaches, such as genetic approaches \citep[e.g.][]{stewart_genetic_2004,cao_spatial_2011}, which are adept at finding diverse sets of solutions, including those in non-convex spaces, their computational complexity causes model runs to be very time consuming \citep{geertman_participatory_2002}.
-Algorithms continue to grow more efficient using more sophisticated algorithms, but even these new routines compromise the number of outputs they are able to generate within a given time, which is problematic especially as the size of the inputs grows.
+More sophisticated approaches, such as genetic approaches \citep[e.g.][]{stewart_genetic_2004,cao_spatial_2011}, which are adept at finding diverse sets of solutions, including those in non-convex spaces, have high computational complexity that causes model runs to be very time consuming \citep{geertman_participatory_2002}.
+Algorithms continue to grow in efficiency and sophistication, but even these newer routines compromise the number of outputs they are able to generate within a given time, which is problematic especially as the size of the inputs grows.
 The advantage of our approach is that it enables quick determination of land use changes, by applying a hybrid set of weights to multiple objectives in order to search for qualitatively appealing land use change scenarios.
@@
 
-We ran the model at 40 weather stations for each of the dominant soils associated in the 10X10 km area \cite{romero_reanalysis_2012} around those stations, for each growing season between 1979 and 2010, resulting in a total of 5,859 climate X soil X weather year combinations.
+We ran the model at 40 weather stations for each of the dominant soils associated in the 10x10 km area \cite{romero_reanalysis_2012} around those stations, for each growing season between 1979 and 2010 (which totals 30 seasons, given that Zambia's growing season begins in November-December and ends in March-April), resulting in a total of 5,859 climate x soil x weather year combinations.
 This was repeated for each cultivar, after which we averaged the results at each station and soil types and for each year across cultivars.
@@
 
-We fit generalized additive models \citep{hastie_generalized_1990,wood_mgcv:_2001} to the cultivar averaged yields, which we further reduced to a mean, 10th percentile, and 90th percentile yield value for each weather station-soil combination, using the predictors described in the main text.
+To fit generalized additive models \citep[GAM;][]{hastie_generalized_1990,wood_mgcv_2001} to the cultivar averaged yields, we first derived predictors from DSSAT's primary weather and soil inputs (Table 1).
+For each of the 30 growing seasons in the time series, we calculated the following variables from the daily weather data \citep{sheffield_development_2006, chaney_spatial_2014, estes_changing_2014} used to run DSSAT: the cumulative rainfall for three different portions of the growing season (November-March; December-March; January-March); the mean daily solar radiation for the same three portions of the growing season; growing degree days (GDD).
+GDD was calculated by first calculating daily mean temperature from the daily maximum and minimum, and then summing the amount of each day's mean temperature exceeding 8$^{\circ}$ \citep[per][]{licker_mind_2010} for each year.
+From the soil profiles \citep{harvestchoice_converting_2010} that we used to run DSSAT, we calculated for the top 1 m of the soil profile average values of the following parameters: percent soil clay content; percent organic carbon; bulk density; pH.
+Of these, the clay parameter is not used directly by DSSAT, but is strongly correlated (along with percent organic carbon, which is directly used by DSSAT) with wilting point and field capacity \citep{saxton_estimating_1986,saxton_soil_2006,hutson_estimation_1983}, two soil hydraulic parameters that DSSAT requires but which are not available in the 1 km soil maps \citep{hengl_mapping_2015}.
+DSSAT in fact provides a tool that uses clay and other texture variables to calculate these hydraulic properties when they are missing using pedotransfer functions, which is a common practice when pre-processing soil inputs for using the model \citep[e.g.][]{estes_comparing_2013}.
+
+\begin{table}[htdp]
+\captionsetup{width=0.9\linewidth}
+\caption{Correspondence between several primary weather and soil variables used by the DSSAT model \citep{jones_potential_2003} and the predictors derived from them that were used in fitting the generalized additive model.}
+\vspace{-10pt}
+\begin{center}
+\begin{tabular}{p{1.5cm}p{6cm}p{6cm}}
+  \hline
+  Type & DSSAT variable & GAM variable \\
+  \hline\hline
+  \multirow{8}{*}{Weather} & \multirow{3}{*}{\parbox{6cm}{Daily rainfall}} & Nov-March cumulative rainfall \\
+                  &    & Dec-March cumulative rainfall \\
+                 &    & Jan-March cumulative rainfall \\\cline{2-3}
+                 & \multirow{3}{*}{\parbox{6cm}{Daily shortwave radiation }} &  Nov-March mean daily solar radiation \\
+                 & & Dec-March mean daily solar radiation\\
+                 & & Jan-March mean daily solar radiation \\\cline{2-3}
+                 & \multirow{1}{*}{\parbox{6cm}{Daily max \& min temperatures }} & 
+                    Growing degree days (GDD) \\
+  \hline                  
+  \multirow{8}{*}{Soil} & Field capacity in \emph{n} horizons  & - \\
+                                 %  \multirow{2}{*}{\parbox{7cm}{Mean \% clay content in top 1 m}}\\ 
+                  &  Wilting point in \emph{n} horizons & - \\\cline{2-3}
+                  & \% clay content in \emph{n horizons} & Mean \% clay content in top 1 m \\
+                  & \% organic carbon in \emph{n} horizons & Mean \% organic carbon in top 1 m \\
+                  & Bulk density in \emph{n} horizons & Mean bulk density in top 1 m \\
+                  & pH in \emph{n} horizons  & Mean pH in top 1 m \\\cline{2-3}
+%                  & Root growth factor in \emph{n} horizons  & - \\
+%                  & Drainage coefficient  & - \\
+  \hline
+\end{tabular}
+\end{center}
+\label{default}
+\end{table}%
+
+After calculating these predictor variables from DSSAT's inputs, we condensed the DSSAT predicted yields by calculating the mean and 10th and 90th percentile yield values for each weather-soil combination.
+To make sure the derived weather predictors (Table 1) aligned with these yield values, we calculated the 30-year means of each weather predictor and associated these with the DSSAT mean yields, and extracted the weather variables for the years corresponding to the 10th and 90th percentile yield values.
+Derived soil predictors did not vary by year.
+We then fit GAM models using the derived variables as predictors of these summarized DSSAT yield values.
+
 We adopted the GAM approach because it makes no assumptions about the distribution of the data, and can fit complex non-linear relationships.
@@
 The soybean used the same predictors, except the precipitation was the sum of the January-March period (JFMP) and solar radiation was averaged over the same time period (SSRMIII).
-With the exception of GDD in soybean (which was near significant at the p<0.1 level and was retained for consistency with the maize model), all predictors were significant on one or both of the linear and squared term of the predictor, while the geographic terms was significant in both models.
+With the exception of GDD in soybean (which was near significant at the p$<$0.1 level and was retained for consistency with the maize model), all predictors were significant on one or both of the linear and squared term of the predictor, while the geographic terms was significant in both models.
 
-Using the gridded predictor variables, we mapped predicted yield values for all of Zambia.
-We then adjusted these data to match the FAO-projected yields as described in the main text.
+To map predicted yields across Zambia, we first derived gridded versions of the GAM predictors (Table 1) from the gridded daily weather data \citep{sheffield_development_2006,chaney_spatial_2014,estes_changing_2014}, and from the 1 km soil maps \citep{hengl_mapping_2015}.
+The soil maps provide values for 6 different layers at standard depths of 0-5 cm, 5-15 cm, 15-30 cm, 30-60 cm, 60-100 cm, and 100-200 cm.
+We therefore calculated the mean of the first 5 layers, weighted by each layer's depth, to calculate the equivalent of the predictors derived from DSSAT's soil inputs (Table 1).
+The GAM coefficients were then applied to the gridded predictor sets to create the yield maps, after which we adjusted the predicted yields to match the FAO-projected yields as described in the main text.
 We used two sources to calculate the likely future yield improvement trend for Zambia.
@@
 \begin{equation}
-y_{2050} = (1 + \beta(2050-2014))y_{current}
+\textrm{y}_{2050} = (1 + \beta(2050-2014))\textrm{y}_{\textrm{current}}
 \end{equation}
@@
 
-The then adjusted the yields in the GAM-derived maps so that their means matched these projected means.
+We then adjusted the yields in the GAM-derived maps so that their means matched these projected means.
 However, since the yield rates and average values on which they were based were obtained from existing cropland, we extracted the gridded yield values that corresponded with existing cropland (determined using the landcover dataset, see main text).
@@
 The choice of landcover product can have a critical impact on the results of any spatial analysis.
-This is particularly true with respect to trying to locate the current extent of existing cropland, where existing landcover datasets are particularly discrepant and in wide disagreement \cite{fritz_comparison_2010} because the average cropland extent is small and, from a nadir perspective, are often difficult to distinguish from the surrounding savanna vegetation \cite{estes_platform_2016,debats_generalized_2015}.
-The most accurate landcover datasets tend to be country-specific Landsat-based maps \cite[][and per a forthcoming South Africa-focused landcover intercomparison]{fritz_comparison_2010}, therefore we selected the 30 m resolution type II landcover map for Zambia for the year 2010, developed under the NASA SERVIR program\footnote{downloaded from http://apps.rcmrd.org/landcoverviewer/}.
-These products have undergone extensive validation and provide the most current and best available landcover maps for Zambia and several other eastern and southern African countries, and their consistency across the region is another advantage in that it facilitates the transferability of our model.
+This is particularly true with respect to the distribution and extent of current cropland.
+Most existing landcover datasets are notoriously inaccurate for this task \cite{fritz_comparison_2010}, because the average field size is small and, from a nadir perspective, often difficult to distinguish from the surrounding savanna vegetation \cite{estes_platform_2016,debats_generalized_2015}.
+The most accurate landcover datasets tend to be country-specific Landsat-based maps \cite[][and per a forthcoming South Africa-focused landcover intercomparison]{fritz_comparison_2010}, therefore we selected the 30 m resolution scheme II landcover map for Zambia for the year 2010, developed under the NASA SERVIR program\footnote{downloaded from http://apps.rcmrd.org/landcoverviewer/}.
+These products have undergone extensive validation and provide the most current and best available landcover maps for Zambia and several other eastern and southern African countries.
+The regional consistency of these maps provides a further advantage in that it facilitates the transferability of our model.
 
 We used this landcover map as the basis for determining how much land in each 1 km$^2$ grid cell was unconverted to cropland or urban areas.
-This information formed the basis for determining what land was potentially available for new cropland (although it does not factor in ownership or other factors that cannot be seen from space, which determine true availability), and also the degree to which habitats remained intact and un-fragmented.
+This information formed the basis for determining what land was potentially available for new cropland (although it cannot factor in ownership and features that are invisible from space, which determine true availability), and also the degree to which habitats remained intact and un-fragmented.
 To calculate both measures, we first removed cropland and urban cover types from the 30 m resolution landcover map, and then aggregated the remainder to 1 km$^2$, producing a fractional map expressing how much untransformed land was available in each grid cell.
@@
 We wanted to give equal weight to intactness because the overwhelmingly dominant habitat types in Zambia are miombo woodlands (Table 1), which to date have remained relatively intact.
-Applying just the rarity/threat index, would penalize these habitats, yet their relative intactness, together with high mammal species richness, is one of the factors that make them of conservation interest \cite{olson_terrestrial_2001,searchinger_high_2015,brooks_global_2006}.
+Applying just the rarity/threat index, would penalize these habitats, yet their relative intactness, together with high mammal species richness, is one of the factors that make them of conservation interest \cite{olson_terrestrial_2001, searchinger_high_2015, brooks_global_2006}.
 Furthermore, a unique feature of miombo woodlands is that, due to the digestibility of forage, megaherbivores (buffalo, elephant) are over-represented in the assemblage of large mammalian herbivores, and correspondingly require large range sizes to maintain viable populations \cite{mcnaughton_ecology_1986}.
@@
 
-To calculate the time to travel to the nearest major town, we considered not just Zambia but also portions of the neighboring countries, under the assumption that for certain areas within Zambia near its borders, the closest market center might be in a neighboring country.
-We therefore calculated travel time within a region defined by longitudes 19.5 to 34.5 degrees, and latitudes -19.5 to -8 degrees.
-
-We obtained a vector dataset of roads from the Zambian government, and downloaded an additional dataset available for broader region from OpenStreetMap.
-We merged these two sets into a single dataset in order to capture the road network in and around Zambia, and classified roads into trunk roads, primary roads, secondary roads, and tertiary roads.
-
-We obtained vector datasets for lakes and major towns in Zambia from the government.
-For towns outside Zambia, we used the Gridded Population of the World Version 3 \citep{ciesin_gridded_2005} town vectors.
-We used the USGS Hydrosheds database to indicate the position and size of rivers, based on their contributing area.
-
-The roads, rivers, and lakes were converted to a raster of 1 km resolution (the same as our analysis), together with Zambia's borders.
-Travel times were then assigned to each of the features, based on how long it would take to cross a 1 km pixel:
+To create the cost distance analysis, features used to define the input friction surface (see main text) were assigned different estimates for how long it took to travel 1 km.
+These were assigned as follows:
 

### 2016-05-05 paper/supplementary_results.tex
Full paper revisions: main manuscript, fully revised, with edits and additions to SI methods and results.  New figure created (figures.Rmd), slight edit of loose code on main-analysis.Rmd

@@
 
-\section*{\large Allocation of land between crops}
-
-To compare the degree to which the cropland allocation routine of \emph{agroEcoTradeoff} causes land that is best for maize to be allocated to soybean and vice versa, we ran the model for each crop separately (with 100\% weight on yield maximization) and compared the overlaps between the resulting conversion maps (Figure 1).
-The two crops were allocated to the same areas in $<$1\% of converted pixels.
+\section*{\large Potential displacements caused by cropland allocation routine}
+To assess the degree to which the model's cropland allocation routine impacts the areas selected for conversion for each crop (see SI Methods for details), we ran the model separately for each crop, using the same parameters applied in the main analysis for the yield maximization scenario.
+We then assessed the overlap in the resulting areas selected for conversion for each crop in the resulting output maps (Figure 1).
+The results indicated that the two crops were allocated to the same areas in $<$1\% of converted pixels.
 
